### 2018-10-19 paper_1.tex
Added tex source

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is the template for submission to ISCA 2018
% The cls file is a modified from  'sig-alternate.cls'
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{sig-alternate} 
\usepackage{mathptmx} % This is Times font

\newcommand{\ignore}[1]{}
\usepackage{fancyhdr}
\usepackage[normalem]{ulem}
\usepackage[hyphens]{url}
\usepackage{microtype}
\usepackage{fixltx2e}
\usepackage{array}

% Always include hyperref last
\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=black,citecolor=blue,urlcolor=black]{hyperref}

% Ensure letter paper
\pdfpagewidth=8.5in
\pdfpageheight=11in

%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
\newcommand{\iscasubmissionnumber}{NaN}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\fancypagestyle{firstpage}{
  \fancyhf{}
\renewcommand{\headrulewidth}{0pt}
  \fancyhead[C]{\normalsize{}} 
  \fancyfoot[C]{\thepage}
}

\pagenumbering{arabic}

%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
\title{SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher} 
\author{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle
\thispagestyle{firstpage}
\pagestyle{plain}

\begin{abstract}

Hardware prefetching has been introduced in modern processors as a way to reduce compulsory cache misses. An efficient prefetcher should be able to identify complex memory access patterns. This enables it to fetch a block ahead of its demand access.

In this paper, we introduce perceptron learning to help make this prefetching decision efficiently. We've taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine. The perceptron layer filters out the unnecessary prefetches recommended by SPP. We've also explored a range of features that can be used to train the perceptron layer. Our results show that SPP-Perc improves performance on SPEC 2017 Benchmark suite by 3.2\% on single-core and 5.5\% on 4-core simulation, as compared to just SPP.

\end{abstract}

\section{Introduction}
Since past few decades, processor and memory scaling has been enabled with different philosophies in mind. While processor scaling focused on speed improvements, memory scaling was centred around increasing the storage capacity. The process technology reaching scaling limitations led to Memory Wall [cite] - the increasing performance gap between the processor speed and the memory speeds. Various techniques have been developed to combat this widening gap. Data prefetching is one such important technique. Data prefetching exploits the fact that most applications, data is accessed with some spatial or temporal locality.

An ideal prefetching scheme involves capturing the memory access patterns to predict future prefetches in a timely manner. Other important decisions are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block. Memory accesses patterns can be as simple as fetching the next cache line or a complex but predictable pattern, like in pointer chasing applications. Predicting this often requires a trade-off between coverage and accuracy. For a prefetcher to capture highly irregular access patterns, it needs to be highly aggressive (high coverage). That also means the prefetcher would end up recommending a lot of useless prefetches, leading to cache pollution (low accuracy). This effect would get even more pronounced in multi-core scenario where the last level cache is a shared resource. Conversely, a conservative prefetcher would be highly accurate but might not make enough useful predictions.

To this effect we propose <come up with a catchy name> <SPP-Perc>. It's an enhancement to the existing state of the art prefetcher and overcomes the coverage vs trade-off issue in a robust manner. The idea is to modify SPP design to make it as aggressive as it can get. This enables SPP to capture complex memory access patterns hence leading to increased coverage. The prefetch suggestions recommended by SPP pass through the perceptron based throttler. Over time, perceptron layer learns to correlate the prefetches recommended by SPP with various available features and eliminate useless prefetches. This leads to an overall increased accuracy of the prefetch scheme and reduction in DRAM traffic.

\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and future scope for expansion. The contributions made by this paper are:

\begin{itemize}
\item This is the first time where a truly on-line learning neural network based model has been used when it comes to prefetching. The prior works in this area either relied on program semantics [cite] or were application specific [cite] or could not yield a significant improvement over the baseline prefetcher used. [cite]

\item Coming up with a robust and accurate prefetching mechanism which can yield increased accuracy as compared to the other state of the art prefetchers

\item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind. We've used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers. We've also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.

\item We came up with an efficient model to use the perceptron output as a representation of the prefetch confidence using an on-line histogram technique. We show how the simple sum can replace the existing mechanism which was essentially based on counting the number of hits for a particular entry in SPP.

\end{itemize}

We tested SPP-Perc on Champsim, an open source simulator which models an Out-of-Order CPU using. For evaluation, we used SPEC 2017 benchmarks. On a single core configuration, SPP-Perc increases performance of the system by XX\% as compared to the baseline with no prefetching. This outperforms SPP by 3.2\%. On a 4-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement of XX\% over the baseline. This was ahead of SPP by 5.5\%. 
<Allude to 8 core results>
<Write about other workloads like Cloudsuite, MLPack>

\vspace{1ex}The rest of the paper is organized as follows:

\begin{itemize}
\item Section 2 talks about background work and recaptures the SPP architecture
\item Section 3 discusses enhancements made to the SPP design and the detailed implementation of the perceptron mechanism
\item Section 4 consists of the methodology for evaluating SPP-Perc and exploring the feature space for perceptron learnng
\item Section 5 comprises results and evaluation in terms of IPC, MPKI, coverage and DRAM traffic, for a variety of benchmark suites
\item Section 6 is the conclusion and future scope of expansion on this work
\end{itemize}

\section{Background Work and Motivation}

In this section, we compare the existing work done in the domain of prefetching. We also discuss about some of the interesting memory access patterns seen in the workloads in SPEC 2017 benchmark. The second half of this section deals with a background of Signature Path Prefetcher architecture.

\subsection{Motivation}
Some of the earliest data prefetchers developed to leverage this locality were stride based [cite:\newline
- A. J. Smith. Sequential Program Prefetching in Memory Hierarchies (1978)\newline
- J.-L. Baer and T.-F. Chen. An Effective On-Chip Preloading Scheme to Reduce Data Access Penalty (1991)\newline
]\newline

Major research in prefetching domain was based on stride length and depth prediction [cite:\newline
- Decoupled Predictor-Directed Stream Prefetching
Architecture (2003)\newline
- Memory Prefetching Using Adaptive Stream Detection (2006)\newline
]\newline

With time newer and more intricate prefetchers were introduced. [Cite:\newline
- Temporal Streaming of Shared Memory (2005)\newline
- Stealth Prefetching (2006)\newline
- Feedback Directed Prefetching: Improving the Performance and Bandwidth-Efficiency of Hardware Prefetchers (2007)\newline
- Coordinated Control of Multiple Prefetchers in Multi-Core Systems (2009)\newline
- Techniques for Bandwidth-Efficient Prefetching of Linked Data Structures in Hybrid Prefetching Systems (2009)\newline
- PACMan: PrefetchAware Cache Management for High Performance Caching (2011)\newline
- TLB Improvements for Chip Multiprocessors: Inter-Core Cooperative Prefetchers and Shared Last-Level TLBs (2013)\newline
- Linearizing Irregular Memory Accesses for Improved Correlated Prefetching (2013)\newline
- Sandbox Prefetching: Safe run-time evaluation of aggressive prefetchers\newline
]\newline

Prefetchers often tend to correlate predictions with the past memory access addresses [cite: Making Address-Correlated Prefetching Practical (2010)].  A whole class of streaming prefetchers developed on the lines of Spatial / Temporal Memory Streaming
[ Cite:\newline
- Accurate and Complexity-Effective Spatial Pattern Prediction (2004)\newline
- Spatial Memory Streaming (2006)\newline
- Temporal Instruction Fetch Streaming (2008)\newline
- Practical Off-Chip Meta-Data for Temporal Memory Streaming (2009)\newline
- Spatio-Temporal Memory Streaming (2009)\newline
- Spatial Memory Streaming with Rotated Patterns (2009)\newline
- Spatial Memory Streaming.” Journal of Instruction-Level Parallelism" (2011)\newline
]\newline

More modern prefetchers are often offset based <cite BOP and other offset based> or lookahead based <cite SPP, KPC and a few more>.

A related line of work also explored cache replacement: insertion policies and dead block prediction, with an eye towards prefetching. [Cite:\newline
- Dead-Block Prediction and Dead-Block Correlating Prefetchers (2001)\newline
- Cache Bursts: A New Approach for Eliminating Dead Blocks and Increasing Cache Efficiency (2008)\newline
- KPC (2017)\newline
- Rethinking Belady's Algorithm to Accommodate Prefetching (2018)\newline
]

\subsection{Prefetching with SPEC 2017}
<Discuss SPEC 2017 traces here>
<What are the possible graphs / analysis tools we can use for SPEC traces?>

\subsection{Perceptron Learning in Architecture}

Perceptron learning for computer architecture design has been around for a while. It was popularized by [cite initial branch prediction paper]. In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the sum is assumed to indicate a  high confidence taken branch \textit{true}. Else the confidence for branch taken is not high enough and the prediction is made as \textit{false}.

The structure that we use is derived from the model introduced in [cite piecewise linear branch prediction]. Here the feature itself is used to index into a hash of perceptron weights. This form of indexing makes sure that the hyper-plane learned by the perceptron weights is able to differentiate between linearly inseparable outcomes. <Is it true? Not sure>

Perceptron Learning for Reuse Prediction [cite] extended the hash perceptron architecture to perform prediction in context of cache replacement. Perceptron layer can learn to correlate the cache replacement behaviour with a variety of features. Each feature has its dedicated table of perceptron weights which it indexes into. Hence number of tables equals the number of features. Once the perceptron weights from all the tables are obtained, they're summed and thresholded based on a pre-decided value. This led to a prediction mechanism that is highly accurate, has quick learning rate and is adaptable to changing program behaviour. The features introduced in that paper were derived from Program Counters of the current and the last few instructions and the memory address of the current access. Even with such basic features, high level of correlation could be established by perceptrons.

The work was further extended by Multiperspective Reuse Prediction[cite], which introduced a varied set of parametric features. Although design space exploration for feature and parameter selection was a non-trivial task, once fixed, it could give a cache replacement predictor highly suitable for the given set of applications.
\newline
\newline
\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron prediction involves multiplying the vector of input features: F\textsubscript{1xN}, with the corresponding weight vector: W\textsubscript{Nx1} in a dot product fashion to obtain the sum y\textsubscript{out}. Here what we use is a perceptron-like structure. The feature is used to hash into the weight of perceptrons and the retrieved weights are added straight away. This can be seen a dot product of $\vec{1}$\textsubscript{1xN} with W\textsubscript{Nx1}. This way, the perceptron algorithm doesn't introduce any multiplication operations in the inference or training process. Hence what we adapt in this work, is a perceptron-like learning algorithm as it involves the same principle involved in perceptron inferencing and training. For simplicity, we'll call our implementation as perceptron implementation.
\newline
\newline
\textit{Perceptron Update Rule:} In all the above implementations, a uniform perceptron update principle is being followed. The weights need to be updated if the prediction was wrong or the predicted sum does not cross a certain threshold. After a certain training period, these weights are proportional to the probability of the outcome being true. This training threshold makes sure that the perceptrons are trained till a certain level of confidence and yet they are not over-trained to the given set. Care is taken that the weights saturate at certain positive and negative values so that they remain confined to the bit-width. The same rules will be applied in the discussion done in section 3.2 on SPP-Perc learning algorithm.

\subsection{Signature Path Prefetcher Architecture}
Since the work we've done builds heavily on SPP[cite], we're presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work. Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists of following structural components:
\newline
\newline
\textbf{Signature Table:} ST keeps track of 256 most recently accessed pages. It is meant to capture memory access patterns within a page boundary. SPP indexes into an entry of ST using the page number. For each entry corresponding to a page, ST stores the 'last block offset' and 'old signature'. Last block offset is the block offset of the last memory access of that given page. The block offset is calculated with respect to the page boundary. The signature is a 12-bit compressed representation of the past few memory accesses for that page. The signature is calculated as:
$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
In case a matching page entry is found, the stored signature in the corresponding entry is retrieved.
\newline
\newline
Consider the case that incoming page number 10 with a block offset 3 finds a match in the ST. The retrieved pattern signature is 0x30 and the Last Offset is 1. Since now there is the Last Block Offset (1), Incoming Block Offset (3), Old Signature (0x30), and New Signature calculated as per above equation (0x182), SPP can infer non-speculatively that the given pattern of memory accesses (as captured in the Old Signature) leads to the particular Delta. In general, Delta is defined as the difference between the prefetch suggestion block and the initial block which triggered the prefetch. In this case, since SPP is in learning phase, it is defined as the difference between the Incoming Block Offset and the Last Block Offset (+2 in this case). This, in turn generates the new memory pattern (New Signature). This newly learned signature pattern and the delta is stored in the Pattern Table.
\newline
\newline
\textbf{Pattern Table:} The PT is indexed by the signature generated from ST. PT holds predicted delta patterns and their confidence estimates. Each entry indexed by the signature holds up to 4 unique delta predictions. This is implemented by making PT as a 4-way associative table. To measure an estimate of a delta confidence, PT implements two counters: C\textsubscript{sig} for each entry indexed by the signature and C\textsubscript{delta} for each delta entry for a given signature. More details of this on-line score-keeping mechanism is described under Confidence Tracking discussion.

\textit{Lookahead Prefetching:} For each trigger to the prefetcher, the first time that the SPP indexes into the PT, it does so in the non-speculative manner. We call it non-speculative because the parameters which SPP uses to predict, like last block offset, incoming block offset and signature are well-defined. Assuming PT generates a valid delta prediction (rather, up to 4 valid deltas), SPP enters into the lookahead mode.

In the lookahead mode, SPP looks at the last generated delta(s), chooses the one with the highest confidence and speculatively generates the new effective address and the new confidence. Using these new parameters, it re-accesses the PT to generate further prefetches. SPP keeps on repeating this cycle of accessing PT and updating the signature based on highest confidence prefetch from the last iteration. These number of iterations are characterized by the 'depth' till which SPP manages to predict prefetch entries in the lookahead manner. While doing this, SPP also keeps compounding the confidence in each depth. Thus as depth increases, overall confidence keeps decreasing. This makes sense since we are going deeper into the speculation path. When the confidence falls below a pre-defined threshold, SPP stops further lookahead iterations
\newline
\newline
\textbf{Global History Register:} One of the innovative ideas introduced in SPP was GHR, which bootstraps learning across page boundaries. GHR is a small 8-entry fully associative structure.

Consider a scenario where the prefetch delta suggested by block offset 60 in page A is +6 and at that instant say the signature was 0x25. It means that the prefetch crosses page boundary. Instead of discarding this as a useless prefetch attempt, SPP stores the signature (0x25), last offset (60), delta (+6) and the confidence (say 0.72) in the GHR. This corresponds to block offset of 66 (=60 + 6). Since there are only 64 blocks per page, block 66 corresponds to 66 \% 64 = 2 \textit{i.e.}, block offset 2 in the new page. This new page can be any arbitrary new page which has not yet been caches in the Signature Table.
This is demonstrated in Fig. \_\_ <insert GHR Figure>

Now suppose that after some arbitrary number of cycles, first prefetch request from block 2 in page B is initiated. This will lead to a miss in ST as no signature corresponding to page B exists in ST. Instead of not attempting any prefetch, SPP will search the GHR for a possible match. It does so by calculating the effective offset pointed by the last offset and the delta entries in the GHR. It sees that the signature 0x25 has the last offset of 60 and predicted delta as +6 and that corresponds to effective block offset of +2. Since this matches the block offset of page B that saw a miss in the ST, SPP will retrieve the corresponding signature (0x25) from the GHR and proceed to form the new signature: with 0x25 and +2 as the inputs \textit{i.e.}, 0x12A. SPP continues prefetching using this signature and also stores this in ST corresponding to page B. <Insert GHR image>
\newline
\newline
\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch Filter - a 1024 entry 1-way associative table which keeps a record of last few entries prefetched by SPP. This proves useful for following reasons:
\begin{itemize}
    \item \textbf{Eliminating duplicate prefetches}\newline 
    Any subsequent prefetches recommended by SPP mechanism is checked against the existing entries in Prefetch Filter. If a valid entry is found, it means that the particular cache line is already either in L2 Cache or being fetched. Hence there is no need to issue a duplicate prefetch.

    \item \textbf{Updating prefetcher}\newline
    If a subsequent demand access comes before eviction for a prefetch that was recommended by SPP, various components of SPP are updated to reflect that this was a good prefetch. Correspondingly, if a prefetched cache line is evicted without being demand hit even once, SPP learns that this was a useless prefetch and updates the various tables accordingly

    \item \textbf{Tracking Global Accuracy}\newline
    PF keeps a track of usefulness of the prefetches by introducing 2 counters: C\textsubscript{total} to track the total number of prefetches getting recorded in the PF and C\textsubscript{useful} to keep a track of which prefetches recorded in the PF led to a demand hit. The ratio C\textsubscript{useful} / C\textsubscript{total} is the global accuracy for the prefetching scheme. It is represented as $\alpha$ in the paper. The global accuracy acts a real-time throttler for the path confidence value as depicted in the following discussion.
\end{itemize}
\textbf{Confidence Tracking}: The PT keeps a track of hits to each signature through a counter C\textsubscript{sig}. The number of hits for a given delta per signature are tracked using a counter C\textsubscript{delta}. The confidence for a given delta is approximated through C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}. When SPP enters into a lookahead mode, the path confidence P\textsubscript{d} is given as:
$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown in the Prefetch Filter section. The range of global accuracy is in [0,1]. The lower the value of $\alpha$ for a given trace or a program phase, the overall confidence is scaled down and this makes SPP less aggressive for the upcoming prefetches.

Here d is the lookahead depth. For d = 1, when SPP is in non-speculative mode, P\textsubscript{0} can be thought of as 1. Once SPP enters into speculative mode, it calculates the path confidence as a product of confidence of the individual prefetches encountered along that path, further scaled using $\alpha$. The final P\textsubscript{d} is thresholded against prefetching-threshold (T\textsubscript{p}) to decide whether to prefetch or not. If yes, then P\textsubscript{d} is thresholded against a numerically bigger fill-threshold (T\textsubscript{f}) to decide whether the prefetch should be sent to L2 Cache (high confidence prefetch) or Last Level Cache (low confidence prefetch)
\newline<Insert a comprehensive figure about SPP structures>
\newline<Insert another detailed picture about SPP data-path flow>

\subsection{Case for an On-line Throttler}
As compared to some of the other state of the art prefetchers [BOP], SPP is a far less aggressive prefetcher. In the single core environment, this gives BOP an edge as there is no resource contention among the different cores. Hence more aggressive prefetching is bound to prove beneficial. As we increase the core count, we observe that SPP starts outperforming rest of the prefetchers. This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. <Figure showing the variation of SPP vs BOP wrt core count>. For 4-core applications, SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.

The above analysis shows that with a more careful throttling mechanism, SPP can be tuned to become much more aggressive, leading to increased coverage. The onus of maintaining the accuracy now falls on the independent throttler. To test the hypothesis, we tuned down SPP to the minimum possible threshold. Only by doing this we managed to increase the SPP coverage by XX\%. Obviously this came at a cost of increased DRAM traffic and cache pollution. To that effect, we introduce an independent on-line learning perceptron based throttling mechanism.

\section{SPP-Perc}

\subsection{Changes made to original SPP}
To modify the SPP Design to suit our scheme, the following changes were made:

\begin{itemize}
\item \textbf{FILL\_THRESHOLD  discarded}\newline
In the original SPP design, internal confidence mechanism was developed. That confidence was compared against a preset FILL \textunderscore THRESHOLD (T\textsubscript{f}) to decide whether a prefetch should be sent to L2 Cache or Last Level Cache. Its value was 90 on the scale of 0 to 100. In SPP-Perc, the perceptron sum is used to make that decision. Hence FILL\_THRESHOLD was no longer needed

\item \textbf{PF\_THRESHOLD reduced}\newline
SPP design used this threshold to decide whether to do prefetch or not. This threshold (T\textsubscript{p}) was set to 25. Since the new perceptron mechanism is used to make this decision now, this was reduced all the way down to 1. It was not set to 0 for preventing non-throttled lookahead running. SPP confidence mechanism multiplies the confidence of each lookahead depth with the previous one to get the cumulative confidence - on the integer scale of 0 to 100. It was observed that once the lookahead depth falls below one, it essentially becomes 0 since we're limiting the multiplication to integer only. Confidence multiplication for subsequent lookahead depths still yield 0 and prefetching keeps happening. This causes a lot of unnecessary prefetches, leading to cache pollution. To mitigate this, the PF\_THRESHOLD was set to the next possible value i.e. 1.

\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
Original SPP would not consider the slots available in the L2 MSHR queue. It had an assumption that at no time can the number of prefetches suggested by SPP could exceed the capacity of the queue. While it was not proven in the paper, the internal confidence mechanism along with the thresholds made sure that the above assumption was maintained. In SPP-Perc since there can be no assumption on the degree of aggressiveness - because SPP is being tuned to its lowest thresholds, we had to explicitly make sure that at no point do the number of suggested prefetches exceed L2 MSHR queue.

\item \textbf{Enhanced Filter}\newline
Original SPP introduces a concept of Prefetch Filter, which keeps a track of up to 1024 L2 Cache prefetches suggested by SPP. The Filter was modified to store meta-data like program counter, memory access address etc. This is needed to train the perceptron when the result of that particular prefetch gets available. At training time, this data is used to retrieve the state of the prefetcher when the prediction was made.

\item \textbf{Reject Filter introduced}\newline
The Prefetch Filter mentioned above keeps a track of the prefetches that actually happened in the L2 Cache. Correspondingly, we introduced a reject filter. It keeps a track of the prefetches that were suggested by the base SPP engine but were rejected by the perceptron. In case it is detected that a future demand fetch could have used this rejected prefetch, the perceptron weights are updated to reflect the same

\end{itemize}

\subsection{The Perceptron Layer}
As discussed earlier, the perceptron layer was introduced to act as a throttler to the underlying SPP engine. This section explores the details of the perceptron layer and the range of features that are ultimately used by the perceptron for correlating the prefetches.

The neural network weights are stored in a 12-bit indexed (4096 entries deep) table. A single entry in the table is a 5 bit counter - saturating at -16 or +15. Each feature gets its own dedicated table. Our proposed design uses 17 features. The design space exploration and the intiotion behind these features is explained in section\_\_. Hence there are 4096 * 17 5-bit entries which hold the weights of the perceptron layer. As the program execution begins, all the weights are initialized to 0.
\newline
\newline
\textbf{INFERENCING}\newline
In the Champsim simulator, the prefetching method is triggered at L2 Cache demand access. At that point, the prefetcher has the option to do or not do prefetch - in case it does, then how many cache lines to prefetch. The suggested prefetches can be places either in L2 Cache or in Last Level Cache. The prefetcher gets to decide this, and is usually done based on an internal confidence mechanism.

When SPP gets triggered, it suggests up to 4 (=Associativity of the Pattern Table) candidates for prefetch based on the demand access address. In each subsequent lookahead depth, SPP suggests up to 4 more prefetch candidates, as explained in section\_\_. All these recommended prefetches are tested in the perceptron throttler before they can access the Prefetch Filter, as in the original SPP data-path.

The perceptron looks at the micro-architectural state, or 'features', at that instant. Each feature indexes into the hash of 4096 entry table dedicated to itself. Feature \#1 indexes into a particular entry in table \#1 and so on. The choice of each feature and the intuition behind them is discussed in section 4.5.

Once all the weights are retrieved, they are added together, much like the traditional perceptron dot product. On obtaining the overall sum, it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO. Only the prefetch candidates with the perceptron sum higher that the threshold qualify for prefetching. The prefetches that qualified the perceptron throttler stage go through the PF stage, just as in SPP design. In SPP-Perc, all the meta-data that is required to recreate the features is also stored in the PF. At a later stage, when the feedback of the current prefetched line is available, the stored data can be used to train the perceptron.

In addition to the Prefetch Filter mentioned above, SPP-Perc also maintains a 'Reject Filter'. Reject filter is a 1024-entry deep 1-way associative table. In case that a prefetch suggestion is rejected by perceptron throttler, it instead gets logged into the 'Reject Filter'. RF keeps track of 1024 <can be reduced considerably> entries that were suggested by the Pattern Table but got rejected by the perceptron throttler. This can be used to train the perceptron to avoid false negatives \textit{i.e.}, cases where prediction was to reject the prefetch but the prefetch would have been useful.
\begin{table}[]
    \centering
    \begin{tabular}{|l|c|p{4.6cm}|}
    \hline
    Field &
    Bits &
    Comments \\
    \hline
         Valid & 1 & Indicates a valid entry in PF\\
         Tag & 6 & Identifier for the entry in PF\\
         Useful & 1 & To show if the given entry led to a useful demand fetch\\
         PC & 12 & Features for perceptron training \\
         Address & 30 & \\
         Delta & 7 & \\
         Curr Signature & 12 & \\
         Last Signature & 12 & \\
         Confidence & 7 & \\
    \hline
    \end{tabular}
    \caption{Meta-data Stored in PF}
    \label{tab:PF_metadata}
\end{table}

The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch. A high confidence prefetch is represented by a higher perceptron sum; and vice-versa. We introduce another threshold - PERC\_THRESHOLD \_HI. Any prefetch with the perceptron sum higher than this threshold becomes a L2 Cache prefetch. Any prefetch with the perceptron sum lower than this threshold goes to Last Level Cache.

The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI were empirically set. It is ensured that numerically PERC\_ THRESHOLD\_HI is greater than or equal to PERC\_THRESHOLD\_Lo.

As previously stated, an enhancement to the original SPP design includes storing all the meta-data describing the program state at the instant of prefetching, in the Prefetch Filter. This information is useful when the perceptron needs to be updated subsequently. Table \ref{tab:PF_metadata} depicts the meta-data stored for each entry in the Prefetch Filter.\newline
\newline
\textbf{TRAINING}\newline
In the prefetching environment, feedback of a prefetch is received whenever there is an eviction or a demand access from L2 Cache. This acts as a trigger to update all the baseline SPP entries and to also train the perceptron throttler. Training of the perceptron throttler involves restoring the meta-data that was stored in Prefetch Filter or the Reject Filter. This is done by indexing into the corresponding filters using the cache line address of the block being used to train. (10-bits indexing + 6-bits tag matching). Once the state of the program at the time of prefetching is available, it is used to index into the perceptron weights table.

If a demand access triggers the training and the block was tagged as a valid prefetch in the Prefetch Filter, it means that the earlier prefetch prediction was correct. In that case the perceptron weights are incremented by 1 if the predicted sum does not cross a pre-defined threshold.  In case it is a cache block eviction that led to training and the corresponding valid entry was found in the Prefetch Filter, it shows that the prediction made by the perceptron was wrong. Perceptron should have ideally rejected the prefetch suggestion as a low-confidence prefetch. Here the weights are decremented by 1 to reflect the same. In either case, weights are saturated at -16 or +15.

A secondary training mechanism also kicks in during demand fetches. Before the demand access triggers the next set of prefetches, it checks the Reject Filter for the valid entry. In case it is a hit, that means that the corresponding cache line was initially suggested by the underlying SPP engine but rejected by the perceptron throttler. This means that perceptron should have been more confident about that particular prefetch. Once such a scenario is identified, the state of the execution at the time of prefetch is retrieved from the Reject Table. The retrieved data is used to index into the various weights tables of the perceptron and the corresponding values are updated by +1, saturating between and -16 and +15. This update reflects increased confidence for the prediction corresponding to that prefetch.

\subsection{Features used by Perceptron}

Here we discuss the various features which help correlate some aspect of prefetching with the program behaviour
\begin{itemize}
\item Base address of the demand access which triggered the prefetch. Since prefetch especially tend to be triggered from the L2 demand access, they tend to be correlated with the triggering address.
\item The cache line: It is derived from the base address as base\_addr >> LOG2\_BLOCK\_SIZE
\item Page address : Calculated as base\_addr >> LOG2\_PAGE\_ SIZE
\item Current Signature : The 12-bit signature used by SPP to index into the Pattern Table
\item Confidence : The integer confidence on the scale of 0 to 100 which was used in the original SPP design
\item Delta : The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched.
\item Last Signature : SPP keeps a track of per-page last signature from which it derives the current signature
\item Program Counter of the Load instruction that triggers that particular prefetch
\item Program Counter of the previous i instructions that access LLC, shifted right by an increasing number. This can be represented as: PC\textsubscript{i} >> i, i in \{1,2,3\}. The rational here is that shifted bits of the previous PCs allow the perceptron to have an approximate idea of the neighbourhood of the current instruction. By right shifting the PCs which look back further into the past, we can gain a wider perspective of the past events. Example: rather than correlating with PC\textsubscript{2}, this allows SPP-Perc to correlate program behaviour with the 4-byte region around PC\textsubscript{2}.
\item PC\textsubscript{1} XOR PC\textsubscript{2} XOR PC\textsubscript{3}: This tells SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block.
\item Base address XOR Delta : This feature captures if any particular delta is highly favoured by a given trigger address
\item Cache line XOR Delta : This feature captures if any particular delta is highly favoured by a given cache line address
\item Page address XOR Delta : This feature captures if any particular delta is highly favoured by a given page number
\item Confidence XOR Delta : This gives a combined insight into that fact that certain deltas might be predicted by SPP and being high confidence ones
\item Current Signature XOR Page Address : This correlates the fact that some pages might favour a particular SPP signature
\item Program Counter XOR Delta : This tells us if a given PC favours a particular value(s) of delta
\item Base address XOR Page address
\item Current Signature XOR Delta
\item Confidence XOR Page address
\end{itemize} 
As noted above, each feature indexes into its independent entry of perceptron weights.

\section{EVALUATION}
\subsection{Performance Model}
For testing and comparing SPP-Perc, we use Champsim[cite] simulator. Champsim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship[cite]. We model model a 1-core / 4-core / 8-core Out of Order machine with the following configuration:

\begin{itemize}
    \item Cache block size: 64 KB
    \item 8 way separate 32 KB L1 I-Cache and D-Cache
    \item 8 way 256 KB L2 Cache
    \item 16 way 2 MB / core shared Last Level Cache
    \item 4 GB Single Channel DRAM for single-core and 8 GB Double Channel DRAM for multi-core
\end{itemize}
Prefetching was only triggered at L2 Cache demand accesses but could be sent to L2 Cache or Last Level Cache. No L1 Data level prefetching was done. LRU replacement policy was used on all levels of cache hierarchies. Branch prediction was done using a gshare branch predictor <Try perceptron predictor here>. Page size was kept as 4KB. Champsim operates all the prefetchers strictly in the physical address space.

\subsection{Testing Under Additional Memory Constraints}
The default single-core configuration simulates a 2MB LLC and a single channel DRAM with 12.8GB/s bandwidth. We extend the simulations to include memory constraints introduced in DPC-2. Specifically we look at the following two variations:
\begin{itemize}
    \item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited to 3.2 GB/s
    \item \textit{Small LLC}: In this scenario, LLC size is reduced to 512 KB
\end{itemize}
All the multi-core simulations are only done in the default configuration.

\subsection{Workloads}
This is the first time that SPEC 2017 benchmark suite[cite] has been used to characterize and measure the prefetch performance. We use all the 20 workloads available in the SPEC 2017 Suite. Using Simpoint[cite] methodology, we identified program segments of 1 Billion instructions each.

Besides SPEC 2017, we also used server level workloads from CouldSuite[cite]: <insert the workloads which should be tested>. <Insert mlpack?> In total, we had XX traces <95 currently> representing workloads across XX benchmarks. <20 currently>

\textit{Single-core performance} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and the next 1 Billion instructions to do detailed simulations and collect run-time statistics. We report the performance as speedup over baseline \textit{i.e.}, no prefetching. We take the Instruction Per Cycle (IPC) of the prefetcher and divide it with the IPC of baseline. The final number reported is the geometric mean of the speed-up attained on individual traces.

\textit{Multi-core performance} For 4-core multi-programmed workloads, we generated 20 unique workloads comprising of a mix of memory intensive traces from SPEC 2017. For each workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. As different workloads have different baseline IPCs, not all of the 4 workloads reach the 1 Billion instruction count at the same time. In that case, the workloads get wrapped-over and are still simulated till the last core completes 1 Billion instructions after warm-up. For collecting IPC and other data, only the first billion instructions are considered as "region of interest". Here we report the weighted speedup normalized to baseline \textit{i.e.}, no prefetching. For each of the workloads running on a particular core of the 4-core 8 MB LLC system, we compute IPC\textsubscript{i}. We then find the IPC\_isolated\textsubscript{i} of the same workload running in isolated 1-core 8 MB LLC environment. Then we calculate the total weighted-IPC for a given workload mix as $\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}). For each of the 20 workload-mix, the sub obtained is normalized to the weighted-IPC calculated similarly for baseline case \textit{i.e.}, no prefetching, to get the weighted-IPC-speedup. Finally the geometric mean of these 20 weighted-IPC-speedup is reported as the effective speedup obtained by the prefetching scheme.

\subsection{Preferchers Simulated}
We compared SPP-Perc against three of the latest state of the art prefetchers: Best Offset Prefetcher (BOP)[cite], DRAM Aware - Access Map Pattern Matching (DA-AMPM) [cite] and Signature Path Prefecher (SPP)[cite]. BOP was the winner in 2nd Data Prefetching Championship[cite]. DA-AMPM is the enhanced version of AMPM[cite], modified to account for DRAM row buffer locality. SPP has been shown to outperform BOP on SPEC 2006 traces[cite]. For each of these, we compare their speedups taking no prefetching as the baseline.

\subsection{Overhead for SPP-Perc}
In this section, we analyze the hardware overhead required to implement SPP-Perc. The exiting structires from SPP - Signature Table, Pattern Table, Global History Register and the Prefetch Table occupy 44060 bits. The Prefetch Filter was enhanced to accommodate storing of meta-data for perceptron training. New Prefetch Filter now consumes 95232 bits. The Reject Filter occupies 89088 bits. The perceptron weights occupy 348160 bits. All in all, SPP-Perc occupies 568348 bits, which comes out to 69.3 KB.

<Inset a neat table showing analysis of area overheads of individual components>

The hardware budget for 2nd Data Prefetching championship was 32 KB. Keeping that in mind and the speedup SPP-Perc obtained over the winner, the extra hardware budget can be accounted for.

\subsection{Developing Features for SPP-Perc}
This section describes the intuition and analysis that went behind finalizing the features.

\section{RESULTS}
This section discusses the results obtained from running SPP-Perc, in terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per Cycle (IPC) metrics. First we present the results for single-threaded workloads then the results for multi-core workloads are presented.

\subsection{Single-core Results}
Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated. All the results have been normalized to baseline \textit{i.e.}, no prefetching. As can be seen, SPP-Perc yields a speedup of XX\% over the baseline. This is equivalent to XX\% over DA-AMPM, xx\% over BOP and XX\% over SPP. Out of the XX benchmarks, SPP-Perc nearly matches or outperforms most of the prefetchers on XX traces.

At its peak, SPP-Perc manages to get the IPC speedup of a factor of over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}. This also corresponds to speedup gain of \textbf{22.4\%} over the next best prefetcher - SPP. In general, benchmarks \textit{602.gcc\_s, 603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s} benefit the most from SPP-Perc, as compared to SPP only prefetching.

Figure \_\_ shows the performance of SPP-Perc against the other prefetchers in tighter constraint environments. For sake of simplicity, we show 2 traces where SPP-Perc gives the most advantage over baseline, 2 traces where SPP-Perc doesn't perform as well as the best prefetcher *<-replace with the prefetcher name once results are available* and finally GeoMean across all the traces. <Hopefully BOP would degrade more here as it is super aggressive. By definition, SPP-Perc should be better than SPP as the throttler would better control the aggressiveness. In SPP paper DA-AMPM performs close enough to BOP. That should get replicated here too.>

<Discuss the extreme behaviour of \textit{623.xalancbmk\_s}: Improvement over SPP is as high as 31\% and even reduces to -1.5\% on traces of this benchmark>

<Discuss about coverage on all traces>

<Discuss cases where coverage gets specially increased>

<Discuss accuracy on all traces>

<Discuss about compelling increase in accuracy - details of traces. Plot linear separability>

<Any case where accuracy got reduced - go into the trace. See if linear separability is the issue. Other possible issues?>

\subsection{Multi-core Results}
In this section, we demonstrate the improvement achieved by SPP-Perc for a mix of multi-programmed workloads. 
\newline \newline
\textit{4-core environment}
Figure \_\_ shows a comparison of speedups obtained on workload mix of memory intensive subset of SPEC 2017. We plot all the 4 prefetchers, normalized to baseline. The workloads have been sorted in the increasing order of performance benefit, as measured in terms of normalized weighted speedup. As can be seen, SPP-Perc offers an IPC improvement by 33.5\% on these traces. This is 5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP and DA-AMPM respectively.

This increased improvement by SPP-Perc on a 4-core machine is not surprising. Since SPP-Perc has an extremely intelligent throttler working to eliminate useless prefetches before they can create cache pollution. This further amplifies the edge that SPP had over BOP. With the advantage of an orthogonal prefetcher, SPP can be as aggressive as possible and leave the accuracy aspect to perceptron throttler.
\newline \newline
\textit{8-core environment}
To test the effectiveness of SPP-Perc in a tighter constrained multi-core environment, we simulated an 8-core machine with 16 MB last Level Cache. Other parameters were kept same as the ones discussed in Methodology section.

\section{CONCLUSION}
In this paper, we introduced perceptron based learning for data prefetching in the form of SPP-Perc. Perceptron acts an orthogonal throttler to the underlying prefetch engine. SPP-Perc improves performance by over XX\% over the baseline, which corresponds to YY\% over the next best performing prefetcher. We believe this paper can lead to interesting future work like generalizing SPP-Perc to work with any prefetcher with minimal knowledge of the underlying prefetcher architecture.

%%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%

%%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
%%\bibliographystyle{ieeetr}
%%\bibliography{ref}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2018-10-19 paper_1.tex
Latex source update

@@
 In this paper, we introduce perceptron learning to help make this prefetching decision efficiently.
-We've taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine.
+We have taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine.
 The perceptron layer filters out the unnecessary prefetches recommended by SPP.
-We've also explored a range of features that can be used to train the perceptron layer.
+We have also explored a range of features that can be used to train the perceptron layer.
 Our results show that SPP-Perc improves performance on SPEC 2017 Benchmark suite by 3.2\% on single-core and 5.5\% on 4-core simulation, as compared to just SPP.
@@
 An ideal prefetching scheme involves capturing the memory access patterns to predict future prefetches in a timely manner.
-Other important decisions are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block.
+Other important decisions faced by the prefetcher are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block.
 Memory accesses patterns can be as simple as fetching the next cache line or a complex but predictable pattern, like in pointer chasing applications.
@@
 To this effect we propose <come up with a catchy name> <SPP-Perc>.
-It's an enhancement to the existing state of the art prefetcher and overcomes the coverage vs trade-off issue in a robust manner.
+It's an enhancement to the existing state of the art prefetcher and overcomes the coverage vs accuracy trade-off in a robust manner.
 The idea is to modify SPP design to make it as aggressive as it can get.
@@
 \item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind.
-We've used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers.
-We've also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
+We have used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers.
+We have also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
 
@@
 Hence what we adapt in this work, is a perceptron-like learning algorithm as it involves the same principle involved in perceptron inferencing and training.
-For simplicity, we'll call our implementation as perceptron implementation.
+For simplicity, we will call our implementation as perceptron implementation.
 \newline
@@
 \subsection{Signature Path Prefetcher Architecture}
-Since the work we've done builds heavily on SPP[cite], we're presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work.
+Since the work we have done builds heavily on SPP[cite], we are presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work.
 Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists of following structural components:
@@
 SPP confidence mechanism multiplies the confidence of each lookahead depth with the previous one to get the cumulative confidence - on the integer scale of 0 to 100.
-It was observed that once the lookahead depth falls below one, it essentially becomes 0 since we're limiting the multiplication to integer only.
+It was observed that once the lookahead depth falls below one, it essentially becomes 0 since we are limiting the multiplication to integer only.
 Confidence multiplication for subsequent lookahead depths still yield 0 and prefetching keeps happening.
@@
 Our proposed design uses 17 features.
-The design space exploration and the intiotion behind these features is explained in section\_\_.
+The design space exploration and the intuition behind these features is explained in section\_\_.
 Hence there are 4096 * 17 5-bit entries which hold the weights of the perceptron layer.
@@
 
-If a demand access triggers the training and the block was tagged as a valid prefetch in the Prefetch Filter, it means that the earlier prefetch prediction was correct.
+If a demand access block which triggers the training  was tagged as a valid prefetch in the Prefetch Filter, it means that the earlier prefetch prediction was correct.
 In that case the perceptron weights are incremented by 1 if the predicted sum does not cross a pre-defined threshold.

### 2018-11-13 paper_1.tex
Tex file: Multiple lines per paragraph
Version-2
- Tables for system config and area overheads
- Statistical analysis added
- General revision / text fill-ins

@@
 \usepackage{array}
+\usepackage{multirow}
 
@@
 \begin{abstract}
-
-Hardware prefetching has been introduced in modern processors as a way to reduce compulsory cache misses.
+\textbf{Hardware prefetching has been introduced in modern processors as a way to reduce compulsory cache misses. 
 An efficient prefetcher should be able to identify complex memory access patterns. 
 This enables it to fetch a block ahead of its demand access. 
+}
 
-In this paper, we introduce perceptron learning to help make this prefetching decision efficiently.
+\textbf{In this paper, we introduce perceptron learning to help make this prefetching decision efficiently. 
 We have taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine. 
-The perceptron layer filters out the unnecessary prefetches recommended by SPP.
+The perceptron layer acts as a throttler filters out the unnecessary prefetches recommended by SPP. 
 We have also explored a range of features that can be used to train the perceptron layer. 
-Our results show that SPP-Perc improves performance on SPEC 2017 Benchmark suite by 3.2\% on single-core and 5.5\% on 4-core simulation, as compared to just SPP.
-
+Our results show that SPP-Perc improves performance on the memory intensive subset of the SPEC 2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared to just SPP.
+We also demonstrate that the performance gained from using our efficient throttler scales better with increasing number of cores.
+}
 \end{abstract}
@@
 
-To this effect we propose <come up with a catchy name> <SPP-Perc>.
-It's an enhancement to the existing state of the art prefetcher and overcomes the coverage vs accuracy trade-off in a robust manner.
+To this effect we propose SPP-Perc <come up with a catchy name>. 
+It is an enhancement to the existing state of the art prefetcher and overcomes the coverage vs accuracy trade-off in a robust manner. 
 The idea is to modify SPP design to make it as aggressive as it can get. 
 This enables SPP to capture complex memory access patterns hence leading to increased coverage.
+This comes at a cost of increased DRAM traffic, hence reduction in overall IPC.
 The prefetch suggestions recommended by SPP pass through the perceptron based throttler. 
@@
 
-\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and future scope for expansion. The contributions made by this paper are:
+\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and future scope for expansion. 
+The contributions made by this paper are:
 
 \begin{itemize}
-\item This is the first time where a truly on-line learning neural network based model has been used when it comes to prefetching.
-The prior works in this area either relied on program semantics [cite] or were application specific [cite] or could not yield a significant improvement over the baseline prefetcher used.
-[cite]
+\item This is the first time where a truly on-line learning based neural network model has been used when it comes to hardware data prefetching. 
+The prior works in this area either relied on program semantics [cite] or were application specific [cite] or could not yield a significant improvement over the baseline prefetcher used. [cite]
 
-\item Coming up with a robust and accurate prefetching mechanism which can yield increased accuracy as compared to the other state of the art prefetchers
+\item Implementing the pereptron throttler in a robust and accurate practical prefetching mechanism. 
+The overall scheme yields increased accuracy as compared to the other state of the art prefetchers.
+Moreover, the perceptron throttler learns to adapt itself to shared resource constraints - leading to an even increased performance in multi-core and lower capacity / bandwidth environments.
 
-\item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind.
-We have used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers.
-We have also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
+\item Examining the perceptron weights doing a statistical study to quantify the contribution made by each feature.
+We have also used this idea to develop a comprehensive set of features where each feature contributes some 'new' information to the perceptron.
+The details are explained under methodology, in Section XX.
 
-\item We came up with an efficient model to use the perceptron output as a representation of the prefetch confidence using an on-line histogram technique.
-We show how the simple sum can replace the existing mechanism which was essentially based on counting the number of hits for a particular entry in SPP.
+\item Cross-validating the design across benchmarks. 
+All the development and parameter tuning was done on the SPEC 2017 benchmark. 
+The results have been shown to be consistent across unseen benchmarks.
+More details can be found in the results section (Section XX).
+
+%\item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind. 
+%We have used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers. 
+%We have also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
+
+%\item We came up with an efficient model to use the perceptron output as a representation of the prefetch confidence using an on-line histogram technique. 
+%We show how the simple sum can replace the existing mechanism which was essentially based on counting the number of hits for a particular entry in SPP.
 
@@
 For evaluation, we used SPEC 2017 benchmarks. 
-On a single core configuration, SPP-Perc increases performance of the system by XX\% as compared to the baseline with no prefetching.
-This outperforms SPP by 3.2\%.
-On a 4-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement of XX\% over the baseline.
-This was ahead of SPP by 5.5\%.
-<Allude to 8 core results>
-<Write about other workloads like Cloudsuite, MLPack>
+SPEC 2006 and CloudSuite were used for cross-validation of the prefetcher implementation.
+On a single core configuration, SPP-Perc increases performance of the system by 39.5\% as compared to the baseline with no prefetching, on the memory intensive subset of the benchmark.
+This outperforms the next best prefetcher, SPP,  by 6.84\%. 
+On a multi-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement 5.95\% over SPP for 4-core system and 10.67\% for 8-core system.
+
+In the Results part in section XX, we show a detailed analysis of the speedup achieved and also talk about speedup achieved on SPEC 2006 and CloudSuite benchmarks.
 
@@
 \item Section 3 discusses enhancements made to the SPP design and the detailed implementation of the perceptron mechanism
-\item Section 4 consists of the methodology for evaluating SPP-Perc and exploring the feature space for perceptron learnng
+\item Section 4 consists of the methodology for evaluating SPP-Perc and exploring the feature space for perceptron learning
 \item Section 5 comprises results and evaluation in terms of IPC, MPKI, coverage and DRAM traffic, for a variety of benchmark suites
@@
 This form of indexing makes sure that the hyper-plane learned by the perceptron weights is able to differentiate between linearly inseparable outcomes. 
-<Is it true?
-Not sure>
+<Is it true? Not sure>
 
@@
 \newline
-\textbf{Global History Register:} One of the innovative ideas introduced in SPP was GHR, which bootstraps learning across page boundaries. GHR is a small 8-entry fully associative structure.
-
-Consider a scenario where the prefetch delta suggested by block offset 60 in page A is +6 and at that instant say the signature was 0x25.
-It means that the prefetch crosses page boundary.
-Instead of discarding this as a useless prefetch attempt, SPP stores the signature (0x25), last offset (60), delta (+6) and the confidence (say 0.72) in the GHR.
-This corresponds to block offset of 66 (=60 + 6).
-Since there are only 64 blocks per page, block 66 corresponds to 66 \% 64 = 2 \textit{i.e.}, block offset 2 in the new page.
-This new page can be any arbitrary new page which has not yet been caches in the Signature Table.
-This is demonstrated in Fig. \_\_ <insert GHR Figure>
-
-Now suppose that after some arbitrary number of cycles, first prefetch request from block 2 in page B is initiated.
-This will lead to a miss in ST as no signature corresponding to page B exists in ST.
-Instead of not attempting any prefetch, SPP will search the GHR for a possible match.
-It does so by calculating the effective offset pointed by the last offset and the delta entries in the GHR.
-It sees that the signature 0x25 has the last offset of 60 and predicted delta as +6 and that corresponds to effective block offset of +2.
-Since this matches the block offset of page B that saw a miss in the ST, SPP will retrieve the corresponding signature (0x25) from the GHR and proceed to form the new signature: with 0x25 and +2 as the inputs \textit{i.e.}, 0x12A.
-SPP continues prefetching using this signature and also stores this in ST corresponding to page B.
-<Insert GHR image>
+\textbf{Global History Register:} One of the innovative ideas introduced in SPP was the Global History Register, which bootstraps learning across page boundaries.
+GHR is a small 8-entry fully associative structure. 
+Whenever a prefetch suggestion from the Pattern Table crosses page boundaries, instead of completely discarding it, GHR saves that information for inter-page learning.
+Since the SPP-Perc implementation in this paper doesn't affect GHR functionality in any ways, we won't be going into more details about GHR.
 \newline
@@
 As we increase the core count, we observe that SPP starts outperforming rest of the prefetchers. 
-This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution.
-<Figure showing the variation of SPP vs BOP wrt core count>.
+This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. <Figure showing the variation of SPP vs BOP wrt core count>. 
 For 4-core applications, SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
@@
 Obviously this came at a cost of increased DRAM traffic and cache pollution.
-To that effect, we introduce an independent on-line learning perceptron based throttling mechanism.
+
+Moreover, the on-line confidence mechanism introduced in SPP was very rudimentary.
+It was based on taking a ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig} as explained above.
+The same confidence was used to make the decision whether to prefetch or not to prefetch; and which level to prefetch.
+While this approximation was shown to work in the original SPP, we believe that a better form of on-line decision making was possible.
+Hence, it was necessary to build a robust and adaptable learning mechanism to accept / reject the prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level Cache).
+
+To that effect, we introduce an independent on-line perceptron based throttling mechanism.
 
@@
 \begin{itemize}
-\item \textbf{FILL\_THRESHOLD  discarded}\newline
-In the original SPP design, internal confidence mechanism was developed.
-That confidence was compared against a preset FILL \textunderscore THRESHOLD (T\textsubscript{f}) to decide whether a prefetch should be sent to L2 Cache or Last Level Cache.
-Its value was 90 on the scale of 0 to 100.
-In SPP-Perc, the perceptron sum is used to make that decision.
-Hence FILL\_THRESHOLD was no longer needed
-
-\item \textbf{PF\_THRESHOLD reduced}\newline
-SPP design used this threshold to decide whether to do prefetch or not.
-This threshold (T\textsubscript{p}) was set to 25.
-Since the new perceptron mechanism is used to make this decision now, this was reduced all the way down to 1.
-It was not set to 0 for preventing non-throttled lookahead running.
-SPP confidence mechanism multiplies the confidence of each lookahead depth with the previous one to get the cumulative confidence - on the integer scale of 0 to 100.
-It was observed that once the lookahead depth falls below one, it essentially becomes 0 since we are limiting the multiplication to integer only.
-Confidence multiplication for subsequent lookahead depths still yield 0 and prefetching keeps happening.
-This causes a lot of unnecessary prefetches, leading to cache pollution.
-To mitigate this, the PF\_THRESHOLD was set to the next possible value i.e.
-1.
+\item \textbf{Original Thresholds discarded}\newline
+In the original SPP design, an internal confidence mechanism was developed which was compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the prefetch suggestion.
+Furthermore, the accepted suggestions were thresholded against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache vs Last Level Cache.
+T\textsubscript{p} and T\textsubscript{f} were set to 25 and 90 respectively, on the scale of 0 to 100.
+In SPP-Perc, the perceptron sum is used to make these decisions. 
+Hence these two thresholds are no longer needed
 
@@
 
-The neural network weights are stored in a 12-bit indexed (4096 entries deep) table.
-A single entry in the table is a 5 bit counter - saturating at -16 or +15.
+The neural network weights are stored in separate tables for each feature.
+The indexing of each feature is done using different number of bits.
+At most 12 bits of a feature are used to index into the perceptron table.
+Certain features require more resolution power \textit{i.e.}, full 12 bits of indexing.
+On the other hand, some features requiring lower resolution required as low as 7 bits of indexing.
+The variable indexing was determined by studying the features and fine-tuned empirically so as to achieve a good accuracy vs area trade-off.
+Exact details are mentioned under `Area Overheads' in Section XX.
+
+A single entry in the table corresponds to a perceptron weight.
+Each weight is a 5 bit counter - saturating at -16 and +15. 
 Each feature gets its own dedicated table. 
-Our proposed design uses 17 features.
-The design space exploration and the intuition behind these features is explained in section\_\_.
-Hence there are 4096 * 17 5-bit entries which hold the weights of the perceptron layer.
+Our proposed design uses 14 features. 
 As the program execution begins, all the weights are initialized to 0.
-\newline
-\newline
+\newline \newline
 \textbf{INFERENCING}\newline
-In the Champsim simulator, the prefetching method is triggered at L2 Cache demand access.
+In the Champsim simulator, the prefetching method is triggered on every L2 Cache demand access. 
 At that point, the prefetcher has the option to do or not do prefetch - in case it does, then how many cache lines to prefetch. 
@@
 The perceptron looks at the micro-architectural state, or 'features', at that instant. 
-Each feature indexes into the hash of 4096 entry table dedicated to itself.
+Each feature indexes into the hash of up to 4096-entry table dedicated for that feature. 
 Feature \#1 indexes into a particular entry in table \#1 and so on.
-The choice of each feature and the intuition behind them is discussed in section 4.5.
 
@@
 Only the prefetch candidates with the perceptron sum higher that the threshold qualify for prefetching. 
-The prefetches that qualified the perceptron throttler stage go through the PF stage, just as in SPP design.
-In SPP-Perc, all the meta-data that is required to recreate the features is also stored in the PF.
+The prefetches that qualify the perceptron throttler stage go through the Prefetch Filter stage, just as in the SPP design.
+The Prefetch Filter eliminates any redundant prefetch suggestions.
+In SPP-Perc, all the meta-data that is required to recreate the features is also stored in the Prefetch Filter. 
 At a later stage, when the feedback of the current prefetched line is available, the stored data can be used to train the perceptron.
 
-In addition to the Prefetch Filter mentioned above, SPP-Perc also maintains a 'Reject Filter'.
+The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch. 
+All prefetch candidates qualified till this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill level.
+The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
+
+The perceptron sum can also be seen as the sum of the individual contribution per feature.
+The value of each contribution corresponds to the extent of confidence of the final output with that given feature.
+By summing the individual contributions, the final perceptron sum denotes the overall confidence for that prefetch suggestion. 
+In a way, by thresholding the perceptron sum on two different thresholds, we are diving the confidence scale of the prefetch suggestion into three bins.
+The first bin corresponding to the lowest confidence leads to prefetch candidate being rejected.
+The next bin corresponds to prefetches with a moderate confidence level. 
+Such prefetches are directed to fill the bigger Last Level Cache and potentially not pollute the more scarce L2 Cache.
+The highest confidence prefetches are filled in the L2 Cache.
+
+In addition to the Prefetch Filter mentioned above, SPP-Perc also maintains a `Reject Filter'. 
 Reject filter is a 1024-entry deep 1-way associative table. 
-In case that a prefetch suggestion is rejected by perceptron throttler, it instead gets logged into the 'Reject Filter'.
-RF keeps track of 1024 <can be reduced considerably> entries that were suggested by the Pattern Table but got rejected by the perceptron throttler.
+In case that a prefetch suggestion is rejected by perceptron throttler, it instead gets logged into the Reject Filter. 
+The Reject Filter keeps track of 1024 entries that were suggested by the Pattern Table but got rejected by the perceptron throttler. 
 This can be used to train the perceptron to avoid false negatives \textit{i.e.}, cases where prediction was to reject the prefetch but the prefetch would have been useful.
-\begin{table}[]
-    \centering
-    \begin{tabular}{|l|c|p{4.6cm}|}
-    \hline
-    Field &
-    Bits &
-    Comments \\
-    \hline
-         Valid & 1 & Indicates a valid entry in PF\\
-         Tag & 6 & Identifier for the entry in PF\\
-         Useful & 1 & To show if the given entry led to a useful demand fetch\\
-         PC & 12 & Features for perceptron training \\
-         Address & 30 & \\
-         Delta & 7 & \\
-         Curr Signature & 12 & \\
-         Last Signature & 12 & \\
-         Confidence & 7 & \\
-    \hline
-    \end{tabular}
-    \caption{Meta-data Stored in PF}
-    \label{tab:PF_metadata}
-\end{table}
-
-The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch.
-A high confidence prefetch is represented by a higher perceptron sum; and vice-versa.
-We introduce another threshold - PERC\_THRESHOLD \_HI.
-Any prefetch with the perceptron sum higher than this threshold becomes a L2 Cache prefetch.
-Any prefetch with the perceptron sum lower than this threshold goes to Last Level Cache.
-
-The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI were empirically set. It is ensured that numerically PERC\_ THRESHOLD\_HI is greater than or equal to PERC\_THRESHOLD\_Lo.
 
@@
 This information is useful when the perceptron needs to be updated subsequently. 
-Table \ref{tab:PF_metadata} depicts the meta-data stored for each entry in the Prefetch Filter.\newline
-\newline
+\newline \newline
 \textbf{TRAINING}\newline
@@
 Training of the perceptron throttler involves restoring the meta-data that was stored in Prefetch Filter or the Reject Filter. 
-This is done by indexing into the corresponding filters using the cache line address of the block being used to train.
-(10-bits indexing + 6-bits tag matching).
+This is done by indexing into the corresponding filters using the cache line address of the block being used to train. (10-bits indexing + 6-bits tag matching). 
 Once the state of the program at the time of prefetching is available, it is used to index into the perceptron weights table.
@@
 
-Here we discuss the various features which help correlate some aspect of prefetching with the program behaviour
+Here we discuss the various features which help correlate the prefetching decision with the program behaviour
 \begin{itemize}
-\item Base address of the demand access which triggered the prefetch. Since prefetch especially tend to be triggered from the L2 demand access, they tend to be correlated with the triggering address.
-\item The cache line: It is derived from the base address as base\_addr >> LOG2\_BLOCK\_SIZE
-\item Page address : Calculated as base\_addr >> LOG2\_PAGE\_ SIZE
-\item Current Signature : The 12-bit signature used by SPP to index into the Pattern Table
-\item Confidence : The integer confidence on the scale of 0 to 100 which was used in the original SPP design
-\item Delta : The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched.
-\item Last Signature : SPP keeps a track of per-page last signature from which it derives the current signature
-\item Program Counter of the Load instruction that triggers that particular prefetch
-\item Program Counter of the previous i instructions that access LLC, shifted right by an increasing number.
-This can be represented as: PC\textsubscript{i} >> i, i in \{1,2,3\}.
-The rational here is that shifted bits of the previous PCs allow the perceptron to have an approximate idea of the neighbourhood of the current instruction.
-By right shifting the PCs which look back further into the past, we can gain a wider perspective of the past events.
-Example: rather than correlating with PC\textsubscript{2}, this allows SPP-Perc to correlate program behaviour with the 4-byte region around PC\textsubscript{2}.
-\item PC\textsubscript{1} XOR PC\textsubscript{2} XOR PC\textsubscript{3}: This tells SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block.
-\item Base address XOR Delta : This feature captures if any particular delta is highly favoured by a given trigger address
-\item Cache line XOR Delta : This feature captures if any particular delta is highly favoured by a given cache line address
-\item Page address XOR Delta : This feature captures if any particular delta is highly favoured by a given page number
-\item Confidence XOR Delta : This gives a combined insight into that fact that certain deltas might be predicted by SPP and being high confidence ones
-\item Current Signature XOR Page Address : This correlates the fact that some pages might favour a particular SPP signature
-\item Program Counter XOR Delta : This tells us if a given PC favours a particular value(s) of delta
-\item Base address XOR Page address
-\item Current Signature XOR Delta
-\item Confidence XOR Page address
+\item \textbf{Base address} of the demand access which triggered the prefetch. 
+Since prefetch especially tend to be triggered from the L2 demand access, they tend to be correlated with the triggering address.
+\item \textbf{Cache line}: It is derived from the base address as base\_addr >> LOG2\_BLOCK\_SIZE
+\item \textbf{Page address}: Calculated as base\_addr >> LOG2\_PAGE\_ SIZE
+\item \textbf{Current Signature}: The 12-bit signature used by SPP to index into the Pattern Table
+\item \textbf{Confidence}: The integer confidence on the scale of 0 to 100 which was used in the original SPP design
+\item \textbf{Delta}: The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched
+\item \textbf{PC}: Program Counter of the Load instruction that triggers that particular prefetch
+\item \textbf{PC XOR Delta}: This helps resolve a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases
+\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2} XOR PC\textsubscript{3}}: This tells SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block
+\item \textbf{Base address XOR Delta}: If any particular value(s) of delta are highly favoured by a given address which triggered the prefetch, this composite feature can capture that information
+\item \textbf{Cache line XOR Delta}: This feature captures if any particular delta is highly favoured by a given cache line address
+\item \textbf{Confidence XOR Delta}: This gives a combined insight into that fact that certain delta-confidence pairs predicted by SPP might correlate more with the prefetch outcome
+\item \textbf{Program Counter XOR Delta}: This tells us if a given PC favours particular value(s) of delta
+\item \textbf{Current Signature XOR Delta}: 
 \end{itemize}
-As noted above, each feature indexes into its independent entry of perceptron weights.
 
-\section{EVALUATION}
+Some of the composite features are derive from simple hashing (XOR) of two primary features.
+There is always a question of usefulness of such composite features and the `fresh' information added.
+We try to justify the choice of each feature by quantifying the contribution made towards predicting prefetch behaviour, in Section XX.
+Finally, as noted above, each feature indexes into its independent entry of perceptron weights.
+
+\section{METHODOLOGY}
 \subsection{Performance Model}
@@
 Champsim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship[cite]. 
-We model model a 1-core / 4-core / 8-core Out of Order machine with the following configuration:
+We model model a 1-core / 4-core / 8-core Out of Order machine. The details of the configuration parameters is summarized in Table  \ref{tab:Sim_params}.
 
-\begin{itemize}
-    \item Cache block size: 64 KB
-    \item 8 way separate 32 KB L1 I-Cache and D-Cache
-    \item 8 way 256 KB L2 Cache
-    \item 16 way 2 MB / core shared Last Level Cache
-    \item 4 GB Single Channel DRAM for single-core and 8 GB Double Channel DRAM for multi-core
-\end{itemize}
-Prefetching was only triggered at L2 Cache demand accesses but could be sent to L2 Cache or Last Level Cache.
+\begin{table}[]
+    \centering
+    \begin{tabular}{|l|p{3.6cm}|c|}
+    \hline
+    Level & Configuration & Access Latency \\
+    \hline
+         L1 Cache & Separate I-cache and D-cache, 32 KB, 8-way & 4 cycles\\
+         L2 Cache & Private, 256 KB & 8 cycles\\
+         LLC & Shared, 2MB / core, 16-way & 20 cycles\\
+         DRAM & 4 GB Single Channel for single-core, 8 GB Double Channel for multi-core & N?? cycles\\
+    \hline
+    \end{tabular}
+    \caption{Memory Model Parameters}
+    \label{tab:Sim_params}
+\end{table}
+
+The block size was fixed 64KB. Prefetching was only triggered at L2 Cache demand accesses but could be sent to L2 Cache or Last Level Cache. 
 No L1 Data level prefetching was done. 
 LRU replacement policy was used on all levels of cache hierarchies. 
-Branch prediction was done using a gshare branch predictor <Try perceptron predictor here>.
+Branch prediction was done using the perceptron branch predictor [cite]. 
 Page size was kept as 4KB. 
@@
 We use all the 20 workloads available in the SPEC 2017 Suite. 
-Using Simpoint[cite] methodology, we identified program segments of 1 Billion instructions each.
+Using SimPoint[cite] methodology, we identified 95 different program segments of 1 Billion instructions each.
 
-Besides SPEC 2017, we also used server level workloads from CouldSuite[cite]: <insert the workloads which should be tested>.
-<Insert mlpack?> In total, we had XX traces <95 currently> representing workloads across XX benchmarks.
-<20 currently>
+We performed cross-validation of our SPP-Perc model using SPEC 2006 and CloudSuite benchmarks.
+For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
+For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2) [cite].
+The traces include 4 multi-core applications with 6 distinct phases per application. 
+In total, we used 285 traces representing workloads across 53 applications.
 
-\textit{Single-core performance} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and the next 1 Billion instructions to do detailed simulations and collect run-time statistics.
+\textit{Single-core performance} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics.
 We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
@@
 
-\textit{Multi-core performance} For 4-core multi-programmed workloads, we generated 20 unique workloads comprising of a mix of memory intensive traces from SPEC 2017.
-For each workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics.
-As different workloads have different baseline IPCs, not all of the 4 workloads reach the 1 Billion instruction count at the same time.
-In that case, the workloads get wrapped-over and are still simulated till the last core completes 1 Billion instructions after warm-up.
-For collecting IPC and other data, only the first billion instructions are considered as "region of interest".
+\textit{Multi-core performance} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
+For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 
+Each CPU keeps executing its workload till the last CPU completes 1 Billion instructions after warm-up. 
+For collecting IPC and other data, only the first billion instructions are considered as the ``region of interest''.
+
 Here we report the weighted speedup normalized to baseline \textit{i.e.}, no prefetching. 
@@
 Then we calculate the total weighted-IPC for a given workload mix as $\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}). 
-For each of the 20 workload-mix, the sub obtained is normalized to the weighted-IPC calculated similarly for baseline case \textit{i.e.}, no prefetching, to get the weighted-IPC-speedup.
-Finally the geometric mean of these 20 weighted-IPC-speedup is reported as the effective speedup obtained by the prefetching scheme.
+For each of the 100 workload-mix, the sum obtained is normalized to the weighted-IPC calculated similarly for baseline case \textit{i.e.}, no prefetching, to get the weighted-IPC-speedup. 
+Finally the geometric mean of these 100 weighted-IPC-speedup is reported as the effective speedup obtained by the prefetching scheme.
+
+We repeat the same process for 8-core workloads with the exception that 20 Million warm-up instructions and 100 Million full instructions were executed.
+This was done so as to keep the simulation run-time within reasonable limits as single 8-core mix was taking up to 5 days of execution time.
 
@@
 
-\subsection{Overhead for SPP-Perc}
-In this section, we analyze the hardware overhead required to implement SPP-Perc.
-The exiting structires from SPP - Signature Table, Pattern Table, Global History Register and the Prefetch Table occupy 44060 bits.
-The Prefetch Filter was enhanced to accommodate storing of meta-data for perceptron training.
-New Prefetch Filter now consumes 95232 bits.
-The Reject Filter occupies 89088 bits.
-The perceptron weights occupy 348160 bits.
-All in all, SPP-Perc occupies 568348 bits, which comes out to 69.3 KB.
-
-<Inset a neat table showing analysis of area overheads of individual components>
-
-The hardware budget for 2nd Data Prefetching championship was 32 KB. Keeping that in mind and the speedup SPP-Perc obtained over the winner, the extra hardware budget can be accounted for.
-
 \subsection{Developing Features for SPP-Perc}
 This section describes the intuition and analysis that went behind finalizing the features. 
+As noted earlier, we have developed a set of 14 features which allow the perceptron throttler to correlate prefetching decision with the program context. 
+To study the correlation across each feature, we study statistically the perceptron weights and try to interpret their distribution.
+\newline  \newline
+\textbf{GLOBAL PEARSON'S CORRELATION}\newline
+For this experiment, we create a dump of the state of perceptron weights at the end of all trace execution.
+The weights obtained from running all the SPEC 2017 traces are concatenated together. 
+Since the weights are collected at the end of individual trace execution, the perceptron weights have attained a relatively stable value by now.
+Hence, this dump represents a `snap' of the trained perceptron weights across all the SPEC 2017 traces.
+The intuition here is that the feature with bulk of the perceptron weights concentrated around 0 or small magnitude numbers show a weak correlation with the prefetching outcome. 
+On the other hand, features with most of the weights saturated around highest value (+15) show a high positive correlation and the features with weights close to the lowest value (-16) show a strong negative correlation.
+
+We plot a histogram for each feature depicting weights distribution from -16 to +15 and use this histogram to generate the Pearson's correlation factor for that feature.
+In statistics, Pearson's factor is a numerical measure of the degree of linear correlation between two variables.
+It ranges from -1 to +1.
+The magnitude of Pearson's factor depicts the extent of correlation and the sign depicts whether it is a positive correlation or a negative correlation.
+Values close to 0 suggest a low correlation or at times, noisy data.
+A value of +1/-1 suggests a perfectly linear positive / negative correlation respectively.
+Figure XX depicts the histogram and the Pearson's coefficient for two of the features, YY and ZZ.
+
+Figure XX shows all the features used sorted in increasing order of their Pearson's factor.
+As can be seen X out of 14 features provide a moderate to high correlation, with the magnitude of P-value > 0.4.
+The single most important feature, \_\_ helps provide a correlation to prefetch outcome with a factor of \_\_.
+\newline \newline
+\textbf{PER TRACE CORRELATION} \newline
+Another important way to look at the perceptron features is to see how much their contribution varies across the traces.
+Here we give special attention to features with low P-values in the previous experiment.
+Figure XX shows the variation of P-values three features : \_\_, \_\_ and \_\_; across all the SPEC 2017 traces.
+For simplicity, the traces are arranged in an increasing order of contribution made by the feature and the trace names have been omitted.
+It can be seen that even features with a low overall correlation provide useful correlation (magnitude > 0.4) for XX out of the 95 simpoints developed for SPEC 2017.
+\newline \newline
+\textbf{USING PEARSON'S CORRELATION} \newline
+Besides providing interesting insights into prefetching behaviour, P-value can also be used for feature selection and prefetcher tuning.
+
+Here, we introduce the concept of cross-correlation across features.
+Just as we examined correlation of each feature with the final outcome, we can also study correlation between the features.
+We used the above methodology to eliminate features which provide a little or no information that has already been captured in other features.
+As a part of feature selection, we initially came up a mix of 23 features- primary and composite.
+By studying cross correlation of each of these features against others in a 23x23 matrix, we identified pairs of features with correlation factor > 0.95 in magnitude.
+Using this approach, we managed to reduce the feature count to 14.
+Thus, in the final implementation of SPP-Perc, no two features have a high correlation between them
+
+[TODO: VALIDATE THIS] Secondly, studying the relative importance of each feature enabled us to vary the number of entries dedicated for each feature.
+Features with higher Pearson's correlation (like \_\_ and \_\_) were given most importance and allowed full 12-bits of indexing.
+Features like \_\_ and \_\_, with a low overall P-value were allocated fewer indexes in the feature table.
+
+To conclude, in the above discussion, we justify the features for perceptron from a statistical viewpoint.
+We also show how this information can also be used for prefetcher tuning.
+All this study was made possible only because we used on-line perceptron learning for prefetching and that enabled us to examine the weights in detail.
+
+\subsection{Overhead for SPP-Perc}
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{|c|c|m{4.8cm}|}
+    \hline
+        \textbf{Field} &
+        \textbf{Bits} &
+        \textbf{Comment} \\
+    \hline
+         Valid & 1 & Indicates a valid entry in the filter\\
+         Tag & 6 & Identifier for the entry in the filter\\
+         Useful & 1 & To show if the given entry led to a useful demand fetch\\
+         Perc Decision & 1 & Prefetched vs Not-prefetched \\
+    \hline
+        PC & 12 & \multirow{5}{4.8cm}{Meta-data required for perceptron training}\\
+        Address & 24 & \\
+        Delta & 7 & \\
+        Curr Signature & 12 & \\
+        Confidence & 7 & \\
+    \hline
+        \multicolumn{3}{|c|}{\textbf{Total} 84 bits}\\
+    \hline
+    \end{tabular}
+    \caption{Meta-data Stored in PF and RF}
+    \label{tab:PF_metadata}
+\end{table}
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{|p{1.5cm}|p{1cm}|c|p{1.5cm}|}
+    \hline
+        \textbf{Structure} &
+        \textbf{Entry} &
+        \textbf{Components} &
+        \textbf{Total} \\
+    \hline
+        \multirow{5}{1.5cm}{Signature Table} & \multirow{5}{1.5cm}{256} & Valid (1 bit) & \multirow{5}{1cm}{11008  bits}\\
+        & & Tag (16 bits) &\\
+        & & Last Offset (6 bits) &\\  
+        & & Signature (12 bits) &\\
+        & & LRU (6 bits) &\\
+    \hline
+        \multirow{3}{1.5cm}{Pattern Table} & \multirow{3}{1.5cm}{512} & $C_{sig}$ (4bits) &\multirow{3}{1cm}{24576 bits}\\
+        & & $C_{delta}$ (4*4 bits) &\\
+        & & Delta (4*7 bits) &\\
+    \hline
+        \multirow{4}{1.5cm}{Perceptron Tables} & 4096*7 & \multirow{4}{1.5cm}{5 bits} & \multirow{4}{1.5cm}{186880 bits}\\
+        & 2048*4 & &\\
+        & 256*1 & &\\
+        & 128*2 & &\\
+    \hline
+        \multirow{2}{1.5cm}{Prefetch Filter \footnotemark[1]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{84 bits} & \multirow{2}{1.5cm}{86016 bits}\\
+        & & &\\
+    \hline
+     \multirow{2}{1.5cm}{Reject Filter \footnotemark[2]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{83 bits} & \multirow{2}{1.5cm}{84992 bits}\\
+        & & &\\
+    \hline
+        \multirow{4}{1.5cm}{Global History Register} & \multirow{4}{1.5cm}{8} & Signature (12 bits) & \multirow{4}{1.5cm}{264 bits}\\
+        & & Confidence (8 bits) & \\
+        & & Last Offset (6 bits) & \\
+        & & Delta (7 bits) & \\
+    \hline
+        \multirow{2}{1.5cm}{Accuracy Counters} & 1 & C$_{total}$ & 10 bits\\
+        & 1 & C$_{useful}$ & 10 bits\\
+    \hline
+        \multirow{3}{1.5cm}{Global PC Trackers} & \multirow{3}{1cm}{3} & $PC_1$ (12 bits)& \multirow{3}{1.cm}{36 bits}\\
+        & & $PC_2$ (12 bits) &\\
+        & & $PC_3$ (12 bits) &\\
+    \hline
+        \multicolumn{4}{|c|}{\textbf{Total: 393792 bits = 48.07 KB}}\\
+    \hline
+    \end{tabular}
+    \caption{SPP-Perc Storage Overhead}
+    \label{tab:SPPPerc_overhead}
+\end{table}
+\footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}
+\footnotetext[2]{RF does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
+
+In this section, we analyze the hardware overhead required to implement SPP-Perc. 
+The Prefetch Filter was enhanced to accommodate storing of meta-data for perceptron training. 
+Table \ref{tab:PF_metadata} depicts the meta-data stored for each entry in the Prefetch Filter.
+Table \ref{tab:SPPPerc_overhead} shows the total storage overhead of SPP-Perc implementation.
+The hardware budget for 2nd Data Prefetching championship was 32 KB. 
+Keeping that in mind and the speedup SPP-Perc obtained over the winner, the extra hardware budget can be accounted for.
+The extra hardware also makes the overall scheme more scalable than SPP. 
+In the original SPP paper, it was demonstrated that adding extra hardware brings little advantage in terms of performance gain.
+The newly added perceptron tables can be scaled to increase / decrease features depending on the permitted budget.
+
+In terms of computations, the perceptron mechanism only introduces an extra adder tree.
+The hash perceptron mechanism makes sure than there is no actual vector multiplication happening in the hardware.
+Obtaining the perceptron sum requires addition of 14 5-bit numbers. 
+Using an adder tree of 7 5-bit adders, this can be done in ceil($log_214$) = 4 steps.
+Perceptron update only requires weight update by +1 or -1.
+Thus, all the operations required for perceptron inferencing or updating the states of the perceptron throttler can be easily done in the time constraints of L2 Cache Accesses.
 
@@
 This section discusses the results obtained from running SPP-Perc, in terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per Cycle (IPC) metrics. 
-First we present the results for single-threaded workloads then the results for multi-core workloads are presented.
+For SPEC 2017 suite, first we present the results for single-threaded workloads then the results for multi-core workloads are presented.
 
@@
 In general, benchmarks \textit{602.gcc\_s, 603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s} benefit the most from SPP-Perc, as compared to SPP only prefetching.
-
-Figure \_\_ shows the performance of SPP-Perc against the other prefetchers in tighter constraint environments.
-For sake of simplicity, we show 2 traces where SPP-Perc gives the most advantage over baseline, 2 traces where SPP-Perc doesn't perform as well as the best prefetcher *<-replace with the prefetcher name once results are available* and finally GeoMean across all the traces.
-<Hopefully BOP would degrade more here as it is super aggressive.
-By definition, SPP-Perc should be better than SPP as the throttler would better control the aggressiveness.
-In SPP paper DA-AMPM performs close enough to BOP.
-That should get replicated here too.>
 
@@
 To test the effectiveness of SPP-Perc in a tighter constrained multi-core environment, we simulated an 8-core machine with 16 MB last Level Cache. 
-Other parameters were kept same as the ones discussed in Methodology section.
+Other parameters were kept same as discussed under Methodology in Section XX.
+
+\subsection{Additional Memory Constraints}
+Figure \_\_ shows the performance of SPP-Perc against the other prefetchers in tighter constraint environments. 
+For sake of simplicity, we show 2 traces where SPP-Perc gives the most advantage over baseline, 2 traces where SPP-Perc doesn't perform as well as the best prefetcher *<-replace with the prefetcher name once results are available* and finally GeoMean across all the traces.
+<Hopefully BOP would degrade more here as it is super aggressive. 
+By definition, SPP-Perc should be better than SPP as the throttler would better control the aggressiveness. 
+In SPP paper DA-AMPM performs close enough to BOP. That should get replicated here too.>
+
+\subsection{Cross Validation}
 

### 2018-11-14 paper_1.tex
Version 2.1
Papers added in the related work section
Features: 14 -> 9
Description of features added

@@
 \begin{abstract}
-\textbf{Hardware prefetching has been introduced in modern processors as a way to reduce compulsory cache misses. 
-An efficient prefetcher should be able to identify complex memory access patterns. 
-This enables it to fetch a block ahead of its demand access. 
+\textbf{Hardware prefetching has been introduced in modern processors as a way to hide cache latencies. 
+An efficient prefetcher should be able to identify complex memory access patterns during a program execution. 
+This enables it to fetch a block ahead of its demand access, potentially saving a cache miss.
 }
@@
 We have taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine. 
-The perceptron layer acts as a throttler filters out the unnecessary prefetches recommended by SPP. 
+The perceptron layer acts as a throttler to filter out the unnecessary prefetches recommended by SPP. 
 We have also explored a range of features that can be used to train the perceptron layer. 
@@
 Data prefetching is one such important technique. 
-Data prefetching exploits the fact that most applications, data is accessed with some spatial or temporal locality.
+Data prefetching exploits the fact that in most applications, memory accesses have a well-defined pattern.
 
-An ideal prefetching scheme involves capturing the memory access patterns to predict future prefetches in a timely manner. 
+An ideal prefetching scheme involves capturing this memory access patterns to predict future prefetches in a timely manner. 
 Other important decisions faced by the prefetcher are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block.
@@
 The prefetch suggestions recommended by SPP pass through the perceptron based throttler. 
-Over time, perceptron layer learns to correlate the prefetches recommended by SPP with various available features and eliminate useless prefetches. 
+Over time, perceptron layer learns to correlate the prefetches recommended by SPP with various available features, leading to elimination of useless prefetches. 
 This leads to an overall increased accuracy of the prefetch scheme and reduction in DRAM traffic.
@@
 \item Examining the perceptron weights doing a statistical study to quantify the contribution made by each feature.
-We have also used this idea to develop a comprehensive set of features where each feature contributes some 'new' information to the perceptron.
+We have also used this idea to develop a comprehensive set of features where each feature contributes some `new' information to the perceptron.
 The details are explained under methodology, in Section XX.
@@
 For evaluation, we used SPEC 2017 benchmarks. 
-SPEC 2006 and CloudSuite were used for cross-validation of the prefetcher implementation.
 On a single core configuration, SPP-Perc increases performance of the system by 39.5\% as compared to the baseline with no prefetching, on the memory intensive subset of the benchmark.
 This outperforms the next best prefetcher, SPP,  by 6.84\%. 
-On a multi-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement 5.95\% over SPP for 4-core system and 10.67\% for 8-core system.
+On a multi-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement of 11.9\% over SPP for 4-core system and 10.67\% for 8-core system.
 
-In the Results part in section XX, we show a detailed analysis of the speedup achieved and also talk about speedup achieved on SPEC 2006 and CloudSuite benchmarks.
+In the results part in section XX, we show a detailed analysis of the speedup achieved.
+We also used SPEC 2006 and CloudSuite benchmarks for cross-validation of the prefetcher implementation.
 
@@
 
-\subsection{Motivation}
+\subsection{Related Work}
 Some of the earliest data prefetchers developed to leverage this locality were stride based [cite:\newline
@@
 
+There have been attempts at introducing Machine Learning based elements into prefetching. 
+The paper Semantic Locality and Context-based Prefetching Using Reinforcement Learning[cite] by Peled \textit{et.al.} introduces interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
+The biggest challenge here is that it relies on compiler support for getting features to build the context.
+
+Another notable research in the domain was done by Liao \textit{et. al.} [cite Machine Learning-Based Prefetch Optimization for Data Center Applications].
+Their work focuses specifically on prefetching for data center applications. 
+They use offline Machine Learning Algorithms like Support Vector Machines, Logistic Regressions, and Tree-based Classification for parametric search for coming up with an optimal prefetcher configuration.
+
+\textit{<Note: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?>}
+Finally, the paper on Data Cache Prefetching with Perceptron Learning[cite] talks about the idea of two step prefetching.
+The first step is an existing prefetcher like Stride Prefetcher or a Markov Prefetcher while the second step is a perceptron based throttler.
+The paper has lots of limitations in terms of design and implementation.
+That reflects in the fact that it did not lead to any significant performance gain over even the basic prefetchers like the Stride Prefetcher[cite] and the Markov Prefetcher[cite]
+
 \subsection{Prefetching with SPEC 2017}
-<Discuss SPEC 2017 traces here>
-<What are the possible graphs / analysis tools we can use for SPEC traces?>
+\textit{<Discuss SPEC 2017 traces here>
+<What are the possible graphs / analysis tools we can use for SPEC traces?>}
 
@@
 The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. 
-The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the sum is assumed to indicate a  high confidence taken branch \textit{true}. 
-Else the confidence for branch taken is not high enough and the prediction is made as \textit{false}.
+The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the prediction is made as \textit{true}, or branch taken.
+Else if the sum is < 0, the prediction is made as \textit{false}.
 
@@
 This form of indexing makes sure that the hyper-plane learned by the perceptron weights is able to differentiate between linearly inseparable outcomes. 
-<Is it true? Not sure>
+\textit{<Is it true? Not sure>}
 
@@
 Care is taken that the weights saturate at certain positive and negative values so that they remain confined to the bit-width. 
-The same rules will be applied in the discussion done in section 3.2 on SPP-Perc learning algorithm.
+The same rules will be applied in the discussion done in section XX on SPP-Perc learning algorithm.
 
@@
 SPP indexes into an entry of ST using the page number. 
-For each entry corresponding to a page, ST stores the 'last block offset' and 'old signature'. 
+For each entry corresponding to a page, ST stores the `last block offset' and `old signature'. 
 Last block offset is the block offset of the last memory access of that given page. 
@@
 
-In the lookahead mode, SPP looks at the last generated delta(s), chooses the one with the highest confidence and speculatively generates the new effective address and the new confidence. 
+In the lookahead mode, SPP looks at the last generated delta(s), chooses the one with the highest confidence and speculatively generates the new effective address and the new signature. 
 Using these new parameters, it re-accesses the PT to generate further prefetches. 
 SPP keeps on repeating this cycle of accessing PT and updating the signature based on highest confidence prefetch from the last iteration. 
-These number of iterations are characterized by the 'depth' till which SPP manages to predict prefetch entries in the lookahead manner. 
+These iteration count till which SPP manages to predict prefetch entries in the lookahead manner is characterized as its `depth'. 
 While doing this, SPP also keeps compounding the confidence in each depth. 
@@
 The global accuracy acts a real-time throttler for the path confidence value as depicted in the following discussion.
+\newline
 \end{itemize}
@@
 As we increase the core count, we observe that SPP starts outperforming rest of the prefetchers. 
-This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. <Figure showing the variation of SPP vs BOP wrt core count>. 
+This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. 
+\textit{<Figure showing the variation of SPP vs BOP wrt core count>} 
 For 4-core applications, SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
@@
 To test the hypothesis, we tuned down SPP to the minimum possible threshold. 
-Only by doing this we managed to increase the SPP coverage by XX\%. 
+Only by doing this we managed to increase the prefetch suggestions made by SPP, by XX\%. 
+\textit{<Comparison of SPP-Unleased with SPP / BOP>}
 Obviously this came at a cost of increased DRAM traffic and cache pollution.
@@
 While it was not proven in the paper, the internal confidence mechanism along with the thresholds made sure that the above assumption was maintained. 
-In SPP-Perc since there can be no assumption on the degree of aggressiveness - because SPP is being tuned to its lowest thresholds, we had to explicitly make sure that at no point do the number of suggested prefetches exceed L2 MSHR queue.
+In SPP-Perc since there can be no assumption on the degree of aggressiveness - because SPP is being tuned to its lowest thresholds, we explicitly check that at no point should the number of suggested prefetches exceed L2 MSHR queue.
 
@@
 Each feature gets its own dedicated table. 
-Our proposed design uses 14 features. 
+Our proposed design uses 9 features. 
 As the program execution begins, all the weights are initialized to 0.
@@
 
-The perceptron looks at the micro-architectural state, or 'features', at that instant. 
+The perceptron looks at the micro-architectural state, or `features', at that instant. 
 Each feature indexes into the hash of up to 4096-entry table dedicated for that feature. 
@@
 
-The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch. 
+The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch cache line. 
 All prefetch candidates qualified till this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill level.
@@
 
+This mechanism allows us to solve the classical problem of exploiting `lost opportunity'. 
+In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
+Our two-step implementation allows us to overcome that issue in a simple manner.
+
 \subsection{Features used by Perceptron}
@@
 Since prefetch especially tend to be triggered from the L2 demand access, they tend to be correlated with the triggering address.
-\item \textbf{Cache line}: It is derived from the base address as base\_addr >> LOG2\_BLOCK\_SIZE
-\item \textbf{Page address}: Calculated as base\_addr >> LOG2\_PAGE\_ SIZE
-\item \textbf{Current Signature}: The 12-bit signature used by SPP to index into the Pattern Table
-\item \textbf{Confidence}: The integer confidence on the scale of 0 to 100 which was used in the original SPP design
-\item \textbf{Delta}: The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched
-\item \textbf{PC}: Program Counter of the Load instruction that triggers that particular prefetch
-\item \textbf{PC XOR Delta}: This helps resolve a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2} XOR PC\textsubscript{3}}: This tells SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block
-\item \textbf{Base address XOR Delta}: If any particular value(s) of delta are highly favoured by a given address which triggered the prefetch, this composite feature can capture that information
-\item \textbf{Cache line XOR Delta}: This feature captures if any particular delta is highly favoured by a given cache line address
-\item \textbf{Confidence XOR Delta}: This gives a combined insight into that fact that certain delta-confidence pairs predicted by SPP might correlate more with the prefetch outcome
-\item \textbf{Program Counter XOR Delta}: This tells us if a given PC favours particular value(s) of delta
-\item \textbf{Current Signature XOR Delta}: 
+
+\item \textbf{Cache line} and \textbf{Page address}: These two separate features are derived from the base address which triggered the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and base\_addr >> LOG2\_PAGE\_ SIZE respectively.
+The idea behind using three different shifted versions of the same feature is that it allows us to look into a wider range of bits than with a single version.
+It also helps give more importance to the overlapping bits and lesser importance to most and least significance bits.
+This approach can also eliminate destructive interference that can be caused by directly folding the bits into half.
+
+\item \textbf{Confidence}: The integer confidence on the scale of 0 to 100 which was used in the original SPP design.
+While the original confidence is not used directly for decision making, it can still contribute to the final decision made.
+
+\item \textbf{Program Counter XOR Depth}: PC is for the instruction which triggered the prefetch chain.
+Depth refers to the iteration count of the lookahead stages.
+In general, PC is not considered as a good basis for doing lookahead prefetching as all the prefetches with depth >= 1 gets aliased onto the same PC which won't be true in actual demand fetch case.
+This feature resolves a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases.
+This is akin to the concept of Virtual Program Counters [cite] introduced by Kim \textit{et. al.} for indirect branch prediction.
+
+\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC before the instruction that triggered the current prefetch. 
+Hashing together the last three PCs tell SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block.
+PCs are shifted in the increasing order of history before being hashed together.
+This is done to avoid the resultant value of zero in case 2 or more PCs are the same. 
+Additionally, obfuscating the information as it gets old allows us to get a wider and an approximate look into the program history.
+
+\item \textbf{Program Counter XOR Delta}: This tells us if a given PC favours particular value(s) of delta.
+As noted earlier, while PC in itself does not much useful information, this hash resolves PC into different values based on the tendency of that PC to favour a certain delta.
+This way, dynamic nature of different instances of the same memory access instruction can be captured here.
+
+\item \textbf{Current Signature XOR Delta}: It can be recalled from the discussion of SPP in section XX that the new signature is generated using the old signature and the blobk delta.
+While `Current Signature XOR Delta' is not the exact formula for generating the future signature, it gives an approximate idea of the path that the combination of these two values can lead to.
+
+%%%%%%% Below are the features rejected when downsizing 14 -> 9
+%\item \textbf{Current Signature}: The 12-bit signature used by SPP to index into the Pattern Table
+%\item \textbf{Delta}: The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched
+%\item \textbf{PC}: Program Counter of the Load instruction that triggers that particular prefetch
+%\item \textbf{Base address XOR Delta}: If any particular value(s) of delta are highly favoured by a given address which triggered the prefetch, this composite feature can capture that information
+%\item \textbf{Cache line XOR Delta}: This feature captures if any particular delta is highly favoured by a given cache line address
+%\item \textbf{Confidence XOR Delta}: This gives a combined insight into that fact that certain delta-confidence pairs predicted by SPP might correlate more with the prefetch outcome
+%%%%%%%%%%%%%%%%%%%%%%
+
 \end{itemize}
@@
 \subsection{Performance Model}
-For testing and comparing SPP-Perc, we use Champsim[cite] simulator. 
-Champsim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship[cite]. 
+For testing and comparing SPP-Perc, we used ChampSim [cite] simulator. 
+ChampSim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship (DPC-2)[cite]. 
 We model model a 1-core / 4-core / 8-core Out of Order machine. The details of the configuration parameters is summarized in Table  \ref{tab:Sim_params}.
@@
 We repeat the same process for 8-core workloads with the exception that 20 Million warm-up instructions and 100 Million full instructions were executed.
-This was done so as to keep the simulation run-time within reasonable limits as single 8-core mix was taking up to 5 days of execution time.
+This was done so as to keep the simulation run-time within reasonable limits as a single 8-core mix was taking up to 5 days to simulate 1 Billion instructions.
 
@@
 This section describes the intuition and analysis that went behind finalizing the features. 
-As noted earlier, we have developed a set of 14 features which allow the perceptron throttler to correlate prefetching decision with the program context. 
+As noted earlier, we developed a set of 9 features which allow the perceptron throttler to correlate prefetching decision with the program context. 
 To study the correlation across each feature, we study statistically the perceptron weights and try to interpret their distribution.
@@
 
-Figure XX shows all the features used sorted in increasing order of their Pearson's factor.
-As can be seen X out of 14 features provide a moderate to high correlation, with the magnitude of P-value > 0.4.
+Figure XX shows all the features used, arranged in the increasing order of their Pearson's factor.
+As can be seen X out of the 9 features provide a moderate to high correlation, with the magnitude of P-value > 0.4.
 The single most important feature, \_\_ helps provide a correlation to prefetch outcome with a factor of \_\_.
@@
 \newline \newline
-\textbf{USING PEARSON'S CORRELATION} \newline
+\textbf{TRIMMING FEATURES USING CORRELATION} \newline
 Besides providing interesting insights into prefetching behaviour, P-value can also be used for feature selection and prefetcher tuning.
@@
 We used the above methodology to eliminate features which provide a little or no information that has already been captured in other features.
+
 As a part of feature selection, we initially came up a mix of 23 features- primary and composite.
-By studying cross correlation of each of these features against others in a 23x23 matrix, we identified pairs of features with correlation factor > 0.95 in magnitude.
-Using this approach, we managed to reduce the feature count to 14.
-Thus, in the final implementation of SPP-Perc, no two features have a high correlation between them
+By studying cross correlation of each of these features against others in a 23x23 matrix, we identified pairs of features with correlation factor > 0.9 in magnitude and carefully eliminated redundant features.
+Using this approach, we managed to reduce the feature count to 9.
+Thus, in the final implementation of SPP-Perc, no two features have a high correlation between them.
+This way we can be sure that each feature is bringing in contribution that cannot be captured using other features.
 
-[TODO: VALIDATE THIS] Secondly, studying the relative importance of each feature enabled us to vary the number of entries dedicated for each feature.
+\textit{<TODO: VALIDATE THIS>} 
+Secondly, studying the relative importance of each feature enabled us to vary the number of entries dedicated for each feature.
 Features with higher Pearson's correlation (like \_\_ and \_\_) were given most importance and allowed full 12-bits of indexing.
@@
 
-To conclude, in the above discussion, we justify the features for perceptron from a statistical viewpoint.
+To conclude, in the above discussion we justify the features for perceptron from a statistical viewpoint.
 We also show how this information can also be used for prefetcher tuning.
@@
     \hline
-        \multicolumn{3}{|c|}{\textbf{Total} 84 bits}\\
+        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
     \hline
     \end{tabular}
-    \caption{Meta-data Stored in PF and RF}
+    \caption{Meta-data Stored in Prefetch Filter}
     \label{tab:PF_metadata}
@@
     \hline
-        \multirow{4}{1.5cm}{Perceptron Tables} & 4096*7 & \multirow{4}{1.5cm}{5 bits} & \multirow{4}{1.5cm}{186880 bits}\\
-        & 2048*4 & &\\
-        & 256*1 & &\\
-        & 128*2 & &\\
+        \multirow{3}{1.5cm}{Perceptron Tables} & \multirow{3}{1.5cm}{4096*8 128*1} & \multirow{3}{1.5cm}{5 bits} & \multirow{3}{1.5cm}{164480 bits}\\
+        & & &\\
+        & & &\\
     \hline
@@
     \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 393792 bits = 48.07 KB}}\\
+        \multicolumn{4}{|c|}{\textbf{Total: 371356 bits = 45.33 KB}}\\
     \hline
@@
 The hash perceptron mechanism makes sure than there is no actual vector multiplication happening in the hardware.
-Obtaining the perceptron sum requires addition of 14 5-bit numbers. 
-Using an adder tree of 7 5-bit adders, this can be done in ceil($log_214$) = 4 steps.
+Obtaining the perceptron sum requires addition of 9 5-bit numbers. 
+Using an adder tree of 4 5-bit adders, this can be done in ceil($log_{2}9$) = 4 steps.
 Perceptron update only requires weight update by +1 or -1.
@@
 Other parameters were kept same as discussed under Methodology in Section XX.
+<See if we can use speed-up over BOP as a measure here>
 
@@
 \subsection{Cross Validation}
+Here we try to demonstrate the robustness of SPP-Perc by testing it on completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
+
+Fig \_\_ shows the speed-up achieved by BOP, SPP and SPP-Perc on selected traces of SPEC 2006, along with the overall geometric mean over whole of the benchmark and over memory-intensive subset of the SPEC 2006 benchmark.
+As can be seen, SPP-Perc peovides a speed-up of 44.4\% over baseline, on the memory intensive subset of SPEC 2006 benchmark.
+This corresponds to 6.84\% over SPP and XX\% over BOP.
+On the whole of the SPEC 2006 suite, the speedup obtained is 22.4\%, which is 3.34\% ahead of SPP.
+
+The 4-core results for SPEC 2006 are \_\_.
+
+As can be seen, the performance is very consistent given no prior parameter tuning was done keeping SPEC 2006 in mind. 
+We attribute this to the inherent adaptability of the perceptron model.
+In general, perceptron weights are able to adjust in real-time so as to find the best possible correlation between the output and the features.
 

### 2018-11-15 paper_1.tex
Version 2.3
Consolidated the bibliography

@@
 While processor scaling focused on speed improvements, memory scaling was centred around increasing the storage capacity. 
-The process technology reaching scaling limitations led to Memory Wall [cite] - the increasing performance gap between the processor speed and the memory speeds. 
+The process technology reaching scaling limitations led to Memory Wall\cite{MemWall} - the increasing performance gap between the processor speed and the memory speeds. 
 Various techniques have been developed to combat this widening gap. 
@@
 \item This is the first time where a truly on-line learning based neural network model has been used when it comes to hardware data prefetching. 
-The prior works in this area either relied on program semantics [cite] or were application specific [cite] or could not yield a significant improvement over the baseline prefetcher used. [cite]
+The prior works in this area either relied on program semantics\cite{Semantics} or were application specific\cite{Datacenter} or could not yield a significant improvement over the baseline prefetcher used\cite{BadPerc}.
 
@@
 \subsection{Related Work}
-Some of the earliest data prefetchers developed to leverage this locality were stride based [cite:\newline
-- A. J. Smith. Sequential Program Prefetching in Memory Hierarchies (1978)\newline
-- J.-L. Baer and T.-F. Chen. An Effective On-Chip Preloading Scheme to Reduce Data Access Penalty (1991)\newline
-]\newline
 
-Major research in prefetching domain was based on stride length and depth prediction [cite:\newline
-- Decoupled Predictor-Directed Stream Prefetching
-Architecture (2003)\newline
-- Memory Prefetching Using Adaptive Stream Detection (2006)\newline
-]\newline
+The idea of prefetching has been around since Instruction Stream Buffers were proposed by Jouppi.
+\cite{ISB}
+Some of the earliest data prefetchers developed to try to identify memory access patterns with a constant stride pattern.
+\cite{Smith} 
+\cite{Baer}
+\cite{Stride}
+Major research was dedicated to stride distance and depth prediction.
+\cite{Decoupled}
+\cite{Adaptive}
+Newer prefetchers aimed at correlating predictions with the past memory access addresses.
+\cite{Address_Correlated} 
+\cite{AMPM}
 
-With time newer and more intricate prefetchers were introduced. [Cite:\newline
-- Temporal Streaming of Shared Memory (2005)\newline
-- Stealth Prefetching (2006)\newline
-- Feedback Directed Prefetching: Improving the Performance and Bandwidth-Efficiency of Hardware Prefetchers (2007)\newline
-- Coordinated Control of Multiple Prefetchers in Multi-Core Systems (2009)\newline
-- Techniques for Bandwidth-Efficient Prefetching of Linked Data Structures in Hybrid Prefetching Systems (2009)\newline
-- PACMan: PrefetchAware Cache Management for High Performance Caching (2011)\newline
-- TLB Improvements for Chip Multiprocessors: Inter-Core Cooperative Prefetchers and Shared Last-Level TLBs (2013)\newline
-- Linearizing Irregular Memory Accesses for Improved Correlated Prefetching (2013)\newline
-- Sandbox Prefetching: Safe run-time evaluation of aggressive prefetchers\newline
-]\newline
+With time newer and more intricate prefetchers were introduced.
+\cite{Wenisch_Temporal_Streaming}
+\cite{Stealth}
+\cite{Feedback_Directed}
+\cite{Coordinated}
+\cite{Bandwidth_Efficient}
+\cite{Pacman}
+\cite{TLB}
+\cite{Linearizing}
+\cite{Sandbox}
+\cite{VLDP}
+\cite{DoL}
+\cite{Domino}
+A whole class of streaming prefetchers developed on the lines of Spatial / Temporal Memory Streaming.
+\cite{Spatial_Pattern}
+\cite{SMS}
+\cite{Temporal_Instruction_Fetch}
+\cite{Off_Chip}
+\cite{STMS}
+\cite{SMS_JILP}
 
-Prefetchers often tend to correlate predictions with the past memory access addresses [cite: Making Address-Correlated Prefetching Practical (2010)]. 
- A whole class of streaming prefetchers developed on the lines of Spatial / Temporal Memory Streaming
-[ Cite:\newline
-- Accurate and Complexity-Effective Spatial Pattern Prediction (2004)\newline
-- Spatial Memory Streaming (2006)\newline
-- Temporal Instruction Fetch Streaming (2008)\newline
-- Practical Off-Chip Meta-Data for Temporal Memory Streaming (2009)\newline
-- Spatio-Temporal Memory Streaming (2009)\newline
-- Spatial Memory Streaming with Rotated Patterns (2009)\newline
-- Spatial Memory Streaming.” Journal of Instruction-Level Parallelism" (2011)\newline
-]\newline
+Most modern prefetchers aim at identifying complex memory access patterns in a given application. 
+This allows them to capture some of the irregular patterns seen in pointer chasing data-structures.
+Offset based prefetchers are a generalization of the next-line prefetcher.
+Some of the prominent prefetchers exploiting this idea are
+Sandbox Prefetcher \cite{Sandbox}
+and the Best Offset Prefetcher \cite{BOP}
+Another class of prefetchers is lookahead Prefetchers.
+Such prefetchers try to speculate deep into the application's memory access trail.
+These include Path Confidence based Lookahead Prefetching \cite{SPP}
+and Kill the Program Counter \cite{KPC}.
 
-More modern prefetchers are often offset based <cite BOP and other offset based> or lookahead based <cite SPP, KPC and a few more>.
+A related line of work also explores cache replacement - insertion policies and dead block prediction, with an eye towards prefetching. 
+\cite{DB_Pred}
+\cite{Cache_Burst}
+\cite{KPC}
+\cite{Harmony}
+Nori \textit{et.al.} introduced prefetching with  respect to runtime criticality.
+\cite{CATCH}
+\newline
 
-A related line of work also explored cache replacement: insertion policies and dead block prediction, with an eye towards prefetching. [Cite:\newline
-- Dead-Block Prediction and Dead-Block Correlating Prefetchers (2001)\newline
-- Cache Bursts: A New Approach for Eliminating Dead Blocks and Increasing Cache Efficiency (2008)\newline
-- KPC (2017)\newline
-- Rethinking Belady's Algorithm to Accommodate Prefetching (2018)\newline
-]
-
-There have been attempts at introducing Machine Learning based elements into prefetching. 
-The paper Semantic Locality and Context-based Prefetching Using Reinforcement Learning[cite] by Peled \textit{et.al.} introduces interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
+\textit{Machine Learning and Prefetching:}
+Peled \textit{et.al.} introduce interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
+\cite{Semantics}
 The biggest challenge here is that it relies on compiler support for getting features to build the context.
 
-Another notable research in the domain was done by Liao \textit{et. al.} [cite Machine Learning-Based Prefetch Optimization for Data Center Applications].
-Their work focuses specifically on prefetching for data center applications. 
-They use offline Machine Learning Algorithms like Support Vector Machines, Logistic Regressions, and Tree-based Classification for parametric search for coming up with an optimal prefetcher configuration.
+The paper by Liao \textit{et. al.} focuses on prefetching for data center applications.
+\cite{Datacenter}
+They use offline Machine Learning Algorithms like SVMs and Logistic Regressions to do parametric search for an optimal prefetcher configuration.
 
-\textit{<Note: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?>}
-Finally, the paper on Data Cache Prefetching with Perceptron Learning[cite] talks about the idea of two step prefetching.
-The first step is an existing prefetcher like Stride Prefetcher or a Markov Prefetcher while the second step is a perceptron based throttler.
+%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?
+Finally, the paper on Data Cache Prefetching with Perceptron Learning\cite{BadPerc} talks about the idea of two step prefetching.
+The first step is an existing baseline prefetcher while the second step is a perceptron based throttler.
 The paper has lots of limitations in terms of design and implementation.
-That reflects in the fact that it did not lead to any significant performance gain over even the basic prefetchers like the Stride Prefetcher[cite] and the Markov Prefetcher[cite]
+That reflects in the fact that it did not lead to any significant performance gain over even the basic prefetchers like the Stride Prefetcher\cite{Stride} and the Markov Prefetcher\cite{Markov}.
+
+et. al.
 
@@
 Perceptron learning for computer architecture design has been around for a while. 
-It was popularized by [cite initial branch prediction paper]. 
+It was first popularized by Jimenez \textit{et. al.} in for doing branch prediction. \cite{Perc_Branch} 
 In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
@@
 
-The structure that we use is derived from the model introduced in [cite piecewise linear branch prediction]. 
+The structure that we use is derived from the model introduced in Piecewise Linear Branch Prediction\cite{Piece_Linear}. 
 Here the feature itself is used to index into a hash of perceptron weights. 
@@
 
-Perceptron Learning for Reuse Prediction [cite] extended the hash perceptron architecture to perform prediction in context of cache replacement. 
+Perceptron Learning for Reuse Prediction\cite{Perc_Reuse} extended the hash perceptron architecture to perform prediction in context of cache replacement. 
 Perceptron layer can learn to correlate the cache replacement behaviour with a variety of features. 
@@
 
-The work was further extended by Multiperspective Reuse Prediction[cite], which introduced a varied set of parametric features. 
+The work was further extended by Multiperspective Reuse Prediction\cite{Multiperspective}, which introduced a varied set of parametric features. 
 Although design space exploration for feature and parameter selection was a non-trivial task, once fixed, it could give a cache replacement predictor highly suitable for the given set of applications.
@@
 \subsection{Signature Path Prefetcher Architecture}
-Since the work we have done builds heavily on SPP[cite], we are presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work. 
+Since the work we have done builds heavily on SPP\cite{SPP}, we are presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work. 
 Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists of following structural components:
@@
 This feature resolves a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases.
-This is akin to the concept of Virtual Program Counters [cite] introduced by Kim \textit{et. al.} for indirect branch prediction.
+This is akin to the concept of Virtual Program Counters\cite{VPC} introduced by Kim \textit{et. al.} for indirect branch prediction.
 
@@
 \subsection{Performance Model}
-For testing and comparing SPP-Perc, we used ChampSim [cite] simulator. 
-ChampSim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship (DPC-2)[cite]. 
+For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim} simulator. 
+ChampSim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} 
 We model model a 1-core / 4-core / 8-core Out of Order machine. The details of the configuration parameters is summarized in Table  \ref{tab:Sim_params}.
@@
 LRU replacement policy was used on all levels of cache hierarchies. 
-Branch prediction was done using the perceptron branch predictor [cite]. 
+Branch prediction was done using the perceptron branch predictor\cite{Perc_Branch}. 
 Page size was kept as 4KB. 
@@
 \subsection{Workloads}
-This is the first time that SPEC 2017 benchmark suite[cite] has been used to characterize and measure the prefetch performance. 
+This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017} has been used to characterize and measure the prefetch performance. 
 We use all the 20 workloads available in the SPEC 2017 Suite. 
-Using SimPoint[cite] methodology, we identified 95 different program segments of 1 Billion instructions each.
+Using SimPoint\cite{SimPoint} methodology, we identified 95 different program segments of 1 Billion instructions each.
 
-We performed cross-validation of our SPP-Perc model using SPEC 2006 and CloudSuite benchmarks.
+We performed cross-validation of our SPP-Perc model using SPEC 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.
 For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
-For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2) [cite].
+For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.
 The traces include 4 multi-core applications with 6 distinct phases per application. 
@@
 \subsection{Preferchers Simulated}
-We compared SPP-Perc against three of the latest state of the art prefetchers: Best Offset Prefetcher (BOP)[cite], DRAM Aware - Access Map Pattern Matching (DA-AMPM) [cite] and Signature Path Prefecher (SPP)[cite].
-BOP was the winner in 2nd Data Prefetching Championship[cite]. 
-DA-AMPM is the enhanced version of AMPM[cite], modified to account for DRAM row buffer locality. 
-SPP has been shown to outperform BOP on SPEC 2006 traces[cite]. 
+We compared SPP-Perc against three of the latest state of the art prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher (SPP).
+BOP was the winner of 2nd Data Prefetching Championship. 
+DA-AMPM is the enhanced version of AMPM, modified to account for DRAM row buffer locality. 
+SPP has been shown to outperform BOP on SPEC 2006 traces. 
 For each of these, we compare their speedups taking no prefetching as the baseline.
@@
 %%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
-%%\bibliographystyle{ieeetr}
-%%\bibliography{ref}
+\bibliographystyle{ieeetr}
+\bibliography{ref}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-11-15 paper_1.tex
Version 2.3
Typos / misc.

@@
 
-\vspace{1ex}The rest of the paper is organized as follows:
-
-\begin{itemize}
-\item Section 2 talks about background work and recaptures the SPP architecture
-\item Section 3 discusses enhancements made to the SPP design and the detailed implementation of the perceptron mechanism
-\item Section 4 consists of the methodology for evaluating SPP-Perc and exploring the feature space for perceptron learning
-\item Section 5 comprises results and evaluation in terms of IPC, MPKI, coverage and DRAM traffic, for a variety of benchmark suites
-\item Section 6 is the conclusion and future scope of expansion on this work
-\end{itemize}
+The rest of the paper is organized as follows.
+Section 2 talks about background work and recaptures the SPP architecture
+Section 3 discusses the detailed implementation of the perceptron mechanism and the features used.
+Section 4 consists of the evaluation methodology and explores the feature space for perceptron learning.
+Section 5 is the performance evaluation of our prefetcher and Section 6 concludes the paper.
 
@@
 \newline
-
+\newline
 \textit{Machine Learning and Prefetching:}
@@
 
-et. al.
-
-\subsection{Prefetching with SPEC 2017}
-\textit{<Discuss SPEC 2017 traces here>
-<What are the possible graphs / analysis tools we can use for SPEC traces?>}
-
 \subsection{Perceptron Learning in Architecture}
@@
 Perceptron learning for computer architecture design has been around for a while. 
-It was first popularized by Jimenez \textit{et. al.} in for doing branch prediction. \cite{Perc_Branch} 
+It was first popularized by Jimenez \textit{et. al.} for doing branch prediction. \cite{Perc_Branch} 
 In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
@@
 The same rules will be applied in the discussion done in section XX on SPP-Perc learning algorithm.
+
+\subsection{Prefetching with SPEC 2017}
+\textit{<Discuss SPEC 2017 traces here>
+<What are the possible graphs / analysis tools we can use for SPEC traces?>}
 
@@
 In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
-Our two-step implementation allows us to overcome that issue in a simple manner.
+Our two-step SPP-Perc allows us to overcome that issue in a simple manner.
 
@@
 \item \textbf{Base address} of the demand access which triggered the prefetch. 
-Since prefetch especially tend to be triggered from the L2 demand access, they tend to be correlated with the triggering address.
+Since prefetch are triggered from the L2 demand access, they tend to be correlated with the triggering address.
 
@@
 In total, we used 285 traces representing workloads across 53 applications. 
-
-\textit{Single-core performance} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics.
+\newline
+\newline
+\textit{Single-core performance:} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics.
 We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
@@
 The final number reported is the geometric mean of the speed-up attained on individual traces.
-
-\textit{Multi-core performance} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
+\newline
+\newline
+\textit{Multi-core performance:} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
 For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 

### 2018-11-15 abs.tex
reorg

--- 2018-11-15 paper_1.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is the template for submission to ISCA 2018
-% The cls file is a modified from  'sig-alternate.cls'
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{sig-alternate} 
-\usepackage{mathptmx} % This is Times font
-
-\newcommand{\ignore}[1]{}
-\usepackage{fancyhdr}
-\usepackage[normalem]{ulem}
-\usepackage[hyphens]{url}
-\usepackage{microtype}
-\usepackage{fixltx2e}
-\usepackage{array}
-\usepackage{multirow}
-
-% Always include hyperref last
-\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=black,citecolor=blue,urlcolor=black]{hyperref}
-
-% Ensure letter paper
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\newcommand{\iscasubmissionnumber}{NaN}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\fancypagestyle{firstpage}{
-  \fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-  \fancyhead[C]{\normalsize{}} 
-  \fancyfoot[C]{\thepage}
+\begin{abstract}
+  \textbf{Hardware prefetching has been introduced in modern
+    processors as a way to hide cache latencies.  An efficient
+    prefetcher should be able to identify complex memory access
+    patterns during a program execution.  This enables it to fetch a
+    block ahead of its demand access, potentially saving a cache miss.
   }
 
-\pagenumbering{arabic}
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\title{SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher} 
-\author{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\maketitle
-\thispagestyle{firstpage}
-\pagestyle{plain}
-
-\begin{abstract}
-\textbf{Hardware prefetching has been introduced in modern processors as a way to hide cache latencies. 
-An efficient prefetcher should be able to identify complex memory access patterns during a program execution. 
-This enables it to fetch a block ahead of its demand access, potentially saving a cache miss.
-}
-
-\textbf{In this paper, we introduce perceptron learning to help make this prefetching decision efficiently. 
-We have taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine. 
-The perceptron layer acts as a throttler to filter out the unnecessary prefetches recommended by SPP. 
-We have also explored a range of features that can be used to train the perceptron layer. 
-Our results show that SPP-Perc improves performance on the memory intensive subset of the SPEC 2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared to just SPP.
-We also demonstrate that the performance gained from using our efficient throttler scales better with increasing number of cores.
-}
+  \textbf{In this paper, we introduce perceptron learning to help make
+    this prefetching decision efficiently.  We have taken a state of
+    the art prefetcher - Signature Path Prefetcher (SPP) as the
+    underlying prefetch engine.  The perceptron layer acts as a
+    throttler to filter out the unnecessary prefetches recommended by
+    SPP.  We have also explored a range of features that can be used
+    to train the perceptron layer.  Our results show that SPP-Perc
+    improves performance on the memory intensive subset of the SPEC
+    2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on
+    multi-core traces, as compared to just SPP.  We also demonstrate
+    that the performance gained from using our efficient throttler
+    scales better with increasing number of cores.  }
 \end{abstract}
-
-\section{Introduction}
-Since past few decades, processor and memory scaling has been enabled with different philosophies in mind. 
-While processor scaling focused on speed improvements, memory scaling was centred around increasing the storage capacity. 
-The process technology reaching scaling limitations led to Memory Wall\cite{MemWall} - the increasing performance gap between the processor speed and the memory speeds. 
-Various techniques have been developed to combat this widening gap. 
-Data prefetching is one such important technique. 
-Data prefetching exploits the fact that in most applications, memory accesses have a well-defined pattern.
-
-An ideal prefetching scheme involves capturing this memory access patterns to predict future prefetches in a timely manner. 
-Other important decisions faced by the prefetcher are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block.
-Memory accesses patterns can be as simple as fetching the next cache line or a complex but predictable pattern, like in pointer chasing applications. 
-Predicting this often requires a trade-off between coverage and accuracy. 
-For a prefetcher to capture highly irregular access patterns, it needs to be highly aggressive (high coverage). 
-That also means the prefetcher would end up recommending a lot of useless prefetches, leading to cache pollution (low accuracy). 
-This effect would get even more pronounced in multi-core scenario where the last level cache is a shared resource. 
-Conversely, a conservative prefetcher would be highly accurate but might not make enough useful predictions.
-
-To this effect we propose SPP-Perc <come up with a catchy name>. 
-It is an enhancement to the existing state of the art prefetcher and overcomes the coverage vs accuracy trade-off in a robust manner. 
-The idea is to modify SPP design to make it as aggressive as it can get. 
-This enables SPP to capture complex memory access patterns hence leading to increased coverage.
-This comes at a cost of increased DRAM traffic, hence reduction in overall IPC.
-The prefetch suggestions recommended by SPP pass through the perceptron based throttler. 
-Over time, perceptron layer learns to correlate the prefetches recommended by SPP with various available features, leading to elimination of useless prefetches. 
-This leads to an overall increased accuracy of the prefetch scheme and reduction in DRAM traffic.
-
-\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and future scope for expansion. 
-The contributions made by this paper are:
-
-\begin{itemize}
-\item This is the first time where a truly on-line learning based neural network model has been used when it comes to hardware data prefetching. 
-The prior works in this area either relied on program semantics\cite{Semantics} or were application specific\cite{Datacenter} or could not yield a significant improvement over the baseline prefetcher used\cite{BadPerc}.
-
-\item Implementing the pereptron throttler in a robust and accurate practical prefetching mechanism. 
-The overall scheme yields increased accuracy as compared to the other state of the art prefetchers.
-Moreover, the perceptron throttler learns to adapt itself to shared resource constraints - leading to an even increased performance in multi-core and lower capacity / bandwidth environments.
-
-\item Examining the perceptron weights doing a statistical study to quantify the contribution made by each feature.
-We have also used this idea to develop a comprehensive set of features where each feature contributes some `new' information to the perceptron.
-The details are explained under methodology, in Section XX.
-
-\item Cross-validating the design across benchmarks. 
-All the development and parameter tuning was done on the SPEC 2017 benchmark. 
-The results have been shown to be consistent across unseen benchmarks.
-More details can be found in the results section (Section XX).
-
-%\item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind. 
-%We have used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers. 
-%We have also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
-
-%\item We came up with an efficient model to use the perceptron output as a representation of the prefetch confidence using an on-line histogram technique. 
-%We show how the simple sum can replace the existing mechanism which was essentially based on counting the number of hits for a particular entry in SPP.
-
-\end{itemize}
-
-We tested SPP-Perc on Champsim, an open source simulator which models an Out-of-Order CPU using. 
-For evaluation, we used SPEC 2017 benchmarks. 
-On a single core configuration, SPP-Perc increases performance of the system by 39.5\% as compared to the baseline with no prefetching, on the memory intensive subset of the benchmark.
-This outperforms the next best prefetcher, SPP,  by 6.84\%. 
-On a multi-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement of 11.9\% over SPP for 4-core system and 10.67\% for 8-core system.
-
-In the results part in section XX, we show a detailed analysis of the speedup achieved.
-We also used SPEC 2006 and CloudSuite benchmarks for cross-validation of the prefetcher implementation.
-
-The rest of the paper is organized as follows.
-Section 2 talks about background work and recaptures the SPP architecture
-Section 3 discusses the detailed implementation of the perceptron mechanism and the features used.
-Section 4 consists of the evaluation methodology and explores the feature space for perceptron learning.
-Section 5 is the performance evaluation of our prefetcher and Section 6 concludes the paper.
-
-\section{Background Work and Motivation}
-
-In this section, we compare the existing work done in the domain of prefetching. 
-We also discuss about some of the interesting memory access patterns seen in the workloads in SPEC 2017 benchmark. 
-The second half of this section deals with a background of Signature Path Prefetcher architecture.
-
-\subsection{Related Work}
-
-The idea of prefetching has been around since Instruction Stream Buffers were proposed by Jouppi.
-\cite{ISB}
-Some of the earliest data prefetchers developed to try to identify memory access patterns with a constant stride pattern.
-\cite{Smith} 
-\cite{Baer}
-\cite{Stride}
-Major research was dedicated to stride distance and depth prediction.
-\cite{Decoupled}
-\cite{Adaptive}
-Newer prefetchers aimed at correlating predictions with the past memory access addresses.
-\cite{Address_Correlated} 
-\cite{AMPM}
-
-With time newer and more intricate prefetchers were introduced.
-\cite{Wenisch_Temporal_Streaming}
-\cite{Stealth}
-\cite{Feedback_Directed}
-\cite{Coordinated}
-\cite{Bandwidth_Efficient}
-\cite{Pacman}
-\cite{TLB}
-\cite{Linearizing}
-\cite{Sandbox}
-\cite{VLDP}
-\cite{DoL}
-\cite{Domino}
-A whole class of streaming prefetchers developed on the lines of Spatial / Temporal Memory Streaming.
-\cite{Spatial_Pattern}
-\cite{SMS}
-\cite{Temporal_Instruction_Fetch}
-\cite{Off_Chip}
-\cite{STMS}
-\cite{SMS_JILP}
-
-Most modern prefetchers aim at identifying complex memory access patterns in a given application. 
-This allows them to capture some of the irregular patterns seen in pointer chasing data-structures.
-Offset based prefetchers are a generalization of the next-line prefetcher.
-Some of the prominent prefetchers exploiting this idea are
-Sandbox Prefetcher \cite{Sandbox}
-and the Best Offset Prefetcher \cite{BOP}
-Another class of prefetchers is lookahead Prefetchers.
-Such prefetchers try to speculate deep into the application's memory access trail.
-These include Path Confidence based Lookahead Prefetching \cite{SPP}
-and Kill the Program Counter \cite{KPC}.
-
-A related line of work also explores cache replacement - insertion policies and dead block prediction, with an eye towards prefetching. 
-\cite{DB_Pred}
-\cite{Cache_Burst}
-\cite{KPC}
-\cite{Harmony}
-Nori \textit{et.al.} introduced prefetching with  respect to runtime criticality.
-\cite{CATCH}
-\newline
-\newline
-\textit{Machine Learning and Prefetching:}
-Peled \textit{et.al.} introduce interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
-\cite{Semantics}
-The biggest challenge here is that it relies on compiler support for getting features to build the context.
-
-The paper by Liao \textit{et. al.} focuses on prefetching for data center applications.
-\cite{Datacenter}
-They use offline Machine Learning Algorithms like SVMs and Logistic Regressions to do parametric search for an optimal prefetcher configuration.
-
-%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?
-Finally, the paper on Data Cache Prefetching with Perceptron Learning\cite{BadPerc} talks about the idea of two step prefetching.
-The first step is an existing baseline prefetcher while the second step is a perceptron based throttler.
-The paper has lots of limitations in terms of design and implementation.
-That reflects in the fact that it did not lead to any significant performance gain over even the basic prefetchers like the Stride Prefetcher\cite{Stride} and the Markov Prefetcher\cite{Markov}.
-
-\subsection{Perceptron Learning in Architecture}
-
-Perceptron learning for computer architecture design has been around for a while. 
-It was first popularized by Jimenez \textit{et. al.} for doing branch prediction. \cite{Perc_Branch} 
-In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
-The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. 
-The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the prediction is made as \textit{true}, or branch taken.
-Else if the sum is < 0, the prediction is made as \textit{false}.
-
-The structure that we use is derived from the model introduced in Piecewise Linear Branch Prediction\cite{Piece_Linear}. 
-Here the feature itself is used to index into a hash of perceptron weights. 
-This form of indexing makes sure that the hyper-plane learned by the perceptron weights is able to differentiate between linearly inseparable outcomes. 
-\textit{<Is it true? Not sure>}
-
-Perceptron Learning for Reuse Prediction\cite{Perc_Reuse} extended the hash perceptron architecture to perform prediction in context of cache replacement. 
-Perceptron layer can learn to correlate the cache replacement behaviour with a variety of features. 
-Each feature has its dedicated table of perceptron weights which it indexes into. 
-Hence number of tables equals the number of features. 
-Once the perceptron weights from all the tables are obtained, they're summed and thresholded based on a pre-decided value. 
-This led to a prediction mechanism that is highly accurate, has quick learning rate and is adaptable to changing program behaviour. 
-The features introduced in that paper were derived from Program Counters of the current and the last few instructions and the memory address of the current access. 
-Even with such basic features, high level of correlation could be established by perceptrons.
-
-The work was further extended by Multiperspective Reuse Prediction\cite{Multiperspective}, which introduced a varied set of parametric features. 
-Although design space exploration for feature and parameter selection was a non-trivial task, once fixed, it could give a cache replacement predictor highly suitable for the given set of applications.
-\newline
-\newline
-\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron prediction involves multiplying the vector of input features: F\textsubscript{1xN}, with the corresponding weight vector: W\textsubscript{Nx1} in a dot product fashion to obtain the sum y\textsubscript{out}.
-Here what we use is a perceptron-like structure. 
-The feature is used to hash into the weight of perceptrons and the retrieved weights are added straight away. 
-This can be seen a dot product of $\vec{1}$\textsubscript{1xN} with W\textsubscript{Nx1}. 
-This way, the perceptron algorithm doesn't introduce any multiplication operations in the inference or training process. 
-Hence what we adapt in this work, is a perceptron-like learning algorithm as it involves the same principle involved in perceptron inferencing and training. 
-For simplicity, we will call our implementation as perceptron implementation.
-\newline
-\newline
-\textit{Perceptron Update Rule:} In all the above implementations, a uniform perceptron update principle is being followed. 
-The weights need to be updated if the prediction was wrong or the predicted sum does not cross a certain threshold. 
-After a certain training period, these weights are proportional to the probability of the outcome being true. 
-This training threshold makes sure that the perceptrons are trained till a certain level of confidence and yet they are not over-trained to the given set. 
-Care is taken that the weights saturate at certain positive and negative values so that they remain confined to the bit-width. 
-The same rules will be applied in the discussion done in section XX on SPP-Perc learning algorithm.
-
-\subsection{Prefetching with SPEC 2017}
-\textit{<Discuss SPEC 2017 traces here>
-<What are the possible graphs / analysis tools we can use for SPEC traces?>}
-
-\subsection{Signature Path Prefetcher Architecture}
-Since the work we have done builds heavily on SPP\cite{SPP}, we are presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work. 
-Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists of following structural components:
-\newline
-\newline
-\textbf{Signature Table:} ST keeps track of 256 most recently accessed pages. 
-It is meant to capture memory access patterns within a page boundary. 
-SPP indexes into an entry of ST using the page number. 
-For each entry corresponding to a page, ST stores the `last block offset' and `old signature'. 
-Last block offset is the block offset of the last memory access of that given page. 
-The block offset is calculated with respect to the page boundary. 
-The signature is a 12-bit compressed representation of the past few memory accesses for that page. 
-The signature is calculated as:
-$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
-In case a matching page entry is found, the stored signature in the corresponding entry is retrieved.
-\newline
-\newline
-Consider the case that incoming page number 10 with a block offset 3 finds a match in the ST. 
-The retrieved pattern signature is 0x30 and the Last Offset is 1. 
-Since now there is the Last Block Offset (1), Incoming Block Offset (3), Old Signature (0x30), and New Signature calculated as per above equation (0x182), SPP can infer non-speculatively that the given pattern of memory accesses (as captured in the Old Signature) leads to the particular Delta.
-In general, Delta is defined as the difference between the prefetch suggestion block and the initial block which triggered the prefetch. 
-In this case, since SPP is in learning phase, it is defined as the difference between the Incoming Block Offset and the Last Block Offset (+2 in this case). 
-This, in turn generates the new memory pattern (New Signature). 
-This newly learned signature pattern and the delta is stored in the Pattern Table.
-\newline
-\newline
-\textbf{Pattern Table:} The PT is indexed by the signature generated from ST. 
-PT holds predicted delta patterns and their confidence estimates. 
-Each entry indexed by the signature holds up to 4 unique delta predictions. 
-This is implemented by making PT as a 4-way associative table. 
-To measure an estimate of a delta confidence, PT implements two counters: C\textsubscript{sig} for each entry indexed by the signature and C\textsubscript{delta} for each delta entry for a given signature.
-More details of this on-line score-keeping mechanism is described under Confidence Tracking discussion.
-
-\textit{Lookahead Prefetching:} For each trigger to the prefetcher, the first time that the SPP indexes into the PT, it does so in the non-speculative manner. 
-We call it non-speculative because the parameters which SPP uses to predict, like last block offset, incoming block offset and signature are well-defined. 
-Assuming PT generates a valid delta prediction (rather, up to 4 valid deltas), SPP enters into the lookahead mode.
-
-In the lookahead mode, SPP looks at the last generated delta(s), chooses the one with the highest confidence and speculatively generates the new effective address and the new signature. 
-Using these new parameters, it re-accesses the PT to generate further prefetches. 
-SPP keeps on repeating this cycle of accessing PT and updating the signature based on highest confidence prefetch from the last iteration. 
-These iteration count till which SPP manages to predict prefetch entries in the lookahead manner is characterized as its `depth'. 
-While doing this, SPP also keeps compounding the confidence in each depth. 
-Thus as depth increases, overall confidence keeps decreasing. 
-This makes sense since we are going deeper into the speculation path. 
-When the confidence falls below a pre-defined threshold, SPP stops further lookahead iterations
-\newline
-\newline
-\textbf{Global History Register:} One of the innovative ideas introduced in SPP was the Global History Register, which bootstraps learning across page boundaries.
-GHR is a small 8-entry fully associative structure. 
-Whenever a prefetch suggestion from the Pattern Table crosses page boundaries, instead of completely discarding it, GHR saves that information for inter-page learning.
-Since the SPP-Perc implementation in this paper doesn't affect GHR functionality in any ways, we won't be going into more details about GHR.
-\newline
-\newline
-\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch Filter - a 1024 entry 1-way associative table which keeps a record of last few entries prefetched by SPP. 
-This proves useful for following reasons:
-\begin{itemize}
-    \item \textbf{Eliminating duplicate prefetches}\newline 
-    Any subsequent prefetches recommended by SPP mechanism is checked against the existing entries in Prefetch Filter. 
-If a valid entry is found, it means that the particular cache line is already either in L2 Cache or being fetched. 
-Hence there is no need to issue a duplicate prefetch.
-
-    \item \textbf{Updating prefetcher}\newline
-    If a subsequent demand access comes before eviction for a prefetch that was recommended by SPP, various components of SPP are updated to reflect that this was a good prefetch. 
-Correspondingly, if a prefetched cache line is evicted without being demand hit even once, SPP learns that this was a useless prefetch and updates the various tables accordingly
-
-    \item \textbf{Tracking Global Accuracy}\newline
-    PF keeps a track of usefulness of the prefetches by introducing 2 counters: C\textsubscript{total} to track the total number of prefetches getting recorded in the PF and C\textsubscript{useful} to keep a track of which prefetches recorded in the PF led to a demand hit.
-The ratio C\textsubscript{useful} / C\textsubscript{total} is the global accuracy for the prefetching scheme. 
-It is represented as $\alpha$ in the paper. 
-The global accuracy acts a real-time throttler for the path confidence value as depicted in the following discussion.
-\newline
-\end{itemize}
-\textbf{Confidence Tracking}: The PT keeps a track of hits to each signature through a counter C\textsubscript{sig}. 
-The number of hits for a given delta per signature are tracked using a counter C\textsubscript{delta}. 
-The confidence for a given delta is approximated through C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}. 
-When SPP enters into a lookahead mode, the path confidence P\textsubscript{d} is given as:
-$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown in the Prefetch Filter section. 
-The range of global accuracy is in [0,1]. 
-The lower the value of $\alpha$ for a given trace or a program phase, the overall confidence is scaled down and this makes SPP less aggressive for the upcoming prefetches.
-
-Here d is the lookahead depth. 
-For d = 1, when SPP is in non-speculative mode, P\textsubscript{0} can be thought of as 1. 
-Once SPP enters into speculative mode, it calculates the path confidence as a product of confidence of the individual prefetches encountered along that path, further scaled using $\alpha$. 
-The final P\textsubscript{d} is thresholded against prefetching-threshold (T\textsubscript{p}) to decide whether to prefetch or not. 
-If yes, then P\textsubscript{d} is thresholded against a numerically bigger fill-threshold (T\textsubscript{f}) to decide whether the prefetch should be sent to L2 Cache (high confidence prefetch) or Last Level Cache (low confidence prefetch)
-\newline<Insert a comprehensive figure about SPP structures>
-\newline<Insert another detailed picture about SPP data-path flow>
-
-\subsection{Case for an On-line Throttler}
-As compared to some of the other state of the art prefetchers [BOP], SPP is a far less aggressive prefetcher. 
-In the single core environment, this gives BOP an edge as there is no resource contention among the different cores. 
-Hence more aggressive prefetching is bound to prove beneficial. 
-As we increase the core count, we observe that SPP starts outperforming rest of the prefetchers. 
-This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. 
-\textit{<Figure showing the variation of SPP vs BOP wrt core count>} 
-For 4-core applications, SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
-
-The above analysis shows that with a more careful throttling mechanism, SPP can be tuned to become much more aggressive, leading to increased coverage. 
-The onus of maintaining the accuracy now falls on the independent throttler. 
-To test the hypothesis, we tuned down SPP to the minimum possible threshold. 
-Only by doing this we managed to increase the prefetch suggestions made by SPP, by XX\%. 
-\textit{<Comparison of SPP-Unleased with SPP / BOP>}
-Obviously this came at a cost of increased DRAM traffic and cache pollution.
-
-Moreover, the on-line confidence mechanism introduced in SPP was very rudimentary.
-It was based on taking a ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig} as explained above.
-The same confidence was used to make the decision whether to prefetch or not to prefetch; and which level to prefetch.
-While this approximation was shown to work in the original SPP, we believe that a better form of on-line decision making was possible.
-Hence, it was necessary to build a robust and adaptable learning mechanism to accept / reject the prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level Cache).
-
-To that effect, we introduce an independent on-line perceptron based throttling mechanism.
-
-\section{SPP-Perc}
-
-\subsection{Changes made to original SPP}
-To modify the SPP Design to suit our scheme, the following changes were made:
-
-\begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline
-In the original SPP design, an internal confidence mechanism was developed which was compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the prefetch suggestion.
-Furthermore, the accepted suggestions were thresholded against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache vs Last Level Cache.
-T\textsubscript{p} and T\textsubscript{f} were set to 25 and 90 respectively, on the scale of 0 to 100.
-In SPP-Perc, the perceptron sum is used to make these decisions. 
-Hence these two thresholds are no longer needed
-
-\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
-Original SPP would not consider the slots available in the L2 MSHR queue. 
-It had an assumption that at no time can the number of prefetches suggested by SPP could exceed the capacity of the queue. 
-While it was not proven in the paper, the internal confidence mechanism along with the thresholds made sure that the above assumption was maintained. 
-In SPP-Perc since there can be no assumption on the degree of aggressiveness - because SPP is being tuned to its lowest thresholds, we explicitly check that at no point should the number of suggested prefetches exceed L2 MSHR queue.
-
-\item \textbf{Enhanced Filter}\newline
-Original SPP introduces a concept of Prefetch Filter, which keeps a track of up to 1024 L2 Cache prefetches suggested by SPP. 
-The Filter was modified to store meta-data like program counter, memory access address etc. 
-This is needed to train the perceptron when the result of that particular prefetch gets available. 
-At training time, this data is used to retrieve the state of the prefetcher when the prediction was made.
-
-\item \textbf{Reject Filter introduced}\newline
-The Prefetch Filter mentioned above keeps a track of the prefetches that actually happened in the L2 Cache. 
-Correspondingly, we introduced a reject filter. 
-It keeps a track of the prefetches that were suggested by the base SPP engine but were rejected by the perceptron. 
-In case it is detected that a future demand fetch could have used this rejected prefetch, the perceptron weights are updated to reflect the same
-
-\end{itemize}
-
-\subsection{The Perceptron Layer}
-As discussed earlier, the perceptron layer was introduced to act as a throttler to the underlying SPP engine. 
-This section explores the details of the perceptron layer and the range of features that are ultimately used by the perceptron for correlating the prefetches.
-
-The neural network weights are stored in separate tables for each feature.
-The indexing of each feature is done using different number of bits.
-At most 12 bits of a feature are used to index into the perceptron table.
-Certain features require more resolution power \textit{i.e.}, full 12 bits of indexing.
-On the other hand, some features requiring lower resolution required as low as 7 bits of indexing.
-The variable indexing was determined by studying the features and fine-tuned empirically so as to achieve a good accuracy vs area trade-off.
-Exact details are mentioned under `Area Overheads' in Section XX.
-
-A single entry in the table corresponds to a perceptron weight.
-Each weight is a 5 bit counter - saturating at -16 and +15. 
-Each feature gets its own dedicated table. 
-Our proposed design uses 9 features. 
-As the program execution begins, all the weights are initialized to 0.
-\newline \newline
-\textbf{INFERENCING}\newline
-In the Champsim simulator, the prefetching method is triggered on every L2 Cache demand access. 
-At that point, the prefetcher has the option to do or not do prefetch - in case it does, then how many cache lines to prefetch. 
-The suggested prefetches can be places either in L2 Cache or in Last Level Cache. 
-The prefetcher gets to decide this, and is usually done based on an internal confidence mechanism.
-
-When SPP gets triggered, it suggests up to 4 (=Associativity of the Pattern Table) candidates for prefetch based on the demand access address. 
-In each subsequent lookahead depth, SPP suggests up to 4 more prefetch candidates, as explained in section\_\_. 
-All these recommended prefetches are tested in the perceptron throttler before they can access the Prefetch Filter, as in the original SPP data-path.
-
-The perceptron looks at the micro-architectural state, or `features', at that instant. 
-Each feature indexes into the hash of up to 4096-entry table dedicated for that feature. 
-Feature \#1 indexes into a particular entry in table \#1 and so on.
-
-Once all the weights are retrieved, they are added together, much like the traditional perceptron dot product. 
-On obtaining the overall sum, it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO. 
-Only the prefetch candidates with the perceptron sum higher that the threshold qualify for prefetching. 
-The prefetches that qualify the perceptron throttler stage go through the Prefetch Filter stage, just as in the SPP design.
-The Prefetch Filter eliminates any redundant prefetch suggestions.
-In SPP-Perc, all the meta-data that is required to recreate the features is also stored in the Prefetch Filter. 
-At a later stage, when the feedback of the current prefetched line is available, the stored data can be used to train the perceptron.
-
-The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch cache line. 
-All prefetch candidates qualified till this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill level.
-The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
-
-The perceptron sum can also be seen as the sum of the individual contribution per feature.
-The value of each contribution corresponds to the extent of confidence of the final output with that given feature.
-By summing the individual contributions, the final perceptron sum denotes the overall confidence for that prefetch suggestion. 
-In a way, by thresholding the perceptron sum on two different thresholds, we are diving the confidence scale of the prefetch suggestion into three bins.
-The first bin corresponding to the lowest confidence leads to prefetch candidate being rejected.
-The next bin corresponds to prefetches with a moderate confidence level. 
-Such prefetches are directed to fill the bigger Last Level Cache and potentially not pollute the more scarce L2 Cache.
-The highest confidence prefetches are filled in the L2 Cache.
-
-In addition to the Prefetch Filter mentioned above, SPP-Perc also maintains a `Reject Filter'. 
-Reject filter is a 1024-entry deep 1-way associative table. 
-In case that a prefetch suggestion is rejected by perceptron throttler, it instead gets logged into the Reject Filter. 
-The Reject Filter keeps track of 1024 entries that were suggested by the Pattern Table but got rejected by the perceptron throttler. 
-This can be used to train the perceptron to avoid false negatives \textit{i.e.}, cases where prediction was to reject the prefetch but the prefetch would have been useful.
-
-As previously stated, an enhancement to the original SPP design includes storing all the meta-data describing the program state at the instant of prefetching, in the Prefetch Filter. 
-This information is useful when the perceptron needs to be updated subsequently. 
-\newline \newline
-\textbf{TRAINING}\newline
-In the prefetching environment, feedback of a prefetch is received whenever there is an eviction or a demand access from L2 Cache. 
-This acts as a trigger to update all the baseline SPP entries and to also train the perceptron throttler. 
-Training of the perceptron throttler involves restoring the meta-data that was stored in Prefetch Filter or the Reject Filter. 
-This is done by indexing into the corresponding filters using the cache line address of the block being used to train. (10-bits indexing + 6-bits tag matching). 
-Once the state of the program at the time of prefetching is available, it is used to index into the perceptron weights table.
-
-If a demand access block which triggers the training  was tagged as a valid prefetch in the Prefetch Filter, it means that the earlier prefetch prediction was correct. 
-In that case the perceptron weights are incremented by 1 if the predicted sum does not cross a pre-defined threshold. 
- In case it is a cache block eviction that led to training and the corresponding valid entry was found in the Prefetch Filter, it shows that the prediction made by the perceptron was wrong. 
-Perceptron should have ideally rejected the prefetch suggestion as a low-confidence prefetch. 
-Here the weights are decremented by 1 to reflect the same. 
-In either case, weights are saturated at -16 or +15.
-
-A secondary training mechanism also kicks in during demand fetches. 
-Before the demand access triggers the next set of prefetches, it checks the Reject Filter for the valid entry. 
-In case it is a hit, that means that the corresponding cache line was initially suggested by the underlying SPP engine but rejected by the perceptron throttler. 
-This means that perceptron should have been more confident about that particular prefetch. 
-Once such a scenario is identified, the state of the execution at the time of prefetch is retrieved from the Reject Table. 
-The retrieved data is used to index into the various weights tables of the perceptron and the corresponding values are updated by +1, saturating between and -16 and +15. 
-This update reflects increased confidence for the prediction corresponding to that prefetch.
-
-This mechanism allows us to solve the classical problem of exploiting `lost opportunity'. 
-In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
-Our two-step SPP-Perc allows us to overcome that issue in a simple manner.
-
-\subsection{Features used by Perceptron}
-
-Here we discuss the various features which help correlate the prefetching decision with the program behaviour
-\begin{itemize}
-\item \textbf{Base address} of the demand access which triggered the prefetch. 
-Since prefetch are triggered from the L2 demand access, they tend to be correlated with the triggering address.
-
-\item \textbf{Cache line} and \textbf{Page address}: These two separate features are derived from the base address which triggered the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and base\_addr >> LOG2\_PAGE\_ SIZE respectively.
-The idea behind using three different shifted versions of the same feature is that it allows us to look into a wider range of bits than with a single version.
-It also helps give more importance to the overlapping bits and lesser importance to most and least significance bits.
-This approach can also eliminate destructive interference that can be caused by directly folding the bits into half.
-
-\item \textbf{Confidence}: The integer confidence on the scale of 0 to 100 which was used in the original SPP design.
-While the original confidence is not used directly for decision making, it can still contribute to the final decision made.
-
-\item \textbf{Program Counter XOR Depth}: PC is for the instruction which triggered the prefetch chain.
-Depth refers to the iteration count of the lookahead stages.
-In general, PC is not considered as a good basis for doing lookahead prefetching as all the prefetches with depth >= 1 gets aliased onto the same PC which won't be true in actual demand fetch case.
-This feature resolves a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases.
-This is akin to the concept of Virtual Program Counters\cite{VPC} introduced by Kim \textit{et. al.} for indirect branch prediction.
-
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC before the instruction that triggered the current prefetch. 
-Hashing together the last three PCs tell SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block.
-PCs are shifted in the increasing order of history before being hashed together.
-This is done to avoid the resultant value of zero in case 2 or more PCs are the same. 
-Additionally, obfuscating the information as it gets old allows us to get a wider and an approximate look into the program history.
-
-\item \textbf{Program Counter XOR Delta}: This tells us if a given PC favours particular value(s) of delta.
-As noted earlier, while PC in itself does not much useful information, this hash resolves PC into different values based on the tendency of that PC to favour a certain delta.
-This way, dynamic nature of different instances of the same memory access instruction can be captured here.
-
-\item \textbf{Current Signature XOR Delta}: It can be recalled from the discussion of SPP in section XX that the new signature is generated using the old signature and the blobk delta.
-While `Current Signature XOR Delta' is not the exact formula for generating the future signature, it gives an approximate idea of the path that the combination of these two values can lead to.
-
-%%%%%%% Below are the features rejected when downsizing 14 -> 9
-%\item \textbf{Current Signature}: The 12-bit signature used by SPP to index into the Pattern Table
-%\item \textbf{Delta}: The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched
-%\item \textbf{PC}: Program Counter of the Load instruction that triggers that particular prefetch
-%\item \textbf{Base address XOR Delta}: If any particular value(s) of delta are highly favoured by a given address which triggered the prefetch, this composite feature can capture that information
-%\item \textbf{Cache line XOR Delta}: This feature captures if any particular delta is highly favoured by a given cache line address
-%\item \textbf{Confidence XOR Delta}: This gives a combined insight into that fact that certain delta-confidence pairs predicted by SPP might correlate more with the prefetch outcome
-%%%%%%%%%%%%%%%%%%%%%%
-
-\end{itemize}
-
-Some of the composite features are derive from simple hashing (XOR) of two primary features.
-There is always a question of usefulness of such composite features and the `fresh' information added.
-We try to justify the choice of each feature by quantifying the contribution made towards predicting prefetch behaviour, in Section XX.
-Finally, as noted above, each feature indexes into its independent entry of perceptron weights.
-
-\section{METHODOLOGY}
-\subsection{Performance Model}
-For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim} simulator. 
-ChampSim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} 
-We model model a 1-core / 4-core / 8-core Out of Order machine. The details of the configuration parameters is summarized in Table  \ref{tab:Sim_params}.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|l|p{3.6cm}|c|}
-    \hline
-    Level & Configuration & Access Latency \\
-    \hline
-         L1 Cache & Separate I-cache and D-cache, 32 KB, 8-way & 4 cycles\\
-         L2 Cache & Private, 256 KB & 8 cycles\\
-         LLC & Shared, 2MB / core, 16-way & 20 cycles\\
-         DRAM & 4 GB Single Channel for single-core, 8 GB Double Channel for multi-core & N?? cycles\\
-    \hline
-    \end{tabular}
-    \caption{Memory Model Parameters}
-    \label{tab:Sim_params}
-\end{table}
-
-The block size was fixed 64KB. Prefetching was only triggered at L2 Cache demand accesses but could be sent to L2 Cache or Last Level Cache. 
-No L1 Data level prefetching was done. 
-LRU replacement policy was used on all levels of cache hierarchies. 
-Branch prediction was done using the perceptron branch predictor\cite{Perc_Branch}. 
-Page size was kept as 4KB. 
-Champsim operates all the prefetchers strictly in the physical address space.
-
-\subsection{Testing Under Additional Memory Constraints}
-The default single-core configuration simulates a 2MB LLC and a single channel DRAM with 12.8GB/s bandwidth. 
-We extend the simulations to include memory constraints introduced in DPC-2. 
-Specifically we look at the following two variations:
-\begin{itemize}
-    \item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited to 3.2 GB/s
-    \item \textit{Small LLC}: In this scenario, LLC size is reduced to 512 KB
-\end{itemize}
-All the multi-core simulations are only done in the default configuration.
-
-\subsection{Workloads}
-This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017} has been used to characterize and measure the prefetch performance. 
-We use all the 20 workloads available in the SPEC 2017 Suite. 
-Using SimPoint\cite{SimPoint} methodology, we identified 95 different program segments of 1 Billion instructions each.
-
-We performed cross-validation of our SPP-Perc model using SPEC 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.
-For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
-For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.
-The traces include 4 multi-core applications with 6 distinct phases per application. 
-In total, we used 285 traces representing workloads across 53 applications. 
-\newline
-\newline
-\textit{Single-core performance:} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics.
-We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
-We take the Instruction Per Cycle (IPC) of the prefetcher and divide it with the IPC of baseline. 
-The final number reported is the geometric mean of the speed-up attained on individual traces.
-\newline
-\newline
-\textit{Multi-core performance:} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
-For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 
-Each CPU keeps executing its workload till the last CPU completes 1 Billion instructions after warm-up. 
-For collecting IPC and other data, only the first billion instructions are considered as the ``region of interest''.
-
-Here we report the weighted speedup normalized to baseline \textit{i.e.}, no prefetching. 
-For each of the workloads running on a particular core of the 4-core 8 MB LLC system, we compute IPC\textsubscript{i}. 
-We then find the IPC\_isolated\textsubscript{i} of the same workload running in isolated 1-core 8 MB LLC environment. 
-Then we calculate the total weighted-IPC for a given workload mix as $\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}). 
-For each of the 100 workload-mix, the sum obtained is normalized to the weighted-IPC calculated similarly for baseline case \textit{i.e.}, no prefetching, to get the weighted-IPC-speedup. 
-Finally the geometric mean of these 100 weighted-IPC-speedup is reported as the effective speedup obtained by the prefetching scheme.
-
-We repeat the same process for 8-core workloads with the exception that 20 Million warm-up instructions and 100 Million full instructions were executed.
-This was done so as to keep the simulation run-time within reasonable limits as a single 8-core mix was taking up to 5 days to simulate 1 Billion instructions.
-
-\subsection{Preferchers Simulated}
-We compared SPP-Perc against three of the latest state of the art prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher (SPP).
-BOP was the winner of 2nd Data Prefetching Championship. 
-DA-AMPM is the enhanced version of AMPM, modified to account for DRAM row buffer locality. 
-SPP has been shown to outperform BOP on SPEC 2006 traces. 
-For each of these, we compare their speedups taking no prefetching as the baseline.
-
-\subsection{Developing Features for SPP-Perc}
-This section describes the intuition and analysis that went behind finalizing the features. 
-As noted earlier, we developed a set of 9 features which allow the perceptron throttler to correlate prefetching decision with the program context. 
-To study the correlation across each feature, we study statistically the perceptron weights and try to interpret their distribution.
-\newline  \newline
-\textbf{GLOBAL PEARSON'S CORRELATION}\newline
-For this experiment, we create a dump of the state of perceptron weights at the end of all trace execution.
-The weights obtained from running all the SPEC 2017 traces are concatenated together. 
-Since the weights are collected at the end of individual trace execution, the perceptron weights have attained a relatively stable value by now.
-Hence, this dump represents a `snap' of the trained perceptron weights across all the SPEC 2017 traces.
-The intuition here is that the feature with bulk of the perceptron weights concentrated around 0 or small magnitude numbers show a weak correlation with the prefetching outcome. 
-On the other hand, features with most of the weights saturated around highest value (+15) show a high positive correlation and the features with weights close to the lowest value (-16) show a strong negative correlation.
-
-We plot a histogram for each feature depicting weights distribution from -16 to +15 and use this histogram to generate the Pearson's correlation factor for that feature.
-In statistics, Pearson's factor is a numerical measure of the degree of linear correlation between two variables.
-It ranges from -1 to +1.
-The magnitude of Pearson's factor depicts the extent of correlation and the sign depicts whether it is a positive correlation or a negative correlation.
-Values close to 0 suggest a low correlation or at times, noisy data.
-A value of +1/-1 suggests a perfectly linear positive / negative correlation respectively.
-Figure XX depicts the histogram and the Pearson's coefficient for two of the features, YY and ZZ.
-
-Figure XX shows all the features used, arranged in the increasing order of their Pearson's factor.
-As can be seen X out of the 9 features provide a moderate to high correlation, with the magnitude of P-value > 0.4.
-The single most important feature, \_\_ helps provide a correlation to prefetch outcome with a factor of \_\_.
-\newline \newline
-\textbf{PER TRACE CORRELATION} \newline
-Another important way to look at the perceptron features is to see how much their contribution varies across the traces.
-Here we give special attention to features with low P-values in the previous experiment.
-Figure XX shows the variation of P-values three features : \_\_, \_\_ and \_\_; across all the SPEC 2017 traces.
-For simplicity, the traces are arranged in an increasing order of contribution made by the feature and the trace names have been omitted.
-It can be seen that even features with a low overall correlation provide useful correlation (magnitude > 0.4) for XX out of the 95 simpoints developed for SPEC 2017.
-\newline \newline
-\textbf{TRIMMING FEATURES USING CORRELATION} \newline
-Besides providing interesting insights into prefetching behaviour, P-value can also be used for feature selection and prefetcher tuning.
-
-Here, we introduce the concept of cross-correlation across features.
-Just as we examined correlation of each feature with the final outcome, we can also study correlation between the features.
-We used the above methodology to eliminate features which provide a little or no information that has already been captured in other features.
-
-As a part of feature selection, we initially came up a mix of 23 features- primary and composite.
-By studying cross correlation of each of these features against others in a 23x23 matrix, we identified pairs of features with correlation factor > 0.9 in magnitude and carefully eliminated redundant features.
-Using this approach, we managed to reduce the feature count to 9.
-Thus, in the final implementation of SPP-Perc, no two features have a high correlation between them.
-This way we can be sure that each feature is bringing in contribution that cannot be captured using other features.
-
-\textit{<TODO: VALIDATE THIS>} 
-Secondly, studying the relative importance of each feature enabled us to vary the number of entries dedicated for each feature.
-Features with higher Pearson's correlation (like \_\_ and \_\_) were given most importance and allowed full 12-bits of indexing.
-Features like \_\_ and \_\_, with a low overall P-value were allocated fewer indexes in the feature table.
-
-To conclude, in the above discussion we justify the features for perceptron from a statistical viewpoint.
-We also show how this information can also be used for prefetcher tuning.
-All this study was made possible only because we used on-line perceptron learning for prefetching and that enabled us to examine the weights in detail.
-
-\subsection{Overhead for SPP-Perc}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|c|c|m{4.8cm}|}
-    \hline
-        \textbf{Field} &
-        \textbf{Bits} &
-        \textbf{Comment} \\
-    \hline
-         Valid & 1 & Indicates a valid entry in the filter\\
-         Tag & 6 & Identifier for the entry in the filter\\
-         Useful & 1 & To show if the given entry led to a useful demand fetch\\
-         Perc Decision & 1 & Prefetched vs Not-prefetched \\
-    \hline
-        PC & 12 & \multirow{5}{4.8cm}{Meta-data required for perceptron training}\\
-        Address & 24 & \\
-        Delta & 7 & \\
-        Curr Signature & 12 & \\
-        Confidence & 7 & \\
-    \hline
-        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
-    \hline
-    \end{tabular}
-    \caption{Meta-data Stored in Prefetch Filter}
-    \label{tab:PF_metadata}
-\end{table}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|p{1.5cm}|p{1cm}|c|p{1.5cm}|}
-    \hline
-        \textbf{Structure} &
-        \textbf{Entry} &
-        \textbf{Components} &
-        \textbf{Total} \\
-    \hline
-        \multirow{5}{1.5cm}{Signature Table} & \multirow{5}{1.5cm}{256} & Valid (1 bit) & \multirow{5}{1cm}{11008  bits}\\
-        & & Tag (16 bits) &\\
-        & & Last Offset (6 bits) &\\  
-        & & Signature (12 bits) &\\
-        & & LRU (6 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Pattern Table} & \multirow{3}{1.5cm}{512} & $C_{sig}$ (4bits) &\multirow{3}{1cm}{24576 bits}\\
-        & & $C_{delta}$ (4*4 bits) &\\
-        & & Delta (4*7 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Perceptron Tables} & \multirow{3}{1.5cm}{4096*8 128*1} & \multirow{3}{1.5cm}{5 bits} & \multirow{3}{1.5cm}{164480 bits}\\
-        & & &\\
-        & & &\\
-    \hline
-        \multirow{2}{1.5cm}{Prefetch Filter \footnotemark[1]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{84 bits} & \multirow{2}{1.5cm}{86016 bits}\\
-        & & &\\
-    \hline
-     \multirow{2}{1.5cm}{Reject Filter \footnotemark[2]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{83 bits} & \multirow{2}{1.5cm}{84992 bits}\\
-        & & &\\
-    \hline
-        \multirow{4}{1.5cm}{Global History Register} & \multirow{4}{1.5cm}{8} & Signature (12 bits) & \multirow{4}{1.5cm}{264 bits}\\
-        & & Confidence (8 bits) & \\
-        & & Last Offset (6 bits) & \\
-        & & Delta (7 bits) & \\
-    \hline
-        \multirow{2}{1.5cm}{Accuracy Counters} & 1 & C$_{total}$ & 10 bits\\
-        & 1 & C$_{useful}$ & 10 bits\\
-    \hline
-        \multirow{3}{1.5cm}{Global PC Trackers} & \multirow{3}{1cm}{3} & $PC_1$ (12 bits)& \multirow{3}{1.cm}{36 bits}\\
-        & & $PC_2$ (12 bits) &\\
-        & & $PC_3$ (12 bits) &\\
-    \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 371356 bits = 45.33 KB}}\\
-    \hline
-    \end{tabular}
-    \caption{SPP-Perc Storage Overhead}
-    \label{tab:SPPPerc_overhead}
-\end{table}
-\footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}
-\footnotetext[2]{RF does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
-
-In this section, we analyze the hardware overhead required to implement SPP-Perc. 
-The Prefetch Filter was enhanced to accommodate storing of meta-data for perceptron training. 
-Table \ref{tab:PF_metadata} depicts the meta-data stored for each entry in the Prefetch Filter.
-Table \ref{tab:SPPPerc_overhead} shows the total storage overhead of SPP-Perc implementation.
-The hardware budget for 2nd Data Prefetching championship was 32 KB. 
-Keeping that in mind and the speedup SPP-Perc obtained over the winner, the extra hardware budget can be accounted for.
-The extra hardware also makes the overall scheme more scalable than SPP. 
-In the original SPP paper, it was demonstrated that adding extra hardware brings little advantage in terms of performance gain.
-The newly added perceptron tables can be scaled to increase / decrease features depending on the permitted budget.
-
-In terms of computations, the perceptron mechanism only introduces an extra adder tree.
-The hash perceptron mechanism makes sure than there is no actual vector multiplication happening in the hardware.
-Obtaining the perceptron sum requires addition of 9 5-bit numbers. 
-Using an adder tree of 4 5-bit adders, this can be done in ceil($log_{2}9$) = 4 steps.
-Perceptron update only requires weight update by +1 or -1.
-Thus, all the operations required for perceptron inferencing or updating the states of the perceptron throttler can be easily done in the time constraints of L2 Cache Accesses.
-
-\section{RESULTS}
-This section discusses the results obtained from running SPP-Perc, in terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per Cycle (IPC) metrics. 
-For SPEC 2017 suite, first we present the results for single-threaded workloads then the results for multi-core workloads are presented.
-
-\subsection{Single-core Results}
-Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated. 
-All the results have been normalized to baseline \textit{i.e.}, no prefetching. 
-As can be seen, SPP-Perc yields a speedup of XX\% over the baseline. 
-This is equivalent to XX\% over DA-AMPM, xx\% over BOP and XX\% over SPP. 
-Out of the XX benchmarks, SPP-Perc nearly matches or outperforms most of the prefetchers on XX traces.
-
-At its peak, SPP-Perc manages to get the IPC speedup of a factor of over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}. 
-This also corresponds to speedup gain of \textbf{22.4\%} over the next best prefetcher - SPP. 
-In general, benchmarks \textit{602.gcc\_s, 603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s} benefit the most from SPP-Perc, as compared to SPP only prefetching.
-
-<Discuss the extreme behaviour of \textit{623.xalancbmk\_s}: Improvement over SPP is as high as 31\% and even reduces to -1.5\% on traces of this benchmark>
-
-<Discuss about coverage on all traces>
-
-<Discuss cases where coverage gets specially increased>
-
-<Discuss accuracy on all traces>
-
-<Discuss about compelling increase in accuracy - details of traces. Plot linear separability>
-
-<Any case where accuracy got reduced - go into the trace. See if linear separability is the issue. Other possible issues?>
-
-\subsection{Multi-core Results}
-In this section, we demonstrate the improvement achieved by SPP-Perc for a mix of multi-programmed workloads. 
-\newline \newline
-\textit{4-core environment}
-Figure \_\_ shows a comparison of speedups obtained on workload mix of memory intensive subset of SPEC 2017. 
-We plot all the 4 prefetchers, normalized to baseline. 
-The workloads have been sorted in the increasing order of performance benefit, as measured in terms of normalized weighted speedup. 
-As can be seen, SPP-Perc offers an IPC improvement by 33.5\% on these traces. 
-This is 5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP and DA-AMPM respectively.
-
-This increased improvement by SPP-Perc on a 4-core machine is not surprising. 
-Since SPP-Perc has an extremely intelligent throttler working to eliminate useless prefetches before they can create cache pollution. 
-This further amplifies the edge that SPP had over BOP. 
-With the advantage of an orthogonal prefetcher, SPP can be as aggressive as possible and leave the accuracy aspect to perceptron throttler.
-\newline \newline
-\textit{8-core environment}
-To test the effectiveness of SPP-Perc in a tighter constrained multi-core environment, we simulated an 8-core machine with 16 MB last Level Cache. 
-Other parameters were kept same as discussed under Methodology in Section XX.
-<See if we can use speed-up over BOP as a measure here>
-
-\subsection{Additional Memory Constraints}
-Figure \_\_ shows the performance of SPP-Perc against the other prefetchers in tighter constraint environments. 
-For sake of simplicity, we show 2 traces where SPP-Perc gives the most advantage over baseline, 2 traces where SPP-Perc doesn't perform as well as the best prefetcher *<-replace with the prefetcher name once results are available* and finally GeoMean across all the traces.
-<Hopefully BOP would degrade more here as it is super aggressive. 
-By definition, SPP-Perc should be better than SPP as the throttler would better control the aggressiveness. 
-In SPP paper DA-AMPM performs close enough to BOP. That should get replicated here too.>
-
-\subsection{Cross Validation}
-Here we try to demonstrate the robustness of SPP-Perc by testing it on completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
-
-Fig \_\_ shows the speed-up achieved by BOP, SPP and SPP-Perc on selected traces of SPEC 2006, along with the overall geometric mean over whole of the benchmark and over memory-intensive subset of the SPEC 2006 benchmark.
-As can be seen, SPP-Perc peovides a speed-up of 44.4\% over baseline, on the memory intensive subset of SPEC 2006 benchmark.
-This corresponds to 6.84\% over SPP and XX\% over BOP.
-On the whole of the SPEC 2006 suite, the speedup obtained is 22.4\%, which is 3.34\% ahead of SPP.
-
-The 4-core results for SPEC 2006 are \_\_.
-
-As can be seen, the performance is very consistent given no prior parameter tuning was done keeping SPEC 2006 in mind. 
-We attribute this to the inherent adaptability of the perceptron model.
-In general, perceptron weights are able to adjust in real-time so as to find the best possible correlation between the output and the features.
-
-\section{CONCLUSION}
-In this paper, we introduced perceptron based learning for data prefetching in the form of SPP-Perc. 
-Perceptron acts an orthogonal throttler to the underlying prefetch engine. 
-SPP-Perc improves performance by over XX\% over the baseline, which corresponds to YY\% over the next best performing prefetcher. 
-We believe this paper can lead to interesting future work like generalizing SPP-Perc to work with any prefetcher with minimal knowledge of the underlying prefetcher architecture.
-
-%%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%
-
-%%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
-\bibliographystyle{ieeetr}
-\bibliography{ref}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}

### 2018-11-15 back.tex
reorg

--- 2018-11-15 paper_1.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is the template for submission to ISCA 2018
-% The cls file is a modified from  'sig-alternate.cls'
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{sig-alternate} 
-\usepackage{mathptmx} % This is Times font
-
-\newcommand{\ignore}[1]{}
-\usepackage{fancyhdr}
-\usepackage[normalem]{ulem}
-\usepackage[hyphens]{url}
-\usepackage{microtype}
-\usepackage{fixltx2e}
-\usepackage{array}
-\usepackage{multirow}
-
-% Always include hyperref last
-\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=black,citecolor=blue,urlcolor=black]{hyperref}
-
-% Ensure letter paper
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\newcommand{\iscasubmissionnumber}{NaN}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\fancypagestyle{firstpage}{
-  \fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-  \fancyhead[C]{\normalsize{}} 
-  \fancyfoot[C]{\thepage}
-}
-
-\pagenumbering{arabic}
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\title{SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher} 
-\author{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\maketitle
-\thispagestyle{firstpage}
-\pagestyle{plain}
-
-\begin{abstract}
-\textbf{Hardware prefetching has been introduced in modern processors as a way to hide cache latencies. 
-An efficient prefetcher should be able to identify complex memory access patterns during a program execution. 
-This enables it to fetch a block ahead of its demand access, potentially saving a cache miss.
-}
-
-\textbf{In this paper, we introduce perceptron learning to help make this prefetching decision efficiently. 
-We have taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine. 
-The perceptron layer acts as a throttler to filter out the unnecessary prefetches recommended by SPP. 
-We have also explored a range of features that can be used to train the perceptron layer. 
-Our results show that SPP-Perc improves performance on the memory intensive subset of the SPEC 2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared to just SPP.
-We also demonstrate that the performance gained from using our efficient throttler scales better with increasing number of cores.
-}
-\end{abstract}
-
-\section{Introduction}
-Since past few decades, processor and memory scaling has been enabled with different philosophies in mind. 
-While processor scaling focused on speed improvements, memory scaling was centred around increasing the storage capacity. 
-The process technology reaching scaling limitations led to Memory Wall\cite{MemWall} - the increasing performance gap between the processor speed and the memory speeds. 
-Various techniques have been developed to combat this widening gap. 
-Data prefetching is one such important technique. 
-Data prefetching exploits the fact that in most applications, memory accesses have a well-defined pattern.
-
-An ideal prefetching scheme involves capturing this memory access patterns to predict future prefetches in a timely manner. 
-Other important decisions faced by the prefetcher are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block.
-Memory accesses patterns can be as simple as fetching the next cache line or a complex but predictable pattern, like in pointer chasing applications. 
-Predicting this often requires a trade-off between coverage and accuracy. 
-For a prefetcher to capture highly irregular access patterns, it needs to be highly aggressive (high coverage). 
-That also means the prefetcher would end up recommending a lot of useless prefetches, leading to cache pollution (low accuracy). 
-This effect would get even more pronounced in multi-core scenario where the last level cache is a shared resource. 
-Conversely, a conservative prefetcher would be highly accurate but might not make enough useful predictions.
-
-To this effect we propose SPP-Perc <come up with a catchy name>. 
-It is an enhancement to the existing state of the art prefetcher and overcomes the coverage vs accuracy trade-off in a robust manner. 
-The idea is to modify SPP design to make it as aggressive as it can get. 
-This enables SPP to capture complex memory access patterns hence leading to increased coverage.
-This comes at a cost of increased DRAM traffic, hence reduction in overall IPC.
-The prefetch suggestions recommended by SPP pass through the perceptron based throttler. 
-Over time, perceptron layer learns to correlate the prefetches recommended by SPP with various available features, leading to elimination of useless prefetches. 
-This leads to an overall increased accuracy of the prefetch scheme and reduction in DRAM traffic.
-
-\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and future scope for expansion. 
-The contributions made by this paper are:
-
-\begin{itemize}
-\item This is the first time where a truly on-line learning based neural network model has been used when it comes to hardware data prefetching. 
-The prior works in this area either relied on program semantics\cite{Semantics} or were application specific\cite{Datacenter} or could not yield a significant improvement over the baseline prefetcher used\cite{BadPerc}.
-
-\item Implementing the pereptron throttler in a robust and accurate practical prefetching mechanism. 
-The overall scheme yields increased accuracy as compared to the other state of the art prefetchers.
-Moreover, the perceptron throttler learns to adapt itself to shared resource constraints - leading to an even increased performance in multi-core and lower capacity / bandwidth environments.
-
-\item Examining the perceptron weights doing a statistical study to quantify the contribution made by each feature.
-We have also used this idea to develop a comprehensive set of features where each feature contributes some `new' information to the perceptron.
-The details are explained under methodology, in Section XX.
-
-\item Cross-validating the design across benchmarks. 
-All the development and parameter tuning was done on the SPEC 2017 benchmark. 
-The results have been shown to be consistent across unseen benchmarks.
-More details can be found in the results section (Section XX).
-
-%\item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind. 
-%We have used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers. 
-%We have also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
-
-%\item We came up with an efficient model to use the perceptron output as a representation of the prefetch confidence using an on-line histogram technique. 
-%We show how the simple sum can replace the existing mechanism which was essentially based on counting the number of hits for a particular entry in SPP.
-
-\end{itemize}
-
-We tested SPP-Perc on Champsim, an open source simulator which models an Out-of-Order CPU using. 
-For evaluation, we used SPEC 2017 benchmarks. 
-On a single core configuration, SPP-Perc increases performance of the system by 39.5\% as compared to the baseline with no prefetching, on the memory intensive subset of the benchmark.
-This outperforms the next best prefetcher, SPP,  by 6.84\%. 
-On a multi-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement of 11.9\% over SPP for 4-core system and 10.67\% for 8-core system.
-
-In the results part in section XX, we show a detailed analysis of the speedup achieved.
-We also used SPEC 2006 and CloudSuite benchmarks for cross-validation of the prefetcher implementation.
-
-The rest of the paper is organized as follows.
-Section 2 talks about background work and recaptures the SPP architecture
-Section 3 discusses the detailed implementation of the perceptron mechanism and the features used.
-Section 4 consists of the evaluation methodology and explores the feature space for perceptron learning.
-Section 5 is the performance evaluation of our prefetcher and Section 6 concludes the paper.
-
 \section{Background Work and Motivation}
 
-In this section, we compare the existing work done in the domain of prefetching. 
-We also discuss about some of the interesting memory access patterns seen in the workloads in SPEC 2017 benchmark. 
-The second half of this section deals with a background of Signature Path Prefetcher architecture.
+In this section, we compare the existing work done in the domain of
+prefetching.  We also discuss about some of the interesting memory
+access patterns seen in the workloads in SPEC 2017 benchmark.  The
+second half of this section deals with a background of Signature Path
+Prefetcher architecture.
 
@@
 
-The idea of prefetching has been around since Instruction Stream Buffers were proposed by Jouppi.
-\cite{ISB}
-Some of the earliest data prefetchers developed to try to identify memory access patterns with a constant stride pattern.
-\cite{Smith} 
-\cite{Baer}
-\cite{Stride}
-Major research was dedicated to stride distance and depth prediction.
-\cite{Decoupled}
-\cite{Adaptive}
-Newer prefetchers aimed at correlating predictions with the past memory access addresses.
-\cite{Address_Correlated} 
-\cite{AMPM}
+The idea of prefetching has been around since Instruction Stream
+Buffers were proposed by Jouppi.  \cite{ISB} Some of the earliest data
+prefetchers developed to try to identify memory access patterns with a
+constant stride pattern.  \cite{Smith,Baer,Stride} Major research was
+dedicated to stride distance and depth prediction.
+\cite{Decoupled,Adaptive} Newer prefetchers aimed at correlating
+predictions with the past memory access addresses.
+\cite{Address_Correlated,AMPM}
 
-With time newer and more intricate prefetchers were introduced.
-\cite{Wenisch_Temporal_Streaming}
-\cite{Stealth}
-\cite{Feedback_Directed}
-\cite{Coordinated}
-\cite{Bandwidth_Efficient}
-\cite{Pacman}
-\cite{TLB}
-\cite{Linearizing}
-\cite{Sandbox}
-\cite{VLDP}
-\cite{DoL}
-\cite{Domino}
-A whole class of streaming prefetchers developed on the lines of Spatial / Temporal Memory Streaming.
-\cite{Spatial_Pattern}
-\cite{SMS}
-\cite{Temporal_Instruction_Fetch}
-\cite{Off_Chip}
-\cite{STMS}
-\cite{SMS_JILP}
+With time newer and more intricate prefetchers were
+introduced~\cite{Wenisch_Temporal_Streaming,Stealth,Feedback_Directed,Coordinated,Bandwidth_Efficient,Pacman,TLB,Linearizing,Sandbox,VLDP,DoL,Domino}.
+A whole class of streaming prefetchers developed on the lines of
+Spatial / Temporal Memory
+Streaming~\cite{Spatial_Pattern,SMS,Temporal_Instruction_Fetch,Off_Chip,STMS,SMS_JILP}.
 
-Most modern prefetchers aim at identifying complex memory access patterns in a given application. 
-This allows them to capture some of the irregular patterns seen in pointer chasing data-structures.
-Offset based prefetchers are a generalization of the next-line prefetcher.
-Some of the prominent prefetchers exploiting this idea are
-Sandbox Prefetcher \cite{Sandbox}
-and the Best Offset Prefetcher \cite{BOP}
-Another class of prefetchers is lookahead Prefetchers.
-Such prefetchers try to speculate deep into the application's memory access trail.
-These include Path Confidence based Lookahead Prefetching \cite{SPP}
-and Kill the Program Counter \cite{KPC}.
+Most modern prefetchers aim at identifying complex memory access
+patterns in a given application.  This allows them to capture some of
+the irregular patterns seen in pointer chasing data-structures.
+Offset based prefetchers are a generalization of the next-line
+prefetcher.  Some of the prominent prefetchers exploiting this idea
+are Sandbox Prefetcher \cite{Sandbox} and the Best Offset Prefetcher
+\cite{BOP} Another class of prefetchers is lookahead Prefetchers.
+Such prefetchers try to speculate deep into the application's memory
+access trail.  These include Path Confidence based Lookahead
+Prefetching \cite{SPP} and Kill the Program Counter \cite{KPC}.
 
-A related line of work also explores cache replacement - insertion policies and dead block prediction, with an eye towards prefetching. 
-\cite{DB_Pred}
-\cite{Cache_Burst}
-\cite{KPC}
-\cite{Harmony}
-Nori \textit{et.al.} introduced prefetching with  respect to runtime criticality.
-\cite{CATCH}
-\newline
-\newline
-\textit{Machine Learning and Prefetching:}
-Peled \textit{et.al.} introduce interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
-\cite{Semantics}
-The biggest challenge here is that it relies on compiler support for getting features to build the context.
+A related line of work also explores cache replacement - insertion
+policies and dead block prediction, with an eye towards
+prefetching~\cite{DB_Pred,Cache_Burst,KPC,Harmony}.  Nori
+\textit{et.al.} introduced prefetching with respect to runtime
+criticality~\cite{CATCH}.
 
-The paper by Liao \textit{et. al.} focuses on prefetching for data center applications.
-\cite{Datacenter}
-They use offline Machine Learning Algorithms like SVMs and Logistic Regressions to do parametric search for an optimal prefetcher configuration.
+\textit{Machine Learning and Prefetching:} Peled \textit{et.al.}
+introduce interesting ideas for on-line Reinforcement Learning and
+dynamically scaling the magnitude of feedback given to the baseline
+prefetcher.  \cite{Semantics} The biggest challenge here is that it
+relies on compiler support for getting features to build the context.
+
+The paper by Liao \textit{et. al.} focuses on prefetching for data
+center applications.  \cite{Datacenter} They use offline Machine
+Learning Algorithms like SVMs and Logistic Regressions to do
+parametric search for an optimal prefetcher configuration.
 
 %% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?
-Finally, the paper on Data Cache Prefetching with Perceptron Learning\cite{BadPerc} talks about the idea of two step prefetching.
-The first step is an existing baseline prefetcher while the second step is a perceptron based throttler.
-The paper has lots of limitations in terms of design and implementation.
-That reflects in the fact that it did not lead to any significant performance gain over even the basic prefetchers like the Stride Prefetcher\cite{Stride} and the Markov Prefetcher\cite{Markov}.
+Finally, the paper on Data Cache Prefetching with Perceptron
+Learning~\cite{BadPerc} talks about the idea of two step prefetching.
+The first step is an existing baseline prefetcher while the second
+step is a perceptron based throttler.  The paper has lots of
+limitations in terms of design and implementation.  That reflects in
+the fact that it did not lead to any significant performance gain over
+even the basic prefetchers like the Stride Prefetcher\cite{Stride} and
+the Markov Prefetcher\cite{Markov}.
 
@@
 
-Perceptron learning for computer architecture design has been around for a while. 
-It was first popularized by Jimenez \textit{et. al.} for doing branch prediction. \cite{Perc_Branch} 
-In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
-The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. 
-The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the prediction is made as \textit{true}, or branch taken.
-Else if the sum is < 0, the prediction is made as \textit{false}.
+Perceptron learning for computer architecture design has been around
+for a while.  It was first popularized by Jimenez \textit{et. al.} for
+doing branch prediction. \cite{Perc_Branch} In it's original
+implementation, the Program Counter of the instruction looking for a
+branch prediction would index into a given table of perceptron
+weights.  The retrieved weights would then be multiplied with the
+history of branch outcomes stored in the History Register (which is a
+feature) in a dot product fashion.  The sum obtained is thresholded at
+0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the
+prediction is made as \textit{true}, or branch taken.  Else if the sum
+is < 0, the prediction is made as \textit{false}.
 
-The structure that we use is derived from the model introduced in Piecewise Linear Branch Prediction\cite{Piece_Linear}. 
-Here the feature itself is used to index into a hash of perceptron weights. 
-This form of indexing makes sure that the hyper-plane learned by the perceptron weights is able to differentiate between linearly inseparable outcomes. 
-\textit{<Is it true? Not sure>}
+The structure that we use is derived from the model introduced in
+Piecewise Linear Branch Prediction\cite{Piece_Linear}.  Here the
+feature itself is used to index into a hash of perceptron weights.
+This form of indexing makes sure that the hyper-plane learned by the
+perceptron weights is able to differentiate between linearly
+inseparable outcomes.  \textit{<Is it true? Not sure>}
 
-Perceptron Learning for Reuse Prediction\cite{Perc_Reuse} extended the hash perceptron architecture to perform prediction in context of cache replacement. 
-Perceptron layer can learn to correlate the cache replacement behaviour with a variety of features. 
-Each feature has its dedicated table of perceptron weights which it indexes into. 
-Hence number of tables equals the number of features. 
-Once the perceptron weights from all the tables are obtained, they're summed and thresholded based on a pre-decided value. 
-This led to a prediction mechanism that is highly accurate, has quick learning rate and is adaptable to changing program behaviour. 
-The features introduced in that paper were derived from Program Counters of the current and the last few instructions and the memory address of the current access. 
-Even with such basic features, high level of correlation could be established by perceptrons.
+Perceptron Learning for Reuse Prediction\cite{Perc_Reuse} extended the
+hash perceptron architecture to perform prediction in context of cache
+replacement.  Perceptron layer can learn to correlate the cache
+replacement behaviour with a variety of features.  Each feature has
+its dedicated table of perceptron weights which it indexes into.
+Hence number of tables equals the number of features.  Once the
+perceptron weights from all the tables are obtained, they're summed
+and thresholded based on a pre-decided value.  This led to a
+prediction mechanism that is highly accurate, has quick learning rate
+and is adaptable to changing program behaviour.  The features
+introduced in that paper were derived from Program Counters of the
+current and the last few instructions and the memory address of the
+current access.  Even with such basic features, high level of
+correlation could be established by perceptrons.
 
-The work was further extended by Multiperspective Reuse Prediction\cite{Multiperspective}, which introduced a varied set of parametric features. 
-Although design space exploration for feature and parameter selection was a non-trivial task, once fixed, it could give a cache replacement predictor highly suitable for the given set of applications.
-\newline
-\newline
-\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron prediction involves multiplying the vector of input features: F\textsubscript{1xN}, with the corresponding weight vector: W\textsubscript{Nx1} in a dot product fashion to obtain the sum y\textsubscript{out}.
-Here what we use is a perceptron-like structure. 
-The feature is used to hash into the weight of perceptrons and the retrieved weights are added straight away. 
-This can be seen a dot product of $\vec{1}$\textsubscript{1xN} with W\textsubscript{Nx1}. 
-This way, the perceptron algorithm doesn't introduce any multiplication operations in the inference or training process. 
-Hence what we adapt in this work, is a perceptron-like learning algorithm as it involves the same principle involved in perceptron inferencing and training. 
-For simplicity, we will call our implementation as perceptron implementation.
-\newline
-\newline
-\textit{Perceptron Update Rule:} In all the above implementations, a uniform perceptron update principle is being followed. 
-The weights need to be updated if the prediction was wrong or the predicted sum does not cross a certain threshold. 
-After a certain training period, these weights are proportional to the probability of the outcome being true. 
-This training threshold makes sure that the perceptrons are trained till a certain level of confidence and yet they are not over-trained to the given set. 
-Care is taken that the weights saturate at certain positive and negative values so that they remain confined to the bit-width. 
-The same rules will be applied in the discussion done in section XX on SPP-Perc learning algorithm.
+The work was further extended by Multiperspective Reuse
+Prediction\cite{Multiperspective}, which introduced a varied set of
+parametric features.  Although design space exploration for feature
+and parameter selection was a non-trivial task, once fixed, it could
+give a cache replacement predictor highly suitable for the given set
+of applications.
+
+\textit{Deviations From Actual Perceptrons:} Traditionally, a
+perceptron prediction involves multiplying the vector of input
+features: F\textsubscript{1xN}, with the corresponding weight vector:
+W\textsubscript{Nx1} in a dot product fashion to obtain the sum
+y\textsubscript{out}.  Here what we use is a perceptron-like
+structure.  The feature is used to hash into the weight of perceptrons
+and the retrieved weights are added straight away.  This can be seen a
+dot product of $\vec{1}$\textsubscript{1xN} with W\textsubscript{Nx1}.
+This way, the perceptron algorithm doesn't introduce any
+multiplication operations in the inference or training process.  Hence
+what we adapt in this work, is a perceptron-like learning algorithm as
+it involves the same principle involved in perceptron inferencing and
+training.  For simplicity, we will call our implementation as
+perceptron implementation.
+
+\textit{Perceptron Update Rule:} In all the above implementations, a
+uniform perceptron update principle is being followed.  The weights
+need to be updated if the prediction was wrong or the predicted sum
+does not cross a certain threshold.  After a certain training period,
+these weights are proportional to the probability of the outcome being
+true.  This training threshold makes sure that the perceptrons are
+trained till a certain level of confidence and yet they are not
+over-trained to the given set.  Care is taken that the weights
+saturate at certain positive and negative values so that they remain
+confined to the bit-width.  The same rules will be applied in the
+discussion done in section XX on SPP-Perc learning algorithm.
 
@@
 \subsection{Signature Path Prefetcher Architecture}
-Since the work we have done builds heavily on SPP\cite{SPP}, we are presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work. 
-Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists of following structural components:
-\newline
-\newline
-\textbf{Signature Table:} ST keeps track of 256 most recently accessed pages. 
-It is meant to capture memory access patterns within a page boundary. 
-SPP indexes into an entry of ST using the page number. 
-For each entry corresponding to a page, ST stores the `last block offset' and `old signature'. 
-Last block offset is the block offset of the last memory access of that given page. 
-The block offset is calculated with respect to the page boundary. 
-The signature is a 12-bit compressed representation of the past few memory accesses for that page. 
-The signature is calculated as:
+Since the work we have done builds heavily on SPP\cite{SPP}, we are
+presenting a brief recap of the ideas introduced in the original
+paper, especially with an eye towards their relevance in our work.
+Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists
+of following structural components:
+
+\textbf{Signature Table:} ST keeps track of 256 most recently accessed
+pages.  It is meant to capture memory access patterns within a page
+boundary.  SPP indexes into an entry of ST using the page number.  For
+each entry corresponding to a page, ST stores the `last block offset'
+and `old signature'.  Last block offset is the block offset of the
+last memory access of that given page.  The block offset is calculated
+with respect to the page boundary.  The signature is a 12-bit
+compressed representation of the past few memory accesses for that
+page.  The signature is calculated as:
 $$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
 In case a matching page entry is found, the stored signature in the corresponding entry is retrieved.
-\newline
-\newline
-Consider the case that incoming page number 10 with a block offset 3 finds a match in the ST. 
-The retrieved pattern signature is 0x30 and the Last Offset is 1. 
-Since now there is the Last Block Offset (1), Incoming Block Offset (3), Old Signature (0x30), and New Signature calculated as per above equation (0x182), SPP can infer non-speculatively that the given pattern of memory accesses (as captured in the Old Signature) leads to the particular Delta.
-In general, Delta is defined as the difference between the prefetch suggestion block and the initial block which triggered the prefetch. 
-In this case, since SPP is in learning phase, it is defined as the difference between the Incoming Block Offset and the Last Block Offset (+2 in this case). 
-This, in turn generates the new memory pattern (New Signature). 
-This newly learned signature pattern and the delta is stored in the Pattern Table.
-\newline
-\newline
-\textbf{Pattern Table:} The PT is indexed by the signature generated from ST. 
-PT holds predicted delta patterns and their confidence estimates. 
-Each entry indexed by the signature holds up to 4 unique delta predictions. 
-This is implemented by making PT as a 4-way associative table. 
-To measure an estimate of a delta confidence, PT implements two counters: C\textsubscript{sig} for each entry indexed by the signature and C\textsubscript{delta} for each delta entry for a given signature.
-More details of this on-line score-keeping mechanism is described under Confidence Tracking discussion.
 
-\textit{Lookahead Prefetching:} For each trigger to the prefetcher, the first time that the SPP indexes into the PT, it does so in the non-speculative manner. 
-We call it non-speculative because the parameters which SPP uses to predict, like last block offset, incoming block offset and signature are well-defined. 
-Assuming PT generates a valid delta prediction (rather, up to 4 valid deltas), SPP enters into the lookahead mode.
+Consider the case that incoming page number 10 with a block offset 3
+finds a match in the ST.  The retrieved pattern signature is 0x30 and
+the Last Offset is 1.  Since now there is the Last Block Offset (1),
+Incoming Block Offset (3), Old Signature (0x30), and New Signature
+calculated as per above equation (0x182), SPP can infer
+non-speculatively that the given pattern of memory accesses (as
+captured in the Old Signature) leads to the particular Delta.  In
+general, Delta is defined as the difference between the prefetch
+suggestion block and the initial block which triggered the prefetch.
+In this case, since SPP is in learning phase, it is defined as the
+difference between the Incoming Block Offset and the Last Block Offset
+(+2 in this case).  This, in turn generates the new memory pattern
+(New Signature).  This newly learned signature pattern and the delta
+is stored in the Pattern Table.
 
-In the lookahead mode, SPP looks at the last generated delta(s), chooses the one with the highest confidence and speculatively generates the new effective address and the new signature. 
-Using these new parameters, it re-accesses the PT to generate further prefetches. 
-SPP keeps on repeating this cycle of accessing PT and updating the signature based on highest confidence prefetch from the last iteration. 
-These iteration count till which SPP manages to predict prefetch entries in the lookahead manner is characterized as its `depth'. 
-While doing this, SPP also keeps compounding the confidence in each depth. 
-Thus as depth increases, overall confidence keeps decreasing. 
-This makes sense since we are going deeper into the speculation path. 
-When the confidence falls below a pre-defined threshold, SPP stops further lookahead iterations
-\newline
-\newline
-\textbf{Global History Register:} One of the innovative ideas introduced in SPP was the Global History Register, which bootstraps learning across page boundaries.
-GHR is a small 8-entry fully associative structure. 
-Whenever a prefetch suggestion from the Pattern Table crosses page boundaries, instead of completely discarding it, GHR saves that information for inter-page learning.
-Since the SPP-Perc implementation in this paper doesn't affect GHR functionality in any ways, we won't be going into more details about GHR.
-\newline
-\newline
-\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch Filter - a 1024 entry 1-way associative table which keeps a record of last few entries prefetched by SPP. 
-This proves useful for following reasons:
+\textbf{Pattern Table:} The PT is indexed by the signature generated
+from ST.  PT holds predicted delta patterns and their confidence
+estimates.  Each entry indexed by the signature holds up to 4 unique
+delta predictions.  This is implemented by making PT as a 4-way
+associative table.  To measure an estimate of a delta confidence, PT
+implements two counters: C\textsubscript{sig} for each entry indexed
+by the signature and C\textsubscript{delta} for each delta entry for a
+given signature.  More details of this on-line score-keeping mechanism
+is described under Confidence Tracking discussion.
+
+\textit{Lookahead Prefetching:} For each trigger to the prefetcher,
+the first time that the SPP indexes into the PT, it does so in the
+non-speculative manner.  We call it non-speculative because the
+parameters which SPP uses to predict, like last block offset, incoming
+block offset and signature are well-defined.  Assuming PT generates a
+valid delta prediction (rather, up to 4 valid deltas), SPP enters into
+the lookahead mode.
+
+In the lookahead mode, SPP looks at the last generated delta(s),
+chooses the one with the highest confidence and speculatively
+generates the new effective address and the new signature.  Using
+these new parameters, it re-accesses the PT to generate further
+prefetches.  SPP keeps on repeating this cycle of accessing PT and
+updating the signature based on highest confidence prefetch from the
+last iteration.  These iteration count till which SPP manages to
+predict prefetch entries in the lookahead manner is characterized as
+its `depth'.  While doing this, SPP also keeps compounding the
+confidence in each depth.  Thus as depth increases, overall confidence
+keeps decreasing.  This makes sense since we are going deeper into the
+speculation path.  When the confidence falls below a pre-defined
+threshold, SPP stops further lookahead iterations.
+
+\textbf{Global History Register:} One of the innovative ideas
+introduced in SPP was the Global History Register, which bootstraps
+learning across page boundaries.  GHR is a small 8-entry fully
+associative structure.  Whenever a prefetch suggestion from the
+Pattern Table crosses page boundaries, instead of completely
+discarding it, GHR saves that information for inter-page learning.
+Since the SPP-Perc implementation in this paper doesn't affect GHR
+functionality in any ways, we won't be going into more details about
+GHR.
+
+\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch
+Filter - a 1024 entry 1-way associative table which keeps a record of
+last few entries prefetched by SPP.  This proves useful for following
+reasons:
 \begin{itemize}
-    \item \textbf{Eliminating duplicate prefetches}\newline 
-    Any subsequent prefetches recommended by SPP mechanism is checked against the existing entries in Prefetch Filter. 
-If a valid entry is found, it means that the particular cache line is already either in L2 Cache or being fetched. 
-Hence there is no need to issue a duplicate prefetch.
+\item \textbf{Eliminating duplicate prefetches}\newline Any subsequent
+  prefetches recommended by SPP mechanism is checked against the
+  existing entries in Prefetch Filter.  If a valid entry is found, it
+  means that the particular cache line is already either in L2 Cache
+  or being fetched.  Hence there is no need to issue a duplicate
+  prefetch.
 
-    \item \textbf{Updating prefetcher}\newline
-    If a subsequent demand access comes before eviction for a prefetch that was recommended by SPP, various components of SPP are updated to reflect that this was a good prefetch. 
-Correspondingly, if a prefetched cache line is evicted without being demand hit even once, SPP learns that this was a useless prefetch and updates the various tables accordingly
+\item \textbf{Updating prefetcher}\newline If a subsequent demand
+  access comes before eviction for a prefetch that was recommended by
+  SPP, various components of SPP are updated to reflect that this was
+  a good prefetch.  Correspondingly, if a prefetched cache line is
+  evicted without being demand hit even once, SPP learns that this was
+  a useless prefetch and updates the various tables accordingly
 
-    \item \textbf{Tracking Global Accuracy}\newline
-    PF keeps a track of usefulness of the prefetches by introducing 2 counters: C\textsubscript{total} to track the total number of prefetches getting recorded in the PF and C\textsubscript{useful} to keep a track of which prefetches recorded in the PF led to a demand hit.
-The ratio C\textsubscript{useful} / C\textsubscript{total} is the global accuracy for the prefetching scheme. 
-It is represented as $\alpha$ in the paper. 
-The global accuracy acts a real-time throttler for the path confidence value as depicted in the following discussion.
-\newline
+\item \textbf{Tracking Global Accuracy}\newline PF keeps a track of
+  usefulness of the prefetches by introducing 2 counters:
+  C\textsubscript{total} to track the total number of prefetches
+  getting recorded in the PF and C\textsubscript{useful} to keep a
+  track of which prefetches recorded in the PF led to a demand hit.
+  The ratio C\textsubscript{useful} / C\textsubscript{total} is the
+  global accuracy for the prefetching scheme.  It is represented as
+  $\alpha$ in the paper.  The global accuracy acts a real-time
+  throttler for the path confidence value as depicted in the following
+  discussion.  \newline
 \end{itemize}
-\textbf{Confidence Tracking}: The PT keeps a track of hits to each signature through a counter C\textsubscript{sig}. 
-The number of hits for a given delta per signature are tracked using a counter C\textsubscript{delta}. 
-The confidence for a given delta is approximated through C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}. 
-When SPP enters into a lookahead mode, the path confidence P\textsubscript{d} is given as:
+\textbf{Confidence Tracking}: The PT keeps a track of hits to each
+signature through a counter C\textsubscript{sig}.  The number of hits
+for a given delta per signature are tracked using a counter
+C\textsubscript{delta}.  The confidence for a given delta is
+approximated through C\textsubscript{d} = C\textsubscript{delta} /
+C\textsubscript{sig}.  When SPP enters into a lookahead mode, the path
+confidence P\textsubscript{d} is given as:
 $$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown in the Prefetch Filter section. 
-The range of global accuracy is in [0,1]. 
-The lower the value of $\alpha$ for a given trace or a program phase, the overall confidence is scaled down and this makes SPP less aggressive for the upcoming prefetches.
+The range of global accuracy is in [0,1].  The lower the value of
+$\alpha$ for a given trace or a program phase, the overall confidence
+is scaled down and this makes SPP less aggressive for the upcoming
+prefetches.
 
-Here d is the lookahead depth. 
-For d = 1, when SPP is in non-speculative mode, P\textsubscript{0} can be thought of as 1. 
-Once SPP enters into speculative mode, it calculates the path confidence as a product of confidence of the individual prefetches encountered along that path, further scaled using $\alpha$. 
-The final P\textsubscript{d} is thresholded against prefetching-threshold (T\textsubscript{p}) to decide whether to prefetch or not. 
-If yes, then P\textsubscript{d} is thresholded against a numerically bigger fill-threshold (T\textsubscript{f}) to decide whether the prefetch should be sent to L2 Cache (high confidence prefetch) or Last Level Cache (low confidence prefetch)
-\newline<Insert a comprehensive figure about SPP structures>
-\newline<Insert another detailed picture about SPP data-path flow>
+Here d is the lookahead depth.  For d = 1, when SPP is in
+non-speculative mode, P\textsubscript{0} can be thought of as 1.  Once
+SPP enters into speculative mode, it calculates the path confidence as
+a product of confidence of the individual prefetches encountered along
+that path, further scaled using $\alpha$.  The final
+P\textsubscript{d} is thresholded against prefetching-threshold
+(T\textsubscript{p}) to decide whether to prefetch or not.  If yes,
+then P\textsubscript{d} is thresholded against a numerically bigger
+fill-threshold (T\textsubscript{f}) to decide whether the prefetch
+should be sent to L2 Cache (high confidence prefetch) or Last Level
+Cache (low confidence prefetch)
+
+<Insert a comprehensive figure about SPP structures>
+
+<Insert another detailed picture about SPP data-path flow>
 
 \subsection{Case for an On-line Throttler}
-As compared to some of the other state of the art prefetchers [BOP], SPP is a far less aggressive prefetcher. 
-In the single core environment, this gives BOP an edge as there is no resource contention among the different cores. 
-Hence more aggressive prefetching is bound to prove beneficial. 
-As we increase the core count, we observe that SPP starts outperforming rest of the prefetchers. 
-This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. 
+As compared to some of the other state of the art prefetchers [BOP],
+SPP is a far less aggressive prefetcher.  In the single core
+environment, this gives BOP an edge as there is no resource contention
+among the different cores.  Hence more aggressive prefetching is bound
+to prove beneficial.  As we increase the core count, we observe that
+SPP starts outperforming rest of the prefetchers.  This can be
+attributed to the fact that each prefetch suggested by SPP is a
+carefully calculated one and that prevents cache pollution.
 \textit{<Figure showing the variation of SPP vs BOP wrt core count>}
-For 4-core applications, SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
+For 4-core applications, SPP suggests XX\% fewer prefetches than BOP
+and yet leads to higher IPC.
 
-The above analysis shows that with a more careful throttling mechanism, SPP can be tuned to become much more aggressive, leading to increased coverage. 
-The onus of maintaining the accuracy now falls on the independent throttler. 
-To test the hypothesis, we tuned down SPP to the minimum possible threshold. 
-Only by doing this we managed to increase the prefetch suggestions made by SPP, by XX\%. 
-\textit{<Comparison of SPP-Unleased with SPP / BOP>}
-Obviously this came at a cost of increased DRAM traffic and cache pollution.
+The above analysis shows that with a more careful throttling
+mechanism, SPP can be tuned to become much more aggressive, leading to
+increased coverage.  The onus of maintaining the accuracy now falls on
+the independent throttler.  To test the hypothesis, we tuned down SPP
+to the minimum possible threshold.  Only by doing this we managed to
+increase the prefetch suggestions made by SPP, by XX\%.
+\textit{<Comparison of SPP-Unleased with SPP / BOP>} Obviously this
+came at a cost of increased DRAM traffic and cache pollution.
 
-Moreover, the on-line confidence mechanism introduced in SPP was very rudimentary.
-It was based on taking a ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig} as explained above.
-The same confidence was used to make the decision whether to prefetch or not to prefetch; and which level to prefetch.
-While this approximation was shown to work in the original SPP, we believe that a better form of on-line decision making was possible.
-Hence, it was necessary to build a robust and adaptable learning mechanism to accept / reject the prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level Cache).
+Moreover, the on-line confidence mechanism introduced in SPP was very
+rudimentary.  It was based on taking a ratio C\textsubscript{d} =
+C\textsubscript{delta} / C\textsubscript{sig} as explained above.  The
+same confidence was used to make the decision whether to prefetch or
+not to prefetch; and which level to prefetch.  While this
+approximation was shown to work in the original SPP, we believe that a
+better form of on-line decision making was possible.  Hence, it was
+necessary to build a robust and adaptable learning mechanism to accept
+/ reject the prefetch suggestions; and to decide the fill level (L2
+Cache vs Last Level Cache).
 
 To that effect, we introduce an independent on-line perceptron based throttling mechanism.
-
-\section{SPP-Perc}
-
-\subsection{Changes made to original SPP}
-To modify the SPP Design to suit our scheme, the following changes were made:
-
-\begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline
-In the original SPP design, an internal confidence mechanism was developed which was compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the prefetch suggestion.
-Furthermore, the accepted suggestions were thresholded against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache vs Last Level Cache.
-T\textsubscript{p} and T\textsubscript{f} were set to 25 and 90 respectively, on the scale of 0 to 100.
-In SPP-Perc, the perceptron sum is used to make these decisions. 
-Hence these two thresholds are no longer needed
-
-\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
-Original SPP would not consider the slots available in the L2 MSHR queue. 
-It had an assumption that at no time can the number of prefetches suggested by SPP could exceed the capacity of the queue. 
-While it was not proven in the paper, the internal confidence mechanism along with the thresholds made sure that the above assumption was maintained. 
-In SPP-Perc since there can be no assumption on the degree of aggressiveness - because SPP is being tuned to its lowest thresholds, we explicitly check that at no point should the number of suggested prefetches exceed L2 MSHR queue.
-
-\item \textbf{Enhanced Filter}\newline
-Original SPP introduces a concept of Prefetch Filter, which keeps a track of up to 1024 L2 Cache prefetches suggested by SPP. 
-The Filter was modified to store meta-data like program counter, memory access address etc. 
-This is needed to train the perceptron when the result of that particular prefetch gets available. 
-At training time, this data is used to retrieve the state of the prefetcher when the prediction was made.
-
-\item \textbf{Reject Filter introduced}\newline
-The Prefetch Filter mentioned above keeps a track of the prefetches that actually happened in the L2 Cache. 
-Correspondingly, we introduced a reject filter. 
-It keeps a track of the prefetches that were suggested by the base SPP engine but were rejected by the perceptron. 
-In case it is detected that a future demand fetch could have used this rejected prefetch, the perceptron weights are updated to reflect the same
-
-\end{itemize}
-
-\subsection{The Perceptron Layer}
-As discussed earlier, the perceptron layer was introduced to act as a throttler to the underlying SPP engine. 
-This section explores the details of the perceptron layer and the range of features that are ultimately used by the perceptron for correlating the prefetches.
-
-The neural network weights are stored in separate tables for each feature.
-The indexing of each feature is done using different number of bits.
-At most 12 bits of a feature are used to index into the perceptron table.
-Certain features require more resolution power \textit{i.e.}, full 12 bits of indexing.
-On the other hand, some features requiring lower resolution required as low as 7 bits of indexing.
-The variable indexing was determined by studying the features and fine-tuned empirically so as to achieve a good accuracy vs area trade-off.
-Exact details are mentioned under `Area Overheads' in Section XX.
-
-A single entry in the table corresponds to a perceptron weight.
-Each weight is a 5 bit counter - saturating at -16 and +15. 
-Each feature gets its own dedicated table. 
-Our proposed design uses 9 features. 
-As the program execution begins, all the weights are initialized to 0.
-\newline \newline
-\textbf{INFERENCING}\newline
-In the Champsim simulator, the prefetching method is triggered on every L2 Cache demand access. 
-At that point, the prefetcher has the option to do or not do prefetch - in case it does, then how many cache lines to prefetch. 
-The suggested prefetches can be places either in L2 Cache or in Last Level Cache. 
-The prefetcher gets to decide this, and is usually done based on an internal confidence mechanism.
-
-When SPP gets triggered, it suggests up to 4 (=Associativity of the Pattern Table) candidates for prefetch based on the demand access address. 
-In each subsequent lookahead depth, SPP suggests up to 4 more prefetch candidates, as explained in section\_\_. 
-All these recommended prefetches are tested in the perceptron throttler before they can access the Prefetch Filter, as in the original SPP data-path.
-
-The perceptron looks at the micro-architectural state, or `features', at that instant. 
-Each feature indexes into the hash of up to 4096-entry table dedicated for that feature. 
-Feature \#1 indexes into a particular entry in table \#1 and so on.
-
-Once all the weights are retrieved, they are added together, much like the traditional perceptron dot product. 
-On obtaining the overall sum, it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO. 
-Only the prefetch candidates with the perceptron sum higher that the threshold qualify for prefetching. 
-The prefetches that qualify the perceptron throttler stage go through the Prefetch Filter stage, just as in the SPP design.
-The Prefetch Filter eliminates any redundant prefetch suggestions.
-In SPP-Perc, all the meta-data that is required to recreate the features is also stored in the Prefetch Filter. 
-At a later stage, when the feedback of the current prefetched line is available, the stored data can be used to train the perceptron.
-
-The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch cache line. 
-All prefetch candidates qualified till this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill level.
-The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
-
-The perceptron sum can also be seen as the sum of the individual contribution per feature.
-The value of each contribution corresponds to the extent of confidence of the final output with that given feature.
-By summing the individual contributions, the final perceptron sum denotes the overall confidence for that prefetch suggestion. 
-In a way, by thresholding the perceptron sum on two different thresholds, we are diving the confidence scale of the prefetch suggestion into three bins.
-The first bin corresponding to the lowest confidence leads to prefetch candidate being rejected.
-The next bin corresponds to prefetches with a moderate confidence level. 
-Such prefetches are directed to fill the bigger Last Level Cache and potentially not pollute the more scarce L2 Cache.
-The highest confidence prefetches are filled in the L2 Cache.
-
-In addition to the Prefetch Filter mentioned above, SPP-Perc also maintains a `Reject Filter'. 
-Reject filter is a 1024-entry deep 1-way associative table. 
-In case that a prefetch suggestion is rejected by perceptron throttler, it instead gets logged into the Reject Filter. 
-The Reject Filter keeps track of 1024 entries that were suggested by the Pattern Table but got rejected by the perceptron throttler. 
-This can be used to train the perceptron to avoid false negatives \textit{i.e.}, cases where prediction was to reject the prefetch but the prefetch would have been useful.
-
-As previously stated, an enhancement to the original SPP design includes storing all the meta-data describing the program state at the instant of prefetching, in the Prefetch Filter. 
-This information is useful when the perceptron needs to be updated subsequently. 
-\newline \newline
-\textbf{TRAINING}\newline
-In the prefetching environment, feedback of a prefetch is received whenever there is an eviction or a demand access from L2 Cache. 
-This acts as a trigger to update all the baseline SPP entries and to also train the perceptron throttler. 
-Training of the perceptron throttler involves restoring the meta-data that was stored in Prefetch Filter or the Reject Filter. 
-This is done by indexing into the corresponding filters using the cache line address of the block being used to train. (10-bits indexing + 6-bits tag matching). 
-Once the state of the program at the time of prefetching is available, it is used to index into the perceptron weights table.
-
-If a demand access block which triggers the training  was tagged as a valid prefetch in the Prefetch Filter, it means that the earlier prefetch prediction was correct. 
-In that case the perceptron weights are incremented by 1 if the predicted sum does not cross a pre-defined threshold. 
- In case it is a cache block eviction that led to training and the corresponding valid entry was found in the Prefetch Filter, it shows that the prediction made by the perceptron was wrong. 
-Perceptron should have ideally rejected the prefetch suggestion as a low-confidence prefetch. 
-Here the weights are decremented by 1 to reflect the same. 
-In either case, weights are saturated at -16 or +15.
-
-A secondary training mechanism also kicks in during demand fetches. 
-Before the demand access triggers the next set of prefetches, it checks the Reject Filter for the valid entry. 
-In case it is a hit, that means that the corresponding cache line was initially suggested by the underlying SPP engine but rejected by the perceptron throttler. 
-This means that perceptron should have been more confident about that particular prefetch. 
-Once such a scenario is identified, the state of the execution at the time of prefetch is retrieved from the Reject Table. 
-The retrieved data is used to index into the various weights tables of the perceptron and the corresponding values are updated by +1, saturating between and -16 and +15. 
-This update reflects increased confidence for the prediction corresponding to that prefetch.
-
-This mechanism allows us to solve the classical problem of exploiting `lost opportunity'. 
-In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
-Our two-step SPP-Perc allows us to overcome that issue in a simple manner.
-
-\subsection{Features used by Perceptron}
-
-Here we discuss the various features which help correlate the prefetching decision with the program behaviour
-\begin{itemize}
-\item \textbf{Base address} of the demand access which triggered the prefetch. 
-Since prefetch are triggered from the L2 demand access, they tend to be correlated with the triggering address.
-
-\item \textbf{Cache line} and \textbf{Page address}: These two separate features are derived from the base address which triggered the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and base\_addr >> LOG2\_PAGE\_ SIZE respectively.
-The idea behind using three different shifted versions of the same feature is that it allows us to look into a wider range of bits than with a single version.
-It also helps give more importance to the overlapping bits and lesser importance to most and least significance bits.
-This approach can also eliminate destructive interference that can be caused by directly folding the bits into half.
-
-\item \textbf{Confidence}: The integer confidence on the scale of 0 to 100 which was used in the original SPP design.
-While the original confidence is not used directly for decision making, it can still contribute to the final decision made.
-
-\item \textbf{Program Counter XOR Depth}: PC is for the instruction which triggered the prefetch chain.
-Depth refers to the iteration count of the lookahead stages.
-In general, PC is not considered as a good basis for doing lookahead prefetching as all the prefetches with depth >= 1 gets aliased onto the same PC which won't be true in actual demand fetch case.
-This feature resolves a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases.
-This is akin to the concept of Virtual Program Counters\cite{VPC} introduced by Kim \textit{et. al.} for indirect branch prediction.
-
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC before the instruction that triggered the current prefetch. 
-Hashing together the last three PCs tell SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block.
-PCs are shifted in the increasing order of history before being hashed together.
-This is done to avoid the resultant value of zero in case 2 or more PCs are the same. 
-Additionally, obfuscating the information as it gets old allows us to get a wider and an approximate look into the program history.
-
-\item \textbf{Program Counter XOR Delta}: This tells us if a given PC favours particular value(s) of delta.
-As noted earlier, while PC in itself does not much useful information, this hash resolves PC into different values based on the tendency of that PC to favour a certain delta.
-This way, dynamic nature of different instances of the same memory access instruction can be captured here.
-
-\item \textbf{Current Signature XOR Delta}: It can be recalled from the discussion of SPP in section XX that the new signature is generated using the old signature and the blobk delta.
-While `Current Signature XOR Delta' is not the exact formula for generating the future signature, it gives an approximate idea of the path that the combination of these two values can lead to.
-
-%%%%%%% Below are the features rejected when downsizing 14 -> 9
-%\item \textbf{Current Signature}: The 12-bit signature used by SPP to index into the Pattern Table
-%\item \textbf{Delta}: The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched
-%\item \textbf{PC}: Program Counter of the Load instruction that triggers that particular prefetch
-%\item \textbf{Base address XOR Delta}: If any particular value(s) of delta are highly favoured by a given address which triggered the prefetch, this composite feature can capture that information
-%\item \textbf{Cache line XOR Delta}: This feature captures if any particular delta is highly favoured by a given cache line address
-%\item \textbf{Confidence XOR Delta}: This gives a combined insight into that fact that certain delta-confidence pairs predicted by SPP might correlate more with the prefetch outcome
-%%%%%%%%%%%%%%%%%%%%%%
-
-\end{itemize}
-
-Some of the composite features are derive from simple hashing (XOR) of two primary features.
-There is always a question of usefulness of such composite features and the `fresh' information added.
-We try to justify the choice of each feature by quantifying the contribution made towards predicting prefetch behaviour, in Section XX.
-Finally, as noted above, each feature indexes into its independent entry of perceptron weights.
-
-\section{METHODOLOGY}
-\subsection{Performance Model}
-For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim} simulator. 
-ChampSim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} 
-We model model a 1-core / 4-core / 8-core Out of Order machine. The details of the configuration parameters is summarized in Table  \ref{tab:Sim_params}.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|l|p{3.6cm}|c|}
-    \hline
-    Level & Configuration & Access Latency \\
-    \hline
-         L1 Cache & Separate I-cache and D-cache, 32 KB, 8-way & 4 cycles\\
-         L2 Cache & Private, 256 KB & 8 cycles\\
-         LLC & Shared, 2MB / core, 16-way & 20 cycles\\
-         DRAM & 4 GB Single Channel for single-core, 8 GB Double Channel for multi-core & N?? cycles\\
-    \hline
-    \end{tabular}
-    \caption{Memory Model Parameters}
-    \label{tab:Sim_params}
-\end{table}
-
-The block size was fixed 64KB. Prefetching was only triggered at L2 Cache demand accesses but could be sent to L2 Cache or Last Level Cache. 
-No L1 Data level prefetching was done. 
-LRU replacement policy was used on all levels of cache hierarchies. 
-Branch prediction was done using the perceptron branch predictor\cite{Perc_Branch}. 
-Page size was kept as 4KB. 
-Champsim operates all the prefetchers strictly in the physical address space.
-
-\subsection{Testing Under Additional Memory Constraints}
-The default single-core configuration simulates a 2MB LLC and a single channel DRAM with 12.8GB/s bandwidth. 
-We extend the simulations to include memory constraints introduced in DPC-2. 
-Specifically we look at the following two variations:
-\begin{itemize}
-    \item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited to 3.2 GB/s
-    \item \textit{Small LLC}: In this scenario, LLC size is reduced to 512 KB
-\end{itemize}
-All the multi-core simulations are only done in the default configuration.
-
-\subsection{Workloads}
-This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017} has been used to characterize and measure the prefetch performance. 
-We use all the 20 workloads available in the SPEC 2017 Suite. 
-Using SimPoint\cite{SimPoint} methodology, we identified 95 different program segments of 1 Billion instructions each.
-
-We performed cross-validation of our SPP-Perc model using SPEC 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.
-For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
-For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.
-The traces include 4 multi-core applications with 6 distinct phases per application. 
-In total, we used 285 traces representing workloads across 53 applications. 
-\newline
-\newline
-\textit{Single-core performance:} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics.
-We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
-We take the Instruction Per Cycle (IPC) of the prefetcher and divide it with the IPC of baseline. 
-The final number reported is the geometric mean of the speed-up attained on individual traces.
-\newline
-\newline
-\textit{Multi-core performance:} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
-For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 
-Each CPU keeps executing its workload till the last CPU completes 1 Billion instructions after warm-up. 
-For collecting IPC and other data, only the first billion instructions are considered as the ``region of interest''.
-
-Here we report the weighted speedup normalized to baseline \textit{i.e.}, no prefetching. 
-For each of the workloads running on a particular core of the 4-core 8 MB LLC system, we compute IPC\textsubscript{i}. 
-We then find the IPC\_isolated\textsubscript{i} of the same workload running in isolated 1-core 8 MB LLC environment. 
-Then we calculate the total weighted-IPC for a given workload mix as $\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}). 
-For each of the 100 workload-mix, the sum obtained is normalized to the weighted-IPC calculated similarly for baseline case \textit{i.e.}, no prefetching, to get the weighted-IPC-speedup. 
-Finally the geometric mean of these 100 weighted-IPC-speedup is reported as the effective speedup obtained by the prefetching scheme.
-
-We repeat the same process for 8-core workloads with the exception that 20 Million warm-up instructions and 100 Million full instructions were executed.
-This was done so as to keep the simulation run-time within reasonable limits as a single 8-core mix was taking up to 5 days to simulate 1 Billion instructions.
-
-\subsection{Preferchers Simulated}
-We compared SPP-Perc against three of the latest state of the art prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher (SPP).
-BOP was the winner of 2nd Data Prefetching Championship. 
-DA-AMPM is the enhanced version of AMPM, modified to account for DRAM row buffer locality. 
-SPP has been shown to outperform BOP on SPEC 2006 traces. 
-For each of these, we compare their speedups taking no prefetching as the baseline.
-
-\subsection{Developing Features for SPP-Perc}
-This section describes the intuition and analysis that went behind finalizing the features. 
-As noted earlier, we developed a set of 9 features which allow the perceptron throttler to correlate prefetching decision with the program context. 
-To study the correlation across each feature, we study statistically the perceptron weights and try to interpret their distribution.
-\newline  \newline
-\textbf{GLOBAL PEARSON'S CORRELATION}\newline
-For this experiment, we create a dump of the state of perceptron weights at the end of all trace execution.
-The weights obtained from running all the SPEC 2017 traces are concatenated together. 
-Since the weights are collected at the end of individual trace execution, the perceptron weights have attained a relatively stable value by now.
-Hence, this dump represents a `snap' of the trained perceptron weights across all the SPEC 2017 traces.
-The intuition here is that the feature with bulk of the perceptron weights concentrated around 0 or small magnitude numbers show a weak correlation with the prefetching outcome. 
-On the other hand, features with most of the weights saturated around highest value (+15) show a high positive correlation and the features with weights close to the lowest value (-16) show a strong negative correlation.
-
-We plot a histogram for each feature depicting weights distribution from -16 to +15 and use this histogram to generate the Pearson's correlation factor for that feature.
-In statistics, Pearson's factor is a numerical measure of the degree of linear correlation between two variables.
-It ranges from -1 to +1.
-The magnitude of Pearson's factor depicts the extent of correlation and the sign depicts whether it is a positive correlation or a negative correlation.
-Values close to 0 suggest a low correlation or at times, noisy data.
-A value of +1/-1 suggests a perfectly linear positive / negative correlation respectively.
-Figure XX depicts the histogram and the Pearson's coefficient for two of the features, YY and ZZ.
-
-Figure XX shows all the features used, arranged in the increasing order of their Pearson's factor.
-As can be seen X out of the 9 features provide a moderate to high correlation, with the magnitude of P-value > 0.4.
-The single most important feature, \_\_ helps provide a correlation to prefetch outcome with a factor of \_\_.
-\newline \newline
-\textbf{PER TRACE CORRELATION} \newline
-Another important way to look at the perceptron features is to see how much their contribution varies across the traces.
-Here we give special attention to features with low P-values in the previous experiment.
-Figure XX shows the variation of P-values three features : \_\_, \_\_ and \_\_; across all the SPEC 2017 traces.
-For simplicity, the traces are arranged in an increasing order of contribution made by the feature and the trace names have been omitted.
-It can be seen that even features with a low overall correlation provide useful correlation (magnitude > 0.4) for XX out of the 95 simpoints developed for SPEC 2017.
-\newline \newline
-\textbf{TRIMMING FEATURES USING CORRELATION} \newline
-Besides providing interesting insights into prefetching behaviour, P-value can also be used for feature selection and prefetcher tuning.
-
-Here, we introduce the concept of cross-correlation across features.
-Just as we examined correlation of each feature with the final outcome, we can also study correlation between the features.
-We used the above methodology to eliminate features which provide a little or no information that has already been captured in other features.
-
-As a part of feature selection, we initially came up a mix of 23 features- primary and composite.
-By studying cross correlation of each of these features against others in a 23x23 matrix, we identified pairs of features with correlation factor > 0.9 in magnitude and carefully eliminated redundant features.
-Using this approach, we managed to reduce the feature count to 9.
-Thus, in the final implementation of SPP-Perc, no two features have a high correlation between them.
-This way we can be sure that each feature is bringing in contribution that cannot be captured using other features.
-
-\textit{<TODO: VALIDATE THIS>} 
-Secondly, studying the relative importance of each feature enabled us to vary the number of entries dedicated for each feature.
-Features with higher Pearson's correlation (like \_\_ and \_\_) were given most importance and allowed full 12-bits of indexing.
-Features like \_\_ and \_\_, with a low overall P-value were allocated fewer indexes in the feature table.
-
-To conclude, in the above discussion we justify the features for perceptron from a statistical viewpoint.
-We also show how this information can also be used for prefetcher tuning.
-All this study was made possible only because we used on-line perceptron learning for prefetching and that enabled us to examine the weights in detail.
-
-\subsection{Overhead for SPP-Perc}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|c|c|m{4.8cm}|}
-    \hline
-        \textbf{Field} &
-        \textbf{Bits} &
-        \textbf{Comment} \\
-    \hline
-         Valid & 1 & Indicates a valid entry in the filter\\
-         Tag & 6 & Identifier for the entry in the filter\\
-         Useful & 1 & To show if the given entry led to a useful demand fetch\\
-         Perc Decision & 1 & Prefetched vs Not-prefetched \\
-    \hline
-        PC & 12 & \multirow{5}{4.8cm}{Meta-data required for perceptron training}\\
-        Address & 24 & \\
-        Delta & 7 & \\
-        Curr Signature & 12 & \\
-        Confidence & 7 & \\
-    \hline
-        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
-    \hline
-    \end{tabular}
-    \caption{Meta-data Stored in Prefetch Filter}
-    \label{tab:PF_metadata}
-\end{table}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|p{1.5cm}|p{1cm}|c|p{1.5cm}|}
-    \hline
-        \textbf{Structure} &
-        \textbf{Entry} &
-        \textbf{Components} &
-        \textbf{Total} \\
-    \hline
-        \multirow{5}{1.5cm}{Signature Table} & \multirow{5}{1.5cm}{256} & Valid (1 bit) & \multirow{5}{1cm}{11008  bits}\\
-        & & Tag (16 bits) &\\
-        & & Last Offset (6 bits) &\\  
-        & & Signature (12 bits) &\\
-        & & LRU (6 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Pattern Table} & \multirow{3}{1.5cm}{512} & $C_{sig}$ (4bits) &\multirow{3}{1cm}{24576 bits}\\
-        & & $C_{delta}$ (4*4 bits) &\\
-        & & Delta (4*7 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Perceptron Tables} & \multirow{3}{1.5cm}{4096*8 128*1} & \multirow{3}{1.5cm}{5 bits} & \multirow{3}{1.5cm}{164480 bits}\\
-        & & &\\
-        & & &\\
-    \hline
-        \multirow{2}{1.5cm}{Prefetch Filter \footnotemark[1]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{84 bits} & \multirow{2}{1.5cm}{86016 bits}\\
-        & & &\\
-    \hline
-     \multirow{2}{1.5cm}{Reject Filter \footnotemark[2]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{83 bits} & \multirow{2}{1.5cm}{84992 bits}\\
-        & & &\\
-    \hline
-        \multirow{4}{1.5cm}{Global History Register} & \multirow{4}{1.5cm}{8} & Signature (12 bits) & \multirow{4}{1.5cm}{264 bits}\\
-        & & Confidence (8 bits) & \\
-        & & Last Offset (6 bits) & \\
-        & & Delta (7 bits) & \\
-    \hline
-        \multirow{2}{1.5cm}{Accuracy Counters} & 1 & C$_{total}$ & 10 bits\\
-        & 1 & C$_{useful}$ & 10 bits\\
-    \hline
-        \multirow{3}{1.5cm}{Global PC Trackers} & \multirow{3}{1cm}{3} & $PC_1$ (12 bits)& \multirow{3}{1.cm}{36 bits}\\
-        & & $PC_2$ (12 bits) &\\
-        & & $PC_3$ (12 bits) &\\
-    \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 371356 bits = 45.33 KB}}\\
-    \hline
-    \end{tabular}
-    \caption{SPP-Perc Storage Overhead}
-    \label{tab:SPPPerc_overhead}
-\end{table}
-\footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}
-\footnotetext[2]{RF does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
-
-In this section, we analyze the hardware overhead required to implement SPP-Perc. 
-The Prefetch Filter was enhanced to accommodate storing of meta-data for perceptron training. 
-Table \ref{tab:PF_metadata} depicts the meta-data stored for each entry in the Prefetch Filter.
-Table \ref{tab:SPPPerc_overhead} shows the total storage overhead of SPP-Perc implementation.
-The hardware budget for 2nd Data Prefetching championship was 32 KB. 
-Keeping that in mind and the speedup SPP-Perc obtained over the winner, the extra hardware budget can be accounted for.
-The extra hardware also makes the overall scheme more scalable than SPP. 
-In the original SPP paper, it was demonstrated that adding extra hardware brings little advantage in terms of performance gain.
-The newly added perceptron tables can be scaled to increase / decrease features depending on the permitted budget.
-
-In terms of computations, the perceptron mechanism only introduces an extra adder tree.
-The hash perceptron mechanism makes sure than there is no actual vector multiplication happening in the hardware.
-Obtaining the perceptron sum requires addition of 9 5-bit numbers. 
-Using an adder tree of 4 5-bit adders, this can be done in ceil($log_{2}9$) = 4 steps.
-Perceptron update only requires weight update by +1 or -1.
-Thus, all the operations required for perceptron inferencing or updating the states of the perceptron throttler can be easily done in the time constraints of L2 Cache Accesses.
-
-\section{RESULTS}
-This section discusses the results obtained from running SPP-Perc, in terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per Cycle (IPC) metrics. 
-For SPEC 2017 suite, first we present the results for single-threaded workloads then the results for multi-core workloads are presented.
-
-\subsection{Single-core Results}
-Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated. 
-All the results have been normalized to baseline \textit{i.e.}, no prefetching. 
-As can be seen, SPP-Perc yields a speedup of XX\% over the baseline. 
-This is equivalent to XX\% over DA-AMPM, xx\% over BOP and XX\% over SPP. 
-Out of the XX benchmarks, SPP-Perc nearly matches or outperforms most of the prefetchers on XX traces.
-
-At its peak, SPP-Perc manages to get the IPC speedup of a factor of over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}. 
-This also corresponds to speedup gain of \textbf{22.4\%} over the next best prefetcher - SPP. 
-In general, benchmarks \textit{602.gcc\_s, 603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s} benefit the most from SPP-Perc, as compared to SPP only prefetching.
-
-<Discuss the extreme behaviour of \textit{623.xalancbmk\_s}: Improvement over SPP is as high as 31\% and even reduces to -1.5\% on traces of this benchmark>
-
-<Discuss about coverage on all traces>
-
-<Discuss cases where coverage gets specially increased>
-
-<Discuss accuracy on all traces>
-
-<Discuss about compelling increase in accuracy - details of traces. Plot linear separability>
-
-<Any case where accuracy got reduced - go into the trace. See if linear separability is the issue. Other possible issues?>
-
-\subsection{Multi-core Results}
-In this section, we demonstrate the improvement achieved by SPP-Perc for a mix of multi-programmed workloads. 
-\newline \newline
-\textit{4-core environment}
-Figure \_\_ shows a comparison of speedups obtained on workload mix of memory intensive subset of SPEC 2017. 
-We plot all the 4 prefetchers, normalized to baseline. 
-The workloads have been sorted in the increasing order of performance benefit, as measured in terms of normalized weighted speedup. 
-As can be seen, SPP-Perc offers an IPC improvement by 33.5\% on these traces. 
-This is 5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP and DA-AMPM respectively.
-
-This increased improvement by SPP-Perc on a 4-core machine is not surprising. 
-Since SPP-Perc has an extremely intelligent throttler working to eliminate useless prefetches before they can create cache pollution. 
-This further amplifies the edge that SPP had over BOP. 
-With the advantage of an orthogonal prefetcher, SPP can be as aggressive as possible and leave the accuracy aspect to perceptron throttler.
-\newline \newline
-\textit{8-core environment}
-To test the effectiveness of SPP-Perc in a tighter constrained multi-core environment, we simulated an 8-core machine with 16 MB last Level Cache. 
-Other parameters were kept same as discussed under Methodology in Section XX.
-<See if we can use speed-up over BOP as a measure here>
-
-\subsection{Additional Memory Constraints}
-Figure \_\_ shows the performance of SPP-Perc against the other prefetchers in tighter constraint environments. 
-For sake of simplicity, we show 2 traces where SPP-Perc gives the most advantage over baseline, 2 traces where SPP-Perc doesn't perform as well as the best prefetcher *<-replace with the prefetcher name once results are available* and finally GeoMean across all the traces.
-<Hopefully BOP would degrade more here as it is super aggressive. 
-By definition, SPP-Perc should be better than SPP as the throttler would better control the aggressiveness. 
-In SPP paper DA-AMPM performs close enough to BOP. That should get replicated here too.>
-
-\subsection{Cross Validation}
-Here we try to demonstrate the robustness of SPP-Perc by testing it on completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
-
-Fig \_\_ shows the speed-up achieved by BOP, SPP and SPP-Perc on selected traces of SPEC 2006, along with the overall geometric mean over whole of the benchmark and over memory-intensive subset of the SPEC 2006 benchmark.
-As can be seen, SPP-Perc peovides a speed-up of 44.4\% over baseline, on the memory intensive subset of SPEC 2006 benchmark.
-This corresponds to 6.84\% over SPP and XX\% over BOP.
-On the whole of the SPEC 2006 suite, the speedup obtained is 22.4\%, which is 3.34\% ahead of SPP.
-
-The 4-core results for SPEC 2006 are \_\_.
-
-As can be seen, the performance is very consistent given no prior parameter tuning was done keeping SPEC 2006 in mind. 
-We attribute this to the inherent adaptability of the perceptron model.
-In general, perceptron weights are able to adjust in real-time so as to find the best possible correlation between the output and the features.
-
-\section{CONCLUSION}
-In this paper, we introduced perceptron based learning for data prefetching in the form of SPP-Perc. 
-Perceptron acts an orthogonal throttler to the underlying prefetch engine. 
-SPP-Perc improves performance by over XX\% over the baseline, which corresponds to YY\% over the next best performing prefetcher. 
-We believe this paper can lead to interesting future work like generalizing SPP-Perc to work with any prefetcher with minimal knowledge of the underlying prefetcher architecture.
-
-%%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%
-
-%%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
-\bibliographystyle{ieeetr}
-\bibliography{ref}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}

### 2018-11-15 conc.tex
reorg

--- 2018-11-15 paper_1.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is the template for submission to ISCA 2018
-% The cls file is a modified from  'sig-alternate.cls'
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{sig-alternate} 
-\usepackage{mathptmx} % This is Times font
-
-\newcommand{\ignore}[1]{}
-\usepackage{fancyhdr}
-\usepackage[normalem]{ulem}
-\usepackage[hyphens]{url}
-\usepackage{microtype}
-\usepackage{fixltx2e}
-\usepackage{array}
-\usepackage{multirow}
-
-% Always include hyperref last
-\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=black,citecolor=blue,urlcolor=black]{hyperref}
-
-% Ensure letter paper
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\newcommand{\iscasubmissionnumber}{NaN}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\fancypagestyle{firstpage}{
-  \fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-  \fancyhead[C]{\normalsize{}} 
-  \fancyfoot[C]{\thepage}
-}
-
-\pagenumbering{arabic}
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\title{SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher} 
-\author{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\maketitle
-\thispagestyle{firstpage}
-\pagestyle{plain}
-
-\begin{abstract}
-\textbf{Hardware prefetching has been introduced in modern processors as a way to hide cache latencies. 
-An efficient prefetcher should be able to identify complex memory access patterns during a program execution. 
-This enables it to fetch a block ahead of its demand access, potentially saving a cache miss.
-}
-
-\textbf{In this paper, we introduce perceptron learning to help make this prefetching decision efficiently. 
-We have taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine. 
-The perceptron layer acts as a throttler to filter out the unnecessary prefetches recommended by SPP. 
-We have also explored a range of features that can be used to train the perceptron layer. 
-Our results show that SPP-Perc improves performance on the memory intensive subset of the SPEC 2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared to just SPP.
-We also demonstrate that the performance gained from using our efficient throttler scales better with increasing number of cores.
-}
-\end{abstract}
-
-\section{Introduction}
-Since past few decades, processor and memory scaling has been enabled with different philosophies in mind. 
-While processor scaling focused on speed improvements, memory scaling was centred around increasing the storage capacity. 
-The process technology reaching scaling limitations led to Memory Wall\cite{MemWall} - the increasing performance gap between the processor speed and the memory speeds. 
-Various techniques have been developed to combat this widening gap. 
-Data prefetching is one such important technique. 
-Data prefetching exploits the fact that in most applications, memory accesses have a well-defined pattern.
-
-An ideal prefetching scheme involves capturing this memory access patterns to predict future prefetches in a timely manner. 
-Other important decisions faced by the prefetcher are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block.
-Memory accesses patterns can be as simple as fetching the next cache line or a complex but predictable pattern, like in pointer chasing applications. 
-Predicting this often requires a trade-off between coverage and accuracy. 
-For a prefetcher to capture highly irregular access patterns, it needs to be highly aggressive (high coverage). 
-That also means the prefetcher would end up recommending a lot of useless prefetches, leading to cache pollution (low accuracy). 
-This effect would get even more pronounced in multi-core scenario where the last level cache is a shared resource. 
-Conversely, a conservative prefetcher would be highly accurate but might not make enough useful predictions.
-
-To this effect we propose SPP-Perc <come up with a catchy name>. 
-It is an enhancement to the existing state of the art prefetcher and overcomes the coverage vs accuracy trade-off in a robust manner. 
-The idea is to modify SPP design to make it as aggressive as it can get. 
-This enables SPP to capture complex memory access patterns hence leading to increased coverage.
-This comes at a cost of increased DRAM traffic, hence reduction in overall IPC.
-The prefetch suggestions recommended by SPP pass through the perceptron based throttler. 
-Over time, perceptron layer learns to correlate the prefetches recommended by SPP with various available features, leading to elimination of useless prefetches. 
-This leads to an overall increased accuracy of the prefetch scheme and reduction in DRAM traffic.
-
-\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and future scope for expansion. 
-The contributions made by this paper are:
-
-\begin{itemize}
-\item This is the first time where a truly on-line learning based neural network model has been used when it comes to hardware data prefetching. 
-The prior works in this area either relied on program semantics\cite{Semantics} or were application specific\cite{Datacenter} or could not yield a significant improvement over the baseline prefetcher used\cite{BadPerc}.
-
-\item Implementing the pereptron throttler in a robust and accurate practical prefetching mechanism. 
-The overall scheme yields increased accuracy as compared to the other state of the art prefetchers.
-Moreover, the perceptron throttler learns to adapt itself to shared resource constraints - leading to an even increased performance in multi-core and lower capacity / bandwidth environments.
-
-\item Examining the perceptron weights doing a statistical study to quantify the contribution made by each feature.
-We have also used this idea to develop a comprehensive set of features where each feature contributes some `new' information to the perceptron.
-The details are explained under methodology, in Section XX.
-
-\item Cross-validating the design across benchmarks. 
-All the development and parameter tuning was done on the SPEC 2017 benchmark. 
-The results have been shown to be consistent across unseen benchmarks.
-More details can be found in the results section (Section XX).
-
-%\item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind. 
-%We have used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers. 
-%We have also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
-
-%\item We came up with an efficient model to use the perceptron output as a representation of the prefetch confidence using an on-line histogram technique. 
-%We show how the simple sum can replace the existing mechanism which was essentially based on counting the number of hits for a particular entry in SPP.
-
-\end{itemize}
-
-We tested SPP-Perc on Champsim, an open source simulator which models an Out-of-Order CPU using. 
-For evaluation, we used SPEC 2017 benchmarks. 
-On a single core configuration, SPP-Perc increases performance of the system by 39.5\% as compared to the baseline with no prefetching, on the memory intensive subset of the benchmark.
-This outperforms the next best prefetcher, SPP,  by 6.84\%. 
-On a multi-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement of 11.9\% over SPP for 4-core system and 10.67\% for 8-core system.
-
-In the results part in section XX, we show a detailed analysis of the speedup achieved.
-We also used SPEC 2006 and CloudSuite benchmarks for cross-validation of the prefetcher implementation.
-
-The rest of the paper is organized as follows.
-Section 2 talks about background work and recaptures the SPP architecture
-Section 3 discusses the detailed implementation of the perceptron mechanism and the features used.
-Section 4 consists of the evaluation methodology and explores the feature space for perceptron learning.
-Section 5 is the performance evaluation of our prefetcher and Section 6 concludes the paper.
-
-\section{Background Work and Motivation}
-
-In this section, we compare the existing work done in the domain of prefetching. 
-We also discuss about some of the interesting memory access patterns seen in the workloads in SPEC 2017 benchmark. 
-The second half of this section deals with a background of Signature Path Prefetcher architecture.
-
-\subsection{Related Work}
-
-The idea of prefetching has been around since Instruction Stream Buffers were proposed by Jouppi.
-\cite{ISB}
-Some of the earliest data prefetchers developed to try to identify memory access patterns with a constant stride pattern.
-\cite{Smith} 
-\cite{Baer}
-\cite{Stride}
-Major research was dedicated to stride distance and depth prediction.
-\cite{Decoupled}
-\cite{Adaptive}
-Newer prefetchers aimed at correlating predictions with the past memory access addresses.
-\cite{Address_Correlated} 
-\cite{AMPM}
-
-With time newer and more intricate prefetchers were introduced.
-\cite{Wenisch_Temporal_Streaming}
-\cite{Stealth}
-\cite{Feedback_Directed}
-\cite{Coordinated}
-\cite{Bandwidth_Efficient}
-\cite{Pacman}
-\cite{TLB}
-\cite{Linearizing}
-\cite{Sandbox}
-\cite{VLDP}
-\cite{DoL}
-\cite{Domino}
-A whole class of streaming prefetchers developed on the lines of Spatial / Temporal Memory Streaming.
-\cite{Spatial_Pattern}
-\cite{SMS}
-\cite{Temporal_Instruction_Fetch}
-\cite{Off_Chip}
-\cite{STMS}
-\cite{SMS_JILP}
-
-Most modern prefetchers aim at identifying complex memory access patterns in a given application. 
-This allows them to capture some of the irregular patterns seen in pointer chasing data-structures.
-Offset based prefetchers are a generalization of the next-line prefetcher.
-Some of the prominent prefetchers exploiting this idea are
-Sandbox Prefetcher \cite{Sandbox}
-and the Best Offset Prefetcher \cite{BOP}
-Another class of prefetchers is lookahead Prefetchers.
-Such prefetchers try to speculate deep into the application's memory access trail.
-These include Path Confidence based Lookahead Prefetching \cite{SPP}
-and Kill the Program Counter \cite{KPC}.
-
-A related line of work also explores cache replacement - insertion policies and dead block prediction, with an eye towards prefetching. 
-\cite{DB_Pred}
-\cite{Cache_Burst}
-\cite{KPC}
-\cite{Harmony}
-Nori \textit{et.al.} introduced prefetching with  respect to runtime criticality.
-\cite{CATCH}
-\newline
-\newline
-\textit{Machine Learning and Prefetching:}
-Peled \textit{et.al.} introduce interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
-\cite{Semantics}
-The biggest challenge here is that it relies on compiler support for getting features to build the context.
-
-The paper by Liao \textit{et. al.} focuses on prefetching for data center applications.
-\cite{Datacenter}
-They use offline Machine Learning Algorithms like SVMs and Logistic Regressions to do parametric search for an optimal prefetcher configuration.
-
-%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?
-Finally, the paper on Data Cache Prefetching with Perceptron Learning\cite{BadPerc} talks about the idea of two step prefetching.
-The first step is an existing baseline prefetcher while the second step is a perceptron based throttler.
-The paper has lots of limitations in terms of design and implementation.
-That reflects in the fact that it did not lead to any significant performance gain over even the basic prefetchers like the Stride Prefetcher\cite{Stride} and the Markov Prefetcher\cite{Markov}.
-
-\subsection{Perceptron Learning in Architecture}
-
-Perceptron learning for computer architecture design has been around for a while. 
-It was first popularized by Jimenez \textit{et. al.} for doing branch prediction. \cite{Perc_Branch} 
-In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
-The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. 
-The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the prediction is made as \textit{true}, or branch taken.
-Else if the sum is < 0, the prediction is made as \textit{false}.
-
-The structure that we use is derived from the model introduced in Piecewise Linear Branch Prediction\cite{Piece_Linear}. 
-Here the feature itself is used to index into a hash of perceptron weights. 
-This form of indexing makes sure that the hyper-plane learned by the perceptron weights is able to differentiate between linearly inseparable outcomes. 
-\textit{<Is it true? Not sure>}
-
-Perceptron Learning for Reuse Prediction\cite{Perc_Reuse} extended the hash perceptron architecture to perform prediction in context of cache replacement. 
-Perceptron layer can learn to correlate the cache replacement behaviour with a variety of features. 
-Each feature has its dedicated table of perceptron weights which it indexes into. 
-Hence number of tables equals the number of features. 
-Once the perceptron weights from all the tables are obtained, they're summed and thresholded based on a pre-decided value. 
-This led to a prediction mechanism that is highly accurate, has quick learning rate and is adaptable to changing program behaviour. 
-The features introduced in that paper were derived from Program Counters of the current and the last few instructions and the memory address of the current access. 
-Even with such basic features, high level of correlation could be established by perceptrons.
-
-The work was further extended by Multiperspective Reuse Prediction\cite{Multiperspective}, which introduced a varied set of parametric features. 
-Although design space exploration for feature and parameter selection was a non-trivial task, once fixed, it could give a cache replacement predictor highly suitable for the given set of applications.
-\newline
-\newline
-\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron prediction involves multiplying the vector of input features: F\textsubscript{1xN}, with the corresponding weight vector: W\textsubscript{Nx1} in a dot product fashion to obtain the sum y\textsubscript{out}.
-Here what we use is a perceptron-like structure. 
-The feature is used to hash into the weight of perceptrons and the retrieved weights are added straight away. 
-This can be seen a dot product of $\vec{1}$\textsubscript{1xN} with W\textsubscript{Nx1}. 
-This way, the perceptron algorithm doesn't introduce any multiplication operations in the inference or training process. 
-Hence what we adapt in this work, is a perceptron-like learning algorithm as it involves the same principle involved in perceptron inferencing and training. 
-For simplicity, we will call our implementation as perceptron implementation.
-\newline
-\newline
-\textit{Perceptron Update Rule:} In all the above implementations, a uniform perceptron update principle is being followed. 
-The weights need to be updated if the prediction was wrong or the predicted sum does not cross a certain threshold. 
-After a certain training period, these weights are proportional to the probability of the outcome being true. 
-This training threshold makes sure that the perceptrons are trained till a certain level of confidence and yet they are not over-trained to the given set. 
-Care is taken that the weights saturate at certain positive and negative values so that they remain confined to the bit-width. 
-The same rules will be applied in the discussion done in section XX on SPP-Perc learning algorithm.
-
-\subsection{Prefetching with SPEC 2017}
-\textit{<Discuss SPEC 2017 traces here>
-<What are the possible graphs / analysis tools we can use for SPEC traces?>}
-
-\subsection{Signature Path Prefetcher Architecture}
-Since the work we have done builds heavily on SPP\cite{SPP}, we are presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work. 
-Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists of following structural components:
-\newline
-\newline
-\textbf{Signature Table:} ST keeps track of 256 most recently accessed pages. 
-It is meant to capture memory access patterns within a page boundary. 
-SPP indexes into an entry of ST using the page number. 
-For each entry corresponding to a page, ST stores the `last block offset' and `old signature'. 
-Last block offset is the block offset of the last memory access of that given page. 
-The block offset is calculated with respect to the page boundary. 
-The signature is a 12-bit compressed representation of the past few memory accesses for that page. 
-The signature is calculated as:
-$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
-In case a matching page entry is found, the stored signature in the corresponding entry is retrieved.
-\newline
-\newline
-Consider the case that incoming page number 10 with a block offset 3 finds a match in the ST. 
-The retrieved pattern signature is 0x30 and the Last Offset is 1. 
-Since now there is the Last Block Offset (1), Incoming Block Offset (3), Old Signature (0x30), and New Signature calculated as per above equation (0x182), SPP can infer non-speculatively that the given pattern of memory accesses (as captured in the Old Signature) leads to the particular Delta.
-In general, Delta is defined as the difference between the prefetch suggestion block and the initial block which triggered the prefetch. 
-In this case, since SPP is in learning phase, it is defined as the difference between the Incoming Block Offset and the Last Block Offset (+2 in this case). 
-This, in turn generates the new memory pattern (New Signature). 
-This newly learned signature pattern and the delta is stored in the Pattern Table.
-\newline
-\newline
-\textbf{Pattern Table:} The PT is indexed by the signature generated from ST. 
-PT holds predicted delta patterns and their confidence estimates. 
-Each entry indexed by the signature holds up to 4 unique delta predictions. 
-This is implemented by making PT as a 4-way associative table. 
-To measure an estimate of a delta confidence, PT implements two counters: C\textsubscript{sig} for each entry indexed by the signature and C\textsubscript{delta} for each delta entry for a given signature.
-More details of this on-line score-keeping mechanism is described under Confidence Tracking discussion.
-
-\textit{Lookahead Prefetching:} For each trigger to the prefetcher, the first time that the SPP indexes into the PT, it does so in the non-speculative manner. 
-We call it non-speculative because the parameters which SPP uses to predict, like last block offset, incoming block offset and signature are well-defined. 
-Assuming PT generates a valid delta prediction (rather, up to 4 valid deltas), SPP enters into the lookahead mode.
-
-In the lookahead mode, SPP looks at the last generated delta(s), chooses the one with the highest confidence and speculatively generates the new effective address and the new signature. 
-Using these new parameters, it re-accesses the PT to generate further prefetches. 
-SPP keeps on repeating this cycle of accessing PT and updating the signature based on highest confidence prefetch from the last iteration. 
-These iteration count till which SPP manages to predict prefetch entries in the lookahead manner is characterized as its `depth'. 
-While doing this, SPP also keeps compounding the confidence in each depth. 
-Thus as depth increases, overall confidence keeps decreasing. 
-This makes sense since we are going deeper into the speculation path. 
-When the confidence falls below a pre-defined threshold, SPP stops further lookahead iterations
-\newline
-\newline
-\textbf{Global History Register:} One of the innovative ideas introduced in SPP was the Global History Register, which bootstraps learning across page boundaries.
-GHR is a small 8-entry fully associative structure. 
-Whenever a prefetch suggestion from the Pattern Table crosses page boundaries, instead of completely discarding it, GHR saves that information for inter-page learning.
-Since the SPP-Perc implementation in this paper doesn't affect GHR functionality in any ways, we won't be going into more details about GHR.
-\newline
-\newline
-\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch Filter - a 1024 entry 1-way associative table which keeps a record of last few entries prefetched by SPP. 
-This proves useful for following reasons:
-\begin{itemize}
-    \item \textbf{Eliminating duplicate prefetches}\newline 
-    Any subsequent prefetches recommended by SPP mechanism is checked against the existing entries in Prefetch Filter. 
-If a valid entry is found, it means that the particular cache line is already either in L2 Cache or being fetched. 
-Hence there is no need to issue a duplicate prefetch.
-
-    \item \textbf{Updating prefetcher}\newline
-    If a subsequent demand access comes before eviction for a prefetch that was recommended by SPP, various components of SPP are updated to reflect that this was a good prefetch. 
-Correspondingly, if a prefetched cache line is evicted without being demand hit even once, SPP learns that this was a useless prefetch and updates the various tables accordingly
-
-    \item \textbf{Tracking Global Accuracy}\newline
-    PF keeps a track of usefulness of the prefetches by introducing 2 counters: C\textsubscript{total} to track the total number of prefetches getting recorded in the PF and C\textsubscript{useful} to keep a track of which prefetches recorded in the PF led to a demand hit.
-The ratio C\textsubscript{useful} / C\textsubscript{total} is the global accuracy for the prefetching scheme. 
-It is represented as $\alpha$ in the paper. 
-The global accuracy acts a real-time throttler for the path confidence value as depicted in the following discussion.
-\newline
-\end{itemize}
-\textbf{Confidence Tracking}: The PT keeps a track of hits to each signature through a counter C\textsubscript{sig}. 
-The number of hits for a given delta per signature are tracked using a counter C\textsubscript{delta}. 
-The confidence for a given delta is approximated through C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}. 
-When SPP enters into a lookahead mode, the path confidence P\textsubscript{d} is given as:
-$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown in the Prefetch Filter section. 
-The range of global accuracy is in [0,1]. 
-The lower the value of $\alpha$ for a given trace or a program phase, the overall confidence is scaled down and this makes SPP less aggressive for the upcoming prefetches.
-
-Here d is the lookahead depth. 
-For d = 1, when SPP is in non-speculative mode, P\textsubscript{0} can be thought of as 1. 
-Once SPP enters into speculative mode, it calculates the path confidence as a product of confidence of the individual prefetches encountered along that path, further scaled using $\alpha$. 
-The final P\textsubscript{d} is thresholded against prefetching-threshold (T\textsubscript{p}) to decide whether to prefetch or not. 
-If yes, then P\textsubscript{d} is thresholded against a numerically bigger fill-threshold (T\textsubscript{f}) to decide whether the prefetch should be sent to L2 Cache (high confidence prefetch) or Last Level Cache (low confidence prefetch)
-\newline<Insert a comprehensive figure about SPP structures>
-\newline<Insert another detailed picture about SPP data-path flow>
-
-\subsection{Case for an On-line Throttler}
-As compared to some of the other state of the art prefetchers [BOP], SPP is a far less aggressive prefetcher. 
-In the single core environment, this gives BOP an edge as there is no resource contention among the different cores. 
-Hence more aggressive prefetching is bound to prove beneficial. 
-As we increase the core count, we observe that SPP starts outperforming rest of the prefetchers. 
-This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. 
-\textit{<Figure showing the variation of SPP vs BOP wrt core count>} 
-For 4-core applications, SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
-
-The above analysis shows that with a more careful throttling mechanism, SPP can be tuned to become much more aggressive, leading to increased coverage. 
-The onus of maintaining the accuracy now falls on the independent throttler. 
-To test the hypothesis, we tuned down SPP to the minimum possible threshold. 
-Only by doing this we managed to increase the prefetch suggestions made by SPP, by XX\%. 
-\textit{<Comparison of SPP-Unleased with SPP / BOP>}
-Obviously this came at a cost of increased DRAM traffic and cache pollution.
-
-Moreover, the on-line confidence mechanism introduced in SPP was very rudimentary.
-It was based on taking a ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig} as explained above.
-The same confidence was used to make the decision whether to prefetch or not to prefetch; and which level to prefetch.
-While this approximation was shown to work in the original SPP, we believe that a better form of on-line decision making was possible.
-Hence, it was necessary to build a robust and adaptable learning mechanism to accept / reject the prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level Cache).
-
-To that effect, we introduce an independent on-line perceptron based throttling mechanism.
-
-\section{SPP-Perc}
-
-\subsection{Changes made to original SPP}
-To modify the SPP Design to suit our scheme, the following changes were made:
-
-\begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline
-In the original SPP design, an internal confidence mechanism was developed which was compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the prefetch suggestion.
-Furthermore, the accepted suggestions were thresholded against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache vs Last Level Cache.
-T\textsubscript{p} and T\textsubscript{f} were set to 25 and 90 respectively, on the scale of 0 to 100.
-In SPP-Perc, the perceptron sum is used to make these decisions. 
-Hence these two thresholds are no longer needed
-
-\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
-Original SPP would not consider the slots available in the L2 MSHR queue. 
-It had an assumption that at no time can the number of prefetches suggested by SPP could exceed the capacity of the queue. 
-While it was not proven in the paper, the internal confidence mechanism along with the thresholds made sure that the above assumption was maintained. 
-In SPP-Perc since there can be no assumption on the degree of aggressiveness - because SPP is being tuned to its lowest thresholds, we explicitly check that at no point should the number of suggested prefetches exceed L2 MSHR queue.
-
-\item \textbf{Enhanced Filter}\newline
-Original SPP introduces a concept of Prefetch Filter, which keeps a track of up to 1024 L2 Cache prefetches suggested by SPP. 
-The Filter was modified to store meta-data like program counter, memory access address etc. 
-This is needed to train the perceptron when the result of that particular prefetch gets available. 
-At training time, this data is used to retrieve the state of the prefetcher when the prediction was made.
-
-\item \textbf{Reject Filter introduced}\newline
-The Prefetch Filter mentioned above keeps a track of the prefetches that actually happened in the L2 Cache. 
-Correspondingly, we introduced a reject filter. 
-It keeps a track of the prefetches that were suggested by the base SPP engine but were rejected by the perceptron. 
-In case it is detected that a future demand fetch could have used this rejected prefetch, the perceptron weights are updated to reflect the same
-
-\end{itemize}
-
-\subsection{The Perceptron Layer}
-As discussed earlier, the perceptron layer was introduced to act as a throttler to the underlying SPP engine. 
-This section explores the details of the perceptron layer and the range of features that are ultimately used by the perceptron for correlating the prefetches.
-
-The neural network weights are stored in separate tables for each feature.
-The indexing of each feature is done using different number of bits.
-At most 12 bits of a feature are used to index into the perceptron table.
-Certain features require more resolution power \textit{i.e.}, full 12 bits of indexing.
-On the other hand, some features requiring lower resolution required as low as 7 bits of indexing.
-The variable indexing was determined by studying the features and fine-tuned empirically so as to achieve a good accuracy vs area trade-off.
-Exact details are mentioned under `Area Overheads' in Section XX.
-
-A single entry in the table corresponds to a perceptron weight.
-Each weight is a 5 bit counter - saturating at -16 and +15. 
-Each feature gets its own dedicated table. 
-Our proposed design uses 9 features. 
-As the program execution begins, all the weights are initialized to 0.
-\newline \newline
-\textbf{INFERENCING}\newline
-In the Champsim simulator, the prefetching method is triggered on every L2 Cache demand access. 
-At that point, the prefetcher has the option to do or not do prefetch - in case it does, then how many cache lines to prefetch. 
-The suggested prefetches can be places either in L2 Cache or in Last Level Cache. 
-The prefetcher gets to decide this, and is usually done based on an internal confidence mechanism.
-
-When SPP gets triggered, it suggests up to 4 (=Associativity of the Pattern Table) candidates for prefetch based on the demand access address. 
-In each subsequent lookahead depth, SPP suggests up to 4 more prefetch candidates, as explained in section\_\_. 
-All these recommended prefetches are tested in the perceptron throttler before they can access the Prefetch Filter, as in the original SPP data-path.
-
-The perceptron looks at the micro-architectural state, or `features', at that instant. 
-Each feature indexes into the hash of up to 4096-entry table dedicated for that feature. 
-Feature \#1 indexes into a particular entry in table \#1 and so on.
-
-Once all the weights are retrieved, they are added together, much like the traditional perceptron dot product. 
-On obtaining the overall sum, it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO. 
-Only the prefetch candidates with the perceptron sum higher that the threshold qualify for prefetching. 
-The prefetches that qualify the perceptron throttler stage go through the Prefetch Filter stage, just as in the SPP design.
-The Prefetch Filter eliminates any redundant prefetch suggestions.
-In SPP-Perc, all the meta-data that is required to recreate the features is also stored in the Prefetch Filter. 
-At a later stage, when the feedback of the current prefetched line is available, the stored data can be used to train the perceptron.
-
-The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch cache line. 
-All prefetch candidates qualified till this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill level.
-The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
-
-The perceptron sum can also be seen as the sum of the individual contribution per feature.
-The value of each contribution corresponds to the extent of confidence of the final output with that given feature.
-By summing the individual contributions, the final perceptron sum denotes the overall confidence for that prefetch suggestion. 
-In a way, by thresholding the perceptron sum on two different thresholds, we are diving the confidence scale of the prefetch suggestion into three bins.
-The first bin corresponding to the lowest confidence leads to prefetch candidate being rejected.
-The next bin corresponds to prefetches with a moderate confidence level. 
-Such prefetches are directed to fill the bigger Last Level Cache and potentially not pollute the more scarce L2 Cache.
-The highest confidence prefetches are filled in the L2 Cache.
-
-In addition to the Prefetch Filter mentioned above, SPP-Perc also maintains a `Reject Filter'. 
-Reject filter is a 1024-entry deep 1-way associative table. 
-In case that a prefetch suggestion is rejected by perceptron throttler, it instead gets logged into the Reject Filter. 
-The Reject Filter keeps track of 1024 entries that were suggested by the Pattern Table but got rejected by the perceptron throttler. 
-This can be used to train the perceptron to avoid false negatives \textit{i.e.}, cases where prediction was to reject the prefetch but the prefetch would have been useful.
-
-As previously stated, an enhancement to the original SPP design includes storing all the meta-data describing the program state at the instant of prefetching, in the Prefetch Filter. 
-This information is useful when the perceptron needs to be updated subsequently. 
-\newline \newline
-\textbf{TRAINING}\newline
-In the prefetching environment, feedback of a prefetch is received whenever there is an eviction or a demand access from L2 Cache. 
-This acts as a trigger to update all the baseline SPP entries and to also train the perceptron throttler. 
-Training of the perceptron throttler involves restoring the meta-data that was stored in Prefetch Filter or the Reject Filter. 
-This is done by indexing into the corresponding filters using the cache line address of the block being used to train. (10-bits indexing + 6-bits tag matching). 
-Once the state of the program at the time of prefetching is available, it is used to index into the perceptron weights table.
-
-If a demand access block which triggers the training  was tagged as a valid prefetch in the Prefetch Filter, it means that the earlier prefetch prediction was correct. 
-In that case the perceptron weights are incremented by 1 if the predicted sum does not cross a pre-defined threshold. 
- In case it is a cache block eviction that led to training and the corresponding valid entry was found in the Prefetch Filter, it shows that the prediction made by the perceptron was wrong. 
-Perceptron should have ideally rejected the prefetch suggestion as a low-confidence prefetch. 
-Here the weights are decremented by 1 to reflect the same. 
-In either case, weights are saturated at -16 or +15.
-
-A secondary training mechanism also kicks in during demand fetches. 
-Before the demand access triggers the next set of prefetches, it checks the Reject Filter for the valid entry. 
-In case it is a hit, that means that the corresponding cache line was initially suggested by the underlying SPP engine but rejected by the perceptron throttler. 
-This means that perceptron should have been more confident about that particular prefetch. 
-Once such a scenario is identified, the state of the execution at the time of prefetch is retrieved from the Reject Table. 
-The retrieved data is used to index into the various weights tables of the perceptron and the corresponding values are updated by +1, saturating between and -16 and +15. 
-This update reflects increased confidence for the prediction corresponding to that prefetch.
-
-This mechanism allows us to solve the classical problem of exploiting `lost opportunity'. 
-In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
-Our two-step SPP-Perc allows us to overcome that issue in a simple manner.
-
-\subsection{Features used by Perceptron}
-
-Here we discuss the various features which help correlate the prefetching decision with the program behaviour
-\begin{itemize}
-\item \textbf{Base address} of the demand access which triggered the prefetch. 
-Since prefetch are triggered from the L2 demand access, they tend to be correlated with the triggering address.
-
-\item \textbf{Cache line} and \textbf{Page address}: These two separate features are derived from the base address which triggered the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and base\_addr >> LOG2\_PAGE\_ SIZE respectively.
-The idea behind using three different shifted versions of the same feature is that it allows us to look into a wider range of bits than with a single version.
-It also helps give more importance to the overlapping bits and lesser importance to most and least significance bits.
-This approach can also eliminate destructive interference that can be caused by directly folding the bits into half.
-
-\item \textbf{Confidence}: The integer confidence on the scale of 0 to 100 which was used in the original SPP design.
-While the original confidence is not used directly for decision making, it can still contribute to the final decision made.
-
-\item \textbf{Program Counter XOR Depth}: PC is for the instruction which triggered the prefetch chain.
-Depth refers to the iteration count of the lookahead stages.
-In general, PC is not considered as a good basis for doing lookahead prefetching as all the prefetches with depth >= 1 gets aliased onto the same PC which won't be true in actual demand fetch case.
-This feature resolves a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases.
-This is akin to the concept of Virtual Program Counters\cite{VPC} introduced by Kim \textit{et. al.} for indirect branch prediction.
-
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC before the instruction that triggered the current prefetch. 
-Hashing together the last three PCs tell SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block.
-PCs are shifted in the increasing order of history before being hashed together.
-This is done to avoid the resultant value of zero in case 2 or more PCs are the same. 
-Additionally, obfuscating the information as it gets old allows us to get a wider and an approximate look into the program history.
-
-\item \textbf{Program Counter XOR Delta}: This tells us if a given PC favours particular value(s) of delta.
-As noted earlier, while PC in itself does not much useful information, this hash resolves PC into different values based on the tendency of that PC to favour a certain delta.
-This way, dynamic nature of different instances of the same memory access instruction can be captured here.
-
-\item \textbf{Current Signature XOR Delta}: It can be recalled from the discussion of SPP in section XX that the new signature is generated using the old signature and the blobk delta.
-While `Current Signature XOR Delta' is not the exact formula for generating the future signature, it gives an approximate idea of the path that the combination of these two values can lead to.
-
-%%%%%%% Below are the features rejected when downsizing 14 -> 9
-%\item \textbf{Current Signature}: The 12-bit signature used by SPP to index into the Pattern Table
-%\item \textbf{Delta}: The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched
-%\item \textbf{PC}: Program Counter of the Load instruction that triggers that particular prefetch
-%\item \textbf{Base address XOR Delta}: If any particular value(s) of delta are highly favoured by a given address which triggered the prefetch, this composite feature can capture that information
-%\item \textbf{Cache line XOR Delta}: This feature captures if any particular delta is highly favoured by a given cache line address
-%\item \textbf{Confidence XOR Delta}: This gives a combined insight into that fact that certain delta-confidence pairs predicted by SPP might correlate more with the prefetch outcome
-%%%%%%%%%%%%%%%%%%%%%%
-
-\end{itemize}
-
-Some of the composite features are derive from simple hashing (XOR) of two primary features.
-There is always a question of usefulness of such composite features and the `fresh' information added.
-We try to justify the choice of each feature by quantifying the contribution made towards predicting prefetch behaviour, in Section XX.
-Finally, as noted above, each feature indexes into its independent entry of perceptron weights.
-
-\section{METHODOLOGY}
-\subsection{Performance Model}
-For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim} simulator. 
-ChampSim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} 
-We model model a 1-core / 4-core / 8-core Out of Order machine. The details of the configuration parameters is summarized in Table  \ref{tab:Sim_params}.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|l|p{3.6cm}|c|}
-    \hline
-    Level & Configuration & Access Latency \\
-    \hline
-         L1 Cache & Separate I-cache and D-cache, 32 KB, 8-way & 4 cycles\\
-         L2 Cache & Private, 256 KB & 8 cycles\\
-         LLC & Shared, 2MB / core, 16-way & 20 cycles\\
-         DRAM & 4 GB Single Channel for single-core, 8 GB Double Channel for multi-core & N?? cycles\\
-    \hline
-    \end{tabular}
-    \caption{Memory Model Parameters}
-    \label{tab:Sim_params}
-\end{table}
-
-The block size was fixed 64KB. Prefetching was only triggered at L2 Cache demand accesses but could be sent to L2 Cache or Last Level Cache. 
-No L1 Data level prefetching was done. 
-LRU replacement policy was used on all levels of cache hierarchies. 
-Branch prediction was done using the perceptron branch predictor\cite{Perc_Branch}. 
-Page size was kept as 4KB. 
-Champsim operates all the prefetchers strictly in the physical address space.
-
-\subsection{Testing Under Additional Memory Constraints}
-The default single-core configuration simulates a 2MB LLC and a single channel DRAM with 12.8GB/s bandwidth. 
-We extend the simulations to include memory constraints introduced in DPC-2. 
-Specifically we look at the following two variations:
-\begin{itemize}
-    \item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited to 3.2 GB/s
-    \item \textit{Small LLC}: In this scenario, LLC size is reduced to 512 KB
-\end{itemize}
-All the multi-core simulations are only done in the default configuration.
-
-\subsection{Workloads}
-This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017} has been used to characterize and measure the prefetch performance. 
-We use all the 20 workloads available in the SPEC 2017 Suite. 
-Using SimPoint\cite{SimPoint} methodology, we identified 95 different program segments of 1 Billion instructions each.
-
-We performed cross-validation of our SPP-Perc model using SPEC 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.
-For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
-For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.
-The traces include 4 multi-core applications with 6 distinct phases per application. 
-In total, we used 285 traces representing workloads across 53 applications. 
-\newline
-\newline
-\textit{Single-core performance:} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics.
-We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
-We take the Instruction Per Cycle (IPC) of the prefetcher and divide it with the IPC of baseline. 
-The final number reported is the geometric mean of the speed-up attained on individual traces.
-\newline
-\newline
-\textit{Multi-core performance:} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
-For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 
-Each CPU keeps executing its workload till the last CPU completes 1 Billion instructions after warm-up. 
-For collecting IPC and other data, only the first billion instructions are considered as the ``region of interest''.
-
-Here we report the weighted speedup normalized to baseline \textit{i.e.}, no prefetching. 
-For each of the workloads running on a particular core of the 4-core 8 MB LLC system, we compute IPC\textsubscript{i}. 
-We then find the IPC\_isolated\textsubscript{i} of the same workload running in isolated 1-core 8 MB LLC environment. 
-Then we calculate the total weighted-IPC for a given workload mix as $\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}). 
-For each of the 100 workload-mix, the sum obtained is normalized to the weighted-IPC calculated similarly for baseline case \textit{i.e.}, no prefetching, to get the weighted-IPC-speedup. 
-Finally the geometric mean of these 100 weighted-IPC-speedup is reported as the effective speedup obtained by the prefetching scheme.
-
-We repeat the same process for 8-core workloads with the exception that 20 Million warm-up instructions and 100 Million full instructions were executed.
-This was done so as to keep the simulation run-time within reasonable limits as a single 8-core mix was taking up to 5 days to simulate 1 Billion instructions.
-
-\subsection{Preferchers Simulated}
-We compared SPP-Perc against three of the latest state of the art prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher (SPP).
-BOP was the winner of 2nd Data Prefetching Championship. 
-DA-AMPM is the enhanced version of AMPM, modified to account for DRAM row buffer locality. 
-SPP has been shown to outperform BOP on SPEC 2006 traces. 
-For each of these, we compare their speedups taking no prefetching as the baseline.
-
-\subsection{Developing Features for SPP-Perc}
-This section describes the intuition and analysis that went behind finalizing the features. 
-As noted earlier, we developed a set of 9 features which allow the perceptron throttler to correlate prefetching decision with the program context. 
-To study the correlation across each feature, we study statistically the perceptron weights and try to interpret their distribution.
-\newline  \newline
-\textbf{GLOBAL PEARSON'S CORRELATION}\newline
-For this experiment, we create a dump of the state of perceptron weights at the end of all trace execution.
-The weights obtained from running all the SPEC 2017 traces are concatenated together. 
-Since the weights are collected at the end of individual trace execution, the perceptron weights have attained a relatively stable value by now.
-Hence, this dump represents a `snap' of the trained perceptron weights across all the SPEC 2017 traces.
-The intuition here is that the feature with bulk of the perceptron weights concentrated around 0 or small magnitude numbers show a weak correlation with the prefetching outcome. 
-On the other hand, features with most of the weights saturated around highest value (+15) show a high positive correlation and the features with weights close to the lowest value (-16) show a strong negative correlation.
-
-We plot a histogram for each feature depicting weights distribution from -16 to +15 and use this histogram to generate the Pearson's correlation factor for that feature.
-In statistics, Pearson's factor is a numerical measure of the degree of linear correlation between two variables.
-It ranges from -1 to +1.
-The magnitude of Pearson's factor depicts the extent of correlation and the sign depicts whether it is a positive correlation or a negative correlation.
-Values close to 0 suggest a low correlation or at times, noisy data.
-A value of +1/-1 suggests a perfectly linear positive / negative correlation respectively.
-Figure XX depicts the histogram and the Pearson's coefficient for two of the features, YY and ZZ.
-
-Figure XX shows all the features used, arranged in the increasing order of their Pearson's factor.
-As can be seen X out of the 9 features provide a moderate to high correlation, with the magnitude of P-value > 0.4.
-The single most important feature, \_\_ helps provide a correlation to prefetch outcome with a factor of \_\_.
-\newline \newline
-\textbf{PER TRACE CORRELATION} \newline
-Another important way to look at the perceptron features is to see how much their contribution varies across the traces.
-Here we give special attention to features with low P-values in the previous experiment.
-Figure XX shows the variation of P-values three features : \_\_, \_\_ and \_\_; across all the SPEC 2017 traces.
-For simplicity, the traces are arranged in an increasing order of contribution made by the feature and the trace names have been omitted.
-It can be seen that even features with a low overall correlation provide useful correlation (magnitude > 0.4) for XX out of the 95 simpoints developed for SPEC 2017.
-\newline \newline
-\textbf{TRIMMING FEATURES USING CORRELATION} \newline
-Besides providing interesting insights into prefetching behaviour, P-value can also be used for feature selection and prefetcher tuning.
-
-Here, we introduce the concept of cross-correlation across features.
-Just as we examined correlation of each feature with the final outcome, we can also study correlation between the features.
-We used the above methodology to eliminate features which provide a little or no information that has already been captured in other features.
-
-As a part of feature selection, we initially came up a mix of 23 features- primary and composite.
-By studying cross correlation of each of these features against others in a 23x23 matrix, we identified pairs of features with correlation factor > 0.9 in magnitude and carefully eliminated redundant features.
-Using this approach, we managed to reduce the feature count to 9.
-Thus, in the final implementation of SPP-Perc, no two features have a high correlation between them.
-This way we can be sure that each feature is bringing in contribution that cannot be captured using other features.
-
-\textit{<TODO: VALIDATE THIS>} 
-Secondly, studying the relative importance of each feature enabled us to vary the number of entries dedicated for each feature.
-Features with higher Pearson's correlation (like \_\_ and \_\_) were given most importance and allowed full 12-bits of indexing.
-Features like \_\_ and \_\_, with a low overall P-value were allocated fewer indexes in the feature table.
-
-To conclude, in the above discussion we justify the features for perceptron from a statistical viewpoint.
-We also show how this information can also be used for prefetcher tuning.
-All this study was made possible only because we used on-line perceptron learning for prefetching and that enabled us to examine the weights in detail.
-
-\subsection{Overhead for SPP-Perc}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|c|c|m{4.8cm}|}
-    \hline
-        \textbf{Field} &
-        \textbf{Bits} &
-        \textbf{Comment} \\
-    \hline
-         Valid & 1 & Indicates a valid entry in the filter\\
-         Tag & 6 & Identifier for the entry in the filter\\
-         Useful & 1 & To show if the given entry led to a useful demand fetch\\
-         Perc Decision & 1 & Prefetched vs Not-prefetched \\
-    \hline
-        PC & 12 & \multirow{5}{4.8cm}{Meta-data required for perceptron training}\\
-        Address & 24 & \\
-        Delta & 7 & \\
-        Curr Signature & 12 & \\
-        Confidence & 7 & \\
-    \hline
-        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
-    \hline
-    \end{tabular}
-    \caption{Meta-data Stored in Prefetch Filter}
-    \label{tab:PF_metadata}
-\end{table}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|p{1.5cm}|p{1cm}|c|p{1.5cm}|}
-    \hline
-        \textbf{Structure} &
-        \textbf{Entry} &
-        \textbf{Components} &
-        \textbf{Total} \\
-    \hline
-        \multirow{5}{1.5cm}{Signature Table} & \multirow{5}{1.5cm}{256} & Valid (1 bit) & \multirow{5}{1cm}{11008  bits}\\
-        & & Tag (16 bits) &\\
-        & & Last Offset (6 bits) &\\  
-        & & Signature (12 bits) &\\
-        & & LRU (6 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Pattern Table} & \multirow{3}{1.5cm}{512} & $C_{sig}$ (4bits) &\multirow{3}{1cm}{24576 bits}\\
-        & & $C_{delta}$ (4*4 bits) &\\
-        & & Delta (4*7 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Perceptron Tables} & \multirow{3}{1.5cm}{4096*8 128*1} & \multirow{3}{1.5cm}{5 bits} & \multirow{3}{1.5cm}{164480 bits}\\
-        & & &\\
-        & & &\\
-    \hline
-        \multirow{2}{1.5cm}{Prefetch Filter \footnotemark[1]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{84 bits} & \multirow{2}{1.5cm}{86016 bits}\\
-        & & &\\
-    \hline
-     \multirow{2}{1.5cm}{Reject Filter \footnotemark[2]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{83 bits} & \multirow{2}{1.5cm}{84992 bits}\\
-        & & &\\
-    \hline
-        \multirow{4}{1.5cm}{Global History Register} & \multirow{4}{1.5cm}{8} & Signature (12 bits) & \multirow{4}{1.5cm}{264 bits}\\
-        & & Confidence (8 bits) & \\
-        & & Last Offset (6 bits) & \\
-        & & Delta (7 bits) & \\
-    \hline
-        \multirow{2}{1.5cm}{Accuracy Counters} & 1 & C$_{total}$ & 10 bits\\
-        & 1 & C$_{useful}$ & 10 bits\\
-    \hline
-        \multirow{3}{1.5cm}{Global PC Trackers} & \multirow{3}{1cm}{3} & $PC_1$ (12 bits)& \multirow{3}{1.cm}{36 bits}\\
-        & & $PC_2$ (12 bits) &\\
-        & & $PC_3$ (12 bits) &\\
-    \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 371356 bits = 45.33 KB}}\\
-    \hline
-    \end{tabular}
-    \caption{SPP-Perc Storage Overhead}
-    \label{tab:SPPPerc_overhead}
-\end{table}
-\footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}
-\footnotetext[2]{RF does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
-
-In this section, we analyze the hardware overhead required to implement SPP-Perc. 
-The Prefetch Filter was enhanced to accommodate storing of meta-data for perceptron training. 
-Table \ref{tab:PF_metadata} depicts the meta-data stored for each entry in the Prefetch Filter.
-Table \ref{tab:SPPPerc_overhead} shows the total storage overhead of SPP-Perc implementation.
-The hardware budget for 2nd Data Prefetching championship was 32 KB. 
-Keeping that in mind and the speedup SPP-Perc obtained over the winner, the extra hardware budget can be accounted for.
-The extra hardware also makes the overall scheme more scalable than SPP. 
-In the original SPP paper, it was demonstrated that adding extra hardware brings little advantage in terms of performance gain.
-The newly added perceptron tables can be scaled to increase / decrease features depending on the permitted budget.
-
-In terms of computations, the perceptron mechanism only introduces an extra adder tree.
-The hash perceptron mechanism makes sure than there is no actual vector multiplication happening in the hardware.
-Obtaining the perceptron sum requires addition of 9 5-bit numbers. 
-Using an adder tree of 4 5-bit adders, this can be done in ceil($log_{2}9$) = 4 steps.
-Perceptron update only requires weight update by +1 or -1.
-Thus, all the operations required for perceptron inferencing or updating the states of the perceptron throttler can be easily done in the time constraints of L2 Cache Accesses.
-
-\section{RESULTS}
-This section discusses the results obtained from running SPP-Perc, in terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per Cycle (IPC) metrics. 
-For SPEC 2017 suite, first we present the results for single-threaded workloads then the results for multi-core workloads are presented.
-
-\subsection{Single-core Results}
-Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated. 
-All the results have been normalized to baseline \textit{i.e.}, no prefetching. 
-As can be seen, SPP-Perc yields a speedup of XX\% over the baseline. 
-This is equivalent to XX\% over DA-AMPM, xx\% over BOP and XX\% over SPP. 
-Out of the XX benchmarks, SPP-Perc nearly matches or outperforms most of the prefetchers on XX traces.
-
-At its peak, SPP-Perc manages to get the IPC speedup of a factor of over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}. 
-This also corresponds to speedup gain of \textbf{22.4\%} over the next best prefetcher - SPP. 
-In general, benchmarks \textit{602.gcc\_s, 603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s} benefit the most from SPP-Perc, as compared to SPP only prefetching.
-
-<Discuss the extreme behaviour of \textit{623.xalancbmk\_s}: Improvement over SPP is as high as 31\% and even reduces to -1.5\% on traces of this benchmark>
-
-<Discuss about coverage on all traces>
-
-<Discuss cases where coverage gets specially increased>
-
-<Discuss accuracy on all traces>
-
-<Discuss about compelling increase in accuracy - details of traces. Plot linear separability>
-
-<Any case where accuracy got reduced - go into the trace. See if linear separability is the issue. Other possible issues?>
-
-\subsection{Multi-core Results}
-In this section, we demonstrate the improvement achieved by SPP-Perc for a mix of multi-programmed workloads. 
-\newline \newline
-\textit{4-core environment}
-Figure \_\_ shows a comparison of speedups obtained on workload mix of memory intensive subset of SPEC 2017. 
-We plot all the 4 prefetchers, normalized to baseline. 
-The workloads have been sorted in the increasing order of performance benefit, as measured in terms of normalized weighted speedup. 
-As can be seen, SPP-Perc offers an IPC improvement by 33.5\% on these traces. 
-This is 5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP and DA-AMPM respectively.
-
-This increased improvement by SPP-Perc on a 4-core machine is not surprising. 
-Since SPP-Perc has an extremely intelligent throttler working to eliminate useless prefetches before they can create cache pollution. 
-This further amplifies the edge that SPP had over BOP. 
-With the advantage of an orthogonal prefetcher, SPP can be as aggressive as possible and leave the accuracy aspect to perceptron throttler.
-\newline \newline
-\textit{8-core environment}
-To test the effectiveness of SPP-Perc in a tighter constrained multi-core environment, we simulated an 8-core machine with 16 MB last Level Cache. 
-Other parameters were kept same as discussed under Methodology in Section XX.
-<See if we can use speed-up over BOP as a measure here>
-
-\subsection{Additional Memory Constraints}
-Figure \_\_ shows the performance of SPP-Perc against the other prefetchers in tighter constraint environments. 
-For sake of simplicity, we show 2 traces where SPP-Perc gives the most advantage over baseline, 2 traces where SPP-Perc doesn't perform as well as the best prefetcher *<-replace with the prefetcher name once results are available* and finally GeoMean across all the traces.
-<Hopefully BOP would degrade more here as it is super aggressive. 
-By definition, SPP-Perc should be better than SPP as the throttler would better control the aggressiveness. 
-In SPP paper DA-AMPM performs close enough to BOP. That should get replicated here too.>
-
-\subsection{Cross Validation}
-Here we try to demonstrate the robustness of SPP-Perc by testing it on completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
-
-Fig \_\_ shows the speed-up achieved by BOP, SPP and SPP-Perc on selected traces of SPEC 2006, along with the overall geometric mean over whole of the benchmark and over memory-intensive subset of the SPEC 2006 benchmark.
-As can be seen, SPP-Perc peovides a speed-up of 44.4\% over baseline, on the memory intensive subset of SPEC 2006 benchmark.
-This corresponds to 6.84\% over SPP and XX\% over BOP.
-On the whole of the SPEC 2006 suite, the speedup obtained is 22.4\%, which is 3.34\% ahead of SPP.
-
-The 4-core results for SPEC 2006 are \_\_.
-
-As can be seen, the performance is very consistent given no prior parameter tuning was done keeping SPEC 2006 in mind. 
-We attribute this to the inherent adaptability of the perceptron model.
-In general, perceptron weights are able to adjust in real-time so as to find the best possible correlation between the output and the features.
-
 \section{CONCLUSION}
-In this paper, we introduced perceptron based learning for data prefetching in the form of SPP-Perc. 
-Perceptron acts an orthogonal throttler to the underlying prefetch engine. 
-SPP-Perc improves performance by over XX\% over the baseline, which corresponds to YY\% over the next best performing prefetcher. 
-We believe this paper can lead to interesting future work like generalizing SPP-Perc to work with any prefetcher with minimal knowledge of the underlying prefetcher architecture.
-
-%%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%
-
-%%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
-\bibliographystyle{ieeetr}
-\bibliography{ref}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
+In this paper, we introduced perceptron based learning for data
+prefetching in the form of SPP-Perc.  Perceptron acts an orthogonal
+throttler to the underlying prefetch engine.  SPP-Perc improves
+performance by over XX\% over the baseline, which corresponds to YY\%
+over the next best performing prefetcher.  We believe this paper can
+lead to interesting future work like generalizing SPP-Perc to work
+with any prefetcher with minimal knowledge of the underlying
+prefetcher architecture.

### 2018-11-15 design.tex
reorg

--- 2018-11-15 paper_1.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is the template for submission to ISCA 2018
-% The cls file is a modified from  'sig-alternate.cls'
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{sig-alternate} 
-\usepackage{mathptmx} % This is Times font
-
-\newcommand{\ignore}[1]{}
-\usepackage{fancyhdr}
-\usepackage[normalem]{ulem}
-\usepackage[hyphens]{url}
-\usepackage{microtype}
-\usepackage{fixltx2e}
-\usepackage{array}
-\usepackage{multirow}
-
-% Always include hyperref last
-\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=black,citecolor=blue,urlcolor=black]{hyperref}
-
-% Ensure letter paper
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\newcommand{\iscasubmissionnumber}{NaN}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\fancypagestyle{firstpage}{
-  \fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-  \fancyhead[C]{\normalsize{}} 
-  \fancyfoot[C]{\thepage}
-}
-
-\pagenumbering{arabic}
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\title{SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher} 
-\author{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\maketitle
-\thispagestyle{firstpage}
-\pagestyle{plain}
-
-\begin{abstract}
-\textbf{Hardware prefetching has been introduced in modern processors as a way to hide cache latencies. 
-An efficient prefetcher should be able to identify complex memory access patterns during a program execution. 
-This enables it to fetch a block ahead of its demand access, potentially saving a cache miss.
-}
-
-\textbf{In this paper, we introduce perceptron learning to help make this prefetching decision efficiently. 
-We have taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine. 
-The perceptron layer acts as a throttler to filter out the unnecessary prefetches recommended by SPP. 
-We have also explored a range of features that can be used to train the perceptron layer. 
-Our results show that SPP-Perc improves performance on the memory intensive subset of the SPEC 2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared to just SPP.
-We also demonstrate that the performance gained from using our efficient throttler scales better with increasing number of cores.
-}
-\end{abstract}
-
-\section{Introduction}
-Since past few decades, processor and memory scaling has been enabled with different philosophies in mind. 
-While processor scaling focused on speed improvements, memory scaling was centred around increasing the storage capacity. 
-The process technology reaching scaling limitations led to Memory Wall\cite{MemWall} - the increasing performance gap between the processor speed and the memory speeds. 
-Various techniques have been developed to combat this widening gap. 
-Data prefetching is one such important technique. 
-Data prefetching exploits the fact that in most applications, memory accesses have a well-defined pattern.
-
-An ideal prefetching scheme involves capturing this memory access patterns to predict future prefetches in a timely manner. 
-Other important decisions faced by the prefetcher are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block.
-Memory accesses patterns can be as simple as fetching the next cache line or a complex but predictable pattern, like in pointer chasing applications. 
-Predicting this often requires a trade-off between coverage and accuracy. 
-For a prefetcher to capture highly irregular access patterns, it needs to be highly aggressive (high coverage). 
-That also means the prefetcher would end up recommending a lot of useless prefetches, leading to cache pollution (low accuracy). 
-This effect would get even more pronounced in multi-core scenario where the last level cache is a shared resource. 
-Conversely, a conservative prefetcher would be highly accurate but might not make enough useful predictions.
-
-To this effect we propose SPP-Perc <come up with a catchy name>. 
-It is an enhancement to the existing state of the art prefetcher and overcomes the coverage vs accuracy trade-off in a robust manner. 
-The idea is to modify SPP design to make it as aggressive as it can get. 
-This enables SPP to capture complex memory access patterns hence leading to increased coverage.
-This comes at a cost of increased DRAM traffic, hence reduction in overall IPC.
-The prefetch suggestions recommended by SPP pass through the perceptron based throttler. 
-Over time, perceptron layer learns to correlate the prefetches recommended by SPP with various available features, leading to elimination of useless prefetches. 
-This leads to an overall increased accuracy of the prefetch scheme and reduction in DRAM traffic.
-
-\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and future scope for expansion. 
-The contributions made by this paper are:
-
-\begin{itemize}
-\item This is the first time where a truly on-line learning based neural network model has been used when it comes to hardware data prefetching. 
-The prior works in this area either relied on program semantics\cite{Semantics} or were application specific\cite{Datacenter} or could not yield a significant improvement over the baseline prefetcher used\cite{BadPerc}.
-
-\item Implementing the pereptron throttler in a robust and accurate practical prefetching mechanism. 
-The overall scheme yields increased accuracy as compared to the other state of the art prefetchers.
-Moreover, the perceptron throttler learns to adapt itself to shared resource constraints - leading to an even increased performance in multi-core and lower capacity / bandwidth environments.
-
-\item Examining the perceptron weights doing a statistical study to quantify the contribution made by each feature.
-We have also used this idea to develop a comprehensive set of features where each feature contributes some `new' information to the perceptron.
-The details are explained under methodology, in Section XX.
-
-\item Cross-validating the design across benchmarks. 
-All the development and parameter tuning was done on the SPEC 2017 benchmark. 
-The results have been shown to be consistent across unseen benchmarks.
-More details can be found in the results section (Section XX).
-
-%\item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind. 
-%We have used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers. 
-%We have also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
-
-%\item We came up with an efficient model to use the perceptron output as a representation of the prefetch confidence using an on-line histogram technique. 
-%We show how the simple sum can replace the existing mechanism which was essentially based on counting the number of hits for a particular entry in SPP.
-
-\end{itemize}
-
-We tested SPP-Perc on Champsim, an open source simulator which models an Out-of-Order CPU using. 
-For evaluation, we used SPEC 2017 benchmarks. 
-On a single core configuration, SPP-Perc increases performance of the system by 39.5\% as compared to the baseline with no prefetching, on the memory intensive subset of the benchmark.
-This outperforms the next best prefetcher, SPP,  by 6.84\%. 
-On a multi-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement of 11.9\% over SPP for 4-core system and 10.67\% for 8-core system.
-
-In the results part in section XX, we show a detailed analysis of the speedup achieved.
-We also used SPEC 2006 and CloudSuite benchmarks for cross-validation of the prefetcher implementation.
-
-The rest of the paper is organized as follows.
-Section 2 talks about background work and recaptures the SPP architecture
-Section 3 discusses the detailed implementation of the perceptron mechanism and the features used.
-Section 4 consists of the evaluation methodology and explores the feature space for perceptron learning.
-Section 5 is the performance evaluation of our prefetcher and Section 6 concludes the paper.
-
-\section{Background Work and Motivation}
-
-In this section, we compare the existing work done in the domain of prefetching. 
-We also discuss about some of the interesting memory access patterns seen in the workloads in SPEC 2017 benchmark. 
-The second half of this section deals with a background of Signature Path Prefetcher architecture.
-
-\subsection{Related Work}
-
-The idea of prefetching has been around since Instruction Stream Buffers were proposed by Jouppi.
-\cite{ISB}
-Some of the earliest data prefetchers developed to try to identify memory access patterns with a constant stride pattern.
-\cite{Smith} 
-\cite{Baer}
-\cite{Stride}
-Major research was dedicated to stride distance and depth prediction.
-\cite{Decoupled}
-\cite{Adaptive}
-Newer prefetchers aimed at correlating predictions with the past memory access addresses.
-\cite{Address_Correlated} 
-\cite{AMPM}
-
-With time newer and more intricate prefetchers were introduced.
-\cite{Wenisch_Temporal_Streaming}
-\cite{Stealth}
-\cite{Feedback_Directed}
-\cite{Coordinated}
-\cite{Bandwidth_Efficient}
-\cite{Pacman}
-\cite{TLB}
-\cite{Linearizing}
-\cite{Sandbox}
-\cite{VLDP}
-\cite{DoL}
-\cite{Domino}
-A whole class of streaming prefetchers developed on the lines of Spatial / Temporal Memory Streaming.
-\cite{Spatial_Pattern}
-\cite{SMS}
-\cite{Temporal_Instruction_Fetch}
-\cite{Off_Chip}
-\cite{STMS}
-\cite{SMS_JILP}
-
-Most modern prefetchers aim at identifying complex memory access patterns in a given application. 
-This allows them to capture some of the irregular patterns seen in pointer chasing data-structures.
-Offset based prefetchers are a generalization of the next-line prefetcher.
-Some of the prominent prefetchers exploiting this idea are
-Sandbox Prefetcher \cite{Sandbox}
-and the Best Offset Prefetcher \cite{BOP}
-Another class of prefetchers is lookahead Prefetchers.
-Such prefetchers try to speculate deep into the application's memory access trail.
-These include Path Confidence based Lookahead Prefetching \cite{SPP}
-and Kill the Program Counter \cite{KPC}.
-
-A related line of work also explores cache replacement - insertion policies and dead block prediction, with an eye towards prefetching. 
-\cite{DB_Pred}
-\cite{Cache_Burst}
-\cite{KPC}
-\cite{Harmony}
-Nori \textit{et.al.} introduced prefetching with  respect to runtime criticality.
-\cite{CATCH}
-\newline
-\newline
-\textit{Machine Learning and Prefetching:}
-Peled \textit{et.al.} introduce interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
-\cite{Semantics}
-The biggest challenge here is that it relies on compiler support for getting features to build the context.
-
-The paper by Liao \textit{et. al.} focuses on prefetching for data center applications.
-\cite{Datacenter}
-They use offline Machine Learning Algorithms like SVMs and Logistic Regressions to do parametric search for an optimal prefetcher configuration.
-
-%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?
-Finally, the paper on Data Cache Prefetching with Perceptron Learning\cite{BadPerc} talks about the idea of two step prefetching.
-The first step is an existing baseline prefetcher while the second step is a perceptron based throttler.
-The paper has lots of limitations in terms of design and implementation.
-That reflects in the fact that it did not lead to any significant performance gain over even the basic prefetchers like the Stride Prefetcher\cite{Stride} and the Markov Prefetcher\cite{Markov}.
-
-\subsection{Perceptron Learning in Architecture}
-
-Perceptron learning for computer architecture design has been around for a while. 
-It was first popularized by Jimenez \textit{et. al.} for doing branch prediction. \cite{Perc_Branch} 
-In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
-The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. 
-The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the prediction is made as \textit{true}, or branch taken.
-Else if the sum is < 0, the prediction is made as \textit{false}.
-
-The structure that we use is derived from the model introduced in Piecewise Linear Branch Prediction\cite{Piece_Linear}. 
-Here the feature itself is used to index into a hash of perceptron weights. 
-This form of indexing makes sure that the hyper-plane learned by the perceptron weights is able to differentiate between linearly inseparable outcomes. 
-\textit{<Is it true? Not sure>}
-
-Perceptron Learning for Reuse Prediction\cite{Perc_Reuse} extended the hash perceptron architecture to perform prediction in context of cache replacement. 
-Perceptron layer can learn to correlate the cache replacement behaviour with a variety of features. 
-Each feature has its dedicated table of perceptron weights which it indexes into. 
-Hence number of tables equals the number of features. 
-Once the perceptron weights from all the tables are obtained, they're summed and thresholded based on a pre-decided value. 
-This led to a prediction mechanism that is highly accurate, has quick learning rate and is adaptable to changing program behaviour. 
-The features introduced in that paper were derived from Program Counters of the current and the last few instructions and the memory address of the current access. 
-Even with such basic features, high level of correlation could be established by perceptrons.
-
-The work was further extended by Multiperspective Reuse Prediction\cite{Multiperspective}, which introduced a varied set of parametric features. 
-Although design space exploration for feature and parameter selection was a non-trivial task, once fixed, it could give a cache replacement predictor highly suitable for the given set of applications.
-\newline
-\newline
-\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron prediction involves multiplying the vector of input features: F\textsubscript{1xN}, with the corresponding weight vector: W\textsubscript{Nx1} in a dot product fashion to obtain the sum y\textsubscript{out}.
-Here what we use is a perceptron-like structure. 
-The feature is used to hash into the weight of perceptrons and the retrieved weights are added straight away. 
-This can be seen a dot product of $\vec{1}$\textsubscript{1xN} with W\textsubscript{Nx1}. 
-This way, the perceptron algorithm doesn't introduce any multiplication operations in the inference or training process. 
-Hence what we adapt in this work, is a perceptron-like learning algorithm as it involves the same principle involved in perceptron inferencing and training. 
-For simplicity, we will call our implementation as perceptron implementation.
-\newline
-\newline
-\textit{Perceptron Update Rule:} In all the above implementations, a uniform perceptron update principle is being followed. 
-The weights need to be updated if the prediction was wrong or the predicted sum does not cross a certain threshold. 
-After a certain training period, these weights are proportional to the probability of the outcome being true. 
-This training threshold makes sure that the perceptrons are trained till a certain level of confidence and yet they are not over-trained to the given set. 
-Care is taken that the weights saturate at certain positive and negative values so that they remain confined to the bit-width. 
-The same rules will be applied in the discussion done in section XX on SPP-Perc learning algorithm.
-
-\subsection{Prefetching with SPEC 2017}
-\textit{<Discuss SPEC 2017 traces here>
-<What are the possible graphs / analysis tools we can use for SPEC traces?>}
-
-\subsection{Signature Path Prefetcher Architecture}
-Since the work we have done builds heavily on SPP\cite{SPP}, we are presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work. 
-Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists of following structural components:
-\newline
-\newline
-\textbf{Signature Table:} ST keeps track of 256 most recently accessed pages. 
-It is meant to capture memory access patterns within a page boundary. 
-SPP indexes into an entry of ST using the page number. 
-For each entry corresponding to a page, ST stores the `last block offset' and `old signature'. 
-Last block offset is the block offset of the last memory access of that given page. 
-The block offset is calculated with respect to the page boundary. 
-The signature is a 12-bit compressed representation of the past few memory accesses for that page. 
-The signature is calculated as:
-$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
-In case a matching page entry is found, the stored signature in the corresponding entry is retrieved.
-\newline
-\newline
-Consider the case that incoming page number 10 with a block offset 3 finds a match in the ST. 
-The retrieved pattern signature is 0x30 and the Last Offset is 1. 
-Since now there is the Last Block Offset (1), Incoming Block Offset (3), Old Signature (0x30), and New Signature calculated as per above equation (0x182), SPP can infer non-speculatively that the given pattern of memory accesses (as captured in the Old Signature) leads to the particular Delta.
-In general, Delta is defined as the difference between the prefetch suggestion block and the initial block which triggered the prefetch. 
-In this case, since SPP is in learning phase, it is defined as the difference between the Incoming Block Offset and the Last Block Offset (+2 in this case). 
-This, in turn generates the new memory pattern (New Signature). 
-This newly learned signature pattern and the delta is stored in the Pattern Table.
-\newline
-\newline
-\textbf{Pattern Table:} The PT is indexed by the signature generated from ST. 
-PT holds predicted delta patterns and their confidence estimates. 
-Each entry indexed by the signature holds up to 4 unique delta predictions. 
-This is implemented by making PT as a 4-way associative table. 
-To measure an estimate of a delta confidence, PT implements two counters: C\textsubscript{sig} for each entry indexed by the signature and C\textsubscript{delta} for each delta entry for a given signature.
-More details of this on-line score-keeping mechanism is described under Confidence Tracking discussion.
-
-\textit{Lookahead Prefetching:} For each trigger to the prefetcher, the first time that the SPP indexes into the PT, it does so in the non-speculative manner. 
-We call it non-speculative because the parameters which SPP uses to predict, like last block offset, incoming block offset and signature are well-defined. 
-Assuming PT generates a valid delta prediction (rather, up to 4 valid deltas), SPP enters into the lookahead mode.
-
-In the lookahead mode, SPP looks at the last generated delta(s), chooses the one with the highest confidence and speculatively generates the new effective address and the new signature. 
-Using these new parameters, it re-accesses the PT to generate further prefetches. 
-SPP keeps on repeating this cycle of accessing PT and updating the signature based on highest confidence prefetch from the last iteration. 
-These iteration count till which SPP manages to predict prefetch entries in the lookahead manner is characterized as its `depth'. 
-While doing this, SPP also keeps compounding the confidence in each depth. 
-Thus as depth increases, overall confidence keeps decreasing. 
-This makes sense since we are going deeper into the speculation path. 
-When the confidence falls below a pre-defined threshold, SPP stops further lookahead iterations
-\newline
-\newline
-\textbf{Global History Register:} One of the innovative ideas introduced in SPP was the Global History Register, which bootstraps learning across page boundaries.
-GHR is a small 8-entry fully associative structure. 
-Whenever a prefetch suggestion from the Pattern Table crosses page boundaries, instead of completely discarding it, GHR saves that information for inter-page learning.
-Since the SPP-Perc implementation in this paper doesn't affect GHR functionality in any ways, we won't be going into more details about GHR.
-\newline
-\newline
-\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch Filter - a 1024 entry 1-way associative table which keeps a record of last few entries prefetched by SPP. 
-This proves useful for following reasons:
-\begin{itemize}
-    \item \textbf{Eliminating duplicate prefetches}\newline 
-    Any subsequent prefetches recommended by SPP mechanism is checked against the existing entries in Prefetch Filter. 
-If a valid entry is found, it means that the particular cache line is already either in L2 Cache or being fetched. 
-Hence there is no need to issue a duplicate prefetch.
-
-    \item \textbf{Updating prefetcher}\newline
-    If a subsequent demand access comes before eviction for a prefetch that was recommended by SPP, various components of SPP are updated to reflect that this was a good prefetch. 
-Correspondingly, if a prefetched cache line is evicted without being demand hit even once, SPP learns that this was a useless prefetch and updates the various tables accordingly
-
-    \item \textbf{Tracking Global Accuracy}\newline
-    PF keeps a track of usefulness of the prefetches by introducing 2 counters: C\textsubscript{total} to track the total number of prefetches getting recorded in the PF and C\textsubscript{useful} to keep a track of which prefetches recorded in the PF led to a demand hit.
-The ratio C\textsubscript{useful} / C\textsubscript{total} is the global accuracy for the prefetching scheme. 
-It is represented as $\alpha$ in the paper. 
-The global accuracy acts a real-time throttler for the path confidence value as depicted in the following discussion.
-\newline
-\end{itemize}
-\textbf{Confidence Tracking}: The PT keeps a track of hits to each signature through a counter C\textsubscript{sig}. 
-The number of hits for a given delta per signature are tracked using a counter C\textsubscript{delta}. 
-The confidence for a given delta is approximated through C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}. 
-When SPP enters into a lookahead mode, the path confidence P\textsubscript{d} is given as:
-$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown in the Prefetch Filter section. 
-The range of global accuracy is in [0,1]. 
-The lower the value of $\alpha$ for a given trace or a program phase, the overall confidence is scaled down and this makes SPP less aggressive for the upcoming prefetches.
-
-Here d is the lookahead depth. 
-For d = 1, when SPP is in non-speculative mode, P\textsubscript{0} can be thought of as 1. 
-Once SPP enters into speculative mode, it calculates the path confidence as a product of confidence of the individual prefetches encountered along that path, further scaled using $\alpha$. 
-The final P\textsubscript{d} is thresholded against prefetching-threshold (T\textsubscript{p}) to decide whether to prefetch or not. 
-If yes, then P\textsubscript{d} is thresholded against a numerically bigger fill-threshold (T\textsubscript{f}) to decide whether the prefetch should be sent to L2 Cache (high confidence prefetch) or Last Level Cache (low confidence prefetch)
-\newline<Insert a comprehensive figure about SPP structures>
-\newline<Insert another detailed picture about SPP data-path flow>
-
-\subsection{Case for an On-line Throttler}
-As compared to some of the other state of the art prefetchers [BOP], SPP is a far less aggressive prefetcher. 
-In the single core environment, this gives BOP an edge as there is no resource contention among the different cores. 
-Hence more aggressive prefetching is bound to prove beneficial. 
-As we increase the core count, we observe that SPP starts outperforming rest of the prefetchers. 
-This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. 
-\textit{<Figure showing the variation of SPP vs BOP wrt core count>} 
-For 4-core applications, SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
-
-The above analysis shows that with a more careful throttling mechanism, SPP can be tuned to become much more aggressive, leading to increased coverage. 
-The onus of maintaining the accuracy now falls on the independent throttler. 
-To test the hypothesis, we tuned down SPP to the minimum possible threshold. 
-Only by doing this we managed to increase the prefetch suggestions made by SPP, by XX\%. 
-\textit{<Comparison of SPP-Unleased with SPP / BOP>}
-Obviously this came at a cost of increased DRAM traffic and cache pollution.
-
-Moreover, the on-line confidence mechanism introduced in SPP was very rudimentary.
-It was based on taking a ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig} as explained above.
-The same confidence was used to make the decision whether to prefetch or not to prefetch; and which level to prefetch.
-While this approximation was shown to work in the original SPP, we believe that a better form of on-line decision making was possible.
-Hence, it was necessary to build a robust and adaptable learning mechanism to accept / reject the prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level Cache).
-
-To that effect, we introduce an independent on-line perceptron based throttling mechanism.
-
 \section{SPP-Perc}
@@
 \begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline
-In the original SPP design, an internal confidence mechanism was developed which was compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the prefetch suggestion.
-Furthermore, the accepted suggestions were thresholded against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache vs Last Level Cache.
-T\textsubscript{p} and T\textsubscript{f} were set to 25 and 90 respectively, on the scale of 0 to 100.
-In SPP-Perc, the perceptron sum is used to make these decisions. 
-Hence these two thresholds are no longer needed
+\item \textbf{Original Thresholds discarded}\newline In the original
+  SPP design, an internal confidence mechanism was developed which was
+  compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to
+  decide whether to accept or reject the prefetch suggestion.
+  Furthermore, the accepted suggestions were thresholded against FILL
+  \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache
+  vs Last Level Cache.  T\textsubscript{p} and T\textsubscript{f} were
+  set to 25 and 90 respectively, on the scale of 0 to 100.  In
+  SPP-Perc, the perceptron sum is used to make these decisions.  Hence
+  these two thresholds are no longer needed
 
 \item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
-Original SPP would not consider the slots available in the L2 MSHR queue. 
-It had an assumption that at no time can the number of prefetches suggested by SPP could exceed the capacity of the queue. 
-While it was not proven in the paper, the internal confidence mechanism along with the thresholds made sure that the above assumption was maintained. 
-In SPP-Perc since there can be no assumption on the degree of aggressiveness - because SPP is being tuned to its lowest thresholds, we explicitly check that at no point should the number of suggested prefetches exceed L2 MSHR queue.
+  Original SPP would not consider the slots available in the L2 MSHR
+  queue.  It had an assumption that at no time can the number of
+  prefetches suggested by SPP could exceed the capacity of the queue.
+  While it was not proven in the paper, the internal confidence
+  mechanism along with the thresholds made sure that the above
+  assumption was maintained.  In SPP-Perc since there can be no
+  assumption on the degree of aggressiveness - because SPP is being
+  tuned to its lowest thresholds, we explicitly check that at no point
+  should the number of suggested prefetches exceed L2 MSHR queue.
 
-\item \textbf{Enhanced Filter}\newline
-Original SPP introduces a concept of Prefetch Filter, which keeps a track of up to 1024 L2 Cache prefetches suggested by SPP. 
-The Filter was modified to store meta-data like program counter, memory access address etc. 
-This is needed to train the perceptron when the result of that particular prefetch gets available. 
-At training time, this data is used to retrieve the state of the prefetcher when the prediction was made.
+\item \textbf{Enhanced Filter}\newline Original SPP introduces a
+  concept of Prefetch Filter, which keeps a track of up to 1024 L2
+  Cache prefetches suggested by SPP.  The Filter was modified to store
+  meta-data like program counter, memory access address etc.  This is
+  needed to train the perceptron when the result of that particular
+  prefetch gets available.  At training time, this data is used to
+  retrieve the state of the prefetcher when the prediction was made.
 
-\item \textbf{Reject Filter introduced}\newline
-The Prefetch Filter mentioned above keeps a track of the prefetches that actually happened in the L2 Cache. 
-Correspondingly, we introduced a reject filter. 
-It keeps a track of the prefetches that were suggested by the base SPP engine but were rejected by the perceptron. 
-In case it is detected that a future demand fetch could have used this rejected prefetch, the perceptron weights are updated to reflect the same
+\item \textbf{Reject Filter introduced}\newline The Prefetch Filter
+  mentioned above keeps a track of the prefetches that actually
+  happened in the L2 Cache.  Correspondingly, we introduced a reject
+  filter.  It keeps a track of the prefetches that were suggested by
+  the base SPP engine but were rejected by the perceptron.  In case it
+  is detected that a future demand fetch could have used this rejected
+  prefetch, the perceptron weights are updated to reflect the same
 
@@
 \subsection{The Perceptron Layer}
-As discussed earlier, the perceptron layer was introduced to act as a throttler to the underlying SPP engine. 
-This section explores the details of the perceptron layer and the range of features that are ultimately used by the perceptron for correlating the prefetches.
+As discussed earlier, the perceptron layer was introduced to act as a
+throttler to the underlying SPP engine.  This section explores the
+details of the perceptron layer and the range of features that are
+ultimately used by the perceptron for correlating the prefetches.
 
-The neural network weights are stored in separate tables for each feature.
-The indexing of each feature is done using different number of bits.
-At most 12 bits of a feature are used to index into the perceptron table.
-Certain features require more resolution power \textit{i.e.}, full 12 bits of indexing.
-On the other hand, some features requiring lower resolution required as low as 7 bits of indexing.
-The variable indexing was determined by studying the features and fine-tuned empirically so as to achieve a good accuracy vs area trade-off.
-Exact details are mentioned under `Area Overheads' in Section XX.
+The neural network weights are stored in separate tables for each
+feature.  The indexing of each feature is done using different number
+of bits.  At most 12 bits of a feature are used to index into the
+perceptron table.  Certain features require more resolution power
+\textit{i.e.}, full 12 bits of indexing.  On the other hand, some
+features requiring lower resolution required as low as 7 bits of
+indexing.  The variable indexing was determined by studying the
+features and fine-tuned empirically so as to achieve a good accuracy
+vs area trade-off.  Exact details are mentioned under `Area Overheads'
+in Section XX.
 
-A single entry in the table corresponds to a perceptron weight.
-Each weight is a 5 bit counter - saturating at -16 and +15. 
-Each feature gets its own dedicated table. 
-Our proposed design uses 9 features. 
+A single entry in the table corresponds to a perceptron weight.  Each
+weight is a 5 bit counter - saturating at -16 and +15.  Each feature
+gets its own dedicated table.  Our proposed design uses 9 features.
 As the program execution begins, all the weights are initialized to 0.
-\newline \newline
-\textbf{INFERENCING}\newline
-In the Champsim simulator, the prefetching method is triggered on every L2 Cache demand access. 
-At that point, the prefetcher has the option to do or not do prefetch - in case it does, then how many cache lines to prefetch. 
-The suggested prefetches can be places either in L2 Cache or in Last Level Cache. 
-The prefetcher gets to decide this, and is usually done based on an internal confidence mechanism.
 
-When SPP gets triggered, it suggests up to 4 (=Associativity of the Pattern Table) candidates for prefetch based on the demand access address. 
-In each subsequent lookahead depth, SPP suggests up to 4 more prefetch candidates, as explained in section\_\_. 
-All these recommended prefetches are tested in the perceptron throttler before they can access the Prefetch Filter, as in the original SPP data-path.
+\textbf{INFERENCING}\newline In the Champsim simulator, the
+prefetching method is triggered on every L2 Cache demand access.  At
+that point, the prefetcher has the option to do or not do prefetch -
+in case it does, then how many cache lines to prefetch.  The suggested
+prefetches can be places either in L2 Cache or in Last Level Cache.
+The prefetcher gets to decide this, and is usually done based on an
+internal confidence mechanism.
 
-The perceptron looks at the micro-architectural state, or `features', at that instant. 
-Each feature indexes into the hash of up to 4096-entry table dedicated for that feature. 
-Feature \#1 indexes into a particular entry in table \#1 and so on.
+When SPP gets triggered, it suggests up to 4 (=Associativity of the
+Pattern Table) candidates for prefetch based on the demand access
+address.  In each subsequent lookahead depth, SPP suggests up to 4
+more prefetch candidates, as explained in section\_\_.  All these
+recommended prefetches are tested in the perceptron throttler before
+they can access the Prefetch Filter, as in the original SPP data-path.
 
-Once all the weights are retrieved, they are added together, much like the traditional perceptron dot product. 
-On obtaining the overall sum, it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO. 
-Only the prefetch candidates with the perceptron sum higher that the threshold qualify for prefetching. 
-The prefetches that qualify the perceptron throttler stage go through the Prefetch Filter stage, just as in the SPP design.
-The Prefetch Filter eliminates any redundant prefetch suggestions.
-In SPP-Perc, all the meta-data that is required to recreate the features is also stored in the Prefetch Filter. 
-At a later stage, when the feedback of the current prefetched line is available, the stored data can be used to train the perceptron.
+The perceptron looks at the micro-architectural state, or `features',
+at that instant.  Each feature indexes into the hash of up to
+4096-entry table dedicated for that feature.  Feature \#1 indexes into
+a particular entry in table \#1 and so on.
 
-The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch cache line. 
-All prefetch candidates qualified till this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill level.
-The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
+Once all the weights are retrieved, they are added together, much like
+the traditional perceptron dot product.  On obtaining the overall sum,
+it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO.
+Only the prefetch candidates with the perceptron sum higher that the
+threshold qualify for prefetching.  The prefetches that qualify the
+perceptron throttler stage go through the Prefetch Filter stage, just
+as in the SPP design.  The Prefetch Filter eliminates any redundant
+prefetch suggestions.  In SPP-Perc, all the meta-data that is required
+to recreate the features is also stored in the Prefetch Filter.  At a
+later stage, when the feedback of the current prefetched line is
+available, the stored data can be used to train the perceptron.
 
-The perceptron sum can also be seen as the sum of the individual contribution per feature.
-The value of each contribution corresponds to the extent of confidence of the final output with that given feature.
-By summing the individual contributions, the final perceptron sum denotes the overall confidence for that prefetch suggestion. 
-In a way, by thresholding the perceptron sum on two different thresholds, we are diving the confidence scale of the prefetch suggestion into three bins.
-The first bin corresponding to the lowest confidence leads to prefetch candidate being rejected.
-The next bin corresponds to prefetches with a moderate confidence level. 
-Such prefetches are directed to fill the bigger Last Level Cache and potentially not pollute the more scarce L2 Cache.
-The highest confidence prefetches are filled in the L2 Cache.
+The above generated perceptron sum is also used to decide L2 Cache vs
+Last Level Cache placement of the prefetch cache line.  All prefetch
+candidates qualified till this stage are thresholded against
+PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
+PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
 
-In addition to the Prefetch Filter mentioned above, SPP-Perc also maintains a `Reject Filter'. 
-Reject filter is a 1024-entry deep 1-way associative table. 
-In case that a prefetch suggestion is rejected by perceptron throttler, it instead gets logged into the Reject Filter. 
-The Reject Filter keeps track of 1024 entries that were suggested by the Pattern Table but got rejected by the perceptron throttler. 
-This can be used to train the perceptron to avoid false negatives \textit{i.e.}, cases where prediction was to reject the prefetch but the prefetch would have been useful.
+The perceptron sum can also be seen as the sum of the individual
+contribution per feature.  The value of each contribution corresponds
+to the extent of confidence of the final output with that given
+feature.  By summing the individual contributions, the final
+perceptron sum denotes the overall confidence for that prefetch
+suggestion.  In a way, by thresholding the perceptron sum on two
+different thresholds, we are diving the confidence scale of the
+prefetch suggestion into three bins.  The first bin corresponding to
+the lowest confidence leads to prefetch candidate being rejected.  The
+next bin corresponds to prefetches with a moderate confidence level.
+Such prefetches are directed to fill the bigger Last Level Cache and
+potentially not pollute the more scarce L2 Cache.  The highest
+confidence prefetches are filled in the L2 Cache.
 
-As previously stated, an enhancement to the original SPP design includes storing all the meta-data describing the program state at the instant of prefetching, in the Prefetch Filter. 
-This information is useful when the perceptron needs to be updated subsequently. 
-\newline \newline
-\textbf{TRAINING}\newline
-In the prefetching environment, feedback of a prefetch is received whenever there is an eviction or a demand access from L2 Cache. 
-This acts as a trigger to update all the baseline SPP entries and to also train the perceptron throttler. 
-Training of the perceptron throttler involves restoring the meta-data that was stored in Prefetch Filter or the Reject Filter. 
-This is done by indexing into the corresponding filters using the cache line address of the block being used to train. (10-bits indexing + 6-bits tag matching). 
-Once the state of the program at the time of prefetching is available, it is used to index into the perceptron weights table.
+In addition to the Prefetch Filter mentioned above, SPP-Perc also
+maintains a `Reject Filter'.  Reject filter is a 1024-entry deep 1-way
+associative table.  In case that a prefetch suggestion is rejected by
+perceptron throttler, it instead gets logged into the Reject Filter.
+The Reject Filter keeps track of 1024 entries that were suggested by
+the Pattern Table but got rejected by the perceptron throttler.  This
+can be used to train the perceptron to avoid false negatives
+\textit{i.e.}, cases where prediction was to reject the prefetch but
+the prefetch would have been useful.
 
-If a demand access block which triggers the training  was tagged as a valid prefetch in the Prefetch Filter, it means that the earlier prefetch prediction was correct. 
-In that case the perceptron weights are incremented by 1 if the predicted sum does not cross a pre-defined threshold. 
- In case it is a cache block eviction that led to training and the corresponding valid entry was found in the Prefetch Filter, it shows that the prediction made by the perceptron was wrong. 
-Perceptron should have ideally rejected the prefetch suggestion as a low-confidence prefetch. 
-Here the weights are decremented by 1 to reflect the same. 
-In either case, weights are saturated at -16 or +15.
+As previously stated, an enhancement to the original SPP design
+includes storing all the meta-data describing the program state at the
+instant of prefetching, in the Prefetch Filter.  This information is
+useful when the perceptron needs to be updated subsequently.
+
+\textbf{TRAINING}\newline In the prefetching environment, feedback of
+a prefetch is received whenever there is an eviction or a demand
+access from L2 Cache.  This acts as a trigger to update all the
+baseline SPP entries and to also train the perceptron throttler.
+Training of the perceptron throttler involves restoring the meta-data
+that was stored in Prefetch Filter or the Reject Filter.  This is done
+by indexing into the corresponding filters using the cache line
+address of the block being used to train. (10-bits indexing + 6-bits
+tag matching).  Once the state of the program at the time of
+prefetching is available, it is used to index into the perceptron
+weights table.
+
+If a demand access block which triggers the training was tagged as a
+valid prefetch in the Prefetch Filter, it means that the earlier
+prefetch prediction was correct.  In that case the perceptron weights
+are incremented by 1 if the predicted sum does not cross a pre-defined
+threshold.  In case it is a cache block eviction that led to training
+and the corresponding valid entry was found in the Prefetch Filter, it
+shows that the prediction made by the perceptron was wrong.
+Perceptron should have ideally rejected the prefetch suggestion as a
+low-confidence prefetch.  Here the weights are decremented by 1 to
+reflect the same.  In either case, weights are saturated at -16 or
++15.
 
 A secondary training mechanism also kicks in during demand fetches.
-Before the demand access triggers the next set of prefetches, it checks the Reject Filter for the valid entry. 
-In case it is a hit, that means that the corresponding cache line was initially suggested by the underlying SPP engine but rejected by the perceptron throttler. 
-This means that perceptron should have been more confident about that particular prefetch. 
-Once such a scenario is identified, the state of the execution at the time of prefetch is retrieved from the Reject Table. 
-The retrieved data is used to index into the various weights tables of the perceptron and the corresponding values are updated by +1, saturating between and -16 and +15. 
-This update reflects increased confidence for the prediction corresponding to that prefetch.
+Before the demand access triggers the next set of prefetches, it
+checks the Reject Filter for the valid entry.  In case it is a hit,
+that means that the corresponding cache line was initially suggested
+by the underlying SPP engine but rejected by the perceptron throttler.
+This means that perceptron should have been more confident about that
+particular prefetch.  Once such a scenario is identified, the state of
+the execution at the time of prefetch is retrieved from the Reject
+Table.  The retrieved data is used to index into the various weights
+tables of the perceptron and the corresponding values are updated by
++1, saturating between and -16 and +15.  This update reflects
+increased confidence for the prediction corresponding to that
+prefetch.
 
-This mechanism allows us to solve the classical problem of exploiting `lost opportunity'. 
-In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
-Our two-step SPP-Perc allows us to overcome that issue in a simple manner.
+This mechanism allows us to solve the classical problem of exploiting
+`lost opportunity'.  In prior perceptron based implementations (and in
+general, prefetching algorithms), there is usually no way of knowing
+the result of not prefetch a particular line.  Our two-step SPP-Perc
+allows us to overcome that issue in a simple manner.
 
@@
 
-Here we discuss the various features which help correlate the prefetching decision with the program behaviour
+Here we discuss the various features which help correlate the
+prefetching decision with the program behaviour
 \begin{itemize}
-\item \textbf{Base address} of the demand access which triggered the prefetch. 
-Since prefetch are triggered from the L2 demand access, they tend to be correlated with the triggering address.
+\item \textbf{Base address} of the demand access which triggered the
+  prefetch.  Since prefetch are triggered from the L2 demand access,
+  they tend to be correlated with the triggering address.
 
-\item \textbf{Cache line} and \textbf{Page address}: These two separate features are derived from the base address which triggered the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and base\_addr >> LOG2\_PAGE\_ SIZE respectively.
-The idea behind using three different shifted versions of the same feature is that it allows us to look into a wider range of bits than with a single version.
-It also helps give more importance to the overlapping bits and lesser importance to most and least significance bits.
-This approach can also eliminate destructive interference that can be caused by directly folding the bits into half.
+\item \textbf{Cache line} and \textbf{Page address}: These two
+  separate features are derived from the base address which triggered
+  the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and
+  base\_addr >> LOG2\_PAGE\_ SIZE respectively.  The idea behind using
+  three different shifted versions of the same feature is that it
+  allows us to look into a wider range of bits than with a single
+  version.  It also helps give more importance to the overlapping bits
+  and lesser importance to most and least significance bits.  This
+  approach can also eliminate destructive interference that can be
+  caused by directly folding the bits into half.
 
-\item \textbf{Confidence}: The integer confidence on the scale of 0 to 100 which was used in the original SPP design.
-While the original confidence is not used directly for decision making, it can still contribute to the final decision made.
+\item \textbf{Confidence}: The integer confidence on the scale of 0 to
+  100 which was used in the original SPP design.  While the original
+  confidence is not used directly for decision making, it can still
+  contribute to the final decision made.
 
-\item \textbf{Program Counter XOR Depth}: PC is for the instruction which triggered the prefetch chain.
-Depth refers to the iteration count of the lookahead stages.
-In general, PC is not considered as a good basis for doing lookahead prefetching as all the prefetches with depth >= 1 gets aliased onto the same PC which won't be true in actual demand fetch case.
-This feature resolves a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases.
-This is akin to the concept of Virtual Program Counters\cite{VPC} introduced by Kim \textit{et. al.} for indirect branch prediction.
+\item \textbf{Program Counter XOR Depth}: PC is for the instruction
+  which triggered the prefetch chain.  Depth refers to the iteration
+  count of the lookahead stages.  In general, PC is not considered as
+  a good basis for doing lookahead prefetching as all the prefetches
+  with depth >= 1 gets aliased onto the same PC which won't be true in
+  actual demand fetch case.  This feature resolves a PC into a
+  different value for each lookahead depth of SPP speculation, giving
+  a more accurate correlation in lookahead cases.  This is akin to the
+  concept of Virtual Program Counters\cite{VPC} introduced by Kim
+  \textit{et. al.} for indirect branch prediction.
 
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC before the instruction that triggered the current prefetch. 
-Hashing together the last three PCs tell SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block.
-PCs are shifted in the increasing order of history before being hashed together.
-This is done to avoid the resultant value of zero in case 2 or more PCs are the same. 
-Additionally, obfuscating the information as it gets old allows us to get a wider and an approximate look into the program history.
+\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
+    PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$
+  PC before the instruction that triggered the current prefetch.
+  Hashing together the last three PCs tell SPP-Perc about the path
+  that led to the current demand access and helps capture and
+  branching information of the current basic block.  PCs are shifted
+  in the increasing order of history before being hashed together.
+  This is done to avoid the resultant value of zero in case 2 or more
+  PCs are the same.  Additionally, obfuscating the information as it
+  gets old allows us to get a wider and an approximate look into the
+  program history.
 
-\item \textbf{Program Counter XOR Delta}: This tells us if a given PC favours particular value(s) of delta.
-As noted earlier, while PC in itself does not much useful information, this hash resolves PC into different values based on the tendency of that PC to favour a certain delta.
-This way, dynamic nature of different instances of the same memory access instruction can be captured here.
+\item \textbf{Program Counter XOR Delta}: This tells us if a given PC
+  favours particular value(s) of delta.  As noted earlier, while PC in
+  itself does not much useful information, this hash resolves PC into
+  different values based on the tendency of that PC to favour a
+  certain delta.  This way, dynamic nature of different instances of
+  the same memory access instruction can be captured here.
 
-\item \textbf{Current Signature XOR Delta}: It can be recalled from the discussion of SPP in section XX that the new signature is generated using the old signature and the blobk delta.
-While `Current Signature XOR Delta' is not the exact formula for generating the future signature, it gives an approximate idea of the path that the combination of these two values can lead to.
+\item \textbf{Current Signature XOR Delta}: It can be recalled from
+  the discussion of SPP in section XX that the new signature is
+  generated using the old signature and the blobk delta.  While
+  `Current Signature XOR Delta' is not the exact formula for
+  generating the future signature, it gives an approximate idea of the
+  path that the combination of these two values can lead to.
 
 %%%%%%% Below are the features rejected when downsizing 14 -> 9
-%\item \textbf{Current Signature}: The 12-bit signature used by SPP to index into the Pattern Table
-%\item \textbf{Delta}: The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched
-%\item \textbf{PC}: Program Counter of the Load instruction that triggers that particular prefetch
-%\item \textbf{Base address XOR Delta}: If any particular value(s) of delta are highly favoured by a given address which triggered the prefetch, this composite feature can capture that information
-%\item \textbf{Cache line XOR Delta}: This feature captures if any particular delta is highly favoured by a given cache line address
-%\item \textbf{Confidence XOR Delta}: This gives a combined insight into that fact that certain delta-confidence pairs predicted by SPP might correlate more with the prefetch outcome
+% \item \textbf{Current Signature}: The 12-bit signature used by SPP
+%   to index into the Pattern Table
+% \item \textbf{Delta}: The signed version of the difference in the
+%   block address that triggers the prefetch and the block address of
+%   the data being prefetched
+% \item \textbf{PC}: Program Counter of the Load instruction that
+%   triggers that particular prefetch
+% \item \textbf{Base address XOR Delta}: If any particular value(s) of
+%   delta are highly favoured by a given address which triggered the
+%   prefetch, this composite feature can capture that information
+% \item \textbf{Cache line XOR Delta}: qThis feature captures if any
+%   particular delta is highly favoured by a given cache line address
+%\item \textbf{Confidence XOR Delta}: This gives a combined insight
+%  into that fact that certain delta-confidence pairqs predicted by
+%  SPP might correlate more with the prefetch outcome
 %%%%%%%%%%%%%%%%%%%%%%
@@
 
-Some of the composite features are derive from simple hashing (XOR) of two primary features.
-There is always a question of usefulness of such composite features and the `fresh' information added.
-We try to justify the choice of each feature by quantifying the contribution made towards predicting prefetch behaviour, in Section XX.
-Finally, as noted above, each feature indexes into its independent entry of perceptron weights.
-
-\section{METHODOLOGY}
-\subsection{Performance Model}
-For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim} simulator. 
-ChampSim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} 
-We model model a 1-core / 4-core / 8-core Out of Order machine. The details of the configuration parameters is summarized in Table  \ref{tab:Sim_params}.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|l|p{3.6cm}|c|}
-    \hline
-    Level & Configuration & Access Latency \\
-    \hline
-         L1 Cache & Separate I-cache and D-cache, 32 KB, 8-way & 4 cycles\\
-         L2 Cache & Private, 256 KB & 8 cycles\\
-         LLC & Shared, 2MB / core, 16-way & 20 cycles\\
-         DRAM & 4 GB Single Channel for single-core, 8 GB Double Channel for multi-core & N?? cycles\\
-    \hline
-    \end{tabular}
-    \caption{Memory Model Parameters}
-    \label{tab:Sim_params}
-\end{table}
-
-The block size was fixed 64KB. Prefetching was only triggered at L2 Cache demand accesses but could be sent to L2 Cache or Last Level Cache. 
-No L1 Data level prefetching was done. 
-LRU replacement policy was used on all levels of cache hierarchies. 
-Branch prediction was done using the perceptron branch predictor\cite{Perc_Branch}. 
-Page size was kept as 4KB. 
-Champsim operates all the prefetchers strictly in the physical address space.
-
-\subsection{Testing Under Additional Memory Constraints}
-The default single-core configuration simulates a 2MB LLC and a single channel DRAM with 12.8GB/s bandwidth. 
-We extend the simulations to include memory constraints introduced in DPC-2. 
-Specifically we look at the following two variations:
-\begin{itemize}
-    \item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited to 3.2 GB/s
-    \item \textit{Small LLC}: In this scenario, LLC size is reduced to 512 KB
-\end{itemize}
-All the multi-core simulations are only done in the default configuration.
-
-\subsection{Workloads}
-This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017} has been used to characterize and measure the prefetch performance. 
-We use all the 20 workloads available in the SPEC 2017 Suite. 
-Using SimPoint\cite{SimPoint} methodology, we identified 95 different program segments of 1 Billion instructions each.
-
-We performed cross-validation of our SPP-Perc model using SPEC 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.
-For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
-For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.
-The traces include 4 multi-core applications with 6 distinct phases per application. 
-In total, we used 285 traces representing workloads across 53 applications. 
-\newline
-\newline
-\textit{Single-core performance:} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics.
-We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
-We take the Instruction Per Cycle (IPC) of the prefetcher and divide it with the IPC of baseline. 
-The final number reported is the geometric mean of the speed-up attained on individual traces.
-\newline
-\newline
-\textit{Multi-core performance:} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
-For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 
-Each CPU keeps executing its workload till the last CPU completes 1 Billion instructions after warm-up. 
-For collecting IPC and other data, only the first billion instructions are considered as the ``region of interest''.
-
-Here we report the weighted speedup normalized to baseline \textit{i.e.}, no prefetching. 
-For each of the workloads running on a particular core of the 4-core 8 MB LLC system, we compute IPC\textsubscript{i}. 
-We then find the IPC\_isolated\textsubscript{i} of the same workload running in isolated 1-core 8 MB LLC environment. 
-Then we calculate the total weighted-IPC for a given workload mix as $\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}). 
-For each of the 100 workload-mix, the sum obtained is normalized to the weighted-IPC calculated similarly for baseline case \textit{i.e.}, no prefetching, to get the weighted-IPC-speedup. 
-Finally the geometric mean of these 100 weighted-IPC-speedup is reported as the effective speedup obtained by the prefetching scheme.
-
-We repeat the same process for 8-core workloads with the exception that 20 Million warm-up instructions and 100 Million full instructions were executed.
-This was done so as to keep the simulation run-time within reasonable limits as a single 8-core mix was taking up to 5 days to simulate 1 Billion instructions.
-
-\subsection{Preferchers Simulated}
-We compared SPP-Perc against three of the latest state of the art prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher (SPP).
-BOP was the winner of 2nd Data Prefetching Championship. 
-DA-AMPM is the enhanced version of AMPM, modified to account for DRAM row buffer locality. 
-SPP has been shown to outperform BOP on SPEC 2006 traces. 
-For each of these, we compare their speedups taking no prefetching as the baseline.
-
-\subsection{Developing Features for SPP-Perc}
-This section describes the intuition and analysis that went behind finalizing the features. 
-As noted earlier, we developed a set of 9 features which allow the perceptron throttler to correlate prefetching decision with the program context. 
-To study the correlation across each feature, we study statistically the perceptron weights and try to interpret their distribution.
-\newline  \newline
-\textbf{GLOBAL PEARSON'S CORRELATION}\newline
-For this experiment, we create a dump of the state of perceptron weights at the end of all trace execution.
-The weights obtained from running all the SPEC 2017 traces are concatenated together. 
-Since the weights are collected at the end of individual trace execution, the perceptron weights have attained a relatively stable value by now.
-Hence, this dump represents a `snap' of the trained perceptron weights across all the SPEC 2017 traces.
-The intuition here is that the feature with bulk of the perceptron weights concentrated around 0 or small magnitude numbers show a weak correlation with the prefetching outcome. 
-On the other hand, features with most of the weights saturated around highest value (+15) show a high positive correlation and the features with weights close to the lowest value (-16) show a strong negative correlation.
-
-We plot a histogram for each feature depicting weights distribution from -16 to +15 and use this histogram to generate the Pearson's correlation factor for that feature.
-In statistics, Pearson's factor is a numerical measure of the degree of linear correlation between two variables.
-It ranges from -1 to +1.
-The magnitude of Pearson's factor depicts the extent of correlation and the sign depicts whether it is a positive correlation or a negative correlation.
-Values close to 0 suggest a low correlation or at times, noisy data.
-A value of +1/-1 suggests a perfectly linear positive / negative correlation respectively.
-Figure XX depicts the histogram and the Pearson's coefficient for two of the features, YY and ZZ.
-
-Figure XX shows all the features used, arranged in the increasing order of their Pearson's factor.
-As can be seen X out of the 9 features provide a moderate to high correlation, with the magnitude of P-value > 0.4.
-The single most important feature, \_\_ helps provide a correlation to prefetch outcome with a factor of \_\_.
-\newline \newline
-\textbf{PER TRACE CORRELATION} \newline
-Another important way to look at the perceptron features is to see how much their contribution varies across the traces.
-Here we give special attention to features with low P-values in the previous experiment.
-Figure XX shows the variation of P-values three features : \_\_, \_\_ and \_\_; across all the SPEC 2017 traces.
-For simplicity, the traces are arranged in an increasing order of contribution made by the feature and the trace names have been omitted.
-It can be seen that even features with a low overall correlation provide useful correlation (magnitude > 0.4) for XX out of the 95 simpoints developed for SPEC 2017.
-\newline \newline
-\textbf{TRIMMING FEATURES USING CORRELATION} \newline
-Besides providing interesting insights into prefetching behaviour, P-value can also be used for feature selection and prefetcher tuning.
-
-Here, we introduce the concept of cross-correlation across features.
-Just as we examined correlation of each feature with the final outcome, we can also study correlation between the features.
-We used the above methodology to eliminate features which provide a little or no information that has already been captured in other features.
-
-As a part of feature selection, we initially came up a mix of 23 features- primary and composite.
-By studying cross correlation of each of these features against others in a 23x23 matrix, we identified pairs of features with correlation factor > 0.9 in magnitude and carefully eliminated redundant features.
-Using this approach, we managed to reduce the feature count to 9.
-Thus, in the final implementation of SPP-Perc, no two features have a high correlation between them.
-This way we can be sure that each feature is bringing in contribution that cannot be captured using other features.
-
-\textit{<TODO: VALIDATE THIS>} 
-Secondly, studying the relative importance of each feature enabled us to vary the number of entries dedicated for each feature.
-Features with higher Pearson's correlation (like \_\_ and \_\_) were given most importance and allowed full 12-bits of indexing.
-Features like \_\_ and \_\_, with a low overall P-value were allocated fewer indexes in the feature table.
-
-To conclude, in the above discussion we justify the features for perceptron from a statistical viewpoint.
-We also show how this information can also be used for prefetcher tuning.
-All this study was made possible only because we used on-line perceptron learning for prefetching and that enabled us to examine the weights in detail.
-
-\subsection{Overhead for SPP-Perc}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|c|c|m{4.8cm}|}
-    \hline
-        \textbf{Field} &
-        \textbf{Bits} &
-        \textbf{Comment} \\
-    \hline
-         Valid & 1 & Indicates a valid entry in the filter\\
-         Tag & 6 & Identifier for the entry in the filter\\
-         Useful & 1 & To show if the given entry led to a useful demand fetch\\
-         Perc Decision & 1 & Prefetched vs Not-prefetched \\
-    \hline
-        PC & 12 & \multirow{5}{4.8cm}{Meta-data required for perceptron training}\\
-        Address & 24 & \\
-        Delta & 7 & \\
-        Curr Signature & 12 & \\
-        Confidence & 7 & \\
-    \hline
-        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
-    \hline
-    \end{tabular}
-    \caption{Meta-data Stored in Prefetch Filter}
-    \label{tab:PF_metadata}
-\end{table}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|p{1.5cm}|p{1cm}|c|p{1.5cm}|}
-    \hline
-        \textbf{Structure} &
-        \textbf{Entry} &
-        \textbf{Components} &
-        \textbf{Total} \\
-    \hline
-        \multirow{5}{1.5cm}{Signature Table} & \multirow{5}{1.5cm}{256} & Valid (1 bit) & \multirow{5}{1cm}{11008  bits}\\
-        & & Tag (16 bits) &\\
-        & & Last Offset (6 bits) &\\  
-        & & Signature (12 bits) &\\
-        & & LRU (6 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Pattern Table} & \multirow{3}{1.5cm}{512} & $C_{sig}$ (4bits) &\multirow{3}{1cm}{24576 bits}\\
-        & & $C_{delta}$ (4*4 bits) &\\
-        & & Delta (4*7 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Perceptron Tables} & \multirow{3}{1.5cm}{4096*8 128*1} & \multirow{3}{1.5cm}{5 bits} & \multirow{3}{1.5cm}{164480 bits}\\
-        & & &\\
-        & & &\\
-    \hline
-        \multirow{2}{1.5cm}{Prefetch Filter \footnotemark[1]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{84 bits} & \multirow{2}{1.5cm}{86016 bits}\\
-        & & &\\
-    \hline
-     \multirow{2}{1.5cm}{Reject Filter \footnotemark[2]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{83 bits} & \multirow{2}{1.5cm}{84992 bits}\\
-        & & &\\
-    \hline
-        \multirow{4}{1.5cm}{Global History Register} & \multirow{4}{1.5cm}{8} & Signature (12 bits) & \multirow{4}{1.5cm}{264 bits}\\
-        & & Confidence (8 bits) & \\
-        & & Last Offset (6 bits) & \\
-        & & Delta (7 bits) & \\
-    \hline
-        \multirow{2}{1.5cm}{Accuracy Counters} & 1 & C$_{total}$ & 10 bits\\
-        & 1 & C$_{useful}$ & 10 bits\\
-    \hline
-        \multirow{3}{1.5cm}{Global PC Trackers} & \multirow{3}{1cm}{3} & $PC_1$ (12 bits)& \multirow{3}{1.cm}{36 bits}\\
-        & & $PC_2$ (12 bits) &\\
-        & & $PC_3$ (12 bits) &\\
-    \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 371356 bits = 45.33 KB}}\\
-    \hline
-    \end{tabular}
-    \caption{SPP-Perc Storage Overhead}
-    \label{tab:SPPPerc_overhead}
-\end{table}
-\footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}
-\footnotetext[2]{RF does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
-
-In this section, we analyze the hardware overhead required to implement SPP-Perc. 
-The Prefetch Filter was enhanced to accommodate storing of meta-data for perceptron training. 
-Table \ref{tab:PF_metadata} depicts the meta-data stored for each entry in the Prefetch Filter.
-Table \ref{tab:SPPPerc_overhead} shows the total storage overhead of SPP-Perc implementation.
-The hardware budget for 2nd Data Prefetching championship was 32 KB. 
-Keeping that in mind and the speedup SPP-Perc obtained over the winner, the extra hardware budget can be accounted for.
-The extra hardware also makes the overall scheme more scalable than SPP. 
-In the original SPP paper, it was demonstrated that adding extra hardware brings little advantage in terms of performance gain.
-The newly added perceptron tables can be scaled to increase / decrease features depending on the permitted budget.
-
-In terms of computations, the perceptron mechanism only introduces an extra adder tree.
-The hash perceptron mechanism makes sure than there is no actual vector multiplication happening in the hardware.
-Obtaining the perceptron sum requires addition of 9 5-bit numbers. 
-Using an adder tree of 4 5-bit adders, this can be done in ceil($log_{2}9$) = 4 steps.
-Perceptron update only requires weight update by +1 or -1.
-Thus, all the operations required for perceptron inferencing or updating the states of the perceptron throttler can be easily done in the time constraints of L2 Cache Accesses.
-
-\section{RESULTS}
-This section discusses the results obtained from running SPP-Perc, in terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per Cycle (IPC) metrics. 
-For SPEC 2017 suite, first we present the results for single-threaded workloads then the results for multi-core workloads are presented.
-
-\subsection{Single-core Results}
-Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated. 
-All the results have been normalized to baseline \textit{i.e.}, no prefetching. 
-As can be seen, SPP-Perc yields a speedup of XX\% over the baseline. 
-This is equivalent to XX\% over DA-AMPM, xx\% over BOP and XX\% over SPP. 
-Out of the XX benchmarks, SPP-Perc nearly matches or outperforms most of the prefetchers on XX traces.
-
-At its peak, SPP-Perc manages to get the IPC speedup of a factor of over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}. 
-This also corresponds to speedup gain of \textbf{22.4\%} over the next best prefetcher - SPP. 
-In general, benchmarks \textit{602.gcc\_s, 603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s} benefit the most from SPP-Perc, as compared to SPP only prefetching.
-
-<Discuss the extreme behaviour of \textit{623.xalancbmk\_s}: Improvement over SPP is as high as 31\% and even reduces to -1.5\% on traces of this benchmark>
-
-<Discuss about coverage on all traces>
-
-<Discuss cases where coverage gets specially increased>
-
-<Discuss accuracy on all traces>
-
-<Discuss about compelling increase in accuracy - details of traces. Plot linear separability>
-
-<Any case where accuracy got reduced - go into the trace. See if linear separability is the issue. Other possible issues?>
-
-\subsection{Multi-core Results}
-In this section, we demonstrate the improvement achieved by SPP-Perc for a mix of multi-programmed workloads. 
-\newline \newline
-\textit{4-core environment}
-Figure \_\_ shows a comparison of speedups obtained on workload mix of memory intensive subset of SPEC 2017. 
-We plot all the 4 prefetchers, normalized to baseline. 
-The workloads have been sorted in the increasing order of performance benefit, as measured in terms of normalized weighted speedup. 
-As can be seen, SPP-Perc offers an IPC improvement by 33.5\% on these traces. 
-This is 5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP and DA-AMPM respectively.
-
-This increased improvement by SPP-Perc on a 4-core machine is not surprising. 
-Since SPP-Perc has an extremely intelligent throttler working to eliminate useless prefetches before they can create cache pollution. 
-This further amplifies the edge that SPP had over BOP. 
-With the advantage of an orthogonal prefetcher, SPP can be as aggressive as possible and leave the accuracy aspect to perceptron throttler.
-\newline \newline
-\textit{8-core environment}
-To test the effectiveness of SPP-Perc in a tighter constrained multi-core environment, we simulated an 8-core machine with 16 MB last Level Cache. 
-Other parameters were kept same as discussed under Methodology in Section XX.
-<See if we can use speed-up over BOP as a measure here>
-
-\subsection{Additional Memory Constraints}
-Figure \_\_ shows the performance of SPP-Perc against the other prefetchers in tighter constraint environments. 
-For sake of simplicity, we show 2 traces where SPP-Perc gives the most advantage over baseline, 2 traces where SPP-Perc doesn't perform as well as the best prefetcher *<-replace with the prefetcher name once results are available* and finally GeoMean across all the traces.
-<Hopefully BOP would degrade more here as it is super aggressive. 
-By definition, SPP-Perc should be better than SPP as the throttler would better control the aggressiveness. 
-In SPP paper DA-AMPM performs close enough to BOP. That should get replicated here too.>
-
-\subsection{Cross Validation}
-Here we try to demonstrate the robustness of SPP-Perc by testing it on completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
-
-Fig \_\_ shows the speed-up achieved by BOP, SPP and SPP-Perc on selected traces of SPEC 2006, along with the overall geometric mean over whole of the benchmark and over memory-intensive subset of the SPEC 2006 benchmark.
-As can be seen, SPP-Perc peovides a speed-up of 44.4\% over baseline, on the memory intensive subset of SPEC 2006 benchmark.
-This corresponds to 6.84\% over SPP and XX\% over BOP.
-On the whole of the SPEC 2006 suite, the speedup obtained is 22.4\%, which is 3.34\% ahead of SPP.
-
-The 4-core results for SPEC 2006 are \_\_.
-
-As can be seen, the performance is very consistent given no prior parameter tuning was done keeping SPEC 2006 in mind. 
-We attribute this to the inherent adaptability of the perceptron model.
-In general, perceptron weights are able to adjust in real-time so as to find the best possible correlation between the output and the features.
-
-\section{CONCLUSION}
-In this paper, we introduced perceptron based learning for data prefetching in the form of SPP-Perc. 
-Perceptron acts an orthogonal throttler to the underlying prefetch engine. 
-SPP-Perc improves performance by over XX\% over the baseline, which corresponds to YY\% over the next best performing prefetcher. 
-We believe this paper can lead to interesting future work like generalizing SPP-Perc to work with any prefetcher with minimal knowledge of the underlying prefetcher architecture.
-
-%%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%
-
-%%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
-\bibliographystyle{ieeetr}
-\bibliography{ref}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
+Some of the composite features are derive from simple hashing (XOR) of
+two primary features.  There is always a question of usefulness of
+such composite features and the `fresh' information added.  We try to
+justify the choice of each feature by quantifying the contribution
+made towards predicting prefetch behaviour, in Section XX.  Finally,
+as noted above, each feature indexes into its independent entry of
+perceptron weights.

### 2018-11-15 intro.tex
reorg

--- 2018-11-15 paper_1.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is the template for submission to ISCA 2018
-% The cls file is a modified from  'sig-alternate.cls'
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+Since past few decades, processor and memory scaling has been enabled
+with different philosophies in mind.  While processor scaling focused
+on speed improvements, memory scaling was centred around increasing
+the storage capacity.  The process technology reaching scaling
+limitations led to Memory Wall\cite{MemWall} - the increasing
+performance gap between the processor speed and the memory speeds.
+Various techniques have been developed to combat this widening gap.
+Data prefetching is one such important technique.  Data prefetching
+exploits the fact that in most applications, memory accesses have a
+well-defined pattern.
 
-\documentclass{sig-alternate} 
-\usepackage{mathptmx} % This is Times font
+An ideal prefetching scheme involves capturing this memory access
+patterns to predict future prefetches in a timely manner.  Other
+important decisions faced by the prefetcher are when to bring in the
+prefetch block of data, which level of cache hierarchy to place the
+block in and finally, what block to evict to accommodate the incoming
+block.  Memory accesses patterns can be as simple as fetching the next
+cache line or a complex but predictable pattern, like in pointer
+chasing applications.  Predicting this often requires a trade-off
+between coverage and accuracy.  For a prefetcher to capture highly
+irregular access patterns, it needs to be highly aggressive (high
+coverage).  That also means the prefetcher would end up recommending a
+lot of useless prefetches, leading to cache pollution (low accuracy).
+This effect would get even more pronounced in multi-core scenario
+where the last level cache is a shared resource.  Conversely, a
+conservative prefetcher would be highly accurate but might not make
+enough useful predictions.
 
-\newcommand{\ignore}[1]{}
-\usepackage{fancyhdr}
-\usepackage[normalem]{ulem}
-\usepackage[hyphens]{url}
-\usepackage{microtype}
-\usepackage{fixltx2e}
-\usepackage{array}
-\usepackage{multirow}
+To this effect we propose SPP-Perc <come up with a catchy name>.  It
+is an enhancement to the existing state of the art prefetcher and
+overcomes the coverage vs accuracy trade-off in a robust manner.  The
+idea is to modify SPP design to make it as aggressive as it can get.
+This enables SPP to capture complex memory access patterns hence
+leading to increased coverage.  This comes at a cost of increased DRAM
+traffic, hence reduction in overall IPC.  The prefetch suggestions
+recommended by SPP pass through the perceptron based throttler.  Over
+time, perceptron layer learns to correlate the prefetches recommended
+by SPP with various available features, leading to elimination of
+useless prefetches.  This leads to an overall increased accuracy of
+the prefetch scheme and reduction in DRAM traffic.
 
-% Always include hyperref last
-\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=black,citecolor=blue,urlcolor=black]{hyperref}
-
-% Ensure letter paper
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\newcommand{\iscasubmissionnumber}{NaN}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\fancypagestyle{firstpage}{
-  \fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-  \fancyhead[C]{\normalsize{}} 
-  \fancyfoot[C]{\thepage}
-}
-
-\pagenumbering{arabic}
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\title{SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher} 
-\author{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\maketitle
-\thispagestyle{firstpage}
-\pagestyle{plain}
-
-\begin{abstract}
-\textbf{Hardware prefetching has been introduced in modern processors as a way to hide cache latencies. 
-An efficient prefetcher should be able to identify complex memory access patterns during a program execution. 
-This enables it to fetch a block ahead of its demand access, potentially saving a cache miss.
-}
-
-\textbf{In this paper, we introduce perceptron learning to help make this prefetching decision efficiently. 
-We have taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine. 
-The perceptron layer acts as a throttler to filter out the unnecessary prefetches recommended by SPP. 
-We have also explored a range of features that can be used to train the perceptron layer. 
-Our results show that SPP-Perc improves performance on the memory intensive subset of the SPEC 2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared to just SPP.
-We also demonstrate that the performance gained from using our efficient throttler scales better with increasing number of cores.
-}
-\end{abstract}
-
-\section{Introduction}
-Since past few decades, processor and memory scaling has been enabled with different philosophies in mind. 
-While processor scaling focused on speed improvements, memory scaling was centred around increasing the storage capacity. 
-The process technology reaching scaling limitations led to Memory Wall\cite{MemWall} - the increasing performance gap between the processor speed and the memory speeds. 
-Various techniques have been developed to combat this widening gap. 
-Data prefetching is one such important technique. 
-Data prefetching exploits the fact that in most applications, memory accesses have a well-defined pattern.
-
-An ideal prefetching scheme involves capturing this memory access patterns to predict future prefetches in a timely manner. 
-Other important decisions faced by the prefetcher are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block.
-Memory accesses patterns can be as simple as fetching the next cache line or a complex but predictable pattern, like in pointer chasing applications. 
-Predicting this often requires a trade-off between coverage and accuracy. 
-For a prefetcher to capture highly irregular access patterns, it needs to be highly aggressive (high coverage). 
-That also means the prefetcher would end up recommending a lot of useless prefetches, leading to cache pollution (low accuracy). 
-This effect would get even more pronounced in multi-core scenario where the last level cache is a shared resource. 
-Conversely, a conservative prefetcher would be highly accurate but might not make enough useful predictions.
-
-To this effect we propose SPP-Perc <come up with a catchy name>. 
-It is an enhancement to the existing state of the art prefetcher and overcomes the coverage vs accuracy trade-off in a robust manner. 
-The idea is to modify SPP design to make it as aggressive as it can get. 
-This enables SPP to capture complex memory access patterns hence leading to increased coverage.
-This comes at a cost of increased DRAM traffic, hence reduction in overall IPC.
-The prefetch suggestions recommended by SPP pass through the perceptron based throttler. 
-Over time, perceptron layer learns to correlate the prefetches recommended by SPP with various available features, leading to elimination of useless prefetches. 
-This leads to an overall increased accuracy of the prefetch scheme and reduction in DRAM traffic.
-
-\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and future scope for expansion. 
-The contributions made by this paper are:
+\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and
+future scope for expansion.  The contributions made by this paper are:
 
 \begin{itemize}
-\item This is the first time where a truly on-line learning based neural network model has been used when it comes to hardware data prefetching. 
-The prior works in this area either relied on program semantics\cite{Semantics} or were application specific\cite{Datacenter} or could not yield a significant improvement over the baseline prefetcher used\cite{BadPerc}.
+\item This is the first time where a truly on-line learning based
+  neural network model has been used when it comes to hardware data
+  prefetching.  The prior works in this area either relied on program
+  semantics\cite{Semantics} or were application
+  specific\cite{Datacenter} or could not yield a significant
+  improvement over the baseline prefetcher used\cite{BadPerc}.
 
-\item Implementing the pereptron throttler in a robust and accurate practical prefetching mechanism. 
-The overall scheme yields increased accuracy as compared to the other state of the art prefetchers.
-Moreover, the perceptron throttler learns to adapt itself to shared resource constraints - leading to an even increased performance in multi-core and lower capacity / bandwidth environments.
+\item Implementing the pereptron throttler in a robust and accurate
+  practical prefetching mechanism.  The overall scheme yields
+  increased accuracy as compared to the other state of the art
+  prefetchers.  Moreover, the perceptron throttler learns to adapt
+  itself to shared resource constraints - leading to an even increased
+  performance in multi-core and lower capacity / bandwidth
+  environments.
 
-\item Examining the perceptron weights doing a statistical study to quantify the contribution made by each feature.
-We have also used this idea to develop a comprehensive set of features where each feature contributes some `new' information to the perceptron.
-The details are explained under methodology, in Section XX.
+\item Examining the perceptron weights doing a statistical study to
+  quantify the contribution made by each feature.  We have also used
+  this idea to develop a comprehensive set of features where each
+  feature contributes some `new' information to the perceptron.  The
+  details are explained under methodology, in Section XX.
 
-\item Cross-validating the design across benchmarks. 
-All the development and parameter tuning was done on the SPEC 2017 benchmark. 
-The results have been shown to be consistent across unseen benchmarks.
-More details can be found in the results section (Section XX).
+\item Cross-validating the design across benchmarks.  All the
+  development and parameter tuning was done on the SPEC 2017
+  benchmark.  The results have been shown to be consistent across
+  unseen benchmarks.  More details can be found in the results section
+  (Section XX).
 
-%\item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind. 
-%We have used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers. 
-%We have also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
+  % \item Designing the prefetcher keeping SPEC 2017 benchmark suite
+  %   in mind.  We have used SPEC 2017 to compare SPP-Perc against the
+  %   highest performing prefetchers.  We have also used our
+  %   prefetcher to characterize memory intensive benchmarks of the
+  %   SPEC 2017 suite, with respect to prefetching.
 
-%\item We came up with an efficient model to use the perceptron output as a representation of the prefetch confidence using an on-line histogram technique. 
-%We show how the simple sum can replace the existing mechanism which was essentially based on counting the number of hits for a particular entry in SPP.
+  % \item We came up with an efficient model to use the perceptron
+  %   output as a representation of the prefetch confidence using an
+  %   on-line histogram technique.  We show how the simple sum can
+  %   replace the existing mechanism which was essentially based on
+  %   counting the number of hits for a particular entry in SPP.
 
@@
 
-We tested SPP-Perc on Champsim, an open source simulator which models an Out-of-Order CPU using. 
-For evaluation, we used SPEC 2017 benchmarks. 
-On a single core configuration, SPP-Perc increases performance of the system by 39.5\% as compared to the baseline with no prefetching, on the memory intensive subset of the benchmark.
-This outperforms the next best prefetcher, SPP,  by 6.84\%. 
-On a multi-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement of 11.9\% over SPP for 4-core system and 10.67\% for 8-core system.
+We tested SPP-Perc on Champsim, an open source simulator which models
+an Out-of-Order CPU using.  For evaluation, we used SPEC 2017
+benchmarks.  On a single core configuration, SPP-Perc increases
+performance of the system by 39.5\% as compared to the baseline with
+no prefetching, on the memory intensive subset of the benchmark.  This
+outperforms the next best prefetcher, SPP, by 6.84\%.  On a multi-core
+system running a mix of memory intensive subset of the SPEC 2017
+traces, SPP-Perc saw an improvement of 11.9\% over SPP for 4-core
+system and 10.67\% for 8-core system.
 
-In the results part in section XX, we show a detailed analysis of the speedup achieved.
-We also used SPEC 2006 and CloudSuite benchmarks for cross-validation of the prefetcher implementation.
+In the results part in section XX, we show a detailed analysis of the
+speedup achieved.  We also used SPEC 2006 and CloudSuite benchmarks
+for cross-validation of the prefetcher implementation.
 
-The rest of the paper is organized as follows.
-Section 2 talks about background work and recaptures the SPP architecture
-Section 3 discusses the detailed implementation of the perceptron mechanism and the features used.
-Section 4 consists of the evaluation methodology and explores the feature space for perceptron learning.
-Section 5 is the performance evaluation of our prefetcher and Section 6 concludes the paper.
-
-\section{Background Work and Motivation}
-
-In this section, we compare the existing work done in the domain of prefetching. 
-We also discuss about some of the interesting memory access patterns seen in the workloads in SPEC 2017 benchmark. 
-The second half of this section deals with a background of Signature Path Prefetcher architecture.
-
-\subsection{Related Work}
-
-The idea of prefetching has been around since Instruction Stream Buffers were proposed by Jouppi.
-\cite{ISB}
-Some of the earliest data prefetchers developed to try to identify memory access patterns with a constant stride pattern.
-\cite{Smith} 
-\cite{Baer}
-\cite{Stride}
-Major research was dedicated to stride distance and depth prediction.
-\cite{Decoupled}
-\cite{Adaptive}
-Newer prefetchers aimed at correlating predictions with the past memory access addresses.
-\cite{Address_Correlated} 
-\cite{AMPM}
-
-With time newer and more intricate prefetchers were introduced.
-\cite{Wenisch_Temporal_Streaming}
-\cite{Stealth}
-\cite{Feedback_Directed}
-\cite{Coordinated}
-\cite{Bandwidth_Efficient}
-\cite{Pacman}
-\cite{TLB}
-\cite{Linearizing}
-\cite{Sandbox}
-\cite{VLDP}
-\cite{DoL}
-\cite{Domino}
-A whole class of streaming prefetchers developed on the lines of Spatial / Temporal Memory Streaming.
-\cite{Spatial_Pattern}
-\cite{SMS}
-\cite{Temporal_Instruction_Fetch}
-\cite{Off_Chip}
-\cite{STMS}
-\cite{SMS_JILP}
-
-Most modern prefetchers aim at identifying complex memory access patterns in a given application. 
-This allows them to capture some of the irregular patterns seen in pointer chasing data-structures.
-Offset based prefetchers are a generalization of the next-line prefetcher.
-Some of the prominent prefetchers exploiting this idea are
-Sandbox Prefetcher \cite{Sandbox}
-and the Best Offset Prefetcher \cite{BOP}
-Another class of prefetchers is lookahead Prefetchers.
-Such prefetchers try to speculate deep into the application's memory access trail.
-These include Path Confidence based Lookahead Prefetching \cite{SPP}
-and Kill the Program Counter \cite{KPC}.
-
-A related line of work also explores cache replacement - insertion policies and dead block prediction, with an eye towards prefetching. 
-\cite{DB_Pred}
-\cite{Cache_Burst}
-\cite{KPC}
-\cite{Harmony}
-Nori \textit{et.al.} introduced prefetching with  respect to runtime criticality.
-\cite{CATCH}
-\newline
-\newline
-\textit{Machine Learning and Prefetching:}
-Peled \textit{et.al.} introduce interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
-\cite{Semantics}
-The biggest challenge here is that it relies on compiler support for getting features to build the context.
-
-The paper by Liao \textit{et. al.} focuses on prefetching for data center applications.
-\cite{Datacenter}
-They use offline Machine Learning Algorithms like SVMs and Logistic Regressions to do parametric search for an optimal prefetcher configuration.
-
-%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?
-Finally, the paper on Data Cache Prefetching with Perceptron Learning\cite{BadPerc} talks about the idea of two step prefetching.
-The first step is an existing baseline prefetcher while the second step is a perceptron based throttler.
-The paper has lots of limitations in terms of design and implementation.
-That reflects in the fact that it did not lead to any significant performance gain over even the basic prefetchers like the Stride Prefetcher\cite{Stride} and the Markov Prefetcher\cite{Markov}.
-
-\subsection{Perceptron Learning in Architecture}
-
-Perceptron learning for computer architecture design has been around for a while. 
-It was first popularized by Jimenez \textit{et. al.} for doing branch prediction. \cite{Perc_Branch} 
-In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
-The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. 
-The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the prediction is made as \textit{true}, or branch taken.
-Else if the sum is < 0, the prediction is made as \textit{false}.
-
-The structure that we use is derived from the model introduced in Piecewise Linear Branch Prediction\cite{Piece_Linear}. 
-Here the feature itself is used to index into a hash of perceptron weights. 
-This form of indexing makes sure that the hyper-plane learned by the perceptron weights is able to differentiate between linearly inseparable outcomes. 
-\textit{<Is it true? Not sure>}
-
-Perceptron Learning for Reuse Prediction\cite{Perc_Reuse} extended the hash perceptron architecture to perform prediction in context of cache replacement. 
-Perceptron layer can learn to correlate the cache replacement behaviour with a variety of features. 
-Each feature has its dedicated table of perceptron weights which it indexes into. 
-Hence number of tables equals the number of features. 
-Once the perceptron weights from all the tables are obtained, they're summed and thresholded based on a pre-decided value. 
-This led to a prediction mechanism that is highly accurate, has quick learning rate and is adaptable to changing program behaviour. 
-The features introduced in that paper were derived from Program Counters of the current and the last few instructions and the memory address of the current access. 
-Even with such basic features, high level of correlation could be established by perceptrons.
-
-The work was further extended by Multiperspective Reuse Prediction\cite{Multiperspective}, which introduced a varied set of parametric features. 
-Although design space exploration for feature and parameter selection was a non-trivial task, once fixed, it could give a cache replacement predictor highly suitable for the given set of applications.
-\newline
-\newline
-\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron prediction involves multiplying the vector of input features: F\textsubscript{1xN}, with the corresponding weight vector: W\textsubscript{Nx1} in a dot product fashion to obtain the sum y\textsubscript{out}.
-Here what we use is a perceptron-like structure. 
-The feature is used to hash into the weight of perceptrons and the retrieved weights are added straight away. 
-This can be seen a dot product of $\vec{1}$\textsubscript{1xN} with W\textsubscript{Nx1}. 
-This way, the perceptron algorithm doesn't introduce any multiplication operations in the inference or training process. 
-Hence what we adapt in this work, is a perceptron-like learning algorithm as it involves the same principle involved in perceptron inferencing and training. 
-For simplicity, we will call our implementation as perceptron implementation.
-\newline
-\newline
-\textit{Perceptron Update Rule:} In all the above implementations, a uniform perceptron update principle is being followed. 
-The weights need to be updated if the prediction was wrong or the predicted sum does not cross a certain threshold. 
-After a certain training period, these weights are proportional to the probability of the outcome being true. 
-This training threshold makes sure that the perceptrons are trained till a certain level of confidence and yet they are not over-trained to the given set. 
-Care is taken that the weights saturate at certain positive and negative values so that they remain confined to the bit-width. 
-The same rules will be applied in the discussion done in section XX on SPP-Perc learning algorithm.
-
-\subsection{Prefetching with SPEC 2017}
-\textit{<Discuss SPEC 2017 traces here>
-<What are the possible graphs / analysis tools we can use for SPEC traces?>}
-
-\subsection{Signature Path Prefetcher Architecture}
-Since the work we have done builds heavily on SPP\cite{SPP}, we are presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work. 
-Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists of following structural components:
-\newline
-\newline
-\textbf{Signature Table:} ST keeps track of 256 most recently accessed pages. 
-It is meant to capture memory access patterns within a page boundary. 
-SPP indexes into an entry of ST using the page number. 
-For each entry corresponding to a page, ST stores the `last block offset' and `old signature'. 
-Last block offset is the block offset of the last memory access of that given page. 
-The block offset is calculated with respect to the page boundary. 
-The signature is a 12-bit compressed representation of the past few memory accesses for that page. 
-The signature is calculated as:
-$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
-In case a matching page entry is found, the stored signature in the corresponding entry is retrieved.
-\newline
-\newline
-Consider the case that incoming page number 10 with a block offset 3 finds a match in the ST. 
-The retrieved pattern signature is 0x30 and the Last Offset is 1. 
-Since now there is the Last Block Offset (1), Incoming Block Offset (3), Old Signature (0x30), and New Signature calculated as per above equation (0x182), SPP can infer non-speculatively that the given pattern of memory accesses (as captured in the Old Signature) leads to the particular Delta.
-In general, Delta is defined as the difference between the prefetch suggestion block and the initial block which triggered the prefetch. 
-In this case, since SPP is in learning phase, it is defined as the difference between the Incoming Block Offset and the Last Block Offset (+2 in this case). 
-This, in turn generates the new memory pattern (New Signature). 
-This newly learned signature pattern and the delta is stored in the Pattern Table.
-\newline
-\newline
-\textbf{Pattern Table:} The PT is indexed by the signature generated from ST. 
-PT holds predicted delta patterns and their confidence estimates. 
-Each entry indexed by the signature holds up to 4 unique delta predictions. 
-This is implemented by making PT as a 4-way associative table. 
-To measure an estimate of a delta confidence, PT implements two counters: C\textsubscript{sig} for each entry indexed by the signature and C\textsubscript{delta} for each delta entry for a given signature.
-More details of this on-line score-keeping mechanism is described under Confidence Tracking discussion.
-
-\textit{Lookahead Prefetching:} For each trigger to the prefetcher, the first time that the SPP indexes into the PT, it does so in the non-speculative manner. 
-We call it non-speculative because the parameters which SPP uses to predict, like last block offset, incoming block offset and signature are well-defined. 
-Assuming PT generates a valid delta prediction (rather, up to 4 valid deltas), SPP enters into the lookahead mode.
-
-In the lookahead mode, SPP looks at the last generated delta(s), chooses the one with the highest confidence and speculatively generates the new effective address and the new signature. 
-Using these new parameters, it re-accesses the PT to generate further prefetches. 
-SPP keeps on repeating this cycle of accessing PT and updating the signature based on highest confidence prefetch from the last iteration. 
-These iteration count till which SPP manages to predict prefetch entries in the lookahead manner is characterized as its `depth'. 
-While doing this, SPP also keeps compounding the confidence in each depth. 
-Thus as depth increases, overall confidence keeps decreasing. 
-This makes sense since we are going deeper into the speculation path. 
-When the confidence falls below a pre-defined threshold, SPP stops further lookahead iterations
-\newline
-\newline
-\textbf{Global History Register:} One of the innovative ideas introduced in SPP was the Global History Register, which bootstraps learning across page boundaries.
-GHR is a small 8-entry fully associative structure. 
-Whenever a prefetch suggestion from the Pattern Table crosses page boundaries, instead of completely discarding it, GHR saves that information for inter-page learning.
-Since the SPP-Perc implementation in this paper doesn't affect GHR functionality in any ways, we won't be going into more details about GHR.
-\newline
-\newline
-\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch Filter - a 1024 entry 1-way associative table which keeps a record of last few entries prefetched by SPP. 
-This proves useful for following reasons:
-\begin{itemize}
-    \item \textbf{Eliminating duplicate prefetches}\newline 
-    Any subsequent prefetches recommended by SPP mechanism is checked against the existing entries in Prefetch Filter. 
-If a valid entry is found, it means that the particular cache line is already either in L2 Cache or being fetched. 
-Hence there is no need to issue a duplicate prefetch.
-
-    \item \textbf{Updating prefetcher}\newline
-    If a subsequent demand access comes before eviction for a prefetch that was recommended by SPP, various components of SPP are updated to reflect that this was a good prefetch. 
-Correspondingly, if a prefetched cache line is evicted without being demand hit even once, SPP learns that this was a useless prefetch and updates the various tables accordingly
-
-    \item \textbf{Tracking Global Accuracy}\newline
-    PF keeps a track of usefulness of the prefetches by introducing 2 counters: C\textsubscript{total} to track the total number of prefetches getting recorded in the PF and C\textsubscript{useful} to keep a track of which prefetches recorded in the PF led to a demand hit.
-The ratio C\textsubscript{useful} / C\textsubscript{total} is the global accuracy for the prefetching scheme. 
-It is represented as $\alpha$ in the paper. 
-The global accuracy acts a real-time throttler for the path confidence value as depicted in the following discussion.
-\newline
-\end{itemize}
-\textbf{Confidence Tracking}: The PT keeps a track of hits to each signature through a counter C\textsubscript{sig}. 
-The number of hits for a given delta per signature are tracked using a counter C\textsubscript{delta}. 
-The confidence for a given delta is approximated through C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}. 
-When SPP enters into a lookahead mode, the path confidence P\textsubscript{d} is given as:
-$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown in the Prefetch Filter section. 
-The range of global accuracy is in [0,1]. 
-The lower the value of $\alpha$ for a given trace or a program phase, the overall confidence is scaled down and this makes SPP less aggressive for the upcoming prefetches.
-
-Here d is the lookahead depth. 
-For d = 1, when SPP is in non-speculative mode, P\textsubscript{0} can be thought of as 1. 
-Once SPP enters into speculative mode, it calculates the path confidence as a product of confidence of the individual prefetches encountered along that path, further scaled using $\alpha$. 
-The final P\textsubscript{d} is thresholded against prefetching-threshold (T\textsubscript{p}) to decide whether to prefetch or not. 
-If yes, then P\textsubscript{d} is thresholded against a numerically bigger fill-threshold (T\textsubscript{f}) to decide whether the prefetch should be sent to L2 Cache (high confidence prefetch) or Last Level Cache (low confidence prefetch)
-\newline<Insert a comprehensive figure about SPP structures>
-\newline<Insert another detailed picture about SPP data-path flow>
-
-\subsection{Case for an On-line Throttler}
-As compared to some of the other state of the art prefetchers [BOP], SPP is a far less aggressive prefetcher. 
-In the single core environment, this gives BOP an edge as there is no resource contention among the different cores. 
-Hence more aggressive prefetching is bound to prove beneficial. 
-As we increase the core count, we observe that SPP starts outperforming rest of the prefetchers. 
-This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. 
-\textit{<Figure showing the variation of SPP vs BOP wrt core count>} 
-For 4-core applications, SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
-
-The above analysis shows that with a more careful throttling mechanism, SPP can be tuned to become much more aggressive, leading to increased coverage. 
-The onus of maintaining the accuracy now falls on the independent throttler. 
-To test the hypothesis, we tuned down SPP to the minimum possible threshold. 
-Only by doing this we managed to increase the prefetch suggestions made by SPP, by XX\%. 
-\textit{<Comparison of SPP-Unleased with SPP / BOP>}
-Obviously this came at a cost of increased DRAM traffic and cache pollution.
-
-Moreover, the on-line confidence mechanism introduced in SPP was very rudimentary.
-It was based on taking a ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig} as explained above.
-The same confidence was used to make the decision whether to prefetch or not to prefetch; and which level to prefetch.
-While this approximation was shown to work in the original SPP, we believe that a better form of on-line decision making was possible.
-Hence, it was necessary to build a robust and adaptable learning mechanism to accept / reject the prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level Cache).
-
-To that effect, we introduce an independent on-line perceptron based throttling mechanism.
-
-\section{SPP-Perc}
-
-\subsection{Changes made to original SPP}
-To modify the SPP Design to suit our scheme, the following changes were made:
-
-\begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline
-In the original SPP design, an internal confidence mechanism was developed which was compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the prefetch suggestion.
-Furthermore, the accepted suggestions were thresholded against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache vs Last Level Cache.
-T\textsubscript{p} and T\textsubscript{f} were set to 25 and 90 respectively, on the scale of 0 to 100.
-In SPP-Perc, the perceptron sum is used to make these decisions. 
-Hence these two thresholds are no longer needed
-
-\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
-Original SPP would not consider the slots available in the L2 MSHR queue. 
-It had an assumption that at no time can the number of prefetches suggested by SPP could exceed the capacity of the queue. 
-While it was not proven in the paper, the internal confidence mechanism along with the thresholds made sure that the above assumption was maintained. 
-In SPP-Perc since there can be no assumption on the degree of aggressiveness - because SPP is being tuned to its lowest thresholds, we explicitly check that at no point should the number of suggested prefetches exceed L2 MSHR queue.
-
-\item \textbf{Enhanced Filter}\newline
-Original SPP introduces a concept of Prefetch Filter, which keeps a track of up to 1024 L2 Cache prefetches suggested by SPP. 
-The Filter was modified to store meta-data like program counter, memory access address etc. 
-This is needed to train the perceptron when the result of that particular prefetch gets available. 
-At training time, this data is used to retrieve the state of the prefetcher when the prediction was made.
-
-\item \textbf{Reject Filter introduced}\newline
-The Prefetch Filter mentioned above keeps a track of the prefetches that actually happened in the L2 Cache. 
-Correspondingly, we introduced a reject filter. 
-It keeps a track of the prefetches that were suggested by the base SPP engine but were rejected by the perceptron. 
-In case it is detected that a future demand fetch could have used this rejected prefetch, the perceptron weights are updated to reflect the same
-
-\end{itemize}
-
-\subsection{The Perceptron Layer}
-As discussed earlier, the perceptron layer was introduced to act as a throttler to the underlying SPP engine. 
-This section explores the details of the perceptron layer and the range of features that are ultimately used by the perceptron for correlating the prefetches.
-
-The neural network weights are stored in separate tables for each feature.
-The indexing of each feature is done using different number of bits.
-At most 12 bits of a feature are used to index into the perceptron table.
-Certain features require more resolution power \textit{i.e.}, full 12 bits of indexing.
-On the other hand, some features requiring lower resolution required as low as 7 bits of indexing.
-The variable indexing was determined by studying the features and fine-tuned empirically so as to achieve a good accuracy vs area trade-off.
-Exact details are mentioned under `Area Overheads' in Section XX.
-
-A single entry in the table corresponds to a perceptron weight.
-Each weight is a 5 bit counter - saturating at -16 and +15. 
-Each feature gets its own dedicated table. 
-Our proposed design uses 9 features. 
-As the program execution begins, all the weights are initialized to 0.
-\newline \newline
-\textbf{INFERENCING}\newline
-In the Champsim simulator, the prefetching method is triggered on every L2 Cache demand access. 
-At that point, the prefetcher has the option to do or not do prefetch - in case it does, then how many cache lines to prefetch. 
-The suggested prefetches can be places either in L2 Cache or in Last Level Cache. 
-The prefetcher gets to decide this, and is usually done based on an internal confidence mechanism.
-
-When SPP gets triggered, it suggests up to 4 (=Associativity of the Pattern Table) candidates for prefetch based on the demand access address. 
-In each subsequent lookahead depth, SPP suggests up to 4 more prefetch candidates, as explained in section\_\_. 
-All these recommended prefetches are tested in the perceptron throttler before they can access the Prefetch Filter, as in the original SPP data-path.
-
-The perceptron looks at the micro-architectural state, or `features', at that instant. 
-Each feature indexes into the hash of up to 4096-entry table dedicated for that feature. 
-Feature \#1 indexes into a particular entry in table \#1 and so on.
-
-Once all the weights are retrieved, they are added together, much like the traditional perceptron dot product. 
-On obtaining the overall sum, it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO. 
-Only the prefetch candidates with the perceptron sum higher that the threshold qualify for prefetching. 
-The prefetches that qualify the perceptron throttler stage go through the Prefetch Filter stage, just as in the SPP design.
-The Prefetch Filter eliminates any redundant prefetch suggestions.
-In SPP-Perc, all the meta-data that is required to recreate the features is also stored in the Prefetch Filter. 
-At a later stage, when the feedback of the current prefetched line is available, the stored data can be used to train the perceptron.
-
-The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch cache line. 
-All prefetch candidates qualified till this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill level.
-The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
-
-The perceptron sum can also be seen as the sum of the individual contribution per feature.
-The value of each contribution corresponds to the extent of confidence of the final output with that given feature.
-By summing the individual contributions, the final perceptron sum denotes the overall confidence for that prefetch suggestion. 
-In a way, by thresholding the perceptron sum on two different thresholds, we are diving the confidence scale of the prefetch suggestion into three bins.
-The first bin corresponding to the lowest confidence leads to prefetch candidate being rejected.
-The next bin corresponds to prefetches with a moderate confidence level. 
-Such prefetches are directed to fill the bigger Last Level Cache and potentially not pollute the more scarce L2 Cache.
-The highest confidence prefetches are filled in the L2 Cache.
-
-In addition to the Prefetch Filter mentioned above, SPP-Perc also maintains a `Reject Filter'. 
-Reject filter is a 1024-entry deep 1-way associative table. 
-In case that a prefetch suggestion is rejected by perceptron throttler, it instead gets logged into the Reject Filter. 
-The Reject Filter keeps track of 1024 entries that were suggested by the Pattern Table but got rejected by the perceptron throttler. 
-This can be used to train the perceptron to avoid false negatives \textit{i.e.}, cases where prediction was to reject the prefetch but the prefetch would have been useful.
-
-As previously stated, an enhancement to the original SPP design includes storing all the meta-data describing the program state at the instant of prefetching, in the Prefetch Filter. 
-This information is useful when the perceptron needs to be updated subsequently. 
-\newline \newline
-\textbf{TRAINING}\newline
-In the prefetching environment, feedback of a prefetch is received whenever there is an eviction or a demand access from L2 Cache. 
-This acts as a trigger to update all the baseline SPP entries and to also train the perceptron throttler. 
-Training of the perceptron throttler involves restoring the meta-data that was stored in Prefetch Filter or the Reject Filter. 
-This is done by indexing into the corresponding filters using the cache line address of the block being used to train. (10-bits indexing + 6-bits tag matching). 
-Once the state of the program at the time of prefetching is available, it is used to index into the perceptron weights table.
-
-If a demand access block which triggers the training  was tagged as a valid prefetch in the Prefetch Filter, it means that the earlier prefetch prediction was correct. 
-In that case the perceptron weights are incremented by 1 if the predicted sum does not cross a pre-defined threshold. 
- In case it is a cache block eviction that led to training and the corresponding valid entry was found in the Prefetch Filter, it shows that the prediction made by the perceptron was wrong. 
-Perceptron should have ideally rejected the prefetch suggestion as a low-confidence prefetch. 
-Here the weights are decremented by 1 to reflect the same. 
-In either case, weights are saturated at -16 or +15.
-
-A secondary training mechanism also kicks in during demand fetches. 
-Before the demand access triggers the next set of prefetches, it checks the Reject Filter for the valid entry. 
-In case it is a hit, that means that the corresponding cache line was initially suggested by the underlying SPP engine but rejected by the perceptron throttler. 
-This means that perceptron should have been more confident about that particular prefetch. 
-Once such a scenario is identified, the state of the execution at the time of prefetch is retrieved from the Reject Table. 
-The retrieved data is used to index into the various weights tables of the perceptron and the corresponding values are updated by +1, saturating between and -16 and +15. 
-This update reflects increased confidence for the prediction corresponding to that prefetch.
-
-This mechanism allows us to solve the classical problem of exploiting `lost opportunity'. 
-In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
-Our two-step SPP-Perc allows us to overcome that issue in a simple manner.
-
-\subsection{Features used by Perceptron}
-
-Here we discuss the various features which help correlate the prefetching decision with the program behaviour
-\begin{itemize}
-\item \textbf{Base address} of the demand access which triggered the prefetch. 
-Since prefetch are triggered from the L2 demand access, they tend to be correlated with the triggering address.
-
-\item \textbf{Cache line} and \textbf{Page address}: These two separate features are derived from the base address which triggered the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and base\_addr >> LOG2\_PAGE\_ SIZE respectively.
-The idea behind using three different shifted versions of the same feature is that it allows us to look into a wider range of bits than with a single version.
-It also helps give more importance to the overlapping bits and lesser importance to most and least significance bits.
-This approach can also eliminate destructive interference that can be caused by directly folding the bits into half.
-
-\item \textbf{Confidence}: The integer confidence on the scale of 0 to 100 which was used in the original SPP design.
-While the original confidence is not used directly for decision making, it can still contribute to the final decision made.
-
-\item \textbf{Program Counter XOR Depth}: PC is for the instruction which triggered the prefetch chain.
-Depth refers to the iteration count of the lookahead stages.
-In general, PC is not considered as a good basis for doing lookahead prefetching as all the prefetches with depth >= 1 gets aliased onto the same PC which won't be true in actual demand fetch case.
-This feature resolves a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases.
-This is akin to the concept of Virtual Program Counters\cite{VPC} introduced by Kim \textit{et. al.} for indirect branch prediction.
-
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC before the instruction that triggered the current prefetch. 
-Hashing together the last three PCs tell SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block.
-PCs are shifted in the increasing order of history before being hashed together.
-This is done to avoid the resultant value of zero in case 2 or more PCs are the same. 
-Additionally, obfuscating the information as it gets old allows us to get a wider and an approximate look into the program history.
-
-\item \textbf{Program Counter XOR Delta}: This tells us if a given PC favours particular value(s) of delta.
-As noted earlier, while PC in itself does not much useful information, this hash resolves PC into different values based on the tendency of that PC to favour a certain delta.
-This way, dynamic nature of different instances of the same memory access instruction can be captured here.
-
-\item \textbf{Current Signature XOR Delta}: It can be recalled from the discussion of SPP in section XX that the new signature is generated using the old signature and the blobk delta.
-While `Current Signature XOR Delta' is not the exact formula for generating the future signature, it gives an approximate idea of the path that the combination of these two values can lead to.
-
-%%%%%%% Below are the features rejected when downsizing 14 -> 9
-%\item \textbf{Current Signature}: The 12-bit signature used by SPP to index into the Pattern Table
-%\item \textbf{Delta}: The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched
-%\item \textbf{PC}: Program Counter of the Load instruction that triggers that particular prefetch
-%\item \textbf{Base address XOR Delta}: If any particular value(s) of delta are highly favoured by a given address which triggered the prefetch, this composite feature can capture that information
-%\item \textbf{Cache line XOR Delta}: This feature captures if any particular delta is highly favoured by a given cache line address
-%\item \textbf{Confidence XOR Delta}: This gives a combined insight into that fact that certain delta-confidence pairs predicted by SPP might correlate more with the prefetch outcome
-%%%%%%%%%%%%%%%%%%%%%%
-
-\end{itemize}
-
-Some of the composite features are derive from simple hashing (XOR) of two primary features.
-There is always a question of usefulness of such composite features and the `fresh' information added.
-We try to justify the choice of each feature by quantifying the contribution made towards predicting prefetch behaviour, in Section XX.
-Finally, as noted above, each feature indexes into its independent entry of perceptron weights.
-
-\section{METHODOLOGY}
-\subsection{Performance Model}
-For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim} simulator. 
-ChampSim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} 
-We model model a 1-core / 4-core / 8-core Out of Order machine. The details of the configuration parameters is summarized in Table  \ref{tab:Sim_params}.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|l|p{3.6cm}|c|}
-    \hline
-    Level & Configuration & Access Latency \\
-    \hline
-         L1 Cache & Separate I-cache and D-cache, 32 KB, 8-way & 4 cycles\\
-         L2 Cache & Private, 256 KB & 8 cycles\\
-         LLC & Shared, 2MB / core, 16-way & 20 cycles\\
-         DRAM & 4 GB Single Channel for single-core, 8 GB Double Channel for multi-core & N?? cycles\\
-    \hline
-    \end{tabular}
-    \caption{Memory Model Parameters}
-    \label{tab:Sim_params}
-\end{table}
-
-The block size was fixed 64KB. Prefetching was only triggered at L2 Cache demand accesses but could be sent to L2 Cache or Last Level Cache. 
-No L1 Data level prefetching was done. 
-LRU replacement policy was used on all levels of cache hierarchies. 
-Branch prediction was done using the perceptron branch predictor\cite{Perc_Branch}. 
-Page size was kept as 4KB. 
-Champsim operates all the prefetchers strictly in the physical address space.
-
-\subsection{Testing Under Additional Memory Constraints}
-The default single-core configuration simulates a 2MB LLC and a single channel DRAM with 12.8GB/s bandwidth. 
-We extend the simulations to include memory constraints introduced in DPC-2. 
-Specifically we look at the following two variations:
-\begin{itemize}
-    \item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited to 3.2 GB/s
-    \item \textit{Small LLC}: In this scenario, LLC size is reduced to 512 KB
-\end{itemize}
-All the multi-core simulations are only done in the default configuration.
-
-\subsection{Workloads}
-This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017} has been used to characterize and measure the prefetch performance. 
-We use all the 20 workloads available in the SPEC 2017 Suite. 
-Using SimPoint\cite{SimPoint} methodology, we identified 95 different program segments of 1 Billion instructions each.
-
-We performed cross-validation of our SPP-Perc model using SPEC 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.
-For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
-For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.
-The traces include 4 multi-core applications with 6 distinct phases per application. 
-In total, we used 285 traces representing workloads across 53 applications. 
-\newline
-\newline
-\textit{Single-core performance:} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics.
-We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
-We take the Instruction Per Cycle (IPC) of the prefetcher and divide it with the IPC of baseline. 
-The final number reported is the geometric mean of the speed-up attained on individual traces.
-\newline
-\newline
-\textit{Multi-core performance:} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
-For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 
-Each CPU keeps executing its workload till the last CPU completes 1 Billion instructions after warm-up. 
-For collecting IPC and other data, only the first billion instructions are considered as the ``region of interest''.
-
-Here we report the weighted speedup normalized to baseline \textit{i.e.}, no prefetching. 
-For each of the workloads running on a particular core of the 4-core 8 MB LLC system, we compute IPC\textsubscript{i}. 
-We then find the IPC\_isolated\textsubscript{i} of the same workload running in isolated 1-core 8 MB LLC environment. 
-Then we calculate the total weighted-IPC for a given workload mix as $\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}). 
-For each of the 100 workload-mix, the sum obtained is normalized to the weighted-IPC calculated similarly for baseline case \textit{i.e.}, no prefetching, to get the weighted-IPC-speedup. 
-Finally the geometric mean of these 100 weighted-IPC-speedup is reported as the effective speedup obtained by the prefetching scheme.
-
-We repeat the same process for 8-core workloads with the exception that 20 Million warm-up instructions and 100 Million full instructions were executed.
-This was done so as to keep the simulation run-time within reasonable limits as a single 8-core mix was taking up to 5 days to simulate 1 Billion instructions.
-
-\subsection{Preferchers Simulated}
-We compared SPP-Perc against three of the latest state of the art prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher (SPP).
-BOP was the winner of 2nd Data Prefetching Championship. 
-DA-AMPM is the enhanced version of AMPM, modified to account for DRAM row buffer locality. 
-SPP has been shown to outperform BOP on SPEC 2006 traces. 
-For each of these, we compare their speedups taking no prefetching as the baseline.
-
-\subsection{Developing Features for SPP-Perc}
-This section describes the intuition and analysis that went behind finalizing the features. 
-As noted earlier, we developed a set of 9 features which allow the perceptron throttler to correlate prefetching decision with the program context. 
-To study the correlation across each feature, we study statistically the perceptron weights and try to interpret their distribution.
-\newline  \newline
-\textbf{GLOBAL PEARSON'S CORRELATION}\newline
-For this experiment, we create a dump of the state of perceptron weights at the end of all trace execution.
-The weights obtained from running all the SPEC 2017 traces are concatenated together. 
-Since the weights are collected at the end of individual trace execution, the perceptron weights have attained a relatively stable value by now.
-Hence, this dump represents a `snap' of the trained perceptron weights across all the SPEC 2017 traces.
-The intuition here is that the feature with bulk of the perceptron weights concentrated around 0 or small magnitude numbers show a weak correlation with the prefetching outcome. 
-On the other hand, features with most of the weights saturated around highest value (+15) show a high positive correlation and the features with weights close to the lowest value (-16) show a strong negative correlation.
-
-We plot a histogram for each feature depicting weights distribution from -16 to +15 and use this histogram to generate the Pearson's correlation factor for that feature.
-In statistics, Pearson's factor is a numerical measure of the degree of linear correlation between two variables.
-It ranges from -1 to +1.
-The magnitude of Pearson's factor depicts the extent of correlation and the sign depicts whether it is a positive correlation or a negative correlation.
-Values close to 0 suggest a low correlation or at times, noisy data.
-A value of +1/-1 suggests a perfectly linear positive / negative correlation respectively.
-Figure XX depicts the histogram and the Pearson's coefficient for two of the features, YY and ZZ.
-
-Figure XX shows all the features used, arranged in the increasing order of their Pearson's factor.
-As can be seen X out of the 9 features provide a moderate to high correlation, with the magnitude of P-value > 0.4.
-The single most important feature, \_\_ helps provide a correlation to prefetch outcome with a factor of \_\_.
-\newline \newline
-\textbf{PER TRACE CORRELATION} \newline
-Another important way to look at the perceptron features is to see how much their contribution varies across the traces.
-Here we give special attention to features with low P-values in the previous experiment.
-Figure XX shows the variation of P-values three features : \_\_, \_\_ and \_\_; across all the SPEC 2017 traces.
-For simplicity, the traces are arranged in an increasing order of contribution made by the feature and the trace names have been omitted.
-It can be seen that even features with a low overall correlation provide useful correlation (magnitude > 0.4) for XX out of the 95 simpoints developed for SPEC 2017.
-\newline \newline
-\textbf{TRIMMING FEATURES USING CORRELATION} \newline
-Besides providing interesting insights into prefetching behaviour, P-value can also be used for feature selection and prefetcher tuning.
-
-Here, we introduce the concept of cross-correlation across features.
-Just as we examined correlation of each feature with the final outcome, we can also study correlation between the features.
-We used the above methodology to eliminate features which provide a little or no information that has already been captured in other features.
-
-As a part of feature selection, we initially came up a mix of 23 features- primary and composite.
-By studying cross correlation of each of these features against others in a 23x23 matrix, we identified pairs of features with correlation factor > 0.9 in magnitude and carefully eliminated redundant features.
-Using this approach, we managed to reduce the feature count to 9.
-Thus, in the final implementation of SPP-Perc, no two features have a high correlation between them.
-This way we can be sure that each feature is bringing in contribution that cannot be captured using other features.
-
-\textit{<TODO: VALIDATE THIS>} 
-Secondly, studying the relative importance of each feature enabled us to vary the number of entries dedicated for each feature.
-Features with higher Pearson's correlation (like \_\_ and \_\_) were given most importance and allowed full 12-bits of indexing.
-Features like \_\_ and \_\_, with a low overall P-value were allocated fewer indexes in the feature table.
-
-To conclude, in the above discussion we justify the features for perceptron from a statistical viewpoint.
-We also show how this information can also be used for prefetcher tuning.
-All this study was made possible only because we used on-line perceptron learning for prefetching and that enabled us to examine the weights in detail.
-
-\subsection{Overhead for SPP-Perc}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|c|c|m{4.8cm}|}
-    \hline
-        \textbf{Field} &
-        \textbf{Bits} &
-        \textbf{Comment} \\
-    \hline
-         Valid & 1 & Indicates a valid entry in the filter\\
-         Tag & 6 & Identifier for the entry in the filter\\
-         Useful & 1 & To show if the given entry led to a useful demand fetch\\
-         Perc Decision & 1 & Prefetched vs Not-prefetched \\
-    \hline
-        PC & 12 & \multirow{5}{4.8cm}{Meta-data required for perceptron training}\\
-        Address & 24 & \\
-        Delta & 7 & \\
-        Curr Signature & 12 & \\
-        Confidence & 7 & \\
-    \hline
-        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
-    \hline
-    \end{tabular}
-    \caption{Meta-data Stored in Prefetch Filter}
-    \label{tab:PF_metadata}
-\end{table}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|p{1.5cm}|p{1cm}|c|p{1.5cm}|}
-    \hline
-        \textbf{Structure} &
-        \textbf{Entry} &
-        \textbf{Components} &
-        \textbf{Total} \\
-    \hline
-        \multirow{5}{1.5cm}{Signature Table} & \multirow{5}{1.5cm}{256} & Valid (1 bit) & \multirow{5}{1cm}{11008  bits}\\
-        & & Tag (16 bits) &\\
-        & & Last Offset (6 bits) &\\  
-        & & Signature (12 bits) &\\
-        & & LRU (6 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Pattern Table} & \multirow{3}{1.5cm}{512} & $C_{sig}$ (4bits) &\multirow{3}{1cm}{24576 bits}\\
-        & & $C_{delta}$ (4*4 bits) &\\
-        & & Delta (4*7 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Perceptron Tables} & \multirow{3}{1.5cm}{4096*8 128*1} & \multirow{3}{1.5cm}{5 bits} & \multirow{3}{1.5cm}{164480 bits}\\
-        & & &\\
-        & & &\\
-    \hline
-        \multirow{2}{1.5cm}{Prefetch Filter \footnotemark[1]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{84 bits} & \multirow{2}{1.5cm}{86016 bits}\\
-        & & &\\
-    \hline
-     \multirow{2}{1.5cm}{Reject Filter \footnotemark[2]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{83 bits} & \multirow{2}{1.5cm}{84992 bits}\\
-        & & &\\
-    \hline
-        \multirow{4}{1.5cm}{Global History Register} & \multirow{4}{1.5cm}{8} & Signature (12 bits) & \multirow{4}{1.5cm}{264 bits}\\
-        & & Confidence (8 bits) & \\
-        & & Last Offset (6 bits) & \\
-        & & Delta (7 bits) & \\
-    \hline
-        \multirow{2}{1.5cm}{Accuracy Counters} & 1 & C$_{total}$ & 10 bits\\
-        & 1 & C$_{useful}$ & 10 bits\\
-    \hline
-        \multirow{3}{1.5cm}{Global PC Trackers} & \multirow{3}{1cm}{3} & $PC_1$ (12 bits)& \multirow{3}{1.cm}{36 bits}\\
-        & & $PC_2$ (12 bits) &\\
-        & & $PC_3$ (12 bits) &\\
-    \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 371356 bits = 45.33 KB}}\\
-    \hline
-    \end{tabular}
-    \caption{SPP-Perc Storage Overhead}
-    \label{tab:SPPPerc_overhead}
-\end{table}
-\footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}
-\footnotetext[2]{RF does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
-
-In this section, we analyze the hardware overhead required to implement SPP-Perc. 
-The Prefetch Filter was enhanced to accommodate storing of meta-data for perceptron training. 
-Table \ref{tab:PF_metadata} depicts the meta-data stored for each entry in the Prefetch Filter.
-Table \ref{tab:SPPPerc_overhead} shows the total storage overhead of SPP-Perc implementation.
-The hardware budget for 2nd Data Prefetching championship was 32 KB. 
-Keeping that in mind and the speedup SPP-Perc obtained over the winner, the extra hardware budget can be accounted for.
-The extra hardware also makes the overall scheme more scalable than SPP. 
-In the original SPP paper, it was demonstrated that adding extra hardware brings little advantage in terms of performance gain.
-The newly added perceptron tables can be scaled to increase / decrease features depending on the permitted budget.
-
-In terms of computations, the perceptron mechanism only introduces an extra adder tree.
-The hash perceptron mechanism makes sure than there is no actual vector multiplication happening in the hardware.
-Obtaining the perceptron sum requires addition of 9 5-bit numbers. 
-Using an adder tree of 4 5-bit adders, this can be done in ceil($log_{2}9$) = 4 steps.
-Perceptron update only requires weight update by +1 or -1.
-Thus, all the operations required for perceptron inferencing or updating the states of the perceptron throttler can be easily done in the time constraints of L2 Cache Accesses.
-
-\section{RESULTS}
-This section discusses the results obtained from running SPP-Perc, in terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per Cycle (IPC) metrics. 
-For SPEC 2017 suite, first we present the results for single-threaded workloads then the results for multi-core workloads are presented.
-
-\subsection{Single-core Results}
-Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated. 
-All the results have been normalized to baseline \textit{i.e.}, no prefetching. 
-As can be seen, SPP-Perc yields a speedup of XX\% over the baseline. 
-This is equivalent to XX\% over DA-AMPM, xx\% over BOP and XX\% over SPP. 
-Out of the XX benchmarks, SPP-Perc nearly matches or outperforms most of the prefetchers on XX traces.
-
-At its peak, SPP-Perc manages to get the IPC speedup of a factor of over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}. 
-This also corresponds to speedup gain of \textbf{22.4\%} over the next best prefetcher - SPP. 
-In general, benchmarks \textit{602.gcc\_s, 603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s} benefit the most from SPP-Perc, as compared to SPP only prefetching.
-
-<Discuss the extreme behaviour of \textit{623.xalancbmk\_s}: Improvement over SPP is as high as 31\% and even reduces to -1.5\% on traces of this benchmark>
-
-<Discuss about coverage on all traces>
-
-<Discuss cases where coverage gets specially increased>
-
-<Discuss accuracy on all traces>
-
-<Discuss about compelling increase in accuracy - details of traces. Plot linear separability>
-
-<Any case where accuracy got reduced - go into the trace. See if linear separability is the issue. Other possible issues?>
-
-\subsection{Multi-core Results}
-In this section, we demonstrate the improvement achieved by SPP-Perc for a mix of multi-programmed workloads. 
-\newline \newline
-\textit{4-core environment}
-Figure \_\_ shows a comparison of speedups obtained on workload mix of memory intensive subset of SPEC 2017. 
-We plot all the 4 prefetchers, normalized to baseline. 
-The workloads have been sorted in the increasing order of performance benefit, as measured in terms of normalized weighted speedup. 
-As can be seen, SPP-Perc offers an IPC improvement by 33.5\% on these traces. 
-This is 5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP and DA-AMPM respectively.
-
-This increased improvement by SPP-Perc on a 4-core machine is not surprising. 
-Since SPP-Perc has an extremely intelligent throttler working to eliminate useless prefetches before they can create cache pollution. 
-This further amplifies the edge that SPP had over BOP. 
-With the advantage of an orthogonal prefetcher, SPP can be as aggressive as possible and leave the accuracy aspect to perceptron throttler.
-\newline \newline
-\textit{8-core environment}
-To test the effectiveness of SPP-Perc in a tighter constrained multi-core environment, we simulated an 8-core machine with 16 MB last Level Cache. 
-Other parameters were kept same as discussed under Methodology in Section XX.
-<See if we can use speed-up over BOP as a measure here>
-
-\subsection{Additional Memory Constraints}
-Figure \_\_ shows the performance of SPP-Perc against the other prefetchers in tighter constraint environments. 
-For sake of simplicity, we show 2 traces where SPP-Perc gives the most advantage over baseline, 2 traces where SPP-Perc doesn't perform as well as the best prefetcher *<-replace with the prefetcher name once results are available* and finally GeoMean across all the traces.
-<Hopefully BOP would degrade more here as it is super aggressive. 
-By definition, SPP-Perc should be better than SPP as the throttler would better control the aggressiveness. 
-In SPP paper DA-AMPM performs close enough to BOP. That should get replicated here too.>
-
-\subsection{Cross Validation}
-Here we try to demonstrate the robustness of SPP-Perc by testing it on completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
-
-Fig \_\_ shows the speed-up achieved by BOP, SPP and SPP-Perc on selected traces of SPEC 2006, along with the overall geometric mean over whole of the benchmark and over memory-intensive subset of the SPEC 2006 benchmark.
-As can be seen, SPP-Perc peovides a speed-up of 44.4\% over baseline, on the memory intensive subset of SPEC 2006 benchmark.
-This corresponds to 6.84\% over SPP and XX\% over BOP.
-On the whole of the SPEC 2006 suite, the speedup obtained is 22.4\%, which is 3.34\% ahead of SPP.
-
-The 4-core results for SPEC 2006 are \_\_.
-
-As can be seen, the performance is very consistent given no prior parameter tuning was done keeping SPEC 2006 in mind. 
-We attribute this to the inherent adaptability of the perceptron model.
-In general, perceptron weights are able to adjust in real-time so as to find the best possible correlation between the output and the features.
-
-\section{CONCLUSION}
-In this paper, we introduced perceptron based learning for data prefetching in the form of SPP-Perc. 
-Perceptron acts an orthogonal throttler to the underlying prefetch engine. 
-SPP-Perc improves performance by over XX\% over the baseline, which corresponds to YY\% over the next best performing prefetcher. 
-We believe this paper can lead to interesting future work like generalizing SPP-Perc to work with any prefetcher with minimal knowledge of the underlying prefetcher architecture.
-
-%%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%
-
-%%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
-\bibliographystyle{ieeetr}
-\bibliography{ref}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
+The rest of the paper is organized as follows.  Section 2 talks about
+background work and recaptures the SPP architecture Section 3
+discusses the detailed implementation of the perceptron mechanism and
+the features used.  Section 4 consists of the evaluation methodology
+and explores the feature space for perceptron learning.  Section 5 is
+the performance evaluation of our prefetcher and Section 6 concludes
+the paper.

### 2018-11-15 method.tex
reorg

--- 2018-11-15 paper_1.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is the template for submission to ISCA 2018
-% The cls file is a modified from  'sig-alternate.cls'
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{sig-alternate} 
-\usepackage{mathptmx} % This is Times font
-
-\newcommand{\ignore}[1]{}
-\usepackage{fancyhdr}
-\usepackage[normalem]{ulem}
-\usepackage[hyphens]{url}
-\usepackage{microtype}
-\usepackage{fixltx2e}
-\usepackage{array}
-\usepackage{multirow}
-
-% Always include hyperref last
-\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=black,citecolor=blue,urlcolor=black]{hyperref}
-
-% Ensure letter paper
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\newcommand{\iscasubmissionnumber}{NaN}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\fancypagestyle{firstpage}{
-  \fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-  \fancyhead[C]{\normalsize{}} 
-  \fancyfoot[C]{\thepage}
-}
-
-\pagenumbering{arabic}
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\title{SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher} 
-\author{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\maketitle
-\thispagestyle{firstpage}
-\pagestyle{plain}
-
-\begin{abstract}
-\textbf{Hardware prefetching has been introduced in modern processors as a way to hide cache latencies. 
-An efficient prefetcher should be able to identify complex memory access patterns during a program execution. 
-This enables it to fetch a block ahead of its demand access, potentially saving a cache miss.
-}
-
-\textbf{In this paper, we introduce perceptron learning to help make this prefetching decision efficiently. 
-We have taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine. 
-The perceptron layer acts as a throttler to filter out the unnecessary prefetches recommended by SPP. 
-We have also explored a range of features that can be used to train the perceptron layer. 
-Our results show that SPP-Perc improves performance on the memory intensive subset of the SPEC 2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared to just SPP.
-We also demonstrate that the performance gained from using our efficient throttler scales better with increasing number of cores.
-}
-\end{abstract}
-
-\section{Introduction}
-Since past few decades, processor and memory scaling has been enabled with different philosophies in mind. 
-While processor scaling focused on speed improvements, memory scaling was centred around increasing the storage capacity. 
-The process technology reaching scaling limitations led to Memory Wall\cite{MemWall} - the increasing performance gap between the processor speed and the memory speeds. 
-Various techniques have been developed to combat this widening gap. 
-Data prefetching is one such important technique. 
-Data prefetching exploits the fact that in most applications, memory accesses have a well-defined pattern.
-
-An ideal prefetching scheme involves capturing this memory access patterns to predict future prefetches in a timely manner. 
-Other important decisions faced by the prefetcher are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block.
-Memory accesses patterns can be as simple as fetching the next cache line or a complex but predictable pattern, like in pointer chasing applications. 
-Predicting this often requires a trade-off between coverage and accuracy. 
-For a prefetcher to capture highly irregular access patterns, it needs to be highly aggressive (high coverage). 
-That also means the prefetcher would end up recommending a lot of useless prefetches, leading to cache pollution (low accuracy). 
-This effect would get even more pronounced in multi-core scenario where the last level cache is a shared resource. 
-Conversely, a conservative prefetcher would be highly accurate but might not make enough useful predictions.
-
-To this effect we propose SPP-Perc <come up with a catchy name>. 
-It is an enhancement to the existing state of the art prefetcher and overcomes the coverage vs accuracy trade-off in a robust manner. 
-The idea is to modify SPP design to make it as aggressive as it can get. 
-This enables SPP to capture complex memory access patterns hence leading to increased coverage.
-This comes at a cost of increased DRAM traffic, hence reduction in overall IPC.
-The prefetch suggestions recommended by SPP pass through the perceptron based throttler. 
-Over time, perceptron layer learns to correlate the prefetches recommended by SPP with various available features, leading to elimination of useless prefetches. 
-This leads to an overall increased accuracy of the prefetch scheme and reduction in DRAM traffic.
-
-\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and future scope for expansion. 
-The contributions made by this paper are:
-
-\begin{itemize}
-\item This is the first time where a truly on-line learning based neural network model has been used when it comes to hardware data prefetching. 
-The prior works in this area either relied on program semantics\cite{Semantics} or were application specific\cite{Datacenter} or could not yield a significant improvement over the baseline prefetcher used\cite{BadPerc}.
-
-\item Implementing the pereptron throttler in a robust and accurate practical prefetching mechanism. 
-The overall scheme yields increased accuracy as compared to the other state of the art prefetchers.
-Moreover, the perceptron throttler learns to adapt itself to shared resource constraints - leading to an even increased performance in multi-core and lower capacity / bandwidth environments.
-
-\item Examining the perceptron weights doing a statistical study to quantify the contribution made by each feature.
-We have also used this idea to develop a comprehensive set of features where each feature contributes some `new' information to the perceptron.
-The details are explained under methodology, in Section XX.
-
-\item Cross-validating the design across benchmarks. 
-All the development and parameter tuning was done on the SPEC 2017 benchmark. 
-The results have been shown to be consistent across unseen benchmarks.
-More details can be found in the results section (Section XX).
-
-%\item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind. 
-%We have used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers. 
-%We have also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
-
-%\item We came up with an efficient model to use the perceptron output as a representation of the prefetch confidence using an on-line histogram technique. 
-%We show how the simple sum can replace the existing mechanism which was essentially based on counting the number of hits for a particular entry in SPP.
-
-\end{itemize}
-
-We tested SPP-Perc on Champsim, an open source simulator which models an Out-of-Order CPU using. 
-For evaluation, we used SPEC 2017 benchmarks. 
-On a single core configuration, SPP-Perc increases performance of the system by 39.5\% as compared to the baseline with no prefetching, on the memory intensive subset of the benchmark.
-This outperforms the next best prefetcher, SPP,  by 6.84\%. 
-On a multi-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement of 11.9\% over SPP for 4-core system and 10.67\% for 8-core system.
-
-In the results part in section XX, we show a detailed analysis of the speedup achieved.
-We also used SPEC 2006 and CloudSuite benchmarks for cross-validation of the prefetcher implementation.
-
-The rest of the paper is organized as follows.
-Section 2 talks about background work and recaptures the SPP architecture
-Section 3 discusses the detailed implementation of the perceptron mechanism and the features used.
-Section 4 consists of the evaluation methodology and explores the feature space for perceptron learning.
-Section 5 is the performance evaluation of our prefetcher and Section 6 concludes the paper.
-
-\section{Background Work and Motivation}
-
-In this section, we compare the existing work done in the domain of prefetching. 
-We also discuss about some of the interesting memory access patterns seen in the workloads in SPEC 2017 benchmark. 
-The second half of this section deals with a background of Signature Path Prefetcher architecture.
-
-\subsection{Related Work}
-
-The idea of prefetching has been around since Instruction Stream Buffers were proposed by Jouppi.
-\cite{ISB}
-Some of the earliest data prefetchers developed to try to identify memory access patterns with a constant stride pattern.
-\cite{Smith} 
-\cite{Baer}
-\cite{Stride}
-Major research was dedicated to stride distance and depth prediction.
-\cite{Decoupled}
-\cite{Adaptive}
-Newer prefetchers aimed at correlating predictions with the past memory access addresses.
-\cite{Address_Correlated} 
-\cite{AMPM}
-
-With time newer and more intricate prefetchers were introduced.
-\cite{Wenisch_Temporal_Streaming}
-\cite{Stealth}
-\cite{Feedback_Directed}
-\cite{Coordinated}
-\cite{Bandwidth_Efficient}
-\cite{Pacman}
-\cite{TLB}
-\cite{Linearizing}
-\cite{Sandbox}
-\cite{VLDP}
-\cite{DoL}
-\cite{Domino}
-A whole class of streaming prefetchers developed on the lines of Spatial / Temporal Memory Streaming.
-\cite{Spatial_Pattern}
-\cite{SMS}
-\cite{Temporal_Instruction_Fetch}
-\cite{Off_Chip}
-\cite{STMS}
-\cite{SMS_JILP}
-
-Most modern prefetchers aim at identifying complex memory access patterns in a given application. 
-This allows them to capture some of the irregular patterns seen in pointer chasing data-structures.
-Offset based prefetchers are a generalization of the next-line prefetcher.
-Some of the prominent prefetchers exploiting this idea are
-Sandbox Prefetcher \cite{Sandbox}
-and the Best Offset Prefetcher \cite{BOP}
-Another class of prefetchers is lookahead Prefetchers.
-Such prefetchers try to speculate deep into the application's memory access trail.
-These include Path Confidence based Lookahead Prefetching \cite{SPP}
-and Kill the Program Counter \cite{KPC}.
-
-A related line of work also explores cache replacement - insertion policies and dead block prediction, with an eye towards prefetching. 
-\cite{DB_Pred}
-\cite{Cache_Burst}
-\cite{KPC}
-\cite{Harmony}
-Nori \textit{et.al.} introduced prefetching with  respect to runtime criticality.
-\cite{CATCH}
-\newline
-\newline
-\textit{Machine Learning and Prefetching:}
-Peled \textit{et.al.} introduce interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
-\cite{Semantics}
-The biggest challenge here is that it relies on compiler support for getting features to build the context.
-
-The paper by Liao \textit{et. al.} focuses on prefetching for data center applications.
-\cite{Datacenter}
-They use offline Machine Learning Algorithms like SVMs and Logistic Regressions to do parametric search for an optimal prefetcher configuration.
-
-%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?
-Finally, the paper on Data Cache Prefetching with Perceptron Learning\cite{BadPerc} talks about the idea of two step prefetching.
-The first step is an existing baseline prefetcher while the second step is a perceptron based throttler.
-The paper has lots of limitations in terms of design and implementation.
-That reflects in the fact that it did not lead to any significant performance gain over even the basic prefetchers like the Stride Prefetcher\cite{Stride} and the Markov Prefetcher\cite{Markov}.
-
-\subsection{Perceptron Learning in Architecture}
-
-Perceptron learning for computer architecture design has been around for a while. 
-It was first popularized by Jimenez \textit{et. al.} for doing branch prediction. \cite{Perc_Branch} 
-In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
-The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. 
-The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the prediction is made as \textit{true}, or branch taken.
-Else if the sum is < 0, the prediction is made as \textit{false}.
-
-The structure that we use is derived from the model introduced in Piecewise Linear Branch Prediction\cite{Piece_Linear}. 
-Here the feature itself is used to index into a hash of perceptron weights. 
-This form of indexing makes sure that the hyper-plane learned by the perceptron weights is able to differentiate between linearly inseparable outcomes. 
-\textit{<Is it true? Not sure>}
-
-Perceptron Learning for Reuse Prediction\cite{Perc_Reuse} extended the hash perceptron architecture to perform prediction in context of cache replacement. 
-Perceptron layer can learn to correlate the cache replacement behaviour with a variety of features. 
-Each feature has its dedicated table of perceptron weights which it indexes into. 
-Hence number of tables equals the number of features. 
-Once the perceptron weights from all the tables are obtained, they're summed and thresholded based on a pre-decided value. 
-This led to a prediction mechanism that is highly accurate, has quick learning rate and is adaptable to changing program behaviour. 
-The features introduced in that paper were derived from Program Counters of the current and the last few instructions and the memory address of the current access. 
-Even with such basic features, high level of correlation could be established by perceptrons.
-
-The work was further extended by Multiperspective Reuse Prediction\cite{Multiperspective}, which introduced a varied set of parametric features. 
-Although design space exploration for feature and parameter selection was a non-trivial task, once fixed, it could give a cache replacement predictor highly suitable for the given set of applications.
-\newline
-\newline
-\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron prediction involves multiplying the vector of input features: F\textsubscript{1xN}, with the corresponding weight vector: W\textsubscript{Nx1} in a dot product fashion to obtain the sum y\textsubscript{out}.
-Here what we use is a perceptron-like structure. 
-The feature is used to hash into the weight of perceptrons and the retrieved weights are added straight away. 
-This can be seen a dot product of $\vec{1}$\textsubscript{1xN} with W\textsubscript{Nx1}. 
-This way, the perceptron algorithm doesn't introduce any multiplication operations in the inference or training process. 
-Hence what we adapt in this work, is a perceptron-like learning algorithm as it involves the same principle involved in perceptron inferencing and training. 
-For simplicity, we will call our implementation as perceptron implementation.
-\newline
-\newline
-\textit{Perceptron Update Rule:} In all the above implementations, a uniform perceptron update principle is being followed. 
-The weights need to be updated if the prediction was wrong or the predicted sum does not cross a certain threshold. 
-After a certain training period, these weights are proportional to the probability of the outcome being true. 
-This training threshold makes sure that the perceptrons are trained till a certain level of confidence and yet they are not over-trained to the given set. 
-Care is taken that the weights saturate at certain positive and negative values so that they remain confined to the bit-width. 
-The same rules will be applied in the discussion done in section XX on SPP-Perc learning algorithm.
-
-\subsection{Prefetching with SPEC 2017}
-\textit{<Discuss SPEC 2017 traces here>
-<What are the possible graphs / analysis tools we can use for SPEC traces?>}
-
-\subsection{Signature Path Prefetcher Architecture}
-Since the work we have done builds heavily on SPP\cite{SPP}, we are presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work. 
-Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists of following structural components:
-\newline
-\newline
-\textbf{Signature Table:} ST keeps track of 256 most recently accessed pages. 
-It is meant to capture memory access patterns within a page boundary. 
-SPP indexes into an entry of ST using the page number. 
-For each entry corresponding to a page, ST stores the `last block offset' and `old signature'. 
-Last block offset is the block offset of the last memory access of that given page. 
-The block offset is calculated with respect to the page boundary. 
-The signature is a 12-bit compressed representation of the past few memory accesses for that page. 
-The signature is calculated as:
-$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
-In case a matching page entry is found, the stored signature in the corresponding entry is retrieved.
-\newline
-\newline
-Consider the case that incoming page number 10 with a block offset 3 finds a match in the ST. 
-The retrieved pattern signature is 0x30 and the Last Offset is 1. 
-Since now there is the Last Block Offset (1), Incoming Block Offset (3), Old Signature (0x30), and New Signature calculated as per above equation (0x182), SPP can infer non-speculatively that the given pattern of memory accesses (as captured in the Old Signature) leads to the particular Delta.
-In general, Delta is defined as the difference between the prefetch suggestion block and the initial block which triggered the prefetch. 
-In this case, since SPP is in learning phase, it is defined as the difference between the Incoming Block Offset and the Last Block Offset (+2 in this case). 
-This, in turn generates the new memory pattern (New Signature). 
-This newly learned signature pattern and the delta is stored in the Pattern Table.
-\newline
-\newline
-\textbf{Pattern Table:} The PT is indexed by the signature generated from ST. 
-PT holds predicted delta patterns and their confidence estimates. 
-Each entry indexed by the signature holds up to 4 unique delta predictions. 
-This is implemented by making PT as a 4-way associative table. 
-To measure an estimate of a delta confidence, PT implements two counters: C\textsubscript{sig} for each entry indexed by the signature and C\textsubscript{delta} for each delta entry for a given signature.
-More details of this on-line score-keeping mechanism is described under Confidence Tracking discussion.
-
-\textit{Lookahead Prefetching:} For each trigger to the prefetcher, the first time that the SPP indexes into the PT, it does so in the non-speculative manner. 
-We call it non-speculative because the parameters which SPP uses to predict, like last block offset, incoming block offset and signature are well-defined. 
-Assuming PT generates a valid delta prediction (rather, up to 4 valid deltas), SPP enters into the lookahead mode.
-
-In the lookahead mode, SPP looks at the last generated delta(s), chooses the one with the highest confidence and speculatively generates the new effective address and the new signature. 
-Using these new parameters, it re-accesses the PT to generate further prefetches. 
-SPP keeps on repeating this cycle of accessing PT and updating the signature based on highest confidence prefetch from the last iteration. 
-These iteration count till which SPP manages to predict prefetch entries in the lookahead manner is characterized as its `depth'. 
-While doing this, SPP also keeps compounding the confidence in each depth. 
-Thus as depth increases, overall confidence keeps decreasing. 
-This makes sense since we are going deeper into the speculation path. 
-When the confidence falls below a pre-defined threshold, SPP stops further lookahead iterations
-\newline
-\newline
-\textbf{Global History Register:} One of the innovative ideas introduced in SPP was the Global History Register, which bootstraps learning across page boundaries.
-GHR is a small 8-entry fully associative structure. 
-Whenever a prefetch suggestion from the Pattern Table crosses page boundaries, instead of completely discarding it, GHR saves that information for inter-page learning.
-Since the SPP-Perc implementation in this paper doesn't affect GHR functionality in any ways, we won't be going into more details about GHR.
-\newline
-\newline
-\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch Filter - a 1024 entry 1-way associative table which keeps a record of last few entries prefetched by SPP. 
-This proves useful for following reasons:
-\begin{itemize}
-    \item \textbf{Eliminating duplicate prefetches}\newline 
-    Any subsequent prefetches recommended by SPP mechanism is checked against the existing entries in Prefetch Filter. 
-If a valid entry is found, it means that the particular cache line is already either in L2 Cache or being fetched. 
-Hence there is no need to issue a duplicate prefetch.
-
-    \item \textbf{Updating prefetcher}\newline
-    If a subsequent demand access comes before eviction for a prefetch that was recommended by SPP, various components of SPP are updated to reflect that this was a good prefetch. 
-Correspondingly, if a prefetched cache line is evicted without being demand hit even once, SPP learns that this was a useless prefetch and updates the various tables accordingly
-
-    \item \textbf{Tracking Global Accuracy}\newline
-    PF keeps a track of usefulness of the prefetches by introducing 2 counters: C\textsubscript{total} to track the total number of prefetches getting recorded in the PF and C\textsubscript{useful} to keep a track of which prefetches recorded in the PF led to a demand hit.
-The ratio C\textsubscript{useful} / C\textsubscript{total} is the global accuracy for the prefetching scheme. 
-It is represented as $\alpha$ in the paper. 
-The global accuracy acts a real-time throttler for the path confidence value as depicted in the following discussion.
-\newline
-\end{itemize}
-\textbf{Confidence Tracking}: The PT keeps a track of hits to each signature through a counter C\textsubscript{sig}. 
-The number of hits for a given delta per signature are tracked using a counter C\textsubscript{delta}. 
-The confidence for a given delta is approximated through C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}. 
-When SPP enters into a lookahead mode, the path confidence P\textsubscript{d} is given as:
-$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown in the Prefetch Filter section. 
-The range of global accuracy is in [0,1]. 
-The lower the value of $\alpha$ for a given trace or a program phase, the overall confidence is scaled down and this makes SPP less aggressive for the upcoming prefetches.
-
-Here d is the lookahead depth. 
-For d = 1, when SPP is in non-speculative mode, P\textsubscript{0} can be thought of as 1. 
-Once SPP enters into speculative mode, it calculates the path confidence as a product of confidence of the individual prefetches encountered along that path, further scaled using $\alpha$. 
-The final P\textsubscript{d} is thresholded against prefetching-threshold (T\textsubscript{p}) to decide whether to prefetch or not. 
-If yes, then P\textsubscript{d} is thresholded against a numerically bigger fill-threshold (T\textsubscript{f}) to decide whether the prefetch should be sent to L2 Cache (high confidence prefetch) or Last Level Cache (low confidence prefetch)
-\newline<Insert a comprehensive figure about SPP structures>
-\newline<Insert another detailed picture about SPP data-path flow>
-
-\subsection{Case for an On-line Throttler}
-As compared to some of the other state of the art prefetchers [BOP], SPP is a far less aggressive prefetcher. 
-In the single core environment, this gives BOP an edge as there is no resource contention among the different cores. 
-Hence more aggressive prefetching is bound to prove beneficial. 
-As we increase the core count, we observe that SPP starts outperforming rest of the prefetchers. 
-This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. 
-\textit{<Figure showing the variation of SPP vs BOP wrt core count>} 
-For 4-core applications, SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
-
-The above analysis shows that with a more careful throttling mechanism, SPP can be tuned to become much more aggressive, leading to increased coverage. 
-The onus of maintaining the accuracy now falls on the independent throttler. 
-To test the hypothesis, we tuned down SPP to the minimum possible threshold. 
-Only by doing this we managed to increase the prefetch suggestions made by SPP, by XX\%. 
-\textit{<Comparison of SPP-Unleased with SPP / BOP>}
-Obviously this came at a cost of increased DRAM traffic and cache pollution.
-
-Moreover, the on-line confidence mechanism introduced in SPP was very rudimentary.
-It was based on taking a ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig} as explained above.
-The same confidence was used to make the decision whether to prefetch or not to prefetch; and which level to prefetch.
-While this approximation was shown to work in the original SPP, we believe that a better form of on-line decision making was possible.
-Hence, it was necessary to build a robust and adaptable learning mechanism to accept / reject the prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level Cache).
-
-To that effect, we introduce an independent on-line perceptron based throttling mechanism.
-
-\section{SPP-Perc}
-
-\subsection{Changes made to original SPP}
-To modify the SPP Design to suit our scheme, the following changes were made:
-
-\begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline
-In the original SPP design, an internal confidence mechanism was developed which was compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the prefetch suggestion.
-Furthermore, the accepted suggestions were thresholded against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache vs Last Level Cache.
-T\textsubscript{p} and T\textsubscript{f} were set to 25 and 90 respectively, on the scale of 0 to 100.
-In SPP-Perc, the perceptron sum is used to make these decisions. 
-Hence these two thresholds are no longer needed
-
-\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
-Original SPP would not consider the slots available in the L2 MSHR queue. 
-It had an assumption that at no time can the number of prefetches suggested by SPP could exceed the capacity of the queue. 
-While it was not proven in the paper, the internal confidence mechanism along with the thresholds made sure that the above assumption was maintained. 
-In SPP-Perc since there can be no assumption on the degree of aggressiveness - because SPP is being tuned to its lowest thresholds, we explicitly check that at no point should the number of suggested prefetches exceed L2 MSHR queue.
-
-\item \textbf{Enhanced Filter}\newline
-Original SPP introduces a concept of Prefetch Filter, which keeps a track of up to 1024 L2 Cache prefetches suggested by SPP. 
-The Filter was modified to store meta-data like program counter, memory access address etc. 
-This is needed to train the perceptron when the result of that particular prefetch gets available. 
-At training time, this data is used to retrieve the state of the prefetcher when the prediction was made.
-
-\item \textbf{Reject Filter introduced}\newline
-The Prefetch Filter mentioned above keeps a track of the prefetches that actually happened in the L2 Cache. 
-Correspondingly, we introduced a reject filter. 
-It keeps a track of the prefetches that were suggested by the base SPP engine but were rejected by the perceptron. 
-In case it is detected that a future demand fetch could have used this rejected prefetch, the perceptron weights are updated to reflect the same
-
-\end{itemize}
-
-\subsection{The Perceptron Layer}
-As discussed earlier, the perceptron layer was introduced to act as a throttler to the underlying SPP engine. 
-This section explores the details of the perceptron layer and the range of features that are ultimately used by the perceptron for correlating the prefetches.
-
-The neural network weights are stored in separate tables for each feature.
-The indexing of each feature is done using different number of bits.
-At most 12 bits of a feature are used to index into the perceptron table.
-Certain features require more resolution power \textit{i.e.}, full 12 bits of indexing.
-On the other hand, some features requiring lower resolution required as low as 7 bits of indexing.
-The variable indexing was determined by studying the features and fine-tuned empirically so as to achieve a good accuracy vs area trade-off.
-Exact details are mentioned under `Area Overheads' in Section XX.
-
-A single entry in the table corresponds to a perceptron weight.
-Each weight is a 5 bit counter - saturating at -16 and +15. 
-Each feature gets its own dedicated table. 
-Our proposed design uses 9 features. 
-As the program execution begins, all the weights are initialized to 0.
-\newline \newline
-\textbf{INFERENCING}\newline
-In the Champsim simulator, the prefetching method is triggered on every L2 Cache demand access. 
-At that point, the prefetcher has the option to do or not do prefetch - in case it does, then how many cache lines to prefetch. 
-The suggested prefetches can be places either in L2 Cache or in Last Level Cache. 
-The prefetcher gets to decide this, and is usually done based on an internal confidence mechanism.
-
-When SPP gets triggered, it suggests up to 4 (=Associativity of the Pattern Table) candidates for prefetch based on the demand access address. 
-In each subsequent lookahead depth, SPP suggests up to 4 more prefetch candidates, as explained in section\_\_. 
-All these recommended prefetches are tested in the perceptron throttler before they can access the Prefetch Filter, as in the original SPP data-path.
-
-The perceptron looks at the micro-architectural state, or `features', at that instant. 
-Each feature indexes into the hash of up to 4096-entry table dedicated for that feature. 
-Feature \#1 indexes into a particular entry in table \#1 and so on.
-
-Once all the weights are retrieved, they are added together, much like the traditional perceptron dot product. 
-On obtaining the overall sum, it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO. 
-Only the prefetch candidates with the perceptron sum higher that the threshold qualify for prefetching. 
-The prefetches that qualify the perceptron throttler stage go through the Prefetch Filter stage, just as in the SPP design.
-The Prefetch Filter eliminates any redundant prefetch suggestions.
-In SPP-Perc, all the meta-data that is required to recreate the features is also stored in the Prefetch Filter. 
-At a later stage, when the feedback of the current prefetched line is available, the stored data can be used to train the perceptron.
-
-The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch cache line. 
-All prefetch candidates qualified till this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill level.
-The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
-
-The perceptron sum can also be seen as the sum of the individual contribution per feature.
-The value of each contribution corresponds to the extent of confidence of the final output with that given feature.
-By summing the individual contributions, the final perceptron sum denotes the overall confidence for that prefetch suggestion. 
-In a way, by thresholding the perceptron sum on two different thresholds, we are diving the confidence scale of the prefetch suggestion into three bins.
-The first bin corresponding to the lowest confidence leads to prefetch candidate being rejected.
-The next bin corresponds to prefetches with a moderate confidence level. 
-Such prefetches are directed to fill the bigger Last Level Cache and potentially not pollute the more scarce L2 Cache.
-The highest confidence prefetches are filled in the L2 Cache.
-
-In addition to the Prefetch Filter mentioned above, SPP-Perc also maintains a `Reject Filter'. 
-Reject filter is a 1024-entry deep 1-way associative table. 
-In case that a prefetch suggestion is rejected by perceptron throttler, it instead gets logged into the Reject Filter. 
-The Reject Filter keeps track of 1024 entries that were suggested by the Pattern Table but got rejected by the perceptron throttler. 
-This can be used to train the perceptron to avoid false negatives \textit{i.e.}, cases where prediction was to reject the prefetch but the prefetch would have been useful.
-
-As previously stated, an enhancement to the original SPP design includes storing all the meta-data describing the program state at the instant of prefetching, in the Prefetch Filter. 
-This information is useful when the perceptron needs to be updated subsequently. 
-\newline \newline
-\textbf{TRAINING}\newline
-In the prefetching environment, feedback of a prefetch is received whenever there is an eviction or a demand access from L2 Cache. 
-This acts as a trigger to update all the baseline SPP entries and to also train the perceptron throttler. 
-Training of the perceptron throttler involves restoring the meta-data that was stored in Prefetch Filter or the Reject Filter. 
-This is done by indexing into the corresponding filters using the cache line address of the block being used to train. (10-bits indexing + 6-bits tag matching). 
-Once the state of the program at the time of prefetching is available, it is used to index into the perceptron weights table.
-
-If a demand access block which triggers the training  was tagged as a valid prefetch in the Prefetch Filter, it means that the earlier prefetch prediction was correct. 
-In that case the perceptron weights are incremented by 1 if the predicted sum does not cross a pre-defined threshold. 
- In case it is a cache block eviction that led to training and the corresponding valid entry was found in the Prefetch Filter, it shows that the prediction made by the perceptron was wrong. 
-Perceptron should have ideally rejected the prefetch suggestion as a low-confidence prefetch. 
-Here the weights are decremented by 1 to reflect the same. 
-In either case, weights are saturated at -16 or +15.
-
-A secondary training mechanism also kicks in during demand fetches. 
-Before the demand access triggers the next set of prefetches, it checks the Reject Filter for the valid entry. 
-In case it is a hit, that means that the corresponding cache line was initially suggested by the underlying SPP engine but rejected by the perceptron throttler. 
-This means that perceptron should have been more confident about that particular prefetch. 
-Once such a scenario is identified, the state of the execution at the time of prefetch is retrieved from the Reject Table. 
-The retrieved data is used to index into the various weights tables of the perceptron and the corresponding values are updated by +1, saturating between and -16 and +15. 
-This update reflects increased confidence for the prediction corresponding to that prefetch.
-
-This mechanism allows us to solve the classical problem of exploiting `lost opportunity'. 
-In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
-Our two-step SPP-Perc allows us to overcome that issue in a simple manner.
-
-\subsection{Features used by Perceptron}
-
-Here we discuss the various features which help correlate the prefetching decision with the program behaviour
-\begin{itemize}
-\item \textbf{Base address} of the demand access which triggered the prefetch. 
-Since prefetch are triggered from the L2 demand access, they tend to be correlated with the triggering address.
-
-\item \textbf{Cache line} and \textbf{Page address}: These two separate features are derived from the base address which triggered the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and base\_addr >> LOG2\_PAGE\_ SIZE respectively.
-The idea behind using three different shifted versions of the same feature is that it allows us to look into a wider range of bits than with a single version.
-It also helps give more importance to the overlapping bits and lesser importance to most and least significance bits.
-This approach can also eliminate destructive interference that can be caused by directly folding the bits into half.
-
-\item \textbf{Confidence}: The integer confidence on the scale of 0 to 100 which was used in the original SPP design.
-While the original confidence is not used directly for decision making, it can still contribute to the final decision made.
-
-\item \textbf{Program Counter XOR Depth}: PC is for the instruction which triggered the prefetch chain.
-Depth refers to the iteration count of the lookahead stages.
-In general, PC is not considered as a good basis for doing lookahead prefetching as all the prefetches with depth >= 1 gets aliased onto the same PC which won't be true in actual demand fetch case.
-This feature resolves a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases.
-This is akin to the concept of Virtual Program Counters\cite{VPC} introduced by Kim \textit{et. al.} for indirect branch prediction.
-
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC before the instruction that triggered the current prefetch. 
-Hashing together the last three PCs tell SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block.
-PCs are shifted in the increasing order of history before being hashed together.
-This is done to avoid the resultant value of zero in case 2 or more PCs are the same. 
-Additionally, obfuscating the information as it gets old allows us to get a wider and an approximate look into the program history.
-
-\item \textbf{Program Counter XOR Delta}: This tells us if a given PC favours particular value(s) of delta.
-As noted earlier, while PC in itself does not much useful information, this hash resolves PC into different values based on the tendency of that PC to favour a certain delta.
-This way, dynamic nature of different instances of the same memory access instruction can be captured here.
-
-\item \textbf{Current Signature XOR Delta}: It can be recalled from the discussion of SPP in section XX that the new signature is generated using the old signature and the blobk delta.
-While `Current Signature XOR Delta' is not the exact formula for generating the future signature, it gives an approximate idea of the path that the combination of these two values can lead to.
-
-%%%%%%% Below are the features rejected when downsizing 14 -> 9
-%\item \textbf{Current Signature}: The 12-bit signature used by SPP to index into the Pattern Table
-%\item \textbf{Delta}: The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched
-%\item \textbf{PC}: Program Counter of the Load instruction that triggers that particular prefetch
-%\item \textbf{Base address XOR Delta}: If any particular value(s) of delta are highly favoured by a given address which triggered the prefetch, this composite feature can capture that information
-%\item \textbf{Cache line XOR Delta}: This feature captures if any particular delta is highly favoured by a given cache line address
-%\item \textbf{Confidence XOR Delta}: This gives a combined insight into that fact that certain delta-confidence pairs predicted by SPP might correlate more with the prefetch outcome
-%%%%%%%%%%%%%%%%%%%%%%
-
-\end{itemize}
-
-Some of the composite features are derive from simple hashing (XOR) of two primary features.
-There is always a question of usefulness of such composite features and the `fresh' information added.
-We try to justify the choice of each feature by quantifying the contribution made towards predicting prefetch behaviour, in Section XX.
-Finally, as noted above, each feature indexes into its independent entry of perceptron weights.
-
 \section{METHODOLOGY}
 \subsection{Performance Model}
-For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim} simulator. 
-ChampSim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} 
-We model model a 1-core / 4-core / 8-core Out of Order machine. The details of the configuration parameters is summarized in Table  \ref{tab:Sim_params}.
+For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim}
+simulator.  ChampSim is an enhanced version of the framework that was
+used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} We
+model model a 1-core / 4-core / 8-core Out of Order machine. The
+details of the configuration parameters is summarized in Table
+\ref{tab:Sim_params}.
 
@@
 
-The block size was fixed 64KB. Prefetching was only triggered at L2 Cache demand accesses but could be sent to L2 Cache or Last Level Cache. 
-No L1 Data level prefetching was done. 
-LRU replacement policy was used on all levels of cache hierarchies. 
-Branch prediction was done using the perceptron branch predictor\cite{Perc_Branch}. 
-Page size was kept as 4KB. 
-Champsim operates all the prefetchers strictly in the physical address space.
+The block size was fixed 64KB. Prefetching was only triggered at L2
+Cache demand accesses but could be sent to L2 Cache or Last Level
+Cache.  No L1 Data level prefetching was done.  LRU replacement policy
+was used on all levels of cache hierarchies.  Branch prediction was
+done using the perceptron branch predictor\cite{Perc_Branch}.  Page
+size was kept as 4KB.  Champsim operates all the prefetchers strictly
+in the physical address space.
 
 \subsection{Testing Under Additional Memory Constraints}
-The default single-core configuration simulates a 2MB LLC and a single channel DRAM with 12.8GB/s bandwidth. 
-We extend the simulations to include memory constraints introduced in DPC-2. 
-Specifically we look at the following two variations:
+The default single-core configuration simulates a 2MB LLC and a single
+channel DRAM with 12.8GB/s bandwidth.  We extend the simulations to
+include memory constraints introduced in DPC-2.  Specifically we look
+at the following two variations:
 \begin{itemize}
-    \item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited to 3.2 GB/s
-    \item \textit{Small LLC}: In this scenario, LLC size is reduced to 512 KB
+\item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited
+  to 3.2 GB/s
+\item \textit{Small LLC}: In this scenario, LLC size is reduced to 512
+  KB
 \end{itemize}
-All the multi-core simulations are only done in the default configuration.
+All the multi-core simulations are only done in the default
+configuration.
 
 \subsection{Workloads}
-This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017} has been used to characterize and measure the prefetch performance. 
-We use all the 20 workloads available in the SPEC 2017 Suite. 
-Using SimPoint\cite{SimPoint} methodology, we identified 95 different program segments of 1 Billion instructions each.
+This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017}
+has been used to characterize and measure the prefetch performance.
+We use all the 20 workloads available in the SPEC 2017 Suite.  Using
+SimPoint\cite{SimPoint} methodology, we identified 95 different
+program segments of 1 Billion instructions each.
 
-We performed cross-validation of our SPP-Perc model using SPEC 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.
-For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
-For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.
-The traces include 4 multi-core applications with 6 distinct phases per application. 
-In total, we used 285 traces representing workloads across 53 applications. 
-\newline
-\newline
-\textit{Single-core performance:} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics.
-We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
-We take the Instruction Per Cycle (IPC) of the prefetcher and divide it with the IPC of baseline. 
-The final number reported is the geometric mean of the speed-up attained on individual traces.
-\newline
-\newline
-\textit{Multi-core performance:} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
-For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 
-Each CPU keeps executing its workload till the last CPU completes 1 Billion instructions after warm-up. 
-For collecting IPC and other data, only the first billion instructions are considered as the ``region of interest''.
+We performed cross-validation of our SPP-Perc model using SPEC
+2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.  For
+SPEC 2006, we developed 94 simpoints spread across the 29
+applications.  For CloudSuite, we used the traces made available for
+the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.  The traces
+include 4 multi-core applications with 6 distinct phases per
+application.  In total, we used 285 traces representing workloads
+across 53 applications.
 
-Here we report the weighted speedup normalized to baseline \textit{i.e.}, no prefetching. 
-For each of the workloads running on a particular core of the 4-core 8 MB LLC system, we compute IPC\textsubscript{i}. 
-We then find the IPC\_isolated\textsubscript{i} of the same workload running in isolated 1-core 8 MB LLC environment. 
-Then we calculate the total weighted-IPC for a given workload mix as $\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}). 
-For each of the 100 workload-mix, the sum obtained is normalized to the weighted-IPC calculated similarly for baseline case \textit{i.e.}, no prefetching, to get the weighted-IPC-speedup. 
-Finally the geometric mean of these 100 weighted-IPC-speedup is reported as the effective speedup obtained by the prefetching scheme.
+\textit{Single-core performance:} For single-core simulations, we use
+200 Million instructions to warm-up the microarchitectural structures
+and next 1 Billion instructions to do detailed simulations and collect
+run-time statistics.  We report the performance as speedup over
+baseline \textit{i.e.}, no prefetching.  We take the Instruction Per
+Cycle (IPC) of the prefetcher and divide it with the IPC of baseline.
+The final number reported is the geometric mean of the speed-up
+attained on individual traces.
 
-We repeat the same process for 8-core workloads with the exception that 20 Million warm-up instructions and 100 Million full instructions were executed.
-This was done so as to keep the simulation run-time within reasonable limits as a single 8-core mix was taking up to 5 days to simulate 1 Billion instructions.
+\textit{Multi-core performance:} For multi-application workloads, we
+generate 100 completely random mixes and another 100 mixes from the
+memory intensive subset of SPEC 2017.  For 4-core workload, 200
+Million instructions are used for warm-up and additional 1 Billion
+instruction simulated for collecting statistics.  Each CPU keeps
+executing its workload till the last CPU completes 1 Billion
+instructions after warm-up.  For collecting IPC and other data, only
+the first billion instructions are considered as the ``region of
+interest''.
+
+Here we report the weighted speedup normalized to baseline
+\textit{i.e.}, no prefetching.  For each of the workloads running on a
+particular core of the 4-core 8 MB LLC system, we compute
+IPC\textsubscript{i}.  We then find the IPC\_isolated\textsubscript{i}
+of the same workload running in isolated 1-core 8 MB LLC environment.
+Then we calculate the total weighted-IPC for a given workload mix as
+$\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}).  For
+each of the 100 workload-mix, the sum obtained is normalized to the
+weighted-IPC calculated similarly for baseline case \textit{i.e.}, no
+prefetching, to get the weighted-IPC-speedup.  Finally the geometric
+mean of these 100 weighted-IPC-speedup is reported as the effective
+speedup obtained by the prefetching scheme.
+
+We repeat the same process for 8-core workloads with the exception
+that 20 Million warm-up instructions and 100 Million full instructions
+were executed.  This was done so as to keep the simulation run-time
+within reasonable limits as a single 8-core mix was taking up to 5
+days to simulate 1 Billion instructions.
 
 \subsection{Preferchers Simulated}
-We compared SPP-Perc against three of the latest state of the art prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher (SPP).
-BOP was the winner of 2nd Data Prefetching Championship. 
-DA-AMPM is the enhanced version of AMPM, modified to account for DRAM row buffer locality. 
-SPP has been shown to outperform BOP on SPEC 2006 traces. 
-For each of these, we compare their speedups taking no prefetching as the baseline.
+We compared SPP-Perc against three of the latest state of the art
+prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map
+Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher
+(SPP).  BOP was the winner of 2nd Data Prefetching Championship.
+DA-AMPM is the enhanced version of AMPM, modified to account for DRAM
+row buffer locality.  SPP has been shown to outperform BOP on SPEC
+2006 traces.  For each of these, we compare their speedups taking no
+prefetching as the baseline.
 
 \subsection{Developing Features for SPP-Perc}
-This section describes the intuition and analysis that went behind finalizing the features. 
-As noted earlier, we developed a set of 9 features which allow the perceptron throttler to correlate prefetching decision with the program context. 
-To study the correlation across each feature, we study statistically the perceptron weights and try to interpret their distribution.
-\newline  \newline
-\textbf{GLOBAL PEARSON'S CORRELATION}\newline
-For this experiment, we create a dump of the state of perceptron weights at the end of all trace execution.
-The weights obtained from running all the SPEC 2017 traces are concatenated together. 
-Since the weights are collected at the end of individual trace execution, the perceptron weights have attained a relatively stable value by now.
-Hence, this dump represents a `snap' of the trained perceptron weights across all the SPEC 2017 traces.
-The intuition here is that the feature with bulk of the perceptron weights concentrated around 0 or small magnitude numbers show a weak correlation with the prefetching outcome. 
-On the other hand, features with most of the weights saturated around highest value (+15) show a high positive correlation and the features with weights close to the lowest value (-16) show a strong negative correlation.
+This section describes the intuition and analysis that went behind
+finalizing the features.  As noted earlier, we developed a set of 9
+features which allow the perceptron throttler to correlate prefetching
+decision with the program context.  To study the correlation across
+each feature, we study statistically the perceptron weights and try to
+interpret their distribution.
 
-We plot a histogram for each feature depicting weights distribution from -16 to +15 and use this histogram to generate the Pearson's correlation factor for that feature.
-In statistics, Pearson's factor is a numerical measure of the degree of linear correlation between two variables.
-It ranges from -1 to +1.
-The magnitude of Pearson's factor depicts the extent of correlation and the sign depicts whether it is a positive correlation or a negative correlation.
-Values close to 0 suggest a low correlation or at times, noisy data.
-A value of +1/-1 suggests a perfectly linear positive / negative correlation respectively.
-Figure XX depicts the histogram and the Pearson's coefficient for two of the features, YY and ZZ.
+\textbf{GLOBAL PEARSON'S CORRELATION}\newline For this experiment, we
+create a dump of the state of perceptron weights at the end of all
+trace execution.  The weights obtained from running all the SPEC 2017
+traces are concatenated together.  Since the weights are collected at
+the end of individual trace execution, the perceptron weights have
+attained a relatively stable value by now.  Hence, this dump
+represents a `snap' of the trained perceptron weights across all the
+SPEC 2017 traces.  The intuition here is that the feature with bulk of
+the perceptron weights concentrated around 0 or small magnitude
+numbers show a weak correlation with the prefetching outcome.  On the
+other hand, features with most of the weights saturated around highest
+value (+15) show a high positive correlation and the features with
+weights close to the lowest value (-16) show a strong negative
+correlation.
 
-Figure XX shows all the features used, arranged in the increasing order of their Pearson's factor.
-As can be seen X out of the 9 features provide a moderate to high correlation, with the magnitude of P-value > 0.4.
-The single most important feature, \_\_ helps provide a correlation to prefetch outcome with a factor of \_\_.
-\newline \newline
-\textbf{PER TRACE CORRELATION} \newline
-Another important way to look at the perceptron features is to see how much their contribution varies across the traces.
-Here we give special attention to features with low P-values in the previous experiment.
-Figure XX shows the variation of P-values three features : \_\_, \_\_ and \_\_; across all the SPEC 2017 traces.
-For simplicity, the traces are arranged in an increasing order of contribution made by the feature and the trace names have been omitted.
-It can be seen that even features with a low overall correlation provide useful correlation (magnitude > 0.4) for XX out of the 95 simpoints developed for SPEC 2017.
-\newline \newline
-\textbf{TRIMMING FEATURES USING CORRELATION} \newline
-Besides providing interesting insights into prefetching behaviour, P-value can also be used for feature selection and prefetcher tuning.
+We plot a histogram for each feature depicting weights distribution
+from -16 to +15 and use this histogram to generate the Pearson's
+correlation factor for that feature.  In statistics, Pearson's factor
+is a numerical measure of the degree of linear correlation between two
+variables.  It ranges from -1 to +1.  The magnitude of Pearson's
+factor depicts the extent of correlation and the sign depicts whether
+it is a positive correlation or a negative correlation.  Values close
+to 0 suggest a low correlation or at times, noisy data.  A value of
++1/-1 suggests a perfectly linear positive / negative correlation
+respectively.  Figure XX depicts the histogram and the Pearson's
+coefficient for two of the features, YY and ZZ.
+
+Figure XX shows all the features used, arranged in the increasing
+order of their Pearson's factor.  As can be seen X out of the 9
+features provide a moderate to high correlation, with the magnitude of
+P-value > 0.4.  The single most important feature, \_\_ helps provide
+a correlation to prefetch outcome with a factor of \_\_.
+
+\textbf{PER TRACE CORRELATION} \newline Another important way to look
+at the perceptron features is to see how much their contribution
+varies across the traces.  Here we give special attention to features
+with low P-values in the previous experiment.  Figure XX shows the
+variation of P-values three features : \_\_, \_\_ and \_\_; across all
+the SPEC 2017 traces.  For simplicity, the traces are arranged in an
+increasing order of contribution made by the feature and the trace
+names have been omitted.  It can be seen that even features with a low
+overall correlation provide useful correlation (magnitude > 0.4) for
+XX out of the 95 simpoints developed for SPEC 2017.
+
+\textbf{TRIMMING FEATURES USING CORRELATION} \newline Besides
+providing interesting insights into prefetching behaviour, P-value can
+also be used for feature selection and prefetcher tuning.
 
 Here, we introduce the concept of cross-correlation across features.
-Just as we examined correlation of each feature with the final outcome, we can also study correlation between the features.
-We used the above methodology to eliminate features which provide a little or no information that has already been captured in other features.
+Just as we examined correlation of each feature with the final
+outcome, we can also study correlation between the features.  We used
+the above methodology to eliminate features which provide a little or
+no information that has already been captured in other features.
 
-As a part of feature selection, we initially came up a mix of 23 features- primary and composite.
-By studying cross correlation of each of these features against others in a 23x23 matrix, we identified pairs of features with correlation factor > 0.9 in magnitude and carefully eliminated redundant features.
-Using this approach, we managed to reduce the feature count to 9.
-Thus, in the final implementation of SPP-Perc, no two features have a high correlation between them.
-This way we can be sure that each feature is bringing in contribution that cannot be captured using other features.
+As a part of feature selection, we initially came up a mix of 23
+features- primary and composite.  By studying cross correlation of
+each of these features against others in a 23x23 matrix, we identified
+pairs of features with correlation factor > 0.9 in magnitude and
+carefully eliminated redundant features.  Using this approach, we
+managed to reduce the feature count to 9.  Thus, in the final
+implementation of SPP-Perc, no two features have a high correlation
+between them.  This way we can be sure that each feature is bringing
+in contribution that cannot be captured using other features.
 
-\textit{<TODO: VALIDATE THIS>} 
-Secondly, studying the relative importance of each feature enabled us to vary the number of entries dedicated for each feature.
-Features with higher Pearson's correlation (like \_\_ and \_\_) were given most importance and allowed full 12-bits of indexing.
-Features like \_\_ and \_\_, with a low overall P-value were allocated fewer indexes in the feature table.
+\textit{<TODO: VALIDATE THIS>} Secondly, studying the relative
+importance of each feature enabled us to vary the number of entries
+dedicated for each feature.  Features with higher Pearson's
+correlation (like \_\_ and \_\_) were given most importance and
+allowed full 12-bits of indexing.  Features like \_\_ and \_\_, with a
+low overall P-value were allocated fewer indexes in the feature table.
 
-To conclude, in the above discussion we justify the features for perceptron from a statistical viewpoint.
-We also show how this information can also be used for prefetcher tuning.
-All this study was made possible only because we used on-line perceptron learning for prefetching and that enabled us to examine the weights in detail.
+To conclude, in the above discussion we justify the features for
+perceptron from a statistical viewpoint.  We also show how this
+information can also be used for prefetcher tuning.  All this study
+was made possible only because we used on-line perceptron learning for
+prefetching and that enabled us to examine the weights in detail.
 
@@
 
-In this section, we analyze the hardware overhead required to implement SPP-Perc. 
-The Prefetch Filter was enhanced to accommodate storing of meta-data for perceptron training. 
-Table \ref{tab:PF_metadata} depicts the meta-data stored for each entry in the Prefetch Filter.
-Table \ref{tab:SPPPerc_overhead} shows the total storage overhead of SPP-Perc implementation.
-The hardware budget for 2nd Data Prefetching championship was 32 KB. 
-Keeping that in mind and the speedup SPP-Perc obtained over the winner, the extra hardware budget can be accounted for.
-The extra hardware also makes the overall scheme more scalable than SPP. 
-In the original SPP paper, it was demonstrated that adding extra hardware brings little advantage in terms of performance gain.
-The newly added perceptron tables can be scaled to increase / decrease features depending on the permitted budget.
+In this section, we analyze the hardware overhead required to
+implement SPP-Perc.  The Prefetch Filter was enhanced to accommodate
+storing of meta-data for perceptron training.  Table
+\ref{tab:PF_metadata} depicts the meta-data stored for each entry in
+the Prefetch Filter.  Table \ref{tab:SPPPerc_overhead} shows the total
+storage overhead of SPP-Perc implementation.  The hardware budget for
+2nd Data Prefetching championship was 32 KB.  Keeping that in mind and
+the speedup SPP-Perc obtained over the winner, the extra hardware
+budget can be accounted for.  The extra hardware also makes the
+overall scheme more scalable than SPP.  In the original SPP paper, it
+was demonstrated that adding extra hardware brings little advantage in
+terms of performance gain.  The newly added perceptron tables can be
+scaled to increase / decrease features depending on the permitted
+budget.
 
-In terms of computations, the perceptron mechanism only introduces an extra adder tree.
-The hash perceptron mechanism makes sure than there is no actual vector multiplication happening in the hardware.
+In terms of computations, the perceptron mechanism only introduces an
+extra adder tree.  The hash perceptron mechanism makes sure than there
+is no actual vector multiplication happening in the hardware.
 Obtaining the perceptron sum requires addition of 9 5-bit numbers.
-Using an adder tree of 4 5-bit adders, this can be done in ceil($log_{2}9$) = 4 steps.
-Perceptron update only requires weight update by +1 or -1.
-Thus, all the operations required for perceptron inferencing or updating the states of the perceptron throttler can be easily done in the time constraints of L2 Cache Accesses.
-
-\section{RESULTS}
-This section discusses the results obtained from running SPP-Perc, in terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per Cycle (IPC) metrics. 
-For SPEC 2017 suite, first we present the results for single-threaded workloads then the results for multi-core workloads are presented.
-
-\subsection{Single-core Results}
-Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated. 
-All the results have been normalized to baseline \textit{i.e.}, no prefetching. 
-As can be seen, SPP-Perc yields a speedup of XX\% over the baseline. 
-This is equivalent to XX\% over DA-AMPM, xx\% over BOP and XX\% over SPP. 
-Out of the XX benchmarks, SPP-Perc nearly matches or outperforms most of the prefetchers on XX traces.
-
-At its peak, SPP-Perc manages to get the IPC speedup of a factor of over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}. 
-This also corresponds to speedup gain of \textbf{22.4\%} over the next best prefetcher - SPP. 
-In general, benchmarks \textit{602.gcc\_s, 603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s} benefit the most from SPP-Perc, as compared to SPP only prefetching.
-
-<Discuss the extreme behaviour of \textit{623.xalancbmk\_s}: Improvement over SPP is as high as 31\% and even reduces to -1.5\% on traces of this benchmark>
-
-<Discuss about coverage on all traces>
-
-<Discuss cases where coverage gets specially increased>
-
-<Discuss accuracy on all traces>
-
-<Discuss about compelling increase in accuracy - details of traces. Plot linear separability>
-
-<Any case where accuracy got reduced - go into the trace. See if linear separability is the issue. Other possible issues?>
-
-\subsection{Multi-core Results}
-In this section, we demonstrate the improvement achieved by SPP-Perc for a mix of multi-programmed workloads. 
-\newline \newline
-\textit{4-core environment}
-Figure \_\_ shows a comparison of speedups obtained on workload mix of memory intensive subset of SPEC 2017. 
-We plot all the 4 prefetchers, normalized to baseline. 
-The workloads have been sorted in the increasing order of performance benefit, as measured in terms of normalized weighted speedup. 
-As can be seen, SPP-Perc offers an IPC improvement by 33.5\% on these traces. 
-This is 5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP and DA-AMPM respectively.
-
-This increased improvement by SPP-Perc on a 4-core machine is not surprising. 
-Since SPP-Perc has an extremely intelligent throttler working to eliminate useless prefetches before they can create cache pollution. 
-This further amplifies the edge that SPP had over BOP. 
-With the advantage of an orthogonal prefetcher, SPP can be as aggressive as possible and leave the accuracy aspect to perceptron throttler.
-\newline \newline
-\textit{8-core environment}
-To test the effectiveness of SPP-Perc in a tighter constrained multi-core environment, we simulated an 8-core machine with 16 MB last Level Cache. 
-Other parameters were kept same as discussed under Methodology in Section XX.
-<See if we can use speed-up over BOP as a measure here>
-
-\subsection{Additional Memory Constraints}
-Figure \_\_ shows the performance of SPP-Perc against the other prefetchers in tighter constraint environments. 
-For sake of simplicity, we show 2 traces where SPP-Perc gives the most advantage over baseline, 2 traces where SPP-Perc doesn't perform as well as the best prefetcher *<-replace with the prefetcher name once results are available* and finally GeoMean across all the traces.
-<Hopefully BOP would degrade more here as it is super aggressive. 
-By definition, SPP-Perc should be better than SPP as the throttler would better control the aggressiveness. 
-In SPP paper DA-AMPM performs close enough to BOP. That should get replicated here too.>
-
-\subsection{Cross Validation}
-Here we try to demonstrate the robustness of SPP-Perc by testing it on completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
-
-Fig \_\_ shows the speed-up achieved by BOP, SPP and SPP-Perc on selected traces of SPEC 2006, along with the overall geometric mean over whole of the benchmark and over memory-intensive subset of the SPEC 2006 benchmark.
-As can be seen, SPP-Perc peovides a speed-up of 44.4\% over baseline, on the memory intensive subset of SPEC 2006 benchmark.
-This corresponds to 6.84\% over SPP and XX\% over BOP.
-On the whole of the SPEC 2006 suite, the speedup obtained is 22.4\%, which is 3.34\% ahead of SPP.
-
-The 4-core results for SPEC 2006 are \_\_.
-
-As can be seen, the performance is very consistent given no prior parameter tuning was done keeping SPEC 2006 in mind. 
-We attribute this to the inherent adaptability of the perceptron model.
-In general, perceptron weights are able to adjust in real-time so as to find the best possible correlation between the output and the features.
-
-\section{CONCLUSION}
-In this paper, we introduced perceptron based learning for data prefetching in the form of SPP-Perc. 
-Perceptron acts an orthogonal throttler to the underlying prefetch engine. 
-SPP-Perc improves performance by over XX\% over the baseline, which corresponds to YY\% over the next best performing prefetcher. 
-We believe this paper can lead to interesting future work like generalizing SPP-Perc to work with any prefetcher with minimal knowledge of the underlying prefetcher architecture.
-
-%%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%
-
-%%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
-\bibliographystyle{ieeetr}
-\bibliography{ref}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
+Using an adder tree of 4 5-bit adders, this can be done in
+ceil($log_{2}9$) = 4 steps.  Perceptron update only requires weight
+update by +1 or -1.  Thus, all the operations required for perceptron
+inferencing or updating the states of the perceptron throttler can be
+easily done in the time constraints of L2 Cache Accesses.

### 2018-11-15 results.tex
reorg

--- 2018-11-15 paper_1.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is the template for submission to ISCA 2018
-% The cls file is a modified from  'sig-alternate.cls'
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{sig-alternate} 
-\usepackage{mathptmx} % This is Times font
-
-\newcommand{\ignore}[1]{}
-\usepackage{fancyhdr}
-\usepackage[normalem]{ulem}
-\usepackage[hyphens]{url}
-\usepackage{microtype}
-\usepackage{fixltx2e}
-\usepackage{array}
-\usepackage{multirow}
-
-% Always include hyperref last
-\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=black,citecolor=blue,urlcolor=black]{hyperref}
-
-% Ensure letter paper
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\newcommand{\iscasubmissionnumber}{NaN}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\fancypagestyle{firstpage}{
-  \fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-  \fancyhead[C]{\normalsize{}} 
-  \fancyfoot[C]{\thepage}
-}
-
-\pagenumbering{arabic}
-
-%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\title{SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher} 
-\author{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\maketitle
-\thispagestyle{firstpage}
-\pagestyle{plain}
-
-\begin{abstract}
-\textbf{Hardware prefetching has been introduced in modern processors as a way to hide cache latencies. 
-An efficient prefetcher should be able to identify complex memory access patterns during a program execution. 
-This enables it to fetch a block ahead of its demand access, potentially saving a cache miss.
-}
-
-\textbf{In this paper, we introduce perceptron learning to help make this prefetching decision efficiently. 
-We have taken a state of the art prefetcher - Signature Path Prefetcher (SPP) as the underlying prefetch engine. 
-The perceptron layer acts as a throttler to filter out the unnecessary prefetches recommended by SPP. 
-We have also explored a range of features that can be used to train the perceptron layer. 
-Our results show that SPP-Perc improves performance on the memory intensive subset of the SPEC 2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared to just SPP.
-We also demonstrate that the performance gained from using our efficient throttler scales better with increasing number of cores.
-}
-\end{abstract}
-
-\section{Introduction}
-Since past few decades, processor and memory scaling has been enabled with different philosophies in mind. 
-While processor scaling focused on speed improvements, memory scaling was centred around increasing the storage capacity. 
-The process technology reaching scaling limitations led to Memory Wall\cite{MemWall} - the increasing performance gap between the processor speed and the memory speeds. 
-Various techniques have been developed to combat this widening gap. 
-Data prefetching is one such important technique. 
-Data prefetching exploits the fact that in most applications, memory accesses have a well-defined pattern.
-
-An ideal prefetching scheme involves capturing this memory access patterns to predict future prefetches in a timely manner. 
-Other important decisions faced by the prefetcher are when to bring in the prefetch block of data, which level of cache hierarchy to place the block in and finally, what block to evict to accommodate the incoming block.
-Memory accesses patterns can be as simple as fetching the next cache line or a complex but predictable pattern, like in pointer chasing applications. 
-Predicting this often requires a trade-off between coverage and accuracy. 
-For a prefetcher to capture highly irregular access patterns, it needs to be highly aggressive (high coverage). 
-That also means the prefetcher would end up recommending a lot of useless prefetches, leading to cache pollution (low accuracy). 
-This effect would get even more pronounced in multi-core scenario where the last level cache is a shared resource. 
-Conversely, a conservative prefetcher would be highly accurate but might not make enough useful predictions.
-
-To this effect we propose SPP-Perc <come up with a catchy name>. 
-It is an enhancement to the existing state of the art prefetcher and overcomes the coverage vs accuracy trade-off in a robust manner. 
-The idea is to modify SPP design to make it as aggressive as it can get. 
-This enables SPP to capture complex memory access patterns hence leading to increased coverage.
-This comes at a cost of increased DRAM traffic, hence reduction in overall IPC.
-The prefetch suggestions recommended by SPP pass through the perceptron based throttler. 
-Over time, perceptron layer learns to correlate the prefetches recommended by SPP with various available features, leading to elimination of useless prefetches. 
-This leads to an overall increased accuracy of the prefetch scheme and reduction in DRAM traffic.
-
-\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and future scope for expansion. 
-The contributions made by this paper are:
-
-\begin{itemize}
-\item This is the first time where a truly on-line learning based neural network model has been used when it comes to hardware data prefetching. 
-The prior works in this area either relied on program semantics\cite{Semantics} or were application specific\cite{Datacenter} or could not yield a significant improvement over the baseline prefetcher used\cite{BadPerc}.
-
-\item Implementing the pereptron throttler in a robust and accurate practical prefetching mechanism. 
-The overall scheme yields increased accuracy as compared to the other state of the art prefetchers.
-Moreover, the perceptron throttler learns to adapt itself to shared resource constraints - leading to an even increased performance in multi-core and lower capacity / bandwidth environments.
-
-\item Examining the perceptron weights doing a statistical study to quantify the contribution made by each feature.
-We have also used this idea to develop a comprehensive set of features where each feature contributes some `new' information to the perceptron.
-The details are explained under methodology, in Section XX.
-
-\item Cross-validating the design across benchmarks. 
-All the development and parameter tuning was done on the SPEC 2017 benchmark. 
-The results have been shown to be consistent across unseen benchmarks.
-More details can be found in the results section (Section XX).
-
-%\item Designing the prefetcher keeping SPEC 2017 benchmark suite in mind. 
-%We have used SPEC 2017 to compare SPP-Perc against the highest performing prefetchers. 
-%We have also used our prefetcher to characterize memory intensive benchmarks of the SPEC 2017 suite, with respect to prefetching.
-
-%\item We came up with an efficient model to use the perceptron output as a representation of the prefetch confidence using an on-line histogram technique. 
-%We show how the simple sum can replace the existing mechanism which was essentially based on counting the number of hits for a particular entry in SPP.
-
-\end{itemize}
-
-We tested SPP-Perc on Champsim, an open source simulator which models an Out-of-Order CPU using. 
-For evaluation, we used SPEC 2017 benchmarks. 
-On a single core configuration, SPP-Perc increases performance of the system by 39.5\% as compared to the baseline with no prefetching, on the memory intensive subset of the benchmark.
-This outperforms the next best prefetcher, SPP,  by 6.84\%. 
-On a multi-core system running a mix of memory intensive subset of the SPEC 2017 traces, SPP-Perc saw an improvement of 11.9\% over SPP for 4-core system and 10.67\% for 8-core system.
-
-In the results part in section XX, we show a detailed analysis of the speedup achieved.
-We also used SPEC 2006 and CloudSuite benchmarks for cross-validation of the prefetcher implementation.
-
-The rest of the paper is organized as follows.
-Section 2 talks about background work and recaptures the SPP architecture
-Section 3 discusses the detailed implementation of the perceptron mechanism and the features used.
-Section 4 consists of the evaluation methodology and explores the feature space for perceptron learning.
-Section 5 is the performance evaluation of our prefetcher and Section 6 concludes the paper.
-
-\section{Background Work and Motivation}
-
-In this section, we compare the existing work done in the domain of prefetching. 
-We also discuss about some of the interesting memory access patterns seen in the workloads in SPEC 2017 benchmark. 
-The second half of this section deals with a background of Signature Path Prefetcher architecture.
-
-\subsection{Related Work}
-
-The idea of prefetching has been around since Instruction Stream Buffers were proposed by Jouppi.
-\cite{ISB}
-Some of the earliest data prefetchers developed to try to identify memory access patterns with a constant stride pattern.
-\cite{Smith} 
-\cite{Baer}
-\cite{Stride}
-Major research was dedicated to stride distance and depth prediction.
-\cite{Decoupled}
-\cite{Adaptive}
-Newer prefetchers aimed at correlating predictions with the past memory access addresses.
-\cite{Address_Correlated} 
-\cite{AMPM}
-
-With time newer and more intricate prefetchers were introduced.
-\cite{Wenisch_Temporal_Streaming}
-\cite{Stealth}
-\cite{Feedback_Directed}
-\cite{Coordinated}
-\cite{Bandwidth_Efficient}
-\cite{Pacman}
-\cite{TLB}
-\cite{Linearizing}
-\cite{Sandbox}
-\cite{VLDP}
-\cite{DoL}
-\cite{Domino}
-A whole class of streaming prefetchers developed on the lines of Spatial / Temporal Memory Streaming.
-\cite{Spatial_Pattern}
-\cite{SMS}
-\cite{Temporal_Instruction_Fetch}
-\cite{Off_Chip}
-\cite{STMS}
-\cite{SMS_JILP}
-
-Most modern prefetchers aim at identifying complex memory access patterns in a given application. 
-This allows them to capture some of the irregular patterns seen in pointer chasing data-structures.
-Offset based prefetchers are a generalization of the next-line prefetcher.
-Some of the prominent prefetchers exploiting this idea are
-Sandbox Prefetcher \cite{Sandbox}
-and the Best Offset Prefetcher \cite{BOP}
-Another class of prefetchers is lookahead Prefetchers.
-Such prefetchers try to speculate deep into the application's memory access trail.
-These include Path Confidence based Lookahead Prefetching \cite{SPP}
-and Kill the Program Counter \cite{KPC}.
-
-A related line of work also explores cache replacement - insertion policies and dead block prediction, with an eye towards prefetching. 
-\cite{DB_Pred}
-\cite{Cache_Burst}
-\cite{KPC}
-\cite{Harmony}
-Nori \textit{et.al.} introduced prefetching with  respect to runtime criticality.
-\cite{CATCH}
-\newline
-\newline
-\textit{Machine Learning and Prefetching:}
-Peled \textit{et.al.} introduce interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
-\cite{Semantics}
-The biggest challenge here is that it relies on compiler support for getting features to build the context.
-
-The paper by Liao \textit{et. al.} focuses on prefetching for data center applications.
-\cite{Datacenter}
-They use offline Machine Learning Algorithms like SVMs and Logistic Regressions to do parametric search for an optimal prefetcher configuration.
-
-%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?
-Finally, the paper on Data Cache Prefetching with Perceptron Learning\cite{BadPerc} talks about the idea of two step prefetching.
-The first step is an existing baseline prefetcher while the second step is a perceptron based throttler.
-The paper has lots of limitations in terms of design and implementation.
-That reflects in the fact that it did not lead to any significant performance gain over even the basic prefetchers like the Stride Prefetcher\cite{Stride} and the Markov Prefetcher\cite{Markov}.
-
-\subsection{Perceptron Learning in Architecture}
-
-Perceptron learning for computer architecture design has been around for a while. 
-It was first popularized by Jimenez \textit{et. al.} for doing branch prediction. \cite{Perc_Branch} 
-In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
-The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. 
-The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the prediction is made as \textit{true}, or branch taken.
-Else if the sum is < 0, the prediction is made as \textit{false}.
-
-The structure that we use is derived from the model introduced in Piecewise Linear Branch Prediction\cite{Piece_Linear}. 
-Here the feature itself is used to index into a hash of perceptron weights. 
-This form of indexing makes sure that the hyper-plane learned by the perceptron weights is able to differentiate between linearly inseparable outcomes. 
-\textit{<Is it true? Not sure>}
-
-Perceptron Learning for Reuse Prediction\cite{Perc_Reuse} extended the hash perceptron architecture to perform prediction in context of cache replacement. 
-Perceptron layer can learn to correlate the cache replacement behaviour with a variety of features. 
-Each feature has its dedicated table of perceptron weights which it indexes into. 
-Hence number of tables equals the number of features. 
-Once the perceptron weights from all the tables are obtained, they're summed and thresholded based on a pre-decided value. 
-This led to a prediction mechanism that is highly accurate, has quick learning rate and is adaptable to changing program behaviour. 
-The features introduced in that paper were derived from Program Counters of the current and the last few instructions and the memory address of the current access. 
-Even with such basic features, high level of correlation could be established by perceptrons.
-
-The work was further extended by Multiperspective Reuse Prediction\cite{Multiperspective}, which introduced a varied set of parametric features. 
-Although design space exploration for feature and parameter selection was a non-trivial task, once fixed, it could give a cache replacement predictor highly suitable for the given set of applications.
-\newline
-\newline
-\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron prediction involves multiplying the vector of input features: F\textsubscript{1xN}, with the corresponding weight vector: W\textsubscript{Nx1} in a dot product fashion to obtain the sum y\textsubscript{out}.
-Here what we use is a perceptron-like structure. 
-The feature is used to hash into the weight of perceptrons and the retrieved weights are added straight away. 
-This can be seen a dot product of $\vec{1}$\textsubscript{1xN} with W\textsubscript{Nx1}. 
-This way, the perceptron algorithm doesn't introduce any multiplication operations in the inference or training process. 
-Hence what we adapt in this work, is a perceptron-like learning algorithm as it involves the same principle involved in perceptron inferencing and training. 
-For simplicity, we will call our implementation as perceptron implementation.
-\newline
-\newline
-\textit{Perceptron Update Rule:} In all the above implementations, a uniform perceptron update principle is being followed. 
-The weights need to be updated if the prediction was wrong or the predicted sum does not cross a certain threshold. 
-After a certain training period, these weights are proportional to the probability of the outcome being true. 
-This training threshold makes sure that the perceptrons are trained till a certain level of confidence and yet they are not over-trained to the given set. 
-Care is taken that the weights saturate at certain positive and negative values so that they remain confined to the bit-width. 
-The same rules will be applied in the discussion done in section XX on SPP-Perc learning algorithm.
-
-\subsection{Prefetching with SPEC 2017}
-\textit{<Discuss SPEC 2017 traces here>
-<What are the possible graphs / analysis tools we can use for SPEC traces?>}
-
-\subsection{Signature Path Prefetcher Architecture}
-Since the work we have done builds heavily on SPP\cite{SPP}, we are presenting a brief recap of the ideas introduced in the original paper, especially with an eye towards their relevance in our work. 
-Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists of following structural components:
-\newline
-\newline
-\textbf{Signature Table:} ST keeps track of 256 most recently accessed pages. 
-It is meant to capture memory access patterns within a page boundary. 
-SPP indexes into an entry of ST using the page number. 
-For each entry corresponding to a page, ST stores the `last block offset' and `old signature'. 
-Last block offset is the block offset of the last memory access of that given page. 
-The block offset is calculated with respect to the page boundary. 
-The signature is a 12-bit compressed representation of the past few memory accesses for that page. 
-The signature is calculated as:
-$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
-In case a matching page entry is found, the stored signature in the corresponding entry is retrieved.
-\newline
-\newline
-Consider the case that incoming page number 10 with a block offset 3 finds a match in the ST. 
-The retrieved pattern signature is 0x30 and the Last Offset is 1. 
-Since now there is the Last Block Offset (1), Incoming Block Offset (3), Old Signature (0x30), and New Signature calculated as per above equation (0x182), SPP can infer non-speculatively that the given pattern of memory accesses (as captured in the Old Signature) leads to the particular Delta.
-In general, Delta is defined as the difference between the prefetch suggestion block and the initial block which triggered the prefetch. 
-In this case, since SPP is in learning phase, it is defined as the difference between the Incoming Block Offset and the Last Block Offset (+2 in this case). 
-This, in turn generates the new memory pattern (New Signature). 
-This newly learned signature pattern and the delta is stored in the Pattern Table.
-\newline
-\newline
-\textbf{Pattern Table:} The PT is indexed by the signature generated from ST. 
-PT holds predicted delta patterns and their confidence estimates. 
-Each entry indexed by the signature holds up to 4 unique delta predictions. 
-This is implemented by making PT as a 4-way associative table. 
-To measure an estimate of a delta confidence, PT implements two counters: C\textsubscript{sig} for each entry indexed by the signature and C\textsubscript{delta} for each delta entry for a given signature.
-More details of this on-line score-keeping mechanism is described under Confidence Tracking discussion.
-
-\textit{Lookahead Prefetching:} For each trigger to the prefetcher, the first time that the SPP indexes into the PT, it does so in the non-speculative manner. 
-We call it non-speculative because the parameters which SPP uses to predict, like last block offset, incoming block offset and signature are well-defined. 
-Assuming PT generates a valid delta prediction (rather, up to 4 valid deltas), SPP enters into the lookahead mode.
-
-In the lookahead mode, SPP looks at the last generated delta(s), chooses the one with the highest confidence and speculatively generates the new effective address and the new signature. 
-Using these new parameters, it re-accesses the PT to generate further prefetches. 
-SPP keeps on repeating this cycle of accessing PT and updating the signature based on highest confidence prefetch from the last iteration. 
-These iteration count till which SPP manages to predict prefetch entries in the lookahead manner is characterized as its `depth'. 
-While doing this, SPP also keeps compounding the confidence in each depth. 
-Thus as depth increases, overall confidence keeps decreasing. 
-This makes sense since we are going deeper into the speculation path. 
-When the confidence falls below a pre-defined threshold, SPP stops further lookahead iterations
-\newline
-\newline
-\textbf{Global History Register:} One of the innovative ideas introduced in SPP was the Global History Register, which bootstraps learning across page boundaries.
-GHR is a small 8-entry fully associative structure. 
-Whenever a prefetch suggestion from the Pattern Table crosses page boundaries, instead of completely discarding it, GHR saves that information for inter-page learning.
-Since the SPP-Perc implementation in this paper doesn't affect GHR functionality in any ways, we won't be going into more details about GHR.
-\newline
-\newline
-\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch Filter - a 1024 entry 1-way associative table which keeps a record of last few entries prefetched by SPP. 
-This proves useful for following reasons:
-\begin{itemize}
-    \item \textbf{Eliminating duplicate prefetches}\newline 
-    Any subsequent prefetches recommended by SPP mechanism is checked against the existing entries in Prefetch Filter. 
-If a valid entry is found, it means that the particular cache line is already either in L2 Cache or being fetched. 
-Hence there is no need to issue a duplicate prefetch.
-
-    \item \textbf{Updating prefetcher}\newline
-    If a subsequent demand access comes before eviction for a prefetch that was recommended by SPP, various components of SPP are updated to reflect that this was a good prefetch. 
-Correspondingly, if a prefetched cache line is evicted without being demand hit even once, SPP learns that this was a useless prefetch and updates the various tables accordingly
-
-    \item \textbf{Tracking Global Accuracy}\newline
-    PF keeps a track of usefulness of the prefetches by introducing 2 counters: C\textsubscript{total} to track the total number of prefetches getting recorded in the PF and C\textsubscript{useful} to keep a track of which prefetches recorded in the PF led to a demand hit.
-The ratio C\textsubscript{useful} / C\textsubscript{total} is the global accuracy for the prefetching scheme. 
-It is represented as $\alpha$ in the paper. 
-The global accuracy acts a real-time throttler for the path confidence value as depicted in the following discussion.
-\newline
-\end{itemize}
-\textbf{Confidence Tracking}: The PT keeps a track of hits to each signature through a counter C\textsubscript{sig}. 
-The number of hits for a given delta per signature are tracked using a counter C\textsubscript{delta}. 
-The confidence for a given delta is approximated through C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}. 
-When SPP enters into a lookahead mode, the path confidence P\textsubscript{d} is given as:
-$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown in the Prefetch Filter section. 
-The range of global accuracy is in [0,1]. 
-The lower the value of $\alpha$ for a given trace or a program phase, the overall confidence is scaled down and this makes SPP less aggressive for the upcoming prefetches.
-
-Here d is the lookahead depth. 
-For d = 1, when SPP is in non-speculative mode, P\textsubscript{0} can be thought of as 1. 
-Once SPP enters into speculative mode, it calculates the path confidence as a product of confidence of the individual prefetches encountered along that path, further scaled using $\alpha$. 
-The final P\textsubscript{d} is thresholded against prefetching-threshold (T\textsubscript{p}) to decide whether to prefetch or not. 
-If yes, then P\textsubscript{d} is thresholded against a numerically bigger fill-threshold (T\textsubscript{f}) to decide whether the prefetch should be sent to L2 Cache (high confidence prefetch) or Last Level Cache (low confidence prefetch)
-\newline<Insert a comprehensive figure about SPP structures>
-\newline<Insert another detailed picture about SPP data-path flow>
-
-\subsection{Case for an On-line Throttler}
-As compared to some of the other state of the art prefetchers [BOP], SPP is a far less aggressive prefetcher. 
-In the single core environment, this gives BOP an edge as there is no resource contention among the different cores. 
-Hence more aggressive prefetching is bound to prove beneficial. 
-As we increase the core count, we observe that SPP starts outperforming rest of the prefetchers. 
-This can be attributed to the fact that each prefetch suggested by SPP is a carefully calculated one and that prevents cache pollution. 
-\textit{<Figure showing the variation of SPP vs BOP wrt core count>} 
-For 4-core applications, SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
-
-The above analysis shows that with a more careful throttling mechanism, SPP can be tuned to become much more aggressive, leading to increased coverage. 
-The onus of maintaining the accuracy now falls on the independent throttler. 
-To test the hypothesis, we tuned down SPP to the minimum possible threshold. 
-Only by doing this we managed to increase the prefetch suggestions made by SPP, by XX\%. 
-\textit{<Comparison of SPP-Unleased with SPP / BOP>}
-Obviously this came at a cost of increased DRAM traffic and cache pollution.
-
-Moreover, the on-line confidence mechanism introduced in SPP was very rudimentary.
-It was based on taking a ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig} as explained above.
-The same confidence was used to make the decision whether to prefetch or not to prefetch; and which level to prefetch.
-While this approximation was shown to work in the original SPP, we believe that a better form of on-line decision making was possible.
-Hence, it was necessary to build a robust and adaptable learning mechanism to accept / reject the prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level Cache).
-
-To that effect, we introduce an independent on-line perceptron based throttling mechanism.
-
-\section{SPP-Perc}
-
-\subsection{Changes made to original SPP}
-To modify the SPP Design to suit our scheme, the following changes were made:
-
-\begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline
-In the original SPP design, an internal confidence mechanism was developed which was compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the prefetch suggestion.
-Furthermore, the accepted suggestions were thresholded against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache vs Last Level Cache.
-T\textsubscript{p} and T\textsubscript{f} were set to 25 and 90 respectively, on the scale of 0 to 100.
-In SPP-Perc, the perceptron sum is used to make these decisions. 
-Hence these two thresholds are no longer needed
-
-\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
-Original SPP would not consider the slots available in the L2 MSHR queue. 
-It had an assumption that at no time can the number of prefetches suggested by SPP could exceed the capacity of the queue. 
-While it was not proven in the paper, the internal confidence mechanism along with the thresholds made sure that the above assumption was maintained. 
-In SPP-Perc since there can be no assumption on the degree of aggressiveness - because SPP is being tuned to its lowest thresholds, we explicitly check that at no point should the number of suggested prefetches exceed L2 MSHR queue.
-
-\item \textbf{Enhanced Filter}\newline
-Original SPP introduces a concept of Prefetch Filter, which keeps a track of up to 1024 L2 Cache prefetches suggested by SPP. 
-The Filter was modified to store meta-data like program counter, memory access address etc. 
-This is needed to train the perceptron when the result of that particular prefetch gets available. 
-At training time, this data is used to retrieve the state of the prefetcher when the prediction was made.
-
-\item \textbf{Reject Filter introduced}\newline
-The Prefetch Filter mentioned above keeps a track of the prefetches that actually happened in the L2 Cache. 
-Correspondingly, we introduced a reject filter. 
-It keeps a track of the prefetches that were suggested by the base SPP engine but were rejected by the perceptron. 
-In case it is detected that a future demand fetch could have used this rejected prefetch, the perceptron weights are updated to reflect the same
-
-\end{itemize}
-
-\subsection{The Perceptron Layer}
-As discussed earlier, the perceptron layer was introduced to act as a throttler to the underlying SPP engine. 
-This section explores the details of the perceptron layer and the range of features that are ultimately used by the perceptron for correlating the prefetches.
-
-The neural network weights are stored in separate tables for each feature.
-The indexing of each feature is done using different number of bits.
-At most 12 bits of a feature are used to index into the perceptron table.
-Certain features require more resolution power \textit{i.e.}, full 12 bits of indexing.
-On the other hand, some features requiring lower resolution required as low as 7 bits of indexing.
-The variable indexing was determined by studying the features and fine-tuned empirically so as to achieve a good accuracy vs area trade-off.
-Exact details are mentioned under `Area Overheads' in Section XX.
-
-A single entry in the table corresponds to a perceptron weight.
-Each weight is a 5 bit counter - saturating at -16 and +15. 
-Each feature gets its own dedicated table. 
-Our proposed design uses 9 features. 
-As the program execution begins, all the weights are initialized to 0.
-\newline \newline
-\textbf{INFERENCING}\newline
-In the Champsim simulator, the prefetching method is triggered on every L2 Cache demand access. 
-At that point, the prefetcher has the option to do or not do prefetch - in case it does, then how many cache lines to prefetch. 
-The suggested prefetches can be places either in L2 Cache or in Last Level Cache. 
-The prefetcher gets to decide this, and is usually done based on an internal confidence mechanism.
-
-When SPP gets triggered, it suggests up to 4 (=Associativity of the Pattern Table) candidates for prefetch based on the demand access address. 
-In each subsequent lookahead depth, SPP suggests up to 4 more prefetch candidates, as explained in section\_\_. 
-All these recommended prefetches are tested in the perceptron throttler before they can access the Prefetch Filter, as in the original SPP data-path.
-
-The perceptron looks at the micro-architectural state, or `features', at that instant. 
-Each feature indexes into the hash of up to 4096-entry table dedicated for that feature. 
-Feature \#1 indexes into a particular entry in table \#1 and so on.
-
-Once all the weights are retrieved, they are added together, much like the traditional perceptron dot product. 
-On obtaining the overall sum, it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO. 
-Only the prefetch candidates with the perceptron sum higher that the threshold qualify for prefetching. 
-The prefetches that qualify the perceptron throttler stage go through the Prefetch Filter stage, just as in the SPP design.
-The Prefetch Filter eliminates any redundant prefetch suggestions.
-In SPP-Perc, all the meta-data that is required to recreate the features is also stored in the Prefetch Filter. 
-At a later stage, when the feedback of the current prefetched line is available, the stored data can be used to train the perceptron.
-
-The above generated perceptron sum is also used to decide L2 Cache vs Last Level Cache placement of the prefetch cache line. 
-All prefetch candidates qualified till this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill level.
-The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
-
-The perceptron sum can also be seen as the sum of the individual contribution per feature.
-The value of each contribution corresponds to the extent of confidence of the final output with that given feature.
-By summing the individual contributions, the final perceptron sum denotes the overall confidence for that prefetch suggestion. 
-In a way, by thresholding the perceptron sum on two different thresholds, we are diving the confidence scale of the prefetch suggestion into three bins.
-The first bin corresponding to the lowest confidence leads to prefetch candidate being rejected.
-The next bin corresponds to prefetches with a moderate confidence level. 
-Such prefetches are directed to fill the bigger Last Level Cache and potentially not pollute the more scarce L2 Cache.
-The highest confidence prefetches are filled in the L2 Cache.
-
-In addition to the Prefetch Filter mentioned above, SPP-Perc also maintains a `Reject Filter'. 
-Reject filter is a 1024-entry deep 1-way associative table. 
-In case that a prefetch suggestion is rejected by perceptron throttler, it instead gets logged into the Reject Filter. 
-The Reject Filter keeps track of 1024 entries that were suggested by the Pattern Table but got rejected by the perceptron throttler. 
-This can be used to train the perceptron to avoid false negatives \textit{i.e.}, cases where prediction was to reject the prefetch but the prefetch would have been useful.
-
-As previously stated, an enhancement to the original SPP design includes storing all the meta-data describing the program state at the instant of prefetching, in the Prefetch Filter. 
-This information is useful when the perceptron needs to be updated subsequently. 
-\newline \newline
-\textbf{TRAINING}\newline
-In the prefetching environment, feedback of a prefetch is received whenever there is an eviction or a demand access from L2 Cache. 
-This acts as a trigger to update all the baseline SPP entries and to also train the perceptron throttler. 
-Training of the perceptron throttler involves restoring the meta-data that was stored in Prefetch Filter or the Reject Filter. 
-This is done by indexing into the corresponding filters using the cache line address of the block being used to train. (10-bits indexing + 6-bits tag matching). 
-Once the state of the program at the time of prefetching is available, it is used to index into the perceptron weights table.
-
-If a demand access block which triggers the training  was tagged as a valid prefetch in the Prefetch Filter, it means that the earlier prefetch prediction was correct. 
-In that case the perceptron weights are incremented by 1 if the predicted sum does not cross a pre-defined threshold. 
- In case it is a cache block eviction that led to training and the corresponding valid entry was found in the Prefetch Filter, it shows that the prediction made by the perceptron was wrong. 
-Perceptron should have ideally rejected the prefetch suggestion as a low-confidence prefetch. 
-Here the weights are decremented by 1 to reflect the same. 
-In either case, weights are saturated at -16 or +15.
-
-A secondary training mechanism also kicks in during demand fetches. 
-Before the demand access triggers the next set of prefetches, it checks the Reject Filter for the valid entry. 
-In case it is a hit, that means that the corresponding cache line was initially suggested by the underlying SPP engine but rejected by the perceptron throttler. 
-This means that perceptron should have been more confident about that particular prefetch. 
-Once such a scenario is identified, the state of the execution at the time of prefetch is retrieved from the Reject Table. 
-The retrieved data is used to index into the various weights tables of the perceptron and the corresponding values are updated by +1, saturating between and -16 and +15. 
-This update reflects increased confidence for the prediction corresponding to that prefetch.
-
-This mechanism allows us to solve the classical problem of exploiting `lost opportunity'. 
-In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
-Our two-step SPP-Perc allows us to overcome that issue in a simple manner.
-
-\subsection{Features used by Perceptron}
-
-Here we discuss the various features which help correlate the prefetching decision with the program behaviour
-\begin{itemize}
-\item \textbf{Base address} of the demand access which triggered the prefetch. 
-Since prefetch are triggered from the L2 demand access, they tend to be correlated with the triggering address.
-
-\item \textbf{Cache line} and \textbf{Page address}: These two separate features are derived from the base address which triggered the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and base\_addr >> LOG2\_PAGE\_ SIZE respectively.
-The idea behind using three different shifted versions of the same feature is that it allows us to look into a wider range of bits than with a single version.
-It also helps give more importance to the overlapping bits and lesser importance to most and least significance bits.
-This approach can also eliminate destructive interference that can be caused by directly folding the bits into half.
-
-\item \textbf{Confidence}: The integer confidence on the scale of 0 to 100 which was used in the original SPP design.
-While the original confidence is not used directly for decision making, it can still contribute to the final decision made.
-
-\item \textbf{Program Counter XOR Depth}: PC is for the instruction which triggered the prefetch chain.
-Depth refers to the iteration count of the lookahead stages.
-In general, PC is not considered as a good basis for doing lookahead prefetching as all the prefetches with depth >= 1 gets aliased onto the same PC which won't be true in actual demand fetch case.
-This feature resolves a PC into a different value for each lookahead depth of SPP speculation, giving a more accurate correlation in lookahead cases.
-This is akin to the concept of Virtual Program Counters\cite{VPC} introduced by Kim \textit{et. al.} for indirect branch prediction.
-
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC before the instruction that triggered the current prefetch. 
-Hashing together the last three PCs tell SPP-Perc about the path that led to the current demand access and helps capture and branching information of the current basic block.
-PCs are shifted in the increasing order of history before being hashed together.
-This is done to avoid the resultant value of zero in case 2 or more PCs are the same. 
-Additionally, obfuscating the information as it gets old allows us to get a wider and an approximate look into the program history.
-
-\item \textbf{Program Counter XOR Delta}: This tells us if a given PC favours particular value(s) of delta.
-As noted earlier, while PC in itself does not much useful information, this hash resolves PC into different values based on the tendency of that PC to favour a certain delta.
-This way, dynamic nature of different instances of the same memory access instruction can be captured here.
-
-\item \textbf{Current Signature XOR Delta}: It can be recalled from the discussion of SPP in section XX that the new signature is generated using the old signature and the blobk delta.
-While `Current Signature XOR Delta' is not the exact formula for generating the future signature, it gives an approximate idea of the path that the combination of these two values can lead to.
-
-%%%%%%% Below are the features rejected when downsizing 14 -> 9
-%\item \textbf{Current Signature}: The 12-bit signature used by SPP to index into the Pattern Table
-%\item \textbf{Delta}: The signed version of the difference in the block address that triggers the prefetch and the block address of the data being prefetched
-%\item \textbf{PC}: Program Counter of the Load instruction that triggers that particular prefetch
-%\item \textbf{Base address XOR Delta}: If any particular value(s) of delta are highly favoured by a given address which triggered the prefetch, this composite feature can capture that information
-%\item \textbf{Cache line XOR Delta}: This feature captures if any particular delta is highly favoured by a given cache line address
-%\item \textbf{Confidence XOR Delta}: This gives a combined insight into that fact that certain delta-confidence pairs predicted by SPP might correlate more with the prefetch outcome
-%%%%%%%%%%%%%%%%%%%%%%
-
-\end{itemize}
-
-Some of the composite features are derive from simple hashing (XOR) of two primary features.
-There is always a question of usefulness of such composite features and the `fresh' information added.
-We try to justify the choice of each feature by quantifying the contribution made towards predicting prefetch behaviour, in Section XX.
-Finally, as noted above, each feature indexes into its independent entry of perceptron weights.
-
-\section{METHODOLOGY}
-\subsection{Performance Model}
-For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim} simulator. 
-ChampSim is an enhanced version of the framework that was used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} 
-We model model a 1-core / 4-core / 8-core Out of Order machine. The details of the configuration parameters is summarized in Table  \ref{tab:Sim_params}.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|l|p{3.6cm}|c|}
-    \hline
-    Level & Configuration & Access Latency \\
-    \hline
-         L1 Cache & Separate I-cache and D-cache, 32 KB, 8-way & 4 cycles\\
-         L2 Cache & Private, 256 KB & 8 cycles\\
-         LLC & Shared, 2MB / core, 16-way & 20 cycles\\
-         DRAM & 4 GB Single Channel for single-core, 8 GB Double Channel for multi-core & N?? cycles\\
-    \hline
-    \end{tabular}
-    \caption{Memory Model Parameters}
-    \label{tab:Sim_params}
-\end{table}
-
-The block size was fixed 64KB. Prefetching was only triggered at L2 Cache demand accesses but could be sent to L2 Cache or Last Level Cache. 
-No L1 Data level prefetching was done. 
-LRU replacement policy was used on all levels of cache hierarchies. 
-Branch prediction was done using the perceptron branch predictor\cite{Perc_Branch}. 
-Page size was kept as 4KB. 
-Champsim operates all the prefetchers strictly in the physical address space.
-
-\subsection{Testing Under Additional Memory Constraints}
-The default single-core configuration simulates a 2MB LLC and a single channel DRAM with 12.8GB/s bandwidth. 
-We extend the simulations to include memory constraints introduced in DPC-2. 
-Specifically we look at the following two variations:
-\begin{itemize}
-    \item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited to 3.2 GB/s
-    \item \textit{Small LLC}: In this scenario, LLC size is reduced to 512 KB
-\end{itemize}
-All the multi-core simulations are only done in the default configuration.
-
-\subsection{Workloads}
-This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017} has been used to characterize and measure the prefetch performance. 
-We use all the 20 workloads available in the SPEC 2017 Suite. 
-Using SimPoint\cite{SimPoint} methodology, we identified 95 different program segments of 1 Billion instructions each.
-
-We performed cross-validation of our SPP-Perc model using SPEC 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.
-For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
-For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.
-The traces include 4 multi-core applications with 6 distinct phases per application. 
-In total, we used 285 traces representing workloads across 53 applications. 
-\newline
-\newline
-\textit{Single-core performance:} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics.
-We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
-We take the Instruction Per Cycle (IPC) of the prefetcher and divide it with the IPC of baseline. 
-The final number reported is the geometric mean of the speed-up attained on individual traces.
-\newline
-\newline
-\textit{Multi-core performance:} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
-For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 
-Each CPU keeps executing its workload till the last CPU completes 1 Billion instructions after warm-up. 
-For collecting IPC and other data, only the first billion instructions are considered as the ``region of interest''.
-
-Here we report the weighted speedup normalized to baseline \textit{i.e.}, no prefetching. 
-For each of the workloads running on a particular core of the 4-core 8 MB LLC system, we compute IPC\textsubscript{i}. 
-We then find the IPC\_isolated\textsubscript{i} of the same workload running in isolated 1-core 8 MB LLC environment. 
-Then we calculate the total weighted-IPC for a given workload mix as $\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}). 
-For each of the 100 workload-mix, the sum obtained is normalized to the weighted-IPC calculated similarly for baseline case \textit{i.e.}, no prefetching, to get the weighted-IPC-speedup. 
-Finally the geometric mean of these 100 weighted-IPC-speedup is reported as the effective speedup obtained by the prefetching scheme.
-
-We repeat the same process for 8-core workloads with the exception that 20 Million warm-up instructions and 100 Million full instructions were executed.
-This was done so as to keep the simulation run-time within reasonable limits as a single 8-core mix was taking up to 5 days to simulate 1 Billion instructions.
-
-\subsection{Preferchers Simulated}
-We compared SPP-Perc against three of the latest state of the art prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher (SPP).
-BOP was the winner of 2nd Data Prefetching Championship. 
-DA-AMPM is the enhanced version of AMPM, modified to account for DRAM row buffer locality. 
-SPP has been shown to outperform BOP on SPEC 2006 traces. 
-For each of these, we compare their speedups taking no prefetching as the baseline.
-
-\subsection{Developing Features for SPP-Perc}
-This section describes the intuition and analysis that went behind finalizing the features. 
-As noted earlier, we developed a set of 9 features which allow the perceptron throttler to correlate prefetching decision with the program context. 
-To study the correlation across each feature, we study statistically the perceptron weights and try to interpret their distribution.
-\newline  \newline
-\textbf{GLOBAL PEARSON'S CORRELATION}\newline
-For this experiment, we create a dump of the state of perceptron weights at the end of all trace execution.
-The weights obtained from running all the SPEC 2017 traces are concatenated together. 
-Since the weights are collected at the end of individual trace execution, the perceptron weights have attained a relatively stable value by now.
-Hence, this dump represents a `snap' of the trained perceptron weights across all the SPEC 2017 traces.
-The intuition here is that the feature with bulk of the perceptron weights concentrated around 0 or small magnitude numbers show a weak correlation with the prefetching outcome. 
-On the other hand, features with most of the weights saturated around highest value (+15) show a high positive correlation and the features with weights close to the lowest value (-16) show a strong negative correlation.
-
-We plot a histogram for each feature depicting weights distribution from -16 to +15 and use this histogram to generate the Pearson's correlation factor for that feature.
-In statistics, Pearson's factor is a numerical measure of the degree of linear correlation between two variables.
-It ranges from -1 to +1.
-The magnitude of Pearson's factor depicts the extent of correlation and the sign depicts whether it is a positive correlation or a negative correlation.
-Values close to 0 suggest a low correlation or at times, noisy data.
-A value of +1/-1 suggests a perfectly linear positive / negative correlation respectively.
-Figure XX depicts the histogram and the Pearson's coefficient for two of the features, YY and ZZ.
-
-Figure XX shows all the features used, arranged in the increasing order of their Pearson's factor.
-As can be seen X out of the 9 features provide a moderate to high correlation, with the magnitude of P-value > 0.4.
-The single most important feature, \_\_ helps provide a correlation to prefetch outcome with a factor of \_\_.
-\newline \newline
-\textbf{PER TRACE CORRELATION} \newline
-Another important way to look at the perceptron features is to see how much their contribution varies across the traces.
-Here we give special attention to features with low P-values in the previous experiment.
-Figure XX shows the variation of P-values three features : \_\_, \_\_ and \_\_; across all the SPEC 2017 traces.
-For simplicity, the traces are arranged in an increasing order of contribution made by the feature and the trace names have been omitted.
-It can be seen that even features with a low overall correlation provide useful correlation (magnitude > 0.4) for XX out of the 95 simpoints developed for SPEC 2017.
-\newline \newline
-\textbf{TRIMMING FEATURES USING CORRELATION} \newline
-Besides providing interesting insights into prefetching behaviour, P-value can also be used for feature selection and prefetcher tuning.
-
-Here, we introduce the concept of cross-correlation across features.
-Just as we examined correlation of each feature with the final outcome, we can also study correlation between the features.
-We used the above methodology to eliminate features which provide a little or no information that has already been captured in other features.
-
-As a part of feature selection, we initially came up a mix of 23 features- primary and composite.
-By studying cross correlation of each of these features against others in a 23x23 matrix, we identified pairs of features with correlation factor > 0.9 in magnitude and carefully eliminated redundant features.
-Using this approach, we managed to reduce the feature count to 9.
-Thus, in the final implementation of SPP-Perc, no two features have a high correlation between them.
-This way we can be sure that each feature is bringing in contribution that cannot be captured using other features.
-
-\textit{<TODO: VALIDATE THIS>} 
-Secondly, studying the relative importance of each feature enabled us to vary the number of entries dedicated for each feature.
-Features with higher Pearson's correlation (like \_\_ and \_\_) were given most importance and allowed full 12-bits of indexing.
-Features like \_\_ and \_\_, with a low overall P-value were allocated fewer indexes in the feature table.
-
-To conclude, in the above discussion we justify the features for perceptron from a statistical viewpoint.
-We also show how this information can also be used for prefetcher tuning.
-All this study was made possible only because we used on-line perceptron learning for prefetching and that enabled us to examine the weights in detail.
-
-\subsection{Overhead for SPP-Perc}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|c|c|m{4.8cm}|}
-    \hline
-        \textbf{Field} &
-        \textbf{Bits} &
-        \textbf{Comment} \\
-    \hline
-         Valid & 1 & Indicates a valid entry in the filter\\
-         Tag & 6 & Identifier for the entry in the filter\\
-         Useful & 1 & To show if the given entry led to a useful demand fetch\\
-         Perc Decision & 1 & Prefetched vs Not-prefetched \\
-    \hline
-        PC & 12 & \multirow{5}{4.8cm}{Meta-data required for perceptron training}\\
-        Address & 24 & \\
-        Delta & 7 & \\
-        Curr Signature & 12 & \\
-        Confidence & 7 & \\
-    \hline
-        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
-    \hline
-    \end{tabular}
-    \caption{Meta-data Stored in Prefetch Filter}
-    \label{tab:PF_metadata}
-\end{table}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|p{1.5cm}|p{1cm}|c|p{1.5cm}|}
-    \hline
-        \textbf{Structure} &
-        \textbf{Entry} &
-        \textbf{Components} &
-        \textbf{Total} \\
-    \hline
-        \multirow{5}{1.5cm}{Signature Table} & \multirow{5}{1.5cm}{256} & Valid (1 bit) & \multirow{5}{1cm}{11008  bits}\\
-        & & Tag (16 bits) &\\
-        & & Last Offset (6 bits) &\\  
-        & & Signature (12 bits) &\\
-        & & LRU (6 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Pattern Table} & \multirow{3}{1.5cm}{512} & $C_{sig}$ (4bits) &\multirow{3}{1cm}{24576 bits}\\
-        & & $C_{delta}$ (4*4 bits) &\\
-        & & Delta (4*7 bits) &\\
-    \hline
-        \multirow{3}{1.5cm}{Perceptron Tables} & \multirow{3}{1.5cm}{4096*8 128*1} & \multirow{3}{1.5cm}{5 bits} & \multirow{3}{1.5cm}{164480 bits}\\
-        & & &\\
-        & & &\\
-    \hline
-        \multirow{2}{1.5cm}{Prefetch Filter \footnotemark[1]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{84 bits} & \multirow{2}{1.5cm}{86016 bits}\\
-        & & &\\
-    \hline
-     \multirow{2}{1.5cm}{Reject Filter \footnotemark[2]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{83 bits} & \multirow{2}{1.5cm}{84992 bits}\\
-        & & &\\
-    \hline
-        \multirow{4}{1.5cm}{Global History Register} & \multirow{4}{1.5cm}{8} & Signature (12 bits) & \multirow{4}{1.5cm}{264 bits}\\
-        & & Confidence (8 bits) & \\
-        & & Last Offset (6 bits) & \\
-        & & Delta (7 bits) & \\
-    \hline
-        \multirow{2}{1.5cm}{Accuracy Counters} & 1 & C$_{total}$ & 10 bits\\
-        & 1 & C$_{useful}$ & 10 bits\\
-    \hline
-        \multirow{3}{1.5cm}{Global PC Trackers} & \multirow{3}{1cm}{3} & $PC_1$ (12 bits)& \multirow{3}{1.cm}{36 bits}\\
-        & & $PC_2$ (12 bits) &\\
-        & & $PC_3$ (12 bits) &\\
-    \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 371356 bits = 45.33 KB}}\\
-    \hline
-    \end{tabular}
-    \caption{SPP-Perc Storage Overhead}
-    \label{tab:SPPPerc_overhead}
-\end{table}
-\footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}
-\footnotetext[2]{RF does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
-
-In this section, we analyze the hardware overhead required to implement SPP-Perc. 
-The Prefetch Filter was enhanced to accommodate storing of meta-data for perceptron training. 
-Table \ref{tab:PF_metadata} depicts the meta-data stored for each entry in the Prefetch Filter.
-Table \ref{tab:SPPPerc_overhead} shows the total storage overhead of SPP-Perc implementation.
-The hardware budget for 2nd Data Prefetching championship was 32 KB. 
-Keeping that in mind and the speedup SPP-Perc obtained over the winner, the extra hardware budget can be accounted for.
-The extra hardware also makes the overall scheme more scalable than SPP. 
-In the original SPP paper, it was demonstrated that adding extra hardware brings little advantage in terms of performance gain.
-The newly added perceptron tables can be scaled to increase / decrease features depending on the permitted budget.
-
-In terms of computations, the perceptron mechanism only introduces an extra adder tree.
-The hash perceptron mechanism makes sure than there is no actual vector multiplication happening in the hardware.
-Obtaining the perceptron sum requires addition of 9 5-bit numbers. 
-Using an adder tree of 4 5-bit adders, this can be done in ceil($log_{2}9$) = 4 steps.
-Perceptron update only requires weight update by +1 or -1.
-Thus, all the operations required for perceptron inferencing or updating the states of the perceptron throttler can be easily done in the time constraints of L2 Cache Accesses.
-
 \section{RESULTS}
-This section discusses the results obtained from running SPP-Perc, in terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per Cycle (IPC) metrics. 
-For SPEC 2017 suite, first we present the results for single-threaded workloads then the results for multi-core workloads are presented.
+This section discusses the results obtained from running SPP-Perc, in
+terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per
+Cycle (IPC) metrics.  For SPEC 2017 suite, first we present the
+results for single-threaded workloads then the results for multi-core
+workloads are presented.
 
 \subsection{Single-core Results}
-Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated. 
-All the results have been normalized to baseline \textit{i.e.}, no prefetching. 
-As can be seen, SPP-Perc yields a speedup of XX\% over the baseline. 
-This is equivalent to XX\% over DA-AMPM, xx\% over BOP and XX\% over SPP. 
-Out of the XX benchmarks, SPP-Perc nearly matches or outperforms most of the prefetchers on XX traces.
+Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated.  All
+the results have been normalized to baseline \textit{i.e.}, no
+prefetching.  As can be seen, SPP-Perc yields a speedup of XX\% over
+the baseline.  This is equivalent to XX\% over DA-AMPM, xx\% over BOP
+and XX\% over SPP.  Out of the XX benchmarks, SPP-Perc nearly matches
+or outperforms most of the prefetchers on XX traces.
 
-At its peak, SPP-Perc manages to get the IPC speedup of a factor of over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}. 
-This also corresponds to speedup gain of \textbf{22.4\%} over the next best prefetcher - SPP. 
-In general, benchmarks \textit{602.gcc\_s, 603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s} benefit the most from SPP-Perc, as compared to SPP only prefetching.
+At its peak, SPP-Perc manages to get the IPC speedup of a factor of
+over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}.  This also
+corresponds to speedup gain of \textbf{22.4\%} over the next best
+prefetcher - SPP.  In general, benchmarks \textit{602.gcc\_s,
+  603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s}
+benefit the most from SPP-Perc, as compared to SPP only prefetching.
 
-<Discuss the extreme behaviour of \textit{623.xalancbmk\_s}: Improvement over SPP is as high as 31\% and even reduces to -1.5\% on traces of this benchmark>
+<Discuss the extreme behaviour of \textit{623.xalancbmk\_s}:
+Improvement over SPP is as high as 31\% and even reduces to -1.5\% on
+traces of this benchmark>
 
@@
 
-<Discuss about compelling increase in accuracy - details of traces. Plot linear separability>
+<Discuss about compelling increase in accuracy - details of
+traces. Plot linear separability>
 
-<Any case where accuracy got reduced - go into the trace. See if linear separability is the issue. Other possible issues?>
+<Any case where accuracy got reduced - go into the trace. See if
+linear separability is the issue. Other possible issues?>
 
 \subsection{Multi-core Results}
-In this section, we demonstrate the improvement achieved by SPP-Perc for a mix of multi-programmed workloads. 
-\newline \newline
-\textit{4-core environment}
-Figure \_\_ shows a comparison of speedups obtained on workload mix of memory intensive subset of SPEC 2017. 
-We plot all the 4 prefetchers, normalized to baseline. 
-The workloads have been sorted in the increasing order of performance benefit, as measured in terms of normalized weighted speedup. 
-As can be seen, SPP-Perc offers an IPC improvement by 33.5\% on these traces. 
-This is 5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP and DA-AMPM respectively.
+In this section, we demonstrate the improvement achieved by SPP-Perc
+for a mix of multi-programmed workloads.
 
-This increased improvement by SPP-Perc on a 4-core machine is not surprising. 
-Since SPP-Perc has an extremely intelligent throttler working to eliminate useless prefetches before they can create cache pollution. 
-This further amplifies the edge that SPP had over BOP. 
-With the advantage of an orthogonal prefetcher, SPP can be as aggressive as possible and leave the accuracy aspect to perceptron throttler.
-\newline \newline
-\textit{8-core environment}
-To test the effectiveness of SPP-Perc in a tighter constrained multi-core environment, we simulated an 8-core machine with 16 MB last Level Cache. 
-Other parameters were kept same as discussed under Methodology in Section XX.
-<See if we can use speed-up over BOP as a measure here>
+\textit{4-core environment} Figure \_\_ shows a comparison of speedups
+obtained on workload mix of memory intensive subset of SPEC 2017.  We
+plot all the 4 prefetchers, normalized to baseline.  The workloads
+have been sorted in the increasing order of performance benefit, as
+measured in terms of normalized weighted speedup.  As can be seen,
+SPP-Perc offers an IPC improvement by 33.5\% on these traces.  This is
+5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP
+and DA-AMPM respectively.
+
+This increased improvement by SPP-Perc on a 4-core machine is not
+surprising.  Since SPP-Perc has an extremely intelligent throttler
+working to eliminate useless prefetches before they can create cache
+pollution.  This further amplifies the edge that SPP had over BOP.
+With the advantage of an orthogonal prefetcher, SPP can be as
+aggressive as possible and leave the accuracy aspect to perceptron
+throttler.
+
+\textit{8-core environment} To test the effectiveness of SPP-Perc in a
+tighter constrained multi-core environment, we simulated an 8-core
+machine with 16 MB last Level Cache.  Other parameters were kept same
+as discussed under Methodology in Section XX.  <See if we can use
+speed-up over BOP as a measure here>
 
 \subsection{Additional Memory Constraints}
-Figure \_\_ shows the performance of SPP-Perc against the other prefetchers in tighter constraint environments. 
-For sake of simplicity, we show 2 traces where SPP-Perc gives the most advantage over baseline, 2 traces where SPP-Perc doesn't perform as well as the best prefetcher *<-replace with the prefetcher name once results are available* and finally GeoMean across all the traces.
-<Hopefully BOP would degrade more here as it is super aggressive. 
-By definition, SPP-Perc should be better than SPP as the throttler would better control the aggressiveness. 
-In SPP paper DA-AMPM performs close enough to BOP. That should get replicated here too.>
+Figure \_\_ shows the performance of SPP-Perc against the other
+prefetchers in tighter constraint environments.  For sake of
+simplicity, we show 2 traces where SPP-Perc gives the most advantage
+over baseline, 2 traces where SPP-Perc doesn't perform as well as the
+best prefetcher *<-replace with the prefetcher name once results are
+available* and finally GeoMean across all the traces.  <Hopefully BOP
+would degrade more here as it is super aggressive.  By definition,
+SPP-Perc should be better than SPP as the throttler would better
+control the aggressiveness.  In SPP paper DA-AMPM performs close
+enough to BOP. That should get replicated here too.>
 
 \subsection{Cross Validation}
-Here we try to demonstrate the robustness of SPP-Perc by testing it on completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
+Here we try to demonstrate the robustness of SPP-Perc by testing it on
+completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
 
-Fig \_\_ shows the speed-up achieved by BOP, SPP and SPP-Perc on selected traces of SPEC 2006, along with the overall geometric mean over whole of the benchmark and over memory-intensive subset of the SPEC 2006 benchmark.
-As can be seen, SPP-Perc peovides a speed-up of 44.4\% over baseline, on the memory intensive subset of SPEC 2006 benchmark.
-This corresponds to 6.84\% over SPP and XX\% over BOP.
-On the whole of the SPEC 2006 suite, the speedup obtained is 22.4\%, which is 3.34\% ahead of SPP.
+Fig \_\_ shows the speed-up achieved by BOP, SPP and SPP-Perc on
+selected traces of SPEC 2006, along with the overall geometric mean
+over whole of the benchmark and over memory-intensive subset of the
+SPEC 2006 benchmark.  As can be seen, SPP-Perc peovides a speed-up of
+44.4\% over baseline, on the memory intensive subset of SPEC 2006
+benchmark.  This corresponds to 6.84\% over SPP and XX\% over BOP.  On
+the whole of the SPEC 2006 suite, the speedup obtained is 22.4\%,
+which is 3.34\% ahead of SPP.
 
@@
 
-As can be seen, the performance is very consistent given no prior parameter tuning was done keeping SPEC 2006 in mind. 
-We attribute this to the inherent adaptability of the perceptron model.
-In general, perceptron weights are able to adjust in real-time so as to find the best possible correlation between the output and the features.
-
-\section{CONCLUSION}
-In this paper, we introduced perceptron based learning for data prefetching in the form of SPP-Perc. 
-Perceptron acts an orthogonal throttler to the underlying prefetch engine. 
-SPP-Perc improves performance by over XX\% over the baseline, which corresponds to YY\% over the next best performing prefetcher. 
-We believe this paper can lead to interesting future work like generalizing SPP-Perc to work with any prefetcher with minimal knowledge of the underlying prefetcher architecture.
-
-%%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%
-
-%%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
-\bibliographystyle{ieeetr}
-\bibliography{ref}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
+As can be seen, the performance is very consistent given no prior
+parameter tuning was done keeping SPEC 2006 in mind.  We attribute
+this to the inherent adaptability of the perceptron model.  In
+general, perceptron weights are able to adjust in real-time so as to
+find the best possible correlation between the output and the
+features.

### 2018-11-16 back.tex
Incorporating suggestions

@@
 \section{Background Work and Motivation}
+\label{Background}
 
@@
 \subsection{Related Work}
-
+\label{Background-Related}
 The idea of prefetching has been around since Instruction Stream
-Buffers were proposed by Jouppi.  \cite{ISB} Some of the earliest data
+Buffers were proposed by Jouppi~\cite{ISB}. Some of the earliest data
 prefetchers developed to try to identify memory access patterns with a
-constant stride pattern.  \cite{Smith,Baer,Stride} Major research was
-dedicated to stride distance and depth prediction.
-\cite{Decoupled,Adaptive} Newer prefetchers aimed at correlating
-predictions with the past memory access addresses.
-\cite{Address_Correlated,AMPM}
+constant stride pattern~\cite{Smith,Baer,Stride}. Major research was
+dedicated to stride distance and depth prediction~\cite{Decoupled,Adaptive}. Newer prefetchers aimed at correlating
+predictions with the past memory access addresses~\cite{Address_Correlated,AMPM}.
 
@@
 Streaming~\cite{Spatial_Pattern,SMS,Temporal_Instruction_Fetch,Off_Chip,STMS,SMS_JILP}.
+There have been ideas for control flow speculation directed hardware prefetching~\cite{BFetch,MTBFetch}.
 
@@
 prefetcher.  Some of the prominent prefetchers exploiting this idea
-are Sandbox Prefetcher \cite{Sandbox} and the Best Offset Prefetcher
-\cite{BOP} Another class of prefetchers is lookahead Prefetchers.
+are Sandbox Prefetcher~\cite{Sandbox} and the Best Offset Prefetcher~\cite{BOP}. 
+Another class of prefetchers is lookahead Prefetchers.
 Such prefetchers try to speculate deep into the application's memory
 access trail.  These include Path Confidence based Lookahead
-Prefetching \cite{SPP} and Kill the Program Counter \cite{KPC}.
+Prefetching~\cite{SPP} and Kill the Program Counter~\cite{KPC}.
 
@@
 dynamically scaling the magnitude of feedback given to the baseline
-prefetcher.  \cite{Semantics} The biggest challenge here is that it
+prefetcher~\cite{Semantics}. The biggest challenge here is that it
 relies on compiler support for getting features to build the context.
@@
 The paper by Liao \textit{et. al.} focuses on prefetching for data
-center applications.  \cite{Datacenter} They use offline Machine
+center applications~\cite{Datacenter}. They use offline Machine
 Learning Algorithms like SVMs and Logistic Regressions to do
@@
 the fact that it did not lead to any significant performance gain over
-even the basic prefetchers like the Stride Prefetcher\cite{Stride} and
-the Markov Prefetcher\cite{Markov}.
+even the basic prefetchers like the Stride Prefetcher~\cite{Stride} and
+the Markov Prefetcher~\cite{Markov}.
 
 \subsection{Perceptron Learning in Architecture}
-
+\label{Background-Perceptron}
 Perceptron learning for computer architecture design has been around
 for a while.  It was first popularized by Jimenez \textit{et. al.} for
-doing branch prediction. \cite{Perc_Branch} In it's original
+doing branch prediction~\cite{Perc_Branch}. In it's original
 implementation, the Program Counter of the instruction looking for a
@@
 The structure that we use is derived from the model introduced in
-Piecewise Linear Branch Prediction\cite{Piece_Linear}.  Here the
+Piecewise Linear Branch Prediction~\cite{Piece_Linear}.  Here the
 feature itself is used to index into a hash of perceptron weights.
@@
 
-Perceptron Learning for Reuse Prediction\cite{Perc_Reuse} extended the
+Perceptron Learning for Reuse Prediction~\cite{Perc_Reuse} extended the
 hash perceptron architecture to perform prediction in context of cache
@@
 introduced in that paper were derived from Program Counters of the
-current and the last few instructions and the memory address of the
+current and the last few instructions and from the memory address of the
 current access.  Even with such basic features, high level of
@@
 The work was further extended by Multiperspective Reuse
-Prediction\cite{Multiperspective}, which introduced a varied set of
+Prediction~\cite{Multiperspective}, which introduced a varied set of
 parametric features.  Although design space exploration for feature
@@
 confined to the bit-width.  The same rules will be applied in the
-discussion done in section XX on SPP-Perc learning algorithm.
+discussion done in Section \ref{Design-Perceptron} on NATCH learning algorithm.
 
 \subsection{Prefetching with SPEC 2017}
+\label{Background-SPEC2017}
 \textit{<Discuss SPEC 2017 traces here>
@@
 \subsection{Signature Path Prefetcher Architecture}
-Since the work we have done builds heavily on SPP\cite{SPP}, we are
+\label{Background-SPP}
+Since the work we have done uses SPP~\cite{SPP} as the underlying prefetcher, we are
 presenting a brief recap of the ideas introduced in the original
@@
 
-\textbf{Signature Table:} ST keeps track of 256 most recently accessed
+\textbf{Signature Table:} The ST keeps track of 256 most recently accessed
 pages.  It is meant to capture memory access patterns within a page
@@
 
+%% NOTE: Can we do away with this example? 
+%%       If it looks like we are weiting too much about SPP
 Consider the case that incoming page number 10 with a block offset 3
@@
 
-\textit{Lookahead Prefetching:} For each trigger to the prefetcher,
-the first time that the SPP indexes into the PT, it does so in the
-non-speculative manner.  We call it non-speculative because the
-parameters which SPP uses to predict, like last block offset, incoming
-block offset and signature are well-defined.  Assuming PT generates a
-valid delta prediction (rather, up to 4 valid deltas), SPP enters into
-the lookahead mode.
-
-In the lookahead mode, SPP looks at the last generated delta(s),
-chooses the one with the highest confidence and speculatively
-generates the new effective address and the new signature.  Using
-these new parameters, it re-accesses the PT to generate further
-prefetches.  SPP keeps on repeating this cycle of accessing PT and
+\textit{Lookahead Prefetching:} On each trigger, SPP tries to go down
+the program speculation path using its own prefetch suggestion.
+Using current prefetch as the baseline, it re-accesses the PT to generate further
+prefetches.  It keeps on repeating this cycle of accessing PT and
 updating the signature based on highest confidence prefetch from the
@@
 confidence in each depth.  Thus as depth increases, overall confidence
-keeps decreasing.  This makes sense since we are going deeper into the
-speculation path.  When the confidence falls below a pre-defined
-threshold, SPP stops further lookahead iterations.
+keeps decreasing.
 
-\textbf{Global History Register:} One of the innovative ideas
-introduced in SPP was the Global History Register, which bootstraps
-learning across page boundaries.  GHR is a small 8-entry fully
-associative structure.  Whenever a prefetch suggestion from the
+\textbf{Global History Register:} The GHR is a small 8-entry fully
+associative structure to bootstrap learning across page boundaries. 
+Whenever a prefetch suggestion from the
 Pattern Table crosses page boundaries, instead of completely
 discarding it, GHR saves that information for inter-page learning.
-Since the SPP-Perc implementation in this paper doesn't affect GHR
+Since our NATCH implementation in this paper doesn't affect GHR
 functionality in any ways, we won't be going into more details about
@@
 Filter - a 1024 entry 1-way associative table which keeps a record of
-last few entries prefetched by SPP.  This proves useful for following
+last few entries prefetched.  This proves useful for following
 reasons:
@@
 \item \textbf{Eliminating duplicate prefetches}\newline Any subsequent
-  prefetches recommended by SPP mechanism is checked against the
+  prefetches recommended by SPP mechanism are checked against the
   existing entries in Prefetch Filter.  If a valid entry is found, it
@@
 
-\item \textbf{Updating prefetcher}\newline If a subsequent demand
-  access comes before eviction for a prefetch that was recommended by
-  SPP, various components of SPP are updated to reflect that this was
-  a good prefetch.  Correspondingly, if a prefetched cache line is
-  evicted without being demand hit even once, SPP learns that this was
-  a useless prefetch and updates the various tables accordingly
+\item \textbf{Updating prefetcher}\newline Since the PF keeps a track
+of past few prefetches, it uses that information to see if a given
+prefetch led to a demand hit or a cache eviction. Using this information, 
+SPP can update its internal states.
 
 \item \textbf{Tracking Global Accuracy}\newline PF keeps a track of
-  usefulness of the prefetches by introducing 2 counters:
-  C\textsubscript{total} to track the total number of prefetches
-  getting recorded in the PF and C\textsubscript{useful} to keep a
-  track of which prefetches recorded in the PF led to a demand hit.
-  The ratio C\textsubscript{useful} / C\textsubscript{total} is the
-  global accuracy for the prefetching scheme.  It is represented as
-  $\alpha$ in the paper.  The global accuracy acts a real-time
-  throttler for the path confidence value as depicted in the following
-  discussion.  \newline
+  global usefulness ($\alpha$) of the prefetches by calculating 
+  C\textsubscript{useful} / C\textsubscript{total}, where
+  C\textsubscript{total} is the total number of prefetches
+  getting recorded in the PF and C\textsubscript{useful} is the number 
+  of prefetches that led to a demand hit.
 \end{itemize}
+
 \textbf{Confidence Tracking}: The PT keeps a track of hits to each
@@
 confidence P\textsubscript{d} is given as:
-$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown in the Prefetch Filter section. 
-The range of global accuracy is in [0,1].  The lower the value of
-$\alpha$ for a given trace or a program phase, the overall confidence
-is scaled down and this makes SPP less aggressive for the upcoming
-prefetches.
+$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown previously. 
+The range of global accuracy ($\alpha$) is [0,1].
 
 Here d is the lookahead depth.  For d = 1, when SPP is in
-non-speculative mode, P\textsubscript{0} can be thought of as 1.  Once
-SPP enters into speculative mode, it calculates the path confidence as
-a product of confidence of the individual prefetches encountered along
-that path, further scaled using $\alpha$.  The final
-P\textsubscript{d} is thresholded against prefetching-threshold
+non-speculative mode, P\textsubscript{0} can be thought of as 1. 
+The final P\textsubscript{d} is thresholded against prefetching-threshold
 (T\textsubscript{p}) to decide whether to prefetch or not.  If yes,
@@
 \subsection{Case for an On-line Throttler}
+\label{Background-Case}
 As compared to some of the other state of the art prefetchers [BOP],
-SPP is a far less aggressive prefetcher.  In the single core
+SPP is a less aggressive prefetcher.  In the single core
 environment, this gives BOP an edge as there is no resource contention
@@
 The above analysis shows that with a more careful throttling
-mechanism, SPP can be tuned to become much more aggressive, leading to
+mechanism, any prefetcher like SPP can be tuned to become much more aggressive, leading to
 increased coverage.  The onus of maintaining the accuracy now falls on
@@
 not to prefetch; and which level to prefetch.  While this
-approximation was shown to work in the original SPP, we believe that a
-better form of on-line decision making was possible.  Hence, it was
+approximation was shown to work in the original implementation, we believe that a
+better form of generalised on-line decision making was possible.  
+Hence, it was
 necessary to build a robust and adaptable learning mechanism to accept

### 2018-11-16 design.tex
Incorporating suggestions

@@
-\section{SPP-Perc}
+\section{NATCH Architecture}
+\label{Design}
+
+In this section we describe our implementation of NATCH.
+NATCH is a generalised prefetch throttler and can be adapted to 
+any existing or upcoming prefetcher with minimal modifications.
+For a practical implementation, we have used SPP as the base engine prefetcher.
 
 \subsection{Changes made to original SPP}
-To modify the SPP Design to suit our scheme, the following changes were made:
+\label{Design-Changes}
+To modify the SPP design to suit our scheme, the following changes were made:
 
@@
 \item \textbf{Original Thresholds discarded}\newline In the original
-  SPP design, an internal confidence mechanism was developed which was
+  SPP design, the internal confidence was
   compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to
@@
   set to 25 and 90 respectively, on the scale of 0 to 100.  In
-  SPP-Perc, the perceptron sum is used to make these decisions.  Hence
+  NATCH, the perceptron sum is used to make these decisions.  Hence
   these two thresholds are no longer needed
 
+%%[NOTE] Can this point be removed?
 \item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
@@
   mechanism along with the thresholds made sure that the above
-  assumption was maintained.  In SPP-Perc since there can be no
-  assumption on the degree of aggressiveness - because SPP is being
-  tuned to its lowest thresholds, we explicitly check that at no point
+  assumption was maintained.  In NATCH since there can be no
+  assumption on the degree of aggressiveness of the underlying prefetcher, 
+  we explicitly check that at no point
   should the number of suggested prefetches exceed L2 MSHR queue.
@@
   concept of Prefetch Filter, which keeps a track of up to 1024 L2
-  Cache prefetches suggested by SPP.  The Filter was modified to store
+  Cache prefetches suggestions.  For NATCH, the Filter was modified to store
   meta-data like program counter, memory access address etc.  This is
@@
 \subsection{The Perceptron Layer}
+\label{Design-Perceptron}
 As discussed earlier, the perceptron layer was introduced to act as a
-throttler to the underlying SPP engine.  This section explores the
+throttler to the underlying prefetchh engine.  This section explores the
 details of the perceptron layer and the range of features that are
@@
 vs area trade-off.  Exact details are mentioned under `Area Overheads'
-in Section XX.
+in Section \ref{Method-Overheads}
 
@@
 
-When SPP gets triggered, it suggests up to 4 (=Associativity of the
-Pattern Table) candidates for prefetch based on the demand access
-address.  In each subsequent lookahead depth, SPP suggests up to 4
-more prefetch candidates, as explained in section\_\_.  All these
+When the base prefetcher gets triggered, it starts suggesting candidates for prefetching. 
+All these
 recommended prefetches are tested in the perceptron throttler before
@@
 threshold qualify for prefetching.  The prefetches that qualify the
-perceptron throttler stage go through the Prefetch Filter stage, just
-as in the SPP design.  The Prefetch Filter eliminates any redundant
-prefetch suggestions.  In SPP-Perc, all the meta-data that is required
+perceptron throttler stage go through the Prefetch Filter stage.  
+The Prefetch Filter eliminates any redundant
+prefetch suggestions.  In NATCH, all the meta-data that is required
 to recreate the features is also stored in the Prefetch Filter.  At a
@@
 
-In addition to the Prefetch Filter mentioned above, SPP-Perc also
+In addition to the Prefetch Filter mentioned above, NATCH also
 maintains a `Reject Filter'.  Reject filter is a 1024-entry deep 1-way
@@
 
-As previously stated, an enhancement to the original SPP design
-includes storing all the meta-data describing the program state at the
-instant of prefetching, in the Prefetch Filter.  This information is
-useful when the perceptron needs to be updated subsequently.
+As previously stated all the meta-data describing the program state at the
+instant of prefetching, is stored in the Prefetch Filter or the Reject Filter.  
+This information is useful when the perceptron needs to be updated subsequently.
 
@@
 access from L2 Cache.  This acts as a trigger to update all the
-baseline SPP entries and to also train the perceptron throttler.
+base prefetcher entries and to also train the perceptron throttler.
 Training of the perceptron throttler involves restoring the meta-data
@@
 that means that the corresponding cache line was initially suggested
-by the underlying SPP engine but rejected by the perceptron throttler.
+by the underlying prefetch engine but rejected by the perceptron throttler.
 This means that perceptron should have been more confident about that
@@
 general, prefetching algorithms), there is usually no way of knowing
-the result of not prefetch a particular line.  Our two-step SPP-Perc
-allows us to overcome that issue in a simple manner.
+the result of not prefetch a particular line.  Our two-step NATCH
+architecture allows us to overcome that issue in a simple manner.
 
 \subsection{Features used by Perceptron}
-
+\label{Design-Features}
 Here we discuss the various features which help correlate the
@@
 \begin{itemize}
-\item \textbf{Base address} of the demand access which triggered the
+\item \textbf{Base Address} of the demand access which triggered the
   prefetch.  Since prefetch are triggered from the L2 demand access,
@@
 
-\item \textbf{Cache line} and \textbf{Page address}: These two
+\item \textbf{Cache Line} and \textbf{Page Address}: These two
   separate features are derived from the base address which triggered
@@
 
-\item \textbf{Confidence}: The integer confidence on the scale of 0 to
-  100 which was used in the original SPP design.  While the original
-  confidence is not used directly for decision making, it can still
-  contribute to the final decision made.
-
 \item \textbf{Program Counter XOR Depth}: PC is for the instruction
@@
   actual demand fetch case.  This feature resolves a PC into a
-  different value for each lookahead depth of SPP speculation, giving
+  different value for each lookahead depth of prefetch speculation, giving
   a more accurate correlation in lookahead cases.  This is akin to the
@@
   PC before the instruction that triggered the current prefetch.
-  Hashing together the last three PCs tell SPP-Perc about the path
+  Hashing together the last three PCs tell NATCH about the path
   that led to the current demand access and helps capture and
@@
 
+\item \textbf{Confidence}: The integer confidence on the scale of 0 to
+  100 which was used in the original SPP design.  While the original
+  confidence is not used directly for decision making, it can still
+  contribute to the final decision made.
+
 \item \textbf{Current Signature XOR Delta}: It can be recalled from
-  the discussion of SPP in section XX that the new signature is
+  the discussion of SPP in Section \ref{Background-SPP} that the new signature is
   generated using the old signature and the blobk delta.  While
@@
   path that the combination of these two values can lead to.
+
+\item \textbf{Page Address XOR Confidence}: This feature helps look at the 
+tendency of each page to be prefetch friendly or prefetch averse. It 
+helps resolve a page into different entries depending on its confidence 
+for prefetching, which can vary during phases of a program execution.
 
@@
 justify the choice of each feature by quantifying the contribution
-made towards predicting prefetch behaviour, in Section XX.  Finally,
+made towards predicting prefetch behaviour, in Section \ref{Method-Features}.  Finally,
 as noted above, each feature indexes into its independent entry of
 perceptron weights.
+
+\subsection{Generalizing NATCH for any Prefetcher}
+\label{Design-Generalizing}
+The above implementation of NATCH suggests that it is highly modular and can be 
+adapted to be used over any base prefecher for increased prefetch accuracy. 
+In our implementation, 
+only two `hooks' are required between NATCH and the baseline SPP. First  
+is to make sure that all the prefetch candidates of SPP pass through 
+the perceptron throttler and if qualified, the metadata for perceptron indexing
+be stored. Second is needed when the feedback of a prior prefetch is available in form 
+of a subsequent demand hit or cache eviction. In that case, the stored meta-data needs 
+to be retrieved to update the state of the perceptrons.
+
+In general, NATCH can be adapted to a new base prefetcher with only a few modifications.
+\begin{itemize}
+
+\item \textbf{Enhancing the Base Prefetcher:} By tuning down any internal thresholds 
+to increase its inherent aggressiveness.
+
+\item \textbf{Inferencing and Storing:} All prefetch recommendations to be tested 
+using the perceptron inferencing algorithm.
+Perceptron output: \textit{true} and \textit{false} should be saved appropriately, along with all 
+the meta-data required for perceptron indexing.
+
+\item \textbf{Retrieving and Training:} When the feedback is available, the previously 
+stored meta-data can be used to re-index into the perceptron entries and increment / decrement 
+the weights.
+
+\item \textbf{Feature Selection:} Six out of the nine features we developed use information derived 
+from program execution, irrespective of the baseline SPP prefetcher. Beyond that, the feature set 
+can be expanded to convey any useful information from the prefetcher to the perceptron throttler.
+For example, in our implementation 'Confidence', 'Current Signature XOR Delta', and 
+'Page Address XOR Confidence' are such features.
+
+\end{itemize}

### 2018-11-16 intro.tex
Incorporating suggestions

@@
-\section{Introduction}
+\section{INTRODUCTION}
+\label{Introduction}
 Since past few decades, processor and memory scaling has been enabled
@@
 
-To this effect we propose SPP-Perc <come up with a catchy name>.  It
+To this effect we propose NATCH: NeurAl prefeTCh tHrottler.  It
 is an enhancement to the existing state of the art prefetcher and
 overcomes the coverage vs accuracy trade-off in a robust manner.  The
-idea is to modify SPP design to make it as aggressive as it can get.
-This enables SPP to capture complex memory access patterns hence
+idea is to use a state-of-the-art prefetcher as the base engine.
+For that we choose Signature Path Prefetcher (SPP) and 
+modify the design to make it as aggressive as it can get.
+This enables the base engine to capture complex memory access patterns hence
 leading to increased coverage.  This comes at a cost of increased DRAM
@@
 recommended by SPP pass through the perceptron based throttler. Over
-time, perceptron layer learns to correlate the prefetches recommended
-by SPP with various available features, leading to elimination of
+time, perceptron layer learns to correlate the prefetch recommendations 
+with various available features, leading to elimination of
 useless prefetches. This leads to an overall increased accuracy of
@@
 
-\vspace{1ex}This paper talks about SPP-Perc, its merits, analysis and
+\vspace{1ex}This paper talks about NATCH, its merits, analysis and
 future scope for expansion.  The contributions made by this paper are:
@@
   feature contributes some `new' information to the perceptron.  The
-  details are explained under methodology, in Section XX.
+  details are explained under methodology, in Section \ref{Method-Overheads}.
 
@@
   benchmark.  The results have been shown to be consistent across
-  unseen benchmarks.  More details can be found in the results section
-  (Section XX).
-
-  % \item Designing the prefetcher keeping SPEC 2017 benchmark suite
-  %   in mind.  We have used SPEC 2017 to compare SPP-Perc against the
-  %   highest performing prefetchers.  We have also used our
-  %   prefetcher to characterize memory intensive benchmarks of the
-  %   SPEC 2017 suite, with respect to prefetching.
-
-  % \item We came up with an efficient model to use the perceptron
-  %   output as a representation of the prefetch confidence using an
-  %   on-line histogram technique.  We show how the simple sum can
-  %   replace the existing mechanism which was essentially based on
-  %   counting the number of hits for a particular entry in SPP.
+  unseen benchmarks.  More details can be found in Section \ref{Results-CrossVal}.
 
@@
 
-We tested SPP-Perc on Champsim, an open source simulator which models
+We tested NATCG on Champsim, an open source simulator which models
 an Out-of-Order CPU using.  For evaluation, we used SPEC 2017
-benchmarks.  On a single core configuration, SPP-Perc increases
+benchmarks.  On a single core configuration, NATCH increases
 performance of the system by 39.5\% as compared to the baseline with
@@
 system running a mix of memory intensive subset of the SPEC 2017
-traces, SPP-Perc saw an improvement of 11.9\% over SPP for 4-core
+traces, NATCH saw an improvement of 11.9\% over SPP for 4-core
 system and 10.67\% for 8-core system.
 
-In the results part in section XX, we show a detailed analysis of the
+In the results part in Section \ref{Results}, we show a detailed analysis of the
 speedup achieved.  We also used SPEC 2006 and CloudSuite benchmarks
@@
 
-The rest of the paper is organized as follows.  Section 2 talks about
-background work and recaptures the SPP architecture Section 3
+The rest of the paper is organized as follows.  Section \ref{Background} talks about
+background work and recaptures the SPP architecture Section \ref{Design}
 discusses the detailed implementation of the perceptron mechanism and
-the features used.  Section 4 consists of the evaluation methodology
-and explores the feature space for perceptron learning.  Section 5 is
-the performance evaluation of our prefetcher and Section 6 concludes
+the features used.  Section \ref{Method} consists of the evaluation methodology
+and explores the feature space for perceptron learning.  Section \ref{Results} is
+the performance evaluation of our prefetcher and Section \ref{Conclusion} concludes
 the paper.

### 2018-11-16 method.tex
Incorporating suggestions

@@
 \section{METHODOLOGY}
+\label{Method}
+
 \subsection{Performance Model}
-For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim}
+\label{Method-Model}
+For testing and comparing NATCH, we used ChampSim\cite{Champsim}
 simulator.  ChampSim is an enhanced version of the framework that was
@@
 \subsection{Testing Under Additional Memory Constraints}
+\label{Method-AdditionalMem}
 The default single-core configuration simulates a 2MB LLC and a single
@@
 \subsection{Workloads}
+\label{Method-Workloads}
 This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017}
@@
 
-We performed cross-validation of our SPP-Perc model using SPEC
+We performed cross-validation of our NATCH model using SPEC
 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.  For
@@
 \subsection{Preferchers Simulated}
-We compared SPP-Perc against three of the latest state of the art
-prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map
+\label{Method-Prefetchers}
+We compared NATCH against three of the latest state of the art
+hardware-only prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map
 Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher
@@
 
-\subsection{Developing Features for SPP-Perc}
+\subsection{Developing Features for NATCH}
+\label{Method-Features}
 This section describes the intuition and analysis that went behind
@@
 managed to reduce the feature count to 9.  Thus, in the final
-implementation of SPP-Perc, no two features have a high correlation
+implementation of NATCH, no two features have a high correlation
 between them.  This way we can be sure that each feature is bringing
@@
 
-\subsection{Overhead for SPP-Perc}
+\subsection{Overhead for NATCH}
+\label{Method-Overheads}
 
@@
     \end{tabular}
-    \caption{SPP-Perc Storage Overhead}
-    \label{tab:SPPPerc_overhead}
+    \caption{NATCH Storage Overhead}
+    \label{tab:NATCH_overhead}
 \end{table}
@@
 In this section, we analyze the hardware overhead required to
-implement SPP-Perc.  The Prefetch Filter was enhanced to accommodate
+implement NATCH.  The Prefetch Filter was enhanced to accommodate
 storing of meta-data for perceptron training.  Table
 \ref{tab:PF_metadata} depicts the meta-data stored for each entry in
-the Prefetch Filter.  Table \ref{tab:SPPPerc_overhead} shows the total
-storage overhead of SPP-Perc implementation.  The hardware budget for
-2nd Data Prefetching championship was 32 KB.  Keeping that in mind and
-the speedup SPP-Perc obtained over the winner, the extra hardware
+the Prefetch Filter.  Table \ref{tab:NATCH_overhead} shows the total
+storage overhead of NATCH implementation.  The hardware budget for
+2nd Data Prefetching championship was 32 KB.  Keeping that in mind 
+the considerable speedup NATCH obtained over the winner, the extra hardware
 budget can be accounted for.  The extra hardware also makes the

### 2018-11-17 design.tex
Aligning Table 3

@@
   a more accurate correlation in lookahead cases.  This is akin to the
-  concept of Virtual Program Counters\cite{VPC} introduced by Kim
+  concept of Virtual Program Counters~\cite{VPC} introduced by Kim
   \textit{et. al.} for indirect branch prediction.
@@
   favours particular value(s) of delta.  As noted earlier, while PC in
-  itself does not much useful information, this hash resolves PC into
+  itself does not convey much useful information, this hash resolves PC into
   different values based on the tendency of that PC to favour a
@@
 can be expanded to convey any useful information from the prefetcher to the perceptron throttler.
-For example, in our implementation 'Confidence', 'Current Signature XOR Delta', and 
+For example, in our implementation `Confidence', `Current Signature XOR Delta', and 
 'Page Address XOR Confidence' are such features.

### 2018-11-17 intro.tex
Aligning Table 3

@@
 the storage capacity.  The process technology reaching scaling
-limitations led to Memory Wall\cite{MemWall} - the increasing
+limitations led to Memory Wall~\cite{MemWall} - the increasing
 performance gap between the processor speed and the memory speeds.
@@
   prefetching.  The prior works in this area either relied on program
-  semantics\cite{Semantics} or were application
-  specific\cite{Datacenter} or could not yield a significant
-  improvement over the baseline prefetcher used\cite{BadPerc}.
+  semantics~\cite{Semantics} or were application
+  specific~\cite{Datacenter} or could not yield a significant
+  improvement over the baseline prefetcher used~\cite{BadPerc}.
 
@@
 
-\item Examining the perceptron weights doing a statistical study to
+\item Examining the perceptron weights by doing a statistical study to
   quantify the contribution made by each feature.  We have also used
@@
 
-We tested NATCG on Champsim, an open source simulator which models
+We tested NATCH on Champsim, an open source simulator which models
 an Out-of-Order CPU using.  For evaluation, we used SPEC 2017

### 2018-11-18 abs.tex

@@
 \begin{abstract}
-  \textbf{Hardware prefetching has been introduced in modern
+Hardware prefetching has been introduced in modern
 processors as a way to hide cache latencies.  An efficient
@@
 block ahead of its demand access, potentially saving a cache miss.
-  }
 
-  \textbf{In this paper, we introduce perceptron learning to help make
-    this prefetching decision efficiently.  We have taken a state of
-    the art prefetcher - Signature Path Prefetcher (SPP) as the
-    underlying prefetch engine.  The perceptron layer acts as a
-    throttler to filter out the unnecessary prefetches recommended by
-    SPP.  We have also explored a range of features that can be used
-    to train the perceptron layer.  Our results show that SPP-Perc
+In this paper, we introduce perceptron learning to help make
+this prefetching decision efficiently. The perceptron layer acts 
+as a throttler to filter out the unnecessary prefetches recommended 
+by the underlying prefetcher.  We have also explored a range 
+of features that can be used to train the perceptron layer.  
+Our results show that NATCH
 improves performance on the memory intensive subset of the SPEC
 2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on
-    multi-core traces, as compared to just SPP.  We also demonstrate
-    that the performance gained from using our efficient throttler
-    scales better with increasing number of cores.  }
+multi-core traces, as compared to state-of-the art prefetcher.  
+We also demonstrate that the performance gained from using 
+our efficient throttler scales better with increasing number 
+of cores. 
 \end{abstract}

### 2018-11-18 conc.tex

@@
 \section{CONCLUSION}
-In this paper, we introduced perceptron based learning for data
-prefetching in the form of SPP-Perc.  Perceptron acts an orthogonal
-throttler to the underlying prefetch engine.  SPP-Perc improves
+\label{Conclusion}
+In this paper, we introduced neural network based learning for data
+prefetching in the form of NATCH.  Perceptron acts an orthogonal
+throttler to the underlying prefetch engine.  NATCH improves
 performance by over XX\% over the baseline, which corresponds to YY\%
-over the next best performing prefetcher.  We believe this paper can
-lead to interesting future work like generalizing SPP-Perc to work
-with any prefetcher with minimal knowledge of the underlying
-prefetcher architecture.
+over the next best performing prefetcher.  
+We believe that this is a robust and adaptable technique which can
+be used to enhance any existing or upcoming prefetchers.

### 2018-11-18 method.tex

@@
 \label{Method-Model}
-For testing and comparing NATCH, we used ChampSim\cite{Champsim}
+For testing and comparing NATCH, we used ChampSim~\cite{Champsim}
 simulator.  ChampSim is an enhanced version of the framework that was
-used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} We
+used for the 2nd Data Prefetch Championship (DPC-2)~\cite{DPC_2}. We
 model model a 1-core / 4-core / 8-core Out of Order machine. The
@@
 was used on all levels of cache hierarchies.  Branch prediction was
-done using the perceptron branch predictor\cite{Perc_Branch}.  Page
+done using the perceptron branch predictor~\cite{Perc_Branch}.  Page
 size was kept as 4KB.  Champsim operates all the prefetchers strictly
@@
 \label{Method-Workloads}
-This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017}
+This is the first time that SPEC 2017 benchmark suite~\cite{SPEC2017}
 has been used to characterize and measure the prefetch performance.
 We use all the 20 workloads available in the SPEC 2017 Suite.  Using
-SimPoint\cite{SimPoint} methodology, we identified 95 different
+SimPoint~\cite{SimPoint} methodology, we identified 95 different
 program segments of 1 Billion instructions each.
@@
 We performed cross-validation of our NATCH model using SPEC
-2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.  For
+2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks.  For
 SPEC 2006, we developed 94 simpoints spread across the 29
 applications.  For CloudSuite, we used the traces made available for
-the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.  The traces
+the 2nd Cache Replacement Competition (CRC-2)~\cite{CRC_2}.  The traces
 include 4 multi-core applications with 6 distinct phases per
@@
 hardware-only prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map
-Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher
+Pattern Matching (DA-AMPM)~\cite{DA_AMPM} and Signature Path Prefecher
 (SPP).  BOP was the winner of 2nd Data Prefetching Championship.
@@
     \centering
-    \begin{tabular}{|p{1.5cm}|p{1cm}|c|p{1.5cm}|}
+    \begin{tabular}{|c|c|c|c|}
     \hline
@@
     \hline
-        \multirow{5}{1.5cm}{Signature Table} & \multirow{5}{1.5cm}{256} & Valid (1 bit) & \multirow{5}{1cm}{11008  bits}\\
-        & & Tag (16 bits) &\\
-        & & Last Offset (6 bits) &\\  
-        & & Signature (12 bits) &\\
-        & & LRU (6 bits) &\\
+        \multirow{5}{2.2cm}{Signature Table} &      & Valid (1 bit)        &             \\
+                                             &      & Tag (16 bits)        &             \\
+                                             & 256  & Last Offset (6 bits) & 11008 bits  \\  
+                                             &      & Signature (12 bits)  &             \\
+                                             &      & LRU (6 bits)         &             \\
     \hline
-        \multirow{3}{1.5cm}{Pattern Table} & \multirow{3}{1.5cm}{512} & $C_{sig}$ (4bits) &\multirow{3}{1cm}{24576 bits}\\
-        & & $C_{delta}$ (4*4 bits) &\\
-        & & Delta (4*7 bits) &\\
+        \multirow{3}{2.2cm}{~~Pattern Table} &      & $C_{sig}$ (4bits)      &               \\
+                                             & 512  & $C_{delta}$ (4*4 bits) & 24576 bits    \\
+                                             &      & Delta (4*7 bits)       &               \\
     \hline
-        \multirow{3}{1.5cm}{Perceptron Tables} & \multirow{3}{1.5cm}{4096*8 128*1} & \multirow{3}{1.5cm}{5 bits} & \multirow{3}{1.5cm}{164480 bits}\\
-        & & &\\
-        & & &\\
+        \multirow{3}{1.5cm}{Perceptron\newline}     & 4096*8    &            &              \\
+        \multirow{2}{0.9cm}{Tables}                 & 2048*x    & 5 bits     & 164480 bits  \\
+                                                    & 128*1     &            &              \\
     \hline
-        \multirow{2}{1.5cm}{Prefetch Filter \footnotemark[1]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{84 bits} & \multirow{2}{1.5cm}{86016 bits}\\
-        & & &\\
+        Prefetch Filter\footnotemark[1]             & 1024      & 84 bits    & 86016 bits   \\
     \hline
-     \multirow{2}{1.5cm}{Reject Filter \footnotemark[2]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{83 bits} & \multirow{2}{1.5cm}{84992 bits}\\
-        & & &\\
+        Reject Filter\footnotemark[2]               & 1024      & 83 bits    & 84992 bits   \\
     \hline
-        \multirow{4}{1.5cm}{Global History Register} & \multirow{4}{1.5cm}{8} & Signature (12 bits) & \multirow{4}{1.5cm}{264 bits}\\
-        & & Confidence (8 bits) & \\
-        & & Last Offset (6 bits) & \\
+        \multirow{4}{1cm}{Global\newline\newline}   & \multirow{4}{0.2cm}{8} & Signature (12 bits)  & \multirow{4}{1.1cm}{264 bits} \\
+        \multirow{3}{1.1cm}{History\newline}        &                        & Confidence (8 bits)  &                               \\
+        \multirow{2}{1.2cm}{Register}               &                        & Last Offset (6 bits) &                               \\
                                                     &                        & Delta (7 bits)       &                               \\
     \hline
-        \multirow{2}{1.5cm}{Accuracy Counters} & 1 & C$_{total}$ & 10 bits\\
-        & 1 & C$_{useful}$ & 10 bits\\
+        Accuracy        & 1     & C$_{total}$       & 10 bits   \\
+        Counters        & 1     & C$_{useful}$      & 10 bits   \\
     \hline
-        \multirow{3}{1.5cm}{Global PC Trackers} & \multirow{3}{1cm}{3} & $PC_1$ (12 bits)& \multirow{3}{1.cm}{36 bits}\\
-        & & $PC_2$ (12 bits) &\\
-        & & $PC_3$ (12 bits) &\\
+        \multirow{3}{1.5cm}{Global PC\newline}      &       & $PC_1$ (12 bits)      &           \\
+        \multirow{2}{1.5cm}{~Trackers}              & 3     & $PC_2$ (12 bits)      & 36 bits   \\
+                                                    &       & $PC_3$ (12 bits)      &           \\
     \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 371356 bits = 45.33 KB}}\\
+        \multicolumn{4}{|c|}{\textbf{Total: 371392 bits = 45.33 KB}}\\
     \hline
     \end{tabular}
-    \caption{NATCH Storage Overhead}
-    \label{tab:NATCH_overhead}
+    \caption{SPP-Perc Storage Overhead}
+    \label{tab:SPPPerc_overhead}
 \end{table}
+
 \footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}

### 2018-11-18 results.tex

@@
 \section{RESULTS}
-This section discusses the results obtained from running SPP-Perc, in
+\label{Results}
+
+This section discusses the results obtained from running NATCH, in
 terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per
@@
 \subsection{Single-core Results}
+\label{Results-Single}
 Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated.  All
 the results have been normalized to baseline \textit{i.e.}, no
-prefetching.  As can be seen, SPP-Perc yields a speedup of XX\% over
+prefetching.  As can be seen, NATCH yields a speedup of XX\% over
 the baseline.  This is equivalent to XX\% over DA-AMPM, xx\% over BOP
-and XX\% over SPP.  Out of the XX benchmarks, SPP-Perc nearly matches
+and XX\% over SPP.  Out of the XX benchmarks, NATCH nearly matches
 or outperforms most of the prefetchers on XX traces.
 
-At its peak, SPP-Perc manages to get the IPC speedup of a factor of
+At its peak, NATCH manages to get the IPC speedup of a factor of
 over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}.  This also
@@
   603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s}
-benefit the most from SPP-Perc, as compared to SPP only prefetching.
+benefit the most from NATCH, as compared to SPP only prefetching.
 
@@
 \subsection{Multi-core Results}
-In this section, we demonstrate the improvement achieved by SPP-Perc
+\label{Results-Multi}
+In this section, we demonstrate the improvement achieved by NATCH
 for a mix of multi-programmed workloads.
@@
 measured in terms of normalized weighted speedup.  As can be seen,
-SPP-Perc offers an IPC improvement by 33.5\% on these traces.  This is
+NATCH offers an IPC improvement by 33.5\% on these traces.  This is
 5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP
@@
 
-This increased improvement by SPP-Perc on a 4-core machine is not
-surprising.  Since SPP-Perc has an extremely intelligent throttler
+This increased improvement by NATCH on a 4-core machine is not
+surprising.  Since NATCH has an extremely intelligent throttler
 working to eliminate useless prefetches before they can create cache
@@
 
-\textit{8-core environment} To test the effectiveness of SPP-Perc in a
+\textit{8-core environment} To test the effectiveness of NATCH in a
 tighter constrained multi-core environment, we simulated an 8-core
 machine with 16 MB last Level Cache.  Other parameters were kept same
-as discussed under Methodology in Section XX.  <See if we can use
-speed-up over BOP as a measure here>
+as discussed under Methodology in Section \ref{Method-Workloads}.  
+\textit{<See if we can use
+speed-up over BOP as a measure here>}
 
 \subsection{Additional Memory Constraints}
-Figure \_\_ shows the performance of SPP-Perc against the other
+\label{Results-AdditionalMem}
+Figure \_\_ shows the performance of NATCH against the other
 prefetchers in tighter constraint environments.  For sake of
-simplicity, we show 2 traces where SPP-Perc gives the most advantage
-over baseline, 2 traces where SPP-Perc doesn't perform as well as the
+simplicity, we show 2 traces where NATCH gives the most advantage
+over baseline, 2 traces where NATCH doesn't perform as well as the
 best prefetcher *<-replace with the prefetcher name once results are
-available* and finally GeoMean across all the traces.  <Hopefully BOP
-would degrade more here as it is super aggressive.  By definition,
-SPP-Perc should be better than SPP as the throttler would better
-control the aggressiveness.  In SPP paper DA-AMPM performs close
-enough to BOP. That should get replicated here too.>
+available* and finally GeoMean across all the traces.
 
 \subsection{Cross Validation}
-Here we try to demonstrate the robustness of SPP-Perc by testing it on
+\label{Results-CrossVal}
+Here we try to demonstrate the robustness of NATCH by testing it on
 completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
 
-Fig \_\_ shows the speed-up achieved by BOP, SPP and SPP-Perc on
+Fig \_\_ shows the speed-up achieved by BOP, SPP and NATCH on
 selected traces of SPEC 2006, along with the overall geometric mean
 over whole of the benchmark and over memory-intensive subset of the
-SPEC 2006 benchmark.  As can be seen, SPP-Perc peovides a speed-up of
+SPEC 2006 benchmark.  As can be seen, NATCH provides a speed-up of
 44.4\% over baseline, on the memory intensive subset of SPEC 2006

### 2018-11-19 back.tex
Updated Sections: Intro, Back, Method, Results

@@
 
-%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?
+%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to talk about this paper?
 Finally, the paper on Data Cache Prefetching with Perceptron
@@
 for a while.  It was first popularized by Jimenez \textit{et. al.} for
-doing branch prediction~\cite{Perc_Branch}. In it's original
-implementation, the Program Counter of the instruction looking for a
-branch prediction would index into a given table of perceptron
-weights.  The retrieved weights would then be multiplied with the
-history of branch outcomes stored in the History Register (which is a
-feature) in a dot product fashion.  The sum obtained is thresholded at
-0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the
-prediction is made as \textit{true}, or branch taken.  Else if the sum
-is < 0, the prediction is made as \textit{false}.
-
+doing branch prediction~\cite{Perc_Branch}. The biggest advantage of 
+the perceptron model is that it can adjust its weight dynamically 
+to increase / decrease the importance given to each feature - branch 
+hostories in this case.
 The structure that we use is derived from the model introduced in
@@
 replacement.  Perceptron layer can learn to correlate the cache
-replacement behaviour with a variety of features.  Each feature has
-its dedicated table of perceptron weights which it indexes into.
-Hence number of tables equals the number of features.  Once the
-perceptron weights from all the tables are obtained, they're summed
-and thresholded based on a pre-decided value.  This led to a
+replacement behaviour with a variety of features. The features
+introduced in that paper were derived from Program Counters of the
+current and the last few instructions and from the current memory 
+access address.  This led to a
 prediction mechanism that is highly accurate, has quick learning rate
-and is adaptable to changing program behaviour.  The features
-introduced in that paper were derived from Program Counters of the
-current and the last few instructions and from the memory address of the
-current access.  Even with such basic features, high level of
-correlation could be established by perceptrons.
+and is adaptable to changing program behaviour.
 
@@
 
-\subsection{Signature Path Prefetcher Architecture}
+\subsection{Baseline Prefetcher: The Signature Path Prefetcher}
 \label{Background-SPP}
-Since the work we have done uses SPP~\cite{SPP} as the underlying prefetcher, we are
-presenting a brief recap of the ideas introduced in the original
-paper, especially with an eye towards their relevance in our work.
-Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists
-of following structural components:
+
+While the basic idea of perceptron based prefetch filtering is applicable 
+to any look-ahead prefetcher, we developed a practical implementation 
+of the idea using SPP~\cite{SPP} as the underlying prefetcher. 
+SPP is a simple, lightweight, lookahead prefetcher which can be easily 
+modified to be highly aggressive.  
+Here we describe the basic architecture of SPP.
 
@@
 $$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
-In case a matching page entry is found, the stored signature in the corresponding entry is retrieved.
-
-%% NOTE: Can we do away with this example? 
-%%       If it looks like we are weiting too much about SPP
-Consider the case that incoming page number 10 with a block offset 3
-finds a match in the ST.  The retrieved pattern signature is 0x30 and
-the Last Offset is 1.  Since now there is the Last Block Offset (1),
-Incoming Block Offset (3), Old Signature (0x30), and New Signature
-calculated as per above equation (0x182), SPP can infer
-non-speculatively that the given pattern of memory accesses (as
-captured in the Old Signature) leads to the particular Delta.  In
-general, Delta is defined as the difference between the prefetch
-suggestion block and the initial block which triggered the prefetch.
-In this case, since SPP is in learning phase, it is defined as the
-difference between the Incoming Block Offset and the Last Block Offset
-(+2 in this case).  This, in turn generates the new memory pattern
-(New Signature).  This newly learned signature pattern and the delta
-is stored in the Pattern Table.
+Delta is the numerical difference between the block offset of the 
+current and the previous memory acces. In case a matching page entry 
+is found, the stored signature retrieved and used to index into the 
+Pattern Table.
 
@@
 delta predictions.  This is implemented by making PT as a 4-way
-associative table.  To measure an estimate of a delta confidence, PT
-implements two counters: C\textsubscript{sig} for each entry indexed
-by the signature and C\textsubscript{delta} for each delta entry for a
-given signature.  More details of this on-line score-keeping mechanism
-is described under Confidence Tracking discussion.
+associative table.
 
-\textit{Lookahead Prefetching:} On each trigger, SPP tries to go down
+%% NOTE: Can we do away with this example? 
+%% If it looks like we are writing too much about SPP
+%% EDIT: Decided to reorganize and remove this for now.
+%% 	Consider the case that incoming page number 10 with a block offset 3
+%% 	finds a match in the ST.  The retrieved pattern signature is 0x30 and
+%% 	the Last Offset is 1.  Since now there is the Last Block Offset (1),
+%% 	Incoming Block Offset (3), Old Signature (0x30), and New Signature
+%% 	calculated as per above equation (0x182), SPP can infer
+%% 	non-speculatively that the given pattern of memory accesses (as
+%% 	captured in the Old Signature) leads to the particular Delta.  In
+%% 	general, Delta is defined as the difference between the prefetch
+%% 	suggestion block and the initial block which triggered the prefetch.
+%% 	In this case, since SPP is in learning phase, it is defined as the
+%% 	difference between the Incoming Block Offset and the Last Block Offset
+%% 	(+2 in this case).  This, in turn generates the new memory pattern
+%% 	(New Signature).  This newly learned signature pattern and the delta
+%% 	is stored in the Pattern Table.
+
+\textbf{Prefetch Filter:} The PF is a 1024 entry 1-way associative table 
+which keeps a record of
+last few entries prefetched.  This proves useful for updating the prefetcher 
+states if a tracked prefetch leads to a demand hit or a cache eviction.
+
+\textbf{Lookahead Prefetching:} On each trigger, SPP tries to go down
 the program speculation path using its own prefetch suggestion.
@@
 
-\textbf{Global History Register:} The GHR is a small 8-entry fully
-associative structure to bootstrap learning across page boundaries. 
-Whenever a prefetch suggestion from the
-Pattern Table crosses page boundaries, instead of completely
-discarding it, GHR saves that information for inter-page learning.
-Since our NATCH implementation in this paper doesn't affect GHR
-functionality in any ways, we won't be going into more details about
-GHR.
-
-\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch
-Filter - a 1024 entry 1-way associative table which keeps a record of
-last few entries prefetched.  This proves useful for following
-reasons:
-\begin{itemize}
-\item \textbf{Eliminating duplicate prefetches}\newline Any subsequent
-  prefetches recommended by SPP mechanism are checked against the
-  existing entries in Prefetch Filter.  If a valid entry is found, it
-  means that the particular cache line is already either in L2 Cache
-  or being fetched.  Hence there is no need to issue a duplicate
-  prefetch.
-
-\item \textbf{Updating prefetcher}\newline Since the PF keeps a track
-of past few prefetches, it uses that information to see if a given
-prefetch led to a demand hit or a cache eviction. Using this information, 
-SPP can update its internal states.
-
-\item \textbf{Tracking Global Accuracy}\newline PF keeps a track of
-  global usefulness ($\alpha$) of the prefetches by calculating 
-  C\textsubscript{useful} / C\textsubscript{total}, where
-  C\textsubscript{total} is the total number of prefetches
-  getting recorded in the PF and C\textsubscript{useful} is the number 
-  of prefetches that led to a demand hit.
-\end{itemize}
-
 \textbf{Confidence Tracking}: The PT keeps a track of hits to each
@@
 confidence P\textsubscript{d} is given as:
-$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown previously. 
-The range of global accuracy ($\alpha$) is [0,1].
+$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ 
+Here $\alpha$ represents the global accuracy, calculated as the ratio of 
+the number of prefetches which led to a demand hit to the number of 
+prefetches recommended in total. The range of $\alpha$ is [0,1].
 
-Here d is the lookahead depth.  For d = 1, when SPP is in
+`D' is the lookahead depth.  For d = 1, when SPP is in
 non-speculative mode, P\textsubscript{0} can be thought of as 1. 
@@
 rudimentary.  It was based on taking a ratio C\textsubscript{d} =
-C\textsubscript{delta} / C\textsubscript{sig} as explained above.  The
-same confidence was used to make the decision whether to prefetch or
+C\textsubscript{delta} / C\textsubscript{sig} as explained previously. This 
+confidence was used to make the decision whether to prefetch or
 not to prefetch; and which level to prefetch.  While this

### 2018-11-19 intro.tex
Updated Sections: Intro, Back, Method, Results

@@
 \label{Introduction}
-Since past few decades, processor and memory scaling has been enabled
+Processor scaling and memory scaling have been enabled
 with different philosophies in mind.  While processor scaling focused
@@
 chasing applications.  Predicting this often requires a trade-off
-between coverage and accuracy.  For a prefetcher to capture highly
+between coverage and accuracy.  
+Most prefetchers achieve this by maintaining an internal confidence 
+mechanism by keeping a counters to track events like cache hits / miss, 
+references to an entry in the table \textit{etc}. By thresholding this 
+confidence against different values, the extent of aggressiveness 
+can be adjusted.
+
+For a prefetcher to capture highly
 irregular access patterns, it needs to be highly aggressive (high
@@
 This effect would get even more pronounced in multi-core scenario
-where the last level cache is a shared resource.  Conversely, a
+where the last level cache is a shared resource~\cite{Friendly}.  Conversely, a
 conservative prefetcher would be highly accurate but might not make
@@
 modify the design to make it as aggressive as it can get.
-This enables the base engine to capture complex memory access patterns hence
-leading to increased coverage.  This comes at a cost of increased DRAM
-traffic, hence reduction in overall IPC.  The prefetch suggestions
-recommended by SPP pass through the perceptron based throttler. Over
-time, perceptron layer learns to correlate the prefetch recommendations 
+This enables the base engine to capture complex memory access patterns 
+and go deeper into the speculation path. This
+leading to increased coverage, which comes at a cost of increased DRAM
+traffic.  The prefetch suggestions
+recommended by the base prefetch engine pass through the perceptron based throttler. 
+Over time, perceptron layer learns to correlate the prefetch recommendations 
 with various available features, leading to elimination of
@@
 \item Implementing the pereptron throttler in a robust and accurate
-  practical prefetching mechanism.  The overall scheme yields
-  increased accuracy as compared to the other state of the art
+  practical prefetching mechanism, yielding a significant performance 
+  improvement as compared to the other state of the art
   prefetchers.  Moreover, the perceptron throttler learns to adapt
-  itself to shared resource constraints - leading to an even increased
-  performance in multi-core and lower capacity / bandwidth
-  environments.
+  itself to shared resource constraints - leading to a further increased
+  performance in multi-core and lower capacity environments.
 
-\item Examining the perceptron weights by doing a statistical study to
-  quantify the contribution made by each feature.  We have also used
-  this idea to develop a comprehensive set of features where each
-  feature contributes some `new' information to the perceptron.  The
-  details are explained under methodology, in Section \ref{Method-Overheads}.
-
-\item Cross-validating the design across benchmarks.  All the
-  development and parameter tuning was done on the SPEC 2017
-  benchmark.  The results have been shown to be consistent across
-  unseen benchmarks.  More details can be found in Section \ref{Results-CrossVal}.
+\item Defining a methodology for determining an appropriate set of 
+  features to be used regardless of the underlying prefetcher used.
+  More details are explained in Section \ref{Method-Features}.
 
@@
 
-We tested NATCH on Champsim, an open source simulator which models
-an Out-of-Order CPU using.  For evaluation, we used SPEC 2017
-benchmarks.  On a single core configuration, NATCH increases
-performance of the system by 39.5\% as compared to the baseline with
-no prefetching, on the memory intensive subset of the benchmark.  This
-outperforms the next best prefetcher, SPP, by 6.84\%.  On a multi-core
+On a single core configuration, NATCH increased
+performance of the system by 6.84\% as compared to the 
+next best prefetcher, SPP.  On a multi-core
 system running a mix of memory intensive subset of the SPEC 2017
@@
 
-In the results part in Section \ref{Results}, we show a detailed analysis of the
-speedup achieved.  We also used SPEC 2006 and CloudSuite benchmarks
-for cross-validation of the prefetcher implementation.
-
 The rest of the paper is organized as follows.  Section \ref{Background} talks about
-background work and recaptures the SPP architecture Section \ref{Design}
+background work and recaptures the SPP architecture.  Section \ref{Design}
 discusses the detailed implementation of the perceptron mechanism and

### 2018-11-19 method.tex
Updated Sections: Intro, Back, Method, Results

@@
 used for the 2nd Data Prefetch Championship (DPC-2)~\cite{DPC_2}. We
-model model a 1-core / 4-core / 8-core Out of Order machine. The
+model a 1-core / 4-core / 8-core Out of Order machine. The
 details of the configuration parameters is summarized in Table
@@
 The block size was fixed 64KB. Prefetching was only triggered at L2
-Cache demand accesses but could be sent to L2 Cache or Last Level
+Cache demand accesses but could be directed to L2 Cache or Last Level
 Cache. No L1 Data level prefetching was done. LRU replacement policy
@@
 
-We performed cross-validation of our NATCH model using SPEC
-2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks.  For
-SPEC 2006, we developed 94 simpoints spread across the 29
-applications.  For CloudSuite, we used the traces made available for
-the 2nd Cache Replacement Competition (CRC-2)~\cite{CRC_2}.  The traces
-include 4 multi-core applications with 6 distinct phases per
-application.  In total, we used 285 traces representing workloads
-across 53 applications.
-
 \textit{Single-core performance:} For single-core simulations, we use
@@
 and next 1 Billion instructions to do detailed simulations and collect
-run-time statistics.  We report the performance as speedup over
-baseline \textit{i.e.}, no prefetching.  We take the Instruction Per
-Cycle (IPC) of the prefetcher and divide it with the IPC of baseline.
+run-time statistics. We report the IPC speed over
+baseline \textit{i.e.}, no prefetching.
 The final number reported is the geometric mean of the speed-up
@@
 \textit{Multi-core performance:} For multi-application workloads, we
-generate 100 completely random mixes and another 100 mixes from the
+generate 100 random mixes and another 100 mixes from the
 memory intensive subset of SPEC 2017.  For 4-core workload, 200
@@
 
-We repeat the same process for 8-core workloads with the exception
-that 20 Million warm-up instructions and 100 Million full instructions
-were executed.  This was done so as to keep the simulation run-time
-within reasonable limits as a single 8-core mix was taking up to 5
+We repeat the same process for 8-core workloads, correspondingly with 16MB LLC.
+Only exception is that 20 Million warm-up instructions and 100 Million 
+full instructions were executed.  
+This was done so as to keep the simulation run-time
+within reasonable limits as a single 8-core mix was taking up to 3
 days to simulate 1 Billion instructions.
+
+\textit{Cross Validation:} We cross-validated our NATCH model using SPEC
+2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks.  
+For single core
+SPEC 2006, we developed 94 simpoints spread across all the 29
+applications. For multi-core, 
+we followed the same methodology as SPEC 2017. 
+For CloudSuite, we used the traces made available for
+the 2nd Cache Replacement Competition (CRC-2)~\cite{CRC_2}.  The traces
+include 4 4-core applications with 6 distinct phases per
+application.
+
+In total, we used 285 traces representing workloads
+across 53 applications. Throughout the paper for SPEC 2017, 
+we consider memory intensive subset as the traces with LLC MPKI > \_\_. 
+This includes 48 out of the 95 simpoints developed.
+Since SPEC 2006 differs considerably in terms of memory behaviour, we 
+define memory intensive subset as traces with LLC MPKI > \_\_.
+This includes 47 out of 94 simpoints.
 

### 2018-11-19 results.tex
Updated Sections: Intro, Back, Method, Results

@@
 \label{Results-Single}
-Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated.  All
-the results have been normalized to baseline \textit{i.e.}, no
+
+Figure \_\_ shows the IPC speedup on the memory intensive subset of SPEC 2017. 
+All the results have been normalized to baseline \textit{i.e.}, no
 prefetching.  As can be seen, NATCH yields a speedup of XX\% over
@@
 At its peak, NATCH manages to get the IPC speedup of a factor of
-over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}.  This also
-corresponds to speedup gain of \textbf{22.4\%} over the next best
+over \textbf{XX} on the trace \textit{602.gcc\_s-2226B}.  This also
+corresponds to speedup gain of \textbf{XX\%} over the next best
 prefetcher - SPP.  In general, benchmarks \textit{602.gcc\_s,
@@
 benefit the most from NATCH, as compared to SPP only prefetching.
+
+On the full SPEC 2017 suite, NATCH improves the baseline by XX\%,
+which is YY\% more than just SPP.
 
@@
 \textit{4-core environment} Figure \_\_ shows a comparison of speedups
-obtained on workload mix of memory intensive subset of SPEC 2017.  We
+obtained on 4-core mix of memory intensive subset of SPEC 2017.  We
 plot all the 4 prefetchers, normalized to baseline.  The workloads
-have been sorted in the increasing order of performance benefit, as
-measured in terms of normalized weighted speedup.  As can be seen,
-NATCH offers an IPC improvement by 33.5\% on these traces.  This is
-5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP
+have been sorted in the increasing order of IPC obtained with running NATCH.
+As can be seen,
+NATCH offers an IPC improvement by 51.7\% on these traces.  This is
+11.9\% over the next best prefetcher SPP; and XX\% and YY\% over BOP
 and DA-AMPM respectively.
 
-This increased improvement by NATCH on a 4-core machine is not
-surprising.  Since NATCH has an extremely intelligent throttler
-working to eliminate useless prefetches before they can create cache
-pollution.  This further amplifies the edge that SPP had over BOP.
-With the advantage of an orthogonal prefetcher, SPP can be as
-aggressive as possible and leave the accuracy aspect to perceptron
-throttler.
+On a different set of fully random SPEC 2017 4-core mixes, NATCH provides an IPC speedup of 
+26.2\% over baseline, which is 5.7\% more than just SPP.
 
-\textit{8-core environment} To test the effectiveness of NATCH in a
-tighter constrained multi-core environment, we simulated an 8-core
-machine with 16 MB last Level Cache.  Other parameters were kept same
-as discussed under Methodology in Section \ref{Method-Workloads}.  
+\textit{8-core environment} The sorted comparison of speedup obtained 
+on the memory intensive 8-core mixes, is depicted in Fig \_\_.
+NATCH improves baseline performance by 38.6\% and that is 10.7\% 
+more than using just SPP.
+
+For completely random SPEC 2017 mixes, NATCH improves performance by 
+23.6\% over the baseline, corresponding to 4.8\% over SPP.
+
+This increased improvement by NATCH on a multi-core machines is not
+surprising.  Since NATCH has an extremely intelligent throttler, it ends up 
+eliminating useless prefetches before they can create cache
+pollution. This further amplifies the edge that SPP had over BOP in multi-core environment.
+
 \textit{<See if we can use
-speed-up over BOP as a measure here>}
+speed-up over BOP as a measure here for showing scalability>}
 
@@
 \label{Results-AdditionalMem}
-Figure \_\_ shows the performance of NATCH against the other
-prefetchers in tighter constraint environments.  For sake of
-simplicity, we show 2 traces where NATCH gives the most advantage
-over baseline, 2 traces where NATCH doesn't perform as well as the
-best prefetcher *<-replace with the prefetcher name once results are
-available* and finally GeoMean across all the traces.
+
+\textit{Low-Bandwidth DRAM}
+Figure \_\_(a) shows the performance of NATCH against other prefetchers. 
+For sake of simplicity, we show 2 traces where NATCH gives the most advantage
+over baseline, 2 traces where NATCH doesn't perform as well as the next
+best prefetcher (replace with the prefetcher name once results are
+available) and GeoMean across memory intensive and across all the 
+SPEC 2017 traces. As can be seen, for the memory intensive subset,
+ NATCH improves baseline by XX\% and that is XX\% more than using SPP.
+
+\textit{Small LLC}
+Figure \_\_(b) shows the similar analysis of NATCH.
+For the memory intensive subset, a performance improvement of XX\% 
+over baseline is seen with NATCH, XX\% more than just SPP.
 
@@
 
-The 4-core results for SPEC 2006 are \_\_.
+For 4-core memory intensive mixes, NATCH improves the baseline by 59.2\%, 
+8.7\% ahead of SPP. For 8-core memory intensive mixes, the IPC speedup 
+over baseline is 47.9\%, 11.5\% ahead of SPP.
 

### 2018-11-19 abs.tex
Mostly editing for English, redundancy, some technical changes, and redundancy.

@@
 \begin{abstract}
-Hardware prefetching has been introduced in modern
-processors as a way to hide cache latencies.  An efficient
-prefetcher should be able to identify complex memory access
-patterns during a program execution.  This enables it to fetch a
-block ahead of its demand access, potentially saving a cache miss.
+Hardware prefetching has been introduced in modern processors as a way to hide
+cache latencies.  An efficient prefetcher should be able to identify complex
+memory access patterns during program execution. This ability enables the
+prefetcher to read a block ahead of its demand access, potentially saving a
+cache miss. Accurately identifying the right blocks to prefetch is essential
+to achieving high performance from the prefetcher.
 
-In this paper, we introduce perceptron learning to help make
-this prefetching decision efficiently. The perceptron layer acts 
-as a throttler to filter out the unnecessary prefetches recommended 
-by the underlying prefetcher.  We have also explored a range 
-of features that can be used to train the perceptron layer.  
-Our results show that NATCH
-improves performance on the memory intensive subset of the SPEC
-2017 Benchmark suite by 6.84\% on single-core and by 11.9\% on
-multi-core traces, as compared to state-of-the art prefetcher.  
-We also demonstrate that the performance gained from using 
-our efficient throttler scales better with increasing number 
+In this paper, we introduce Perceptron-based Prefetch Filtering to help make
+this prefetching decision accurately. The perceptron layer acts as a throttler
+to filter out the unnecessary prefetches recommended by the underlying
+prefetcher.  We have also explored a range of features that can be used to
+train the perceptron layer.  Our results show that perceptron-based filtering
+improves performance on the memory intensive subset of the SPEC 2017 benchmark
+suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared
+to a state-of-the art prefetcher.  We also demonstrate that the performance
+gained from using our efficient throttler scales better with increasing numbers
 of cores.
+
 \end{abstract}

### 2018-11-19 back.tex
Mostly editing for English, redundancy, some technical changes, and redundancy.

@@
 In this section, we compare the existing work done in the domain of
-prefetching.  We also discuss about some of the interesting memory
-access patterns seen in the workloads in SPEC 2017 benchmark.  The
-second half of this section deals with a background of Signature Path
-Prefetcher architecture.
+prefetching.  We also discuss interesting memory access patterns seen in the
+workloads in SPEC 2017 benchmark.  The second half of this section gives
+background for the Signature Path Prefetcher (SPP) architecture.
 
@@
 
-With time newer and more intricate prefetchers were
+With time, newer and more intricate prefetchers were
 introduced~\cite{Wenisch_Temporal_Streaming,Stealth,Feedback_Directed,Coordinated,Bandwidth_Efficient,Pacman,TLB,Linearizing,Sandbox,VLDP,DoL,Domino}.
-A whole class of streaming prefetchers developed on the lines of
-Spatial / Temporal Memory
+A whole class of streaming prefetchers developed on the lines of Spatial /
+Temporal Memory
 Streaming~\cite{Spatial_Pattern,SMS,Temporal_Instruction_Fetch,Off_Chip,STMS,SMS_JILP}.
-There have been ideas for control flow speculation directed hardware prefetching~\cite{BFetch,MTBFetch}.
+Control-flow speculation directed hardware prefetching has also been
+proposed~\cite{BFetch,MTBFetch}.
 
-Most modern prefetchers aim at identifying complex memory access
-patterns in a given application.  This allows them to capture some of
-the irregular patterns seen in pointer chasing data-structures.
-Offset based prefetchers are a generalization of the next-line
-prefetcher.  Some of the prominent prefetchers exploiting this idea
-are Sandbox Prefetcher~\cite{Sandbox} and the Best Offset Prefetcher~\cite{BOP}. 
-Another class of prefetchers is lookahead Prefetchers.
-Such prefetchers try to speculate deep into the application's memory
-access trail.  These include Path Confidence based Lookahead
-Prefetching~\cite{SPP} and Kill the Program Counter~\cite{KPC}.
+Most modern prefetchers aim at identifying complex memory access patterns in a
+given application.  Thus, they capture some of the irregular patterns seen in
+pointer chasing data-structures.  Offset based prefetchers are a
+generalization of the next-line prefetcher.  Some of the prominent prefetchers
+exploiting this idea are Sandbox Prefetcher~\cite{Sandbox} and the Best Offset
+Prefetcher~\cite{BOP}.  Another class of prefetchers is lookahead Prefetchers.
+Such prefetchers try to speculate deep into the application's memory access
+path.  These include Path Confidence based Lookahead Prefetching~\cite{SPP}
+and Kill the Program Counter~\cite{KPC}.
 
-A related line of work also explores cache replacement - insertion
-policies and dead block prediction, with an eye towards
-prefetching~\cite{DB_Pred,Cache_Burst,KPC,Harmony}.  Nori
-\textit{et.al.} introduced prefetching with respect to runtime
-criticality~\cite{CATCH}.
+A related line of work also explores cache replacement -- insertion policies
+and dead block prediction, with an eye towards
+prefetching~\cite{DB_Pred,Cache_Burst,KPC,Harmony}.  Nori \textit{et.al.}
+introduced prefetching with respect to runtime criticality~\cite{CATCH}.
 
-\textit{Machine Learning and Prefetching:} Peled \textit{et.al.}
-introduce interesting ideas for on-line Reinforcement Learning and
-dynamically scaling the magnitude of feedback given to the baseline
-prefetcher~\cite{Semantics}. The biggest challenge here is that it
-relies on compiler support for getting features to build the context.
+\textit{Machine Learning and Prefetching:} Peled \textit{et.al.} introduce
+interesting ideas for on-line Reinforcement Learning and dynamically scaling
+the magnitude of feedback given to the baseline prefetcher~\cite{Semantics}.
+The biggest challenge here is that the idea relies on compiler support for
+getting features to build the context.
 
-The paper by Liao \textit{et. al.} focuses on prefetching for data
-center applications~\cite{Datacenter}. They use offline Machine
-Learning Algorithms like SVMs and Logistic Regressions to do
-parametric search for an optimal prefetcher configuration.
+Liao \textit{et. al.} focus on prefetching for data center
+applications~\cite{Datacenter}. They use offline machine learning algorithms
+such as SVMs and logistic regression to do a parametric search for an optimal
+prefetcher configuration.
 
 %% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to talk about this paper?
-Finally, the paper on Data Cache Prefetching with Perceptron
-Learning~\cite{BadPerc} talks about the idea of two step prefetching.
-The first step is an existing baseline prefetcher while the second
-step is a perceptron based throttler.  The paper has lots of
-limitations in terms of design and implementation.  That reflects in
-the fact that it did not lead to any significant performance gain over
-even the basic prefetchers like the Stride Prefetcher~\cite{Stride} and
-the Markov Prefetcher~\cite{Markov}.
+% djimenez: let us not cite this paper. it is in arXiv and we are not
+% obligated to cite it. as far as ISCA is concerned it does not exist. there is
+% a reason it is only in arXiv and not published somewhere: it is crap.
+
+%Finally, the paper on Data Cache Prefetching with Perceptron
+%Learning~\cite{BadPerc} talks about the idea of two step prefetching.
+%The first step is an existing baseline prefetcher while the second
+%step is a perceptron based throttler.  The paper has lots of
+%limitations in terms of design and implementation.  That reflects in
+%the fact that it did not lead to any significant performance gain over
+%even the basic prefetchers like the Stride Prefetcher~\cite{Stride} and
+%the Markov Prefetcher~\cite{Markov}.
 
@@
 \label{Background-Perceptron}
-Perceptron learning for computer architecture design has been around
-for a while.  It was first popularized by Jimenez \textit{et. al.} for
-doing branch prediction~\cite{Perc_Branch}. The biggest advantage of 
-the perceptron model is that it can adjust its weight dynamically 
-to increase / decrease the importance given to each feature - branch 
-hostories in this case.
-The structure that we use is derived from the model introduced in
-Piecewise Linear Branch Prediction~\cite{Piece_Linear}.  Here the
-feature itself is used to index into a hash of perceptron weights.
-This form of indexing makes sure that the hyper-plane learned by the
-perceptron weights is able to differentiate between linearly
-inseparable outcomes.  \textit{<Is it true? Not sure>}
+Perceptron learning for computer architecture design has been around for a
+while.  It was first popularized by Jim\'enez \textit{et. al.} for conditional
+branch prediction~\cite{Perc_Branch}. The biggest advantage of the perceptron
+model is that it can adjust its weights dynamically to increase / decrease the
+importance given to each feature - branch histories in this case.  The
+structure we use is derived from the model introduced for Perceptron-based
+Reuse Prediction~\cite{Perc_Reuse}, refined with many additional features as
+Multiperspective Reuse Prediction{Multiperspective}, and originally derived
+from Hashed Perceptron branch prediction{HashedPerceptron}.  Here, a hash of
+the feature itself is used to index into a table of perceptron weights.
 
-Perceptron Learning for Reuse Prediction~\cite{Perc_Reuse} extended the
-hash perceptron architecture to perform prediction in context of cache
-replacement.  Perceptron layer can learn to correlate the cache
-replacement behaviour with a variety of features. The features
-introduced in that paper were derived from Program Counters of the
-current and the last few instructions and from the current memory 
-access address.  This led to a
-prediction mechanism that is highly accurate, has quick learning rate
-and is adaptable to changing program behaviour.
+% djimenez: this detail is unnecessary and likely to invite questions.
+% reviewers might ask "what about linear separability?" but the proof is in the
+% pudding.
+%This form of indexing makes sure that the hyper-plane learned by the
+%perceptron weights is able to differentiate between linearly
+%inseparable outcomes.
+
+Perceptron Learning for Reuse Prediction~\cite{Perc_Reuse} extended the hashed
+perceptron architecture to perform prediction in context of cache replacement.
+The perceptron layer learns to correlate cache replacement behaviour with a
+variety of features. The features introduced in that paper were derived from
+program counters of the current and the last few memory access instructions
+and from the current memory access address.  This led to a prediction
+mechanism that is highly accurate, has quick learning rate and is adaptable to
+changing program behaviour.
 
@@
 Prediction~\cite{Multiperspective}, which introduced a varied set of
-parametric features.  Although design space exploration for feature
-and parameter selection was a non-trivial task, once fixed, it could
-give a cache replacement predictor highly suitable for the given set
-of applications.
+parametric features.  Although design space exploration for feature and
+parameter selection was a non-trivial task, once fixed, it yields
+replacement predictor highly suitable for the given set of applications.
 
-\textit{Deviations From Actual Perceptrons:} Traditionally, a
-perceptron prediction involves multiplying the vector of input
-features: F\textsubscript{1xN}, with the corresponding weight vector:
-W\textsubscript{Nx1} in a dot product fashion to obtain the sum
-y\textsubscript{out}.  Here what we use is a perceptron-like
-structure.  The feature is used to hash into the weight of perceptrons
-and the retrieved weights are added straight away.  This can be seen a
-dot product of $\vec{1}$\textsubscript{1xN} with W\textsubscript{Nx1}.
-This way, the perceptron algorithm doesn't introduce any
-multiplication operations in the inference or training process.  Hence
-what we adapt in this work, is a perceptron-like learning algorithm as
-it involves the same principle involved in perceptron inferencing and
-training.  For simplicity, we will call our implementation as
-perceptron implementation.
+\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron
+prediction involves multiplying the vector of input features:
+F\textsubscript{1xN}, with the corresponding weight vector:
+W\textsubscript{Nx1} in a dot product to obtain the sum y\textsubscript{out}.
+Here what we use is a perceptron-like structure.  The feature is used to hash
+into the weight of perceptrons and the retrieved weights are added straight
+away.  Thus, the perceptron algorithm doesn't introduce any multiplication
+operations in the inference or training process.  Hence what we adapt in this
+work is a perceptron-like learning algorithm as it involves the same principle
+involved in perceptron inferencing and training.
 
-\textit{Perceptron Update Rule:} In all the above implementations, a
-uniform perceptron update principle is being followed.  The weights
-need to be updated if the prediction was wrong or the predicted sum
-does not cross a certain threshold.  After a certain training period,
-these weights are proportional to the probability of the outcome being
-true.  This training threshold makes sure that the perceptrons are
-trained till a certain level of confidence and yet they are not
-over-trained to the given set.  Care is taken that the weights
-saturate at certain positive and negative values so that they remain
-confined to the bit-width.  The same rules will be applied in the
-discussion done in Section \ref{Design-Perceptron} on NATCH learning algorithm.
+\textit{Perceptron Update Rule:} In all the above implementations, a uniform
+perceptron update principle is followed.  The weights need to be updated if
+the prediction was wrong or the magnituded of the dot product does not exceed
+a certain threshold.  After a certain training period, these weights are
+proportional to the probability of the outcome being true.  This training
+threshold makes sure that the perceptrons are trained until a certain level of
+confidence and yet they are not over-trained to the given set.  Care is taken
+that the weights saturate at the maximum positive and negative values so that
+they remain confined to the bit-width.  The same rules will be applied in the
+discussion done in Section \ref{Design-Perceptron} on PPF learning algorithm.
 
-\subsection{Prefetching with SPEC 2017}
-\label{Background-SPEC2017}
-\textit{<Discuss SPEC 2017 traces here>
-<What are the possible graphs / analysis tools we can use for SPEC traces?>}
+\subsection{Prefetching with SPEC 2017} \label{Background-SPEC2017}
+% djimenez: we might not need to discuss the benchmarks too much as they are
+% standard. we don't have a lot of room anyway.
+
+\textit{<Discuss SPEC 2017 traces here> <What are the possible graphs /
+analysis tools we can use for SPEC traces?>}
 
@@
 
-While the basic idea of perceptron based prefetch filtering is applicable 
-to any look-ahead prefetcher, we developed a practical implementation 
-of the idea using SPP~\cite{SPP} as the underlying prefetcher. 
-SPP is a simple, lightweight, lookahead prefetcher which can be easily 
-modified to be highly aggressive.  
-Here we describe the basic architecture of SPP.
+While the basic idea of perceptron based prefetch filtering is applicable to
+any look-ahead prefetcher, we developed a practical implementation of the idea
+using SPP~\cite{SPP} as the underlying prefetcher.  SPP is a simple,
+lightweight, lookahead prefetcher which can be easily modified to be highly
+aggressive.  Here we describe the basic architecture of SPP.
+
+% djimenez: Paul, is all this stuff accurate? SPP continues to confuse me.
 
@@
 
+% djimenez: it does look like we're pushing it with all the stuff about SPP.
+% Let's describe it enough to give the reader the idea that he/she has a vague
+% understanding of how it works and can refer to the MICRO paper for more
+% details. (I'm not saying cut anything that's already there, just don't expand.)
+
 %% NOTE: Can we do away with this example? 
@@
 
-\textbf{Prefetch Filter:} The PF is a 1024 entry 1-way associative table 
-which keeps a record of
-last few entries prefetched.  This proves useful for updating the prefetcher 
-states if a tracked prefetch leads to a demand hit or a cache eviction.
+\textbf{Prefetch Filter:} The PF is a 1024 entry 1-way associative table that
+keeps a record of last few entries prefetched.  This proves useful for
+updating the prefetcher states if a tracked prefetch leads to a demand hit or
+a cache eviction.
 
@@
 
-`D' is the lookahead depth.  For d = 1, when SPP is in
+`D' is the lookahead depth. For $d = 1$, when SPP is in
 non-speculative mode, P\textsubscript{0} can be thought of as 1. 
@@
 \label{Background-Case}
-As compared to some of the other state of the art prefetchers [BOP],
-SPP is a less aggressive prefetcher.  In the single core
-environment, this gives BOP an edge as there is no resource contention
-among the different cores.  Hence more aggressive prefetching is bound
-to prove beneficial.  As we increase the core count, we observe that
-SPP starts outperforming rest of the prefetchers.  This can be
-attributed to the fact that each prefetch suggested by SPP is a
-carefully calculated one and that prevents cache pollution.
-\textit{<Figure showing the variation of SPP vs BOP wrt core count>}
-For 4-core applications, SPP suggests XX\% fewer prefetches than BOP
-and yet leads to higher IPC.
+As compared to some of the other state of the art prefetchers [BOP], SPP is
+less aggressive.  In the single-core environment, this gives BOP an edge as
+there is no resource contention among the different cores.  Hence more
+aggressive prefetching is bound to prove beneficial.  As we increase the core
+count, we observe that SPP starts outperforming rest of the prefetchers.  This
+can be attributed to the fact that each prefetch suggested by SPP is a
+carefully calculated one and that prevents cache pollution.  \textit{<Figure
+showing the variation of SPP vs BOP wrt core count>} For 4-core applications,
+SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
 
-The above analysis shows that with a more careful throttling
-mechanism, any prefetcher like SPP can be tuned to become much more aggressive, leading to
-increased coverage.  The onus of maintaining the accuracy now falls on
-the independent throttler.  To test the hypothesis, we tuned down SPP
-to the minimum possible threshold.  Only by doing this we managed to
-increase the prefetch suggestions made by SPP, by XX\%.
-\textit{<Comparison of SPP-Unleased with SPP / BOP>} Obviously this
-came at a cost of increased DRAM traffic and cache pollution.
+The above analysis shows that with a more careful throttling mechanism, any
+prefetcher like SPP can be tuned to become much more aggressive, leading to
+increased coverage.  The onus of maintaining the accuracy now falls on the
+independent throttler.  To test the hypothesis, we tuned down SPP to the
+minimum possible threshold, with the effect of increasing prefetch suggestions
+made by SPP by XX\%.  \textit{<Comparison of SPP-Unleased with SPP / BOP>}
+Obviously this came at a cost of increased DRAM traffic and cache pollution.
 
@@
 C\textsubscript{delta} / C\textsubscript{sig} as explained previously. This
-confidence was used to make the decision whether to prefetch or
-not to prefetch; and which level to prefetch.  While this
-approximation was shown to work in the original implementation, we believe that a
-better form of generalised on-line decision making was possible.  
-Hence, it was
-necessary to build a robust and adaptable learning mechanism to accept
-/ reject the prefetch suggestions; and to decide the fill level (L2
-Cache vs Last Level Cache).
+confidence was used to make the decision of whether to prefetch or not to
+prefetch; and which level to prefetch.  While this approximation was shown to
+work in the original implementation, we believe that a better form of
+generalised on-line decision making was possible.  Hence, it was necessary to
+build a robust and adaptable learning mechanism to accept / reject the
+prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level
+Cache).
 
-To that effect, we introduce an independent on-line perceptron based throttling mechanism.
+To that effect, we introduce an independent on-line perceptron based
+throttling mechanism.

### 2018-11-19 conc.tex
Mostly editing for English, redundancy, some technical changes, and redundancy.

@@
-\section{CONCLUSION}
+\section{Conclusion}
 \label{Conclusion}
 In this paper, we introduced neural network based learning for data
-prefetching in the form of NATCH.  Perceptron acts an orthogonal
-throttler to the underlying prefetch engine.  NATCH improves
+prefetching in the form of PPF.  Perceptron acts an orthogonal
+throttler to the underlying prefetch engine.  PPF improves
 performance by over XX\% over the baseline, which corresponds to YY\%

### 2018-11-19 design.tex
Mostly editing for English, redundancy, some technical changes, and redundancy.

--- 2018-11-16 design.tex
@@
-\section{NATCH Architecture}
+\section{PPF Architecture}
 \label{Design}
 
-In this section we describe our implementation of NATCH.
-NATCH is a generalised prefetch throttler and can be adapted to 
-any existing or upcoming prefetcher with minimal modifications.
-For a practical implementation, we have used SPP as the base engine prefetcher.
+In this section we describe our implementation of PPF.  PPF is a generalised
+prefetch throttler and can be adapted to most prefetchers with minimal
+modifications.  For a practical implementation, we have used SPP as the base
+engine prefetcher.
 
@@
 \item \textbf{Original Thresholds discarded}\newline In the original
-  SPP design, the internal confidence was
-  compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to
-  decide whether to accept or reject the prefetch suggestion.
-  Furthermore, the accepted suggestions were thresholded against FILL
-  \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache
-  vs Last Level Cache.  T\textsubscript{p} and T\textsubscript{f} were
-  set to 25 and 90 respectively, on the scale of 0 to 100.  In
-  NATCH, the perceptron sum is used to make these decisions.  Hence
-  these two thresholds are no longer needed
+  SPP design, the internal confidence was compared against a preset
+  PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the
+  prefetch suggestion.  Furthermore, the accepted suggestions were thresholded
+  against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2
+  Cache vs Last Level Cache.  T\textsubscript{p} and T\textsubscript{f} were
+  set to 25 and 90 respectively, on the scale of 0 to 100.  In PPF, the
+  perceptron sum is used to make these decisions.  Hence these two thresholds
+  are no longer needed
 
 %%[NOTE] Can this point be removed?
-\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
-  Original SPP would not consider the slots available in the L2 MSHR
-  queue.  It had an assumption that at no time can the number of
-  prefetches suggested by SPP could exceed the capacity of the queue.
-  While it was not proven in the paper, the internal confidence
-  mechanism along with the thresholds made sure that the above
-  assumption was maintained.  In NATCH since there can be no
-  assumption on the degree of aggressiveness of the underlying prefetcher, 
-  we explicitly check that at no point
-  should the number of suggested prefetches exceed L2 MSHR queue.
+% djimenez: I think this would be necessary to reproduce our results so leave it in.
 
-\item \textbf{Enhanced Filter}\newline Original SPP introduces a
-  concept of Prefetch Filter, which keeps a track of up to 1024 L2
-  Cache prefetches suggestions.  For NATCH, the Filter was modified to store
-  meta-data like program counter, memory access address etc.  This is
-  needed to train the perceptron when the result of that particular
-  prefetch gets available.  At training time, this data is used to
-  retrieve the state of the prefetcher when the prediction was made.
+\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The original
+SPP would not consider the slots available in the L2 MSHR queue.  It had an
+assumption that at no time can the number of prefetches suggested by SPP
+exceed the capacity of the queue.  While it was not proven in the paper, the
+internal confidence mechanism along with the thresholds made sure that the
+above assumption was maintained.  In PPF since there can be no assumption on
+the degree of aggressiveness of the underlying prefetcher, we explicitly check
+that at no point should the number of suggested prefetches exceed the L2 MSHR
+queue.
 
-\item \textbf{Reject Filter introduced}\newline The Prefetch Filter
-  mentioned above keeps a track of the prefetches that actually
-  happened in the L2 Cache.  Correspondingly, we introduced a reject
-  filter.  It keeps a track of the prefetches that were suggested by
-  the base SPP engine but were rejected by the perceptron.  In case it
-  is detected that a future demand fetch could have used this rejected
-  prefetch, the perceptron weights are updated to reflect the same
+\item \textbf{Enhanced Filter}\newline The original SPP introduces a concept of
+the Prefetch Filter that keeps a track of up to 1024 L2 cache prefetch
+suggestions.  For PPF, the Filter was modified to store metadata such as program
+counter, memory access address etc.  These features are needed to train the perceptron
+when the result of that particular prefetch becomes available.  At training time,
+this data is used to retrieve the state of the prefetcher when the prediction
+was made.
+
+\item \textbf{Reject Filter introduced}\newline The Prefetch Filter mentioned
+above keeps track of the prefetches that actually happened in the L2 Cache.
+Correspondingly, we introduced a reject filter.  It keeps track of the
+prefetches that were suggested by the base SPP engine but were rejected by the
+perceptron.  In case it is detected that a future demand fetch could have used
+this rejected prefetch, the perceptron weights are updated to reflect that
+misprediction.
 
@@
 As discussed earlier, the perceptron layer was introduced to act as a
-throttler to the underlying prefetchh engine.  This section explores the
-details of the perceptron layer and the range of features that are
-ultimately used by the perceptron for correlating the prefetches.
+throttler to the underlying prefetch engine.  This section explores the
+details of the perceptron layer and the range of features that are ultimately
+used by the perceptron for correlating the prefetches.
 
-The neural network weights are stored in separate tables for each
-feature.  The indexing of each feature is done using different number
-of bits.  At most 12 bits of a feature are used to index into the
-perceptron table.  Certain features require more resolution power
-\textit{i.e.}, full 12 bits of indexing.  On the other hand, some
-features requiring lower resolution required as low as 7 bits of
+The weights are stored in separate tables for each
+feature.  Features are indexed using different numbers
+of bits.  At most 12 bits of a feature are used to index into a
+weights table.  Certain features require more resolution power
+\textit{i.e.}, the full 12 bits of indexing.  On the other hand, some
+features requiring lower resolution require as few as 7 bits of
 indexing.  The variable indexing was determined by studying the
 features and fine-tuned empirically so as to achieve a good accuracy
-vs area trade-off.  Exact details are mentioned under `Area Overheads'
-in Section \ref{Method-Overheads}
+vs area trade-off.  Exact details are mentioned under ``Area Overheads''
+in Section~\ref{Method-Overheads}
 
@@
 weight is a 5 bit counter - saturating at -16 and +15.  Each feature
-gets its own dedicated table.  Our proposed design uses 9 features.
-As the program execution begins, all the weights are initialized to 0.
+has its own dedicated table.  Our proposed design uses 9 features.
+As program execution begins, all the weights are initialized to 0.
 
-\textbf{INFERENCING}\newline In the Champsim simulator, the
-prefetching method is triggered on every L2 Cache demand access.  At
-that point, the prefetcher has the option to do or not do prefetch -
-in case it does, then how many cache lines to prefetch.  The suggested
-prefetches can be places either in L2 Cache or in Last Level Cache.
-The prefetcher gets to decide this, and is usually done based on an
-internal confidence mechanism.
+\textbf{Inference}\newline In the ChampSim simulator, the prefetching method
+is triggered on every L2 Cache demand access.  At that point, the prefetcher
+has the option to do or not do a prefetch -- if it does, then it has a choice
+in how many cache lines to prefetch.  The suggested prefetches can be either
+in the L2 Cache or the last-level cache.  The prefetcher decides this
+placement, and it is usually done based on an internal confidence mechanism.
 
-When the base prefetcher gets triggered, it starts suggesting candidates for prefetching. 
-All these
-recommended prefetches are tested in the perceptron throttler before
-they can access the Prefetch Filter, as in the original SPP data-path.
+When the base prefetcher is triggered, it starts suggesting candidates for
+prefetching.  All these recommended prefetches are tested in the perceptron
+throttler before they can access the Prefetch Filter, as in the original SPP
+datapath.
 
-The perceptron looks at the micro-architectural state, or `features',
-at that instant.  Each feature indexes into the hash of up to
-4096-entry table dedicated for that feature.  Feature \#1 indexes into
+The perceptron looks at the microarchitectural state, {\em i.e.} the features,
+at that instant.  Each feature is hashed to form an index into a table of up to 
+4096 entries dedicated for that feature.  
+Feature \#1 indexes into
 a particular entry in table \#1 and so on.
 
-Once all the weights are retrieved, they are added together, much like
-the traditional perceptron dot product.  On obtaining the overall sum,
-it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO.
-Only the prefetch candidates with the perceptron sum higher that the
-threshold qualify for prefetching.  The prefetches that qualify the
-perceptron throttler stage go through the Prefetch Filter stage.  
-The Prefetch Filter eliminates any redundant
-prefetch suggestions.  In NATCH, all the meta-data that is required
-to recreate the features is also stored in the Prefetch Filter.  At a
-later stage, when the feedback of the current prefetched line is
-available, the stored data can be used to train the perceptron.
+Once all the weights are retrieved, they are summed.  The sum is thresholded
+based on the preset threshold PERC\_THRESHOLD\_LO.  Only the prefetch
+candidates with perceptron sum higher that the threshold qualify for
+prefetching.  The prefetches that qualify through the perceptron throttler
+stage go through the Prefetch Filter stage.  The Prefetch Filter eliminates
+any redundant prefetch suggestions.  In PPF, all the metadata that is required
+to recreate the features is also stored in the Prefetch Filter.  At a later
+stage, when the feedback of the current prefetched line is available, the
+stored data is used to train the perceptron.
 
-The above generated perceptron sum is also used to decide L2 Cache vs
-Last Level Cache placement of the prefetch cache line.  All prefetch
-candidates qualified till this stage are thresholded against
-PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
-PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
+The perceptron sum is also used to decide L2 Cache vs Last Level Cache
+placement of the prefetch cache line.  All prefetch candidates qualified till
+this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill
+level. The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are
+empirically set.
 
-The perceptron sum can also be seen as the sum of the individual
-contribution per feature.  The value of each contribution corresponds
-to the extent of confidence of the final output with that given
-feature.  By summing the individual contributions, the final
-perceptron sum denotes the overall confidence for that prefetch
-suggestion.  In a way, by thresholding the perceptron sum on two
-different thresholds, we are diving the confidence scale of the
-prefetch suggestion into three bins.  The first bin corresponding to
-the lowest confidence leads to prefetch candidate being rejected.  The
-next bin corresponds to prefetches with a moderate confidence level.
-Such prefetches are directed to fill the bigger Last Level Cache and
-potentially not pollute the more scarce L2 Cache.  The highest
-confidence prefetches are filled in the L2 Cache.
+The perceptron sum can also be seen as the sum of the individual contribution
+per feature.  The value of each contribution corresponds to the extent of
+confidence of the final output with that given feature.  By summing the
+individual contributions, the final perceptron sum denotes the overall
+confidence for that prefetch suggestion.  By thresholding the perceptron sum
+on two different thresholds, we are dividing the confidence scale of the
+prefetch suggestion into three bins.  The first bin corresponds to the lowest
+confidence leads to prefetch candidate being rejected.  The next bin
+corresponds to prefetches with a moderate confidence level.  Such prefetches
+are directed to fill the bigger Last Level Cache and potentially not pollute
+the more scarce L2 Cache.  The highest confidence prefetches are filled in the
+L2 Cache.
 
-In addition to the Prefetch Filter mentioned above, NATCH also
-maintains a `Reject Filter'.  Reject filter is a 1024-entry deep 1-way
-associative table.  In case that a prefetch suggestion is rejected by
-perceptron throttler, it instead gets logged into the Reject Filter.
-The Reject Filter keeps track of 1024 entries that were suggested by
-the Pattern Table but got rejected by the perceptron throttler.  This
-can be used to train the perceptron to avoid false negatives
-\textit{i.e.}, cases where prediction was to reject the prefetch but
-the prefetch would have been useful.
+In addition to the Prefetch Filter mentioned above, PPF also maintains a
+``Reject Filter.''  The reject filter is a 1024-entry deep direct-mapped
+table.  If a prefetch suggestion is rejected by perceptron throttler, it is
+logged into the Reject Filter.  The filer is used to train the perceptron to
+avoid false negatives \textit{i.e.}, cases where prediction was to reject the
+prefetch but the prefetch would have been useful.
 
-As previously stated all the meta-data describing the program state at the
-instant of prefetching, is stored in the Prefetch Filter or the Reject Filter.  
-This information is useful when the perceptron needs to be updated subsequently.
+% djimenez: if it were previously stated then it's redundant here? do we need
+% this statement?
 
-\textbf{TRAINING}\newline In the prefetching environment, feedback of
-a prefetch is received whenever there is an eviction or a demand
-access from L2 Cache.  This acts as a trigger to update all the
-base prefetcher entries and to also train the perceptron throttler.
-Training of the perceptron throttler involves restoring the meta-data
-that was stored in Prefetch Filter or the Reject Filter.  This is done
-by indexing into the corresponding filters using the cache line
-address of the block being used to train. (10-bits indexing + 6-bits
-tag matching).  Once the state of the program at the time of
-prefetching is available, it is used to index into the perceptron
-weights table.
+%As previously stated all the metadata describing the program state at the
+%instant of prefetching is stored in the Prefetch Filter or the Reject Filter.
+%This information is useful when the perceptron needs to be updated
+%subsequently.
 
-If a demand access block which triggers the training was tagged as a
-valid prefetch in the Prefetch Filter, it means that the earlier
-prefetch prediction was correct.  In that case the perceptron weights
-are incremented by 1 if the predicted sum does not cross a pre-defined
-threshold.  In case it is a cache block eviction that led to training
-and the corresponding valid entry was found in the Prefetch Filter, it
-shows that the prediction made by the perceptron was wrong.
-Perceptron should have ideally rejected the prefetch suggestion as a
-low-confidence prefetch.  Here the weights are decremented by 1 to
-reflect the same.  In either case, weights are saturated at -16 or
-+15.
+\textbf{Training}\newline In the prefetching environment, feedback for a
+prefetch is received whenever there is an eviction or a demand access from the
+L2 Cache.  This action triggers training of the base prefetcher as well as the
+perceptron throttler.  Training the perceptron throttler involves restoring
+the metadata that was stored in Prefetch Filter or the Reject Filter.  This is
+done by indexing into the corresponding filters using the cache line address
+of the block being used to train. (10-bits indexing + 6-bits tag matching).
+Once the state of the program at the time of prefetching is available, it is
+used to index into the perceptron weights table.
 
-A secondary training mechanism also kicks in during demand fetches.
-Before the demand access triggers the next set of prefetches, it
-checks the Reject Filter for the valid entry.  In case it is a hit,
-that means that the corresponding cache line was initially suggested
-by the underlying prefetch engine but rejected by the perceptron throttler.
-This means that perceptron should have been more confident about that
-particular prefetch.  Once such a scenario is identified, the state of
-the execution at the time of prefetch is retrieved from the Reject
-Table.  The retrieved data is used to index into the various weights
-tables of the perceptron and the corresponding values are updated by
-+1, saturating between and -16 and +15.  This update reflects
-increased confidence for the prediction corresponding to that
-prefetch.
+If a demand access block that triggers the training was tagged as a valid
+prefetch in the Prefetch Filter, then the earlier prefetch prediction was
+correct.  In that case the perceptron weights are incremented by 1 if the
+predicted sum does not cross a pre-defined threshold.  If a cache block
+eviction led to training and the corresponding valid entry was found in the
+Prefetch Filter, then the prediction made by the perceptron was wrong.  The
+perceptron should have ideally rejected the prefetch suggestion as a
+low-confidence prefetch.  Here the weights are decremented by 1 to reflect the
+misprediction. In either case, weights are saturated at -16 or +15.
 
-This mechanism allows us to solve the classical problem of exploiting
-`lost opportunity'.  In prior perceptron based implementations (and in
-general, prefetching algorithms), there is usually no way of knowing
-the result of not prefetch a particular line.  Our two-step NATCH
-architecture allows us to overcome that issue in a simple manner.
+A secondary training mechanism also kicks in during demand fetches.  Before
+the demand access triggers the next set of prefetches, Reject Filter is
+checked for a valid entry.  A hit means that the corresponding cache line was
+initially suggested by the underlying prefetch engine but rejected by the
+perceptron throttler.  Thus, the perceptron should have been more confident
+about that particular prefetch.  Once such a scenario is identified, the state
+of the execution at the time of prefetch is retrieved from the Reject Table.
+The retrieved data is used to index into the various weights tables of the
+perceptron and the corresponding values are updated by +1, saturating between
+and -16 and +15.  This update reflects increased confidence for the prediction
+corresponding to that prefetch.
+
+This mechanism allows us to solve the classical problem of exploiting a lost
+opportunity.  In prior perceptron based implementations (and in general,
+prefetching algorithms), there is usually no way of knowing the result of not
+prefetching a particular line.  Our two-step PPF architecture allows us to
+overcome that issue.
 
@@
 \label{Design-Features}
-Here we discuss the various features which help correlate the
-prefetching decision with the program behaviour
+Here we discuss the various features that correlate the prefetching decision
+with the program behaviour.
+
 \begin{itemize}
-\item \textbf{Base Address} of the demand access which triggered the
-  prefetch.  Since prefetch are triggered from the L2 demand access,
+\item \textbf{Base Address} of the demand access that triggered the
+  prefetch.  Since prefetches are triggered from the L2 demand access,
   they tend to be correlated with the triggering address.
@@
 \item \textbf{Cache Line} and \textbf{Page Address}: These two
-  separate features are derived from the base address which triggered
+  separate features are derived from the base address that triggered
   the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and
@@
 
-\item \textbf{Program Counter XOR Depth}: PC is for the instruction
-  which triggered the prefetch chain.  Depth refers to the iteration
-  count of the lookahead stages.  In general, PC is not considered as
-  a good basis for doing lookahead prefetching as all the prefetches
-  with depth >= 1 gets aliased onto the same PC which won't be true in
-  actual demand fetch case.  This feature resolves a PC into a
-  different value for each lookahead depth of prefetch speculation, giving
-  a more accurate correlation in lookahead cases.  This is akin to the
-  concept of Virtual Program Counters\cite{VPC} introduced by Kim
-  \textit{et. al.} for indirect branch prediction.
+% djimenez: this is a nice intuition and explanation.
+
+\item \textbf{Program Counter XOR Depth}: The PC is for the instruction
+  that triggered the prefetch chain.  Depth refers to the iteration count of
+  the lookahead stages.  In general, the PC is not considered as a good basis
+  for doing lookahead prefetching as all the prefetches with depth >= 1 are
+  aliased into the same PC which would not be true in an actual demand access.
+  This feature resolves a PC into a different value for each lookahead depth
+  of prefetch speculation, giving a more accurate correlation in lookahead
+  cases.  This is akin to the concept of Virtual Program Counters~\cite{VPC}
+  introduced by Kim \textit{et. al.} for indirect branch prediction.
 
@@
   PC before the instruction that triggered the current prefetch.
-  Hashing together the last three PCs tell NATCH about the path
+  Hashing together the last three PCs tell PPF about the path
   that led to the current demand access and helps capture and
@@
 
-\item \textbf{Program Counter XOR Delta}: This tells us if a given PC
-  favours particular value(s) of delta.  As noted earlier, while PC in
-  itself does not much useful information, this hash resolves PC into
-  different values based on the tendency of that PC to favour a
-  certain delta.  This way, dynamic nature of different instances of
-  the same memory access instruction can be captured here.
+\item \textbf{Program Counter XOR Delta}: This feature tells us if a given PC
+  favours particular value(s) of delta.  As noted earlier, while the PC in
+  itself does not convey much useful information, this hash resolves the PC
+  into different values based on the tendency of that PC to favour a certain
+  delta.  Thus, the dynamic nature of different instances of the same memory
+  access instruction can be captured here.
 
-\item \textbf{Confidence}: The integer confidence on the scale of 0 to
-  100 which was used in the original SPP design.  While the original
-  confidence is not used directly for decision making, it can still
-  contribute to the final decision made.
+\item \textbf{Confidence}: The integer confidence on a scale of 0 to
+  100 that was used in the original SPP design.  While the original confidence
+  is not used directly for decision making, it can still contribute to the
+  final decision made.
 
-\item \textbf{Current Signature XOR Delta}: It can be recalled from
-  the discussion of SPP in Section \ref{Background-SPP} that the new signature is
-  generated using the old signature and the blobk delta.  While
-  `Current Signature XOR Delta' is not the exact formula for
-  generating the future signature, it gives an approximate idea of the
-  path that the combination of these two values can lead to.
+% djimenez: this is a little vague. can you be more precise?
+\item \textbf{Current Signature XOR Delta}: Recall from
+  the discussion of SPP in Section \ref{Background-SPP} that the new signature
+  is generated using the old signature and the block delta.  While ``Current
+  Signature XOR Delta'' is not the exact formula for generating the future
+  signature, it gives an approximate idea of the path that the combination of
+  these two values can lead to.
 
-\item \textbf{Page Address XOR Confidence}: This feature helps look at the 
-tendency of each page to be prefetch friendly or prefetch averse. It 
-helps resolve a page into different entries depending on its confidence 
-for prefetching, which can vary during phases of a program execution.
+\item \textbf{Page Address XOR Confidence}: This feature scores the tendency
+of each page to be prefetch friendly or prefetch averse. It helps resolve a
+page into different entries depending on its confidence for prefetching, which
+can vary during phases of a program execution.
 
@@
 
-Some of the composite features are derive from simple hashing (XOR) of
-two primary features.  There is always a question of usefulness of
-such composite features and the `fresh' information added.  We try to
-justify the choice of each feature by quantifying the contribution
-made towards predicting prefetch behaviour, in Section \ref{Method-Features}.  Finally,
-as noted above, each feature indexes into its independent entry of
-perceptron weights.
+Some of the composite features are derived from simple hashing (XOR) of two
+primary features.  There is always a question of usefulness of such composite
+features and the new information added.  We justify the choice of each feature
+by quantifying the contribution made towards predicting prefetch behaviour, in
+Section \ref{Method-Features}.  Finally, as noted above, each feature indexes
+into its independent entry of perceptron weights.
 
-\subsection{Generalizing NATCH for any Prefetcher}
+\subsection{Generalizing PPF for any Prefetcher}
 \label{Design-Generalizing}
-The above implementation of NATCH suggests that it is highly modular and can be 
+The above implementation of PPF shows that it is highly modular and can be
 adapted to be used over any base prefecher for increased prefetch accuracy.
-In our implementation, 
-only two `hooks' are required between NATCH and the baseline SPP. First  
-is to make sure that all the prefetch candidates of SPP pass through 
-the perceptron throttler and if qualified, the metadata for perceptron indexing
-be stored. Second is needed when the feedback of a prior prefetch is available in form 
-of a subsequent demand hit or cache eviction. In that case, the stored meta-data needs 
-to be retrieved to update the state of the perceptrons.
+In our implementation, only two hooks are required between PPF and the
+baseline SPP. The first is to make sure that all the prefetch candidates of
+SPP pass through the perceptron throttler and if qualified, the metadata for
+perceptron indexing be stored. The second is needed when the feedback of a
+prior prefetch is available in form of a subsequent demand hit or cache
+eviction. In that case, the stored metadata needs to be retrieved to update
+the state of the perceptrons.
 
-In general, NATCH can be adapted to a new base prefetcher with only a few modifications.
+In general, PPF can be adapted to a new base prefetcher with only a few modifications.
 \begin{itemize}
@@
 
-\item \textbf{Inferencing and Storing:} All prefetch recommendations to be tested 
-using the perceptron inferencing algorithm.
-Perceptron output: \textit{true} and \textit{false} should be saved appropriately, along with all 
-the meta-data required for perceptron indexing.
+\item \textbf{Inferencing and Storing:} All prefetch recommendations to be
+tested using the perceptron inferencing algorithm.  Perceptron output:
+\textit{true} and \textit{false} should be saved appropriately, along with all
+the metadata required for perceptron indexing.
 
-\item \textbf{Retrieving and Training:} When the feedback is available, the previously 
-stored meta-data can be used to re-index into the perceptron entries and increment / decrement 
-the weights.
+\item \textbf{Retrieving and Training:} When the feedback is available, the
+previously stored metadata can be used to re-index into the perceptron entries
+and increment / decrement the weights.
 
-\item \textbf{Feature Selection:} Six out of the nine features we developed use information derived 
-from program execution, irrespective of the baseline SPP prefetcher. Beyond that, the feature set 
-can be expanded to convey any useful information from the prefetcher to the perceptron throttler.
-For example, in our implementation 'Confidence', 'Current Signature XOR Delta', and 
-'Page Address XOR Confidence' are such features.
+\item \textbf{Feature Selection:} Six out of the nine features we developed
+use information derived from program execution, irrespective of the baseline
+SPP prefetcher. Beyond that, the feature set can be expanded to convey any
+useful information from the prefetcher to the perceptron throttler.  For
+example, in our implementation ``Confidence,'' ``Current Signature XOR Delta,''
+and ``Page Address XOR Confidence'' are such features.
 

### 2018-11-19 intro.tex
Mostly editing for English, redundancy, some technical changes, and redundancy.

@@
-\section{INTRODUCTION}
+\section{Introduction}
 \label{Introduction}
-Processor scaling and memory scaling have been enabled
-with different philosophies in mind.  While processor scaling focused
-on speed improvements, memory scaling was centred around increasing
-the storage capacity.  The process technology reaching scaling
-limitations led to Memory Wall~\cite{MemWall} - the increasing
-performance gap between the processor speed and the memory speeds.
-Various techniques have been developed to combat this widening gap.
-Data prefetching is one such important technique.  Data prefetching
-exploits the fact that in most applications, memory accesses have a
-well-defined pattern.
+Processor scaling and memory scaling have been enabled with different
+philosophies in mind.  While processor scaling focused on speed improvements,
+memory scaling was centred around increasing the storage capacity.  Process
+technology reaching scaling limitations led to the Memory Wall~\cite{MemWall}
+-- the increasing performance gap between processor speed and the memory
+speeds.  Various techniques have been developed to combat this wide gap. Data
+prefetching is one such important technique. Data prefetching exploits the
+fact that in most applications, memory accesses have a well-defined pattern.
 
-An ideal prefetching scheme involves capturing this memory access
-patterns to predict future prefetches in a timely manner.  Other
-important decisions faced by the prefetcher are when to bring in the
-prefetch block of data, which level of cache hierarchy to place the
-block in and finally, what block to evict to accommodate the incoming
-block.  Memory accesses patterns can be as simple as fetching the next
-cache line or a complex but predictable pattern, like in pointer
-chasing applications.  Predicting this often requires a trade-off
-between coverage and accuracy.  
-Most prefetchers achieve this by maintaining an internal confidence 
-mechanism by keeping a counters to track events like cache hits / miss, 
-references to an entry in the table \textit{etc}. By thresholding this 
-confidence against different values, the extent of aggressiveness 
-can be adjusted.
+An ideal prefetching scheme woudl capture these memory access patterns to
+predict future accesses in a timely manner.  Other important decisions faced
+by the prefetcher are when to bring in the prefetch block of data, which level
+of the cache hierarchy to place the block in, and what block to evict to
+accommodate the incoming block.  Memory accesses patterns can be as simple as
+fetching the next cache line or a complex pattern such as pointer chasing.
+Predicting future accesses requires a trade-off between coverage and accuracy.
+Most prefetchers maintain an internal confidence mechanism by keeping counters
+to track events such as cache hits / misses, references to an entry in the
+table, etc. By thresholding this confidence against different values, the
+degree of aggressiveness can be adjusted.
 
-For a prefetcher to capture highly
-irregular access patterns, it needs to be highly aggressive (high
-coverage).  That also means the prefetcher would end up recommending a
-lot of useless prefetches, leading to cache pollution (low accuracy).
-This effect would get even more pronounced in multi-core scenario
-where the last level cache is a shared resource~\cite{Friendly}.  Conversely, a
-conservative prefetcher would be highly accurate but might not make
-enough useful predictions.
+For a prefetcher to capture highly irregular access patterns, it needs to be
+highly aggressive (high coverage).  Thus, the prefetcher would recommend many
+useless prefetches, leading to cache pollution (low accuracy).  This effect is
+even more pronounced in a multi-core scenario where the last-level cache is a
+shared resource~\cite{Friendly}.  Conversely, a conservative prefetcher would
+be highly accurate but might not make enough useful prefetches.
 
-To this effect we propose NATCH: NeurAl prefeTCh tHrottler.  It
-is an enhancement to the existing state of the art prefetcher and
-overcomes the coverage vs accuracy trade-off in a robust manner.  The
-idea is to use a state-of-the-art prefetcher as the base engine.
-For that we choose Signature Path Prefetcher (SPP) and 
-modify the design to make it as aggressive as it can get.
-This enables the base engine to capture complex memory access patterns 
-and go deeper into the speculation path. This
-leading to increased coverage, which comes at a cost of increased DRAM
-traffic.  The prefetch suggestions
-recommended by the base prefetch engine pass through the perceptron based throttler. 
-Over time, perceptron layer learns to correlate the prefetch recommendations 
-with various available features, leading to elimination of
-useless prefetches. This leads to an overall increased accuracy of
-the prefetch scheme and reduction in DRAM traffic.
+We propose Perceptron-based Prefetch Filtering (PPF).  It is an enhancement to
+the existing state of the art prefetcher and robustly overcomes the coverage
+vs.  accuracy trade-off.  The idea is to use a state-of-the-art prefetcher as
+the base engine.  For that we choose Signature Path Prefetcher (SPP) and
+modify the design to make it as aggressive as possible.  This modification
+enables the base engine to capture complex memory access patterns and go
+deeper into the speculation path, leading to increased coverage. Normally such
+aggressive prefetching would come at the cost of increased DRAM traffic.
+However, prefetch suggestions recommended by the base prefetch engine are
+passed through the perceptron based throttler.  Over time, perceptron layer
+learns to correlate the prefetch recommendations with various available
+features, leading to elimination of useless prefetches. This filtering leads
+to overall increased accuracy of the prefetch scheme and reduction in DRAM
+traffic.
 
-\vspace{1ex}This paper talks about NATCH, its merits, analysis and
-future scope for expansion.  The contributions made by this paper are:
+\vspace{1ex}This paper describes PPF, giving its merits, analysis and
+future scope for expansion.  The contributions include:
 
 \begin{itemize}
-\item This is the first time where a truly on-line learning based
-  neural network model has been used when it comes to hardware data
-  prefetching.  The prior works in this area either relied on program
-  semantics~\cite{Semantics} or were application
-  specific~\cite{Datacenter} or could not yield a significant
-  improvement over the baseline prefetcher used~\cite{BadPerc}.
 
-\item Implementing the pereptron throttler in a robust and accurate
+\item An on-line neural model is used for hardware data
+  prefetching.  Previous work in this area either relied on program
+  semantics~\cite{Semantics} or were application specific~\cite{Datacenter}.
+% djimenez: removing reference to this arXiv paper. we are not obligated to cite it.
+%, or
+%  could not yield a significant improvement over the baseline prefetcher
+%  used~\cite{BadPerc}.
+
+\item Implementing the perceptron throttler in a robust and accurate
   practical prefetching mechanism, yielding a significant performance
-  improvement as compared to the other state of the art
-  prefetchers.  Moreover, the perceptron throttler learns to adapt
-  itself to shared resource constraints - leading to a further increased
-  performance in multi-core and lower capacity environments.
+  improvement as compared to the other state of the art prefetchers.
+  Moreover, the perceptron throttler learns to adapt itself to shared resource
+  constraints, leading to further increased performance in multi-core and
+  lower capacity environments.
 
 \item Defining a methodology for determining an appropriate set of 
-  features to be used regardless of the underlying prefetcher used.
+  features for prediction regardless of the underlying prefetcher used.
   More details are explained in Section \ref{Method-Features}.
@@
 
-On a single core configuration, NATCH increased
-performance of the system by 6.84\% as compared to the 
-next best prefetcher, SPP.  On a multi-core
-system running a mix of memory intensive subset of the SPEC 2017
-traces, NATCH saw an improvement of 11.9\% over SPP for 4-core
-system and 10.67\% for 8-core system.
+On a single core configuration, PPF increased performance by 6.84\% as
+compared to the next best prefetcher, SPP.  On a multi-core system running a
+mix of memory intensive subset of the SPEC 2017 traces, PPF saw an improvement
+of 11.9\% over SPP for a 4-core system and 10.67\% for an 8-core system.
 
-The rest of the paper is organized as follows.  Section \ref{Background} talks about
-background work and recaptures the SPP architecture.  Section \ref{Design}
-discusses the detailed implementation of the perceptron mechanism and
-the features used.  Section \ref{Method} consists of the evaluation methodology
-and explores the feature space for perceptron learning.  Section \ref{Results} is
-the performance evaluation of our prefetcher and Section \ref{Conclusion} concludes
-the paper.
+The paper is organized as follows. Section \ref{Background} describes
+background work and summarizes SPP.  Section \ref{Design} discusses the
+detailed implementation of the perceptron mechanism and the features used.
+Section \ref{Method} gives the evaluation methodology and explores the feature
+space for perceptron learning.  Section \ref{Results} evaluates the
+performance of the prefetcher and Section \ref{Conclusion} concludes the
+paper.

### 2018-11-19 method.tex
Mostly editing for English, redundancy, some technical changes, and redundancy.

@@
-\section{METHODOLOGY}
+\section{Methodology}
 \label{Method}
@@
 \label{Method-Model}
-For testing and comparing NATCH, we used ChampSim~\cite{Champsim}
-simulator.  ChampSim is an enhanced version of the framework that was
-used for the 2nd Data Prefetch Championship (DPC-2)~\cite{DPC_2}. We
-model a 1-core / 4-core / 8-core Out of Order machine. The
-details of the configuration parameters is summarized in Table
-\ref{tab:Sim_params}.
+For testing and comparing PPF, we used the ChampSim~\cite{Champsim} simulator.
+ChampSim is an enhanced version of the framework that was used for the 2nd
+Data Prefetch Championship (DPC-2)~\cite{DPC_2}. We model 1-core, 4-core, and
+8-core out-of-order machines. The details of the configuration parameters are
+summarized in Table~\ref{tab:Sim_params}.
 
@@
 
-The block size was fixed 64KB. Prefetching was only triggered at L2
-Cache demand accesses but could be directed to L2 Cache or Last Level
-Cache. No L1 Data level prefetching was done. LRU replacement policy
-was used on all levels of cache hierarchies. Branch prediction was
-done using the perceptron branch predictor~\cite{Perc_Branch}.  Page
-size was kept as 4KB.  Champsim operates all the prefetchers strictly
-in the physical address space.
+% djimenez: changing all the "was" to "is." when describing our work, use
+% present tense
+
+The block size is fixed at 64 bytes. Prefetching is only triggered at an L2 cache
+demand accesses but could be directed to the L2 or last-level cache. No L1
+data level prefetching is done. The LRU replacement policy is used on all levels
+of cache hierarchies. Branch prediction is done using the perceptron branch
+predictor~\cite{Perc_Branch}.  ThE page size is 4KB.  Champsim operates
+all the prefetchers strictly in the physical address space.
 
@@
 \label{Method-Workloads}
-This is the first time that SPEC 2017 benchmark suite~\cite{SPEC2017}
-has been used to characterize and measure the prefetch performance.
-We use all the 20 workloads available in the SPEC 2017 Suite.  Using
-SimPoint~\cite{SimPoint} methodology, we identified 95 different
-program segments of 1 Billion instructions each.
+% djimenez: don't say this is the first time SPEC 2017 has been used this way.
+% i don't know if that's true, and in any event the value is small.
 
-\textit{Single-core performance:} For single-core simulations, we use
-200 Million instructions to warm-up the microarchitectural structures
-and next 1 Billion instructions to do detailed simulations and collect
-run-time statistics. We report the IPC speed over
-baseline \textit{i.e.}, no prefetching.
-The final number reported is the geometric mean of the speed-up
-attained on individual traces.
+%This is the first time that SPEC 2017 benchmark suite~\cite{SPEC2017}
+%has been used to characterize and measure the prefetch performance.
+We use the that SPEC 2017 benchmark suite~\cite{SPEC2017}.  We use all the 20
+workloads available in the SPEC 2017 suite.  Using the SimPoint~\cite{SimPoint}
+methodology, we identified 95 different program segments of 1 Billion
+instructions each.
 
-\textit{Multi-core performance:} For multi-application workloads, we
-generate 100 random mixes and another 100 mixes from the
-memory intensive subset of SPEC 2017.  For 4-core workload, 200
-Million instructions are used for warm-up and additional 1 Billion
-instruction simulated for collecting statistics.  Each CPU keeps
-executing its workload till the last CPU completes 1 Billion
-instructions after warm-up.  For collecting IPC and other data, only
-the first billion instructions are considered as the ``region of
-interest''.
+\textit{Single-core performance:} For single-core simulations, we use 200
+million instructions to warm-up the microarchitectural structures and the next
+one billion instructions to do detailed simulations and collect run-time
+statistics. We report the IPC speedup over the baseline of no prefetching.
+The final number reported is the geometric mean of the speedup achieved on
+individual traces.
+
+\textit{Multi-core performance:} For multi-application workloads, we generate
+100 random mixes and another 100 mixes from the memory intensive subset of
+SPEC 2017.  For 4-core workload, 200 Million instructions are used for warm-up
+and additional 1 Billion instruction simulated for collecting statistics.
+Each CPU keeps executing its workload till the last CPU completes one billion
+instructions after warm-up.  For collecting IPC and other data, only the first
+billion instructions are considered as the region of interest.
 
@@
 
-We repeat the same process for 8-core workloads, correspondingly with 16MB LLC.
-Only exception is that 20 Million warm-up instructions and 100 Million 
-full instructions were executed.  
-This was done so as to keep the simulation run-time
-within reasonable limits as a single 8-core mix was taking up to 3
-days to simulate 1 Billion instructions.
+We repeat the same process for 8-core workloads, correspondingly with 16MB
+LLC.  The only difference is that 20 million warm-up instructions and 100
+million full instructions are executed.  This is done so as to keep the
+simulation run-time within reasonable limits as a single 8-core mix takes up
+to 3 days to simulate one billion instructions.
 
-\textit{Cross Validation:} We cross-validated our NATCH model using SPEC
-2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks.  
-For single core
-SPEC 2006, we developed 94 simpoints spread across all the 29
-applications. For multi-core, 
-we followed the same methodology as SPEC 2017. 
-For CloudSuite, we used the traces made available for
-the 2nd Cache Replacement Competition (CRC-2)~\cite{CRC_2}.  The traces
-include 4 4-core applications with 6 distinct phases per
-application.
+\textit{Validation:} We cross-validated our PPF model using SPEC
+2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks.  For
+single-core SPEC 2006, we developed 94 simpoints spread across all the 29
+applications. For multi-core, we followed the same methodology as SPEC 2017.
+For CloudSuite, we used the traces made available for the 2nd Cache
+Replacement Competition (CRC-2)~\cite{CRC_2}.  The traces include 4 4-core
+applications with 6 distinct phases per application.
 
-In total, we used 285 traces representing workloads
-across 53 applications. Throughout the paper for SPEC 2017, 
-we consider memory intensive subset as the traces with LLC MPKI > \_\_. 
-This includes 48 out of the 95 simpoints developed.
-Since SPEC 2006 differs considerably in terms of memory behaviour, we 
-define memory intensive subset as traces with LLC MPKI > \_\_.
-This includes 47 out of 94 simpoints.
+In total, we used 285 traces representing workloads across 53 applications.
+Throughout the paper for SPEC 2017, we consider memory intensive subset as the
+traces with LLC MPKI > \_\_.  This includes 48 out of the 95 simpoints
+developed.  Since SPEC 2006 differs considerably in terms of memory behaviour,
+we define memory intensive subset as traces with LLC MPKI > \_\_.  This
+includes 47 out of 94 simpoints.
 
@@
 \label{Method-Prefetchers}
-We compared NATCH against three of the latest state of the art
-hardware-only prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map
-Pattern Matching (DA-AMPM)~\cite{DA_AMPM} and Signature Path Prefecher
-(SPP).  BOP was the winner of 2nd Data Prefetching Championship.
-DA-AMPM is the enhanced version of AMPM, modified to account for DRAM
-row buffer locality.  SPP has been shown to outperform BOP on SPEC
-2006 traces.  For each of these, we compare their speedups taking no
-prefetching as the baseline.
+We compared PPF against three of the latest state of the art hardware-only
+prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern
+Matching (DA-AMPM)~\cite{DA_AMPM} and Signature Path Prefecher (SPP).  BOP was
+the winner of 2nd Data Prefetching Championship.  DA-AMPM is the enhanced
+version of AMPM, modified to account for DRAM row buffer locality.  SPP has
+been shown to outperform BOP on SPEC 2006 traces.  For each of these, we
+compare their speedups taking no prefetching as the baseline.
 
-\subsection{Developing Features for NATCH}
+\subsection{Developing Features for PPF}
 \label{Method-Features}
-This section describes the intuition and analysis that went behind
-finalizing the features.  As noted earlier, we developed a set of 9
-features which allow the perceptron throttler to correlate prefetching
-decision with the program context.  To study the correlation across
-each feature, we study statistically the perceptron weights and try to
-interpret their distribution.
+This section describes the intuition and analysis that went behind developing
+the features.  As noted earlier, we developed a set of 9 features that allow
+the perceptron throttler to correlate prefetching decision with the program
+context.  To study the correlation across each feature, we study statistically
+the perceptron weights and try to interpret their distribution.
 
-\textbf{GLOBAL PEARSON'S CORRELATION}\newline For this experiment, we
-create a dump of the state of perceptron weights at the end of all
-trace execution.  The weights obtained from running all the SPEC 2017
-traces are concatenated together.  Since the weights are collected at
-the end of individual trace execution, the perceptron weights have
-attained a relatively stable value by now.  Hence, this dump
-represents a `snap' of the trained perceptron weights across all the
-SPEC 2017 traces.  The intuition here is that the feature with bulk of
-the perceptron weights concentrated around 0 or small magnitude
-numbers show a weak correlation with the prefetching outcome.  On the
-other hand, features with most of the weights saturated around highest
-value (+15) show a high positive correlation and the features with
-weights close to the lowest value (-16) show a strong negative
-correlation.
+\textbf{Global Pearson's Correlation}\newline For this experiment, we create a
+dump of the state of perceptron weights at the end of all trace execution.
+The weights obtained from running all the SPEC 2017 traces are concatenated.
+Since the weights are collected at the end of individual trace execution, the
+perceptron weights have attained a relatively stable value by now.  Hence,
+this dump represents a `snap' of the trained perceptron weights across all the
+SPEC 2017 traces.  The intuition here is that the feature with bulk of the
+perceptron weights concentrated around 0 or small magnitude numbers show a
+weak correlation with the prefetching outcome.  On the other hand, features
+with most of the weights saturated around highest value (+15) show a high
+positive correlation and the features with weights close to the lowest value
+(-16) show a strong negative correlation.
 
@@
 
-\textbf{PER TRACE CORRELATION} \newline Another important way to look
+\textbf{Per Trace Correlation} \newline Another important way to look
 at the perceptron features is to see how much their contribution
@@
 
-\textbf{TRIMMING FEATURES USING CORRELATION} \newline Besides
+\textbf{Trimming Features Using Correlation} \newline Besides
 providing interesting insights into prefetching behaviour, P-value can
@@
 outcome, we can also study correlation between the features.  We used
-the above methodology to eliminate features which provide a little or
+the above methodology to eliminate features that provide a little or
 no information that has already been captured in other features.
@@
 managed to reduce the feature count to 9.  Thus, in the final
-implementation of NATCH, no two features have a high correlation
+implementation of PPF, no two features have a high correlation
 between them.  This way we can be sure that each feature is bringing
@@
 
-\subsection{Overhead for NATCH}
+\subsection{Overhead for PPF}
 \label{Method-Overheads}
@@
 In this section, we analyze the hardware overhead required to
-implement NATCH.  The Prefetch Filter was enhanced to accommodate
+implement PPF.  The Prefetch Filter was enhanced to accommodate
 storing of meta-data for perceptron training.  Table
 \ref{tab:PF_metadata} depicts the meta-data stored for each entry in
-the Prefetch Filter.  Table \ref{tab:NATCH_overhead} shows the total
-storage overhead of NATCH implementation.  The hardware budget for
+the Prefetch Filter.  Table \ref{tab:PPF_overhead} shows the total
+storage overhead of PPF implementation.  The hardware budget for
 2nd Data Prefetching championship was 32 KB.  Keeping that in mind 
-the considerable speedup NATCH obtained over the winner, the extra hardware
+the considerable speedup PPF obtained over the winner, the extra hardware
 budget can be accounted for.  The extra hardware also makes the

### 2018-11-19 results.tex
Mostly editing for English, redundancy, some technical changes, and redundancy.

@@
-\section{RESULTS}
+\section{Results}
 \label{Results}
 
-This section discusses the results obtained from running NATCH, in
+This section discusses the results obtained from running PPF, in
 terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per
@@
 All the results have been normalized to baseline \textit{i.e.}, no
-prefetching.  As can be seen, NATCH yields a speedup of XX\% over
+prefetching.  As can be seen, PPF yields a speedup of XX\% over
 the baseline.  This is equivalent to XX\% over DA-AMPM, xx\% over BOP
-and XX\% over SPP.  Out of the XX benchmarks, NATCH nearly matches
+and XX\% over SPP.  Out of the XX benchmarks, PPF nearly matches
 or outperforms most of the prefetchers on XX traces.
 
-At its peak, NATCH manages to get the IPC speedup of a factor of
+At its peak, PPF manages to get the IPC speedup of a factor of
 over \textbf{XX} on the trace \textit{602.gcc\_s-2226B}.  This also
@@
   603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s}
-benefit the most from NATCH, as compared to SPP only prefetching.
+benefit the most from PPF, as compared to SPP only prefetching.
 
-On the full SPEC 2017 suite, NATCH improves the baseline by XX\%,
+On the full SPEC 2017 suite, PPF improves the baseline by XX\%,
 which is YY\% more than just SPP.
@@
 \label{Results-Multi}
-In this section, we demonstrate the improvement achieved by NATCH
+In this section, we demonstrate the improvement achieved by PPF
 for a mix of multi-programmed workloads.
@@
 plot all the 4 prefetchers, normalized to baseline.  The workloads
-have been sorted in the increasing order of IPC obtained with running NATCH.
+have been sorted in the increasing order of IPC obtained with running PPF.
 As can be seen,
-NATCH offers an IPC improvement by 51.7\% on these traces.  This is
+PPF offers an IPC improvement by 51.7\% on these traces.  This is
 11.9\% over the next best prefetcher SPP; and XX\% and YY\% over BOP
@@
 
-On a different set of fully random SPEC 2017 4-core mixes, NATCH provides an IPC speedup of 
+On a different set of fully random SPEC 2017 4-core mixes, PPF provides an IPC speedup of 
 26.2\% over baseline, which is 5.7\% more than just SPP.
@@
 on the memory intensive 8-core mixes, is depicted in Fig \_\_.
-NATCH improves baseline performance by 38.6\% and that is 10.7\% 
+PPF improves baseline performance by 38.6\% and that is 10.7\% 
 more than using just SPP.
 
-For completely random SPEC 2017 mixes, NATCH improves performance by 
+For completely random SPEC 2017 mixes, PPF improves performance by 
 23.6\% over the baseline, corresponding to 4.8\% over SPP.
 
-This increased improvement by NATCH on a multi-core machines is not
-surprising.  Since NATCH has an extremely intelligent throttler, it ends up 
+This increased improvement by PPF on a multi-core machines is not
+surprising.  Since PPF has an extremely intelligent throttler, it ends up 
 eliminating useless prefetches before they can create cache
@@
 \textit{Low-Bandwidth DRAM}
-Figure \_\_(a) shows the performance of NATCH against other prefetchers. 
-For sake of simplicity, we show 2 traces where NATCH gives the most advantage
-over baseline, 2 traces where NATCH doesn't perform as well as the next
+Figure \_\_(a) shows the performance of PPF against other prefetchers. 
+For sake of simplicity, we show 2 traces where PPF gives the most advantage
+over baseline, 2 traces where PPF doesn't perform as well as the next
 best prefetcher (replace with the prefetcher name once results are
@@
 SPEC 2017 traces. As can be seen, for the memory intensive subset,
- NATCH improves baseline by XX\% and that is XX\% more than using SPP.
+ PPF improves baseline by XX\% and that is XX\% more than using SPP.
 
 \textit{Small LLC}
-Figure \_\_(b) shows the similar analysis of NATCH.
+Figure \_\_(b) shows the similar analysis of PPF.
 For the memory intensive subset, a performance improvement of XX\% 
-over baseline is seen with NATCH, XX\% more than just SPP.
+over baseline is seen with PPF, XX\% more than just SPP.
 
@@
 \label{Results-CrossVal}
-Here we try to demonstrate the robustness of NATCH by testing it on
+Here we try to demonstrate the robustness of PPF by testing it on
 completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
 
-Fig \_\_ shows the speed-up achieved by BOP, SPP and NATCH on
+Fig \_\_ shows the speed-up achieved by BOP, SPP and PPF on
 selected traces of SPEC 2006, along with the overall geometric mean
 over whole of the benchmark and over memory-intensive subset of the
-SPEC 2006 benchmark.  As can be seen, NATCH provides a speed-up of
+SPEC 2006 benchmark.  As can be seen, PPF provides a speed-up of
 44.4\% over baseline, on the memory intensive subset of SPEC 2006
@@
 
-For 4-core memory intensive mixes, NATCH improves the baseline by 59.2\%, 
+For 4-core memory intensive mixes, PPF improves the baseline by 59.2\%, 
 8.7\% ahead of SPP. For 8-core memory intensive mixes, the IPC speedup 

### 2018-11-20 abs.tex
Separating arch and implementation. Filter-Table and Throttler-Filter renaming

@@
 In this paper, we introduce Perceptron-based Prefetch Filtering to help make
-this prefetching decision accurately. The perceptron layer acts as a throttler
+this prefetching decision accurately. The perceptron layer acts as a check 
 to filter out the unnecessary prefetches recommended by the underlying
@@
 to a state-of-the art prefetcher.  We also demonstrate that the performance
-gained from using our efficient throttler scales better with increasing numbers
+gained from using our efficient filter scales better with increasing numbers
 of cores.

### 2018-11-20 back.tex
Separating arch and implementation. Filter-Table and Throttler-Filter renaming

@@
 In this section, we compare the existing work done in the domain of
-prefetching.  We also discuss interesting memory access patterns seen in the
-workloads in SPEC 2017 benchmark.  The second half of this section gives
-background for the Signature Path Prefetcher (SPP) architecture.
+prefetching.  We also give a brief background for the 
+Signature Path Prefetcher (SPP) architecture.
 
@@
 %The first step is an existing baseline prefetcher while the second
-%step is a perceptron based throttler.  The paper has lots of
+%step is a perceptron based filter.  The paper has lots of
 %limitations in terms of design and implementation.  That reflects in
@@
 model is that it can adjust its weights dynamically to increase / decrease the
-importance given to each feature - branch histories in this case.  The
-structure we use is derived from the model introduced for Perceptron-based
-Reuse Prediction~\cite{Perc_Reuse}, refined with many additional features as
-Multiperspective Reuse Prediction{Multiperspective}, and originally derived
-from Hashed Perceptron branch prediction{HashedPerceptron}.  Here, a hash of
-the feature itself is used to index into a table of perceptron weights.
+importance given to each feature - branch histories in this case.  
+Perceptron-based Reuse Prediction~\cite{Perc_Reuse} implemented the hashed
+perceptron architecture to perform prediction in context of cache replacement.
+The structure  we use is derived from the model introduced in that paper, refined with 
+many additional features as in Multiperspective Reuse Prediction{Multiperspective}.  
+The model was originally derived from Hashed Perceptron branch 
+prediction~\cite{HashedPerceptron}.  Here, a hash of the feature itself is used to 
+index into a table of perceptron weights.
 
@@
 %inseparable outcomes.
-
-Perceptron Learning for Reuse Prediction~\cite{Perc_Reuse} extended the hashed
-perceptron architecture to perform prediction in context of cache replacement.
-The perceptron layer learns to correlate cache replacement behaviour with a
-variety of features. The features introduced in that paper were derived from
-program counters of the current and the last few memory access instructions
-and from the current memory access address.  This led to a prediction
-mechanism that is highly accurate, has quick learning rate and is adaptable to
-changing program behaviour.
-
-The work was further extended by Multiperspective Reuse
-Prediction~\cite{Multiperspective}, which introduced a varied set of
-parametric features.  Although design space exploration for feature and
-parameter selection was a non-trivial task, once fixed, it yields
-replacement predictor highly suitable for the given set of applications.
 
@@
 they remain confined to the bit-width.  The same rules will be applied in the
-discussion done in Section \ref{Design-Perceptron} on PPF learning algorithm.
+discussion done in Section \ref{Arch-Perceptron} on PPF learning algorithm.
 
-\subsection{Prefetching with SPEC 2017} \label{Background-SPEC2017}
+%\subsection{Prefetching with SPEC 2017} \label{Background-SPEC2017}
 % djimenez: we might not need to discuss the benchmarks too much as they are
@@
 
-\textit{<Discuss SPEC 2017 traces here> <What are the possible graphs /
-analysis tools we can use for SPEC traces?>}
+%\textit{<Discuss SPEC 2017 traces here> <What are the possible graphs /
+%analysis tools we can use for SPEC traces?>}
 
@@
 
-\textbf{Signature Table:} The ST keeps track of 256 most recently accessed
+\textbf{Signature Table:} The Signature Table keeps track of 256 most recently accessed
 pages.  It is meant to capture memory access patterns within a page
-boundary.  SPP indexes into an entry of ST using the page number.  For
-each entry corresponding to a page, ST stores the `last block offset'
-and `old signature'.  Last block offset is the block offset of the
+boundary.  SPP indexes into an entry of the Signature Table using the page number.  For
+each entry corresponding to a page, the table stores a `last block offset'
+and an `old signature'.  Last block offset is the block offset of the
 last memory access of that given page.  The block offset is calculated
@@
 
-\textbf{Pattern Table:} The PT is indexed by the signature generated
-from ST.  PT holds predicted delta patterns and their confidence
+\textbf{Pattern Table:} The Pattern Table is indexed by the signature generated
+from the Signature Table.  Pattern Table holds predicted delta patterns and their confidence
 estimates.  Each entry indexed by the signature holds up to 4 unique
-delta predictions.  This is implemented by making PT as a 4-way
+delta predictions.  This is implemented by making the Pattern Table as a 4-way
 associative table.
@@
 
-\textbf{Prefetch Filter:} The PF is a 1024 entry 1-way associative table that
+\textbf{Prefetch Table:} The Prefetch Table is a 1024 entry 1-way associative table that
 keeps a record of last few entries prefetched.  This proves useful for
 updating the prefetcher states if a tracked prefetch leads to a demand hit or
-a cache eviction.
+a cache eviction. \textit{Note:} The original paper refers to this as the Prefetch Filter 
+but since in our context filter refers to the perceptron filter, we will be calling 
+this structre as the Prefetch Table to avoid any confusion.
 
@@
 the program speculation path using its own prefetch suggestion.
-Using current prefetch as the baseline, it re-accesses the PT to generate further
-prefetches.  It keeps on repeating this cycle of accessing PT and
+Using current prefetch as the baseline, it re-accesses the Pattern Table to generate further
+prefetches.  It keeps on repeating this cycle of accessing Pattern Table and
 updating the signature based on highest confidence prefetch from the
@@
 
-\textbf{Confidence Tracking}: The PT keeps a track of hits to each
+\textbf{Confidence Tracking}: The Pattern Table keeps a track of hits to each
 signature through a counter C\textsubscript{sig}.  The number of hits
@@
 non-speculative mode, P\textsubscript{0} can be thought of as 1. 
-The final P\textsubscript{d} is thresholded against prefetching-threshold
-(T\textsubscript{p}) to decide whether to prefetch or not.  If yes,
-then P\textsubscript{d} is thresholded against a numerically bigger
-fill-threshold (T\textsubscript{f}) to decide whether the prefetch
-should be sent to L2 Cache (high confidence prefetch) or Last Level
-Cache (low confidence prefetch)
+The final P\textsubscript{d} is thresholded against PF\_THRESHOLD 
+(T\textsubscript{p}) to reject the low confidence suggestions and 
+then against a numerically bigger FILL\_THRESHOLD (T\textsubscript{f}) to 
+decide whether the prefetch should be sent to L2 Cache 
+(high confidence prefetch) or Last Level Cache (low confidence prefetch).
+The two thresholds were empirically set to 25 and 90 respectively, 
+on the scale of 0 to 100.
 
@@
 
-\subsection{Case for an On-line Throttler}
+\subsection{Case for an On-line Filter}
 \label{Background-Case}
@@
 
-The above analysis shows that with a more careful throttling mechanism, any
+The above analysis shows that with a more careful filtering mechanism, any
 prefetcher like SPP can be tuned to become much more aggressive, leading to
 increased coverage.  The onus of maintaining the accuracy now falls on the
-independent throttler.  To test the hypothesis, we tuned down SPP to the
+independent filter.  To test the hypothesis, we tuned down SPP to the
 minimum possible threshold, with the effect of increasing prefetch suggestions
@@
 To that effect, we introduce an independent on-line perceptron based
-throttling mechanism.
+filtering mechanism.

### 2018-11-20 design_arch.tex
Separating arch and implementation. Filter-Table and Throttler-Filter renaming

--- 2018-11-19 design.tex
@@
 \section{PPF Architecture}
-\label{Design}
+\label{Arch}
 
-In this section we describe our implementation of PPF.  PPF is a generalised
-prefetch throttler and can be adapted to most prefetchers with minimal
-modifications.  For a practical implementation, we have used SPP as the base
-engine prefetcher.
+In this section, we talk about the the general structure of the 
+Perceptron-Based Prefetch Filtering (PPF) for an underlying prefetcher.  
+PPF is a generalised prefetch filtering mechanism and can be adapted to 
+most prefetchers with minimal modifications.
 
-\subsection{Changes made to original SPP}
-\label{Design-Changes}
-To modify the SPP design to suit our scheme, the following changes were made:
-
-\begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline In the original
-  SPP design, the internal confidence was compared against a preset
-  PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the
-  prefetch suggestion.  Furthermore, the accepted suggestions were thresholded
-  against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2
-  Cache vs Last Level Cache.  T\textsubscript{p} and T\textsubscript{f} were
-  set to 25 and 90 respectively, on the scale of 0 to 100.  In PPF, the
-  perceptron sum is used to make these decisions.  Hence these two thresholds
-  are no longer needed
-
-%%[NOTE] Can this point be removed?
-% djimenez: I think this would be necessary to reproduce our results so leave it in.
-
-\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The original
-SPP would not consider the slots available in the L2 MSHR queue.  It had an
-assumption that at no time can the number of prefetches suggested by SPP
-exceed the capacity of the queue.  While it was not proven in the paper, the
-internal confidence mechanism along with the thresholds made sure that the
-above assumption was maintained.  In PPF since there can be no assumption on
-the degree of aggressiveness of the underlying prefetcher, we explicitly check
-that at no point should the number of suggested prefetches exceed the L2 MSHR
-queue.
-
-\item \textbf{Enhanced Filter}\newline The original SPP introduces a concept of
-the Prefetch Filter that keeps a track of up to 1024 L2 cache prefetch
-suggestions.  For PPF, the Filter was modified to store metadata such as program
-counter, memory access address etc.  These features are needed to train the perceptron
-when the result of that particular prefetch becomes available.  At training time,
-this data is used to retrieve the state of the prefetcher when the prediction
-was made.
-
-\item \textbf{Reject Filter introduced}\newline The Prefetch Filter mentioned
-above keeps track of the prefetches that actually happened in the L2 Cache.
-Correspondingly, we introduced a reject filter.  It keeps track of the
-prefetches that were suggested by the base SPP engine but were rejected by the
-perceptron.  In case it is detected that a future demand fetch could have used
-this rejected prefetch, the perceptron weights are updated to reflect that
-misprediction.
-
-\end{itemize}
-
-\subsection{The Perceptron Layer}
-\label{Design-Perceptron}
-As discussed earlier, the perceptron layer was introduced to act as a
-throttler to the underlying prefetch engine.  This section explores the
-details of the perceptron layer and the range of features that are ultimately
-used by the perceptron for correlating the prefetches.
-
-The weights are stored in separate tables for each
-feature.  Features are indexed using different numbers
-of bits.  At most 12 bits of a feature are used to index into a
+\subsection{The Perceptron Filter}
+\label{Arch-Perceptron}
+The Perceptron Filter is organised as tables of perceptron weights, 
+with a separate table for each feature.  
+The feature value is used to index into the corresponding table.
+This indexing is done using different numbers of bits for each feature.  
+At most 12 bits of a feature are used to index into the
 weights table.  Certain features require more resolution power
@@
 A single entry in the table corresponds to a perceptron weight.  Each
-weight is a 5 bit counter - saturating at -16 and +15.  Each feature
-has its own dedicated table.  Our proposed design uses 9 features.
-As program execution begins, all the weights are initialized to 0.
+weight is a 5 bit counter - saturating at -16 and +15.  
+Our proposed design uses nine features, hence there are nine perceptron 
+weight tables. As program execution begins, all the weights are initialized to 0.
 
-\textbf{Inference}\newline In the ChampSim simulator, the prefetching method
-is triggered on every L2 Cache demand access.  At that point, the prefetcher
+\textbf{Inference}\newline In the ChampSim simulator, the prefetching method is 
+invoked on every L2 Cache demand access.  At that point, the base prefetcher
 has the option to do or not do a prefetch -- if it does, then it has a choice
@@
 When the base prefetcher is triggered, it starts suggesting candidates for
-prefetching.  All these recommended prefetches are tested in the perceptron
-throttler before they can access the Prefetch Filter, as in the original SPP
-datapath.
+prefetching.  All these recommended prefetches have to be tested through the perceptron
+filter to qualify for final prefetching.
 
@@
 4096 entries dedicated for that feature.  
-Feature \#1 indexes into
-a particular entry in table \#1 and so on.
+Feature \#1 indexes into a particular entry in table \#1 and so on.
 
 Once all the weights are retrieved, they are summed.  The sum is thresholded
-based on the preset threshold PERC\_THRESHOLD\_LO.  Only the prefetch
+based on the preset threshold PERC\_ THRESHOLD\_LO.  Only the prefetch
 candidates with perceptron sum higher that the threshold qualify for
-prefetching.  The prefetches that qualify through the perceptron throttler
-stage go through the Prefetch Filter stage.  The Prefetch Filter eliminates
-any redundant prefetch suggestions.  In PPF, all the metadata that is required
-to recreate the features is also stored in the Prefetch Filter.  At a later
+prefetching.  The prefetches that qualify through the perceptron 
+stage are recorded in the Prefetch Table.  In PPF, all the metadata that is required
+to recreate the features is also stored in the Prefetch Table.  At a later
 stage, when the feedback of the current prefetched line is available, the
-stored data is used to train the perceptron.
+stored data can be used to train the perceptron.
 
@@
 
-In addition to the Prefetch Filter mentioned above, PPF also maintains a
-``Reject Filter.''  The reject filter is a 1024-entry deep direct-mapped
-table.  If a prefetch suggestion is rejected by perceptron throttler, it is
-logged into the Reject Filter.  The filer is used to train the perceptron to
+In addition to the Prefetch Table mentioned above, PPF also maintains a
+``Reject Table.''  The Reject Table is a 1024-entry deep direct-mapped
+table.  If a prefetch suggestion is rejected by the perceptron layer, it is
+logged into the Reject Table.  The filer is used to train the perceptron to
 avoid false negatives \textit{i.e.}, cases where prediction was to reject the
@@
 %As previously stated all the metadata describing the program state at the
-%instant of prefetching is stored in the Prefetch Filter or the Reject Filter.
+%instant of prefetching is stored in the Prefetch Table or the Reject Table.
 %This information is useful when the perceptron needs to be updated
@@
 L2 Cache.  This action triggers training of the base prefetcher as well as the
-perceptron throttler.  Training the perceptron throttler involves restoring
-the metadata that was stored in Prefetch Filter or the Reject Filter.  This is
-done by indexing into the corresponding filters using the cache line address
+perceptron layer.  Training the perceptrons involves restoring
+the metadata that was stored in Prefetch Table or the Reject Table.  This is
+done by indexing into the corresponding tables using the cache line address
 of the block being used to train. (10-bits indexing + 6-bits tag matching).
@@
 If a demand access block that triggers the training was tagged as a valid
-prefetch in the Prefetch Filter, then the earlier prefetch prediction was
-correct.  In that case the perceptron weights are incremented by 1 if the
-predicted sum does not cross a pre-defined threshold.  If a cache block
+prefetch in the Prefetch Table, then the earlier prefetch prediction was
+correct.  In that case the perceptron weights are incremented by 1 
+if the predicted sum does not cross a pre-defined threshold. These training 
+thresholds are introduced to limit the degree of update on correct predictions.
+Doing this helps avoid overfitting of the perceptron weights to the given program 
+behaviour. These thresholds are referred to as POS\_UPDATE\_THRESHOLD ($\theta_p$) and 
+NEG\_UPDATE\_THRESHOLD ($\theta_n$),  respectively for the positive and 
+negative values of training saturation.
+
+If a cache block
 eviction led to training and the corresponding valid entry was found in the
-Prefetch Filter, then the prediction made by the perceptron was wrong.  The
+Prefetch Table, then the prediction made by the perceptron was wrong.  The
 perceptron should have ideally rejected the prefetch suggestion as a
@@
 A secondary training mechanism also kicks in during demand fetches.  Before
-the demand access triggers the next set of prefetches, Reject Filter is
+the demand access triggers the next set of prefetches, Reject Table is
 checked for a valid entry.  A hit means that the corresponding cache line was
 initially suggested by the underlying prefetch engine but rejected by the
-perceptron throttler.  Thus, the perceptron should have been more confident
+perceptron filter.  Thus, the perceptron should have been more confident
 about that particular prefetch.  Once such a scenario is identified, the state
@@
 
-\subsection{Features used by Perceptron}
-\label{Design-Features}
-Here we discuss the various features that correlate the prefetching decision
-with the program behaviour.
-
-\begin{itemize}
-\item \textbf{Base Address} of the demand access that triggered the
-  prefetch.  Since prefetches are triggered from the L2 demand access,
-  they tend to be correlated with the triggering address.
-
-\item \textbf{Cache Line} and \textbf{Page Address}: These two
-  separate features are derived from the base address that triggered
-  the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and
-  base\_addr >> LOG2\_PAGE\_ SIZE respectively.  The idea behind using
-  three different shifted versions of the same feature is that it
-  allows us to look into a wider range of bits than with a single
-  version.  It also helps give more importance to the overlapping bits
-  and lesser importance to most and least significance bits.  This
-  approach can also eliminate destructive interference that can be
-  caused by directly folding the bits into half.
-
-% djimenez: this is a nice intuition and explanation.
-
-\item \textbf{Program Counter XOR Depth}: The PC is for the instruction
-  that triggered the prefetch chain.  Depth refers to the iteration count of
-  the lookahead stages.  In general, the PC is not considered as a good basis
-  for doing lookahead prefetching as all the prefetches with depth >= 1 are
-  aliased into the same PC which would not be true in an actual demand access.
-  This feature resolves a PC into a different value for each lookahead depth
-  of prefetch speculation, giving a more accurate correlation in lookahead
-  cases.  This is akin to the concept of Virtual Program Counters~\cite{VPC}
-  introduced by Kim \textit{et. al.} for indirect branch prediction.
-
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
-  PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$
-  PC before the instruction that triggered the current prefetch.
-  Hashing together the last three PCs tell PPF about the path
-  that led to the current demand access and helps capture and
-  branching information of the current basic block.  PCs are shifted
-  in the increasing order of history before being hashed together.
-  This is done to avoid the resultant value of zero in case 2 or more
-  PCs are the same.  Additionally, obfuscating the information as it
-  gets old allows us to get a wider and an approximate look into the
-  program history.
-
-\item \textbf{Program Counter XOR Delta}: This feature tells us if a given PC
-  favours particular value(s) of delta.  As noted earlier, while the PC in
-  itself does not convey much useful information, this hash resolves the PC
-  into different values based on the tendency of that PC to favour a certain
-  delta.  Thus, the dynamic nature of different instances of the same memory
-  access instruction can be captured here.
-
-\item \textbf{Confidence}: The integer confidence on a scale of 0 to
-  100 that was used in the original SPP design.  While the original confidence
-  is not used directly for decision making, it can still contribute to the
-  final decision made.
-
-% djimenez: this is a little vague. can you be more precise?
-\item \textbf{Current Signature XOR Delta}: Recall from
-  the discussion of SPP in Section \ref{Background-SPP} that the new signature
-  is generated using the old signature and the block delta.  While ``Current
-  Signature XOR Delta'' is not the exact formula for generating the future
-  signature, it gives an approximate idea of the path that the combination of
-  these two values can lead to.
-
-\item \textbf{Page Address XOR Confidence}: This feature scores the tendency
-of each page to be prefetch friendly or prefetch averse. It helps resolve a
-page into different entries depending on its confidence for prefetching, which
-can vary during phases of a program execution.
-
-%%%%%%% Below are the features rejected when downsizing 14 -> 9
-% \item \textbf{Current Signature}: The 12-bit signature used by SPP
-%   to index into the Pattern Table
-% \item \textbf{Delta}: The signed version of the difference in the
-%   block address that triggers the prefetch and the block address of
-%   the data being prefetched
-% \item \textbf{PC}: Program Counter of the Load instruction that
-%   triggers that particular prefetch
-% \item \textbf{Base address XOR Delta}: If any particular value(s) of
-%   delta are highly favoured by a given address which triggered the
-%   prefetch, this composite feature can capture that information
-% \item \textbf{Cache line XOR Delta}: qThis feature captures if any
-%   particular delta is highly favoured by a given cache line address
-%\item \textbf{Confidence XOR Delta}: This gives a combined insight
-%  into that fact that certain delta-confidence pairqs predicted by
-%  SPP might correlate more with the prefetch outcome
-%%%%%%%%%%%%%%%%%%%%%%
-
-\end{itemize}
-
-Some of the composite features are derived from simple hashing (XOR) of two
-primary features.  There is always a question of usefulness of such composite
-features and the new information added.  We justify the choice of each feature
-by quantifying the contribution made towards predicting prefetch behaviour, in
-Section \ref{Method-Features}.  Finally, as noted above, each feature indexes
-into its independent entry of perceptron weights.
-
 \subsection{Generalizing PPF for any Prefetcher}
-\label{Design-Generalizing}
-The above implementation of PPF shows that it is highly modular and can be
+\label{Arch-Generalizing}
+The above discussion of PPF shows that it is highly modular and can be
 adapted to be used over any base prefecher for increased prefetch accuracy.
-In our implementation, only two hooks are required between PPF and the
-baseline SPP. The first is to make sure that all the prefetch candidates of
-SPP pass through the perceptron throttler and if qualified, the metadata for
+In general, only two hooks are required between PPF and the
+baseline prefetcher. The first is to make sure that all the prefetch candidates of
+the prefetcher pass through the perceptron filter and if qualified, the metadata for
 perceptron indexing be stored. The second is needed when the feedback of a
@@
 
-\item \textbf{Feature Selection:} Six out of the nine features we developed
-use information derived from program execution, irrespective of the baseline
-SPP prefetcher. Beyond that, the feature set can be expanded to convey any
-useful information from the prefetcher to the perceptron throttler.  For
-example, in our implementation ``Confidence,'' ``Current Signature XOR Delta,''
-and ``Page Address XOR Confidence'' are such features.
+\item \textbf{Feature Selection:} In Section~\ref{Impl-Features}, we show the features 
+used for PPF. Six out of the nine features we developed
+use information derived directly from program execution, irrespective of the baseline 
+prefetcher. Beyond that, the feature set can be expanded to convey any
+useful information from the prefetcher to the perceptron layer.  The 
+methodology explained in Section~\ref{Method-Features} talks about developing a minimal 
+feature set for PPF.
+%For example, in our implementation ``Confidence,'' ``Current Signature XOR Delta,''
+%and ``Page Address XOR Confidence'' are such features.
 

### 2018-11-20 design_impl.tex
Separating arch and implementation. Filter-Table and Throttler-Filter renaming

--- 2018-11-19 design.tex
@@
-\section{PPF Architecture}
-\label{Design}
+\section{PPF Implementation Using SPP}
+\label{Impl}
 
-In this section we describe our implementation of PPF.  PPF is a generalised
-prefetch throttler and can be adapted to most prefetchers with minimal
-modifications.  For a practical implementation, we have used SPP as the base
-engine prefetcher.
+This section describes our implementation of PPF and the range of 
+features that are ultimately used by the perceptron for correlating 
+the prefetches.  For a practical implementation, we have used SPP 
+as the base engine prefetcher.
 
 \subsection{Changes made to original SPP}
-\label{Design-Changes}
+\label{Impl-Changes}
 To modify the SPP design to suit our scheme, the following changes were made:
@@
 \begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline In the original
-  SPP design, the internal confidence was compared against a preset
-  PF\_THRESHOLD (T\textsubscript{p}) to decide whether to accept or reject the
-  prefetch suggestion.  Furthermore, the accepted suggestions were thresholded
-  against FILL \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2
-  Cache vs Last Level Cache.  T\textsubscript{p} and T\textsubscript{f} were
-  set to 25 and 90 respectively, on the scale of 0 to 100.  In PPF, the
-  perceptron sum is used to make these decisions.  Hence these two thresholds
-  are no longer needed
+\item \textbf{Original Thresholds discarded}\newline 
+  In PPF, the perceptron sum is used to decide whether to prefetch or not, 
+  and the fill-level in case of prefetch. 
+  Thus, the confidence thresholds --  T\textsubscript{f} and T\textsubscript{p} 
+  are no longer needed.
 
@@
   SPP would not consider the slots available in the L2 MSHR queue.  It had an
-assumption that at no time can the number of prefetches suggested by SPP
+  assumption that at no time can the number of prefetches suggested can
   exceed the capacity of the queue.  While it was not proven in the paper, the
-internal confidence mechanism along with the thresholds made sure that the
-above assumption was maintained.  In PPF since there can be no assumption on
-the degree of aggressiveness of the underlying prefetcher, we explicitly check
-that at no point should the number of suggested prefetches exceed the L2 MSHR
-queue.
+  internal confidence mechanism made sure that the above assumption was maintained.  
+  In PPF since there can be no assumption on the degree of aggressiveness of the 
+  underlying prefetcher, we explicitly check that at no point should the number 
+  of suggested prefetches exceed the L2 MSHR queue.
 
-\item \textbf{Enhanced Filter}\newline The original SPP introduces a concept of
-the Prefetch Filter that keeps a track of up to 1024 L2 cache prefetch
-suggestions.  For PPF, the Filter was modified to store metadata such as program
-counter, memory access address etc.  These features are needed to train the perceptron
-when the result of that particular prefetch becomes available.  At training time,
-this data is used to retrieve the state of the prefetcher when the prediction
-was made.
+\item \textbf{Enhanced Prefetch Table}\newline The Prefetch Table introduced in the 
+  original SPP keeps a track of up to 1024 L2 cache prefetch suggestions.  
+  For PPF, the Table was modified to store metadata such as program counter, 
+  memory access address etc.  This metadata is needed to retrieve the state 
+  of the features to train the perceptron at a later stage in time.
 
-\item \textbf{Reject Filter introduced}\newline The Prefetch Filter mentioned
+\item \textbf{Reject Table introduced}\newline The Prefetch Table mentioned
   above keeps track of the prefetches that actually happened in the L2 Cache.
-Correspondingly, we introduced a reject filter.  It keeps track of the
+  Correspondingly, we introduced the Reject Table to keep a track of the
   prefetches that were suggested by the base SPP engine but were rejected by the
@@
 
-\subsection{The Perceptron Layer}
-\label{Design-Perceptron}
-As discussed earlier, the perceptron layer was introduced to act as a
-throttler to the underlying prefetch engine.  This section explores the
-details of the perceptron layer and the range of features that are ultimately
-used by the perceptron for correlating the prefetches.
-
-The weights are stored in separate tables for each
-feature.  Features are indexed using different numbers
-of bits.  At most 12 bits of a feature are used to index into a
-weights table.  Certain features require more resolution power
-\textit{i.e.}, the full 12 bits of indexing.  On the other hand, some
-features requiring lower resolution require as few as 7 bits of
-indexing.  The variable indexing was determined by studying the
-features and fine-tuned empirically so as to achieve a good accuracy
-vs area trade-off.  Exact details are mentioned under ``Area Overheads''
-in Section~\ref{Method-Overheads}
-
-A single entry in the table corresponds to a perceptron weight.  Each
-weight is a 5 bit counter - saturating at -16 and +15.  Each feature
-has its own dedicated table.  Our proposed design uses 9 features.
-As program execution begins, all the weights are initialized to 0.
-
-\textbf{Inference}\newline In the ChampSim simulator, the prefetching method
-is triggered on every L2 Cache demand access.  At that point, the prefetcher
-has the option to do or not do a prefetch -- if it does, then it has a choice
-in how many cache lines to prefetch.  The suggested prefetches can be either
-in the L2 Cache or the last-level cache.  The prefetcher decides this
-placement, and it is usually done based on an internal confidence mechanism.
-
-When the base prefetcher is triggered, it starts suggesting candidates for
-prefetching.  All these recommended prefetches are tested in the perceptron
-throttler before they can access the Prefetch Filter, as in the original SPP
-datapath.
-
-The perceptron looks at the microarchitectural state, {\em i.e.} the features,
-at that instant.  Each feature is hashed to form an index into a table of up to 
-4096 entries dedicated for that feature.  
-Feature \#1 indexes into
-a particular entry in table \#1 and so on.
-
-Once all the weights are retrieved, they are summed.  The sum is thresholded
-based on the preset threshold PERC\_THRESHOLD\_LO.  Only the prefetch
-candidates with perceptron sum higher that the threshold qualify for
-prefetching.  The prefetches that qualify through the perceptron throttler
-stage go through the Prefetch Filter stage.  The Prefetch Filter eliminates
-any redundant prefetch suggestions.  In PPF, all the metadata that is required
-to recreate the features is also stored in the Prefetch Filter.  At a later
-stage, when the feedback of the current prefetched line is available, the
-stored data is used to train the perceptron.
-
-The perceptron sum is also used to decide L2 Cache vs Last Level Cache
-placement of the prefetch cache line.  All prefetch candidates qualified till
-this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill
-level. The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are
-empirically set.
-
-The perceptron sum can also be seen as the sum of the individual contribution
-per feature.  The value of each contribution corresponds to the extent of
-confidence of the final output with that given feature.  By summing the
-individual contributions, the final perceptron sum denotes the overall
-confidence for that prefetch suggestion.  By thresholding the perceptron sum
-on two different thresholds, we are dividing the confidence scale of the
-prefetch suggestion into three bins.  The first bin corresponds to the lowest
-confidence leads to prefetch candidate being rejected.  The next bin
-corresponds to prefetches with a moderate confidence level.  Such prefetches
-are directed to fill the bigger Last Level Cache and potentially not pollute
-the more scarce L2 Cache.  The highest confidence prefetches are filled in the
-L2 Cache.
-
-In addition to the Prefetch Filter mentioned above, PPF also maintains a
-``Reject Filter.''  The reject filter is a 1024-entry deep direct-mapped
-table.  If a prefetch suggestion is rejected by perceptron throttler, it is
-logged into the Reject Filter.  The filer is used to train the perceptron to
-avoid false negatives \textit{i.e.}, cases where prediction was to reject the
-prefetch but the prefetch would have been useful.
-
-% djimenez: if it were previously stated then it's redundant here? do we need
-% this statement?
-
-%As previously stated all the metadata describing the program state at the
-%instant of prefetching is stored in the Prefetch Filter or the Reject Filter.
-%This information is useful when the perceptron needs to be updated
-%subsequently.
-
-\textbf{Training}\newline In the prefetching environment, feedback for a
-prefetch is received whenever there is an eviction or a demand access from the
-L2 Cache.  This action triggers training of the base prefetcher as well as the
-perceptron throttler.  Training the perceptron throttler involves restoring
-the metadata that was stored in Prefetch Filter or the Reject Filter.  This is
-done by indexing into the corresponding filters using the cache line address
-of the block being used to train. (10-bits indexing + 6-bits tag matching).
-Once the state of the program at the time of prefetching is available, it is
-used to index into the perceptron weights table.
-
-If a demand access block that triggers the training was tagged as a valid
-prefetch in the Prefetch Filter, then the earlier prefetch prediction was
-correct.  In that case the perceptron weights are incremented by 1 if the
-predicted sum does not cross a pre-defined threshold.  If a cache block
-eviction led to training and the corresponding valid entry was found in the
-Prefetch Filter, then the prediction made by the perceptron was wrong.  The
-perceptron should have ideally rejected the prefetch suggestion as a
-low-confidence prefetch.  Here the weights are decremented by 1 to reflect the
-misprediction. In either case, weights are saturated at -16 or +15.
-
-A secondary training mechanism also kicks in during demand fetches.  Before
-the demand access triggers the next set of prefetches, Reject Filter is
-checked for a valid entry.  A hit means that the corresponding cache line was
-initially suggested by the underlying prefetch engine but rejected by the
-perceptron throttler.  Thus, the perceptron should have been more confident
-about that particular prefetch.  Once such a scenario is identified, the state
-of the execution at the time of prefetch is retrieved from the Reject Table.
-The retrieved data is used to index into the various weights tables of the
-perceptron and the corresponding values are updated by +1, saturating between
-and -16 and +15.  This update reflects increased confidence for the prediction
-corresponding to that prefetch.
-
-This mechanism allows us to solve the classical problem of exploiting a lost
-opportunity.  In prior perceptron based implementations (and in general,
-prefetching algorithms), there is usually no way of knowing the result of not
-prefetching a particular line.  Our two-step PPF architecture allows us to
-overcome that issue.
-
 \subsection{Features used by Perceptron}
-\label{Design-Features}
+\label{Impl-Features}
 Here we discuss the various features that correlate the prefetching decision
@@
   the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and
-  base\_addr >> LOG2\_PAGE\_ SIZE respectively.  The idea behind using
+  base\_addr >> LOG2\_PAGE\_SIZE respectively.  The idea behind using
   three different shifted versions of the same feature is that it
@@
   approach can also eliminate destructive interference that can be
-  caused by directly folding the bits into half.
+  caused by directly folding the address bits into half.
 
@@
 
+\item \textbf{Page Address XOR Confidence}: This feature scores the tendency
+  of each page to be prefetch friendly or prefetch averse. It helps resolve a
+  page into different entries depending on its confidence for prefetching, which
+  can vary during phases of a program execution.
+
 % djimenez: this is a little vague. can you be more precise?
@@
   these two values can lead to.
-
-\item \textbf{Page Address XOR Confidence}: This feature scores the tendency
-of each page to be prefetch friendly or prefetch averse. It helps resolve a
-page into different entries depending on its confidence for prefetching, which
-can vary during phases of a program execution.
 
@@
 into its independent entry of perceptron weights.
-
-\subsection{Generalizing PPF for any Prefetcher}
-\label{Design-Generalizing}
-The above implementation of PPF shows that it is highly modular and can be
-adapted to be used over any base prefecher for increased prefetch accuracy.
-In our implementation, only two hooks are required between PPF and the
-baseline SPP. The first is to make sure that all the prefetch candidates of
-SPP pass through the perceptron throttler and if qualified, the metadata for
-perceptron indexing be stored. The second is needed when the feedback of a
-prior prefetch is available in form of a subsequent demand hit or cache
-eviction. In that case, the stored metadata needs to be retrieved to update
-the state of the perceptrons.
-
-In general, PPF can be adapted to a new base prefetcher with only a few modifications.
-\begin{itemize}
-
-\item \textbf{Enhancing the Base Prefetcher:} By tuning down any internal thresholds 
-to increase its inherent aggressiveness.
-
-\item \textbf{Inferencing and Storing:} All prefetch recommendations to be
-tested using the perceptron inferencing algorithm.  Perceptron output:
-\textit{true} and \textit{false} should be saved appropriately, along with all
-the metadata required for perceptron indexing.
-
-\item \textbf{Retrieving and Training:} When the feedback is available, the
-previously stored metadata can be used to re-index into the perceptron entries
-and increment / decrement the weights.
-
-\item \textbf{Feature Selection:} Six out of the nine features we developed
-use information derived from program execution, irrespective of the baseline
-SPP prefetcher. Beyond that, the feature set can be expanded to convey any
-useful information from the prefetcher to the perceptron throttler.  For
-example, in our implementation ``Confidence,'' ``Current Signature XOR Delta,''
-and ``Page Address XOR Confidence'' are such features.
-
-\end{itemize}

### 2018-11-20 intro.tex
Separating arch and implementation. Filter-Table and Throttler-Filter renaming

@@
 
-An ideal prefetching scheme woudl capture these memory access patterns to
+An ideal prefetching scheme would capture these memory access patterns to
 predict future accesses in a timely manner.  Other important decisions faced
@@
 However, prefetch suggestions recommended by the base prefetch engine are
-passed through the perceptron based throttler.  Over time, perceptron layer
+passed through the perceptron based filter.  Over time, perceptron layer
 learns to correlate the prefetch recommendations with various available
@@
 
-\item Implementing the perceptron throttler in a robust and accurate
+\item Implementing the perceptron filter in a robust and accurate
   practical prefetching mechanism, yielding a significant performance
   improvement as compared to the other state of the art prefetchers.
-  Moreover, the perceptron throttler learns to adapt itself to shared resource
+  Moreover, the perceptron learns to adapt itself to shared resource
   constraints, leading to further increased performance in multi-core and
@@
 
-The paper is organized as follows. Section \ref{Background} describes
-background work and summarizes SPP.  Section \ref{Design} discusses the
-detailed implementation of the perceptron mechanism and the features used.
-Section \ref{Method} gives the evaluation methodology and explores the feature
-space for perceptron learning.  Section \ref{Results} evaluates the
-performance of the prefetcher and Section \ref{Conclusion} concludes the
+The paper is organized as follows. Section~\ref{Background} describes
+background work and summarizes SPP.  Section~\ref{Arch} given the 
+architectural overview of PPF. Section~\ref{Impl} discusses the
+implementation of PPF using SPP and the features used.
+Section~\ref{Method} gives the evaluation methodology and explores the feature
+space for perceptron learning.  Section~\ref{Results} evaluates the
+performance of the prefetcher and Section~\ref{Conclusion} concludes the
 paper.

### 2018-11-20 method.tex
Separating arch and implementation. Filter-Table and Throttler-Filter renaming

@@
 This section describes the intuition and analysis that went behind developing
-the features.  As noted earlier, we developed a set of 9 features that allow
-the perceptron throttler to correlate prefetching decision with the program
+the features.  As noted earlier, we developed a set of nine features that allow
+the perceptron layer to correlate prefetching decision with the program
 context.  To study the correlation across each feature, we study statistically
@@
     \hline
-         Valid & 1 & Indicates a valid entry in the filter\\
-         Tag & 6 & Identifier for the entry in the filter\\
+         Valid & 1 & Indicates a valid entry in the table\\
+         Tag & 6 & Identifier for the entry in the table\\
          Useful & 1 & To show if the given entry led to a useful demand fetch\\
@@
     \end{tabular}
-    \caption{Meta-data Stored in Prefetch Filter}
-    \label{tab:PF_metadata}
+    \caption{Meta-data Stored in Prefetch Table}
+    \label{tab:PTable_metadata}
 \end{table}
@@
         \multirow{3}{1.5cm}{Perceptron\newline}     & 4096*8    &            &              \\
-        \multirow{2}{0.9cm}{Tables}                 & 2048*x    & 5 bits     & 164480 bits  \\
+        \multirow{2}{0.9cm}{Weights}                 & 2048*x    & 5 bits     & 164480 bits  \\
                                                     & 128*1     &            &              \\
     \hline
-        Prefetch Filter\footnotemark[1]             & 1024      & 84 bits    & 86016 bits   \\
+        Prefetch Table\footnotemark[1]             & 1024      & 84 bits    & 86016 bits   \\
     \hline
-        Reject Filter\footnotemark[2]               & 1024      & 83 bits    & 84992 bits   \\
+        Reject Table\footnotemark[2]               & 1024      & 83 bits    & 84992 bits   \\
     \hline
@@
     \caption{SPP-Perc Storage Overhead}
-    \label{tab:SPPPerc_overhead}
+    \label{tab:PPF_overhead}
 \end{table}
 
-\footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}
-\footnotetext[2]{RF does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
+\footnotetext[1]{Components of Prefetch Table can be found in Table \ref{tab:PTable_metadata}}
+\footnotetext[2]{the Reject Table does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
 
 In this section, we analyze the hardware overhead required to
-implement PPF.  The Prefetch Filter was enhanced to accommodate
+implement PPF.  The Prefetch Table was enhanced to accommodate
 storing of meta-data for perceptron training.  Table
-\ref{tab:PF_metadata} depicts the meta-data stored for each entry in
-the Prefetch Filter.  Table \ref{tab:PPF_overhead} shows the total
+\ref{tab:PTable_metadata} depicts the meta-data stored for each entry in
+the Prefetch Table.  Table \ref{tab:PPF_overhead} shows the total
 storage overhead of PPF implementation.  The hardware budget for
@@
 update by +1 or -1.  Thus, all the operations required for perceptron
-inferencing or updating the states of the perceptron throttler can be
+inferencing or updating the states of the perceptrons can be
 easily done in the time constraints of L2 Cache Accesses.

### 2018-11-29 abs.tex
Gino's first pass, modifying the abs, intro, design_arch, and design_impl

@@
 Hardware prefetching has been introduced in modern processors as a way to hide
-cache latencies.  An efficient prefetcher should be able to identify complex
+cache latencies.  An efficient prefetcher should identify complex
 memory access patterns during program execution. This ability enables the
-prefetcher to read a block ahead of its demand access, potentially saving a
+prefetcher to read a block ahead of its demand access, potentially preventing a
 cache miss. Accurately identifying the right blocks to prefetch is essential
@@
 In this paper, we introduce Perceptron-based Prefetch Filtering to help make
-this prefetching decision accurately. The perceptron layer acts as a check 
+ prefetching decisions accurately. The perceptron layer acts as a check 
 to filter out the unnecessary prefetches recommended by the underlying
@@
 to a state-of-the art prefetcher.  We also demonstrate that the performance
-gained from using our efficient filter scales better with increasing numbers
-of cores.
+gained from using our efficient filter continues to scale as the number of cores sharing a 
+last level cache increases.
 

### 2018-11-29 design_arch.tex
Gino's first pass, modifying the abs, intro, design_arch, and design_impl

@@
 In this section, we talk about the the general structure of the 
-Perceptron-Based Prefetch Filtering (PPF) for an underlying prefetcher.  
+Perceptron-Based Prefetch Filtering (PPF) for enhancing an underlying prefetcher.  
 PPF is a generalised prefetch filtering mechanism and can be adapted to 
-most prefetchers with minimal modifications.
+most prefetchers with minimal modifications. %Gino: Are these modifications minimal?
 
@@
 \label{Arch-Perceptron}
-The Perceptron Filter is organised as tables of perceptron weights, 
-with a separate table for each feature.  
-The feature value is used to index into the corresponding table.
-This indexing is done using different numbers of bits for each feature.  
-At most 12 bits of a feature are used to index into the
-weights table.  Certain features require more resolution power
+The Perceptron Filter is organized as tables of perceptron weights, 
+with a separate table of weights for each feature.  
+A different number of bits are used for each feature to index into its 
+corresponding table.
+%Gino: repharse below?
+Each feature's value is used to index into the corresponding table.
+This indexing is done using a different number of bits for each feature.  
+At most, 12 bits of a feature are used to index into the
+weights table. Certain features require more resolution
 \textit{i.e.}, the full 12 bits of indexing.  On the other hand, some
-features requiring lower resolution require as few as 7 bits of
-indexing.  The variable indexing was determined by studying the
-features and fine-tuned empirically so as to achieve a good accuracy
-vs area trade-off.  Exact details are mentioned under ``Area Overheads''
+features requiring lower resolution require as few as 7 bits of indexing. 
+The variable indexing was determined by studying the
+features and fine-tuned empirically to achieve a good accuracy
+vs hardware overhead trade-off.  Exact details are mentioned under ``Area Overheads''
 in Section~\ref{Method-Overheads}
@@
 A single entry in the table corresponds to a perceptron weight.  Each
-weight is a 5 bit counter - saturating at -16 and +15.  
-Our proposed design uses nine features, hence there are nine perceptron 
-weight tables. As program execution begins, all the weights are initialized to 0.
+weight is a 5 bit counter - saturating at -16 and +15 and initialized to zero at the 
+beginning of a program's execution. Our proposed design uses nine features, 
+represented in nine perceptron weight tables.
 
@@
 invoked on every L2 Cache demand access.  At that point, the base prefetcher
-has the option to do or not do a prefetch -- if it does, then it has a choice
-in how many cache lines to prefetch.  The suggested prefetches can be either
-in the L2 Cache or the last-level cache.  The prefetcher decides this
-placement, and it is usually done based on an internal confidence mechanism.
+has the option to perform a prefetch -- if it does, then it has a choice
+of how many cache lines to prefetch.  The suggested prefetches can be either placed
+in the L2 cache or the last-level cache based on the prefetcher's internal confidence mechanism.
 
-When the base prefetcher is triggered, it starts suggesting candidates for
-prefetching.  All these recommended prefetches have to be tested through the perceptron
-filter to qualify for final prefetching.
+When the base prefetcher is triggered, it begins to suggest candidates for
+prefetching. All suggested prefetches are passed through the perceptron
+filter to decide if they are qualified for final prefetching. 
+he perceptron decides whether to prefetch a candidate by looking at the microarchitectural state 
+\em} i.e. the features, at that instant. Each feature is hashed to form an index into a table of up to 
+4096 entries dedicated for that feature.
 
-The perceptron looks at the microarchitectural state, {\em i.e.} the features,
-at that instant.  Each feature is hashed to form an index into a table of up to 
-4096 entries dedicated for that feature.  
-Feature \#1 indexes into a particular entry in table \#1 and so on.
+Once all the weights are retrieved, they are summed and compared to a preset threshold 
+,PERC\_ THRESHOLD\_LO.  Only the prefetch candidates with perceptron sum higher than 
+the threshold qualify for prefetching.  The prefetches that qualify through the perceptron 
+stage are recorded in the ``Prefetch Table". The prefetch table is a 1024-entry, direct mapped structure taht
+conatains all metadata required to recreate the features. 	%Gino: Recreate the features?
+When the feedback of the current prefetched line is available at a later stage, the
+stored data is used to train the perceptron.
 
-Once all the weights are retrieved, they are summed.  The sum is thresholded
-based on the preset threshold PERC\_ THRESHOLD\_LO.  Only the prefetch
-candidates with perceptron sum higher that the threshold qualify for
-prefetching.  The prefetches that qualify through the perceptron 
-stage are recorded in the Prefetch Table.  In PPF, all the metadata that is required
-to recreate the features is also stored in the Prefetch Table.  At a later
-stage, when the feedback of the current prefetched line is available, the
-stored data can be used to train the perceptron.
-
-The perceptron sum is also used to decide L2 Cache vs Last Level Cache
-placement of the prefetch cache line.  All prefetch candidates qualified till
-this stage are thresholded against PERC\_THRESHOLD\_HI to decide the fill
+The sum of the perceptron weights are also used to decide whether to place teh 
+data in the L2 or the last level cache. All prefetch candidates that are qualified until
+this stage are compared against PERC\_THRESHOLD\_HI to decide the fill
 level. The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are
@@
 
-The perceptron sum can also be seen as the sum of the individual contribution
-per feature.  The value of each contribution corresponds to the extent of
-confidence of the final output with that given feature.  By summing the
+The perceptron sum can be considered the sum of the individual contribution
+of a feature.  The value of each contribution corresponds to a feature's
+confidence in the final decision on whether to prefetch a cache line. By summing the
 individual contributions, the final perceptron sum denotes the overall
-confidence for that prefetch suggestion.  By thresholding the perceptron sum
-on two different thresholds, we are dividing the confidence scale of the
+confidence for that prefetch suggestion.  By comparing the perceptron sum
+to two different thresholds, we divide the confidence scale of the
 prefetch suggestion into three bins. The first bin corresponds to the lowest
-confidence leads to prefetch candidate being rejected.  The next bin
-corresponds to prefetches with a moderate confidence level.  Such prefetches
-are directed to fill the bigger Last Level Cache and potentially not pollute
-the more scarce L2 Cache.  The highest confidence prefetches are filled in the
-L2 Cache.
+confidence and leads to prefetch candidate being rejected.  The next bin
+corresponds to prefetches with a moderate confidence level.  This bin prefetches
+ directeky to fill the bigger last level cache and prevents polluting
+the smaller L2 Cache.  The highest confidence prefetches are filled in the
+L2 cache.
 
-In addition to the Prefetch Table mentioned above, PPF also maintains a
-``Reject Table.''  The Reject Table is a 1024-entry deep direct-mapped
+In addition to the prefetch table mentioned above, PPF also maintains a
+``Reject Table.''  The reject table is a 1024-entry deep direct-mapped
 table.  If a prefetch suggestion is rejected by the perceptron layer, it is
-logged into the Reject Table.  The filer is used to train the perceptron to
+logged into the reject table.  The filer is used to train the perceptron to %Gino: Filer?
 avoid false negatives \textit{i.e.}, cases where prediction was to reject the
@@
 L2 Cache.  This action triggers training of the base prefetcher as well as the
-perceptron layer.  Training the perceptrons involves restoring
-the metadata that was stored in Prefetch Table or the Reject Table.  This is
-done by indexing into the corresponding tables using the cache line address
-of the block being used to train. (10-bits indexing + 6-bits tag matching).
+perceptron layer.  Training the perceptrons involves accessing
+the metadata that was stored in Prefetch Table or the Reject Table. 
+The cache lines address of the block being trained is used to access a table,
+using 10-bits of the address for indexing and another 6-bits to perform tag matching.
 Once the state of the program at the time of prefetching is available, it is 
+%Gino: By "state of the program" are you referring to the state of the features?
 used to index into the perceptron weights table.
 
-If a demand access block that triggers the training was tagged as a valid
-prefetch in the Prefetch Table, then the earlier prefetch prediction was
+If the demanded block that triggers the training was tagged as a valid
+prefetch in the prefetch table, then the earlier prefetch prediction was
 correct.  In that case the perceptron weights are incremented by 1
 if the predicted sum does not cross a pre-defined threshold. These training 
-thresholds are introduced to limit the degree of update on correct predictions.
-Doing this helps avoid overfitting of the perceptron weights to the given program 
+thresholds are introduced to avoid overfitting of the perceptron weights to the given program 
 behaviour. These thresholds are referred to as POS\_UPDATE\_THRESHOLD ($\theta_p$) and 
@@
 
-If a cache block
-eviction led to training and the corresponding valid entry was found in the
-Prefetch Table, then the prediction made by the perceptron was wrong.  The
-perceptron should have ideally rejected the prefetch suggestion as a
+If a cache block eviction led to training and the corresponding valid entry is found in the
+prefetch table without the block being accessed, then the prediction made by the perceptron was wrong.
+The perceptron should have ideally rejected the prefetch suggestion as a
 low-confidence prefetch.  Here the weights are decremented by 1 to reflect the
-misprediction. In either case, weights are saturated at -16 or +15.
+misprediction. In either case, weights are saturated at -16 or +15. %Gino: Describe how these limits were decided?
 
 A secondary training mechanism also kicks in during demand fetches.  Before
-the demand access triggers the next set of prefetches, Reject Table is
+the demand access triggers the next set of prefetches, the reject table is
 checked for a valid entry.  A hit means that the corresponding cache line was
-initially suggested by the underlying prefetch engine but rejected by the
+initially suggested by the underlying prefetch engine, but rejected by the
 perceptron filter.  Thus, the perceptron should have been more confident
-about that particular prefetch.  Once such a scenario is identified, the state
-of the execution at the time of prefetch is retrieved from the Reject Table.
+about that particular prefetch.  Once such a scenario is identified, the state %Gino: state of execution == value of features?
+of the execution at the time of prefetch is retrieved from the reject table.
 The retrieved data is used to index into the various weights tables of the
@@
 
-This mechanism allows us to solve the classical problem of exploiting a lost
-opportunity.  In prior perceptron based implementations (and in general,
-prefetching algorithms), there is usually no way of knowing the result of not
+This mechanism allows us to exploit a previously lost opportunity.  
+In prior perceptron based implementations, and in general,
+prefetching algorithms, there is usually no way of knowing the result of not
 prefetching a particular line.  Our two-step PPF architecture allows us to
@@
 
-\item \textbf{Inferencing and Storing:} All prefetch recommendations to be
-tested using the perceptron inferencing algorithm.  Perceptron output:
-\textit{true} and \textit{false} should be saved appropriately, along with all
-the metadata required for perceptron indexing.
+\item \textbf{Inferencing and Storing:} All prefetch recommendations are
+tested using the perceptron inferencing algorithm. The perceptron's output,
+\textit{true} or \textit{false}, should be saved appropriately, along with all
+ metadata required for perceptron indexing.
 
-\item \textbf{Retrieving and Training:} When the feedback is available, the
+\item \textbf{Retrieving and Training:} When feedback for a prefetch is available, the
 previously stored metadata can be used to re-index into the perceptron entries
-and increment / decrement the weights.
+and increment or decrement the weights.
 
@@
 prefetcher. Beyond that, the feature set can be expanded to convey any
-useful information from the prefetcher to the perceptron layer.  The 
+useful information from the prefetcher to the perceptron filter.  The 
 methodology explained in Section~\ref{Method-Features} talks about developing a minimal 

### 2018-11-29 design_impl.tex
Gino's first pass, modifying the abs, intro, design_arch, and design_impl

@@
 This section describes our implementation of PPF and the range of 
-features that are ultimately used by the perceptron for correlating 
+features that are used by the perceptron for correlating 
 the prefetches.  For a practical implementation, we have used SPP 
@@
 \item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The original
-  SPP would not consider the slots available in the L2 MSHR queue.  It had an
-  assumption that at no time can the number of prefetches suggested can
-  exceed the capacity of the queue.  While it was not proven in the paper, the
-  internal confidence mechanism made sure that the above assumption was maintained.  
-  In PPF since there can be no assumption on the degree of aggressiveness of the 
+  SPP does not consider the slots available in the L2 MSHR queue.  It has the
+  assumption that at no time can the number of suggested prefetches
+  exceed the capacity of the queue.  While it was not described in the paper, the
+  internal confidence mechanism makes sure that the above assumption was maintained.  
+  In PPF, since there cannot be an assumption on the degree of aggressiveness of the 
   underlying prefetcher, we explicitly check that at no point should the number 
@@
 
-\item \textbf{Enhanced Prefetch Table}\newline The Prefetch Table introduced in the 
-  original SPP keeps a track of up to 1024 L2 cache prefetch suggestions.  
-  For PPF, the Table was modified to store metadata such as program counter, 
-  memory access address etc.  This metadata is needed to retrieve the state 
-  of the features to train the perceptron at a later stage in time.
-
-\item \textbf{Reject Table introduced}\newline The Prefetch Table mentioned
-  above keeps track of the prefetches that actually happened in the L2 Cache.
-  Correspondingly, we introduced the Reject Table to keep a track of the
-  prefetches that were suggested by the base SPP engine but were rejected by the
-  perceptron.  In case it is detected that a future demand fetch could have used
+\item \textbf{Enhanced Prefetch Table}\newline The prefetch table introduced in the 
+  original SPP keeps track of up to 1024 L2 cache prefetch suggestions.  
+  For PPF, the prefetch table is modified to store metadata such as program counter, 
+  memory access address etc.  This metadata is needed to retrieve the state  %Gino: What's the etc.?
+  of the features to train the perceptron at a later stage in time. The prefetch table 
+  mentioned above keeps track of the prefetches that actually happened in the L2 Cache.
+  We introduce the reject table to keep a track of the
+  prefetches that were suggested by the base SPP engine but were rejected by PPF.  
+  In case it is detected that a future demand fetch could have used
   this rejected prefetch, the perceptron weights are updated to reflect that
@@
   prefetch.  Since prefetches are triggered from the L2 demand access,
-  they tend to be correlated with the triggering address.
+  they tend to correlate to the triggering address.
 
@@
   in the increasing order of history before being hashed together.
-  This is done to avoid the resultant value of zero in case 2 or more
+  This is done to avoid the resultant value of zero when 2 or more
   PCs are the same.  Additionally, obfuscating the information as it
-  gets old allows us to get a wider and an approximate look into the
-  program history.
+  gets old allows us to get a wider and approximate look into the
+  program's history.
 
@@
   the discussion of SPP in Section \ref{Background-SPP} that the new signature
-  is generated using the old signature and the block delta.  While ``Current
-  Signature XOR Delta'' is not the exact formula for generating the future
+  is generated using the old signature and the current block delta.  
+%Gino: Added the sentence below, is this correct?
+  The result of this feature is the next delta that is predicted to be accessed based on the 
+  delta predicted by SPP. While ``Current Signature XOR Delta'' is not the exact formula for generating the future
   signature, it gives an approximate idea of the path that the combination of

### 2018-11-29 intro.tex
Gino's first pass, modifying the abs, intro, design_arch, and design_impl

@@
 \label{Introduction}
-Processor scaling and memory scaling have been enabled with different
+Processor scaling and memory technologies have been developed with different
 philosophies in mind.  While processor scaling focused on speed improvements,
-memory scaling was centred around increasing the storage capacity.  Process
-technology reaching scaling limitations led to the Memory Wall~\cite{MemWall}
--- the increasing performance gap between processor speed and the memory
-speeds.  Various techniques have been developed to combat this wide gap. Data
-prefetching is one such important technique. Data prefetching exploits the
-fact that in most applications, memory accesses have a well-defined pattern.
+memory scaling was centred around increasing the storage capacity.  The
+difference in each technology's scaling has led to the Memory Wall~\cite{MemWall}
+-- the increasing performance gap between processor and memory
+speeds. Data prefetching is one important technique that has been developed 
+to bridge this widening performance gap. Data prefetching exploits the fact that in most applications,
+ memory accesses have a repeating and predictable pattern to speculatively fetch useful data from slower levels
+of the cache hierarchy into faster higher levels of cache. %phrasing?
 
-An ideal prefetching scheme would capture these memory access patterns to
-predict future accesses in a timely manner.  Other important decisions faced
+An ideal prefetching scheme would perfectly capture memory access patterns to
+predict and request future accesses in a timely manner. Other important decisions faced
 by the prefetcher are when to bring in the prefetch block of data, which level
@@
 fetching the next cache line or a complex pattern such as pointer chasing.
-Predicting future accesses requires a trade-off between coverage and accuracy.
+Predicting future accesses requires a trade-off between coverage, the percent of predictions that 
+prevented a cache miss, and accuracy, the percent of predictions made that were correct. 
+%Gino: Just added a short explanation of coverage and accuracy tradeoff, not sure if necessary
+A prefetcher may have high coverage by reducing the number of misses by simply requesting a large 
+amount of data, however this does not imply accuracy. Even if the prefetcher brought in data accessed in the future,
+a high percentage of the data requested could never be accessed meaning the prefetcher had low accuracy, 
+resulting in wasted resources as well as cache pollution. Likewise, a prefetcher can have a 
+higher accuracy by being completely sure the data requested is used in the future, 
+but may not affect performance since it did not have the coverage needed to make an impact. %end addition.
 Most prefetchers maintain an internal confidence mechanism by keeping counters
-to track events such as cache hits / misses, references to an entry in the
-table, etc. By thresholding this confidence against different values, the
-degree of aggressiveness can be adjusted.
+to track events such as cache hits, misses, references to an entry in the prefetcher's
+structure, etc. By comparing the confidence indicated by the counters against different 
+threshold values, the aggressiveness of the prefetcher can be adjusted, resulting in lower
+coverage and higher accuracy.
 
 For a prefetcher to capture highly irregular access patterns, it needs to be
-highly aggressive (high coverage).  Thus, the prefetcher would recommend many
-useless prefetches, leading to cache pollution (low accuracy).  This effect is
-even more pronounced in a multi-core scenario where the last-level cache is a
+highly aggressive with high coverage. This causes the prefetcher to recommend many
+low accurracy prefetches that lead to the cache being polluted with useless data.  
+This effect is more pronounced in a multi-core scenario where the last-level cache is a
 shared resource~\cite{Friendly}.  Conversely, a conservative prefetcher would
-be highly accurate but might not make enough useful prefetches.
+be highly accurate but might not make enough useful prefetches due to its lower coverage.
 
-We propose Perceptron-based Prefetch Filtering (PPF).  It is an enhancement to
-the existing state of the art prefetcher and robustly overcomes the coverage
+We propose Perceptron-based Prefetch Filtering (PPF) as an enhancement to
+the existing state of the art prefetchers to robustly overcome the coverage
 vs. accuracy trade-off.  The idea is to use a state-of-the-art prefetcher as
@@
 enables the base engine to capture complex memory access patterns and go
-deeper into the speculation path, leading to increased coverage. Normally such
-aggressive prefetching would come at the cost of increased DRAM traffic.
-However, prefetch suggestions recommended by the base prefetch engine are
-passed through the perceptron based filter.  Over time, perceptron layer
-learns to correlate the prefetch recommendations with various available
+deeper into the speculation path, increasing coverage without sacrificing accuracy. 
+Normally such aggressive prefetching would come at the cost of increased DRAM traffic
+and cache pollution. However, prefetch suggestions recommended by the base prefetch
+engine are passed through the perceptron based filter to ensure an accurate prediction.
+Over time, PPF learns to correlate the prefetch recommendations with various available
 features, leading to elimination of useless prefetches. This filtering leads
 to overall increased accuracy of the prefetch scheme and reduction in DRAM
-traffic.
+traffic and cache pollution.
 
@@
 
-\item An on-line neural model is used for hardware data
+\item An on-line neural model used for hardware data
   prefetching.  Previous work in this area either relied on program
@@
   practical prefetching mechanism, yielding a significant performance
-  improvement as compared to the other state of the art prefetchers.
+  improvement compared to other state of the art prefetchers.
   Moreover, the perceptron learns to adapt itself to shared resource
   constraints, leading to further increased performance in multi-core and
-  lower capacity environments.
+  lower memory capacity environments.
 
-\item Defining a methodology for determining an appropriate set of 
+\item A methodology for determining an appropriate set of 
   features for prediction regardless of the underlying prefetcher used.
@@
 
-On a single core configuration, PPF increased performance by 6.84\% as
-compared to the next best prefetcher, SPP.  On a multi-core system running a
+On a single core configuration, PPF increases performance by 6.84\%
+compared to the baseline prefetcher, SPP.  On a multi-core system running a
 mix of memory intensive subset of the SPEC 2017 traces, PPF saw an improvement

### 2018-11-30 abs.tex
made abstract more "punchy" and direct

--- 2018-11-20 abs.tex
@@
 \begin{abstract}
-Hardware prefetching has been introduced in modern processors as a way to hide
-cache latencies.  An efficient prefetcher should be able to identify complex
-memory access patterns during program execution. This ability enables the
-prefetcher to read a block ahead of its demand access, potentially saving a
-cache miss. Accurately identifying the right blocks to prefetch is essential
-to achieving high performance from the prefetcher.
 
-In this paper, we introduce Perceptron-based Prefetch Filtering to help make
-this prefetching decision accurately. The perceptron layer acts as a check 
-to filter out the unnecessary prefetches recommended by the underlying
-prefetcher.  We have also explored a range of features that can be used to
-train the perceptron layer.  Our results show that perceptron-based filtering
-improves performance on the memory intensive subset of the SPEC 2017 benchmark
-suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared
-to a state-of-the art prefetcher.  We also demonstrate that the performance
-gained from using our efficient filter scales better with increasing numbers
-of cores.
+Hardware prefetching is an effective technique for hiding cache miss latencies
+in modern processor designs.  Prefetcher performance can be characterized by 
+two main metrics, which are generally at odds with one another: coverage 
+(the improvement in the cache miss metric) and accuracy (the fraction of 
+useful prefetches).  An overly aggressive prefetcher may improve coverage, but
+may waste so many resources (e.g., bandwidth) that performance is ultimately 
+harmed.  An ideal prefetcher would have both very high coverage and accuracy.
+
+%Hardware prefetching has been introduced in modern processors as a way to hide
+%cache latencies.  An efficient prefetcher should identify complex
+%memory access patterns during program execution. This ability enables the
+%prefetcher to read a block ahead of its demand access, potentially preventing a
+%cache miss. Accurately identifying the right blocks to prefetch is essential
+%to achieving high performance from the prefetcher.
+
+In this paper, we introduce Perceptron-based Prefetch Filtering (PPF) as a way
+to reject inaccurate candidate prefetches generated by a baseline prefetcher.  
+PPF directly increases prefetcher accuracy, and enables more aggressive tuning
+of the baseline prefetcher, which in turn leads to increased coverage.  We also
+explore a range of features to use to train PPF's perceptron layer.  Our results
+show that PPF improves performance on a memory-intensive subset of SPEC CPU 2017
+benchmarks by 6.84\% for a single-core configuration, and by 11.9\% for a 4-core
+configuration, compared to the baseline prefetcher alone.  We also show that
+performance continues to scale as the number of cores sharing a last level cache
+increases.
+
+%In this paper, we introduce Perceptron-based Prefetch Filtering to help make
+% prefetching decisions accurately. The perceptron layer acts as a check 
+%to filter out the unnecessary prefetches recommended by the underlying
+%prefetcher.  We have also explored a range of features that can be used to
+%train the perceptron layer.  Our results show that perceptron-based filtering
+%improves performance on the memory intensive subset of the SPEC 2017 benchmark
+%suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared
+%to a state-of-the art prefetcher.  We also demonstrate that the performance
+%gained from using our efficient filter continues to scale as the number of cores sharing a 
+%last level cache increases.
 

### 2018-12-01 intro.tex
reworded almost everything

--- 2018-11-20 intro.tex
@@
 \label{Introduction}
-Processor scaling and memory scaling have been enabled with different
-philosophies in mind.  While processor scaling focused on speed improvements,
-memory scaling was centred around increasing the storage capacity.  Process
-technology reaching scaling limitations led to the Memory Wall~\cite{MemWall}
--- the increasing performance gap between processor speed and the memory
-speeds.  Various techniques have been developed to combat this wide gap. Data
-prefetching is one such important technique. Data prefetching exploits the
-fact that in most applications, memory accesses have a well-defined pattern.
 
-An ideal prefetching scheme would capture these memory access patterns to
-predict future accesses in a timely manner.  Other important decisions faced
-by the prefetcher are when to bring in the prefetch block of data, which level
-of the cache hierarchy to place the block in, and what block to evict to
-accommodate the incoming block.  Memory accesses patterns can be as simple as
-fetching the next cache line or a complex pattern such as pointer chasing.
-Predicting future accesses requires a trade-off between coverage and accuracy.
-Most prefetchers maintain an internal confidence mechanism by keeping counters
-to track events such as cache hits / misses, references to an entry in the
-table, etc. By thresholding this confidence against different values, the
-degree of aggressiveness can be adjusted.
+Processor and memory technologies have been developed with different
+goals in mind.  While processor scaling has focused on speed improvements,
+memory scaling has primarily focused on increasing capacity.  The
+difference in each technology's scaling has led to the Memory Wall~\cite{MemWall}
+-- the increasing gap between processor and memory performance. Data prefetching
+is one important technique that has been developed to minimize the effects 
+of this trend.
 
-For a prefetcher to capture highly irregular access patterns, it needs to be
-highly aggressive (high coverage).  Thus, the prefetcher would recommend many
-useless prefetches, leading to cache pollution (low accuracy).  This effect is
-even more pronounced in a multi-core scenario where the last-level cache is a
-shared resource~\cite{Friendly}.  Conversely, a conservative prefetcher would
-be highly accurate but might not make enough useful prefetches.
+%Data prefetching exploits the fact that in most applications,
+% memory accesses have a repeating and predictable pattern to speculatively fetch useful data from slower levels
+%of the cache hierarchy into faster higher levels of cache. %phrasing?
 
-We propose Perceptron-based Prefetch Filtering (PPF).  It is an enhancement to
-the existing state of the art prefetcher and robustly overcomes the coverage
-vs.  accuracy trade-off.  The idea is to use a state-of-the-art prefetcher as
-the base engine.  For that we choose Signature Path Prefetcher (SPP) and
-modify the design to make it as aggressive as possible.  This modification
-enables the base engine to capture complex memory access patterns and go
-deeper into the speculation path, leading to increased coverage. Normally such
-aggressive prefetching would come at the cost of increased DRAM traffic.
-However, prefetch suggestions recommended by the base prefetch engine are
-passed through the perceptron based filter.  Over time, perceptron layer
-learns to correlate the prefetch recommendations with various available
-features, leading to elimination of useless prefetches. This filtering leads
-to overall increased accuracy of the prefetch scheme and reduction in DRAM
-traffic.
+An ideal prefetching scheme would perfectly capture a program's memory access pattern, and then
+predict and pre-load the needed data into the processor's caches in a timely manner. 
+Memory access patterns may be simple, such as accessing every item in an array with a for-loop,
+or very complex, such as chasing pointers through dynamically-allocated memory.
+Practical prefetchers face challenges not only in predicting what data will be useful in the future,
+but also in timing when data should be prefetched, deciding which level of the cache hierarchy
+the data should be stored in, and deciding what data should be evicted from the caches to accomodate
+the prefetched data.
 
-\vspace{1ex}This paper describes PPF, giving its merits, analysis and
-future scope for expansion.  The contributions include:
+All prefetchers are designed around a fundamental trade-off between two important metrics:
+coverage and accuracy.  Prefetcher coverage refers to the number of cache hits that would have
+been cache misses without the prefetcher.  For example, if an application experiences 1,000 cache misses
+without a prefetcher, but only 200 cache misses with a prefetcher, then the prefetcher has 80\%
+coverage for that application.  Prefetcher accuracy refers to the fraction of prefetched cache lines
+that end up being used by the application.  So if a prefetcher prefetches 1,200 cache lines, 
+but only 800 of them are used by the application, then that prefetcher's accuracy is 66.7\%.
+
+Coverage and accuracy are generally at odds with one another, and as one metric improves, the other
+usually gets worse.  For example, when an application accesses a new region of memory for the first
+time, a na\"ive prefetcher may predict that all data in that region will be used by the application.
+This will clearly result in 100\% coverage for that region, but with possibly a very low accuracy.
+In fact, so much bandwidth may be wasted prefetching unused data that performance can ultimately be
+harmed by this strategy.  On the other extreme, another prefetcher may be overly
+conservative and never prefetch anything, wasting no bandwidth, and achieving 0\% prefetch
+coverage.
+
+A delicate balance between coverage and accuracy is required for a prefetcher to maximize
+its performance impact.  Prefetchers are generally designed with internal mechanisms to
+monitor their accuracy, with throttling thresholds that can be tuned for either coverage 
+or accuracy.  The more irregular an application's memory access pattern is, the more difficult
+it is to accurately predict every access, so a prefetcher will have to be tuned more toward
+coverage (and away from accuracy) in order to gain any benefit.  This may be especially dangerous
+to do in the context of a multi-core processor, where overly aggressive prefetching in one core can
+waste shared resources, such as last-level cache capacity, and off-chip bandwidth~\cite{Friendly}.
+
+%Predicting future accesses requires a trade-off between coverage, the percent of predictions that 
+%prevented a cache miss, and accuracy, the percent of predictions made that were correct. 
+%%Gino: Just added a short explanation of coverage and accuracy tradeoff, not sure if necessary
+%A prefetcher may have high coverage by reducing the number of misses by simply requesting a large 
+%amount of data, however this does not imply accuracy. Even if the prefetcher brought in data accessed in the future,
+%a high percentage of the data requested could never be accessed meaning the prefetcher had low accuracy, 
+%resulting in wasted resources as well as cache pollution. Likewise, a prefetcher can have a 
+%higher accuracy by being completely sure the data requested is used in the future, 
+%but may not affect performance since it did not have the coverage needed to make an impact. %end addition.
+%Most prefetchers maintain an internal confidence mechanism by keeping counters
+%to track events such as cache hits, misses, references to an entry in the prefetcher's
+%structure, etc. By comparing the confidence indicated by the counters against different 
+%threshold values, the aggressiveness of the prefetcher can be adjusted, resulting in lower
+%coverage and higher accuracy.
+
+%For a prefetcher to capture highly irregular access patterns, it needs to be
+%highly aggressive with high coverage. This causes the prefetcher to recommend many
+%low accurracy prefetches that lead to the cache being polluted with useless data.  
+%This effect is more pronounced in a multi-core scenario where the last-level cache is a
+%shared resource~\cite{Friendly}.  Conversely, a conservative prefetcher would
+%be highly accurate but might not make enough useful prefetches due to its lower coverage.
+
+We propose Perceptron-based Prefetch Filtering (PPF) as an enhancement to
+existing state-of-the-art prefetchers.  PPF works by observing the stream of
+candidate prefetches generated by a prefetcher, and then rejects those that
+are predicted by the online-trained neural model to be inaccurate.
+The state-of-the-art prefetcher that we 
+use to evaluate PPF in this paper is the Signature Path Prefetcher 
+(SPP)~\cite{SPP}, however in principle, PPF could benefit any prefetcher.
+In this design, PPF replaces SPP's existing confidence-based throttling mechanism,
+which itself was a highly tuned feature of that prefetcher.  Because PPF is
+so much more effective at rejecting inaccurate prefetches than SPP's baseline
+mechanism, we are free to re-tune the rest of SPP's design around maximizing
+coverage.  The result is an increase in both accuracy and coverage, and
+a notable increase in performance.
+
+%We propose Perceptron-based Prefetch Filtering (PPF) as an enhancement to
+%the existing state of the art prefetchers to robustly overcome the coverage
+%vs. accuracy trade-off.  The idea is to use a state-of-the-art prefetcher as
+%the base engine. For that we choose Signature Path Prefetcher (SPP) and
+%modify the design to make it as aggressive as possible.  This modification
+%enables the base engine to capture complex memory access patterns and go
+%deeper into the speculation path, increasing coverage without sacrificing accuracy. 
+%Normally such aggressive prefetching would come at the cost of increased DRAM traffic
+%and cache pollution. However, prefetch suggestions recommended by the base prefetch
+%engine are passed through the perceptron based filter to ensure an accurate prediction.
+%Over time, PPF learns to correlate the prefetch recommendations with various available
+%features, leading to elimination of useless prefetches. This filtering leads
+%to overall increased accuracy of the prefetch scheme and reduction in DRAM
+%traffic and cache pollution.
+
+\vspace{1ex}This paper describes PPF, explains its merits, offers analysis, and
+outlines the scope for future research. Its contributions include:
 
@@
 
-\item Implementing the perceptron filter in a robust and accurate
+\item Implementing PPF in a robust and accurate
   practical prefetching mechanism, yielding a significant performance
-  improvement as compared to the other state of the art prefetchers.
-  Moreover, the perceptron learns to adapt itself to shared resource
+  improvement compared to other state-of-the-art prefetchers.
+  Moreover, PPF learns to adapt itself to shared resource
   constraints, leading to further increased performance in multi-core and
-  lower capacity environments.
+  bandwidth-constrained environments.
 
-\item Defining a methodology for determining an appropriate set of 
-  features for prediction regardless of the underlying prefetcher used.
+\item A methodology for determining an appropriate set of 
+  features for prediction, regardless of the underlying prefetcher used.
   More details are explained in Section \ref{Method-Features}.
@@
 
-On a single core configuration, PPF increased performance by 6.84\% as
-compared to the next best prefetcher, SPP.  On a multi-core system running a
-mix of memory intensive subset of the SPEC 2017 traces, PPF saw an improvement
-of 11.9\% over SPP for a 4-core system and 10.67\% for an 8-core system.
+In a single core configuration, PPF increases performance by 6.84\%
+compared to the baseline prefetcher, SPP.  In a multi-core system running a
+mixes of memory intensive SPEC CPU 2017 traces, PPF saw an improvement
+of 11.9\% over SPP for a 4-core system, and 10.67\% for an 8-core system.
 
 The paper is organized as follows. Section~\ref{Background} describes
-background work and summarizes SPP.  Section~\ref{Arch} given the 
+background work and summarizes SPP.  Section~\ref{Arch} gives the 
 architectural overview of PPF. Section~\ref{Impl} discusses the
-implementation of PPF using SPP and the features used.
+implementation of PPF using SPP and the particular features used.
 Section~\ref{Method} gives the evaluation methodology and explores the feature

### 2018-12-01 method.tex
Minor change
Updated results section & added graphs

@@
 of cache hierarchies. Branch prediction is done using the perceptron branch
-predictor~\cite{Perc_Branch}.  ThE page size is 4KB.  Champsim operates
+predictor~\cite{Perc_Branch}.  The page size is 4KB.  Champsim operates
 all the prefetchers strictly in the physical address space.
@@
 %has been used to characterize and measure the prefetch performance.
-We use the that SPEC 2017 benchmark suite~\cite{SPEC2017}.  We use all the 20
+We use the SPEC 2017 benchmark suite~\cite{SPEC2017}.  We use all the 20
 workloads available in the SPEC 2017 suite.  Using the SimPoint~\cite{SimPoint}
@@
 
-\begin{table}[]
+\begin{table}[t]
     \centering
@@
 
-\begin{table}[]
+In this section, we analyze the hardware overhead required to
+implement PPF.  The Prefetch Table was enhanced to accommodate
+storing of meta-data for perceptron training.  Table
+\ref{tab:PTable_metadata} depicts the meta-data stored for each entry in
+the Prefetch Table.  Table \ref{tab:PPF_overhead} shows the total
+storage overhead of PPF implementation.  The hardware budget for
+2nd Data Prefetching championship was 32 KB.  Keeping that in mind 
+the considerable speedup PPF obtained over the winner, the extra hardware
+budget can be accounted for.  The extra hardware also makes the
+overall scheme more scalable than SPP.  In the original SPP paper, it
+was demonstrated that adding extra hardware brings little advantage in
+terms of performance gain.  The newly added perceptron tables can be
+scaled to increase / decrease features depending on the permitted
+budget.
+
+\begin{table}[h]
+\begin{adjustwidth}{-0.5cm}{}
     \centering
@@
     \hline
-        \multirow{5}{2.2cm}{Signature Table} &      & Valid (1 bit)        &             \\
+        \multirow{5}{1.5cm}{Signature Table} &      & Valid (1 bit)        &             \\
                                              &      & Tag (16 bits)        &             \\
@@
     \hline
-        \multirow{3}{2.2cm}{~~Pattern Table} &      & $C_{sig}$ (4bits)      &               \\
+        \multirow{3}{2cm}{~~Pattern Table} &      & $C_{sig}$ (4bits)      &               \\
                                              & 512  & $C_{delta}$ (4*4 bits) & 24576 bits    \\
@@
     \label{tab:PPF_overhead}
+\end{adjustwidth}
 \end{table}
@@
 \footnotetext[1]{Components of Prefetch Table can be found in Table \ref{tab:PTable_metadata}}
-\footnotetext[2]{the Reject Table does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
+\footnotetext[2]{The Reject Table does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
 
-In this section, we analyze the hardware overhead required to
-implement PPF.  The Prefetch Table was enhanced to accommodate
-storing of meta-data for perceptron training.  Table
-\ref{tab:PTable_metadata} depicts the meta-data stored for each entry in
-the Prefetch Table.  Table \ref{tab:PPF_overhead} shows the total
-storage overhead of PPF implementation.  The hardware budget for
-2nd Data Prefetching championship was 32 KB.  Keeping that in mind 
-the considerable speedup PPF obtained over the winner, the extra hardware
-budget can be accounted for.  The extra hardware also makes the
-overall scheme more scalable than SPP.  In the original SPP paper, it
-was demonstrated that adding extra hardware brings little advantage in
-terms of performance gain.  The newly added perceptron tables can be
-scaled to increase / decrease features depending on the permitted
-budget.
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{SPEC2017}
+\caption{SPEC 2017 Single-Core IPC Speedup}
+\label{Fig:SPEC2017_1core}
+\end{figure}
 

### 2018-12-01 results.tex
Separating arch and implementation. Filter-Table and Throttler-Filter renaming
Updated results section & added graphs

@@
 
-Figure \_\_ shows the IPC speedup on the memory intensive subset of SPEC 2017. 
+Figure~\ref{Fig:SPEC2017_1core} shows the IPC speedup obtained by
+PPF, compred to BOP, DA-AMPM and SPP. 
+In the figure we have depicted selected traces, where PPF shows a 
+noticeable improvement or shows some degradation, along with the 
+geometric mean improvement on the memory intensive subset of SPEC 2017, 
+and finally for the complete SPEC 2017 benchmark. 
 All the results have been normalized to baseline \textit{i.e.}, no
-prefetching.  As can be seen, PPF yields a speedup of XX\% over
-the baseline.  This is equivalent to XX\% over DA-AMPM, xx\% over BOP
-and XX\% over SPP.  Out of the XX benchmarks, PPF nearly matches
-or outperforms most of the prefetchers on XX traces.
+prefetching.
+
+As can be seen, PPF yields a speedup of \textbf{42.8\%} over
+the baseline. This is equivalent to \textbf{7.8\%} over DA-AMPM, \textbf{9.7\%} over BOP
+and \textbf{6.92\%} over SPP.  Out of the 95 simpoints developed for SPEC 2017, 
+PPF nearly matches or outperforms most of the prefetchers on 91 simpoints.
+
+Only in 607.cactuBSSN and 654.roms, PPF fails to match the improvement 
+offered by BOP.
 
 At its peak, PPF manages to get the IPC speedup of a factor of
-over \textbf{XX} on the trace \textit{602.gcc\_s-2226B}.  This also
-corresponds to speedup gain of \textbf{XX\%} over the next best
-prefetcher - SPP.  In general, benchmarks \textit{602.gcc\_s,
-  603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s}
-benefit the most from PPF, as compared to SPP only prefetching.
+over \textbf{3.39x} on the trace \textit{602.gcc\_s-2226B}.  This also
+corresponds to speedup gain of \textbf{60\%} over the next best
+prefetcher -- BOP. 
+In general, benchmarks \textit{603.bwaves\_s, 605.mcf\_s, 623.xalancbmk\_s} 
+and \textit{649.fotonik3d\_s} benefit the most from PPF, with the speedup over
+SPP ranging from \textbf{10\% to 25\%}.
 
-On the full SPEC 2017 suite, PPF improves the baseline by XX\%,
-which is YY\% more than just SPP.
+On the full SPEC 2017 suite, PPF improves the baseline by \textbf{20.9\%},
+which is \textbf{3.44\%} more than the next best prefetcher -- SPP.
 
@@
 Improvement over SPP is as high as 31\% and even reduces to -1.5\% on
-traces of this benchmark>
+traces of this benchmark>\newline
 
-<Discuss about coverage on all traces>
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{Coverage}
+\caption{Fraction of Cache Misses Covered}
+\label{Fig:Coverage}
+\end{figure}
+
+\textbf{Coverage}\newline
+A prefetcher coverage can be defined as the ratio of number of misses 
+avoided through prefetching to the number of misses in presence of no 
+prefetchig.
+Figure~\ref{Fig:Coverage} shows the fraction of misses in L2 and Last
+Level Caches avoided by the various prefetchers.
+PPF has the highest coverage of all the prefetchers simulated. Using PPF 
+on SPEC 2017 benchmark, 
+misses can be reduced by \textbf{62.5\%} and \textbf{82.8\%} in L2 Cache and Last Level 
+Cache respectively. For the same benchmark, the next best prefetcher -- 
+DA-AMPM covers \textbf{48.6\%} and \textbf{72.8\%} of the misses respectively.
+
+This superior coverage of PPF can be attributed to aggressive 
+re-tuning of the underlying SPP. At the same time, the Perceptron Filter
+makes sure than this high coverage does not lead to an increased 
+cache pollution.
 
 <Discuss cases where coverage gets specially increased>
-
 <Discuss accuracy on all traces>
 
-<Discuss about compelling increase in accuracy - details of
-traces. Plot linear separability>
-
-<Any case where accuracy got reduced - go into the trace. See if
-linear separability is the issue. Other possible issues?>
+\begin{figure}[h]
+%%\includegraphics[width=\columnwidth]{4Core_SPEC2017}
+\caption{Normalized speedup for 2 sets of 4-core workloads}
+\label{Fig:4Core_SPEC2017}
+\end{figure}
 
@@
 
-\textit{4-core environment} Figure \_\_ shows a comparison of speedups
+\textit{4-core environment} Figure~\ref{Fig:4Core_SPEC2017} (a) 
+shows a comparison of speedups
 obtained on 4-core mix of memory intensive subset of SPEC 2017.  We
@@
 As can be seen,
-PPF offers an IPC improvement by 51.7\% on these traces.  This is
-11.9\% over the next best prefetcher SPP; and XX\% and YY\% over BOP
-and DA-AMPM respectively.
+PPF offers an IPC improvement by \textbf{51.7\%} on these traces.  This is
+\textbf{XX\%} over the next best prefetcher -- \_\_; and \textbf{XX\%} 
+and \textbf{YY\%} over \_\_ and \_\_ respectively.
 
-On a different set of fully random SPEC 2017 4-core mixes, PPF provides an IPC speedup of 
-26.2\% over baseline, which is 5.7\% more than just SPP.
+On a different set of fully random SPEC 2017 4-core mixes, 
+PPF provides an IPC speedup of 
+\textbf{26.2}\% over baseline, which is \textbf{5.7}\% more than just SPP.
+
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{8Core_SPEC2017}
+\caption{Normalized speedup for 2 sets of 8-core workloads}
+\label{Fig:8Core_SPEC2017}
+\end{figure}
 
 \textit{8-core environment} The sorted comparison of speedup obtained 
-on the memory intensive 8-core mixes, is depicted in Fig \_\_.
-PPF improves baseline performance by 38.6\% and that is 10.7\% 
+on the memory intensive 8-core mixes is depicted in Fig~\ref{Fig:8Core_SPEC2017}.
+PPF improves baseline performance by \textbf{38.6\%} and that is \textbf{10.7\%} 
 more than using just SPP. 
+For completely random set of SPEC 2017 mixes, PPF improves performance by 
+\textbf{23.6\%} over the baseline, corresponding to \textbf{4.8\%} over SPP.
+This increased improvement achieved by PPF over the base engine SPP,
+in a multi-core environment is not surprising. 
+Since PPF has an extremely intelligent filter, it ends up 
+eliminating useless prefetches before they can create cache
+pollution in the shared LLC.
 
-For completely random SPEC 2017 mixes, PPF improves performance by 
-23.6\% over the baseline, corresponding to 4.8\% over SPP.
+As can be seen, BOP offer a better improvement than 
+SPP, for the memory intemsive mix, This can be attributed 
+to BOP's inherent aggressive nature. DA-AMPM is also ahead of
+SPP in both the mixes. Interestingly, in all these cases, 
+PPF consistently outperforms the best performing prefetcher.
 
-This increased improvement by PPF on a multi-core machines is not
-surprising.  Since PPF has an extremely intelligent throttler, it ends up 
-eliminating useless prefetches before they can create cache
-pollution. This further amplifies the edge that SPP had over BOP in multi-core environment.
-
-\textit{<See if we can use
-speed-up over BOP as a measure here for showing scalability>}
+%[EB] How to justify only a small margin over DA-AMPM?
 
@@
 
+\begin{figure}[ht]
+\includegraphics[width=\columnwidth]{SPEC2006}
+\caption{SPEC 2006 Single-Core IPC Speedup}
+\label{Fig:SPEC2006_1core}
+\end{figure}
+
 \subsection{Cross Validation}
 \label{Results-CrossVal}
+
 Here we try to demonstrate the robustness of PPF by testing it on
@@
 
-Fig \_\_ shows the speed-up achieved by BOP, SPP and PPF on
+Fig~\ref{Fig:SPEC2006_1core} shows the speed-up achieved by BOP, SPP and PPF on
 selected traces of SPEC 2006, along with the overall geometric mean
@@
 SPEC 2006 benchmark.  As can be seen, PPF provides a speed-up of
-44.4\% over baseline, on the memory intensive subset of SPEC 2006
-benchmark.  This corresponds to 6.84\% over SPP and XX\% over BOP.  On
-the whole of the SPEC 2006 suite, the speedup obtained is 22.4\%,
-which is 3.34\% ahead of SPP.
+46.1\% over baseline, on the memory intensive subset of SPEC 2006
+benchmark. This corresponds to \textbf{6.66\%} over SPP and {9.86\%} over BOP and DA-AMPM.  
+On the whole of the SPEC 2006 suite, the speedup obtained is \textbf{22.4\%},
+which is \text{3.34\%} ahead of SPP.
 
-For 4-core memory intensive mixes, PPF improves the baseline by 59.2\%, 
-8.7\% ahead of SPP. For 8-core memory intensive mixes, the IPC speedup 
-over baseline is 47.9\%, 11.5\% ahead of SPP.
+For 4-core memory intensive mixes, PPF improves the baseline by \textbf{59.2\%}, 
+\textbf{8.7\%} ahead of SPP. For 8-core memory intensive mixes, the IPC speedup 
+over baseline is \textbf{47.9\%}, \textbf{11.5\%} ahead of SPP.
 
@@
 general, perceptron weights are able to adjust in real-time so as to
-find the best possible correlation between the output and the
+find the best possible correlation between the output and the given set of 
 features.

### 2018-12-01 intro.tex
Including introduction illustration

@@
 
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{Motivation}
+\caption{Variation of IPC, Total Prefetches and Useful Prefetches}
+\label{Fig:Motivation}
+\end{figure}
+
 Coverage and accuracy are generally at odds with one another, and as one metric improves, the other
@@
 
-A delicate balance between coverage and accuracy is required for a prefetcher to maximize
+Figure~\ref{Fig:Motivation} illustrates the above scenario in a working manner.
+Here we consider a state-of-the-art lookahead prefetcher -- SPP. We keep on 
+re-tuning its thresholds, each time getting a variant with an increasing lookahead 
+depth coverage. The figure depicts the behavior of the prefetcher variants running 
+on the trace \textit{603.bwaves\_s-1740B}. The IPC, total number of prefetches 
+suggested by the prefetcher (TOTAL\_PF) and the actual useful predictions (GOOD\_PF),
+all have been normalized to the least aggressive version of the prefetcher. 
+As the look-ahead depth of the prefetch suggestions increase, so do useful prefetches. 
+But that comes at a cost of total prefetches inceasing at an even higher rate. This leads 
+to cache pollution and bandwidth contention, as is clearly visible with the reduction in 
+the IPC. Similar behaviour was observed in 12 other SPEC 2017 simpoints.
+
+Therefore, a delicate balance between coverage and accuracy is required for a prefetcher to maximize
 its performance impact.  Prefetchers are generally designed with internal mechanisms to

### 2018-12-02 design_arch.tex
Updated results section & added graphs
added more general intro

@@
-\section{PPF Architecture}
+\section{PPF Design and Architecture}
 \label{Arch}
+% Really need a high-level overview of the technique before you start
+% talking about tables!
 
-In this section, we talk about the the general structure of the 
-Perceptron-Based Prefetch Filtering (PPF) for enhancing an underlying prefetcher.  
-PPF is a generalised prefetch filtering mechanism and can be adapted to 
-most prefetchers with minimal modifications. %Gino: Are these modifications minimal?
+As previously discussed, ultimately it might be profitable to allow a
+given prefetcher to speculate as deeply as possible, since often some
+useful prefetches are still being generated long after the confidence
+of the prefetcher has fallen below the point at which performance
+improvements are seen, due to the increasing rate of inaccurate
+prefetches.  In order to allow this deep speculation in the
+prefetcher, however, the increasing numbers of bad prefetches must be
+filtered out.  To this end, we propose to leverage perceptron-based
+machine learning as a mechanism to differentiate between potentially
+useful deeply speculated prefetches and likely not-useful ones.  This
+Perceptron Prefetch Filter (PPF) is placed between the prefetcher and
+the prefetch insertion queue, to ensure that predicted bad prefetches
+are not inserted into the memory system, \pg{As shown in figure
+  whatever...}.
+
+As discussed in Section~\ref{sec:Background-Perceptron}, perceptrons
+are useful, light-weight mechanisms to pull together disparate forms
+of information and synthesize a decision from them.  In this case, we
+will leverage a variety of different relative characteristics of a
+given prefetch, such as speculation depth, page address and offset,
+into a binary useful/non-useful prediction.  In the remainder of this
+section, we discus the the general structure of PPF for enhancing a
+given underlying prefetcher.  We will show that PPF is a generalized
+prefetch filtering mechanism which may adapted to any prefetcher with
+minimal modifications. %Gino: Are these modifications minimal?
 
@@
 \label{Arch-Perceptron}
+
 The Perceptron Filter is organized as tables of perceptron weights,
-with a separate table of weights for each feature.  
-A different number of bits are used for each feature to index into its 
-corresponding table.
+with a separate table of weights for each feature.  A different number
+of bits are used for each feature to index into its corresponding
+table.
 %Gino: repharse below?
 Each feature's value is used to index into the corresponding table.
-This indexing is done using a different number of bits for each feature.  
-At most, 12 bits of a feature are used to index into the
-weights table. Certain features require more resolution
-\textit{i.e.}, the full 12 bits of indexing.  On the other hand, some
-features requiring lower resolution require as few as 7 bits of indexing. 
-The variable indexing was determined by studying the
-features and fine-tuned empirically to achieve a good accuracy
-vs hardware overhead trade-off.  Exact details are mentioned under ``Area Overheads''
-in Section~\ref{Method-Overheads}
+This indexing is done using a different number of bits for each
+feature.  At most, 12 bits of a feature are used to index into the
+weights table. Certain features require more resolution \textit{i.e.},
+the full 12 bits of indexing.  On the other hand, some features
+requiring lower resolution require as few as 7 bits of indexing.  The
+variable indexing was determined by studying the features and
+fine-tuned empirically to achieve a good accuracy vs hardware overhead
+trade-off.  Exact details are mentioned under ``Area Overheads'' in
+Section~\ref{Method-Overheads}
 
 A single entry in the table corresponds to a perceptron weight.  Each
-weight is a 5 bit counter - saturating at -16 and +15 and initialized to zero at the 
-beginning of a program's execution. Our proposed design uses nine features, 
-represented in nine perceptron weight tables.
+weight is a 5 bit counter - saturating at -16 and +15 and initialized
+to zero at the beginning of a program's execution. Our proposed design
+uses nine features, represented in nine perceptron weight tables.
 
-\textbf{Inference}\newline In the ChampSim simulator, the prefetching method is 
-invoked on every L2 Cache demand access.  At that point, the base prefetcher
-has the option to perform a prefetch -- if it does, then it has a choice
-of how many cache lines to prefetch.  The suggested prefetches can be either placed
-in the L2 cache or the last-level cache based on the prefetcher's internal confidence mechanism.
+\textbf{Inference}\newline In the ChampSim simulator, the prefetching
+method is invoked on every L2 Cache demand access.  At that point, the
+base prefetcher has the option to perform a prefetch -- if it does,
+then it has a choice of how many cache lines to prefetch.  The
+suggested prefetches can be either placed in the L2 cache or the
+last-level cache based on the prefetcher's internal confidence
+mechanism.
 
-When the base prefetcher is triggered, it begins to suggest candidates for
-prefetching. All suggested prefetches are passed through the perceptron
-filter to decide if they are qualified for final prefetching. 
-he perceptron decides whether to prefetch a candidate by looking at the microarchitectural state 
-\em} i.e. the features, at that instant. Each feature is hashed to form an index into a table of up to 
-4096 entries dedicated for that feature.
+When the base prefetcher is triggered, it begins to suggest candidates
+for prefetching. All suggested prefetches are passed through the
+perceptron filter to decide if they are qualified for final
+prefetching.  The perceptron decides whether to prefetch a candidate
+by looking at the microarchitectural state \textit{i.e.} the features,
+at that instant. Each feature is hashed to form an index into a table
+of up to 4096 entries dedicated for that feature.
 
-Once all the weights are retrieved, they are summed and compared to a preset threshold 
-,PERC\_ THRESHOLD\_LO.  Only the prefetch candidates with perceptron sum higher than 
-the threshold qualify for prefetching.  The prefetches that qualify through the perceptron 
-stage are recorded in the ``Prefetch Table". The prefetch table is a 1024-entry, direct mapped structure taht
-conatains all metadata required to recreate the features. 	%Gino: Recreate the features?
-When the feedback of the current prefetched line is available at a later stage, the
-stored data is used to train the perceptron.
+Once all the weights are retrieved, they are summed and compared to a
+preset threshold ,PERC\_ THRESHOLD\_LO.  Only the prefetch candidates
+with perceptron sum higher than the threshold qualify for prefetching.
+The prefetches that qualify through the perceptron stage are recorded
+in the ``Prefetch Table". The prefetch table is a 1024-entry, direct
+mapped structure taht conatains all metadata required to recreate the
+features.  %Gino: Recreate the features?
+When the feedback of the current prefetched line is available at a
+later stage, the stored data is used to train the perceptron.
 
-The sum of the perceptron weights are also used to decide whether to place teh 
-data in the L2 or the last level cache. All prefetch candidates that are qualified until
-this stage are compared against PERC\_THRESHOLD\_HI to decide the fill
-level. The two thresholds: PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are
-empirically set.
+The sum of the perceptron weights are also used to decide whether to
+place teh data in the L2 or the last level cache. All prefetch
+candidates that are qualified until this stage are compared against
+PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
+PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
 
-The perceptron sum can be considered the sum of the individual contribution
-of a feature.  The value of each contribution corresponds to a feature's
-confidence in the final decision on whether to prefetch a cache line. By summing the
-individual contributions, the final perceptron sum denotes the overall
-confidence for that prefetch suggestion.  By comparing the perceptron sum
-to two different thresholds, we divide the confidence scale of the
-prefetch suggestion into three bins. The first bin corresponds to the lowest
-confidence and leads to prefetch candidate being rejected.  The next bin
-corresponds to prefetches with a moderate confidence level.  This bin prefetches
- directeky to fill the bigger last level cache and prevents polluting
-the smaller L2 Cache.  The highest confidence prefetches are filled in the
-L2 cache.
+The perceptron sum can be considered the sum of the individual
+contribution of a feature.  The value of each contribution corresponds
+to a feature's confidence in the final decision on whether to prefetch
+a cache line. By summing the individual contributions, the final
+perceptron sum denotes the overall confidence for that prefetch
+suggestion.  By comparing the perceptron sum to two different
+thresholds, we divide the confidence scale of the prefetch suggestion
+into three bins. The first bin corresponds to the lowest confidence
+and leads to prefetch candidate being rejected.  The next bin
+corresponds to prefetches with a moderate confidence level.  This bin
+prefetches directeky to fill the bigger last level cache and prevents
+polluting the smaller L2 Cache.  The highest confidence prefetches are
+filled in the L2 cache.
 
-In addition to the prefetch table mentioned above, PPF also maintains a
-``Reject Table.''  The reject table is a 1024-entry deep direct-mapped
-table.  If a prefetch suggestion is rejected by the perceptron layer, it is
-logged into the reject table.  The filer is used to train the perceptron to %Gino: Filer?
-avoid false negatives \textit{i.e.}, cases where prediction was to reject the
-prefetch but the prefetch would have been useful.
+In addition to the prefetch table mentioned above, PPF also maintains
+a ``Reject Table.''  The reject table is a 1024-entry deep
+direct-mapped table.  If a prefetch suggestion is rejected by the
+perceptron layer, it is logged into the reject table.  The filer is
+used to train the perceptron to %Gino: Filer?
+avoid false negatives \textit{i.e.}, cases where prediction was to
+reject the prefetch but the prefetch would have been useful.
 
@@
 
-\textbf{Training}\newline In the prefetching environment, feedback for a
-prefetch is received whenever there is an eviction or a demand access from the
-L2 Cache.  This action triggers training of the base prefetcher as well as the
-perceptron layer.  Training the perceptrons involves accessing
-the metadata that was stored in Prefetch Table or the Reject Table. 
-The cache lines address of the block being trained is used to access a table,
-using 10-bits of the address for indexing and another 6-bits to perform tag matching.
-Once the state of the program at the time of prefetching is available, it is 
+\textbf{Training}\newline In the prefetching environment, feedback for
+a prefetch is received whenever there is an eviction or a demand
+access from the L2 Cache.  This action triggers training of the base
+prefetcher as well as the perceptron layer.  Training the perceptrons
+involves accessing the metadata that was stored in Prefetch Table or
+the Reject Table.  The cache lines address of the block being trained
+is used to access a table, using 10-bits of the address for indexing
+and another 6-bits to perform tag matching.  Once the state of the
+program at the time of prefetching is available, it is
 %Gino: By "state of the program" are you referring to the state of the features?
@@
 If the demanded block that triggers the training was tagged as a valid
-prefetch in the prefetch table, then the earlier prefetch prediction was
-correct.  In that case the perceptron weights are incremented by 1
-if the predicted sum does not cross a pre-defined threshold. These training 
-thresholds are introduced to avoid overfitting of the perceptron weights to the given program 
-behaviour. These thresholds are referred to as POS\_UPDATE\_THRESHOLD ($\theta_p$) and 
+prefetch in the prefetch table, then the earlier prefetch prediction
+was correct.  In that case the perceptron weights are incremented by 1
+if the predicted sum does not cross a pre-defined threshold. These
+training thresholds are introduced to avoid overfitting of the
+perceptron weights to the given program behaviour. These thresholds
+are referred to as POS\_UPDATE\_THRESHOLD ($\theta_p$) and
 NEG\_UPDATE\_THRESHOLD ($\theta_n$), respectively for the positive and
@@
 
-If a cache block eviction led to training and the corresponding valid entry is found in the
-prefetch table without the block being accessed, then the prediction made by the perceptron was wrong.
-The perceptron should have ideally rejected the prefetch suggestion as a
-low-confidence prefetch.  Here the weights are decremented by 1 to reflect the
-misprediction. In either case, weights are saturated at -16 or +15. %Gino: Describe how these limits were decided?
+If a cache block eviction led to training and the corresponding valid
+entry is found in the prefetch table without the block being accessed,
+then the prediction made by the perceptron was wrong.  The perceptron
+should have ideally rejected the prefetch suggestion as a
+low-confidence prefetch.  Here the weights are decremented by 1 to
+reflect the misprediction. In either case, weights are saturated at
+-16 or +15. %Gino: Describe how these limits were decided?
 
-A secondary training mechanism also kicks in during demand fetches.  Before
-the demand access triggers the next set of prefetches, the reject table is
-checked for a valid entry.  A hit means that the corresponding cache line was
-initially suggested by the underlying prefetch engine, but rejected by the
-perceptron filter.  Thus, the perceptron should have been more confident
-about that particular prefetch.  Once such a scenario is identified, the state %Gino: state of execution == value of features?
-of the execution at the time of prefetch is retrieved from the reject table.
-The retrieved data is used to index into the various weights tables of the
-perceptron and the corresponding values are updated by +1, saturating between
-and -16 and +15.  This update reflects increased confidence for the prediction
-corresponding to that prefetch.
+A secondary training mechanism also kicks in during demand fetches.
+Before the demand access triggers the next set of prefetches, the
+reject table is checked for a valid entry.  A hit means that the
+corresponding cache line was initially suggested by the underlying
+prefetch engine, but rejected by the perceptron filter.  Thus, the
+perceptron should have been more confident about that particular
+prefetch.  Once such a scenario is identified, the
+state %Gino: state of execution == value of features?
+of the execution at the time of prefetch is retrieved from the reject
+table.  The retrieved data is used to index into the various weights
+tables of the perceptron and the corresponding values are updated by
++1, saturating between and -16 and +15.  This update reflects
+increased confidence for the prediction corresponding to that
+prefetch.
 
-This mechanism allows us to exploit a previously lost opportunity.  
-In prior perceptron based implementations, and in general,
-prefetching algorithms, there is usually no way of knowing the result of not
-prefetching a particular line.  Our two-step PPF architecture allows us to
-overcome that issue.
+This mechanism allows us to exploit a previously lost opportunity.  In
+prior perceptron based implementations, and in general, prefetching
+algorithms, there is usually no way of knowing the result of not
+prefetching a particular line.  Our two-step PPF architecture allows
+us to overcome that issue.
 
@@
 The above discussion of PPF shows that it is highly modular and can be
-adapted to be used over any base prefecher for increased prefetch accuracy.
-In general, only two hooks are required between PPF and the
-baseline prefetcher. The first is to make sure that all the prefetch candidates of
-the prefetcher pass through the perceptron filter and if qualified, the metadata for
-perceptron indexing be stored. The second is needed when the feedback of a
-prior prefetch is available in form of a subsequent demand hit or cache
-eviction. In that case, the stored metadata needs to be retrieved to update
-the state of the perceptrons.
+adapted to be used over any base prefecher for increased prefetch
+accuracy.  In general, only two hooks are required between PPF and the
+baseline prefetcher. The first is to make sure that all the prefetch
+candidates of the prefetcher pass through the perceptron filter and if
+qualified, the metadata for perceptron indexing be stored. The second
+is needed when the feedback of a prior prefetch is available in form
+of a subsequent demand hit or cache eviction. In that case, the stored
+metadata needs to be retrieved to update the state of the perceptrons.
 
@@
 
-\item \textbf{Enhancing the Base Prefetcher:} By tuning down any internal thresholds 
-to increase its inherent aggressiveness.
+\item \textbf{Enhancing the Base Prefetcher:} By tuning down any
+  internal thresholds to increase its inherent aggressiveness.
 
-\item \textbf{Inferencing and Storing:} All prefetch recommendations are
-tested using the perceptron inferencing algorithm. The perceptron's output,
-\textit{true} or \textit{false}, should be saved appropriately, along with all
- metadata required for perceptron indexing.
+\item \textbf{Inferencing and Storing:} All prefetch recommendations
+  are tested using the perceptron inferencing algorithm. The
+  perceptron's output, \textit{true} or \textit{false}, should be
+  saved appropriately, along with all metadata required for perceptron
+  indexing.
 
-\item \textbf{Retrieving and Training:} When feedback for a prefetch is available, the
-previously stored metadata can be used to re-index into the perceptron entries
-and increment or decrement the weights.
+\item \textbf{Retrieving and Training:} When feedback for a prefetch
+  is available, the previously stored metadata can be used to re-index
+  into the perceptron entries and increment or decrement the weights.
 
-\item \textbf{Feature Selection:} In Section~\ref{Impl-Features}, we show the features 
-used for PPF. Six out of the nine features we developed
-use information derived directly from program execution, irrespective of the baseline 
-prefetcher. Beyond that, the feature set can be expanded to convey any
-useful information from the prefetcher to the perceptron filter.  The 
-methodology explained in Section~\ref{Method-Features} talks about developing a minimal 
+\item \textbf{Feature Selection:} In Section~\ref{Impl-Features}, we
+  show the features used for PPF. Six out of the nine features we
+  developed use information derived directly from program execution,
+  irrespective of the baseline prefetcher. Beyond that, the feature
+  set can be expanded to convey any useful information from the
+  prefetcher to the perceptron filter.  The methodology explained in
+  Section~\ref{Method-Features} talks about developing a minimal
   feature set for PPF.

### 2018-12-02 design_impl.tex
some comments

@@
 This section describes our implementation of PPF and the range of
-features that are used by the perceptron for correlating 
-the prefetches.  For a practical implementation, we have used SPP 
-as the base engine prefetcher.
+features that are used by the perceptron for correlating the
+prefetches.  For a practical implementation, we have used SPP as the
+base engine prefetcher.
 
@@
 \label{Impl-Changes}
-To modify the SPP design to suit our scheme, the following changes were made:
+To modify the SPP design to suit our scheme, the following changes
+were made:
 
 \begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline 
-  In PPF, the perceptron sum is used to decide whether to prefetch or not, 
-  and the fill-level in case of prefetch. 
-  Thus, the confidence thresholds --  T\textsubscript{f} and T\textsubscript{p} 
-  are no longer needed.
+\item \textbf{Original Thresholds discarded}\newline In PPF, the
+  perceptron sum is used to decide whether to prefetch or not, and the
+  fill-level in case of prefetch.  Thus, the confidence thresholds --
+  T\textsubscript{f} and T\textsubscript{p} are no longer needed.
 
@@
 
-\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The original
-  SPP does not consider the slots available in the L2 MSHR queue.  It has the
-  assumption that at no time can the number of suggested prefetches
-  exceed the capacity of the queue.  While it was not described in the paper, the
-  internal confidence mechanism makes sure that the above assumption was maintained.  
-  In PPF, since there cannot be an assumption on the degree of aggressiveness of the 
-  underlying prefetcher, we explicitly check that at no point should the number 
-  of suggested prefetches exceed the L2 MSHR queue.
+\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The
+  original SPP does not consider the slots available in the L2 MSHR
+  queue.  It has the assumption that at no time can the number of
+  suggested prefetches exceed the capacity of the queue.  While it was
+  not described in the paper, the internal confidence mechanism makes
+  sure that the above assumption was maintained.  In PPF, since there
+  cannot be an assumption on the degree of aggressiveness of the
+  underlying prefetcher, we explicitly check that at no point should
+  the number of suggested prefetches exceed the L2 MSHR queue. \pg{is
+    this a big deal?  Seems like a detail}
 
-\item \textbf{Enhanced Prefetch Table}\newline The prefetch table introduced in the 
-  original SPP keeps track of up to 1024 L2 cache prefetch suggestions.  
-  For PPF, the prefetch table is modified to store metadata such as program counter, 
-  memory access address etc.  This metadata is needed to retrieve the state  %Gino: What's the etc.?
-  of the features to train the perceptron at a later stage in time. The prefetch table 
-  mentioned above keeps track of the prefetches that actually happened in the L2 Cache.
-  We introduce the reject table to keep a track of the
-  prefetches that were suggested by the base SPP engine but were rejected by PPF.  
-  In case it is detected that a future demand fetch could have used
-  this rejected prefetch, the perceptron weights are updated to reflect that
+\item \textbf{Enhanced Prefetch Table}\newline The prefetch table
+  introduced in the original SPP keeps track of up to 1024 L2 cache
+  prefetch suggestions.  For PPF, the prefetch table is modified to
+  store metadata such as program counter, memory access address etc.
+  This metadata is needed to retrieve the
+  state %Gino: What's the etc.?
+  of the features to train the perceptron at a later stage in
+  time. The prefetch table mentioned above keeps track of the
+  prefetches that actually happened in the L2 Cache.  We introduce the
+  reject table to keep a track of the prefetches that were suggested
+  by the base SPP engine but were rejected by PPF.  In case it is
+  detected that a future demand fetch could have used this rejected
+  prefetch, the perceptron weights are updated to reflect that
   misprediction.
@@
 \label{Impl-Features}
-Here we discuss the various features that correlate the prefetching decision
-with the program behaviour.
+Here we discuss the various features that correlate the prefetching
+decision with the program behavior. \pg{again you want to get at why
+  these were chosen.  Also mention that you searched over a large
+  space of possible features and these were the best.}
 
@@
 
-\item \textbf{Program Counter XOR Depth}: The PC is for the instruction
-  that triggered the prefetch chain.  Depth refers to the iteration count of
-  the lookahead stages.  In general, the PC is not considered as a good basis
-  for doing lookahead prefetching as all the prefetches with depth >= 1 are
-  aliased into the same PC which would not be true in an actual demand access.
-  This feature resolves a PC into a different value for each lookahead depth
-  of prefetch speculation, giving a more accurate correlation in lookahead
-  cases.  This is akin to the concept of Virtual Program Counters~\cite{VPC}
+\item \textbf{Program Counter XOR Depth}: The PC is for the
+  instruction that triggered the prefetch chain.  Depth refers to the
+  iteration count of the lookahead stages.  In general, the PC is not
+  considered as a good basis for doing lookahead prefetching as all
+  the prefetches with depth >= 1 are aliased into the same PC which
+  would not be true in an actual demand access.  This feature resolves
+  a PC into a different value for each lookahead depth of prefetch
+  speculation, giving a more accurate correlation in lookahead cases.
+  This is akin to the concept of Virtual Program Counters~\cite{VPC}
   introduced by Kim \textit{et. al.} for indirect branch prediction.
@@
   PC before the instruction that triggered the current prefetch.
-  Hashing together the last three PCs tell PPF about the path
-  that led to the current demand access and helps capture and
-  branching information of the current basic block.  PCs are shifted
-  in the increasing order of history before being hashed together.
-  This is done to avoid the resultant value of zero when 2 or more
-  PCs are the same.  Additionally, obfuscating the information as it
-  gets old allows us to get a wider and approximate look into the
-  program's history.
+  Hashing together the last three PCs tell PPF about the path that led
+  to the current demand access and helps capture and branching
+  information of the current basic block.  PCs are shifted in the
+  increasing order of history before being hashed together.  This is
+  done to avoid the resultant value of zero when 2 or more PCs are the
+  same.  Additionally, obfuscating the information as it gets old
+  allows us to get a wider and approximate look into the program's
+  history.
 
-\item \textbf{Program Counter XOR Delta}: This feature tells us if a given PC
-  favours particular value(s) of delta.  As noted earlier, while the PC in
-  itself does not convey much useful information, this hash resolves the PC
-  into different values based on the tendency of that PC to favour a certain
-  delta.  Thus, the dynamic nature of different instances of the same memory
-  access instruction can be captured here.
+\item \textbf{Program Counter XOR Delta}: This feature tells us if a
+  given PC favours particular value(s) of delta.  As noted earlier,
+  while the PC in itself does not convey much useful information, this
+  hash resolves the PC into different values based on the tendency of
+  that PC to favour a certain delta.  Thus, the dynamic nature of
+  different instances of the same memory access instruction can be
+  captured here.
 
 \item \textbf{Confidence}: The integer confidence on a scale of 0 to
-  100 that was used in the original SPP design. While the original confidence
-  is not used directly for decision making, it can still contribute to the
-  final decision made.
+  100 that was used in the original SPP design. While the original
+  confidence is not used directly for decision making, it can still
+  contribute to the final decision made.
 
-\item \textbf{Page Address XOR Confidence}: This feature scores the tendency
-  of each page to be prefetch friendly or prefetch averse. It helps resolve a
-  page into different entries depending on its confidence for prefetching, which
-  can vary during phases of a program execution.
+\item \textbf{Page Address XOR Confidence}: This feature scores the
+  tendency of each page to be prefetch friendly or prefetch averse. It
+  helps resolve a page into different entries depending on its
+  confidence for prefetching, which can vary during phases of a
+  program execution.
 
 % djimenez: this is a little vague. can you be more precise?
-\item \textbf{Current Signature XOR Delta}: Recall from
-  the discussion of SPP in Section \ref{Background-SPP} that the new signature
-  is generated using the old signature and the current block delta.  
+\item \textbf{Current Signature XOR Delta}: Recall from the discussion
+  of SPP in Section \ref{Background-SPP} that the new signature is
+  generated using the old signature and the current block delta.
 %Gino: Added the sentence below, is this correct?
-  The result of this feature is the next delta that is predicted to be accessed based on the 
-  delta predicted by SPP. While ``Current Signature XOR Delta'' is not the exact formula for generating the future
-  signature, it gives an approximate idea of the path that the combination of
-  these two values can lead to.
+  The result of this feature is the next delta that is predicted to be
+  accessed based on the delta predicted by SPP. While ``Current
+  Signature XOR Delta'' is not the exact formula for generating the
+  future signature, it gives an approximate idea of the path that the
+  combination of these two values can lead to.
 
@@
 
-Some of the composite features are derived from simple hashing (XOR) of two
-primary features.  There is always a question of usefulness of such composite
-features and the new information added.  We justify the choice of each feature
-by quantifying the contribution made towards predicting prefetch behaviour, in
-Section \ref{Method-Features}.  Finally, as noted above, each feature indexes
-into its independent entry of perceptron weights.
+Some of the composite features are derived from simple hashing (XOR)
+of two primary features.  There is always a question of usefulness of
+such composite features and the new information added.  We justify the
+choice of each feature by quantifying the contribution made towards
+predicting prefetch behaviour, in Section \ref{Method-Features}.
+Finally, as noted above, each feature indexes into its independent
+entry of perceptron weights.

### 2018-12-02 design_arch.tex
spelling
some comments

@@
 into a binary useful/non-useful prediction.  In the remainder of this
-section, we discus the the general structure of PPF for enhancing a
+section, we discuss the the general structure of PPF for enhancing a
 given underlying prefetcher.  We will show that PPF is a generalized
@@
 \label{Arch-Perceptron}
+\pg{rewrite this section as talking to the figures and walking through
+  them}
 
@@
 trade-off.  Exact details are mentioned under ``Area Overheads'' in
-Section~\ref{Method-Overheads}
+Section~\ref{Method-Overheads}. \pg{this whole ``bits'' discussion
+  belongs in implementation not here IMHO}
 
@@
 to zero at the beginning of a program's execution. Our proposed design
-uses nine features, represented in nine perceptron weight tables.
+uses nine features, represented in nine perceptron weight
+tables. \pg{the fact that its nine is again implementation details}
 
-\textbf{Inference}\newline In the ChampSim simulator, the prefetching
-method is invoked on every L2 Cache demand access.  At that point, the
-base prefetcher has the option to perform a prefetch -- if it does,
-then it has a choice of how many cache lines to prefetch.  The
-suggested prefetches can be either placed in the L2 cache or the
-last-level cache based on the prefetcher's internal confidence
-mechanism.
+\pg{The word ChampSim should not appear in this part of the paper at
+  all.  Remember you are describing how to build this in the general
+  case...} \textbf{Inference}\newline In the ChampSim simulator, the
+prefetching method is invoked on every L2 Cache demand access.  At
+that point, the base prefetcher has the option to perform a prefetch
+-- if it does, then it has a choice of how many cache lines to
+prefetch.  The suggested prefetches can be either placed in the L2
+cache or the last-level cache based on the prefetcher's internal
+confidence mechanism.
 
@@
 Once all the weights are retrieved, they are summed and compared to a
-preset threshold ,PERC\_ THRESHOLD\_LO.  Only the prefetch candidates
+preset threshold, PERC\_ THRESHOLD\_LO.  Only the prefetch candidates
 with perceptron sum higher than the threshold qualify for prefetching.
 The prefetches that qualify through the perceptron stage are recorded
-in the ``Prefetch Table". The prefetch table is a 1024-entry, direct
-mapped structure taht conatains all metadata required to recreate the
+in the ``Prefetch Table''. The prefetch table is a 1024-entry, direct
+mapped structure that contains all metadata required to recreate the
 features.  %Gino: Recreate the features?
@@
 
-The sum of the perceptron weights are also used to decide whether to
-place teh data in the L2 or the last level cache. All prefetch
-candidates that are qualified until this stage are compared against
-PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
-PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
+\pg{this discussion below has redundancy and doesn't flow well.
+  Rework it}
 
@@
 filled in the L2 cache.
+
+The sum of the perceptron weights are also used to decide whether to
+place the data in the L2 or the last level cache.  All prefetch
+candidates that are qualified until this stage are compared against
+PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
+PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
 
@@
 
-\subsection{Generalizing PPF for any Prefetcher}
+\subsection{Optimizing PPF for a Given Prefetcher}
 \label{Arch-Generalizing}
@@
 
-In general, PPF can be adapted to a new base prefetcher with only a few modifications.
+In general, PPF can be adapted to a new base prefetcher with only a
+few modifications.
 \begin{itemize}
@@
   Section~\ref{Method-Features} talks about developing a minimal
-  feature set for PPF.
+  feature set for PPF. \pg{This should be more about how one should
+    choose features than about what you did in the SPP implementation
+    in particular.  Rewrite as a guide on picking features.}
 %For example, in our implementation ``Confidence,'' ``Current Signature XOR Delta,''

### 2018-12-03 results.tex
Results Section and PDF Images

@@
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{SPEC2017}
+\caption{SPEC 2017 Single-Core IPC Speedup}
+\label{Fig:SPEC2017_1core}
+\end{figure}
+
 \section{Results}
@@
 This section discusses the results obtained from running PPF, in
-terms of cache Misses Per Kilo Instructions (MPKI) and Instruction Per
-Cycle (IPC) metrics.  For SPEC 2017 suite, first we present the
-results for single-threaded workloads then the results for multi-core
-workloads are presented.
+terms of prefetch cache coverage and Instruction Per
+Cycle (IPC).  For SPEC 2017 suite, first we present the
+results for single-threaded workloads then for multi-core
+workloads.
 
@@
 PPF nearly matches or outperforms most of the prefetchers on 91 simpoints.
-
 Only in 607.cactuBSSN and 654.roms, PPF fails to match the improvement 
@@
 prefetcher -- BOP. 
-In general, benchmarks \textit{603.bwaves\_s, 605.mcf\_s, 623.xalancbmk\_s} 
-and \textit{649.fotonik3d\_s} benefit the most from PPF, with the speedup over
+In general, benchmarks \textit{bwaves, mcf, xalancbmk} 
+and \textit{fotonik3d} benefit the most from PPF, with the speedup over
 SPP ranging from \textbf{10\% to 25\%}.
@@
 which is \textbf{3.44\%} more than the next best prefetcher -- SPP.
-
-<Discuss the extreme behaviour of \textit{623.xalancbmk\_s}:
-Improvement over SPP is as high as 31\% and even reduces to -1.5\% on
-traces of this benchmark>\newline
 
@@
 
-<Discuss cases where coverage gets specially increased>
-<Discuss accuracy on all traces>
-
 \begin{figure}[h]
-%%\includegraphics[width=\columnwidth]{4Core_SPEC2017}
-\caption{Normalized speedup for 2 sets of 4-core workloads}
+\includegraphics[width=\columnwidth]{4Core_SPEC2017}
+\caption{Normalized speedup for 4-core workloads}
 \label{Fig:4Core_SPEC2017}
@@
 obtained on 4-core mix of memory intensive subset of SPEC 2017.  We
-plot all the 4 prefetchers, normalized to baseline.  The workloads
-have been sorted in the increasing order of IPC obtained with running PPF.
+plot all the 4 prefetchers, normalized to the baseline.  The workloads
+have been sorted in the increasing order of the speedup.
 As can be seen,
-PPF offers an IPC improvement by \textbf{51.7\%} on these traces.  This is
-\textbf{XX\%} over the next best prefetcher -- \_\_; and \textbf{XX\%} 
-and \textbf{YY\%} over \_\_ and \_\_ respectively.
+PPF offers an IPC improvement by \textbf{51.7\%} on these traces.  
+Compared to the base prefetcher -- SPP only, this is \textbf{11.9\%} more.
+As compared to the next best prefetcher -- DA-AMPM, this is 
+This is \textbf{10.2\%} and \textbf{17.4\%} over DA-AMPM, 
+the next best prefetcher, and BOP, respectively.
 
@@
 \includegraphics[width=\columnwidth]{8Core_SPEC2017}
-\caption{Normalized speedup for 2 sets of 8-core workloads}
+\caption{Normalized speedup for 8-core workloads}
 \label{Fig:8Core_SPEC2017}
@@
 As can be seen, BOP offer a better improvement than 
-SPP, for the memory intemsive mix, This can be attributed 
+SPP, for the memory intensive mix, This can be attributed 
 to BOP's inherent aggressive nature. DA-AMPM is also ahead of
@@
 
+\begin{figure}[ht]
+\includegraphics[width=\columnwidth]{AddnConstr}
+\caption{IPC speedup for Small LLC and Low BW}
+\label{Fig:AddnConstr}
+\end{figure}
+
 \subsection{Additional Memory Constraints}
@@
 
-\textit{Low-Bandwidth DRAM}
-Figure \_\_(a) shows the performance of PPF against other prefetchers. 
-For sake of simplicity, we show 2 traces where PPF gives the most advantage
-over baseline, 2 traces where PPF doesn't perform as well as the next
-best prefetcher (replace with the prefetcher name once results are
-available) and GeoMean across memory intensive and across all the 
-SPEC 2017 traces. As can be seen, for the memory intensive subset,
- PPF improves baseline by XX\% and that is XX\% more than using SPP.
-
-\textit{Small LLC}
-Figure \_\_(b) shows the similar analysis of PPF.
-For the memory intensive subset, a performance improvement of XX\% 
-over baseline is seen with PPF, XX\% more than just SPP.
+Figure \ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a 
+reduced LLC and low bandwidth constraints respectively. For the sake of brevity, 
+we have only shownselected traces, followed
+by the overall performance on the complete benchmark.
+As can be seen, \textit{mcf} in low bandwidth conditions is prefetch averse.
+In general, any prefetcher yields a negative speedup on that trace.
+On \textit{roms} and \textit{cactuBSSN}, PPF is unable to match the performance 
+achieved by the best prefetcher. On the other hand, PPF outperforms all the 
+other prefetchers on \textit{xalancbmk} and \textit{imagick} benchmarks.
+Overall, PPF gives a better improvement under small LLC condition and 
+matches the best prefetcher -- BOP under low DRAM bandwidth condition.
 
@@
 \label{Fig:SPEC2006_1core}
+\end{figure}
+
+\begin{figure}[ht]
+\includegraphics[width=\columnwidth,]{CloudSuite}
+\caption{IPC Speedup for Multi-core CloudSuite Workloads}
+\label{Fig:CloudSuite}
 \end{figure}
@@
 SPEC 2006 benchmark.  As can be seen, PPF provides a speed-up of
-46.1\% over baseline, on the memory intensive subset of SPEC 2006
-benchmark. This corresponds to \textbf{6.66\%} over SPP and {9.86\%} over BOP and DA-AMPM.  
+\textbf{46.1\%} over baseline, on the memory intensive subset of SPEC 2006
+benchmark. This corresponds to \textbf{6.66\%} over SPP and \textbf{9.86\%} over BOP and DA-AMPM.  
 On the whole of the SPEC 2006 suite, the speedup obtained is \textbf{22.4\%},
-which is \text{3.34\%} ahead of SPP.
+which is \textbf{3.34\%} ahead of SPP.
 
@@
 
+Figure~\ref{Fig:CloudSuite} shows the performance benefit comparison of all 
+the prefetch schemes on 4 different applications in the CloudSuite benchmark. 
+In general, these applications are prefetch agnostic. Even in that scenario, 
+PPF manages a \textbf{3.7\%} improvement over no prefetching, making it ahead of the 
+next best prefetcher -- SPP, which provides a 3.08\% IPC speedup.
+
 As can be seen, the performance is very consistent given no prior
-parameter tuning was done keeping SPEC 2006 in mind.  We attribute
+parameter tuning was done keeping these benchmarks in mind.  We attribute
 this to the inherent adaptability of the perceptron model.  In

### 2018-12-03 back.tex
background and related work section

@@
-\section{Background Work and Motivation}
+\section{Motivation and Prior Work}
 \label{Background}
 
-In this section, we compare the existing work done in the domain of
-prefetching.  We also give a brief background for the 
-Signature Path Prefetcher (SPP) architecture.
+In this section we discussed the most closely related work to our proposed technique. Other relevant work will be discussed throughout the paper. See section \ref{related}.
 
-\subsection{Related Work}
-\label{Background-Related}
-The idea of prefetching has been around since Instruction Stream
-Buffers were proposed by Jouppi~\cite{ISB}. Some of the earliest data
-prefetchers developed to try to identify memory access patterns with a
-constant stride pattern~\cite{Smith,Baer,Stride}. Major research was
-dedicated to stride distance and depth prediction~\cite{Decoupled,Adaptive}. Newer prefetchers aimed at correlating
-predictions with the past memory access addresses~\cite{Address_Correlated,AMPM}.
+	The idea of prefetching has been around ever since Jouppi introduced \textit{instruction stream buffers}.
+Early prefetchers detected stride access patterns in order to predict future memory requests ~\cite{Smith,Baer,Stride}.
+These days prefetching mechanisms are more sophisticated, they look into past memory behavior~\cite{Address_Correlated,AMPM}, locality ~\cite{Spatial_Pattern,SMS,Temporal_Instruction_Fetch,Off_Chip,STMS,SMS_JILP}, control-flow speculation ~\cite{BFetch,MTBFetch}, between many other aspects in order to detect complex access patterns in modern applications.
 
-With time, newer and more intricate prefetchers were
-introduced~\cite{Wenisch_Temporal_Streaming,Stealth,Feedback_Directed,Coordinated,Bandwidth_Efficient,Pacman,TLB,Linearizing,Sandbox,VLDP,DoL,Domino}.
-A whole class of streaming prefetchers developed on the lines of Spatial /
-Temporal Memory
-Streaming~\cite{Spatial_Pattern,SMS,Temporal_Instruction_Fetch,Off_Chip,STMS,SMS_JILP}.
-Control-flow speculation directed hardware prefetching has also been
-proposed~\cite{BFetch,MTBFetch}.
+\subsection{Lookahead-based Prefetchers}
 
-Most modern prefetchers aim at identifying complex memory access patterns in a
-given application.  Thus, they capture some of the irregular patterns seen in
-pointer chasing data-structures.  Offset based prefetchers are a
-generalization of the next-line prefetcher.  Some of the prominent prefetchers
-exploiting this idea are Sandbox Prefetcher~\cite{Sandbox} and the Best Offset
-Prefetcher~\cite{BOP}.  Another class of prefetchers is lookahead Prefetchers.
-Such prefetchers try to speculate deep into the application's memory access
-path.  These include Path Confidence based Lookahead Prefetching~\cite{SPP}
-and Kill the Program Counter~\cite{KPC}.
-
-A related line of work also explores cache replacement -- insertion policies
-and dead block prediction, with an eye towards
-prefetching~\cite{DB_Pred,Cache_Burst,KPC,Harmony}.  Nori \textit{et.al.}
-introduced prefetching with respect to runtime criticality~\cite{CATCH}.
-
-\textit{Machine Learning and Prefetching:} Peled \textit{et.al.} introduce
-interesting ideas for on-line Reinforcement Learning and dynamically scaling
-the magnitude of feedback given to the baseline prefetcher~\cite{Semantics}.
-The biggest challenge here is that the idea relies on compiler support for
-getting features to build the context.
-
-Liao \textit{et. al.} focus on prefetching for data center
-applications~\cite{Datacenter}. They use offline machine learning algorithms
-such as SVMs and logistic regression to do a parametric search for an optimal
-prefetcher configuration.
-
-%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to talk about this paper?
-% djimenez: let us not cite this paper. it is in arXiv and we are not
-% obligated to cite it. as far as ISCA is concerned it does not exist. there is
-% a reason it is only in arXiv and not published somewhere: it is crap.
-
-%Finally, the paper on Data Cache Prefetching with Perceptron
-%Learning~\cite{BadPerc} talks about the idea of two step prefetching.
-%The first step is an existing baseline prefetcher while the second
-%step is a perceptron based filter.  The paper has lots of
-%limitations in terms of design and implementation.  That reflects in
-%the fact that it did not lead to any significant performance gain over
-%even the basic prefetchers like the Stride Prefetcher~\cite{Stride} and
-%the Markov Prefetcher~\cite{Markov}.
-
-\subsection{Perceptron Learning in Architecture}
-\label{Background-Perceptron}
-Perceptron learning for computer architecture design has been around for a
-while.  It was first popularized by Jim\'enez \textit{et. al.} for conditional
-branch prediction~\cite{Perc_Branch}. The biggest advantage of the perceptron
-model is that it can adjust its weights dynamically to increase / decrease the
-importance given to each feature - branch histories in this case.  
-Perceptron-based Reuse Prediction~\cite{Perc_Reuse} implemented the hashed
-perceptron architecture to perform prediction in context of cache replacement.
-The structure  we use is derived from the model introduced in that paper, refined with 
-many additional features as in Multiperspective Reuse Prediction{Multiperspective}.  
-The model was originally derived from Hashed Perceptron branch 
-prediction~\cite{HashedPerceptron}.  Here, a hash of the feature itself is used to 
-index into a table of perceptron weights.
-
-% djimenez: this detail is unnecessary and likely to invite questions.
-% reviewers might ask "what about linear separability?" but the proof is in the
-% pudding.
-%This form of indexing makes sure that the hyper-plane learned by the
-%perceptron weights is able to differentiate between linearly
-%inseparable outcomes.
-
-\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron
-prediction involves multiplying the vector of input features:
-F\textsubscript{1xN}, with the corresponding weight vector:
-W\textsubscript{Nx1} in a dot product to obtain the sum y\textsubscript{out}.
-Here what we use is a perceptron-like structure.  The feature is used to hash
-into the weight of perceptrons and the retrieved weights are added straight
-away.  Thus, the perceptron algorithm doesn't introduce any multiplication
-operations in the inference or training process.  Hence what we adapt in this
-work is a perceptron-like learning algorithm as it involves the same principle
-involved in perceptron inferencing and training.
-
-\textit{Perceptron Update Rule:} In all the above implementations, a uniform
-perceptron update principle is followed.  The weights need to be updated if
-the prediction was wrong or the magnituded of the dot product does not exceed
-a certain threshold.  After a certain training period, these weights are
-proportional to the probability of the outcome being true.  This training
-threshold makes sure that the perceptrons are trained until a certain level of
-confidence and yet they are not over-trained to the given set.  Care is taken
-that the weights saturate at the maximum positive and negative values so that
-they remain confined to the bit-width.  The same rules will be applied in the
-discussion done in Section \ref{Arch-Perceptron} on PPF learning algorithm.
-
-%\subsection{Prefetching with SPEC 2017} \label{Background-SPEC2017}
-% djimenez: we might not need to discuss the benchmarks too much as they are
-% standard. we don't have a lot of room anyway.
-
-%\textit{<Discuss SPEC 2017 traces here> <What are the possible graphs /
-%analysis tools we can use for SPEC traces?>}
+An optimal prefetcher should be able to detect a wide range of memory access patterns.
+Simple stride prefetching techniques only detect sequences of addresses that differ by a constant value and fail to capture diverse delta patterns.
+Lookahead prefetchers learn patterns by collecting histories of observed data access patterns, and correlating these with the strides, by considering each new page in a vacumm.
+Kim textit{et.
+al.} proposed SPP ~\cite{SPP}, a confidence-based lookahead prefetcher.
+SPP creates a signature associated to a memory request by compressing the history of accesses.
+By correlating the signature with future likely delta patterns, SPP learn both simple and complicated memory access patterns quickly.
+Our proposed technique builds upon SPP, therefore, we now introduce SPP in more detail.
 
@@
 any look-ahead prefetcher, we developed a practical implementation of the idea
-using SPP~\cite{SPP} as the underlying prefetcher.  SPP is a simple,
-lightweight, lookahead prefetcher which can be easily modified to be highly
-aggressive.  Here we describe the basic architecture of SPP.
+using SPP as our baseline.  Here we describe the basic architecture of SPP.
 
@@
 Delta is the numerical difference between the block offset of the 
-current and the previous memory acces. In case a matching page entry 
+current and the previous memory access. In case a matching page entry 
 is found, the stored signature retrieved and used to index into the 
-Pattern Table.
+Pattern Table. This process is illustrated by figure \ref{fig:spp_update}.
 
@@
 but since in our context filter refers to the perceptron filter, we will be calling 
-this structre as the Prefetch Table to avoid any confusion.
+this structure as the Prefetch Table to avoid any confusion.
 
-\textbf{Lookahead Prefetching:} On each trigger, SPP tries to go down
+\textbf{Lookahead Prefetching:} On each trigger, SPP goes down
 the program speculation path using its own prefetch suggestion.
 Using current prefetch as the baseline, it re-accesses the Pattern Table to generate further
-prefetches.  It keeps on repeating this cycle of accessing Pattern Table and
+prefetches. It repeats the cycle of accessing Pattern Table and
 updating the signature based on highest confidence prefetch from the
-last iteration.  These iteration count till which SPP manages to
+last iteration.  The iteration counter on which SPP manages to
 predict prefetch entries in the lookahead manner is characterized as
-its `depth'.  While doing this, SPP also keeps compounding the
+its `depth'.  While doing so, SPP also keeps compounding the
 confidence in each depth.  Thus as depth increases, overall confidence
@@
 
-\textbf{Confidence Tracking}: The Pattern Table keeps a track of hits to each
+\textbf{Confidence Tracking}: The Pattern Table keeps a count of hits to each
 signature through a counter C\textsubscript{sig}.  The number of hits
@@
 
-<Insert a comprehensive figure about SPP structures>
+%<Insert a comprehensive figure about SPP structures>
 
-<Insert another detailed picture about SPP data-path flow>
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=9cm]{figures/SPP_Prefetch_Description.jpg}
+    \label{fig:spp_strcture}
+  \caption{SPP architecture}
+  \end{center}
+\end{figure}
+
+%<Insert another detailed picture about SPP data-path flow>
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=9cm]{figures/SPP_Update_Description.jpg}
+  \label{fig:spp_update}
+  \caption{SPP data-path flow}
+  \end{center}
+\end{figure}
+
+% djimenez: this detail is unnecessary and likely to invite questions.
+% reviewers might ask "what about linear separability?" but the proof is in the
+% pudding.
+%This form of indexing makes sure that the hyper-plane learned by the
+%perceptron weights is able to differentiate between linearly
+%inseparable outcomes.
+
+%\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron
+%prediction involves multiplying the vector of input features:
+%F\textsubscript{1xN}, with the corresponding weight vector:
+%W\textsubscript{Nx1} in a dot product to obtain the sum y\textsubscript{out}.
+%Here what we use is a perceptron-like structure.  The feature is used to hash
+%into the weight of perceptrons and the retrieved weights are added straight
+%away.  Thus, the perceptron algorithm doesn't introduce any multiplication
+%operations in the inference or training process.  Hence what we adapt in this
+%work is a perceptron-like learning algorithm as it involves the same principle
+%involved in perceptron inferencing and training.
 
@@
 \label{Background-Case}
-As compared to some of the other state of the art prefetchers [BOP], SPP is
-less aggressive.  In the single-core environment, this gives BOP an edge as
-there is no resource contention among the different cores.  Hence more
-aggressive prefetching is bound to prove beneficial.  As we increase the core
+
+Compared to some of the other state of the art prefetchers, SPP is
+less aggressive.  In a single-core environment, there is no resource
+contention among the different cores, hence aggressive prefetching is bound to prove beneficial. 
+As we increase the core
 count, we observe that SPP starts outperforming rest of the prefetchers.  This
 can be attributed to the fact that each prefetch suggested by SPP is a
-carefully calculated one and that prevents cache pollution.  \textit{<Figure
+carefully calculated and prevents cache pollution.  \textit{<Figure
 showing the variation of SPP vs BOP wrt core count>} For 4-core applications,
-SPP suggests XX\% fewer prefetches than BOP and yet leads to higher IPC.
+SPP suggests XX\% fewer prefetches than Best Offset Prefetcher ~\cite{bop} and yet leads to higher IPC.
 
 The above analysis shows that with a more careful filtering mechanism, any
-prefetcher like SPP can be tuned to become much more aggressive, leading to
-increased coverage.  The onus of maintaining the accuracy now falls on the
+prefetcher like SPP can be tuned to become much more aggressive, yielding higher coverage.  The responsibility of maintaining high accuracy now falls on the
 independent filter.  To test the hypothesis, we tuned down SPP to the
@@
 filtering mechanism.
+
+\subsection{Supervised Learning}
+
+\textit{Perceptron Learning Rule:} In a perceptron-based mechanism, a uniform
+perceptron update principle is followed.  The weights need to be updated if
+the prediction was wrong or the magnitude of the dot product does not exceed
+a certain threshold.  After a certain training period, these weights are
+proportional to the probability of a correct prediction.  Thresholding the outcome to a
+ magnitude makes sure the perceptrons are trained until a certain level of
+confidence and yet they are not over-trained to the given set.  The perceptron 
+learning rule will be applied to PPF.

### 2018-12-03 related.tex
background and related work section

\section{Related Work}
\label{related}

The idea of using concepts from machine learning to drive cache management mechanism has become widely popular. Perceptrons in microarchitecture were first introduced by Jimenez and Lin ~\cite{Perc_Branch} for branch prediction. Their predictor consists of a table of weights that was indexed by the program counter associated to the branch. The dot product of the corresponding weight and a history of branches is thresholded against a magnitude to make a prediction. The perceptron-based predictor helped differentiate which branch history outcomes were relevant for marking a prediction for a certain branch, which yield higher accuracy than previously proposed branch predictors. This reason motivated the use of perceptrons in other speculation mechanisms.

To drive their replacement policy, Teran \textit{et.al} proposed using a perceptron-based predictor that obeys by the perceptron learning rule. By adding this way of reinforcement learning, their mechanism avoids overtraining and can rapidly adapt to changes in memory behavior. The continuation of this work was later introduced ~\cite{Multiperspective}, where a genetic algorithm was used to find an extensive set of features to be used as an input to the predictor. The extra features were able to cover more of the memory access patterns. The outcomes of these predictors are used to drive optimizations such a replacement, placement and bypass of the LLC.

%%MOVE IT SOMEWHERE
Peled \textit{et.al.} proposed using on-line reinforcement learnign to dynamically scale the maginitude of input information feed into their prefetcher mechanism ~\cite{Semantics}. The major drawback of this work is the need of compiler support. SVMs and logistic regression were used by Liao \textit{et.al.} to do an off-line optimal feature search for their prefetcher.

### 2018-12-04 design_arch.tex
Modified sections 4.2 and the inference section in 3.1

@@
 
-\pg{this discussion below has redundancy and doesn't flow well.
-  Rework it}
+%\pg{this discussion below has redundancy and doesn't flow well.
+%  Rework it}
 
-The perceptron sum can be considered the sum of the individual
-contribution of a feature.  The value of each contribution corresponds
-to a feature's confidence in the final decision on whether to prefetch
-a cache line. By summing the individual contributions, the final
-perceptron sum denotes the overall confidence for that prefetch
-suggestion.  By comparing the perceptron sum to two different
-thresholds, we divide the confidence scale of the prefetch suggestion
-into three bins. The first bin corresponds to the lowest confidence
-and leads to prefetch candidate being rejected.  The next bin
-corresponds to prefetches with a moderate confidence level.  This bin
-prefetches directeky to fill the bigger last level cache and prevents
-polluting the smaller L2 Cache.  The highest confidence prefetches are
-filled in the L2 cache.
+Each feature's confidence contributes to the final decision on 
+whether to prefetch a cache line, with the summation of all features' 
+confidences dentoting the overall confidence for a prefetch. 
+The final sum is compared against two thresholds: 
+ PERC\_THRESHOLD\_HI and  PERC\_ THRESHOLD\_ LO to decide the
+ placement of the data in the cache heirarchy. Prefetches that 
+have a confidence above the higher threshold are placed into 
+the L2 cache since the observed features strongly recommend  
+the prefetch. A prefetch that results in a confidence between 
+PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI are placed in 
+the larger last level cache, as the filter is moderately confident of 
+the cache line's use, but seeks to reduce pollution in the smaller 
+L2 cache. Finally, the remaining prefetches that are below the 
+lower threshold are not prefetched due to a lack of correlation 
+between the current observed features and the predicction from 
+the base prefetch engine.
 
-The sum of the perceptron weights are also used to decide whether to
-place the data in the L2 or the last level cache.  All prefetch
-candidates that are qualified until this stage are compared against
-PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
-PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.
+%GAC: The section's previous contents are below, in case the above 
+%rewritten work is not what we were looking for in terms of a rewording
+%
+%The perceptron sum can be considered the sum of the individual
+%contribution of a feature.  The value of each contribution corresponds
+%to a feature's confidence in the final decision on whether to prefetch
+%a cache line. By summing the individual contributions, the final
+%perceptron sum denotes the overall confidence for that prefetch
+%suggestion.  By comparing the perceptron sum to two different
+%thresholds, we divide the confidence scale of the prefetch suggestion
+%into three bins. The first bin corresponds to the lowest confidence
+%and leads to prefetch candidate being rejected.  The next bin
+%corresponds to prefetches with a moderate confidence level.  This bin
+%prefetches directeky to fill the bigger last level cache and prevents
+%polluting the smaller L2 Cache.  The highest confidence prefetches are
+%filled in the L2 cache.
+%
+%
+%The sum of the perceptron weights are also used to decide whether to
+%place the data in the L2 or the last level cache.  All prefetch
+%candidates that are qualified until this stage are compared against
+%PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
+%PERC\_THRESHOLD\_\_LO and PERC \_THRESHOLD\_HI are empirically set.
 

### 2018-12-04 design_impl.tex
Modified sections 4.2 and the inference section in 3.1

@@
 \begin{itemize}
-\item \textbf{Base Address} of the demand access that triggered the
-  prefetch.  Since prefetches are triggered from the L2 demand access,
-  they tend to correlate to the triggering address.
+\item \textbf{Base Address}: The base address of the demand 
+access that triggers the prefetch coresponds to a stream of 
+future accesses that SPP and the PPF have seen before. 
+Therefore, PPF will correlate the past behavior of this address 
+to the confidence of a prefetch.
 
@@
   the prefetches with depth >= 1 are aliased into the same PC which
-  would not be true in an actual demand access.  This feature resolves
+  is not true in an actual demand access.  This feature resolves
   a PC into a different value for each lookahead depth of prefetch
@@
   given PC favours particular value(s) of delta.  As noted earlier,
-  while the PC in itself does not convey much useful information, this
+  while the PC alone does not convey useful information, this
   hash resolves the PC into different values based on the tendency of
   that PC to favour a certain delta.  Thus, the dynamic nature of
-  different instances of the same memory access instruction can be
+  different instances of the same memory access instruction is 
   captured here.
+ % GAC: Seems more predictable than dynamic if its capturing 
+ % what is essentially a streaming access pattern. Not quite sure how to rephrase,
+ % if I'm correct
 
-\item \textbf{Confidence}: The integer confidence on a scale of 0 to
-  100 that was used in the original SPP design. While the original
-  confidence is not used directly for decision making, it can still
-  contribute to the final decision made.
+ % GAC: Again having trouble phrasing the below point. I was attempting to say
+ % that despite SPP having low confidence due to speculation, PPF may 
+ % reinforce prefetches with lower confidence if SPP has had a tendancy to be 
+ % correct in the past.
+\item \textbf{Confidence}: The integer confidence, on a scale of 0 to
+  100, that was used in the original SPP design. While the original 
+  confidence does not directly make the decision to prefetch,
+  PPF correlates it to the correctness of a proposed prefetch.
+  While the original SPP may have dismissed a prefetch due to running
+  further into speculation, PPF can use the original confidence as indicator
+  of not only when prefetches become less confident, but also how likely
+  a low confident speculation is correct in the context of other features.
 
@@
 
-% djimenez: this is a little vague. can you be more precise?
 \item \textbf{Current Signature XOR Delta}: Recall from the discussion

### 2018-12-04 design_arch.tex
Completing Methodology, Images

--- 2018-12-02 design_arch.tex
@@
 direct-mapped table.  If a prefetch suggestion is rejected by the
-perceptron layer, it is logged into the reject table.  The filer is
-used to train the perceptron to %Gino: Filer?
+perceptron layer, it is logged into the reject table.  The table is
+used to train the perceptron to %Gino: Filer? [EB]: Resolved
 avoid false negatives \textit{i.e.}, cases where prediction was to

### 2018-12-04 design_impl.tex
Completing Methodology, Images

--- 2018-12-02 design_impl.tex
@@
 
-%%[NOTE] Can this point be removed?
+% EB: Can this point be removed?
 % djimenez: I think this would be necessary to reproduce our results so leave it in. 
+% EB: It's ultimately a check to avoid SPP's internal segfault. Won't have a real impact.
 
-\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The
-  original SPP does not consider the slots available in the L2 MSHR
-  queue.  It has the assumption that at no time can the number of
-  suggested prefetches exceed the capacity of the queue.  While it was
-  not described in the paper, the internal confidence mechanism makes
-  sure that the above assumption was maintained.  In PPF, since there
-  cannot be an assumption on the degree of aggressiveness of the
-  underlying prefetcher, we explicitly check that at no point should
-  the number of suggested prefetches exceed the L2 MSHR queue. \pg{is
-    this a big deal?  Seems like a detail}
+%\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The
+%  original SPP does not consider the slots available in the L2 MSHR
+%  queue.  It has the assumption that at no time can the number of
+%  suggested prefetches exceed the capacity of the queue.  While it was
+%  not described in the paper, the internal confidence mechanism makes
+%  sure that the above assumption was maintained.  In PPF, since there
+%  cannot be an assumption on the degree of aggressiveness of the
+%  underlying prefetcher, we explicitly check that at no point should
+%  the number of suggested prefetches exceed the L2 MSHR queue. \pg{is
+%    this a big deal?  Seems like a detail}
 
@@
   prefetch suggestions.  For PPF, the prefetch table is modified to
-  store metadata such as program counter, memory access address etc.
+  store the features used for perceptron indexing..
   This metadata is needed to retrieve the
-  state %Gino: What's the etc.?
-  of the features to train the perceptron at a later stage in
+  values of the features to train the perceptron at a later stage in
   time. The prefetch table mentioned above keeps track of the
@@
   misprediction.
-
+%Gino: What's the etc.?
+%EB: Resolved
 \end{itemize}

### 2018-12-04 method.tex
Results Section and PDF Images
Completing Methodology, Images

@@
     \centering
-    \begin{tabular}{|l|p{3.6cm}|c|}
+    \begin{tabular}{|c|c|}
     \hline
-    Level & Configuration & Access Latency \\
+    	 Block & Configuration\\
     \hline
-         L1 Cache & Separate I-cache and D-cache, 32 KB, 8-way & 4 cycles\\
-         L2 Cache & Private, 256 KB & 8 cycles\\
-         LLC & Shared, 2MB / core, 16-way & 20 cycles\\
-         DRAM & 4 GB Single Channel for single-core, 8 GB Double Channel for multi-core & N?? cycles\\
+	 \multirow{2}{1.5cm}{CPU Core} 		& 1-8 Cores, 4 GHz\\
+						& 256 entry ROB, 4-wide\\
+    \hline
+         \multirow{2}{2.7cm}{Private L1 DCache} & 32 KB, 8-way, 4 cycles\\
+						& 8 MSHRs, LRU\\
+    \hline
+         \multirow{2}{2.4cm}{Private L2 Cache}  & 256 KB, 8-way, 8 cycles\\
+						& 16 MSHRs, LRU, Non-inclusive\\		
+    \hline
+         \multirow{2}{1.7cm}{Shared LLC} 	& 2MB/core, 16-way, 12 cycles\\
+						& 32 MSHRs, LRU, Non-inclusive\\
+    \hline
+         \multirow{3}{1.1cm}{DRAM} 		& 4 GB 1-Channel (single-core)\\ 
+						& 8 GB 2-Channels (multi-core)\\ 
+						& 64-bit channel, 1600MT/s\\
     \hline
     \end{tabular}
-    \caption{Memory Model Parameters}
+    \caption{Simulation Parameters}
     \label{tab:Sim_params}
@@
 of cache hierarchies. Branch prediction is done using the perceptron branch
-predictor~\cite{Perc_Branch}.  The page size is 4KB.  Champsim operates
+predictor~\cite{Perc_Branch}.  The page size is 4KB.  ChampSim operates
 all the prefetchers strictly in the physical address space.
@@
 In total, we used 285 traces representing workloads across 53 applications.
-Throughout the paper for SPEC 2017, we consider memory intensive subset as the
-traces with LLC MPKI > \_\_.  This includes 48 out of the 95 simpoints
-developed.  Since SPEC 2006 differs considerably in terms of memory behaviour,
-we define memory intensive subset as traces with LLC MPKI > \_\_.  This
-includes 47 out of 94 simpoints.
+Throughout the paper, we consider memory intensive subset as the
+traces with LLC MPKI > 1.  This includes 48 out of the 95 simpoints
+developed for SPEC 2017. For SPEC 2006  This includes 48 out of 94 simpoints.
 
@@
 
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=\columnwidth]{Hist}
+    \caption{Distribution of Trained Weights}
+    \label{Fig:Hist}
+  \end{center}
+\end{figure}
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=\columnwidth]{Pvals}
+    \caption{P-Values for all Features}
+    \label{Fig:Pvals}
+  \end{center}
+\end{figure}
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=\columnwidth]{Pvals_Trace}
+    \caption{P-value Variation across Traces for selected Features}
+    \label{Fig:Pvals_Trace}
+  \end{center}
+\end{figure}
+
 We plot a histogram for each feature depicting weights distribution
@@
 +1/-1 suggests a perfectly linear positive / negative correlation
-respectively.  Figure XX depicts the histogram and the Pearson's
-coefficient for two of the features, YY and ZZ.
+respectively.  Figure~\ref{Fig:Hist} depicts the histogram distribution 
+of trained weights, for a feature with high and another with moderate
+Pearson's coefficient.
 
-Figure XX shows all the features used, arranged in the increasing
-order of their Pearson's factor.  As can be seen X out of the 9
+Figure~\ref{Fig:Pvals} shows all the features used, arranged in the increasing
+order of their Pearson's factor.  As can be seen 5 out of the 9
 features provide a moderate to high correlation, with the magnitude of
-P-value > 0.4.  The single most important feature, \_\_ helps provide
-a correlation to prefetch outcome with a factor of \_\_.
+P-value > 0.6.  The single most important feature, 
+\textit{Confidence XOR Page address} helps provide
+a correlation to prefetch outcome with a factor of 0.90.
 
@@
 varies across the traces.  Here we give special attention to features
-with low P-values in the previous experiment.  Figure XX shows the
-variation of P-values three features : \_\_, \_\_ and \_\_; across all
-the SPEC 2017 traces.  For simplicity, the traces are arranged in an
+with low P-values in.  Figure~\ref{Pvals_Traces} shows the
+variation of P-values for three features : \textit{PC XOR Delta},
+\textit{Signature XOR Delta} and 
+\textit{PC XOR Depth}; across all the SPEC 2017 traces.  
+For simplicity, the traces are arranged in an
 increasing order of contribution made by the feature and the trace
 names have been omitted.  It can be seen that even features with a low
-overall correlation provide useful correlation (magnitude > 0.4) for
-XX out of the 95 simpoints developed for SPEC 2017.
+overall correlation provide useful correlation (magnitude > 0.5) for
+a significant number of traces.
 
@@
 
-\textit{<TODO: VALIDATE THIS>} Secondly, studying the relative
-importance of each feature enabled us to vary the number of entries
-dedicated for each feature.  Features with higher Pearson's
-correlation (like \_\_ and \_\_) were given most importance and
-allowed full 12-bits of indexing.  Features like \_\_ and \_\_, with a
-low overall P-value were allocated fewer indexes in the feature table.
+Secondly, studying the relative importance of each feature enabled 
+us to vary the number of entries dedicated for each feature.  
+Features with higher Pearson's correlation, \textit{cache line} 
+and \textit{page address}  were given most importance and
+allowed full 12-bits of indexing.  Features like \textit{ip XOR delta} 
+and \textit{ip XOR depth} with a low overall P-value 
+were allocated fewer indexes in the feature table.
 
@@
 
-\subsection{Overhead for PPF}
-\label{Method-Overheads}
-
-\begin{table}[t]
+\begin{table}[ht]
     \centering
-    \begin{tabular}{|c|c|m{4.8cm}|}
+    \begin{tabular}{|c|c|c|}
     \hline
@@
          Tag 		& 6  & Identifier for the entry in the table\\
-         Useful & 1 & To show if the given entry led to a useful demand fetch\\
+         \multirow{2}{1cm}{Useful} 	& 1  & To show if the given entry led to a\\
+                    	&    & useful demand fetch\\
          Perc Decision 	& 1  & Prefetched vs Not-prefetched \\
     \hline
-        PC & 12 & \multirow{5}{4.8cm}{Meta-data required for perceptron training}\\
-        Address & 24 & \\
+        PC 		& 12 & \multirow{4}{4.8cm}{Meta-data required for perceptron}\\
+        Address 	& 24 & \multirow{4}{1.1cm}{training}\\
         Delta 		& 7  & \\
@@
 \begin{table}[h]
-\begin{adjustwidth}{-0.5cm}{}
+\begin{adjustwidth}{-0.1cm}{}
     \centering
@@
     \hline
-        \multirow{5}{1.5cm}{Signature Table} &      & Valid (1 bit)        &             \\
-                                             &      & Tag (16 bits)        &             \\
-                                             & 256  & Last Offset (6 bits) & 11008 bits  \\  
+                                            &  \multirow{5}{0.5cm}{256}    & Valid (1 bit)  &             \\
+                                             &      & Tag (16 bits)        &  \multirow{2}{0.9cm}{11008}           \\
+                            Signature Table  &   & Last Offset (6 bits) &  \multirow{2}{0.5cm}{bits}  \\  
                                              &      & Signature (12 bits)  &             \\
@@
     \hline
-        \multirow{3}{2cm}{~~Pattern Table} &      & $C_{sig}$ (4bits)      &               \\
-                                             & 512  & $C_{delta}$ (4*4 bits) & 24576 bits    \\
+                                    &  \multirow{3}{0.5cm}{512}    & $C_{sig}$ (4bits)      &\multirow{2}{0.9cm}{24576}               \\
+                       Pattern Table         &   & $C_{delta}$ (4*4 bits) &  \multirow{2}{0.5cm}{bits}  \\
                                              &      & Delta (4*7 bits)       &               \\
     \hline
-        \multirow{3}{1.5cm}{Perceptron\newline}     & 4096*8    &            &              \\
-        \multirow{2}{0.9cm}{Weights}                 & 2048*x    & 5 bits     & 164480 bits  \\
+        \multirow{4}{1.5cm}{Perceptron\newline}     & 4096*4    & \multirow{4}{0.8cm}{5 bits}  & \multirow{3}{1.1cm}{113280}             \\
+        \multirow{3}{1.2cm}{Weights}                & 2048*2    &           &  \multirow{3}{0.5cm}{bits}  \\
+                                                    & 1024*2    &           &               \\
                                                     & 128*1     &           &              \\
     \hline
-        Prefetch Table\footnotemark[1]             & 1024      & 84 bits    & 86016 bits   \\
+        Prefetch                & \multirow{2}{0.7cm}{1024}      & \multirow{2}{1cm}{85 bits}       & 87040 \\
+        Table\footnotemark[1]   &           &               & bits\\
     \hline
-        Reject Table\footnotemark[2]               & 1024      & 83 bits    & 84992 bits   \\
+        Reject                & \multirow{2}{0.7cm}{1024}      & \multirow{2}{1cm}{84 bits}    & 86016 \\
+        Table\footnotemark[2] & & & bits\\
     \hline
-        \multirow{4}{1cm}{Global\newline\newline}   & \multirow{4}{0.2cm}{8} & Signature (12 bits)  & \multirow{4}{1.1cm}{264 bits} \\
+        \multirow{4}{1.0cm}{Global\newline\newline}   & \multirow{4}{0.2cm}{8} & Signature (12 bits)  & \multirow{4}{1.1cm}{264 bits} \\
         \multirow{3}{1.1cm}{History\newline}        &                        & Confidence (8 bits)  &                               \\
@@
     \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 371392 bits = 45.33 KB}}\\
+        \multicolumn{4}{|c|}{\textbf{Total: 3222940 bits = 39.34 KB}}\\
     \hline
@@
 
-\begin{figure}[h]
-\includegraphics[width=\columnwidth]{SPEC2017}
-\caption{SPEC 2017 Single-Core IPC Speedup}
-\label{Fig:SPEC2017_1core}
-\end{figure}
-
 In terms of computations, the perceptron mechanism only introduces an

### 2018-12-04 design_arch.tex
Changes in Design. Minor edits in intro / method / results

--- 2018-12-04 design_arch.tex
@@
 the prefetch insertion queue, to ensure that predicted bad prefetches
-are not inserted into the memory system, \pg{As shown in figure
-  whatever...}.
+are not inserted into the memory system, 
+\pg{As shown in figure whatever...}.
 
@@
 prefetch filtering mechanism which may adapted to any prefetcher with
-minimal modifications. %Gino: Are these modifications minimal?
+appropriate modifications.
+% Gino: Are these modifications minimal?
+% EB: That's a good question :P. Edited to 'appropriate'.
 
@@
 
+%Gino: repharse below?
 The Perceptron Filter is organized as tables of perceptron weights,
-with a separate table of weights for each feature.  A different number
-of bits are used for each feature to index into its corresponding
-table.
-%Gino: repharse below?
-Each feature's value is used to index into the corresponding table.
-This indexing is done using a different number of bits for each
-feature.  At most, 12 bits of a feature are used to index into the
-weights table. Certain features require more resolution \textit{i.e.},
-the full 12 bits of indexing.  On the other hand, some features
-requiring lower resolution require as few as 7 bits of indexing.  The
-variable indexing was determined by studying the features and
+with a separate table of weights for each feature. For a design using 
+`N' number of features, it's weights will be arranged in N different 
+tables. Each feature's value is used to index into the corresponding table.
+Since each table gets indexed independently, different number of bits can 
+be used for indexing, thus giving each table a different depth.
+The variable indexing can be determined by studying the features and
 fine-tuned empirically to achieve a good accuracy vs hardware overhead
-trade-off.  Exact details are mentioned under ``Area Overheads'' in
-Section~\ref{Method-Overheads}. \pg{this whole ``bits'' discussion
-  belongs in implementation not here IMHO}
+trade-off.  Exact details of our PPF implementation are mentioned 
+under ``Area Overheads'' in Section~\ref{Method-Overheads}.
+
+%\pg{this whole ``bits'' discussion belongs in implementation not here IMHO}
+%[EB] Resolved
 
@@
 weight is a 5 bit counter - saturating at -16 and +15 and initialized
-to zero at the beginning of a program's execution. Our proposed design
-uses nine features, represented in nine perceptron weight
-tables. \pg{the fact that its nine is again implementation details}
+to zero at the beginning of a program's execution. It was empirically 
+observed that 5-bits of indexing achieves a good trade-off between 
+hardware overheads and accuracy. 
+%\pg{the fact that its nine is again implementation details}
+% [EB] Resolved
 
-\pg{The word ChampSim should not appear in this part of the paper at
-  all.  Remember you are describing how to build this in the general
-  case...} \textbf{Inference}\newline In the ChampSim simulator, the
-prefetching method is invoked on every L2 Cache demand access.  At
+% \pg{The word ChampSim should not appear in this part of the paper at
+%   all.  Remember you are describing how to build this in the general
+%   case...} 
+% [EB] Resolved
+\textbf{Inference}\newline The prefetcher gets triggered on every 
+L2 Cache demand access.  At
 that point, the base prefetcher has the option to perform a prefetch
@@
 in the ``Prefetch Table''. The prefetch table is a 1024-entry, direct
-mapped structure that contains all metadata required to recreate the
-features.  %Gino: Recreate the features?
+mapped structure that contains all metadata required to re-index the 
+perceptron entries for training.  
 When the feedback of the current prefetched line is available at a
 later stage, the stored data is used to train the perceptron.
-
 %\pg{this discussion below has redundancy and doesn't flow well.
@@
 %corresponds to prefetches with a moderate confidence level.  This bin
-%prefetches directeky to fill the bigger last level cache and prevents
+%prefetches directly to fill the bigger last level cache and prevents
 %polluting the smaller L2 Cache.  The highest confidence prefetches are
@@
 direct-mapped table.  If a prefetch suggestion is rejected by the
-perceptron layer, it is logged into the reject table.  The filer is
-used to train the perceptron to %Gino: Filer?
+perceptron layer, it is logged into the reject table.  The table is
+used to train the perceptron to 
 avoid false negatives \textit{i.e.}, cases where prediction was to
 reject the prefetch but the prefetch would have been useful.
-
-% djimenez: if it were previously stated then it's redundant here? do we need
-% this statement?
 
@@
 is used to access a table, using 10-bits of the address for indexing
-and another 6-bits to perform tag matching.  Once the state of the
-program at the time of prefetching is available, it is
-%Gino: By "state of the program" are you referring to the state of the features?
+and another 6-bits to perform tag matching.  Once the values of the
+features at the time of prefetching is available, it is
 used to index into the perceptron weights table.
@@
 reflect the misprediction. In either case, weights are saturated at
--16 or +15. %Gino: Describe how these limits were decided?
+-16 or +15.
 
@@
 prefetch.  Once such a scenario is identified, the
-state %Gino: state of execution == value of features?
-of the execution at the time of prefetch is retrieved from the reject
+values of the features at the time of prefetch is retrieved from the reject
 table.  The retrieved data is used to index into the various weights
@@
 
-\item \textbf{Feature Selection:} In Section~\ref{Impl-Features}, we
-  show the features used for PPF. Six out of the nine features we
-  developed use information derived directly from program execution,
-  irrespective of the baseline prefetcher. Beyond that, the feature
-  set can be expanded to convey any useful information from the
-  prefetcher to the perceptron filter.  The methodology explained in
-  Section~\ref{Method-Features} talks about developing a minimal
-  feature set for PPF. \pg{This should be more about how one should
-    choose features than about what you did in the SPP implementation
-    in particular.  Rewrite as a guide on picking features.}
-%For example, in our implementation ``Confidence,'' ``Current Signature XOR Delta,''
-%and ``Page Address XOR Confidence'' are such features.
+\item \textbf{Feature Selection:}  Perceptrons essentially normalize
+  contributions from different features to get a single sum representing the
+  final confidence. So the perceptron learning can be as good as the set of 
+  features chosen. Interestingly, this is what makes perceptron learning 
+  scalable, as it can easily learn to incorporate newer information in 
+  form of new features. Some of the features we developed use information 
+  derived directly from program execution, agnostic to the baseline 
+  prefetcher. Beyond that, the feature set can be expanded to convey 
+  any useful information from the prefetcher to the perceptron filter.
+
+  %[EB] This statement fits better in the next section 'Features used by Perceptron'
+  %The methodology explained in
+  %Section~\ref{Method-Features} talks about developing a minimal feature 
+  %set for PPF.
+
+%\pg{This should be more about how one should
+%choose features than about what you did in the SPP implementation
+%in particular.  Rewrite as a guide on picking features.}
+% [EB]: Hopefully this is general enough and not too redundant
 

### 2018-12-04 design_impl.tex
Changes in Design. Minor edits in intro / method / results

@@
   misprediction.
-%Gino: What's the etc.?
-%EB: Resolved
 \end{itemize}
@@
 Here we discuss the various features that correlate the prefetching
-decision with the program behavior. \pg{again you want to get at why
-  these were chosen.  Also mention that you searched over a large
-  space of possible features and these were the best.}
+decision with the program behavior. All the features we used can be 
+derived from the information available in the L2 Cache access stream.
+Our feature selction involved searching over a large space of 
+relevant perceptron features. Using the statistical methodology 
+outlined in Section~\ref{Method-Features}, we pruned the feature 
+set to a minimal yet relevant set of features
+% \pg{again you want to get at why
+% these were chosen.  Also mention that you searched over a large
+% space of possible features and these were the best.}
+
+% [EB] Tried addressing. Does it look reduntant when read after 'Feature Selection'
+% bullet from the previous section?
 
@@
 % djimenez: this is a nice intuition and explanation. 
+% [EB] Thank you!
 
@@
 %Gino: Added the sentence below, is this correct?
-  The result of this feature is the next delta that is predicted to be
+%[EB]: It gives the next signature instead of next delta
+  The result of this feature is the next signature that is predicted to be
   accessed based on the delta predicted by SPP. While ``Current
@@
 entry of perceptron weights.
+
+% [EB]: Think of more implementation specific details
+% 7 to 12 bits of indexing

### 2018-12-04 abs.tex
spell check, editing for grammar, clarity, style, put in some questions preceded by 'djimenez'

@@
 \begin{abstract}
+
+% djimenez: I have taken out parenthetical terms in favor of explicitly
+% mentioning them in the text. some people have a visceral reaction against
+% parentheses that delimit things essential to understanding the text.
 
@@
 in modern processor designs.  Prefetcher performance can be characterized by
-two main metrics, which are generally at odds with one another: coverage 
-(the improvement in the cache miss metric) and accuracy (the fraction of 
-useful prefetches).  An overly aggressive prefetcher may improve coverage, but
-may waste so many resources (e.g., bandwidth) that performance is ultimately 
-harmed.  An ideal prefetcher would have both very high coverage and accuracy.
+two main metrics that are generally at odds with one another: coverage, {\em
+i.e.} the improvement in cache misses; and accuracy, {\em i.e.} the fraction
+of useful prefetches.  An overly aggressive prefetcher may improve coverage,
+but may waste so many resources, {\em e.g.}, bandwidth, that performance is
+harmed.  An ideal prefetcher would have both high coverage and accuracy.
 
@@
 to reject inaccurate candidate prefetches generated by a baseline prefetcher.
-PPF directly increases prefetcher accuracy, and enables more aggressive tuning
-of the baseline prefetcher, which in turn leads to increased coverage.  We also
-explore a range of features to use to train PPF's perceptron layer.  Our results
-show that PPF improves performance on a memory-intensive subset of SPEC CPU 2017
-benchmarks by 6.84\% for a single-core configuration, and by 11.9\% for a 4-core
+PPF directly increases prefetcher accuracy and enables more aggressive tuning
+of the baseline prefetcher, leading to increased coverage.  We also explore a
+range of features to use to train PPF's perceptron layer.  PPF improves
+performance on a memory-intensive subset of the SPEC CPU 2017 benchmarks by
+6.84\% for a single-core configuration, and by 11.9\% for a 4-core
 configuration, compared to the baseline prefetcher alone.  We also show that
-performance continues to scale as the number of cores sharing a last level cache
-increases.
+performance continues to scale as the number of cores sharing a last level
+cache increases.
 
@@
 %train the perceptron layer.  Our results show that perceptron-based filtering
-%improves performance on the memory intensive subset of the SPEC 2017 benchmark
+%improves performance on the memory intensive subset of the SPEC CPU 2017 benchmark
 %suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared

### 2018-12-04 back.tex
Completing Methodology, Images
spell check, editing for grammar, clarity, style, put in some questions preceded by 'djimenez'

@@
 
-In this section we discussed the most closely related work to our proposed technique. Other relevant work will be discussed throughout the paper. See section \ref{related}.
+In this section we discuss the most closely related work to our proposed
+technique. See Section~\ref{related} for other relevant work.
 
-	The idea of prefetching has been around ever since Jouppi introduced \textit{instruction stream buffers}.
-Early prefetchers detected stride access patterns in order to predict future memory requests ~\cite{Smith,Baer,Stride}.
-These days prefetching mechanisms are more sophisticated, they look into past memory behavior~\cite{Address_Correlated,AMPM}, locality ~\cite{Spatial_Pattern,SMS,Temporal_Instruction_Fetch,Off_Chip,STMS,SMS_JILP}, control-flow speculation ~\cite{BFetch,MTBFetch}, between many other aspects in order to detect complex access patterns in modern applications.
+The idea of prefetching begins with Jouppi's \textit{instruction stream
+buffers}. Early prefetchers detected stride access patterns in order to
+predict future memory requests~\cite{Smith,Baer,Stride}.  Modern prefetching
+mechanisms are more sophisticated as they look into past memory
+behavior~\cite{Address_Correlated,AMPM},
+locality~\cite{Spatial_Pattern,SMS,Temporal_Instruction_Fetch,Off_Chip,STMS,SMS_JILP},
+control-flow speculation~\cite{BFetch,MTBFetch}, and many other aspects to
+detect complex access patterns in modern applications.
 
@@
 
-An optimal prefetcher should be able to detect a wide range of memory access patterns.
-Simple stride prefetching techniques only detect sequences of addresses that differ by a constant value and fail to capture diverse delta patterns.
-Lookahead prefetchers learn patterns by collecting histories of observed data access patterns, and correlating these with the strides, by considering each new page in a vacumm.
-Kim textit{et.
-al.} proposed SPP ~\cite{SPP}, a confidence-based lookahead prefetcher.
-SPP creates a signature associated to a memory request by compressing the history of accesses.
-By correlating the signature with future likely delta patterns, SPP learn both simple and complicated memory access patterns quickly.
-Our proposed technique builds upon SPP, therefore, we now introduce SPP in more detail.
+An optimal prefetcher should be able to detect a wide range of memory access
+patterns. Simple stride prefetching techniques only detect sequences of
+addresses that differ by a constant value and fail to capture diverse delta
+patterns. Lookahead prefetchers learn patterns by collecting histories of
+observed data access patterns, and correlating these with the strides, by
+considering each new page in a vacuum. Kim {\em et. al.} proposed
+SPP~\cite{SPP}, a confidence-based lookahead prefetcher. SPP creates a
+signature associated to a memory request by compressing the history of
+accesses. By correlating the signature with future likely delta patterns, SPP
+learn both simple and complicated memory access patterns quickly. Our proposed
+technique builds upon SPP. Thus, we describe SPP in more detail.
 
@@
 While the basic idea of perceptron based prefetch filtering is applicable to
-any look-ahead prefetcher, we developed a practical implementation of the idea
+any look-ahead prefetcher, we develop a practical implementation of the idea
 using SPP as our baseline.  Here we describe the basic architecture of SPP.
 
-% djimenez: Paul, is all this stuff accurate? SPP continues to confuse me.
+\textbf{Signature Table:} The Signature Table keeps track of 256 most recently
+accessed pages.  It is meant to capture memory access patterns within a page
+boundary.  SPP indexes into an entry of the Signature Table using the page
+number.  For each entry corresponding to a page, the table stores a `last
+block offset' and an `old signature'.  Last block offset is the block offset
+of the last memory access of that given page.  The block offset is calculated
+with respect to the page boundary.  The signature is a 12-bit compressed
+representation of the past few memory accesses for that page.  The signature
+is calculated as: $$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\;
+(\,Delta\,)$$ Delta is the numerical difference between the block offset of
+the current and the previous memory access. In case a matching page entry is
+found, the stored signature retrieved and used to index into the Pattern
+Table. This process is illustrated by figure \ref{fig:spp_update}.
 
-\textbf{Signature Table:} The Signature Table keeps track of 256 most recently accessed
-pages.  It is meant to capture memory access patterns within a page
-boundary.  SPP indexes into an entry of the Signature Table using the page number.  For
-each entry corresponding to a page, the table stores a `last block offset'
-and an `old signature'.  Last block offset is the block offset of the
-last memory access of that given page.  The block offset is calculated
-with respect to the page boundary.  The signature is a 12-bit
-compressed representation of the past few memory accesses for that
-page.  The signature is calculated as:
-$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
-Delta is the numerical difference between the block offset of the 
-current and the previous memory access. In case a matching page entry 
-is found, the stored signature retrieved and used to index into the 
-Pattern Table. This process is illustrated by figure \ref{fig:spp_update}.
-
-\textbf{Pattern Table:} The Pattern Table is indexed by the signature generated
-from the Signature Table.  Pattern Table holds predicted delta patterns and their confidence
-estimates.  Each entry indexed by the signature holds up to 4 unique
-delta predictions.  This is implemented by making the Pattern Table as a 4-way
-associative table.
+\textbf{Pattern Table:} The Pattern Table is indexed by the signature
+generated from the Signature Table.  Pattern Table holds predicted delta
+patterns and their confidence estimates. Each entry indexed by the signature
+holds up to 4 unique delta predictions.  This is implemented by making the
+Pattern Table 4-way associative.
 
@@
 
-\textbf{Prefetch Table:} The Prefetch Table is a 1024 entry 1-way associative table that
-keeps a record of last few entries prefetched.  This proves useful for
-updating the prefetcher states if a tracked prefetch leads to a demand hit or
-a cache eviction. \textit{Note:} The original paper refers to this as the Prefetch Filter 
-but since in our context filter refers to the perceptron filter, we will be calling 
-this structure as the Prefetch Table to avoid any confusion.
+\textbf{Prefetch Table:} The Prefetch Table is a 1024 entry 1-way associative
+table that keeps a record of last few entries prefetched.  This proves useful
+for updating the prefetcher states if a tracked prefetch leads to a demand hit
+or a cache eviction. \textit{Note:} The original paper refers to this as the
+Prefetch Filter but since in our context filter refers to the perceptron
+filter, we will be calling this structure as the Prefetch Table to avoid any
+confusion.
 
-\textbf{Lookahead Prefetching:} On each trigger, SPP goes down
-the program speculation path using its own prefetch suggestion.
-Using current prefetch as the baseline, it re-accesses the Pattern Table to generate further
-prefetches. It repeats the cycle of accessing Pattern Table and
-updating the signature based on highest confidence prefetch from the
-last iteration.  The iteration counter on which SPP manages to
-predict prefetch entries in the lookahead manner is characterized as
-its `depth'.  While doing so, SPP also keeps compounding the
-confidence in each depth.  Thus as depth increases, overall confidence
-keeps decreasing.
+\textbf{Lookahead Prefetching:} On each trigger, SPP goes down the program
+speculation path using its own prefetch suggestion.  Using current prefetch as
+the baseline, it re-accesses the Pattern Table to generate further prefetches.
+It repeats the cycle of accessing Pattern Table and updating the signature
+based on highest confidence prefetch from the last iteration.  The iteration
+counter on which SPP manages to predict prefetch entries in the lookahead
+manner is characterized as its `depth'.  While doing so, SPP also keeps
+compounding the confidence in each depth.  Thus as depth increases, overall
+confidence keeps decreasing.
 
 \textbf{Confidence Tracking}: The Pattern Table keeps a count of hits to each
-signature through a counter C\textsubscript{sig}.  The number of hits
-for a given delta per signature are tracked using a counter
-C\textsubscript{delta}.  The confidence for a given delta is
-approximated through C\textsubscript{d} = C\textsubscript{delta} /
-C\textsubscript{sig}.  When SPP enters into a lookahead mode, the path
-confidence P\textsubscript{d} is given as:
-$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ 
-Here $\alpha$ represents the global accuracy, calculated as the ratio of 
-the number of prefetches which led to a demand hit to the number of 
-prefetches recommended in total. The range of $\alpha$ is [0,1].
+signature through a counter C\textsubscript{sig}.  The number of hits for a
+given delta per signature are tracked using a counter C\textsubscript{delta}.
+The confidence for a given delta is approximated through C\textsubscript{d} =
+C\textsubscript{delta} / C\textsubscript{sig}.  When SPP enters into a
+lookahead mode, the path confidence P\textsubscript{d} is given as:
+$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;
+P\textsubscript{d-1}$$ Here $\alpha$ represents the global accuracy,
+calculated as the ratio of the number of prefetches which led to a demand hit
+to the number of prefetches recommended in total. The range of $\alpha$ is
+[0,1].
 
@@
 
-%<Insert a comprehensive figure about SPP structures>
-
 \begin{figure}
   \begin{center}
-  \includegraphics[width=9cm]{figures/SPP_Prefetch_Description.jpg}
+  \includegraphics[width=9cm]{SPP_Prefetch_Description}
     \label{fig:spp_strcture}
@@
 
-%<Insert another detailed picture about SPP data-path flow>
 \begin{figure}
   \begin{center}
-  \includegraphics[width=9cm]{figures/SPP_Update_Description.jpg}
+  \includegraphics[width=9cm]{SPP_Update_Description}
   \label{fig:spp_update}
@@
 work in the original implementation, we believe that a better form of
-generalised on-line decision making was possible.  Hence, it was necessary to
+generalized on-line decision making was possible.  Hence, it was necessary to
 build a robust and adaptable learning mechanism to accept / reject the
@@
 
-\subsection{Supervised Learning}
+\subsection{Perceptron Learning}
+\label{sec:Background-Perceptron}
+Perceptron learning for microarchitectural prediction was introduced for
+branch prediction~\cite{PerceptronPredictor}. Our predictor uses a version of
+microarchitectural perceptron prediction known as the ``hashed perceptron''
+organization~\cite{HashedPerceptron}. As an abstract idea, a hashed perceptron
+predictor hashes several different features into values that index several
+distinct tables. Small integer weights are read out from the tables and
+summed. If the sum exceeds some threshold, a positive prediction is made, {\em
+e.g.} ``predict branch taken'' or ``allow the prefetch.'' Otherwise, a
+negative prediction is made. Once the ground truth is known, the weights
+corredponding to the prediction are incremented if the outcome was positive,
+or decremented if it was negative. This update only occurs if the prediction
+was incorrect or if the magnitude of the sum failed to exceed a threshold.
+Beyond branch prediction, perceptron learning has been applied to last-level
+cache reuse prediction~\cite{Perc_Reuse,Multiperspective}. In this paper, we
+apply it for the first time to prefetch filtering.
 
-\textit{Perceptron Learning Rule:} In a perceptron-based mechanism, a uniform
-perceptron update principle is followed.  The weights need to be updated if
-the prediction was wrong or the magnitude of the dot product does not exceed
-a certain threshold.  After a certain training period, these weights are
-proportional to the probability of a correct prediction.  Thresholding the outcome to a
- magnitude makes sure the perceptrons are trained until a certain level of
-confidence and yet they are not over-trained to the given set.  The perceptron 
-learning rule will be applied to PPF.
+%\textit{Perceptron Learning Rule:} In a perceptron-based mechanism, a uniform
+%perceptron update principle is followed.  The weights need to be updated if
+%the prediction was wrong or the magnitude of the dot product does not exceed a
+%certain threshold.  After a certain training period, these weights are
+%proportional to the probability of a correct prediction.  Thresholding the
+%outcome to a
+% magnitude makes sure the perceptrons are trained until a certain level of
+%confidence and yet they are not over-trained to the given set.  The perceptron
+%learning rule will be applied to PPF.

### 2018-12-04 conc.tex
Separating arch and implementation. Filter-Table and Throttler-Filter renaming
spell check, editing for grammar, clarity, style, put in some questions preceded by 'djimenez'

@@
 \label{Conclusion}
-In this paper, we introduced neural network based learning for data
-prefetching in the form of PPF.  Perceptron acts an orthogonal
-throttler to the underlying prefetch engine.  PPF improves
-performance by over XX\% over the baseline, which corresponds to YY\%
-over the next best performing prefetcher.  
-We believe that this is a robust and adaptable technique which can
-be used to enhance any existing or upcoming prefetchers.
+In this paper, we introduced perceptron-based learning for data prefetching.
+The perceptron filter acts independently as a check on the underlying prefetch
+engine.  PPF improves performance by over XX\% over the baseline,
+corresponding to YY\% over the next best performing prefetcher.  We believe
+that this is a robust and adaptable technique that can be used to enhance any
+existing or upcoming prefetchers.

### 2018-12-04 design_arch.tex
spell check, editing for grammar, clarity, style, put in some questions preceded by 'djimenez'

@@
 
-As previously discussed, ultimately it might be profitable to allow a
-given prefetcher to speculate as deeply as possible, since often some
-useful prefetches are still being generated long after the confidence
-of the prefetcher has fallen below the point at which performance
-improvements are seen, due to the increasing rate of inaccurate
-prefetches.  In order to allow this deep speculation in the
-prefetcher, however, the increasing numbers of bad prefetches must be
-filtered out.  To this end, we propose to leverage perceptron-based
-machine learning as a mechanism to differentiate between potentially
-useful deeply speculated prefetches and likely not-useful ones.  This
-Perceptron Prefetch Filter (PPF) is placed between the prefetcher and
-the prefetch insertion queue, to ensure that predicted bad prefetches
-are not inserted into the memory system, 
+As previously discussed, ultimately it might be profitable to allow a given
+prefetcher to speculate as deeply as possible, since often some useful
+prefetches are generated long after the confidence of the prefetcher has
+fallen below the point at which performance improvements are seen due to the
+increasing rate of inaccurate prefetches.  However, to allow this deep
+speculation in the prefetcher the increasing numbers of bad prefetches must be
+filtered out.  Thus, we propose to leverage perceptron-based learning as a
+mechanism to differentiate between potentially useful deeply speculated
+prefetches and likely not-useful ones.  This Perceptron Prefetch Filter (PPF)
+is placed between the prefetcher and the prefetch insertion queue, to ensure
+that predicted bad prefetches are not inserted into the memory system.
 \pg{As shown in figure whatever...}.
 
-As discussed in Section~\ref{sec:Background-Perceptron}, perceptrons
-are useful, light-weight mechanisms to pull together disparate forms
-of information and synthesize a decision from them.  In this case, we
-will leverage a variety of different relative characteristics of a
-given prefetch, such as speculation depth, page address and offset,
-into a binary useful/non-useful prediction.  In the remainder of this
-section, we discuss the the general structure of PPF for enhancing a
-given underlying prefetcher.  We will show that PPF is a generalized
-prefetch filtering mechanism which may adapted to any prefetcher with
-appropriate modifications.
+As discussed in Section~\ref{sec:Background-Perceptron}, perceptron learning
+is a light-weight mechanism to pull together disparate forms of information
+and synthesize a decision from them.  In this case, we leverage a variety of
+different relative characteristics of a given prefetch, such as speculation
+depth, page address and offset, into a binary useful/non-useful prediction.
+In the remainder of this section, we discuss the the general structure of PPF
+for enhancing a given underlying prefetcher.  We will show that PPF is a
+generalized prefetch filtering mechanism that may be adapted to any prefetcher
+with appropriate modifications.
+
 % Gino: Are these modifications minimal?
@@
 
-%Gino: repharse below?
-The Perceptron Filter is organized as tables of perceptron weights,
-with a separate table of weights for each feature. For a design using 
-`N' number of features, it's weights will be arranged in N different 
-tables. Each feature's value is used to index into the corresponding table.
-Since each table gets indexed independently, different number of bits can 
-be used for indexing, thus giving each table a different depth.
-The variable indexing can be determined by studying the features and
-fine-tuned empirically to achieve a good accuracy vs hardware overhead
-trade-off.  Exact details of our PPF implementation are mentioned 
-under ``Area Overheads'' in Section~\ref{Method-Overheads}.
+%Gino: rephrase below?
+The Perceptron Filter is organized as tables of perceptron weights, with a
+separate table of weights for each feature. For a design using `N' number of
+features, its weights will be arranged in $N$ different tables. Each feature's
+value is used to index into the corresponding table.  Since each table is
+indexed independently, different numbers of bits can be used for indexing,
+thus giving each table a different size.  The variable indexing can be
+determined by studying the features and fine-tuning them empirically to
+achieve a good accuracy versus hardware overhead trade-off.  Exact details of
+our PPF implementation are given under ``Area Overheads'' in
+Section~\ref{Method-Overheads}.
 
@@
 
-A single entry in the table corresponds to a perceptron weight.  Each
-weight is a 5 bit counter - saturating at -16 and +15 and initialized
-to zero at the beginning of a program's execution. It was empirically 
-observed that 5-bits of indexing achieves a good trade-off between 
-hardware overheads and accuracy. 
+A single entry in the table corresponds to a perceptron weight.  Each weight
+is a 5 bit counter saturating at -16 and +15 and initialized to zero at the
+beginning of a program's execution. We observed empirically that 5 bits of
+index achieves a good trade-off between hardware overheads and accuracy.
+
 %\pg{the fact that its nine is again implementation details}
@@
 % [EB] Resolved
-\textbf{Inference}\newline The prefetcher gets triggered on every 
-L2 Cache demand access.  At
-that point, the base prefetcher has the option to perform a prefetch
--- if it does, then it has a choice of how many cache lines to
-prefetch.  The suggested prefetches can be either placed in the L2
-cache or the last-level cache based on the prefetcher's internal
-confidence mechanism.
+\textbf{Inference}\newline The prefetcher is triggered on every L2 cache
+demand access.  At that point, the base prefetcher has the option to perform a
+prefetch -- if it does, then it has a choice of how many cache lines to
+prefetch.  The suggested prefetches can be either placed in the L2 cache or
+the last-level cache based on the prefetcher's internal confidence mechanism.
 
-When the base prefetcher is triggered, it begins to suggest candidates
-for prefetching.  All suggested prefetches are passed through the
-perceptron filter to decide if they are qualified for final
-prefetching.  The perceptron decides whether to prefetch a candidate
-by looking at the microarchitectural state \textit{i.e.} the features,
-at that instant.  Each feature is hashed to form an index into a table
-of up to 4096 entries dedicated for that feature.
+When the base prefetcher is triggered, it begins to suggest candidates for
+prefetching.  All suggested prefetches are passed through the perceptron
+filter to decide whether they are qualified for final prefetching.  The filter
+decides whether to prefetch a candidate by looking at the microarchitectural
+state, \textit{i.e.} the features, at that instant.  Each feature is hashed to
+form an index into a table of up to 4,096 entries dedicated to that feature.
 
-Once all the weights are retrieved, they are summed and compared to a
-preset threshold, PERC\_ THRESHOLD\_LO.  Only the prefetch candidates
-with perceptron sum higher than the threshold qualify for prefetching.
-The prefetches that qualify through the perceptron stage are recorded
-in the ``Prefetch Table''. The prefetch table is a 1024-entry, direct
-mapped structure that contains all metadata required to re-index the 
-perceptron entries for training.  
-When the feedback of the current prefetched line is available at a
-later stage, the stored data is used to train the perceptron.
+Once all the weights are retrieved, they are summed and compared to a preset
+threshold, PERC\_ THRESHOLD\_LO.  Only prefetch candidates with perceptron sum
+higher than the threshold qualify for prefetching.  The prefetches that
+qualify through the perceptron stage are recorded in the ``Prefetch Table''.
+The prefetch table is a 1,024-entry, direct mapped structure that contains all
+metadata required to re-index the perceptron entries for training.  When the
+feedback of the current prefetched line is available at a later stage, the
+stored data is used to train the perceptron.
+
 %\pg{this discussion below has redundancy and doesn't flow well.
@@
 
-Each feature's confidence contributes to the final decision on 
-whether to prefetch a cache line, with the summation of all features' 
-confidences dentoting the overall confidence for a prefetch. 
-The final sum is compared against two thresholds: 
- PERC\_THRESHOLD\_HI and  PERC\_ THRESHOLD\_ LO to decide the
- placement of the data in the cache heirarchy. Prefetches that 
-have a confidence above the higher threshold are placed into 
-the L2 cache since the observed features strongly recommend  
-the prefetch. A prefetch that results in a confidence between 
-PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI are placed in 
-the larger last level cache, as the filter is moderately confident of 
-the cache line's use, but seeks to reduce pollution in the smaller 
-L2 cache. Finally, the remaining prefetches that are below the 
-lower threshold are not prefetched due to a lack of correlation 
-between the current observed features and the predicction from 
-the base prefetch engine.
+Each feature's confidence contributes to the final decision of whether to
+prefetch a cache line, with the summation of all features' confidences
+denoting the overall confidence for a prefetch.  The final sum is compared
+against two thresholds: PERC\_THRESHOLD\_HI and  PERC\_ THRESHOLD\_ LO to
+decide the placement of the data in the cache hierarchy. Prefetches that have
+a confidence above the higher threshold are placed into the L2 cache since the
+observed features strongly recommend  the prefetch. A prefetch that results in
+a confidence between PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI are
+placed in the larger last level cache, as the filter is moderately confident
+of the cache line's use, but seeks to reduce pollution in the smaller L2
+cache. Finally, the remaining prefetches that are below the lower threshold
+are not prefetched due to a lack of correlation between the current observed
+features and the prediction from the base prefetch engine.
 
@@
 
-In addition to the prefetch table mentioned above, PPF also maintains
-a ``Reject Table.''  The reject table is a 1024-entry deep
-direct-mapped table.  If a prefetch suggestion is rejected by the
-perceptron layer, it is logged into the reject table.  The table is
-used to train the perceptron to 
-avoid false negatives \textit{i.e.}, cases where prediction was to
-reject the prefetch but the prefetch would have been useful.
+In addition to the prefetch table mentioned above, PPF also maintains a
+``Reject Table.''  The reject table is a 1,024-entry direct-mapped table.
+If a prefetch suggestion is rejected by the perceptron layer, it is logged
+into the reject table.  The table is used to train the perceptron to avoid
+false negatives \textit{i.e.}, cases where prediction was to reject the
+prefetch but the prefetch would have been useful.
 
@@
 
-\textbf{Training}\newline In the prefetching environment, feedback for
-a prefetch is received whenever there is an eviction or a demand
-access from the L2 Cache.  This action triggers training of the base
-prefetcher as well as the perceptron layer.  Training the perceptrons
-involves accessing the metadata that was stored in Prefetch Table or
-the Reject Table.  The cache lines address of the block being trained
-is used to access a table, using 10-bits of the address for indexing
-and another 6-bits to perform tag matching.  Once the values of the
-features at the time of prefetching is available, it is
-used to index into the perceptron weights table.
+\textbf{Training}\newline In the prefetching environment, feedback for a
+prefetch is received whenever there is an eviction or a demand access from the
+L2 cache.  This action triggers training of the base prefetcher as well as the
+perceptron layer.  Training the perceptrons involves accessing the metadata
+that was stored in Prefetch Table or the Reject Table.  The cache lines
+address of the block being trained is used to access a table, using 10 bits of
+the address for indexing and another 6 bits to perform tag matching.  Once the
+values of the features at the time of prefetching is available, it is used to
+index into the perceptron weights table.
 
 If the demanded block that triggers the training was tagged as a valid
-prefetch in the prefetch table, then the earlier prefetch prediction
-was correct.  In that case the perceptron weights are incremented by 1
-if the predicted sum does not cross a pre-defined threshold. These
-training thresholds are introduced to avoid overfitting of the
-perceptron weights to the given program behaviour. These thresholds
-are referred to as POS\_UPDATE\_THRESHOLD ($\theta_p$) and
-NEG\_UPDATE\_THRESHOLD ($\theta_n$), respectively for the positive and
-negative values of training saturation.
+prefetch in the prefetch table, then the earlier prefetch prediction was
+correct.  In that case the perceptron weights are incremented by 1 if the
+predicted sum does not cross a pre-defined threshold. These training
+thresholds are introduced to avoid overfitting of the perceptron weights to
+the given program behavior. These thresholds are referred to as
+POS\_UPDATE\_THRESHOLD ($\theta_p$) and NEG\_UPDATE\_THRESHOLD ($\theta_n$),
+respectively for the positive and negative values of training saturation.
 
-If a cache block eviction led to training and the corresponding valid
-entry is found in the prefetch table without the block being accessed,
-then the prediction made by the perceptron was wrong.  The perceptron
-should have ideally rejected the prefetch suggestion as a
-low-confidence prefetch.  Here the weights are decremented by 1 to
-reflect the misprediction. In either case, weights are saturated at
--16 or +15.
+If a cache block eviction led to training and the corresponding valid entry is
+found in the prefetch table without the block being accessed, then the
+prediction made by the perceptron was wrong.  The perceptron should have
+ideally rejected the prefetch suggestion as a low-confidence prefetch.  Here
+the weights are decremented by 1 to reflect the misprediction. In either case,
+weights are saturated at -16 or +15.
 
-A secondary training mechanism also kicks in during demand fetches.
-Before the demand access triggers the next set of prefetches, the
-reject table is checked for a valid entry.  A hit means that the
-corresponding cache line was initially suggested by the underlying
-prefetch engine, but rejected by the perceptron filter.  Thus, the
-perceptron should have been more confident about that particular
-prefetch.  Once such a scenario is identified, the
-values of the features at the time of prefetch is retrieved from the reject
-table.  The retrieved data is used to index into the various weights
-tables of the perceptron and the corresponding values are updated by
-+1, saturating between and -16 and +15.  This update reflects
-increased confidence for the prediction corresponding to that
-prefetch.
+A secondary training mechanism also kicks in during demand fetches.  Before
+the demand access triggers the next set of prefetches, the reject table is
+checked for a valid entry.  A hit means that the corresponding cache line was
+initially suggested by the underlying prefetch engine, but rejected by the
+perceptron filter.  Thus, the perceptron should have been more confident about
+that particular prefetch.  Once such a scenario is identified, the values of
+the features at the time of prefetch is retrieved from the reject table.  The
+retrieved data is used to index into the various weights tables of the
+perceptron and the corresponding values are updated by +1, saturating between
+and -16 and +15.  This update reflects increased confidence for the prediction
+corresponding to that prefetch.
 
-This mechanism allows us to exploit a previously lost opportunity.  In
-prior perceptron based implementations, and in general, prefetching
-algorithms, there is usually no way of knowing the result of not
-prefetching a particular line.  Our two-step PPF architecture allows
-us to overcome that issue.
+This mechanism allows us to exploit a previously lost opportunity.  In prior
+perceptron based implementations, and in general, prefetching algorithms,
+there is usually no way of knowing the result of not prefetching a particular
+line.  Our two-step PPF architecture allows us to overcome that issue.
 
@@
 \label{Arch-Generalizing}
-The above discussion of PPF shows that it is highly modular and can be
-adapted to be used over any base prefecher for increased prefetch
-accuracy.  In general, only two hooks are required between PPF and the
-baseline prefetcher. The first is to make sure that all the prefetch
-candidates of the prefetcher pass through the perceptron filter and if
-qualified, the metadata for perceptron indexing be stored. The second
-is needed when the feedback of a prior prefetch is available in form
-of a subsequent demand hit or cache eviction. In that case, the stored
-metadata needs to be retrieved to update the state of the perceptrons.
+The above discussion of PPF shows that it is highly modular and can be adapted
+to be used over any base prefetcher for increased prefetch accuracy.  In
+general, only two hooks are required between PPF and the baseline prefetcher.
+The first is to make sure that all the prefetch candidates of the prefetcher
+pass through the perceptron filter and if qualified, the metadata for
+perceptron indexing be stored. The second is needed when the feedback of a
+prior prefetch is available in form of a subsequent demand hit or cache
+eviction. In that case, the stored metadata needs to be retrieved to update
+the state of the perceptrons.
 
@@
   scalable, as it can easily learn to incorporate newer information in 
-  form of new features. Some of the features we developed use information 
+  the form of new features. Some of the features we developed use information 
   derived directly from program execution, agnostic to the baseline 

### 2018-12-04 design_impl.tex
spell check, editing for grammar, clarity, style, put in some questions preceded by 'djimenez'

@@
 derived from the information available in the L2 Cache access stream.
-Our feature selction involved searching over a large space of 
+Our feature selection involved searching over a large space of 
 relevant perceptron features. Using the statistical methodology 
@@
 
-% [EB] Tried addressing. Does it look reduntant when read after 'Feature Selection'
+% [EB] Tried addressing. Does it look redundant when read after 'Feature Selection'
 % bullet from the previous section?
@@
 \item \textbf{Base Address}: The base address of the demand 
-access that triggers the prefetch coresponds to a stream of 
+access that triggers the prefetch corresponds to a stream of 
 future accesses that SPP and the PPF have seen before. 
@@
 \item \textbf{Program Counter XOR Delta}: This feature tells us if a
-  given PC favours particular value(s) of delta.  As noted earlier,
+  given PC favors particular value(s) of delta.  As noted earlier,
   while the PC alone does not convey useful information, this
   hash resolves the PC into different values based on the tendency of
-  that PC to favour a certain delta.  Thus, the dynamic nature of
+  that PC to favor a certain delta.  Thus, the dynamic nature of
   different instances of the same memory access instruction is 
@@
  % that despite SPP having low confidence due to speculation, PPF may 
- % reinforce prefetches with lower confidence if SPP has had a tendancy to be 
+ % reinforce prefetches with lower confidence if SPP has had a tendency to be 
  % correct in the past.
@@
 % \item \textbf{Base address XOR Delta}: If any particular value(s) of
-%   delta are highly favoured by a given address which triggered the
+%   delta are highly favored by a given address which triggered the
 %   prefetch, this composite feature can capture that information
-% \item \textbf{Cache line XOR Delta}: qThis feature captures if any
-%   particular delta is highly favoured by a given cache line address
+% \item \textbf{Cache line XOR Delta}: this feature captures if any
+%   particular delta is highly favored by a given cache line address
 %\item \textbf{Confidence XOR Delta}: This gives a combined insight
-%  into that fact that certain delta-confidence pairqs predicted by
+%  into that fact that certain delta-confidence pairs predicted by
 %  SPP might correlate more with the prefetch outcome
@@
 choice of each feature by quantifying the contribution made towards
-predicting prefetch behaviour, in Section \ref{Method-Features}.
+predicting prefetch behavior, in Section \ref{Method-Features}.
 Finally, as noted above, each feature indexes into its independent

### 2018-12-04 intro.tex
Changes in Design. Minor edits in intro / method / results
spell check, editing for grammar, clarity, style, put in some questions preceded by 'djimenez'

@@
 but also in timing when data should be prefetched, deciding which level of the cache hierarchy
-the data should be stored in, and deciding what data should be evicted from the caches to accomodate
+the data should be stored in, and deciding what data should be evicted from the caches to accommodate
 the prefetched data.
@@
 As the look-ahead depth of the prefetch suggestions increase, so do useful prefetches. 
-But that comes at a cost of total prefetches inceasing at an even higher rate. This leads 
+But that comes at a cost of total prefetches increasing at an even higher rate. This leads 
 to cache pollution and bandwidth contention, as is clearly visible with the reduction in 
-the IPC. Similar behaviour was observed in 12 other SPEC 2017 simpoints.
+the IPC. Similar behavior was observed in 12 other SPEC CPU 2017 simpoints.
 
@@
 to do in the context of a multi-core processor, where overly aggressive prefetching in one core can
-waste shared resources, such as last-level cache capacity, and off-chip bandwidth~\cite{Friendly}.
+waste shared resources, such as last-level cache (LLC) capacity, and off-chip bandwidth~\cite{Friendly}.
 
@@
 %highly aggressive with high coverage. This causes the prefetcher to recommend many
-%low accurracy prefetches that lead to the cache being polluted with useless data.  
-%This effect is more pronounced in a multi-core scenario where the last-level cache is a
+%low accuracy prefetches that lead to the cache being polluted with useless data.  
+%This effect is more pronounced in a multi-core scenario where the LLC is a
 %shared resource~\cite{Friendly}.  Conversely, a conservative prefetcher would
@@
 In a single core configuration, PPF increases performance by 6.84\%
-compared to the baseline prefetcher, SPP.  In a multi-core system running a
+compared to the underlying prefetcher, SPP.  In a multi-core system running a
 mixes of memory intensive SPEC CPU 2017 traces, PPF saw an improvement

### 2018-12-04 method.tex
Changes in Design. Minor edits in intro / method / results
spell check, editing for grammar, clarity, style, put in some questions preceded by 'djimenez'

@@
 \label{Method-Model}
-For testing and comparing PPF, we used the ChampSim~\cite{Champsim} simulator.
+For testing and comparing PPF, we use the ChampSim~\cite{Champsim} simulator.
 ChampSim is an enhanced version of the framework that was used for the 2nd
@@
 
-The block size is fixed at 64 bytes. Prefetching is only triggered at an L2 cache
-demand accesses but could be directed to the L2 or last-level cache. No L1
-data level prefetching is done. The LRU replacement policy is used on all levels
-of cache hierarchies. Branch prediction is done using the perceptron branch
-predictor~\cite{Perc_Branch}.  The page size is 4KB.  ChampSim operates
+The block size is fixed at 64 bytes. Prefetching is only triggered at an L2
+cache demand accesses but could be directed to the L2 or last-level cache. No
+L1 data level prefetching is done. The LRU replacement policy is used on all
+levels of cache hierarchies. Branch prediction is done using the perceptron
+branch predictor~\cite{Perc_Branch}.  The page size is 4KB.  ChampSim operates
 all the prefetchers strictly in the physical address space.
@@
 \label{Method-Workloads}
-% djimenez: don't say this is the first time SPEC 2017 has been used this way.
+% djimenez: don't say this is the first time SPEC CPU 2017 has been used this way.
 % i don't know if that's true, and in any event the value is small.
 
-%This is the first time that SPEC 2017 benchmark suite~\cite{SPEC2017}
+%This is the first time that SPEC CPU 2017 benchmark suite~\cite{SPEC2017}
 %has been used to characterize and measure the prefetch performance.
-We use the SPEC 2017 benchmark suite~\cite{SPEC2017}.  We use all the 20
-workloads available in the SPEC 2017 suite.  Using the SimPoint~\cite{SimPoint}
+We use the SPEC CPU 2017 benchmark suite~\cite{SPEC2017}.  We use all the 20
+workloads available in the SPEC CPU 2017 suite.  Using the SimPoint~\cite{SimPoint}
 methodology, we identified 95 different program segments of 1 Billion
@@
 100 random mixes and another 100 mixes from the memory intensive subset of
-SPEC 2017.  For 4-core workload, 200 Million instructions are used for warm-up
+SPEC CPU 2017.  For 4-core workload, 200 Million instructions are used for warm-up
 and additional 1 Billion instruction simulated for collecting statistics.
@@
 \textit{Validation:} We cross-validated our PPF model using SPEC
-2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks.  For
-single-core SPEC 2006, we developed 94 simpoints spread across all the 29
-applications. For multi-core, we followed the same methodology as SPEC 2017.
+CPU 2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks.  For
+single-core SPEC CPU 2006, we developed 94 simpoints spread across all the 29
+applications. For multi-core, we followed the same methodology as SPEC CPU 2017.
 For CloudSuite, we used the traces made available for the 2nd Cache
@@
 traces with LLC MPKI > 1.  This includes 48 out of the 95 simpoints
-developed for SPEC 2017. For SPEC 2006  This includes 48 out of 94 simpoints.
+developed for SPEC CPU 2017. For SPEC CPU 2006, this includes 48 out of 94 simpoints.
 
-\subsection{Preferchers Simulated}
+\subsection{Prefetchers Simulated}
 \label{Method-Prefetchers}
@@
 prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern
-Matching (DA-AMPM)~\cite{DA_AMPM} and Signature Path Prefecher (SPP).  BOP was
+Matching (DA-AMPM)~\cite{DA_AMPM} and Signature Path Prefetcher (SPP).  BOP was
 the winner of 2nd Data Prefetching Championship.  DA-AMPM is the enhanced
 version of AMPM, modified to account for DRAM row buffer locality.  SPP has
-been shown to outperform BOP on SPEC 2006 traces.  For each of these, we
+been shown to outperform BOP on SPEC CPU 2006 traces.  For each of these, we
 compare their speedups taking no prefetching as the baseline.
@@
 dump of the state of perceptron weights at the end of all trace execution.
-The weights obtained from running all the SPEC 2017 traces are concatenated.
+The weights obtained from running all the SPEC CPU 2017 traces are concatenated.
 Since the weights are collected at the end of individual trace execution, the
@@
 this dump represents a `snap' of the trained perceptron weights across all the
-SPEC 2017 traces.  The intuition here is that the feature with bulk of the
+SPEC CPU 2017 traces.  The intuition here is that the feature with bulk of the
 perceptron weights concentrated around 0 or small magnitude numbers show a
@@
 \textit{Signature XOR Delta} and 
-\textit{PC XOR Depth}; across all the SPEC 2017 traces.  
+\textit{PC XOR Depth}; across all the SPEC CPU 2017 traces.  
 For simplicity, the traces are arranged in an
@@
 \textbf{Trimming Features Using Correlation} \newline Besides
-providing interesting insights into prefetching behaviour, P-value can
+providing interesting insights into prefetching behavior, P-value can
 also be used for feature selection and prefetcher tuning.
@@
 and \textit{page address}  were given most importance and
-allowed full 12-bits of indexing.  Features like \textit{ip XOR delta} 
-and \textit{ip XOR depth} with a low overall P-value 
+allowed full 12-bits of indexing.  Features like \textit{IP XOR delta} 
+and \textit{IP XOR depth} with a low overall P-value 
 were allocated fewer indexes in the feature table.

### 2018-12-04 related.tex
spell check, editing for grammar, clarity, style, put in some questions preceded by 'djimenez'

@@
 The idea of using concepts from machine learning to drive cache management mechanism has become widely popular.
-Perceptrons in microarchitecture were first introduced by Jimenez and Lin ~\cite{Perc_Branch} for branch prediction.
+Perceptrons in microarchitecture were first introduced by Jim\'enez and Lin ~\cite{Perc_Branch} for branch prediction.
 Their predictor consists of a table of weights that was indexed by the program counter associated to the branch.
@@
 %%MOVE IT SOMEWHERE
-Peled \textit{et.al.} proposed using on-line reinforcement learnign to dynamically scale the maginitude of input information feed into their prefetcher mechanism ~\cite{Semantics}.
+Peled \textit{et.al.} proposed using on-line reinforcement learning to dynamically scale the magnitude of input information feed into their prefetcher mechanism ~\cite{Semantics}.
 The major drawback of this work is the need of compiler support.

### 2018-12-04 results.tex
Completing Methodology, Images
Changes in Design. Minor edits in intro / method / results
spell check, editing for grammar, clarity, style, put in some questions preceded by 'djimenez'

@@
-\begin{figure}[h]
+\begin{figure}[ht]
 \includegraphics[width=\columnwidth]{SPEC2017}
-\caption{SPEC 2017 Single-Core IPC Speedup}
+\caption{SPEC CPU 2017 Single-Core IPC Speedup}
 \label{Fig:SPEC2017_1core}
@@
 
-This section discusses the results obtained from running PPF, in
-terms of prefetch cache coverage and Instruction Per
-Cycle (IPC).  For SPEC 2017 suite, first we present the
-results for single-threaded workloads then for multi-core
+This section discusses the results obtained from running PPF in terms of
+prefetch cache coverage and speedup. For the SPEC CPU 2017 benchmarks, first
+we present the results for single-threaded workloads then for multi-core
 workloads.
@@
 
-Figure~\ref{Fig:SPEC2017_1core} shows the IPC speedup obtained by
-PPF, compred to BOP, DA-AMPM and SPP. 
-In the figure we have depicted selected traces, where PPF shows a 
-noticeable improvement or shows some degradation, along with the 
-geometric mean improvement on the memory intensive subset of SPEC 2017, 
-and finally for the complete SPEC 2017 benchmark. 
-All the results have been normalized to baseline \textit{i.e.}, no
-prefetching.
+% djimenez: why is this figure described this way?  it's not OK to just
+% highlight noticeable improvement. the benchmarks must be selected by some
+% unbiased criteria, e.g. some minimum MPKI under LRU or some minimum speedup
+% under a baseline prefetcher that isn't PPF.
 
-As can be seen, PPF yields a speedup of \textbf{42.8\%} over
-the baseline. This is equivalent to \textbf{7.8\%} over DA-AMPM, \textbf{9.7\%} over BOP
-and \textbf{6.92\%} over SPP.  Out of the 95 simpoints developed for SPEC 2017, 
-PPF nearly matches or outperforms most of the prefetchers on 91 simpoints.
-Only in 607.cactuBSSN and 654.roms, PPF fails to match the improvement 
-offered by BOP.
+Figure~\ref{Fig:SPEC2017_1core} shows the IPC speedup obtained by PPF,
+compared to BOP, DA-AMPM and SPP.  In the figure we have depicted selected
+traces where PPF shows a noticeable improvement or shows some degradation,
+along with the geometric mean improvement on the memory intensive subset of
+SPEC CPU 2017, and finally for the complete SPEC CPU 2017 benchmark.  All the
+results have been normalized to the baseline of no prefetching.
 
-At its peak, PPF manages to get the IPC speedup of a factor of
-over \textbf{3.39x} on the trace \textit{602.gcc\_s-2226B}.  This also
-corresponds to speedup gain of \textbf{60\%} over the next best
-prefetcher -- BOP. 
-In general, benchmarks \textit{bwaves, mcf, xalancbmk} 
-and \textit{fotonik3d} benefit the most from PPF, with the speedup over
-SPP ranging from \textbf{10\% to 25\%}.
+% djimenez: you've inconsistently italicized and not italicized benchmark
+% names. i'll help you out here; i like putting them in a typewriter font so
+% i'll do that throughout this section. also, you've inconsistent put and not put
+% the number of the benchmark. i'll fix that too.
 
-On the full SPEC 2017 suite, PPF improves the baseline by \textbf{20.9\%},
-which is \textbf{3.44\%} more than the next best prefetcher -- SPP.
+PPF yields a speedup of \textbf{42.8\%} over the baseline. This is equivalent
+to \textbf{7.8\%} over DA-AMPM, \textbf{9.7\%} over BOP and \textbf{6.92\%}
+over SPP.  Out of the 95 simpoints developed for SPEC CPU 2017, PPF nearly
+matches or outperforms most of the prefetchers on 91 simpoints.  PPF fails to
+match the improvement offered by BOP only for {\tt 607.cactuBSSN\_s} and {\tt
+654.roms\_s}.
+
+At its peak, PPF manages a speedup of a factor of over \textbf{3.39x} on the
+trace {\tt 602.gcc\_s-2226B}.  This also corresponds to speedup gain of
+\textbf{60\%} over the next best prefetcher -- BOP.  In general, benchmarks
+{\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt 623.xalancbmk\_s} and {\tt
+649.fotonik3d\_s} benefit the most from PPF, with the speedup over SPP ranging
+from \textbf{10\% to 25\%}.
+
+% djimenez: is this geometric mean? if so, say so.
+On the full SPEC CPU 2017 suite, PPF improves the baseline by \textbf{20.9\%},
+which is \textbf{3.44\%} better than the next best prefetcher -- SPP.
 
@@
 
+% djimenez: this is a weird way to make a header for a section. use \paragraph or \subsection
+
 \textbf{Coverage}\newline
-A prefetcher coverage can be defined as the ratio of number of misses 
-avoided through prefetching to the number of misses in presence of no 
-prefetchig.
-Figure~\ref{Fig:Coverage} shows the fraction of misses in L2 and Last
-Level Caches avoided by the various prefetchers.
-PPF has the highest coverage of all the prefetchers simulated. Using PPF 
-on SPEC 2017 benchmark, 
-misses can be reduced by \textbf{62.5\%} and \textbf{82.8\%} in L2 Cache and Last Level 
-Cache respectively. For the same benchmark, the next best prefetcher -- 
-DA-AMPM covers \textbf{48.6\%} and \textbf{72.8\%} of the misses respectively.
+Prefetcher coverage is defined as the ratio of the number of misses avoided
+through prefetching to the number of misses with no prefetching.
 
-This superior coverage of PPF can be attributed to aggressive 
-re-tuning of the underlying SPP. At the same time, the Perceptron Filter
-makes sure than this high coverage does not lead to an increased 
-cache pollution.
+Figure~\ref{Fig:Coverage} shows the fraction of misses in the L2 and LLC
+avoided by the various prefetchers.  PPF has the highest coverage of all the
+prefetchers simulated. On the SPEC CPU 2017 benchmarks, PPF reduces misses by
+\textbf{62.5\%} and \textbf{82.8\%} in the L2 and LLC respectively. For the
+same benchmark, the next best prefetcher, DA-AMPM, covers \textbf{48.6\%} and
+\textbf{72.8\%} of the misses respectively.
+
+This superior coverage of PPF can be attributed to aggressive re-tuning of the
+underlying SPP, enabled by the Perceptron Filter making sure the high coverage
+does not lead to increased cache pollution.
 
@@
 \label{Results-Multi}
-In this section, we demonstrate the improvement achieved by PPF
-for a mix of multi-programmed workloads.
+In this section, we demonstrate the improvement achieved by PPF for a mix of
+multi-programmed workloads.
 
-\textit{4-core environment} Figure~\ref{Fig:4Core_SPEC2017} (a) 
-shows a comparison of speedups
-obtained on 4-core mix of memory intensive subset of SPEC 2017.  We
-plot all the 4 prefetchers, normalized to the baseline.  The workloads
-have been sorted in the increasing order of the speedup.
-As can be seen,
-PPF offers an IPC improvement by \textbf{51.7\%} on these traces.  
-Compared to the base prefetcher -- SPP only, this is \textbf{11.9\%} more.
-As compared to the next best prefetcher -- DA-AMPM, this is 
-This is \textbf{10.2\%} and \textbf{17.4\%} over DA-AMPM, 
-the next best prefetcher, and BOP, respectively.
+\textit{4-core environment} Figure~\ref{Fig:4Core_SPEC2017}(a) shows a
+comparison of speedups obtained on 4-core mixes of a memory intensive subset
+SPEC CPU 2017.  We plot all 4 prefetchers, normalized to the baseline.  The
+workloads have been sorted in increasing order of the speedup.  PPF offers a
+speedup of \textbf{51.7\%} on these traces, and improvement of \textbf{11.9\%}
+over the baseline SPP, \textbf{10.2\%} over the next DA-AMPM, and
+\textbf{17.4\%} over BOP.
 
-On a different set of fully random SPEC 2017 4-core mixes, 
-PPF provides an IPC speedup of 
-\textbf{26.2}\% over baseline, which is \textbf{5.7}\% more than just SPP.
+On a different set of fully random SPEC CPU 2017 4-core mixes, PPF provides an
+IPC speedup of \textbf{26.2}\% over the baseline, which is an improvement of
+\textbf{5.7}\% over SPP.
 
@@
 
-\textit{8-core environment} The sorted comparison of speedup obtained 
-on the memory intensive 8-core mixes is depicted in Fig~\ref{Fig:8Core_SPEC2017}.
-PPF improves baseline performance by \textbf{38.6\%} and that is \textbf{10.7\%} 
-more than using just SPP. 
-For completely random set of SPEC 2017 mixes, PPF improves performance by 
-\textbf{23.6\%} over the baseline, corresponding to \textbf{4.8\%} over SPP.
-This increased improvement achieved by PPF over the base engine SPP,
-in a multi-core environment is not surprising. 
-Since PPF has an extremely intelligent filter, it ends up 
-eliminating useless prefetches before they can create cache
+\textit{8-core environment} The sorted comparison of speedups on the memory
+intensive 8-core mixes is shown in Figure~\ref{Fig:8Core_SPEC2017}.  PPF
+improves baseline performance by \textbf{38.6\%}, an improvement of
+\textbf{10.7\%} over SPP.  For a random set of SPEC CPU 2017 mixes, PPF
+improves performance by \textbf{23.6\%} over the baseline, corresponding to
+\textbf{4.8\%} over SPP.  This increased improvement achieved by PPF over the
+base engine SPP in a multi-core environment is expected as PPF has a very
+accurate filter, it eliminates useless prefetches before they can cause
 pollution in the shared LLC.
 
-As can be seen, BOP offer a better improvement than 
-SPP, for the memory intensive mix, This can be attributed 
-to BOP's inherent aggressive nature. DA-AMPM is also ahead of
-SPP in both the mixes. Interestingly, in all these cases, 
-PPF consistently outperforms the best performing prefetcher.
+BOP offers a better improvement than SPP for the memory intensive mixes. This
+superiority can be attributed to BOP's inherent aggressive nature.  DA-AMPM is
+also ahead of SPP in both the mixes. Interestingly, in all these cases, PPF
+consistently outperforms the best performing prefetcher.
 
@@
 
-Figure \ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a 
-reduced LLC and low bandwidth constraints respectively. For the sake of brevity, 
-we have only shownselected traces, followed
-by the overall performance on the complete benchmark.
-As can be seen, \textit{mcf} in low bandwidth conditions is prefetch averse.
-In general, any prefetcher yields a negative speedup on that trace.
-On \textit{roms} and \textit{cactuBSSN}, PPF is unable to match the performance 
-achieved by the best prefetcher. On the other hand, PPF outperforms all the 
-other prefetchers on \textit{xalancbmk} and \textit{imagick} benchmarks.
-Overall, PPF gives a better improvement under small LLC condition and 
-matches the best prefetcher -- BOP under low DRAM bandwidth condition.
+Figure~\ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a
+reduced LLC and with low bandwidth constraints, respectively. For the sake of
+brevity, we have only shown selected traces followed by the overall
+performance on the complete benchmark.  Benchmark {\tt 605.mcf\_s} in low
+bandwidth conditions is prefetch averse.  In general, any prefetcher yields a
+negative speedup on that trace.  On {\tt 654.roms\_s} and {\tt
+607.cactuBSSN\_s}, PPF is unable to match the performance achieved by the best
+prefetcher. On the other hand, PPF outperforms all the other prefetchers on
+{\tt 623.xalancbmk\_s} and {\tt 638.imagick\_s} benchmarks.  Overall, PPF gives
+a better improvement under small LLC condition and matches the best
+prefetcher, BOP, under low DRAM bandwidth conditions.
 
@@
 \includegraphics[width=\columnwidth]{SPEC2006}
-\caption{SPEC 2006 Single-Core IPC Speedup}
+\caption{SPEC CPU 2006 Single-Core IPC Speedup}
 \label{Fig:SPEC2006_1core}
@@
 
-Here we try to demonstrate the robustness of PPF by testing it on
-completely unseen benchmarks, namely SPEC 2006 and CloudSuite.
+% djimenez: cite cloudsuite here? do we have a reference?
 
-Fig~\ref{Fig:SPEC2006_1core} shows the speed-up achieved by BOP, SPP and PPF on
-selected traces of SPEC 2006, along with the overall geometric mean
-over whole of the benchmark and over memory-intensive subset of the
-SPEC 2006 benchmark.  As can be seen, PPF provides a speed-up of
-\textbf{46.1\%} over baseline, on the memory intensive subset of SPEC 2006
-benchmark. This corresponds to \textbf{6.66\%} over SPP and \textbf{9.86\%} over BOP and DA-AMPM.  
-On the whole of the SPEC 2006 suite, the speedup obtained is \textbf{22.4\%},
-which is \textbf{3.34\%} ahead of SPP.
+Here we demonstrate the robustness of PPF by testing it on different
+benchmarks: SPEC CPU 2006 and CloudSuite.
 
-For 4-core memory intensive mixes, PPF improves the baseline by \textbf{59.2\%}, 
-\textbf{8.7\%} ahead of SPP. For 8-core memory intensive mixes, the IPC speedup 
-over baseline is \textbf{47.9\%}, \textbf{11.5\%} ahead of SPP.
+% djimenez: is the improvement over BOP and DA-AMPM exactly the same 9.86%
+% here? that's weird.
+
+Figure~\ref{Fig:SPEC2006_1core} shows the speed-up achieved by BOP, SPP and
+PPF on traces from SPEC CPU 2006, along with the overall geometric mean over
+the whole suite and over a memory-intensive subset.  PPF provides a speedup of
+\textbf{46.1\%} over the baseline on the memory intensive subset of SPEC CPU
+2006 benchmark, giving an improvement of \textbf{6.66\%} over SPP and
+\textbf{9.86\%} over BOP and DA-AMPM.  On the whole of the SPEC CPU 2006
+suite, the speedup is \textbf{22.4\%}, an improvement of \textbf{3.34\%} over
+SPP.
+
+For 4-core memory intensive mixes, PPF improves the baseline by
+\textbf{59.2\%}, \textbf{8.7\%} ahead of SPP. For 8-core memory intensive
+mixes, the speedup over the baseline is \textbf{47.9\%}, \textbf{11.5\%} ahead
+of SPP.
 
@@
 the prefetch schemes on 4 different applications in the CloudSuite benchmark.
-In general, these applications are prefetch agnostic. Even in that scenario, 
-PPF manages a \textbf{3.7\%} improvement over no prefetching, making it ahead of the 
-next best prefetcher -- SPP, which provides a 3.08\% IPC speedup.
+In general, these applications are prefetch agnostic. Even so, PPF manages a
+\textbf{3.7\%} improvement over no prefetching, putting it ahead of the next
+best prefetcher, SPP, which provides a 3.08\% speedup.
 
-As can be seen, the performance is very consistent given no prior
-parameter tuning was done keeping these benchmarks in mind.  We attribute
-this to the inherent adaptability of the perceptron model.  In
-general, perceptron weights are able to adjust in real-time so as to
-find the best possible correlation between the output and the given set of 
+We developed PPF to yield good performance on the SPEC CPU 2017 benchmarks.
+Nevertheless, the performance is consistently good on other benchmark suites.
+We attribute this fact to the inherent adaptability of the perceptron model.
+In general, perceptron weights are able to adjust in real-time so as to find
+the best possible correlation between the output and the given set of
 features.

### 2018-12-04 back.tex
detail about perc learning rule

@@
 %learning rule will be applied to PPF.
+
+%ETG: this is what I had added about perceptrons, we're trying to 
+%figure out what to leave and what to get rid of ...  
+\subsection{Perceptron-based Prediction}
+
+Perceptrons in microarchitecture were first introduced by Jimenez and Lin ~\cite{Perc_Branch} for branch prediction.
+In this work, a perceptron weights vector is selected by a hash of the branch address and its dot product with an input vector of branch history outcomes.
+If the product is at least 0, the branch is predicted as taken, otherwise it is predicted not taken.
+The weights are updated with the perceptron update rule: if the prediction is incorrect, or the dot product does not exceed some threshold magnitude, then the weights are incremented or decremented based on their correlation to the corresponding history bits.
+Perceptron learning yields to superior accuracy, which results in increased opportunity for optimizations.
+
+\subsubsection{Perceptron Learning}
+
+In a perceptron-based mechanism, a uniform
+perceptron update principle is followed.  Perceptrons are a simple model of neurons
+in neural networks \cite{perceptrons1,perceptrons2} modeled by vectors of signed weights learned through online training.
+The output of a perceptron is the dot product of the weights and a vector of inputs.
+In this work, we make use of the perceptron learning algorithm.
+There are two components to the abstract perceptron learning algorithm:
+
+\begin{enumerate}
+\item To predict true or false, a vector of signed weights is chosen according to some criteria.
+The dot product of the weights and input vectors, called $y_{\mbox{\small out}}$, is computed.
+If $y_{\mbox{\small out}}$ exceeds some threshold, the prediction is true, otherwise it is false.
+
+\item To update the weights after the true outcome of the predicted event is
+known, we first consider the value of $y_{\mbox{\small out}}$. If the
+prediction was correct and $|y_{\mbox{\small out}}|$ exceeds a threshold
+$\theta$, then the weights remain unchanged. Otherwise, the inputs are used to
+update the corresponding weights.  If there is positive correlation between
+the input and the outcome of the event, the corresponding weight is
+incremented; otherwise, it is decremented. Over time, the weights are
+proportional to the probability that the outcome of the event is true in the
+context of the input and the criterion for choosing that weight. The weights
+saturate at maximum and minimum values so they will fit in a fixed bit width.
+\end{enumerate}

### 2018-12-04 related.tex
related work

--- 2018-12-03 related.tex
@@
 
-The idea of using concepts from machine learning to drive cache management mechanism has become widely popular.
-Perceptrons in microarchitecture were first introduced by Jimenez and Lin ~\cite{Perc_Branch} for branch prediction.
-Their predictor consists of a table of weights that was indexed by the program counter associated to the branch.
-The dot product of the corresponding weight and a history of branches is thresholded against a magnitude to make a prediction.
-The perceptron-based predictor helped differentiate which branch history outcomes were relevant for marking a prediction for a certain branch, which yield higher accuracy than previously proposed branch predictors.
-This reason motivated the use of perceptrons in other speculation mechanisms.
+\subsection{Perceptrons in Cache Management}
 
-To drive their replacement policy, Teran \textit{et.al} proposed using a perceptron-based predictor that obeys by the perceptron learning rule.
-By adding this way of reinforcement learning, their mechanism avoids overtraining and can rapidly adapt to changes in memory behavior.
-The continuation of this work was later introduced ~\cite{Multiperspective}, where a genetic algorithm was used to find an extensive set of features to be used as an input to the predictor.
-The extra features were able to cover more of the memory access patterns.
-The outcomes of these predictors are used to drive optimizations such a replacement, placement and bypass of the LLC.
+Perceptron-based prediction mechanisms have proven to yield superior accuracy, which represents an opportunity for optimizations in cache management. 
+Khan \textit{et.
+al.} proposed SDBP \cite{sdbp}, in this work, a sampler structure keeps partial tags of sampled sets separate from the cache.
+Three tables of two-bit saturating counters are accessed using a technique similar to a skewed branch predictor \cite{Piece_Linear}.
+For each hit to a block in the sampled set, the program counter (PC) of the relevant memory instruction is hashed into the three tables and the corresponding counters are decremented.
+For each eviction from a sampled set, the counters corresponding to the PC of the last instruction to access the victim block are incremented.
+For an LLC access, the predictor is consulted by hashing the PC of the memory access instruction into the three tables and taking the sum of the indexed counters.
+When the sum exceeds some threshold, the accessed block is predicted to be dead.
+Tags in sampled sets are managed with true LRU and a reduced associativity, but the LLC may be managed by any policy.
+The paper applies its predictor to replacement and bypass.
 
-%%MOVE IT SOMEWHERE
-Peled \textit{et.al.} proposed using on-line reinforcement learnign to dynamically scale the maginitude of input information feed into their prefetcher mechanism ~\cite{Semantics}.
-The major drawback of this work is the need of compiler support.
-SVMs and logistic regression were used by Liao \textit{et.al.} to do an off-line optimal feature search for their prefetcher.
+Teran \textit{et al.} propose using perceptron learning for reuse prediction driving a bypass and replacement optimization \cite{Perc_Reuse}.
+In that work, perceptron learning is used to set weights selected by hashes of multiple features consisting of the PC, several recent PCs, and two different shifts of the tag of the referenced block.
+Each feature is hashed to index its own table of weights.
+The selected weights are summed and thresholded to make a prediction.
+When a sampled block is reused or evicted, the corresponding weights are decremented or incremented, respectively, according to the perceptron learning rule.
+
+The continuation of Perceptron was later introduced in the work, Multiperspective Reuse Prediction \cite{Multiperspective}.
+Compared to previous work that focuses on one or two features at a time, Multiperspective prediction uses many features, each contributing to the overall prediction.
+The prediction enables placement, promotion and bypass optimizations that improves performance.
+
+\subsection{Base Prefetchers}
+
+Two widely used prefetchers are ?Next-n Lines?
+\cite{nextnlines} and Stride [4], both of which capture regular memory access patterns with very low hardware overhead.
+The ?Next-n Lines?
+prefetcher simply queues prefetches for the next n lines after any given miss, under the expectation that the principle of spatial locality will hold and those cache lines after a missed line are likely to be used in the future.
+The stride prefetcher is slightly more sophisticated; it attempts to identify simple stride reference patterns in programs based upon the past behavior of missing loads.
+Similar to the ?Next-n Lines?
+technique, when a given load misses, cache lines ahead of that miss are fetched in the pattern following the previous be- havior in the hope of avoiding future misses.
+Both prefetch- ers have been widely used due to the simple design and low hardware overhead.
+However, without further knowledge about temporal locality and application characteristics, these prefetchers cannot do more than detecting and prefetching regular memory access patterns with limited spatial locality.
+
+Somogyi et al.
+proposed one of the current top- performing, practical, low-overhead prefetchers, the Spatial Memory Streaming (SMS) prefetcher [23].
+SMS leverages code-based correlation to take advantage of spatial locality in the applications over larger regions of memory (called spatial regions).
+It predicts the future access pattern within a spatial region around a miss, based on a history of access patterns initiated by that missing instruction in the past.
+While the SMS prefetcher is effective, it is indirectly inferring future program control-flow when it speculates on the misses in a spatial region.
+As a result, the state overheads of this predictor can be higher than the others in this class.
+
+\subsection{Lookahead Prefetchers}
+
+%ANMP
+
+While previous work had proposed dynamically finding the best prefetch offset, they have failed to look into prefetch timeliness.
+Michaud proposed the Best-Offset prefetcher \cite{BOP}.
+A prefetch offset is set automatically and dynamically tries to adapt to the application behavior.
+A good prefetch offset is determined by looking at two recent access to a particularly line that do not happen too recently.
+The time between both accesses should be greater than the latency for completing a prefetch request.
+
+\subsection{Aggressive Prefetching}
+
+A low-accuracy aggressive prefetcher can significantly harm performance.
+In order to minimize the interference from prefetching, Wu \textit{et.
+al} proposed PACMan ~\cite{pacman}, a prefetch-aware cache management policy.
+PACMan dedicated a few sets of the LLC to each of the three competing policies that treat demand and prefetch request differently and uses the policy that shows the lowest number of cache misses.
+Similarly, Seshadri  \textit{et.
+al.} propose ICP \cite{icp}.
+ICP is also designed as a comprehensive mechanism to mitigate prefetching interference.
+ICP simply demotes a prefetch blow to the lowest priority on a demand hit based on the observation that most prefetches are dead after their first hit.
+To address prefetcher-caused cache pollution, it also uses a variation of EAF [27] to track prefetching accuracy and inserts only an accurate prefetch to the higher priority position in the LRU stack.
+ICP assumes, however, that all prefetches are inserted only into the LLC, which restricts the maximum benefit of prefetching.

### 2018-12-04 design_arch.tex
Updated 3.1 to be centered around figure discussion

@@
 % talking about tables!
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=9cm]{PPF_Hierarchy.pdf}
+    \label{fig:PPF_Hierarchy}
+  \caption{PPF Architecture in the Memory Hierarchy}
+  \end{center}
+\end{figure}
 
@@
 is placed between the prefetcher and the prefetch insertion queue, to ensure
-that predicted bad prefetches are not inserted into the memory system.
-\pg{As shown in figure whatever...}.
+that predicted bad prefetches are not inserted into the memory system as shown in 
+figure~\ref{fig:PPF_Hierarchy}.
 
@@
 
-% Gino: Are these modifications minimal?
-% EB: That's a good question :P. Edited to 'appropriate'.
-
 \subsection{The Perceptron Filter}
@@
 
-%Gino: rephrase below?
-The Perceptron Filter is organized as tables of perceptron weights, with a
-separate table of weights for each feature. For a design using `N' number of
-features, its weights will be arranged in $N$ different tables. Each feature's
-value is used to index into the corresponding table.  Since each table is
+The Perceptron Filter and the steps used to performed inference 
+and training are shown in figure~\ref{fig:PPF_Datapath}. The filter is organized as tables of 
+perceptron weights, with a separate table of weights for each feature. 
+For a design using `N' number of features, its weights will be arranged
+in $N$ different tables. Each feature's value is used to index 
+into the corresponding table with $K$ entries.  Since each table is
 indexed independently, different numbers of bits can be used for indexing,
@@
 our PPF implementation are given under ``Area Overheads'' in
-Section~\ref{Method-Overheads}.
-
-%\pg{this whole ``bits'' discussion belongs in implementation not here IMHO}
-%[EB] Resolved
+Section~\ref{sec:Method-Overheads}.
 
@@
 is a 5 bit counter saturating at -16 and +15 and initialized to zero at the
-beginning of a program's execution. We observed empirically that 5 bits of
-index achieves a good trade-off between hardware overheads and accuracy.
+beginning of a program's execution. We empirically observe that 5 bits of
+index achieves a good trade-off between hardware overhead and accuracy.\newline
 
-%\pg{the fact that its nine is again implementation details}
-% [EB] Resolved
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=9cm]{Datapath.pdf}
+    \label{fig:PPF_Datapath}
+  \caption{PPF Data Path and Operation}
+  \end{center}
+\end{figure}
 
-% \pg{The word ChampSim should not appear in this part of the paper at
-%   all.  Remember you are describing how to build this in the general
-%   case...} 
-% [EB] Resolved
-\textbf{Inference}\newline The prefetcher is triggered on every L2 cache
+\noindent{\textbf{1. Inference}}\newline
+
+The prefetcher is triggered on every L2 cache
 demand access.  At that point, the base prefetcher has the option to perform a
@@
 When the base prefetcher is triggered, it begins to suggest candidates for
-prefetching.  All suggested prefetches are passed through the perceptron
+prefetching.  All suggested prefetches pass through the perceptron
 filter to decide whether they are qualified for final prefetching.  The filter
@@
 
-Once all the weights are retrieved, they are summed and compared to a preset
-threshold, PERC\_ THRESHOLD\_LO.  Only prefetch candidates with perceptron sum
-higher than the threshold qualify for prefetching.  The prefetches that
-qualify through the perceptron stage are recorded in the ``Prefetch Table''.
-The prefetch table is a 1,024-entry, direct mapped structure that contains all
-metadata required to re-index the perceptron entries for training.  When the
-feedback of the current prefetched line is available at a later stage, the
-stored data is used to train the perceptron.
+Each feature's confidence contributes to 
+the final decision of whether to prefetch a cache line, with the summation 
+of all features' confidences denoting the overall confidence for a prefetch.  
+The final sum is compared against two thresholds: PERC\_THRESHOLD\_HI and  
+PERC\_ THRESHOLD\_ LO to decide the placement of the data in the cache 
+hierarchy. Prefetches that have a confidence above the higher threshold 
+are placed into the L2 cache since the observed features strongly recommend  
+the prefetch. A prefetch that results in a confidence between 
+PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI are placed in the larger 
+last level cache, as the filter is moderately confident of the cache line's 
+use, but seeks to reduce pollution in the smaller L2 cache. Finally, the 
+remaining prefetches that are below the lower threshold are not prefetched 
+due to a lack of correlation between the current observed features and 
+the prediction from the base prefetch engine.\newline
 
-%\pg{this discussion below has redundancy and doesn't flow well.
-%  Rework it}
+\noindent{\textbf{2. Recording}}\newline
 
-Each feature's confidence contributes to the final decision of whether to
-prefetch a cache line, with the summation of all features' confidences
-denoting the overall confidence for a prefetch.  The final sum is compared
-against two thresholds: PERC\_THRESHOLD\_HI and  PERC\_ THRESHOLD\_ LO to
-decide the placement of the data in the cache hierarchy. Prefetches that have
-a confidence above the higher threshold are placed into the L2 cache since the
-observed features strongly recommend  the prefetch. A prefetch that results in
-a confidence between PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI are
-placed in the larger last level cache, as the filter is moderately confident
-of the cache line's use, but seeks to reduce pollution in the smaller L2
-cache. Finally, the remaining prefetches that are below the lower threshold
-are not prefetched due to a lack of correlation between the current observed
-features and the prediction from the base prefetch engine.
+The prefetches that qualify through the inference stage are recorded in the 
+``Prefetch Table''. The prefetch table is a 1,024-entry, direct mapped 
+structure that contains all metadata required to re-index the perceptron 
+entries for training.
 
-%GAC: The section's previous contents are below, in case the above 
-%rewritten work is not what we were looking for in terms of a rewording
-%
-%The perceptron sum can be considered the sum of the individual
-%contribution of a feature.  The value of each contribution corresponds
-%to a feature's confidence in the final decision on whether to prefetch
-%a cache line. By summing the individual contributions, the final
-%perceptron sum denotes the overall confidence for that prefetch
-%suggestion.  By comparing the perceptron sum to two different
-%thresholds, we divide the confidence scale of the prefetch suggestion
-%into three bins. The first bin corresponds to the lowest confidence
-%and leads to prefetch candidate being rejected.  The next bin
-%corresponds to prefetches with a moderate confidence level.  This bin
-%prefetches directly to fill the bigger last level cache and prevents
-%polluting the smaller L2 Cache.  The highest confidence prefetches are
-%filled in the L2 cache.
-%
-%
-%The sum of the perceptron weights are also used to decide whether to
-%place the data in the L2 or the last level cache.  All prefetch
-%candidates that are qualified until this stage are compared against
-%PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
-%PERC\_THRESHOLD\_\_LO and PERC \_THRESHOLD\_HI are empirically set.
+In addition to the prefetch table mentioned above, 
+PPF also maintains a 1,024-entry direct-mapped ``Reject Table.'' 
+If a prefetch suggestion is rejected by the perceptron layer, it is logged into 
+the reject table. The table is used to train the perceptron to avoid false negatives 
+\textit{i.e.}, cases where prediction was to reject the prefetch but 
+the prefetch would have been useful based on the observed demand accesses to the L2. \newline
 
-In addition to the prefetch table mentioned above, PPF also maintains a
-``Reject Table.''  The reject table is a 1,024-entry direct-mapped table.
-If a prefetch suggestion is rejected by the perceptron layer, it is logged
-into the reject table.  The table is used to train the perceptron to avoid
-false negatives \textit{i.e.}, cases where prediction was to reject the
-prefetch but the prefetch would have been useful.
+\noindent{\textbf{3. Feedback and Data Retrieval}} \newline
 
-%As previously stated all the metadata describing the program state at the
-%instant of prefetching is stored in the Prefetch Table or the Reject Table.
-%This information is useful when the perceptron needs to be updated
-%subsequently.
+In the prefetching environment, feedback for a prefetch is received whenever 
+there is an eviction or a demand access from the L2 cache. This action 
+triggers training of the base prefetcher as well as the perceptron layer.  
+The cache line's address of the block being trained is used to access either the Prefetch
+or Reject table, using 10 bits of the address for indexing and another 6 bits to 
+perform tag matching.  Once the values of the features at the time of prefetching is available,
+it is used to index into the perceptron weights table. \newline
 
-\textbf{Training}\newline In the prefetching environment, feedback for a
-prefetch is received whenever there is an eviction or a demand access from the
-L2 cache.  This action triggers training of the base prefetcher as well as the
-perceptron layer.  Training the perceptrons involves accessing the metadata
-that was stored in Prefetch Table or the Reject Table.  The cache lines
-address of the block being trained is used to access a table, using 10 bits of
-the address for indexing and another 6 bits to perform tag matching.  Once the
-values of the features at the time of prefetching is available, it is used to
-index into the perceptron weights table.
+\noindent{\textbf{4. Training}}\newline
+
+Training the perceptrons involves accessing the metadata
+that was stored in Prefetch Table or the Reject Table.
 
@@
 the given program behavior. These thresholds are referred to as
-POS\_UPDATE\_THRESHOLD ($\theta_p$) and NEG\_UPDATE\_THRESHOLD ($\theta_n$),
+POS\_UPDATE\_THRESHOLD ($\theta_p$) and \newline
+NEG\_UPDATE\_THRESHOLD ($\theta_n$),
 respectively for the positive and negative values of training saturation.
@@
 line.  Our two-step PPF architecture allows us to overcome that issue.
+
+%Gino: rephrase below?
+%The Perceptron Filter is organized as tables of perceptron weights, with a
+%separate table of weights for each feature. For a design using `N' number of
+%features, its weights will be arranged in $N$ different tables. Each feature's
+%value is used to index into the corresponding table.  Since each table is
+%indexed independently, different numbers of bits can be used for indexing,
+%thus giving each table a different size.  The variable indexing can be
+%determined by studying the features and fine-tuning them empirically to
+%achieve a good accuracy versus hardware overhead trade-off.  Exact details of
+%our PPF implementation are given under ``Area Overheads'' in
+%Section~\ref{Method-Overheads}.
+%
+%%\pg{this whole ``bits'' discussion belongs in implementation not here IMHO}
+%%[EB] Resolved
+%
+%A single entry in the table corresponds to a perceptron weight.  Each weight
+%is a 5 bit counter saturating at -16 and +15 and initialized to zero at the
+%beginning of a program's execution. We observed empirically that 5 bits of
+%index achieves a good trade-off between hardware overheads and accuracy.
+%
+%%\pg{the fact that its nine is again implementation details}
+%% [EB] Resolved
+%
+%% \pg{The word ChampSim should not appear in this part of the paper at
+%%   all.  Remember you are describing how to build this in the general
+%%   case...} 
+%% [EB] Resolved
+%
+%\begin{figure}
+%  \begin{center}
+%  \includegraphics[width=9cm]{Datapath.pdf}
+%    \label{fig:PPF_Datapath}
+%  \caption{PPF Data Path and Operation}
+%  \end{center}
+%\end{figure}
+%
+%\textbf{Inference}\newline The prefetcher is triggered on every L2 cache
+%demand access.  At that point, the base prefetcher has the option to perform a
+%prefetch -- if it does, then it has a choice of how many cache lines to
+%prefetch.  The suggested prefetches can be either placed in the L2 cache or
+%the last-level cache based on the prefetcher's internal confidence mechanism.
+%
+%When the base prefetcher is triggered, it begins to suggest candidates for
+%prefetching.  All suggested prefetches are passed through the perceptron
+%filter to decide whether they are qualified for final prefetching.  The filter
+%decides whether to prefetch a candidate by looking at the microarchitectural
+%state, \textit{i.e.} the features, at that instant.  Each feature is hashed to
+%form an index into a table of up to 4,096 entries dedicated to that feature.
+%
+%Once all the weights are retrieved, they are summed and compared to a preset
+%threshold, PERC\_ THRESHOLD\_LO.  Only prefetch candidates with perceptron sum
+%higher than the threshold qualify for prefetching.  The prefetches that
+%qualify through the perceptron stage are recorded in the ``Prefetch Table''.
+%The prefetch table is a 1,024-entry, direct mapped structure that contains all
+%metadata required to re-index the perceptron entries for training.  When the
+%feedback of the current prefetched line is available at a later stage, the
+%stored data is used to train the perceptron.
+%
+%%\pg{this discussion below has redundancy and doesn't flow well.
+%%  Rework it}
+%
+%Each feature's confidence contributes to the final decision of whether to
+%prefetch a cache line, with the summation of all features' confidences
+%denoting the overall confidence for a prefetch.  The final sum is compared
+%against two thresholds: PERC\_THRESHOLD\_HI and  PERC\_ THRESHOLD\_ LO to
+%decide the placement of the data in the cache hierarchy. Prefetches that have
+%a confidence above the higher threshold are placed into the L2 cache since the
+%observed features strongly recommend  the prefetch. A prefetch that results in
+%a confidence between PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI are
+%placed in the larger last level cache, as the filter is moderately confident
+%of the cache line's use, but seeks to reduce pollution in the smaller L2
+%cache. Finally, the remaining prefetches that are below the lower threshold
+%are not prefetched due to a lack of correlation between the current observed
+%features and the prediction from the base prefetch engine.
+%
+%In addition to the prefetch table mentioned above, PPF also maintains a
+%``Reject Table.''  The reject table is a 1,024-entry direct-mapped table.
+%If a prefetch suggestion is rejected by the perceptron layer, it is logged
+%into the reject table.  The table is used to train the perceptron to avoid
+%false negatives \textit{i.e.}, cases where prediction was to reject the
+%prefetch but the prefetch would have been useful.
+%
+%%As previously stated all the metadata describing the program state at the
+%%instant of prefetching is stored in the Prefetch Table or the Reject Table.
+%%This information is useful when the perceptron needs to be updated
+%%subsequently.
+%
+%\textbf{Training}\newline In the prefetching environment, feedback for a
+%prefetch is received whenever there is an eviction or a demand access from the
+%L2 cache.  This action triggers training of the base prefetcher as well as the
+%perceptron layer.  Training the perceptrons involves accessing the metadata
+%that was stored in Prefetch Table or the Reject Table.  The cache lines
+%address of the block being trained is used to access a table, using 10 bits of
+%the address for indexing and another 6 bits to perform tag matching.  Once the
+%values of the features at the time of prefetching is available, it is used to
+%index into the perceptron weights table.
+%
+%If the demanded block that triggers the training was tagged as a valid
+%prefetch in the prefetch table, then the earlier prefetch prediction was
+%correct.  In that case the perceptron weights are incremented by 1 if the
+%predicted sum does not cross a pre-defined threshold. These training
+%thresholds are introduced to avoid overfitting of the perceptron weights to
+%the given program behavior. These thresholds are referred to as
+%POS\_UPDATE\_THRESHOLD ($\theta_p$) and NEG\_UPDATE\_THRESHOLD ($\theta_n$),
+%respectively for the positive and negative values of training saturation.
+%
+%If a cache block eviction led to training and the corresponding valid entry is
+%found in the prefetch table without the block being accessed, then the
+%prediction made by the perceptron was wrong.  The perceptron should have
+%ideally rejected the prefetch suggestion as a low-confidence prefetch.  Here
+%the weights are decremented by 1 to reflect the misprediction. In either case,
+%weights are saturated at -16 or +15.
+%
+%A secondary training mechanism also kicks in during demand fetches.  Before
+%the demand access triggers the next set of prefetches, the reject table is
+%checked for a valid entry.  A hit means that the corresponding cache line was
+%initially suggested by the underlying prefetch engine, but rejected by the
+%perceptron filter.  Thus, the perceptron should have been more confident about
+%that particular prefetch.  Once such a scenario is identified, the values of
+%the features at the time of prefetch is retrieved from the reject table.  The
+%retrieved data is used to index into the various weights tables of the
+%perceptron and the corresponding values are updated by +1, saturating between
+%and -16 and +15.  This update reflects increased confidence for the prediction
+%corresponding to that prefetch.
+%
+%This mechanism allows us to exploit a previously lost opportunity.  In prior
+%perceptron based implementations, and in general, prefetching algorithms,
+%there is usually no way of knowing the result of not prefetching a particular
+%line.  Our two-step PPF architecture allows us to overcome that issue.
 

### 2018-12-05 design_arch.tex
gino check this

@@
 depth, page address and offset, into a binary useful/non-useful prediction.
-In the remainder of this section, we discuss the the general structure of PPF
+In the remainder of this section, we discuss the general structure of PPF
 for enhancing a given underlying prefetcher.  We will show that PPF is a
@@
 
-The Perceptron Filter and the steps used to performed inference 
-and training are shown in figure~\ref{fig:PPF_Datapath}. The filter is organized as tables of 
-perceptron weights, with a separate table of weights for each feature. 
-For a design using `N' number of features, its weights will be arranged
-in $N$ different tables. Each feature's value is used to index 
-into the corresponding table with $K$ entries.  Since each table is
-indexed independently, different numbers of bits can be used for indexing,
-thus giving each table a different size.  The variable indexing can be
-determined by studying the features and fine-tuning them empirically to
-achieve a good accuracy versus hardware overhead trade-off.  Exact details of
-our PPF implementation are given under ``Area Overheads'' in
+The Perceptron Filter structure and the steps used to performed inference 
+and training are illustrated in figure~\ref{fig:PPF_Datapath}. 
+The filter is organized as a set of tables, where each table entry holds a  
+perceptron weight. 
+%There exists a distinct table for each of the features. redundant with next sentence?
+
+For a configuration using $N$ number of features, 
+their corresponding weights will be distributed
+in $N$ different tables. Each feature is used to index 
+the corresponding table with $K$ entries.  Each table is
+indexed independently, and the size of table varies, hence, different
+number of bits are used to index different the tables.
+
+%ETG: so not all tables have K entries? K is not a constant,
+
+Variable indexing can be
+determined by studying the features and tuning them empirically to
+achieve a good tradeoff between high accuracy and hardware overhead.
+A detail explanation of the storage overhead of
+our PPF implementation can be found in
 Section~\ref{sec:Method-Overheads}.
@@
 is a 5 bit counter saturating at -16 and +15 and initialized to zero at the
-beginning of a program's execution. We empirically observe that 5 bits of
-index achieves a good trade-off between hardware overhead and accuracy.\newline
+beginning of an applications's execution. We empirically observe having 5 bit weights,
+yields a good trade-off between hardware overhead and accuracy.\newline
 
@@
 demand access.  At that point, the base prefetcher has the option to perform a
-prefetch -- if it does, then it has a choice of how many cache lines to
+prefetch -- if it does, then it makes the decision of how many cache lines to
 prefetch.  The suggested prefetches can be either placed in the L2 cache or
-the last-level cache based on the prefetcher's internal confidence mechanism.
+the last-level cache based on a confidence value.
 
@@
 prefetching.  All suggested prefetches pass through the perceptron
-filter to decide whether they are qualified for final prefetching.  The filter
-decides whether to prefetch a candidate by looking at the microarchitectural
-state, \textit{i.e.} the features, at that instant.  Each feature is hashed to
-form an index into a table of up to 4,096 entries dedicated to that feature.
+filter to decide whether they are qualified for final prefetching. %final prefetching -_-  
+The filter makes the final decision to prefetch a \textit{prefetch candidate}
+suggested by the base prefetcher.
+Each of the feature is used to index a distinct table, 
+the corresponding eeights are summed. The sum denotes the overall confidence for the prefetch. 
+The value is thresholded against two diffent magnitudes:
 
-Each feature's confidence contributes to 
-the final decision of whether to prefetch a cache line, with the summation 
-of all features' confidences denoting the overall confidence for a prefetch.  
-The final sum is compared against two thresholds: PERC\_THRESHOLD\_HI and  
-PERC\_ THRESHOLD\_ LO to decide the placement of the data in the cache 
-hierarchy. Prefetches that have a confidence above the higher threshold 
-are placed into the L2 cache since the observed features strongly recommend  
-the prefetch. A prefetch that results in a confidence between 
-PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI are placed in the larger 
-last level cache, as the filter is moderately confident of the cache line's 
-use, but seeks to reduce pollution in the smaller L2 cache. Finally, the 
-remaining prefetches that are below the lower threshold are not prefetched 
-due to a lack of correlation between the current observed features and 
-the prediction from the base prefetch engine.\newline
+%The filter
+%decides whether to prefetch a candidate by looking at the microarchitectural
+%state, \textit{i.e.} the features.  Each feature is used as an index 
+%to its corresponding table, each table can have up 4,096 entries.
+%Each feature's confidence contributes to 
+%the final decision of whether to prefetch a cache line, with the summation 
+%of all features' confidences denoting the overall confidence for a prefetch.  
+%The final sum is compared against two thresholds:
+
+PERC\_THRESHOLD\_HI and  
+PERC\_ THRESHOLD\_ LO. 
+%to decide the placement of the data in the cache 
+%hierarchy. 
+Prefetches that have a confidence above the higher magnitude 
+are placed into the L2 cache since the obtained summed suggests with high
+confidence this would be a useful prefetch.
+A prefetchs which features result in a confidence value between  
+%A prefetch that results in a confidence between 
+PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI is allocated in the larger 
+last-level cache, the filter is moderately confident of use of this 
+cache line, but seeks to reduce pollution in a significantly smaller L2 cache. 
+Lastly, prefetches which features yield a confident value lower than 
+PERC\_ THRESHOLD\_ LO are not prefetched. \newline 
+%due to a lack of correlation between the current observed features and 
+%the prediction from the base prefetch engine.\newline
 

### 2018-12-05 design_arch.tex
filter mechanism

--- 2018-12-04 design_arch.tex
@@
 
-As previously discussed, ultimately it might be profitable to allow a given
-prefetcher to speculate as deeply as possible, since often some useful
+As mentioned before, ultimately it might be benefitial to allow a
+prefetcher to speculate as deeply as possible. Often, some useful
 prefetches are generated long after the confidence of the prefetcher has
-fallen below the point at which performance improvements are seen due to the
-increasing rate of inaccurate prefetches.  However, to allow this deep
-speculation in the prefetcher the increasing numbers of bad prefetches must be
-filtered out.  Thus, we propose to leverage perceptron-based learning as a
+fallen below a point, in which performance improvements are observed, due to the
+increase of inaccurate prefetches.  However, in order to allow deep
+speculation in the prefetcher, inaccurate prefetches must be
+filtered out.  We propose to leverage perceptron-based learning as a
 mechanism to differentiate between potentially useful deeply speculated
-prefetches and likely not-useful ones.  This Perceptron Prefetch Filter (PPF)
-is placed between the prefetcher and the prefetch insertion queue, to ensure
-that predicted bad prefetches are not inserted into the memory system as shown in 
-figure~\ref{fig:PPF_Hierarchy}.
+prefetches and likely not-useful ones. The Perceptron Prefetch Filter (PPF)
+is placed between the prefetcher and the prefetch insertion queue, as illuastrated
+in figure ~\ref{fig:PPF_Hierarchy}, to prevent
+not-useful prefetches from polluting the higher levels of the memory hierarchy.
 
@@
 is a light-weight mechanism to pull together disparate forms of information
-and synthesize a decision from them.  In this case, we leverage a variety of
-different relative characteristics of a given prefetch, such as speculation
-depth, page address and offset, into a binary useful/non-useful prediction.
-In the remainder of this section, we discuss the the general structure of PPF
-for enhancing a given underlying prefetcher.  We will show that PPF is a
-generalized prefetch filtering mechanism that may be adapted to any prefetcher
-with appropriate modifications.
+and synthesize a decision from them.  Our work considers a number of
+features corresponding to a prefetch, such as speculation
+depth, page address and offset, and use this information as input to our perceptron-based filter in order to predict the usefulness of a prefetch.
+In the remainder of this section, we discuss our proposed technique PPF.
+It enhances a base prefetcher, by filtering out not-useful prefetches. 
+PPF is a generalized prefetch filtering mechanism that may be adapted 
+to any prefetcher with appropriate modifications.
 
@@
 
-The Perceptron Filter and the steps used to performed inference 
-and training are shown in figure~\ref{fig:PPF_Datapath}. The filter is organized as tables of 
-perceptron weights, with a separate table of weights for each feature. 
-For a design using `N' number of features, its weights will be arranged
-in $N$ different tables. Each feature's value is used to index 
-into the corresponding table with $K$ entries.  Since each table is
-indexed independently, different numbers of bits can be used for indexing,
-thus giving each table a different size.  The variable indexing can be
-determined by studying the features and fine-tuning them empirically to
-achieve a good accuracy versus hardware overhead trade-off.  Exact details of
-our PPF implementation are given under ``Area Overheads'' in
-Section~\ref{sec:Method-Overheads}.
-
-A single entry in the table corresponds to a perceptron weight.  Each weight
-is a 5 bit counter saturating at -16 and +15 and initialized to zero at the
-beginning of a program's execution. We empirically observe that 5 bits of
-index achieves a good trade-off between hardware overhead and accuracy.\newline
+Figure~\ref{fig:PPF_Datapath} shows the required structures of PPF, as well as the
+steps followed to filter out not-useful prefetches.
+The filter is organized as a set of tables, where each entry in a tables holds 
+a \textit{weight}. For a configuration of PPF using $N$ number of features, 
+$N$ different tables of weights are needed. Each feature is used to index 
+a distinct table. The number of entries of each table varies according
+to their corresponding feature, hence, different
+number of bits are needed to index different tables.
+Each weight is a 5-bit saturating counter ranging from -16 to +15. 
+We found the having 5-bit weights provide a good trade-off 
+between accuracy and area. A detail explanation of the storage overhead of
+PPF can be found in Section~\ref{sec:Method-Overheads}.
 
@@
 
-The prefetcher is triggered on every L2 cache
-demand access.  At that point, the base prefetcher has the option to perform a
-prefetch -- if it does, then it has a choice of how many cache lines to
-prefetch.  The suggested prefetches can be either placed in the L2 cache or
-the last-level cache based on the prefetcher's internal confidence mechanism.
+%ETG:is this like an explanation of how prefetching happens?
 
-When the base prefetcher is triggered, it begins to suggest candidates for
-prefetching.  All suggested prefetches pass through the perceptron
-filter to decide whether they are qualified for final prefetching.  The filter
-decides whether to prefetch a candidate by looking at the microarchitectural
-state, \textit{i.e.} the features, at that instant.  Each feature is hashed to
-form an index into a table of up to 4,096 entries dedicated to that feature.
+The base prefetcher is triggered on every demand access to the L2.  
+As this point, it has the opportunity to trigger a prefetch  -- if it 
+decides to do so, it will also need to decide how many cache lines
+to prefetch.  These cache lines can be either placed in the L2 or
+L3 according to the confidence of the prefetching mechanism.
+Once the base prefetcher is triggered, the suggested
+prefetch candidates are feeded to the perceptron filter, 
+in order to determine the usefulnes of these prefetches.  
+The filter makes the decision to issue a prefetch request based on 
+the suggestions from the base prefetcher.
+In order to make such decision, each feature corresponding to a 
+suggested prefetch is used to index a table, 
+all corresponding weights are summed. 
+The sum denotes the confidence value for the suggested prefetch, 
+and is thresholded against two diffent magnitudes:  
+PERC\_THRESHOLD\_HI and PERC\_ THRESHOLD\_ LO.
 
-Each feature's confidence contributes to 
-the final decision of whether to prefetch a cache line, with the summation 
-of all features' confidences denoting the overall confidence for a prefetch.  
-The final sum is compared against two thresholds: PERC\_THRESHOLD\_HI and  
-PERC\_ THRESHOLD\_ LO to decide the placement of the data in the cache 
-hierarchy. Prefetches that have a confidence above the higher threshold 
-are placed into the L2 cache since the observed features strongly recommend  
-the prefetch. A prefetch that results in a confidence between 
-PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI are placed in the larger 
-last level cache, as the filter is moderately confident of the cache line's 
-use, but seeks to reduce pollution in the smaller L2 cache. Finally, the 
-remaining prefetches that are below the lower threshold are not prefetched 
-due to a lack of correlation between the current observed features and 
-the prediction from the base prefetch engine.\newline
+Prefetches which sum is above PERC\_THRESHOLD\_HI
+are placed into the L2 cache, the higher confidence value
+hints the prefetch would be useful and should be prioritized.
+A prefetch for which its corresponding features result 
+in a confidence value between  
+PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI is allocated in the larger 
+last-level cache, the filter is moderately confident of the future reuse of the 
+cache line, but not enough to possible pollute a significantly smaller L2. 
+Suggested prefetches which features yield a confident value lower than 
+PERC\_ THRESHOLD\_ LO are not prefetched, a low confidence value 
+represents that the filter has learn that similar set of features
+are associated with not-useful prefetches. \newline
 
@@
 
-The prefetches that qualify through the inference stage are recorded in the 
+As shown in step 2 of figure ~\ref{fig:PPF_Datapath}
+the prefetches that make it through the inference stage are recorded in the 
 ``Prefetch Table''. The prefetch table is a 1,024-entry, direct mapped 
@@
 the reject table. The table is used to train the perceptron to avoid false negatives 
-\textit{i.e.}, cases where prediction was to reject the prefetch but 
-the prefetch would have been useful based on the observed demand accesses to the L2. \newline
+\textit{i.e.}, cases where the prediction suggested to reject the prefetch 
+but the prefetch was proven to be useful based 
+on the observed demand accesses to the L2. \newline
 
@@
 
-In the prefetching environment, feedback for a prefetch is received whenever 
-there is an eviction or a demand access from the L2 cache. This action 
-triggers training of the base prefetcher as well as the perceptron layer.  
-The cache line's address of the block being trained is used to access either the Prefetch
-or Reject table, using 10 bits of the address for indexing and another 6 bits to 
-perform tag matching.  Once the values of the features at the time of prefetching is available,
-it is used to index into the perceptron weights table. \newline
+When there is an eviction or a demand access to the L2, training for both the base 
+prefetcher and our filter mechanism is triggered. 
+The address of the cache line that triggered training is used to index
+both the Prefetch and Reject tables. 10 bits of the address are used to index 
+tables, and another 6 bits to perform tag matching. 
+Once the base prefetcher has suggested a prefetch request, the corresponding 
+features are used to index the tables of weights. \newline
+
+% Which bits? are the tables accesses sequentially?
 
@@
 
-Training the perceptrons involves accessing the metadata
-that was stored in Prefetch Table or the Reject Table.
-
-If the demanded block that triggers the training was tagged as a valid
-prefetch in the prefetch table, then the earlier prefetch prediction was
-correct.  In that case the perceptron weights are incremented by 1 if the
-predicted sum does not cross a pre-defined threshold. These training
-thresholds are introduced to avoid overfitting of the perceptron weights to
-the given program behavior. These thresholds are referred to as
+The address from the demand request that trigered training is lookup in 
+both tables. If the address resides in the prefetch table, and is marked 
+as valid, this hints the previous prediction was correct and this is a useful
+prefetch. We compute the sum of the corresponding weights and
+compare the sum to an specific threshold, if the sum falls below it, 
+training must happen and the corresponding weights are adjusted accordingly.
+These thresholds are introduced inorder to avoid overtraining, 
+which helps the filter adapt quickly to changes in memory behavior. 
+These thresholds are referred to as
 POS\_UPDATE\_THRESHOLD ($\theta_p$) and \newline
@@
 
-If a cache block eviction led to training and the corresponding valid entry is
-found in the prefetch table without the block being accessed, then the
-prediction made by the perceptron was wrong.  The perceptron should have
-ideally rejected the prefetch suggestion as a low-confidence prefetch.  Here
-the weights are decremented by 1 to reflect the misprediction. In either case,
-weights are saturated at -16 or +15.
+On an eviction of a cache block, we look up the corresponding address
+in the prefetch table. If there is a valid entry with this address, 
+this mean the filter made a misprediction. The block was allocated in the
+L2 with a prefetch request that the filter should have categorized 
+not-reuse prefetch. In this scenario, the corresponding features of the
+prefetch request that cause these block to be placed in the cache are used
+to index the tables of weights, and adjust weights accordingly.
 
-A secondary training mechanism also kicks in during demand fetches.  Before
+%wait at the same time or no?
+Paralell to accessing the prefetch table, on a demand access, 
+the reject table is accessed. Before
 the demand access triggers the next set of prefetches, the reject table is
 checked for a valid entry.  A hit means that the corresponding cache line was
-initially suggested by the underlying prefetch engine, but rejected by the
-perceptron filter.  Thus, the perceptron should have been more confident about
-that particular prefetch.  Once such a scenario is identified, the values of
-the features at the time of prefetch is retrieved from the reject table.  The
-retrieved data is used to index into the various weights tables of the
-perceptron and the corresponding values are updated by +1, saturating between
-and -16 and +15.  This update reflects increased confidence for the prediction
-corresponding to that prefetch.
+initially suggested by the base prefetcher, but wrongly rejected by the
+perceptron filter. The perceptron filter learns from this and makes use of the
+correspoding features associated to the original prefetch request, which are
+stored in the reject table, to index the weights tables and adjust the weigths
+accordingly.
 
-This mechanism allows us to exploit a previously lost opportunity.  In prior
-perceptron based implementations, and in general, prefetching algorithms,
-there is usually no way of knowing the result of not prefetching a particular
-line.  Our two-step PPF architecture allows us to overcome that issue.
+Prior work lacks the knowledge of which cache lines should have 
+been prefetched, but were not. The implementation of the reject table,
+allows us to have this information, which can be use to futher optimize
+our prefetching mechanism.
 

### 2018-12-05 related.tex
missing prefetchers

@@
 
-Two widely used prefetchers are ?Next-n Lines?
-\cite{nextnlines} and Stride [4], both of which capture regular memory access patterns with very low hardware overhead.
-The ?Next-n Lines?
-prefetcher simply queues prefetches for the next n lines after any given miss, under the expectation that the principle of spatial locality will hold and those cache lines after a missed line are likely to be used in the future.
+Two widely used prefetchers are Next-n Lines \cite{nextnlines} and Stride \cite{stride}, both of which capture regular memory access patterns with very low hardware overhead.
+The Next-n Lines prefetcher simply queues prefetches for the next n lines after any given miss, under the expectation that the principle of spatial locality will hold and those cache lines after a missed line are likely to be used in the future.
 The stride prefetcher is slightly more sophisticated; it attempts to identify simple stride reference patterns in programs based upon the past behavior of missing loads.
-Similar to the ?Next-n Lines?
-technique, when a given load misses, cache lines ahead of that miss are fetched in the pattern following the previous be- havior in the hope of avoiding future misses.
-Both prefetch- ers have been widely used due to the simple design and low hardware overhead.
+Similar to the Next-n Lines technique, when a given load misses, cache lines ahead of that miss are fetched in the pattern following the previous behavior in the hope of avoiding future misses.
+Both prefetchers have been widely used due to the simple design and low hardware overhead.
 However, without further knowledge about temporal locality and application characteristics, these prefetchers cannot do more than detecting and prefetching regular memory access patterns with limited spatial locality.
@@
 Somogyi et al.
-proposed one of the current top- performing, practical, low-overhead prefetchers, the Spatial Memory Streaming (SMS) prefetcher [23].
+proposed one of the top practical, low-overhead prefetchers, the Spatial Memory Streaming (SMS) prefetcher \cite{sms}.
 SMS leverages code-based correlation to take advantage of spatial locality in the applications over larger regions of memory (called spatial regions).
@@
 
-%ANMP
+To address both prefetching coverage and accuracy, prior work has adopted lookahead 
+mechanisms. Lookahead prefetchers efficiently encode the relationship between accesses to yield future predictions, enabling further speculative lookahead accesses.
+VLDP \cite {vldp} prefetches a static depth ahead of the demand fetch, without considering the prefetching confidence,
+
+DA-AMPM
 
@@
 \subsection{Aggressive Prefetching}
+
+%not sure this is where this section was going, I just kidna went with it.
 

### 2018-12-05 related.tex
addes some references

@@
 
-Two widely used prefetchers are Next-n Lines \cite{nextnlines} and Stride \cite{stride}, both of which capture regular memory access patterns with very low hardware overhead.
+Two widely used prefetchers are Next-n Lines \cite{nextn} and Stride \cite{stride}, both of which capture regular memory access patterns with very low hardware overhead.
 The Next-n Lines prefetcher simply queues prefetches for the next n lines after any given miss, under the expectation that the principle of spatial locality will hold and those cache lines after a missed line are likely to be used in the future.
@@
 Somogyi et al.
-proposed one of the top practical, low-overhead prefetchers, the Spatial Memory Streaming (SMS) prefetcher \cite{sms}.
+proposed one of the top practical, low-overhead prefetchers, the Spatial Memory Streaming (SMS) prefetcher \cite{SMS}.
 SMS leverages code-based correlation to take advantage of spatial locality in the applications over larger regions of memory (called spatial regions).
@@
 mechanisms. Lookahead prefetchers efficiently encode the relationship between accesses to yield future predictions, enabling further speculative lookahead accesses.
-VLDP \cite {vldp} prefetches a static depth ahead of the demand fetch, without considering the prefetching confidence,
+VLDP \cite {VLDP} prefetches a static depth ahead of the demand fetch, without considering the prefetching confidence,
 
@@
 ICP simply demotes a prefetch blow to the lowest priority on a demand hit based on the observation that most prefetches are dead after their first hit.
-To address prefetcher-caused cache pollution, it also uses a variation of EAF [27] to track prefetching accuracy and inserts only an accurate prefetch to the higher priority position in the LRU stack.
+To address prefetcher-caused cache pollution, it also uses a variation of EAF \cite{eaf} to track prefetching accuracy and inserts only an accurate prefetch to the higher priority position in the LRU stack.
 ICP assumes, however, that all prefetches are inserted only into the LLC, which restricts the maximum benefit of prefetching.

### 2018-12-05 results.tex
Worked on Images, Results Section

@@
 
-PPF yields a speedup of \textbf{42.8\%} over the baseline. This is equivalent
-to \textbf{7.8\%} over DA-AMPM, \textbf{9.7\%} over BOP and \textbf{6.92\%}
-over SPP.  Out of the 95 simpoints developed for SPEC CPU 2017, PPF nearly
-matches or outperforms most of the prefetchers on 91 simpoints.  PPF fails to
-match the improvement offered by BOP only for {\tt 607.cactuBSSN\_s} and {\tt
-654.roms\_s}.
+PPF yields a geometric mean speedup of \textbf{42.8\%} over the baseline. 
+This is equivalent to \textbf{7.8\%} over DA-AMPM, \textbf{9.7\%} over BOP 
+and \textbf{6.92\%} over SPP.  Out of the 95 simpoints developed for SPEC 
+CPU 2017, PPF nearly matches or outperforms most of the prefetchers on 91 
+simpoints.  PPF fails to match the improvement offered by BOP only for 
+{\tt 607.cactuBSSN\_s} and {\tt 654.roms\_s}.
 
@@
 \textbf{60\%} over the next best prefetcher -- BOP.  In general, benchmarks
-{\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt 623.xalancbmk\_s} and {\tt
+{\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt 623. xalancbmk\_s} and {\tt
 649.fotonik3d\_s} benefit the most from PPF, with the speedup over SPP ranging
@@
 % djimenez: is this geometric mean? if so, say so.
-On the full SPEC CPU 2017 suite, PPF improves the baseline by \textbf{20.9\%},
-which is \textbf{3.44\%} better than the next best prefetcher -- SPP.
+% [EB] Resolved
+On the full SPEC CPU 2017 suite, PPF improves the geometric mean IPC of 
+the baseline by \textbf{20.9\%}, which is \textbf{3.44\%} better 
+than the next best prefetcher -- SPP.
+\newline
 
@@
 
-\textbf{Coverage}\newline
+\noindent \textbf{COVERAGE}
+\newline
 Prefetcher coverage is defined as the ratio of the number of misses avoided
@@
 
-\begin{figure}[h]
-\includegraphics[width=\columnwidth]{4Core_SPEC2017}
-\caption{Normalized speedup for 4-core workloads}
+\begin{figure*}[ht]
+\includegraphics[width=\textwidth]{4Core_SPEC2017}
+\caption{Normalized Speedup for 4-core SPEC CPU 2017 Workloads}
 \label{Fig:4Core_SPEC2017}
-\end{figure}
+\end{figure*}
 
@@
 multi-programmed workloads.
+\newline
 
-\textit{4-core environment} Figure~\ref{Fig:4Core_SPEC2017}(a) shows a
+\noindent \textbf{4-CORE ENVIRONMENT}
+\newline
+Figure~\ref{Fig:4Core_SPEC2017}(a) shows a
 comparison of speedups obtained on 4-core mixes of a memory intensive subset
@@
 \textbf{5.7}\% over SPP.
+\newline
 
-\begin{figure}[h]
-\includegraphics[width=\columnwidth]{8Core_SPEC2017}
-\caption{Normalized speedup for 8-core workloads}
+\begin{figure*}[ht]
+\includegraphics[width=\textwidth]{8Core_SPEC2017}
+\caption{Normalized Speedup for 8-core SPEC CPU 2017 Workloads}
 \label{Fig:8Core_SPEC2017}
-\end{figure}
+\end{figure*}
 
-\textit{8-core environment} The sorted comparison of speedups on the memory
+\noindent \textbf{8-CORE ENVIRONMENT}
+\newline
+The sorted comparison of speedups on the memory
 intensive 8-core mixes is shown in Figure~\ref{Fig:8Core_SPEC2017}.  PPF
@@
 \begin{figure}[ht]
-\includegraphics[width=\columnwidth]{AddnConstr}
+\begin{adjustwidth}{-1cm}{}
+  \includegraphics[width=1.2\columnwidth]{AddnConstr}
   \caption{IPC speedup for Small LLC and Low BW}
   \label{Fig:AddnConstr}
+\end{adjustwidth}
 \end{figure}
@@
 % djimenez: cite cloudsuite here? do we have a reference?
+% [EB]: Resolved. Cited "Clearing the clouds: A study of emerging 
+%        scale-out workloads on modern hardware" in methodology
 
@@
 % here? that's weird.
+% [EB] Resolved. Increased precision of values. But yes, BOP / DA-AMPM 
+% have really close numbers for SPEC 2006
 
@@
 2006 benchmark, giving an improvement of \textbf{6.66\%} over SPP and
-\textbf{9.86\%} over BOP and DA-AMPM.  On the whole of the SPEC CPU 2006
-suite, the speedup is \textbf{22.4\%}, an improvement of \textbf{3.34\%} over
-SPP.
+\textbf{9.88\%} over DA-AMPM and \textbf{9.82} over BOP.  On the whole of 
+the SPEC CPU 2006 suite, the speedup is \textbf{22.4\%}, an improvement 
+of \textbf{3.34\%} over SPP.
 

### 2018-12-05 design_arch.tex
Some small typos

@@
 
-As mentioned before, ultimately it might be benefitial to allow a
+As mentioned before, ultimately it might be beneficial to allow a
 prefetcher to speculate as deeply as possible. Often, some useful
 prefetches are generated long after the confidence of the prefetcher has
-fallen below a point, in which performance improvements are observed, due to the
+fallen below a point, after which the performance degrades, due to the
 increase of inaccurate prefetches.  However, in order to allow deep
@@
 prefetches and likely not-useful ones. The Perceptron Prefetch Filter (PPF)
-is placed between the prefetcher and the prefetch insertion queue, as illuastrated
+is placed between the prefetcher and the prefetch insertion queue, as illustrated
 in figure ~\ref{fig:PPF_Hierarchy}, to prevent
@@
 features corresponding to a prefetch, such as speculation
-depth, page address and offset, and use this information as input to our perceptron-based filter in order to predict the usefulness of a prefetch.
-In the remainder of this section, we discuss our proposed technique PPF.
+depth, page address and offset, and uses this information as the inputs to our 
+perceptron-based filter, in order to predict the usefulness of a prefetch.
+In the remainder of this section, we discuss our proposed technique -- PPF.
 It enhances a base prefetcher, by filtering out not-useful prefetches. 
@@
 steps followed to filter out not-useful prefetches.
-The filter is organized as a set of tables, where each entry in a tables holds 
+The filter is organized as a set of tables, where each entry in the tables holds 
 a \textit{weight}. For a configuration of PPF using $N$ number of features, 
@@
 Each weight is a 5-bit saturating counter ranging from -16 to +15. 
-We found the having 5-bit weights provide a good trade-off 
+We found that having 5-bit weights provide a good trade-off 
 between accuracy and area. A detail explanation of the storage overhead of
@@
 %ETG:is this like an explanation of how prefetching happens? 
+%EB: I see this as a short context of the datapath leading to the perc filter.
 
-The base prefetcher is triggered on every demand access to the L2.  
+The base prefetcher is triggered on every demand access to the L2 Cache.  
 As this point, it has the opportunity to trigger a prefetch  -- if it 
 decides to do so, it will also need to decide how many cache lines
-to prefetch.  These cache lines can be either placed in the L2 or
-L3 according to the confidence of the prefetching mechanism.
+to prefetch.  These cache lines can be either placed in the L2 Cache or
+the L3 Cache, according to the confidence of the prefetching mechanism.
 Once the base prefetcher is triggered, the suggested
-prefetch candidates are feeded to the perceptron filter, 
-in order to determine the usefulnes of these prefetches.  
-The filter makes the decision to issue a prefetch request based on 
-the suggestions from the base prefetcher.
+prefetch candidates are fed to the perceptron filter, 
+in order to determine the usefulness of these prefetches.  
+The filter makes the decision to issue the prefetch
+suggestions from the base prefetcher.
 In order to make such decision, each feature corresponding to a 
 suggested prefetch is used to index a table, 
-all corresponding weights are summed. 
+all the corresponding weights are summed. 
 The sum denotes the confidence value for the suggested prefetch, 
-and is thresholded against two diffent magnitudes:  
+and is thresholded against two different magnitudes:  
 PERC\_THRESHOLD\_HI and PERC\_ THRESHOLD\_ LO.
@@
 in a confidence value between  
-PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI is allocated in the larger 
-last-level cache, the filter is moderately confident of the future reuse of the 
+PERC\_THRESHOLD\_LO and PERC\_THRESHOLD\_HI is allocated in the larger 
+last-level cache, as the filter is moderately confident of the future reuse of the 
 cache line, but not enough to possible pollute a significantly smaller L2. 
-Suggested prefetches which features yield a confident value lower than 
-PERC\_ THRESHOLD\_ LO are not prefetched, a low confidence value 
-represents that the filter has learn that similar set of features
-are associated with not-useful prefetches. \newline
+Suggested prefetches for which the features lead to a confidence value lower than 
+PERC\_THRESHOLD\_LO are not prefetched, as the low confidence value 
+represents that the perceptron has learn that similar set of features
+are associated with non-useful prefetches. \newline
 
@@
 
-The address from the demand request that trigered training is lookup in 
+The address from the demand request that triggered the training is lookup in 
 both tables. If the address resides in the prefetch table, and is marked 
@@
 training must happen and the corresponding weights are adjusted accordingly.
-These thresholds are introduced inorder to avoid overtraining, 
+These thresholds are introduced in order to avoid over-training, 
 which helps the filter adapt quickly to changes in memory behavior. 
@@
 L2 with a prefetch request that the filter should have categorized 
-not-reuse prefetch. In this scenario, the corresponding features of the
+as a useless prefetch. In this scenario, the corresponding features of the
 prefetch request that cause these block to be placed in the cache are used
-to index the tables of weights, and adjust weights accordingly.
+to index the tables of weights, and those weights are adjusted accordingly.
 
 %wait at the same time or no?
-Paralell to accessing the prefetch table, on a demand access, 
+%Yes, and they are mutex cases
+
+Parallel to accessing the prefetch table, on a demand access, 
 the reject table is accessed. Before
@@
 perceptron filter. The perceptron filter learns from this and makes use of the
-correspoding features associated to the original prefetch request, which are
-stored in the reject table, to index the weights tables and adjust the weigths
+corresponding features associated to the original prefetch request, which are
+stored in the reject table, to index the weights tables and adjust the weights
 accordingly.
 
-Prior work lacks the knowledge of which cache lines should have 
+Prior work in this domain lacks the knowledge of which cache lines should have 
 been prefetched, but were not. The implementation of the reject table,
-allows us to have this information, which can be use to futher optimize
+allows us to have this information, which can be use to further optimize
 our prefetching mechanism.
@@
 The above discussion of PPF shows that it is highly modular and can be adapted
-to be used over any base prefetcher for increased prefetch accuracy.  In
-general, only two hooks are required between PPF and the baseline prefetcher.
+to be used over any base prefetcher for increased prefetch accuracy. Only 
+two hooks are required between PPF and the baseline prefetcher.
 The first is to make sure that all the prefetch candidates of the prefetcher

### 2018-12-05 back.tex
Adding Perc image, motivation section

@@
 
-Compared to some of the other state of the art prefetchers, SPP is
-less aggressive.  In a single-core environment, there is no resource
-contention among the different cores, hence aggressive prefetching is bound to prove beneficial. 
-As we increase the core
-count, we observe that SPP starts outperforming rest of the prefetchers.  This
-can be attributed to the fact that each prefetch suggested by SPP is a
-carefully calculated and prevents cache pollution.  \textit{<Figure
-showing the variation of SPP vs BOP wrt core count>} For 4-core applications,
-SPP suggests XX\% fewer prefetches than Best Offset Prefetcher ~\cite{bop} and yet leads to higher IPC.
+%% Compared to some of the other state of the art prefetchers, SPP is
+%% less aggressive.  In a single-core environment, there is no resource
+%% contention among the different cores, hence aggressive prefetching is bound to prove beneficial. 
+%% As we increase the core
+%% count, we observe that SPP starts outperforming rest of the prefetchers.  This
+%% can be attributed to the fact that each prefetch suggested by SPP is a
+%% carefully calculated and prevents cache pollution.  \textit{<Figure
+%% showing the variation of SPP vs BOP wrt core count>} For 4-core applications,
+%% SPP suggests XX\% fewer prefetches than Best Offset Prefetcher ~\cite{bop} and yet leads to higher IPC.
+%% 
+%% The above analysis shows that with a more careful filtering mechanism, any
+%% prefetcher like SPP can be tuned to become much more aggressive, yielding higher coverage.  The responsibility of maintaining high accuracy now falls on the
+%% independent filter.  To test the hypothesis, we tuned down SPP to the
+%% minimum possible threshold, with the effect of increasing prefetch suggestions
+%% made by SPP by XX\%.  \textit{<Comparison of SPP-Unleased with SPP / BOP>}
+%% Obviously this came at a cost of increased DRAM traffic and cache pollution.
 
-The above analysis shows that with a more careful filtering mechanism, any
-prefetcher like SPP can be tuned to become much more aggressive, yielding higher coverage.  The responsibility of maintaining high accuracy now falls on the
-independent filter.  To test the hypothesis, we tuned down SPP to the
-minimum possible threshold, with the effect of increasing prefetch suggestions
-made by SPP by XX\%.  \textit{<Comparison of SPP-Unleased with SPP / BOP>}
-Obviously this came at a cost of increased DRAM traffic and cache pollution.
+% EB: This remains one of the untouched subsections of the paper
 
-Moreover, the on-line confidence mechanism introduced in SPP was very
-rudimentary.  It was based on taking a ratio C\textsubscript{d} =
-C\textsubscript{delta} / C\textsubscript{sig} as explained previously. This
-confidence was used to make the decision of whether to prefetch or not to
+As was noted in Figure~\ref{Fig:Motivation}, an aggressive lookahead 
+prefetching if done without any accuracy check, can harm the performance 
+of the system. In this given case, the effective performance gained
+by the original prefetcher is lost by almost 9\% by the aggressive version 
+of the same prefetcher. Most modern prefetchers are getting increasingly 
+aggressive as that enables them to capture the complex memory access patterns 
+seen in pointer-chasing applications. Thus, we need a mechanism that is 
+orthogonal to the underlying prefetching scheme and can be used
+to prune out the useful prefetcher from the useless ones.
+
+Moreover, the on-line confidence mechanism used by most prefetchers is very
+rudimentary. For example, SPP's internal confidence mechanism is based on taking 
+the ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}.
+This confidence was used to make the decision of whether to prefetch or not to
 prefetch; and which level to prefetch.  While this approximation was shown to
@@
 
-Perceptrons in microarchitecture were first introduced by Jimenez and Lin ~\cite{Perc_Branch} for branch prediction.
-In this work, a perceptron weights vector is selected by a hash of the branch address and its dot product with an input vector of branch history outcomes.
-If the product is at least 0, the branch is predicted as taken, otherwise it is predicted not taken.
-The weights are updated with the perceptron update rule: if the prediction is incorrect, or the dot product does not exceed some threshold magnitude, then the weights are incremented or decremented based on their correlation to the corresponding history bits.
-Perceptron learning yields to superior accuracy, which results in increased opportunity for optimizations.
+Perceptrons in microarchitecture were first introduced by Jimenez and Lin~\cite{Perc_Branch} 
+for branch prediction. In this work, a perceptron weights vector is selected by a hash of 
+the branch address and its dot product with an input vector of the previous branch history 
+outcomes. If the product is at least 0, the branch is predicted as taken, otherwise it is 
+predicted not taken. The weights are updated with the perceptron update rule: if the prediction 
+is incorrect, or the dot product does not exceed some threshold magnitude, then the weights 
+are incremented or decremented based on their correlation to the corresponding history bits. 
+Perceptron learning yields to superior accuracy, which results in increased opportunity for 
+optimizations.
+
+\begin{figure}[ht]
+\includegraphics[width=\columnwidth]{Perc_Model}
+\caption{The Basic Perceptron Model}
+\label{Fig:Perc_Model}
+\end{figure}
 
@@
 
-In a perceptron-based mechanism, a uniform
-perceptron update principle is followed.  Perceptrons are a simple model of neurons
-in neural networks \cite{perceptrons1,perceptrons2} modeled by vectors of signed weights learned through online training.
-The output of a perceptron is the dot product of the weights and a vector of inputs.
-In this work, we make use of the perceptron learning algorithm.
-There are two components to the abstract perceptron learning algorithm:
+In a perceptron-based mechanism, a uniform perceptron update principle is followed. As
+shown in the figure~\ref{Fig:Perc_Model} Perceptrons 
+are a simple model of neurons in neural networks \cite{perceptrons1,perceptrons2} modeled by 
+vectors of signed weights learned through online training. The output of a perceptron is the 
+dot product of the weights and a vector of inputs. In this work, we make use of the perceptron 
+learning algorithm. There are two components to the abstract perceptron learning algorithm:
 

### 2018-12-05 conc.tex
Adding Perc image, motivation section

@@
 The perceptron filter acts independently as a check on the underlying prefetch
-engine.  PPF improves performance by over XX\% over the baseline,
-corresponding to YY\% over the next best performing prefetcher.  We believe
-that this is a robust and adaptable technique that can be used to enhance any
-existing or upcoming prefetchers.
+engine.  PPF improves performance of the base prefetcher used in isolation, 
+by 11.9\%.  We believe that this is a robust and adaptable technique 
+that can be used to enhance any existing and upcoming prefetchers.

### 2018-12-05 back.tex
BG and Related

@@
 
-The idea of prefetching begins with Jouppi's \textit{instruction stream
-buffers}. Early prefetchers detected stride access patterns in order to
+The idea of prefetching begins with Jouppi's \textit{Instruction Stream
+Buffers}~\cite{ISB}. Early prefetchers detected stride access patterns in order to
 predict future memory requests~\cite{Smith,Baer,Stride}.  Modern prefetching
@@
 
-An optimal prefetcher should be able to detect a wide range of memory access
-patterns. Simple stride prefetching techniques only detect sequences of
-addresses that differ by a constant value and fail to capture diverse delta
-patterns. Lookahead prefetchers learn patterns by collecting histories of
-observed data access patterns, and correlating these with the strides, by
-considering each new page in a vacuum. Kim {\em et. al.} proposed
+%% Simple stride prefetching techniques only detect sequences of addresses that differ by a constant value and fail to capture diverse memory access patterns. 
+Lookahead prefetchers can learn complicated patterns by collecting histories of
+observed data access patterns.
+%EB: Didn't really understand this: 
+%% "and correlating these with the strides, by considering each new page in a vacuum."
+Kim {\em et. al.} proposed
 SPP~\cite{SPP}, a confidence-based lookahead prefetcher. SPP creates a

### 2018-12-05 related.tex
BG and Related

--- 2018-12-05 related.tex
@@
 
-Perceptron-based prediction mechanisms have proven to yield superior accuracy, which represents an opportunity for optimizations in cache management. 
-Khan \textit{et.
-al.} proposed SDBP \cite{sdbp}, in this work, a sampler structure keeps partial tags of sampled sets separate from the cache.
-Three tables of two-bit saturating counters are accessed using a technique similar to a skewed branch predictor \cite{Piece_Linear}.
-For each hit to a block in the sampled set, the program counter (PC) of the relevant memory instruction is hashed into the three tables and the corresponding counters are decremented.
-For each eviction from a sampled set, the counters corresponding to the PC of the last instruction to access the victim block are incremented.
-For an LLC access, the predictor is consulted by hashing the PC of the memory access instruction into the three tables and taking the sum of the indexed counters.
-When the sum exceeds some threshold, the accessed block is predicted to be dead.
-Tags in sampled sets are managed with true LRU and a reduced associativity, but the LLC may be managed by any policy.
-The paper applies its predictor to replacement and bypass.
+Perceptron-based prediction mechanisms have proven to yield superior accuracy, 
+which represents an opportunity for optimizations in cache management. 
+Khan \textit{et. al.} proposed SDBP~\cite{sdbp}, in this work, a sampler 
+structure keeps partial tags of sampled sets separate from the cache. 
+Three tables of two-bit saturating counters are accessed using a technique 
+similar to a skewed branch predictor~\cite{Piece_Linear}. For each hit to 
+a block in the sampled set, the program counter (PC) of the relevant memory 
+instruction is hashed into the three tables and the corresponding counters 
+are decremented. For each eviction from a sampled set, the counters corresponding 
+to the PC of the last instruction to access the victim block are incremented. 
+For an LLC access, the predictor is consulted by hashing the PC of the memory 
+access instruction into the three tables and taking the sum of the indexed 
+counters. When the sum exceeds some threshold, the accessed block is 
+predicted to be dead. Tags in sampled sets are managed with true LRU and 
+a reduced associativity, but the LLC may be managed by any policy. 
+The paper applies its predictor to do block replacement and bypass for LLC.
 
-Teran \textit{et al.} propose using perceptron learning for reuse prediction driving a bypass and replacement optimization \cite{Perc_Reuse}.
-In that work, perceptron learning is used to set weights selected by hashes of multiple features consisting of the PC, several recent PCs, and two different shifts of the tag of the referenced block.
-Each feature is hashed to index its own table of weights.
-The selected weights are summed and thresholded to make a prediction.
-When a sampled block is reused or evicted, the corresponding weights are decremented or incremented, respectively, according to the perceptron learning rule.
+Teran \textit{et al.} propose using perceptron learning for reuse prediction 
+driving a bypass and replacement optimization~\cite{Perc_Reuse}. In that work, 
+perceptron learning is used to set weights selected by hashes of multiple features 
+consisting of the PC, several recent PCs, and two different shifts of the tag 
+of the referenced block. Each feature is hashed to index its own table of weights. 
+The selected weights are summed and thresholded to make a prediction. When a sampled 
+block is reused or evicted, the corresponding weights are decremented or incremented, 
+respectively, according to the perceptron learning rule.
 
-The continuation of Perceptron was later introduced in the work, Multiperspective Reuse Prediction \cite{Multiperspective}.
-Compared to previous work that focuses on one or two features at a time, Multiperspective prediction uses many features, each contributing to the overall prediction.
-The prediction enables placement, promotion and bypass optimizations that improves performance.
+The continuation of Perceptron was later introduced in the work, Multiperspective Reuse 
+Prediction~\cite{Multiperspective}. Compared to previous work that focuses on one or two 
+features at a time, Multiperspective prediction uses many features, each contributing 
+to the overall prediction. The prediction enables placement, promotion and bypass 
+optimizations that further improves performance.
 
@@
 
-Two widely used prefetchers are Next-n Lines \cite{nextnlines} and Stride \cite{stride}, both of which capture regular memory access patterns with very low hardware overhead.
-The Next-n Lines prefetcher simply queues prefetches for the next n lines after any given miss, under the expectation that the principle of spatial locality will hold and those cache lines after a missed line are likely to be used in the future.
-The stride prefetcher is slightly more sophisticated; it attempts to identify simple stride reference patterns in programs based upon the past behavior of missing loads.
-Similar to the Next-n Lines technique, when a given load misses, cache lines ahead of that miss are fetched in the pattern following the previous behavior in the hope of avoiding future misses.
-Both prefetchers have been widely used due to the simple design and low hardware overhead.
-However, without further knowledge about temporal locality and application characteristics, these prefetchers cannot do more than detecting and prefetching regular memory access patterns with limited spatial locality.
+Two widely used prefetchers are Next-n Lines~\cite{nextn} and Stride~\cite{stride}
+prefetchers, both of which capture regular memory access patterns with very low hardware overhead. 
+The Next-n Lines prefetcher simply queues prefetches for the next n lines after any given miss, 
+under the expectation that the principle of spatial locality will hold and those cache lines after 
+a missed line are likely to be used in the future. The stride prefetcher is slightly more 
+sophisticated -- it attempts to identify simple stride reference patterns in programs based 
+upon the past behavior of missing loads. Similar to the Next-n Lines technique, 
+when a given load misses, cache lines ahead of that miss are fetched in the pattern 
+following the previous behavior in the hope of avoiding future misses. Both prefetchers 
+have been widely used due to the simple design and low hardware overhead. However, 
+without further knowledge about temporal locality and application characteristics, 
+these prefetchers cannot do more than detecting and prefetching regular memory access 
+patterns with limited spatial locality.
 
-Somogyi et al.
-proposed one of the top practical, low-overhead prefetchers, the Spatial Memory Streaming (SMS) prefetcher \cite{sms}.
-SMS leverages code-based correlation to take advantage of spatial locality in the applications over larger regions of memory (called spatial regions).
-It predicts the future access pattern within a spatial region around a miss, based on a history of access patterns initiated by that missing instruction in the past.
-While the SMS prefetcher is effective, it is indirectly inferring future program control-flow when it speculates on the misses in a spatial region.
-As a result, the state overheads of this predictor can be higher than the others in this class.
+Somogyi et al. proposed one of the top practical, low-overhead prefetchers, the Spatial 
+Memory Streaming (SMS) prefetcher~\cite{SMS}. SMS leverages code-based correlation 
+to take advantage of spatial locality in the applications over larger regions of memory 
+(called spatial regions). It predicts the future access pattern within a spatial 
+region around a miss, based on a history of access patterns initiated by that 
+missing instruction in the past. While the SMS prefetcher is effective, it is 
+indirectly inferring future program control-flow when it speculates on the misses 
+in a spatial region. As a result, the state overheads of this predictor can 
+be higher than the others in this class. This idea was further extended by 
+the class of related prefetchers~\cite{STMS,Temporal_Instruction_Fetch}.
+
+While previous work had proposed dynamically finding the best prefetch offset, they have failed to 
+look into prefetch timeliness. Michaud proposed the Best-Offset prefetcher~\cite{BOP}. 
+A prefetch offset is set automatically and dynamically tries to adapt to the application behavior. 
+A good prefetch offset is determined by looking at two recent access to a particularly line 
+that did not happen too recently. The time between both accesses should be 
+greater than the latency for completing a prefetch request.
 
 \subsection{Lookahead Prefetchers}
+While the stride and the next-n perefetchers are efficient, they suffer from their limited
+ability to capture memory delta patterns. Moreover, a good prefetch proves to be useful only 
+if the prefetcher manages to bring the block before its next demand access. Both these issues 
+can be overcome by Lookahead Prefetchers. These prefetchers try to use their own predictions 
+to create the spelative path of the program's memory access pattern. By going down the speculation
+trail till a certain depth, the prefetcher can bring in cache lines well before they are demand 
+accessed by the processor. VLDP~\cite{VLDP} prefetches a static depth ahead of the demand fetch, 
+without considering the prefetching confidence. More recent Lookahead Prefetchers include SPP and
+KPC~\cite{KPC}
 
-To address both prefetching coverage and accuracy, prior work has adopted lookahead 
-mechanisms. Lookahead prefetchers efficiently encode the relationship between accesses to yield future predictions, enabling further speculative lookahead accesses.
-VLDP \cite {vldp} prefetches a static depth ahead of the demand fetch, without considering the prefetching confidence,
-
-DA-AMPM
-
-While previous work had proposed dynamically finding the best prefetch offset, they have failed to look into prefetch timeliness.
-Michaud proposed the Best-Offset prefetcher \cite{BOP}.
-A prefetch offset is set automatically and dynamically tries to adapt to the application behavior.
-A good prefetch offset is determined by looking at two recent access to a particularly line that do not happen too recently.
-The time between both accesses should be greater than the latency for completing a prefetch request.
+Talk about DA-AMPM
 
@@
 A low-accuracy aggressive prefetcher can significantly harm performance. 
-In order to minimize the interference from prefetching, Wu \textit{et.
-al} proposed PACMan ~\cite{pacman}, a prefetch-aware cache management policy.
-PACMan dedicated a few sets of the LLC to each of the three competing policies that treat demand and prefetch request differently and uses the policy that shows the lowest number of cache misses.
-Similarly, Seshadri  \textit{et.
-al.} propose ICP \cite{icp}.
-ICP is also designed as a comprehensive mechanism to mitigate prefetching interference.
-ICP simply demotes a prefetch blow to the lowest priority on a demand hit based on the observation that most prefetches are dead after their first hit.
-To address prefetcher-caused cache pollution, it also uses a variation of EAF [27] to track prefetching accuracy and inserts only an accurate prefetch to the higher priority position in the LRU stack.
-ICP assumes, however, that all prefetches are inserted only into the LLC, which restricts the maximum benefit of prefetching.
+In order to minimize the interference from prefetching, Wu \textit{et. al} proposed PACMan~\cite{pacman}, 
+a prefetch-aware cache management policy. PACMan dedicated a few sets of the LLC to each of the three 
+competing policies that treat demand and prefetch request differently and uses the policy that 
+shows the lowest number of cache misses. Similarly, Seshadri  \textit{et. al.} proposed ICP~\cite{icp}. 
+ICP is also designed as a comprehensive mechanism to mitigate prefetching interference. It simply 
+demotes a prefetch to the lowest priority on a demand hit based on the observation that most 
+prefetches are dead after their first hit. To address prefetcher-caused cache pollution, it also 
+uses a variation of EAF~\cite{eaf} to track prefetching accuracy and inserts only an accurate 
+prefetch to the higher priority position in the LRU stack. ICP assumes, however, that all 
+prefetches are inserted only into the LLC, which restricts the maximum benefit of prefetching.
+
+\subsection{Machine Learning for Prefetching}
+Peled \textit{et.al.} introduced
+interesting ideas for on-line Reinforcement Learning and dynamically scaling
+the magnitude of feedback given to the baseline prefetcher~\cite{Semantics}.
+The biggest challenge here is that the prefetcher relies on compiler support for
+getting features to build the context.
+
+Liao \textit{et. al.} focused on prefetching for data center
+applications~\cite{Datacenter}. They used offline machine learning algorithms
+such as SVMs and logistic regression to do a parametric search for an optimal
+prefetcher configuration.

### 2018-12-05 back.tex
went through with a chainsaw and hacked off large chunks, then fine-tuned a lot of words. added Elvira's IPC graphs

@@
 Moreover, the on-line confidence mechanism used by most prefetchers is very
-rudimentary. For example, SPP's internal confidence mechanism is based on taking 
-the ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}.
-This confidence was used to make the decision of whether to prefetch or not to
-prefetch; and which level to prefetch.  While this approximation was shown to
-work in the original implementation, we believe that a better form of
-generalized on-line decision making was possible.  Hence, it was necessary to
-build a robust and adaptable learning mechanism to accept / reject the
-prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level
-Cache).
-
-To that effect, we introduce an independent on-line perceptron based
-filtering mechanism.
+rudimentary. For example, SPP's internal confidence mechanism is based on
+taking the ratio C\textsubscript{d} = C\textsubscript{delta} /
+C\textsubscript{sig}.  This confidence was used to make the decision of
+whether to prefetch or not to prefetch; and which level to prefetch.  While
+this approximation was shown to work in the original implementation, we
+believe that a better form of generalized on-line decision making was
+possible.  Hence, it was necessary to build a robust and adaptable learning
+mechanism to accept / reject the prefetch suggestions; and to decide the fill
+level (L2 Cache vs Last Level Cache).  Thus, we introduce an independent
+on-line perceptron based filtering mechanism.
 
@@
 %figure out what to leave and what to get rid of ...  
-\subsection{Perceptron-based Prediction}
+% djimenez: I ended up cutting everything you wrote as well as the figure. I
+% wrote something similar above, and the perceptron model described here isn't
+% the hashed perceptron that we're using but rather the dot-product model from
+% the original paper.
 
-Perceptrons in microarchitecture were first introduced by Jimenez and Lin~\cite{Perc_Branch} 
-for branch prediction. In this work, a perceptron weights vector is selected by a hash of 
-the branch address and its dot product with an input vector of the previous branch history 
-outcomes. If the product is at least 0, the branch is predicted as taken, otherwise it is 
-predicted not taken. The weights are updated with the perceptron update rule: if the prediction 
-is incorrect, or the dot product does not exceed some threshold magnitude, then the weights 
-are incremented or decremented based on their correlation to the corresponding history bits. 
-Perceptron learning yields to superior accuracy, which results in increased opportunity for 
-optimizations.
+%\subsection{Perceptron-based Prediction}
+%
+%Perceptrons in microarchitecture were first introduced by Jimenez and Lin~\cite{PerceptronPredictor} 
+%for branch prediction. In this work, a perceptron weights vector is selected by a hash of 
+%the branch address and its dot product with an input vector of the previous branch history 
+%outcomes. If the product is at least 0, the branch is predicted as taken, otherwise it is 
+%predicted not taken. The weights are updated with the perceptron update rule: if the prediction 
+%is incorrect, or the dot product does not exceed some threshold magnitude, then the weights 
+%are incremented or decremented based on their correlation to the corresponding history bits. 
+%Perceptron learning yields to superior accuracy, which results in increased opportunity for 
+%optimizations.
 
-\begin{figure}[ht]
-\includegraphics[width=\columnwidth]{Perc_Model}
-\caption{The Basic Perceptron Model}
-\label{Fig:Perc_Model}
-\end{figure}
+%\begin{figure}[ht]
+%\includegraphics[width=\columnwidth]{Perc_Model}
+%\caption{The Basic Perceptron Model}
+%\label{Fig:Perc_Model}
+%\end{figure}
 
-\subsubsection{Perceptron Learning}
-
-In a perceptron-based mechanism, a uniform perceptron update principle is followed. As
-shown in the figure~\ref{Fig:Perc_Model} Perceptrons 
-are a simple model of neurons in neural networks \cite{perceptrons1,perceptrons2} modeled by 
-vectors of signed weights learned through online training. The output of a perceptron is the 
-dot product of the weights and a vector of inputs. In this work, we make use of the perceptron 
-learning algorithm. There are two components to the abstract perceptron learning algorithm:
-
-\begin{enumerate}
-\item To predict true or false, a vector of signed weights is chosen according to some criteria. 
-The dot product of the weights and input vectors, called $y_{\mbox{\small out}}$, is computed. 
-If $y_{\mbox{\small out}}$ exceeds some threshold, the prediction is true, otherwise it is false.
-
-\item To update the weights after the true outcome of the predicted event is
-known, we first consider the value of $y_{\mbox{\small out}}$. If the
-prediction was correct and $|y_{\mbox{\small out}}|$ exceeds a threshold
-$\theta$, then the weights remain unchanged. Otherwise, the inputs are used to
-update the corresponding weights.  If there is positive correlation between
-the input and the outcome of the event, the corresponding weight is
-incremented; otherwise, it is decremented. Over time, the weights are
-proportional to the probability that the outcome of the event is true in the
-context of the input and the criterion for choosing that weight. The weights
-saturate at maximum and minimum values so they will fit in a fixed bit width.
-\end{enumerate}
+%\subsubsection{Perceptron Learning}
+%
+%In a perceptron-based mechanism, a uniform perceptron update principle is
+%followed. As shown in the figure~\ref{Fig:Perc_Model} Perceptrons are a simple
+%model of neurons in neural networks \cite{perceptrons1,perceptrons2} modeled
+%by vectors of signed weights learned through online training. The output of a
+%perceptron is the dot product of the weights and a vector of inputs. In this
+%work, we make use of the perceptron learning algorithm. There are two
+%components to the abstract perceptron learning algorithm:
+%
+%\begin{enumerate}
+%\item To predict true or false, a vector of signed weights is chosen according to some criteria. 
+%The dot product of the weights and input vectors, called $y_{\mbox{\small out}}$, is computed. 
+%If $y_{\mbox{\small out}}$ exceeds some threshold, the prediction is true, otherwise it is false.
+%
+%\item To update the weights after the true outcome of the predicted event is
+%known, we first consider the value of $y_{\mbox{\small out}}$. If the
+%prediction was correct and $|y_{\mbox{\small out}}|$ exceeds a threshold
+%$\theta$, then the weights remain unchanged. Otherwise, the inputs are used to
+%update the corresponding weights.  If there is positive correlation between
+%the input and the outcome of the event, the corresponding weight is
+%incremented; otherwise, it is decremented. Over time, the weights are
+%proportional to the probability that the outcome of the event is true in the
+%context of the input and the criterion for choosing that weight. The weights
+%saturate at maximum and minimum values so they will fit in a fixed bit width.
+%\end{enumerate}

### 2018-12-05 conc.tex
went through with a chainsaw and hacked off large chunks, then fine-tuned a lot of words. added Elvira's IPC graphs

@@
 \label{Conclusion}
-In this paper, we introduced perceptron-based learning for data prefetching.
-The perceptron filter acts independently as a check on the underlying prefetch
-engine.  PPF improves performance of the base prefetcher used in isolation, 
-by 11.9\%.  We believe that this is a robust and adaptable technique 
-that can be used to enhance any existing and upcoming prefetchers.
+In this paper, we introduce perceptron-based learning for prefetching.  The
+perceptron filter acts independently as a check on the underlying prefetch
+engine.  PPF improves performance over the base prefetcher by 11.9\%.  We
+believe that this is a robust and adaptable technique that can be used to
+enhance any existing and upcoming prefetchers.

### 2018-12-05 design_arch.tex
went through with a chainsaw and hacked off large chunks, then fine-tuned a lot of words. added Elvira's IPC graphs

@@
   \begin{center}
-  \includegraphics[width=9cm]{PPF_Hierarchy.pdf}
+  \includegraphics[width=8cm]{PPF_Hierarchy.pdf}
     \label{fig:PPF_Hierarchy}
@@
 
-As mentioned before, ultimately it might be beneficial to allow a
-prefetcher to speculate as deeply as possible. Often, some useful
-prefetches are generated long after the confidence of the prefetcher has
-fallen below a point, after which the performance degrades, due to the
-increase of inaccurate prefetches.  However, in order to allow deep
-speculation in the prefetcher, inaccurate prefetches must be
-filtered out.  We propose to leverage perceptron-based learning as a
+It can be beneficial to allow a prefetcher to speculate as deeply as
+possible. Often, some useful prefetches are generated long after the
+confidence of the prefetcher has fallen below a point, after which the
+performance degrades, due to the increase of inaccurate prefetches.  However,
+in order to allow deep speculation in the prefetcher, inaccurate prefetches
+must be filtered out.  We propose to leverage perceptron-based learning as a
 mechanism to differentiate between potentially useful deeply speculated
-prefetches and likely not-useful ones. The Perceptron Prefetch Filter (PPF)
-is placed between the prefetcher and the prefetch insertion queue, as illustrated
-in figure ~\ref{fig:PPF_Hierarchy}, to prevent
-not-useful prefetches from polluting the higher levels of the memory hierarchy.
+prefetches and likely not-useful ones. The Perceptron Prefetch Filter (PPF) is
+placed between the prefetcher and the prefetch insertion queue, as illustrated
+in figure ~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
+polluting the higher levels of the memory hierarchy.
 
-As discussed in Section~\ref{sec:Background-Perceptron}, perceptron learning
-is a light-weight mechanism to pull together disparate forms of information
-and synthesize a decision from them.  Our work considers a number of
-features corresponding to a prefetch, such as speculation
-depth, page address and offset, and uses this information as the inputs to our 
-perceptron-based filter, in order to predict the usefulness of a prefetch.
-In the remainder of this section, we discuss our proposed technique -- PPF.
-It enhances a base prefetcher, by filtering out not-useful prefetches. 
-PPF is a generalized prefetch filtering mechanism that may be adapted 
-to any prefetcher with appropriate modifications.
+Perceptron learning is a light-weight mechanism to pull together disparate
+forms of information and synthesize a decision from them.  Our work considers
+a number of features corresponding to a prefetch, such as speculation depth,
+page address and offset, and uses this information as the inputs to our
+perceptron-based filter, in order to predict the usefulness of a prefetch.  In
+the remainder of this section, we discuss our proposed technique: PPF.  It
+enhances a base prefetcher, by filtering out not-useful prefetches.  PPF is a
+generalized prefetch filtering mechanism that may be adapted to any prefetcher
+with appropriate modifications.
 
@@
 
-Figure~\ref{fig:PPF_Datapath} shows the required structures of PPF, as well as the
-steps followed to filter out not-useful prefetches.
-The filter is organized as a set of tables, where each entry in the tables holds 
-a \textit{weight}. For a configuration of PPF using $N$ number of features, 
-$N$ different tables of weights are needed. Each feature is used to index 
-a distinct table. The number of entries of each table varies according
-to their corresponding feature, hence, different
-number of bits are needed to index different tables.
-Each weight is a 5-bit saturating counter ranging from -16 to +15. 
-We found that having 5-bit weights provide a good trade-off 
-between accuracy and area. A detail explanation of the storage overhead of
-PPF can be found in Section~\ref{sec:Method-Overheads}.
+Figure~\ref{fig:PPF_Datapath} shows the required structures of PPF, as well as
+the steps followed to filter out not-useful prefetches.  The filter is
+organized as a set of tables, where each entry in the tables holds a
+\textit{weight}. For a configuration of PPF using $N$ number of features, $N$
+different tables of weights are needed. Each feature is used to index a
+distinct table. The number of entries of each table varies according to their
+corresponding feature, hence, different number of bits are needed to index
+different tables.  Each weight is a 5-bit saturating counter ranging from -16
+to +15.  We found that having 5-bit weights provide a good trade-off between
+accuracy and area. A detailed explanation of the storage overhead of PPF can
+be found in Figure~\ref{tab:PPF_overhead}.
 
@@
 
-\noindent{\textbf{1. Inference}}\newline
-
+\noindent{\textbf{Inference}\\
 %ETG:is this like an explanation of how prefetching happens? 
 %EB: I see this as a short context of the datapath leading to the perc filter.
+The base prefetcher is triggered on every demand access to the L2 Cache.  At
+this point, it has the opportunity to trigger a prefetch. If it decides to do
+so, it will also need to decide how many cache blocks to prefetch.  These
+blocks can be either placed in the L2 or L3 cache according to the confidence
+of the prefetching mechanism.  Once the base prefetcher is triggered, the
+suggested prefetch candidates are fed to the perceptron filter to determine
+the usefulness of these prefetches.  The filter makes the decision to issue
+the prefetch suggestions from the base prefetcher.  To make the decision, each
+feature corresponding to a suggested prefetch is used to index a table and all
+the corresponding weights are summed.  The sum denotes the confidence value
+for the suggested prefetch, and is thresholded against two different
+magnitudes: PERC\_THRESHOLD\_HI and PERC\_ THRESHOLD\_ LO.
 
-The base prefetcher is triggered on every demand access to the L2 Cache.  
-As this point, it has the opportunity to trigger a prefetch  -- if it 
-decides to do so, it will also need to decide how many cache lines
-to prefetch.  These cache lines can be either placed in the L2 Cache or
-the L3 Cache, according to the confidence of the prefetching mechanism.
-Once the base prefetcher is triggered, the suggested
-prefetch candidates are fed to the perceptron filter, 
-in order to determine the usefulness of these prefetches.  
-The filter makes the decision to issue the prefetch
-suggestions from the base prefetcher.
-In order to make such decision, each feature corresponding to a 
-suggested prefetch is used to index a table, 
-all the corresponding weights are summed. 
-The sum denotes the confidence value for the suggested prefetch, 
-and is thresholded against two different magnitudes:  
-PERC\_THRESHOLD\_HI and PERC\_ THRESHOLD\_ LO.
+Prefetches whose sum exceeds PERC\_THRESHOLD\_HI are placed into the L2 cache.
+The higher confidence value hints the prefetch would be useful and should be
+prioritized.  A prefetch for which the features result in a confidence value
+between PERC\_THRESHOLD\_LO and PERC\_THRESHOLD\_HI is allocated in the larger
+LLC, as the filter is moderately confident of the future reuse of the cache
+block, but not enough to possible pollute a significantly smaller L2.
+Suggested prefetches for which the features lead to a confidence value lower
+than PERC\_THRESHOLD\_LO are not prefetched, as the low confidence value
+represents that the perceptron has learn that similar set of features are
+associated with non-useful prefetches. \newline
 
-Prefetches which sum is above PERC\_THRESHOLD\_HI
-are placed into the L2 cache, the higher confidence value
-hints the prefetch would be useful and should be prioritized.
-A prefetch for which its corresponding features result 
-in a confidence value between  
-PERC\_THRESHOLD\_LO and PERC\_THRESHOLD\_HI is allocated in the larger 
-last-level cache, as the filter is moderately confident of the future reuse of the 
-cache line, but not enough to possible pollute a significantly smaller L2. 
-Suggested prefetches for which the features lead to a confidence value lower than 
-PERC\_THRESHOLD\_LO are not prefetched, as the low confidence value 
-represents that the perceptron has learn that similar set of features
-are associated with non-useful prefetches. \newline
+\noindent{2. Recording}\\
+As shown in step 2 of figure ~\ref{fig:PPF_Datapath} the prefetches that make
+it through the inference stage are recorded in the ``Prefetch Table''. The
+prefetch table is a 1,024-entry, direct mapped structure that contains all
+metadata required to re-index the perceptron entries for training.
 
-\noindent{\textbf{2. Recording}}\newline
+In addition to the prefetch table mentioned above, PPF also maintains a
+1,024-entry direct-mapped ``Reject Table.'' If a prefetch suggestion is
+rejected by the perceptron layer, it is logged into the reject table. The
+table is used to train the perceptron to avoid false negatives \textit{i.e.},
+cases where the prediction suggested to reject the prefetch but the prefetch
+was proven to be useful based on the observed demand accesses to the L2. \\
 
-As shown in step 2 of figure ~\ref{fig:PPF_Datapath}
-the prefetches that make it through the inference stage are recorded in the 
-``Prefetch Table''. The prefetch table is a 1,024-entry, direct mapped 
-structure that contains all metadata required to re-index the perceptron 
-entries for training.
-
-In addition to the prefetch table mentioned above, 
-PPF also maintains a 1,024-entry direct-mapped ``Reject Table.'' 
-If a prefetch suggestion is rejected by the perceptron layer, it is logged into 
-the reject table. The table is used to train the perceptron to avoid false negatives 
-\textit{i.e.}, cases where the prediction suggested to reject the prefetch 
-but the prefetch was proven to be useful based 
-on the observed demand accesses to the L2. \newline
-
-\noindent{\textbf{3. Feedback and Data Retrieval}} \newline
-
-When there is an eviction or a demand access to the L2, training for both the base 
-prefetcher and our filter mechanism is triggered. 
-The address of the cache line that triggered training is used to index
-both the Prefetch and Reject tables. 10 bits of the address are used to index 
-tables, and another 6 bits to perform tag matching. 
-Once the base prefetcher has suggested a prefetch request, the corresponding 
-features are used to index the tables of weights. \newline
+\noindent{\textbf{Feedback and Data Retrieval}\\
+When there is an eviction or a demand access to the L2, training for both the
+base prefetcher and our filter mechanism is triggered.  The address of the
+cache block that triggered training is used to index both the Prefetch and
+Reject tables. 10 bits of the address are used to index tables, and another 6
+bits to perform tag matching.  Once the base prefetcher has suggested a
+prefetch request, the corresponding features are used to index the tables of
+weights. \newline
 
@@
 
-\noindent{\textbf{4. Training}}\newline
+\noindent{\textbf{Training}\newline
+The address from the demand request triggering the training is looked up in
+both tables. If the address is in the prefetch table and marked as valid, this
+hints the previous prediction was correct and this is a useful prefetch. We
+compute the sum of the corresponding weights.  If the sum falls below a
+specific threshold, training occurs and the corresponding weights are adjusted
+accordingly.  These thresholds are introduced in order to avoid over-training,
+helping the filter adapt quickly to changes in memory behavior.  These
+thresholds are referred to as POS\_UPDATE\_THRESHOLD ($\theta_p$) and \newline
+NEG\_UPDATE\_THRESHOLD ($\theta_n$), respectively for the positive and
+negative values of training saturation.
 
-The address from the demand request that triggered the training is lookup in 
-both tables. If the address resides in the prefetch table, and is marked 
-as valid, this hints the previous prediction was correct and this is a useful
-prefetch. We compute the sum of the corresponding weights and
-compare the sum to an specific threshold, if the sum falls below it, 
-training must happen and the corresponding weights are adjusted accordingly.
-These thresholds are introduced in order to avoid over-training, 
-which helps the filter adapt quickly to changes in memory behavior. 
-These thresholds are referred to as
-POS\_UPDATE\_THRESHOLD ($\theta_p$) and \newline
-NEG\_UPDATE\_THRESHOLD ($\theta_n$),
-respectively for the positive and negative values of training saturation.
-
-On an eviction of a cache block, we look up the corresponding address
-in the prefetch table. If there is a valid entry with this address, 
-this mean the filter made a misprediction. The block was allocated in the
-L2 with a prefetch request that the filter should have categorized 
-as a useless prefetch. In this scenario, the corresponding features of the
-prefetch request that cause these block to be placed in the cache are used
-to index the tables of weights, and those weights are adjusted accordingly.
+On a cache block eviction, we look up the corresponding address in the
+prefetch table. If there is a valid entry with this address, the filter made a
+misprediction. The block was allocated in the L2 with a prefetch request that
+the filter should have categorized as a useless prefetch.  Thus, the
+corresponding features of the prefetch request that cause these block to be
+placed in the cache are used to index the tables of weights, and those weights
+are adjusted accordingly.
 
@@
 
-Parallel to accessing the prefetch table, on a demand access, 
-the reject table is accessed. Before
-the demand access triggers the next set of prefetches, the reject table is
-checked for a valid entry.  A hit means that the corresponding cache line was
-initially suggested by the base prefetcher, but wrongly rejected by the
-perceptron filter. The perceptron filter learns from this and makes use of the
-corresponding features associated to the original prefetch request, which are
-stored in the reject table, to index the weights tables and adjust the weights
-accordingly.
+Parallel to accessing the prefetch table, on a demand access, the reject table
+is accessed. Before the demand access triggers the next set of prefetches, the
+reject table is checked for a valid entry.  A hit means that the corresponding
+cache block was initially suggested by the base prefetcher, but wrongly
+rejected by the perceptron filter. The perceptron filter learns from this and
+makes use of the corresponding features associated to the original prefetch
+request, which are stored in the reject table, to index the weights tables and
+adjust the weights accordingly.
 
-Prior work in this domain lacks the knowledge of which cache lines should have 
+Prior work in this domain lacks the knowledge of which cache blocks should have 
 been prefetched, but were not. The implementation of the reject table,
@@
 %our PPF implementation are given under ``Area Overheads'' in
-%Section~\ref{Method-Overheads}.
-%
+%Section~\ref{Method-Overheads}.  %
 %%\pg{this whole ``bits'' discussion belongs in implementation not here IMHO}
@@
 In general, PPF can be adapted to a new base prefetcher with only a
-few modifications.
-\begin{itemize}
+few modifications: \\ \\
+%
+\textbf{Enhancing the Base Prefetcher:} By tuning down any internal thresholds
+to increase its inherent aggressiveness. \\
+%
+\textbf{Inferencing and Storing:} All prefetch recommendations are tested
+using the perceptron inferencing algorithm. The perceptron's output,
+\textit{true} or \textit{false}, should be saved appropriately, along with all
+metadata required for perceptron indexing. \\
+%
+\textbf{Retrieving and Training:} When feedback for a prefetch is available,
+the previously stored metadata can be used to re-index into the perceptron
+entries and increment or decrement the weights. \\
+%
+\textbf{Feature Selection:}  Perceptrons essentially normalize contributions
+from different features to get a single sum representing the final confidence.
+So the perceptron learning can be as good as the set of features chosen.
+Interestingly, this is what makes perceptron learning scalable, as it can
+easily learn to incorporate newer information in the form of new features.
+Some of the features we developed use information derived directly from
+program execution, agnostic to the baseline prefetcher. Beyond that, the
+feature set can be expanded to convey any useful information from the
+prefetcher to the perceptron filter.
 
-\item \textbf{Enhancing the Base Prefetcher:} By tuning down any
-  internal thresholds to increase its inherent aggressiveness.
-
-\item \textbf{Inferencing and Storing:} All prefetch recommendations
-  are tested using the perceptron inferencing algorithm. The
-  perceptron's output, \textit{true} or \textit{false}, should be
-  saved appropriately, along with all metadata required for perceptron
-  indexing.
-
-\item \textbf{Retrieving and Training:} When feedback for a prefetch
-  is available, the previously stored metadata can be used to re-index
-  into the perceptron entries and increment or decrement the weights.
-
-\item \textbf{Feature Selection:}  Perceptrons essentially normalize
-  contributions from different features to get a single sum representing the
-  final confidence. So the perceptron learning can be as good as the set of 
-  features chosen. Interestingly, this is what makes perceptron learning 
-  scalable, as it can easily learn to incorporate newer information in 
-  the form of new features. Some of the features we developed use information 
-  derived directly from program execution, agnostic to the baseline 
-  prefetcher. Beyond that, the feature set can be expanded to convey 
-  any useful information from the prefetcher to the perceptron filter.
-
-  %[EB] This statement fits better in the next section 'Features used by Perceptron'
-  %The methodology explained in
-  %Section~\ref{Method-Features} talks about developing a minimal feature 
-  %set for PPF.
+%  %[EB] This statement fits better in the next section 'Features used by
+%  Perceptron' %The methodology explained in %Section~\ref{Method-Features}
+%  talks about developing a minimal feature %set for PPF.
 
@@
 % [EB]: Hopefully this is general enough and not too redundant
-
-\end{itemize}

### 2018-12-05 design_impl.tex
went through with a chainsaw and hacked off large chunks, then fine-tuned a lot of words. added Elvira's IPC graphs

@@
 
-\subsection{Changes made to original SPP}
+\subsection{Changes Made to SPP}
 \label{Impl-Changes}
-To modify the SPP design to suit our scheme, the following changes
-were made:
-
-\begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline In PPF, the
-  perceptron sum is used to decide whether to prefetch or not, and the
-  fill-level in case of prefetch.  Thus, the confidence thresholds --
-  T\textsubscript{f} and T\textsubscript{p} are no longer needed.
-
+To modify the SPP design to suit our scheme, the following changes were made: \\ \\
+%
+\textbf{Original Thresholds Discarded:} In PPF, the perceptron sum is used to
+decide whether to prefetch or not, and the fill-level in case of prefetch.
+Thus, the confidence thresholds -- T\textsubscript{f} and T\textsubscript{p}
+are no longer needed. \\
+% djimenez: keeping comments on every line here so it doesn't screw up my
+%"roll-your-own" itemized list. don't put in any newlines.
+%
+%
 % EB: Can this point be removed?
@@
 % EB: It's ultimately a check to avoid SPP's internal segfault. Won't have a real impact.
-
+%
 %\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The
@@
 %    this a big deal?  Seems like a detail}
-
-\item \textbf{Enhanced Prefetch Table}\newline The prefetch table
-  introduced in the original SPP keeps track of up to 1024 L2 cache
-  prefetch suggestions.  For PPF, the prefetch table is modified to
-  store the features used for perceptron indexing..
-  This metadata is needed to retrieve the
-  values of the features to train the perceptron at a later stage in
-  time. The prefetch table mentioned above keeps track of the
-  prefetches that actually happened in the L2 Cache.  We introduce the
-  reject table to keep a track of the prefetches that were suggested
-  by the base SPP engine but were rejected by PPF.  In case it is
-  detected that a future demand fetch could have used this rejected
-  prefetch, the perceptron weights are updated to reflect that
-  misprediction.
-\end{itemize}
+%
+\textbf{Enhanced Prefetch Table:} The prefetch table introduced in the
+original SPP keeps track of up to 1024 L2 cache prefetch suggestions.  For
+PPF, the prefetch table is modified to store the features used for perceptron
+indexing..  This metadata is needed to retrieve the values of the features to
+train the perceptron at a later stage in time. The prefetch table mentioned
+above keeps track of the prefetches that actually happened in the L2 Cache.
+We introduce the reject table to keep a track of the prefetches that were
+suggested by the base SPP engine but were rejected by PPF.  In case it is
+detected that a future demand fetch could have used this rejected prefetch,
+the perceptron weights are updated to reflect that misprediction.
 
@@
 outlined in Section~\ref{Method-Features}, we pruned the feature 
-set to a minimal yet relevant set of features
+set to a minimal yet relevant set of features. \\ \\
 % \pg{again you want to get at why
@@
 % space of possible features and these were the best.}
-
+%
 % [EB] Tried addressing. Does it look redundant when read after 'Feature Selection'
 % bullet from the previous section?
-
-\begin{itemize}
-\item \textbf{Base Address}: The base address of the demand 
-access that triggers the prefetch corresponds to a stream of 
-future accesses that SPP and the PPF have seen before. 
-Therefore, PPF will correlate the past behavior of this address 
-to the confidence of a prefetch.
-
-\item \textbf{Cache Line} and \textbf{Page Address}: These two
-  separate features are derived from the base address that triggered
-  the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and
-  base\_addr >> LOG2\_PAGE\_SIZE respectively. The idea behind using
-  three different shifted versions of the same feature is that it
-  allows us to look into a wider range of bits than with a single
-  version.  It also helps give more importance to the overlapping bits
-  and lesser importance to most and least significance bits. This
-  approach can also eliminate destructive interference that can be
-  caused by directly folding the address bits into half.
-
+%
+\textbf{Base Address}: The base address of the demand access that triggers the
+prefetch corresponds to a stream of future accesses that SPP and the PPF have
+seen before.  Therefore, PPF will correlate the past behavior of this address
+to the confidence of a prefetch. \\
+%
+\textbf{Cache Line} and \textbf{Page Address}: These two separate features are
+derived from the base address that triggered the prefetch, as follows:
+base\_addr >> LOG2\_BLOCK \_SIZE and base\_addr >> LOG2\_PAGE\_SIZE
+respectively. The idea behind using three different shifted versions of the
+same feature is that it allows us to look into a wider range of bits than with
+a single version.  It also helps give more importance to the overlapping bits
+and lesser importance to most and least significance bits. This approach can
+also eliminate destructive interference that can be caused by directly folding
+the address bits into half. \\
+%
 % djimenez: this is a nice intuition and explanation. 
 % [EB] Thank you!
-
-\item \textbf{Program Counter XOR Depth}: The PC is for the
-  instruction that triggered the prefetch chain.  Depth refers to the
-  iteration count of the lookahead stages.  In general, the PC is not
-  considered as a good basis for doing lookahead prefetching as all
-  the prefetches with depth >= 1 are aliased into the same PC which
-  is not true in an actual demand access.  This feature resolves
-  a PC into a different value for each lookahead depth of prefetch
-  speculation, giving a more accurate correlation in lookahead cases.
-  This is akin to the concept of Virtual Program Counters~\cite{VPC}
-  introduced by Kim \textit{et. al.} for indirect branch prediction.
-
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
+\textbf{Program Counter XOR Depth}: The PC is for the instruction that
+triggered the prefetch chain.  Depth refers to the iteration count of the
+lookahead stages.  In general, the PC is not considered as a good basis for
+doing lookahead prefetching as all the prefetches with depth >= 1 are aliased
+into the same PC which is not true in an actual demand access.  This feature
+resolves a PC into a different value for each lookahead depth of prefetch
+speculation, giving a more accurate correlation in lookahead cases.  This is
+akin to the concept of Virtual Program Counters~\cite{VPC} introduced by Kim
+\textit{et. al.} for indirect branch prediction. \\
+%
+\textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
 PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$
@@
 done to avoid the resultant value of zero when 2 or more PCs are the
-  same.  Additionally, obfuscating the information as it gets old
+same.  Additionally, blurring the information as it gets older
 allows us to get a wider and approximate look into the program's
-  history.
-
-\item \textbf{Program Counter XOR Delta}: This feature tells us if a
-  given PC favors particular value(s) of delta.  As noted earlier,
-  while the PC alone does not convey useful information, this
-  hash resolves the PC into different values based on the tendency of
-  that PC to favor a certain delta.  Thus, the dynamic nature of
-  different instances of the same memory access instruction is 
-  captured here.
+history. \\
+%
+\textbf{Program Counter XOR Delta}: This feature tells us if a given PC favors
+particular value(s) of delta.  As noted earlier, while the PC alone does not
+convey useful information, this hash resolves the PC into different values
+based on the tendency of that PC to favor a certain delta.  Thus, the dynamic
+nature of different instances of the same memory access instruction is
+captured here. \\
+%
  % GAC: Seems more predictable than dynamic if its capturing 
@@
  % if I'm correct
-
+%
  % GAC: Again having trouble phrasing the below point. I was attempting to say
@@
  % correct in the past.
-\item \textbf{Confidence}: The integer confidence, on a scale of 0 to
-  100, that was used in the original SPP design. While the original 
-  confidence does not directly make the decision to prefetch,
-  PPF correlates it to the correctness of a proposed prefetch.
-  While the original SPP may have dismissed a prefetch due to running
-  further into speculation, PPF can use the original confidence as indicator
-  of not only when prefetches become less confident, but also how likely
-  a low confident speculation is correct in the context of other features.
-
-\item \textbf{Page Address XOR Confidence}: This feature scores the
-  tendency of each page to be prefetch friendly or prefetch averse. It
-  helps resolve a page into different entries depending on its
-  confidence for prefetching, which can vary during phases of a
-  program execution.
-
-\item \textbf{Current Signature XOR Delta}: Recall from the discussion
-  of SPP in Section \ref{Background-SPP} that the new signature is
-  generated using the old signature and the current block delta.
+\textbf{Confidence}: The integer confidence, on a scale of 0 to 100, that was
+used in the original SPP design. While the original confidence does not
+directly make the decision to prefetch, PPF correlates it to the correctness
+of a proposed prefetch.  While the original SPP may have dismissed a prefetch
+due to running further into speculation, PPF can use the original confidence
+as indicator of not only when prefetches become less confident, but also how
+likely a low confident speculation is correct in the context of other
+features. \\
+\textbf{Page Address XOR Confidence}: This feature scores the tendency of each
+page to be prefetch friendly or prefetch averse. It helps resolve a page into
+different entries depending on its confidence for prefetching, which can vary
+during phases of a program execution. \\
+%
+\textbf{Current Signature XOR Delta}: Recall from the discussion of SPP in
+Section \ref{Background-SPP} that the new signature is generated using the old
+signature and the current block delta.  The result of this feature is the next
+signature that is predicted to be accessed based on the delta predicted by
+SPP. While ``Current Signature XOR Delta'' is not the exact formula for
+generating the future signature, it gives an approximate idea of the path that
+the combination of these two values can lead to. \\ \\
+%
 %Gino: Added the sentence below, is this correct?
 %[EB]: It gives the next signature instead of next delta
-  The result of this feature is the next signature that is predicted to be
-  accessed based on the delta predicted by SPP. While ``Current
-  Signature XOR Delta'' is not the exact formula for generating the
-  future signature, it gives an approximate idea of the path that the
-  combination of these two values can lead to.
-
+%
 %%%%%%% Below are the features rejected when downsizing 14 -> 9
@@
 %%%%%%%%%%%%%%%%%%%%%%
-
-\end{itemize}
-
+%
 Some of the composite features are derived from simple hashing (XOR)

### 2018-12-05 intro.tex
Adding Perc image, motivation section
went through with a chainsaw and hacked off large chunks, then fine-tuned a lot of words. added Elvira's IPC graphs

@@
 depth coverage. The figure depicts the behavior of the prefetcher variants running 
-on the trace \textit{603.bwaves\_s-1740B}. The IPC, total number of prefetches 
+on the trace {\tt 603.bwaves\_s-1740B}. The IPC, total number of prefetches 
 suggested by the prefetcher (TOTAL\_PF) and the actual useful predictions (GOOD\_PF),
@@
 
-\vspace{1ex}This paper describes PPF, explains its merits, offers analysis, and
-outlines the scope for future research. Its contributions include:
+This paper describes PPF, explains its merits, offers analysis, and
+outlines the scope for future research. Its contributions are:
 
@@
 
-\item An on-line neural model used for hardware data
-  prefetching.  Previous work in this area either relied on program
-  semantics~\cite{Semantics} or were application specific~\cite{Datacenter}.
-% djimenez: removing reference to this arXiv paper. we are not obligated to cite it.
-%, or
-%  could not yield a significant improvement over the baseline prefetcher
-%  used~\cite{BadPerc}.
+\item An on-line neural model used for hardware data prefetching.  Previous
+work in this area either relied on program semantics~\cite{Semantics} or were
+application specific~\cite{Datacenter}.
 
-\item Implementing PPF in a robust and accurate
-  practical prefetching mechanism, yielding a significant performance
-  improvement compared to other state-of-the-art prefetchers.
-  Moreover, PPF learns to adapt itself to shared resource
-  constraints, leading to further increased performance in multi-core and
-  bandwidth-constrained environments.
+\item Implementing PPF filtering a state-of-the-art prefetchier, giving a
+significant performance improvement compared to previous work.  PPF learns to
+adapt itself to shared resource constraints, leading to further increased
+performance in multi-core and bandwidth-constrained environments.
 
-\item A methodology for determining an appropriate set of 
-  features for prediction, regardless of the underlying prefetcher used.
-  More details are explained in Section \ref{Method-Features}.
+\item A methodology for determining an appropriate set of features for
+prediction, regardless of the underlying prefetcher used.  More details are
+explained in Section \ref{Method-Features}.
 
@@
 
-The paper is organized as follows. Section~\ref{Background} describes
-background work and summarizes SPP.  Section~\ref{Arch} gives the 
-architectural overview of PPF. Section~\ref{Impl} discusses the
-implementation of PPF using SPP and the particular features used.
-Section~\ref{Method} gives the evaluation methodology and explores the feature
-space for perceptron learning.  Section~\ref{Results} evaluates the
-performance of the prefetcher and Section~\ref{Conclusion} concludes the
-paper.
+% djimenez - cutting this
+%The paper is organized as follows. Section~\ref{Background} describes
+%background work and summarizes SPP.  Section~\ref{Arch} gives the 
+%architectural overview of PPF. Section~\ref{Impl} discusses the
+%implementation of PPF using SPP and the particular features used.
+%Section~\ref{Method} gives the evaluation methodology and explores the feature
+%space for perceptron learning.  Section~\ref{Results} evaluates the
+%performance of the prefetcher and Section~\ref{Conclusion} concludes the
+%paper.

### 2018-12-05 method.tex
went through with a chainsaw and hacked off large chunks, then fine-tuned a lot of words. added Elvira's IPC graphs

@@
 levels of cache hierarchies. Branch prediction is done using the perceptron
-branch predictor~\cite{Perc_Branch}.  The page size is 4KB.  ChampSim operates
+branch predictor~\cite{PerceptronPredictor}.  The page size is 4KB.  ChampSim operates
 all the prefetchers strictly in the physical address space.
@@
 \label{Method-AdditionalMem}
-The default single-core configuration simulates a 2MB LLC and a single
-channel DRAM with 12.8GB/s bandwidth.  We extend the simulations to
-include memory constraints introduced in DPC-2.  Specifically we look
-at the following two variations:
-\begin{itemize}
-\item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited
-  to 3.2 GB/s
-\item \textit{Small LLC}: In this scenario, LLC size is reduced to 512
-  KB
-\end{itemize}
-All the multi-core simulations are only done in the default
-configuration.
+The default single-core configuration simulates a 2MB LLC and a single channel
+DRAM with 12.8GB/s bandwidth.  We extend the simulations to include memory
+constraints introduced in DPC-2.  Specifically we look at the following two
+variations: Low Bandwidth DRAM, where DRAM bandwidth is limited to 3.2 GB/s,
+and small LLC, where the LLC size is reduced to 512 KB All the multi-core
+simulations are only done in the default configuration.
 
@@
 context.  To study the correlation across each feature, we study statistically
-the perceptron weights and try to interpret their distribution.
-
-\textbf{Global Pearson's Correlation}\newline For this experiment, we create a
-dump of the state of perceptron weights at the end of all trace execution.
-The weights obtained from running all the SPEC CPU 2017 traces are concatenated.
-Since the weights are collected at the end of individual trace execution, the
-perceptron weights have attained a relatively stable value by now.  Hence,
-this dump represents a `snap' of the trained perceptron weights across all the
-SPEC CPU 2017 traces.  The intuition here is that the feature with bulk of the
+the perceptron weights and try to interpret their distribution. \\ \\
+%
+\textbf{Global Pearson's Correlation}\newline For this experiment, we examine
+the perceptron weights at the end of all trace execution by which time the
+weights have settled to steady values.  The weights obtained from running all
+the SPEC CPU 2017 traces are concatenated.  The feature with bulk of the
 perceptron weights concentrated around 0 or small magnitude numbers show a
@@
 
-We plot a histogram for each feature depicting weights distribution
-from -16 to +15 and use this histogram to generate the Pearson's
-correlation factor for that feature.  In statistics, Pearson's factor
-is a numerical measure of the degree of linear correlation between two
-variables.  It ranges from -1 to +1.  The magnitude of Pearson's
-factor depicts the extent of correlation and the sign depicts whether
-it is a positive correlation or a negative correlation.  Values close
-to 0 suggest a low correlation or at times, noisy data.  A value of
-+1/-1 suggests a perfectly linear positive / negative correlation
-respectively.  Figure~\ref{Fig:Hist} depicts the histogram distribution 
-of trained weights, for a feature with high and another with moderate
-Pearson's coefficient.
+We plot a histogram for each feature depicting weights distribution from -16
+to +15 and generate the Pearson's correlation factor for that feature.
+Pearson's factor is a numerical measure ranging from -1 to 1 of the degree of
+linear correlation between two variables.  The magnitude of Pearson's factor
+gives the extent of correlation and the sign indicates whether it is a
+positive correlation or a negative correlation.  Values close to 0 suggest a
+low correlation or at times, noisy data.  A value of +1/-1 suggests a
+perfectly linear positive / negative correlation respectively.
+Figure~\ref{Fig:Hist} depicts the histogram distribution of trained weights,
+for a feature with high and another with moderate Pearson's coefficient.
 
@@
 \textit{Confidence XOR Page address} helps provide
-a correlation to prefetch outcome with a factor of 0.90.
-
+a correlation to prefetch outcome with a factor of 0.90. \\ \\
+%
 \textbf{Per Trace Correlation} \newline Another important way to look
@@
 overall correlation provide useful correlation (magnitude > 0.5) for
-a significant number of traces.
-
-\textbf{Trimming Features Using Correlation} \newline Besides
+a significant number of traces. \\ \\
+%
+\textbf{Trimming Features Using Correlation} \newline Beside
 providing interesting insights into prefetching behavior, P-value can
@@
 
-Here, we introduce the concept of cross-correlation across features.
-Just as we examined correlation of each feature with the final
-outcome, we can also study correlation between the features.  We used
-the above methodology to eliminate features that provide a little or
-no information that has already been captured in other features.
+As we examine correlation of each feature with the final outcome, we also
+study correlation between the features.  We used the above methodology to
+eliminate features providing little information that has already been captured
+in other features.
 
-As a part of feature selection, we initially came up a mix of 23
-features- primary and composite.  By studying cross correlation of
-each of these features against others in a 23x23 matrix, we identified
-pairs of features with correlation factor > 0.9 in magnitude and
-carefully eliminated redundant features.  Using this approach, we
-managed to reduce the feature count to 9.  Thus, in the final
-implementation of PPF, no two features have a high correlation
-between them.  This way we can be sure that each feature is bringing
-in contribution that cannot be captured using other features.
+We initially came up a mix of 23 features- primary and composite.  By studying
+cross correlation of each of these features against others in a 23$\times$23
+matrix, we identified pairs of features with correlation factor > 0.9 in
+magnitude and eliminated redundant features, reducing the feature count to 9.
+Thus, in the final implementation of PPF, no two features have high
+correlation between them.  This way we can be sure that each feature makes a
+contribution that cannot be captured using other features.
 
-Secondly, studying the relative importance of each feature enabled 
-us to vary the number of entries dedicated for each feature.  
-Features with higher Pearson's correlation, \textit{cache line} 
-and \textit{page address}  were given most importance and
-allowed full 12-bits of indexing.  Features like \textit{IP XOR delta} 
-and \textit{IP XOR depth} with a low overall P-value 
-were allocated fewer indexes in the feature table.
+Secondly, studying the relative importance of each feature enabled us to vary
+the number of entries dedicated for each feature.  Features with higher
+correlation, \textit{cache line} and \textit{page address} were given most
+importance and allowed full 12-bits of indexing.  Features like \textit{IP XOR
+delta} and \textit{IP XOR depth} with a low overall P-value were allocated
+fewer entries in the feature table.
 
-To conclude, in the above discussion we justify the features for
-perceptron from a statistical viewpoint.  We also show how this
-information can also be used for prefetcher tuning.  All this study
-was made possible only because we used on-line perceptron learning for
-prefetching and that enabled us to examine the weights in detail.
+%To conclude, in the above discussion we justify the features for
+%perceptron from a statistical viewpoint.  We also show how this
+%information can also be used for prefetcher tuning.  All this study
+%was made possible only because we used on-line perceptron learning for
+%prefetching and that enabled us to examine the weights in detail.
 
@@
     \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 3222940 bits = 39.34 KB}}\\
+        \multicolumn{4}{|c|}{\textbf{Total: 3,222,940 bits = 39.34 KB}}\\
     \hline

### 2018-12-05 related.tex
went through with a chainsaw and hacked off large chunks, then fine-tuned a lot of words. added Elvira's IPC graphs

@@
 which represents an opportunity for optimizations in cache management. 
-Khan \textit{et. al.} proposed SDBP~\cite{sdbp}, in this work, a sampler 
-structure keeps partial tags of sampled sets separate from the cache. 
-Three tables of two-bit saturating counters are accessed using a technique 
-similar to a skewed branch predictor~\cite{Piece_Linear}. For each hit to 
-a block in the sampled set, the program counter (PC) of the relevant memory 
-instruction is hashed into the three tables and the corresponding counters 
-are decremented. For each eviction from a sampled set, the counters corresponding 
-to the PC of the last instruction to access the victim block are incremented. 
-For an LLC access, the predictor is consulted by hashing the PC of the memory 
-access instruction into the three tables and taking the sum of the indexed 
-counters. When the sum exceeds some threshold, the accessed block is 
-predicted to be dead. Tags in sampled sets are managed with true LRU and 
-a reduced associativity, but the LLC may be managed by any policy. 
-The paper applies its predictor to do block replacement and bypass for LLC.
+%Khan \textit{et. al.} proposed SDBP~\cite{sdbp}, in this work, a sampler 
+%structure keeps partial tags of sampled sets separate from the cache. 
+%Three tables of two-bit saturating counters are accessed using a technique 
+%similar to a skewed branch predictor~\cite{Piece_Linear}. For each hit to 
+%a block in the sampled set, the program counter (PC) of the relevant memory 
+%instruction is hashed into the three tables and the corresponding counters 
+%are decremented. For each eviction from a sampled set, the counters corresponding 
+%to the PC of the last instruction to access the victim block are incremented. 
+%For an LLC access, the predictor is consulted by hashing the PC of the memory 
+%access instruction into the three tables and taking the sum of the indexed 
+%counters. When the sum exceeds some threshold, the accessed block is 
+%predicted to be dead. Tags in sampled sets are managed with true LRU and 
+%a reduced associativity, but the LLC may be managed by any policy. 
+%The paper applies its predictor to do block replacement and bypass for LLC.
 
-Teran \textit{et al.} propose using perceptron learning for reuse prediction 
-driving a bypass and replacement optimization~\cite{Perc_Reuse}. In that work, 
-perceptron learning is used to set weights selected by hashes of multiple features 
-consisting of the PC, several recent PCs, and two different shifts of the tag 
-of the referenced block. Each feature is hashed to index its own table of weights. 
-The selected weights are summed and thresholded to make a prediction. When a sampled 
-block is reused or evicted, the corresponding weights are decremented or incremented, 
-respectively, according to the perceptron learning rule.
+Teran \textit{et al.} propose perceptron learning for reuse prediction for
+bypass and replacement~\cite{Perc_Reuse}. Perceptron learning trains weights
+selected by hashes of multiple features consisting of the PC of the memory
+access instruction, some recent PCs, and two different shifts of the tag of
+the referenced block. Features are hashed to index weights tables, reading out
+weights that are and thresholded to make a prediction. When a block from one
+of a few sampled sets~\cite{sdbp} is reused or evicted, corresponding weights
+are decremented or incremented, respectively according to the perceptron
+learning rule.  Multiperspective Reuse Prediction~\cite{Multiperspective}
+improves on Perceptron by contributing many new features.
 
-The continuation of Perceptron was later introduced in the work, Multiperspective Reuse 
-Prediction~\cite{Multiperspective}. Compared to previous work that focuses on one or two 
-features at a time, Multiperspective prediction uses many features, each contributing 
-to the overall prediction. The prediction enables placement, promotion and bypass 
-optimizations that further improves performance.
+%features. Compared to previous work that focuses on one or two features at a
+%time, Multiperspective prediction uses many features, each contributing to the
+%overall prediction. The prediction enables placement, promotion and bypass
+%optimizations that further improves performance.
 
@@
 
-Two widely used prefetchers are Next-n Lines~\cite{nextn} and Stride~\cite{stride}
-prefetchers, both of which capture regular memory access patterns with very low hardware overhead. 
-The Next-n Lines prefetcher simply queues prefetches for the next n lines after any given miss, 
-under the expectation that the principle of spatial locality will hold and those cache lines after 
-a missed line are likely to be used in the future. The stride prefetcher is slightly more 
-sophisticated -- it attempts to identify simple stride reference patterns in programs based 
-upon the past behavior of missing loads. Similar to the Next-n Lines technique, 
-when a given load misses, cache lines ahead of that miss are fetched in the pattern 
-following the previous behavior in the hope of avoiding future misses. Both prefetchers 
-have been widely used due to the simple design and low hardware overhead. However, 
-without further knowledge about temporal locality and application characteristics, 
-these prefetchers cannot do more than detecting and prefetching regular memory access 
-patterns with limited spatial locality.
+Two widely used prefetchers are Next-$n$ Lines~\cite{nextn} and
+Stride~\cite{stride} prefetchers. Both capture regular memory access patterns
+with low overhead. The Next-$n$ Lines prefetcher queues prefetches for the
+next $n$ blocks after any given miss, expecting that the principle of spatial
+locality will hold and those cache blocks after a missed block are likely to
+be used in the future. The stride prefetcher identifies strided reference
+patterns in programs based on past behavior of missing loads.  When a load
+misses, cache blocks ahead of that miss are fetched in the pattern following
+the previous behavior in the hope of avoiding future misses.  Without further
+knowledge about temporal locality and application characteristics, these
+prefetchers cannot do more than detecting and prefetching regular memory
+access patterns with limited spatial locality.
 
-Somogyi et al. proposed one of the top practical, low-overhead prefetchers, the Spatial 
-Memory Streaming (SMS) prefetcher~\cite{SMS}. SMS leverages code-based correlation 
-to take advantage of spatial locality in the applications over larger regions of memory 
-(called spatial regions). It predicts the future access pattern within a spatial 
-region around a miss, based on a history of access patterns initiated by that 
-missing instruction in the past. While the SMS prefetcher is effective, it is 
-indirectly inferring future program control-flow when it speculates on the misses 
-in a spatial region. As a result, the state overheads of this predictor can 
-be higher than the others in this class. This idea was further extended by 
-the class of related prefetchers~\cite{STMS,Temporal_Instruction_Fetch}.
+Somogyi et al. proposed a practical, low-overhead prefetcher: Spatial Memory
+Streaming (SMS) prefetcher~\cite{SMS}. SMS leverages code-based correlation to
+take advantage of spatial locality in the applications over larger regions of
+memory called spatial regions. It predicts future access pattern within a
+spatial region around a miss based on a history of access patterns initiated
+by that missing instruction in the past. SMS is effective, but it indirectly
+infers future program control-flow when it speculates on the misses in a
+spatial region. Thus, the state overhead of this predictor can be higher than
+the others in this class. The idea was further extended by the class of
+related prefetchers~\cite{STMS,Temporal_Instruction_Fetch}.
 
-While previous work had proposed dynamically finding the best prefetch offset, they have failed to 
-look into prefetch timeliness. Michaud proposed the Best-Offset prefetcher~\cite{BOP}. 
-A prefetch offset is set automatically and dynamically tries to adapt to the application behavior. 
-A good prefetch offset is determined by looking at two recent access to a particularly line 
+Other work also considers prefetch timeliness. Michaud proposed the
+Best-Offset prefetcher~\cite{BOP}.  A prefetch offset is set automatically and
+dynamically tries to adapt to the application behavior.  A good prefetch
+offset is determined by looking at two recent access to a particular block
 that did not happen too recently. The time between both accesses should be
@@
 \subsection{Lookahead Prefetchers}
-While the stride and the next-n perefetchers are efficient, they suffer from their limited
-ability to capture memory delta patterns. Moreover, a good prefetch proves to be useful only 
-if the prefetcher manages to bring the block before its next demand access. Both these issues 
-can be overcome by Lookahead Prefetchers. These prefetchers try to use their own predictions 
-to create the spelative path of the program's memory access pattern. By going down the speculation
-trail till a certain depth, the prefetcher can bring in cache lines well before they are demand 
-accessed by the processor. VLDP~\cite{VLDP} prefetches a static depth ahead of the demand fetch, 
-without considering the prefetching confidence. More recent Lookahead Prefetchers include SPP and
+Previous prefetchers have only limited ability to capture memory delta
+patterns. Moreover, a good prefetch proves to be useful only if the prefetcher
+manages to bring the block before its next demand access. Both these issues
+are overcome by Lookahead Prefetchers that use their own predictions to create
+the speculative path of the program's memory access pattern. By going down the
+speculation trail to a certain depth, the prefetcher can bring in cache blocks
+well before they are demand accessed by the processor. VLDP~\cite{VLDP}
+prefetches a static depth ahead of the demand fetch, without considering the
+prefetching confidence. More recent Lookahead Prefetchers include SPP and
 KPC~\cite{KPC}
 
-Talk about DA-AMPM
+%Talk about DA-AMPM
 
@@
 
-A low-accuracy aggressive prefetcher can significantly harm performance. 
-In order to minimize the interference from prefetching, Wu \textit{et. al} proposed PACMan~\cite{pacman}, 
-a prefetch-aware cache management policy. PACMan dedicated a few sets of the LLC to each of the three 
-competing policies that treat demand and prefetch request differently and uses the policy that 
-shows the lowest number of cache misses. Similarly, Seshadri  \textit{et. al.} proposed ICP~\cite{icp}. 
-ICP is also designed as a comprehensive mechanism to mitigate prefetching interference. It simply 
-demotes a prefetch to the lowest priority on a demand hit based on the observation that most 
-prefetches are dead after their first hit. To address prefetcher-caused cache pollution, it also 
-uses a variation of EAF~\cite{eaf} to track prefetching accuracy and inserts only an accurate 
-prefetch to the higher priority position in the LRU stack. ICP assumes, however, that all 
-prefetches are inserted only into the LLC, which restricts the maximum benefit of prefetching.
+A low-accuracy aggressive prefetcher can significantly harm performance.  To
+minimize interference from prefetching, Wu \textit{et. al} proposed
+PACMan~\cite{pacman}, a prefetch-aware cache management policy.  PACMan
+dedicates some LLC sets to each of three competing policies that treat demand
+and prefetch requests differently, using the policy that shows the fewest
+misses. Seshadri \textit{et. al.} proposed ICP~\cite{icp} that is also
+designed as a comprehensive mechanism to mitigate prefetching interference. It
+demotes a prefetch to the lowest priority on a demand hit based on the
+observation that most prefetches are dead after their first hit.  To address
+prefetcher-caused cache pollution, it also uses a variation of EAF~\cite{eaf}
+to track prefetching accuracy and inserts only an accurate prefetch to the
+higher priority position in the LRU stack.
 
 \subsection{Machine Learning for Prefetching}
-Peled \textit{et.al.} introduced
-interesting ideas for on-line Reinforcement Learning and dynamically scaling
-the magnitude of feedback given to the baseline prefetcher~\cite{Semantics}.
-The biggest challenge here is that the prefetcher relies on compiler support for
-getting features to build the context.
-
-Liao \textit{et. al.} focused on prefetching for data center
-applications~\cite{Datacenter}. They used offline machine learning algorithms
-such as SVMs and logistic regression to do a parametric search for an optimal
-prefetcher configuration.
+Peled \textit{et.al.} introduced interesting ideas for on-line Reinforcement
+Learning and dynamically scaling the magnitude of feedback given to the
+baseline prefetcher~\cite{Semantics}.  The prefetcher relies on compiler
+support for getting features to build the context.  Liao \textit{et. al.}
+focused on prefetching for data center applications~\cite{Datacenter}. They
+used offline machine learning algorithms such as SVMs and logistic regression
+to do a parametric search for an optimal prefetcher configuration.

### 2018-12-05 results.tex
went through with a chainsaw and hacked off large chunks, then fine-tuned a lot of words. added Elvira's IPC graphs

@@
 
-\begin{figure*}[ht]
-\includegraphics[width=\textwidth]{4Core_SPEC2017}
-\caption{Normalized Speedup for 4-core SPEC CPU 2017 Workloads}
+%\begin{figure*}[ht]
+%\includegraphics[width=\textwidth]{4Core_SPEC2017}
+%\caption{Normalized Speedup for 4-core SPEC CPU 2017 Workloads}
+%\label{Fig:4Core_SPEC2017}
+%\end{figure*}
+
+\begin{figure}[ht]
+\includegraphics[width=2.75in]{4CORE_IPC_PPF}
+\caption{Speedup for 4-core SPEC CPU 2017}
 \label{Fig:4Core_SPEC2017}
-\end{figure*}
+\end{figure}
 
@@
 \newline
-Figure~\ref{Fig:4Core_SPEC2017}(a) shows a
-comparison of speedups obtained on 4-core mixes of a memory intensive subset
-SPEC CPU 2017.  We plot all 4 prefetchers, normalized to the baseline.  The
-workloads have been sorted in increasing order of the speedup.  PPF offers a
-speedup of \textbf{51.7\%} on these traces, and improvement of \textbf{11.9\%}
-over the baseline SPP, \textbf{10.2\%} over the next DA-AMPM, and
-\textbf{17.4\%} over BOP.
+Figure~\ref{Fig:4Core_SPEC2017} shows a comparison of speedups obtained on
+4-core mixes of a memory intensive subset SPEC CPU 2017.  We plot all 4
+prefetchers, normalized to the baseline.  The workloads have been sorted in
+increasing order of the speedup.  PPF offers a speedup of \textbf{51.7\%} on
+these traces, and improvement of \textbf{11.9\%} over the baseline SPP,
+\textbf{10.2\%} over the next DA-AMPM, and \textbf{17.4\%} over BOP.
 
-On a different set of fully random SPEC CPU 2017 4-core mixes, PPF provides an
-IPC speedup of \textbf{26.2}\% over the baseline, which is an improvement of
-\textbf{5.7}\% over SPP.
-\newline
+On a different set of fully random SPEC CPU 2017 4-core mixes (not illustrated
+for space reasons), PPF provides an IPC speedup of \textbf{26.2}\% over the
+baseline, which is an improvement of \textbf{5.7}\% over SPP.
 
-\begin{figure*}[ht]
-\includegraphics[width=\textwidth]{8Core_SPEC2017}
-\caption{Normalized Speedup for 8-core SPEC CPU 2017 Workloads}
+%\begin{figure*}[ht]
+%\includegraphics[width=\textwidth]{8Core_SPEC2017}
+%\caption{Normalized Speedup for 8-core SPEC CPU 2017 Workloads}
+%\label{Fig:8Core_SPEC2017}
+%\end{figure*}
+
+\begin{figure}[ht]
+\includegraphics[width=2.75in]{8CORE_IPC_PPF}
+\caption{Speedup for 8-core SPEC CPU 2017}
 \label{Fig:8Core_SPEC2017}
-\end{figure*}
+\end{figure}
 
@@
 \newline
-The sorted comparison of speedups on the memory
-intensive 8-core mixes is shown in Figure~\ref{Fig:8Core_SPEC2017}.  PPF
-improves baseline performance by \textbf{38.6\%}, an improvement of
-\textbf{10.7\%} over SPP.  For a random set of SPEC CPU 2017 mixes, PPF
-improves performance by \textbf{23.6\%} over the baseline, corresponding to
+The sorted comparison of speedups on the memory intensive 8-core mixes is
+shown in Figure~\ref{Fig:8Core_SPEC2017}.  PPF improves baseline performance
+by \textbf{38.6\%}, an improvement of \textbf{10.7\%} over SPP.  For a random
+set of SPEC CPU 2017 mixes (not illustrated for space reasons), PPF improves
+performance by \textbf{23.6\%} over the baseline, corresponding to
 \textbf{4.8\%} over SPP.  This increased improvement achieved by PPF over the
@@
 
-\begin{figure}[ht]
-\begin{adjustwidth}{-1cm}{}
-  \includegraphics[width=1.2\columnwidth]{AddnConstr}
-  \caption{IPC speedup for Small LLC and Low BW}
-  \label{Fig:AddnConstr}
-\end{adjustwidth}
-\end{figure}
+% djimenez: cut this graph and modified text accordingly
+%\begin{figure}[ht]
+%\begin{adjustwidth}{-1cm}{}
+%  \includegraphics[width=1.2\columnwidth]{AddnConstr}
+%  \caption{IPC speedup for Small LLC and Low BW}
+%  \label{Fig:AddnConstr}
+%\end{adjustwidth}
+%\end{figure}
 
@@
 
-Figure~\ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a
-reduced LLC and with low bandwidth constraints, respectively. For the sake of
-brevity, we have only shown selected traces followed by the overall
-performance on the complete benchmark.  Benchmark {\tt 605.mcf\_s} in low
-bandwidth conditions is prefetch averse.  In general, any prefetcher yields a
-negative speedup on that trace.  On {\tt 654.roms\_s} and {\tt
+%Figure~\ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a
+%followed by the overall performance on the complete benchmark.  
+%For the sake of brevity, we have only shown selected traces
+
+We also model PPF with reduced LLC and with low bandwidth constraints,
+respectively (not illustrated for space reasons).  Benchmark {\tt 605.mcf\_s}
+in low bandwidth conditions is prefetch averse.  In general, any prefetcher
+yields a negative speedup on that trace.  On {\tt 654.roms\_s} and {\tt
 607.cactuBSSN\_s}, PPF is unable to match the performance achieved by the best
 prefetcher. On the other hand, PPF outperforms all the other prefetchers on
-{\tt 623.xalancbmk\_s} and {\tt 638.imagick\_s} benchmarks.  Overall, PPF gives
-a better improvement under small LLC condition and matches the best
+{\tt 623.xalancbmk\_s} and {\tt 638.imagick\_s} benchmarks.  Overall, PPF
+gives a better improvement under small LLC condition and matches the best
 prefetcher, BOP, under low DRAM bandwidth conditions.

### 2018-12-05 related.tex
Last sub-section still needs work!

@@
 
-Perceptron-based prediction mechanisms have proven to yield superior accuracy, 
-which represents an opportunity for optimizations in cache management. 
+%Perceptron-based prediction mechanisms have proven to yield superior accuracy, 
+%which represents an opportunity for optimizations in cache management. 
 %Khan \textit{et. al.} proposed SDBP~\cite{sdbp}, in this work, a sampler 
@@
 
-Teran \textit{et al.} propose perceptron learning for reuse prediction for
-bypass and replacement~\cite{Perc_Reuse}. Perceptron learning trains weights
-selected by hashes of multiple features consisting of the PC of the memory
-access instruction, some recent PCs, and two different shifts of the tag of
-the referenced block. Features are hashed to index weights tables, reading out
-weights that are and thresholded to make a prediction. When a block from one
-of a few sampled sets~\cite{sdbp} is reused or evicted, corresponding weights
-are decremented or incremented, respectively according to the perceptron
+In addition to branch prediction~\cite{PerceptronPredictor}, perceptron-based
+learning has been applied to the area of cache management.
+Teran \textit{et al.} propose using perceptrons to predict cache line resuse, 
+bypass, and replacement~\cite{Perc_Reuse}. Perceptron Learning trains weights
+selected by hashes of multiple features, including the PC of the memory
+access instruction, some other recent PCs, and two different shifts of the tag of
+the referenced block. These features are used to index into weight tables, and
+the weights are then thresholded to generate a prediction. When a block from one
+of a few sampled sets~\cite{sdbp} is reused or evicted, the corresponding weights
+are decremented or incremented, according to the perceptron
 learning rule.  Multiperspective Reuse Prediction~\cite{Multiperspective}
-improves on Perceptron by contributing many new features.
+improves on Perceptron Learning by contributing many new features.
 
@@
 
-\subsection{Base Prefetchers}
+\subsection{Spatial Prefetchers}
 
-Two widely used prefetchers are Next-$n$ Lines~\cite{nextn} and
-Stride~\cite{stride} prefetchers. Both capture regular memory access patterns
-with low overhead. The Next-$n$ Lines prefetcher queues prefetches for the
-next $n$ blocks after any given miss, expecting that the principle of spatial
-locality will hold and those cache blocks after a missed block are likely to
-be used in the future. The stride prefetcher identifies strided reference
-patterns in programs based on past behavior of missing loads.  When a load
-misses, cache blocks ahead of that miss are fetched in the pattern following
-the previous behavior in the hope of avoiding future misses.  Without further
-knowledge about temporal locality and application characteristics, these
-prefetchers cannot do more than detecting and prefetching regular memory
-access patterns with limited spatial locality.
+%Two widely used prefetchers are Next-$n$ Lines~\cite{nextn} and
+%Stride~\cite{stride} prefetchers. Both capture regular memory access patterns
+%with low overhead. The Next-$n$ Lines prefetcher queues prefetches for the
+%next $n$ blocks after any given miss, expecting that the principle of spatial
+%locality will hold and those cache blocks after a missed block are likely to
+%be used in the future. The stride prefetcher identifies strided reference
+%patterns in programs based on past behavior of missing loads.  When a load
+%misses, cache blocks ahead of that miss are fetched in the pattern following
+%the previous behavior in the hope of avoiding future misses.  Without further
+%knowledge about temporal locality and application characteristics, these
+%prefetchers cannot do more than detecting and prefetching regular memory
+%access patterns with limited spatial locality.
 
-Somogyi et al. proposed a practical, low-overhead prefetcher: Spatial Memory
-Streaming (SMS) prefetcher~\cite{SMS}. SMS leverages code-based correlation to
-take advantage of spatial locality in the applications over larger regions of
-memory called spatial regions. It predicts future access pattern within a
-spatial region around a miss based on a history of access patterns initiated
-by that missing instruction in the past. SMS is effective, but it indirectly
-infers future program control-flow when it speculates on the misses in a
-spatial region. Thus, the state overhead of this predictor can be higher than
-the others in this class. The idea was further extended by the class of
-related prefetchers~\cite{STMS,Temporal_Instruction_Fetch}.
+Spatial prefetchers include such well-understood examples as the next-line (or
+next-$n$-line) prefetcher, and the stream prefetcher, and are distinguished by prefetching
+data without regard for the order in which the data will be accessed.
+In addition to simpler these, Somogyi \textit{et al.} 
+propose Spatial Memory Streaming (SMS)~\cite{SMS}.  SMS works by learning
+the spatial footprint of all data used by a program within a region of memory,
+and when that region is revisited by the program, the same spatial footprint
+is prefetched.  Ishii \textit{et al.} propose the Access Map Pattern Matching 
+prefetcher (AMPM)~\cite{AMPM}, which creates a map of all accessed lines within a region
+of memory, and then analyzes this map on every access to see if any 
+fixed-stride pattern can be identified and prefetched that is centered on 
+the current access. Pugsley \textit{et al.} propose the Sandbox Prefetcher~\cite{sandbox},
+which analyzes candidate fixed-offset prefetchers in a sandboxed environment
+to determine which is most suitable for the current program phase.  Michaud
+proposes the Best-Offset Prefetcher~\cite{BOP}, which determines the optimal
+offset to prefetch while considering memory latency and prefetch timeliness.
 
-Other work also considers prefetch timeliness. Michaud proposed the
-Best-Offset prefetcher~\cite{BOP}.  A prefetch offset is set automatically and
-dynamically tries to adapt to the application behavior.  A good prefetch
-offset is determined by looking at two recent access to a particular block
-that did not happen too recently. The time between both accesses should be
-greater than the latency for completing a prefetch request.
+%Somogyi et al. proposed a practical, low-overhead prefetcher: Spatial Memory
+%Streaming (SMS) prefetcher~\cite{SMS}. SMS leverages code-based correlation to
+%take advantage of spatial locality in the applications over larger regions of
+%memory called spatial regions. It predicts future access pattern within a
+%spatial region around a miss based on a history of access patterns initiated
+%by that missing instruction in the past. SMS is effective, but it indirectly
+%infers future program control-flow when it speculates on the misses in a
+%spatial region. Thus, the state overhead of this predictor can be higher than
+%the others in this class. The idea was further extended by the class of
+%related prefetchers~\cite{STMS,Temporal_Instruction_Fetch}.
+
+%Other work also considers prefetch timeliness. Michaud proposed the
+%Best-Offset prefetcher~\cite{BOP}.  A prefetch offset is set automatically and
+%dynamically tries to adapt to the application behavior.  A good prefetch
+%offset is determined by looking at two recent access to a particular block
+%that did not happen too recently. The time between both accesses should be
+%greater than the latency for completing a prefetch request.
 
 \subsection{Lookahead Prefetchers}
-Previous prefetchers have only limited ability to capture memory delta
-patterns. Moreover, a good prefetch proves to be useful only if the prefetcher
-manages to bring the block before its next demand access. Both these issues
-are overcome by Lookahead Prefetchers that use their own predictions to create
-the speculative path of the program's memory access pattern. By going down the
-speculation trail to a certain depth, the prefetcher can bring in cache blocks
-well before they are demand accessed by the processor. VLDP~\cite{VLDP}
-prefetches a static depth ahead of the demand fetch, without considering the
-prefetching confidence. More recent Lookahead Prefetchers include SPP and
-KPC~\cite{KPC}
+
+Unlike spatial prefetchers, lookahead prefetchers take program order into
+account when they make predictions.  Shevgoor \textit{et al.} propose the Variable 
+Length Delta Prefetcher (VLDP)~\cite{VLDP} that correlates histories of
+deltas between cache line accesses within memory pages with the next delta
+within that page.  SPP~\cite{SPP} and KPC's prefetching component~\cite{KPC} are
+more recent examples of lookahead prefetchers.  They try to predict not only what
+data will be used in the future, but also the precise order in which the data
+will be used, within a given page.  Predictions made by lookahead prefetchers can
+be fed back into their prediction mechanisms to predict even further down 
+a speculative path of memory acesses. These prefetchers can also generalize their
+learned patterns from one page, and use those patterns to make predictions in 
+other pages.
+
+%Previous prefetchers have only limited ability to capture memory delta
+%patterns. Moreover, a good prefetch proves to be useful only if the prefetcher
+%manages to bring the block before its next demand access. Both these issues
+%are overcome by Lookahead Prefetchers that use their own predictions to create
+%the speculative path of the program's memory access pattern. By going down the
+%speculation trail to a certain depth, the prefetcher can bring in cache blocks
+%well before they are demand accessed by the processor. VLDP~\cite{VLDP}
+%prefetches a static depth ahead of the demand fetch, without considering the
+%prefetching confidence. More recent Lookahead Prefetchers include SPP and
+%KPC~\cite{KPC}
 
@@
 
-\subsection{Aggressive Prefetching}
+\subsection{Managing Prefetched Data}
 
@@
 A low-accuracy aggressive prefetcher can significantly harm performance.  To
-minimize interference from prefetching, Wu \textit{et. al} proposed
+minimize interference from prefetching, Wu \textit{et al.} propose
 PACMan~\cite{pacman}, a prefetch-aware cache management policy.  PACMan
 dedicates some LLC sets to each of three competing policies that treat demand
-and prefetch requests differently, using the policy that shows the fewest
-misses. Seshadri \textit{et. al.} proposed ICP~\cite{icp} that is also
-designed as a comprehensive mechanism to mitigate prefetching interference. It
-demotes a prefetch to the lowest priority on a demand hit based on the
+and prefetch requests differently, using the policy in the rest of the cache that shows the fewest
+misses. Seshadri \textit{et al.} propose ICP~\cite{icp}, which
+demotes a prefetch to the lowest reuse priority on a demand hit, based on the
 observation that most prefetches are dead after their first hit.  To address
 prefetcher-caused cache pollution, it also uses a variation of EAF~\cite{eaf}
-to track prefetching accuracy and inserts only an accurate prefetch to the
-higher priority position in the LRU stack.
+to track prefetching accuracy, and inserts only accurate prefetches to the
+higher priority position in the LRU stack.  Jain \textit{et al.} propose
+Harmony~\cite{HarmonyHawkeye} to accomodate prefetches in their MIN 
+algorithm-inspired Hawkeye cache management system.
 
 \subsection{Machine Learning for Prefetching}
-Peled \textit{et.al.} introduced interesting ideas for on-line Reinforcement
+
+Peled \textit{et al.} introduce interesting ideas for on-line Reinforcement
 Learning and dynamically scaling the magnitude of feedback given to the
 baseline prefetcher~\cite{Semantics}.  The prefetcher relies on compiler
-support for getting features to build the context.  Liao \textit{et. al.}
-focused on prefetching for data center applications~\cite{Datacenter}. They
-used offline machine learning algorithms such as SVMs and logistic regression
+support for getting features to build the context.  Liao \textit{et al.}
+focuse on prefetching for data center applications~\cite{Datacenter}. They
+use offline machine learning algorithms such as SVMs and logistic regression
 to do a parametric search for an optimal prefetcher configuration.

### 2018-12-05 related.tex
Machine Learning for Prefetching still needs work!

@@
 fixed-stride pattern can be identified and prefetched that is centered on 
-the current access. Pugsley \textit{et al.} propose the Sandbox Prefetcher~\cite{sandbox},
+the current access. Pugsley \textit{et al.} propose the Sandbox Prefetcher~\cite{Sandbox},
 which analyzes candidate fixed-offset prefetchers in a sandboxed environment
@@
 higher priority position in the LRU stack.  Jain \textit{et al.} propose
-Harmony~\cite{HarmonyHawkeye} to accomodate prefetches in their MIN 
+Harmony~\cite{Harmony} to accomodate prefetches in their MIN 
 algorithm-inspired Hawkeye cache management system.

### 2018-12-06 abs.tex
rewording and minor refocus

--- 2018-11-29 abs.tex
@@
 \begin{abstract}
-Hardware prefetching has been introduced in modern processors as a way to hide
-cache latencies.  An efficient prefetcher should identify complex
-memory access patterns during program execution. This ability enables the
-prefetcher to read a block ahead of its demand access, potentially preventing a
-cache miss. Accurately identifying the right blocks to prefetch is essential
-to achieving high performance from the prefetcher.
 
-In this paper, we introduce Perceptron-based Prefetch Filtering to help make
- prefetching decisions accurately. The perceptron layer acts as a check 
-to filter out the unnecessary prefetches recommended by the underlying
-prefetcher.  We have also explored a range of features that can be used to
-train the perceptron layer.  Our results show that perceptron-based filtering
-improves performance on the memory intensive subset of the SPEC 2017 benchmark
-suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared
-to a state-of-the art prefetcher.  We also demonstrate that the performance
-gained from using our efficient filter continues to scale as the number of cores sharing a 
-last level cache increases.
+  % djimenez: I have taken out parenthetical terms in favor of
+  % explicitly mentioning them in the text. some people have a
+  % visceral reaction against parentheses that delimit things
+  % essential to understanding the text.
+
+  Hardware prefetching is an effective technique for hiding cache miss
+  latencies in modern processor designs.  Prefetcher performance can
+  be characterized by two main metrics that are generally at odds with
+  one another: coverage, the fraction of baseline cache misses which
+  the prefetcher brings into the cache; and accuracy, the fraction of
+  prefetches which are ultimately used.  An overly aggressive
+  prefetcher may improve coverage at the cost of reduced accuracy.
+  Thus, performance may be harmed by this over-aggressiveness because
+  many resources are wasted, including cache capacity and bandwidth.
+  An ideal prefetcher would have both high coverage and accuracy.
+
+  % Hardware prefetching has been introduced in modern processors as a
+  % way to hide cache latencies.  An efficient prefetcher should
+  % identify complex memory access patterns during program
+  % execution. This ability enables the prefetcher to read a block
+  % ahead of its demand access, potentially preventing a cache
+  % miss. Accurately identifying the right blocks to prefetch is
+  % essential to achieving high performance from the prefetcher.
+
+  In this paper, we introduce Perceptron-based Prefetch Filtering
+  (PPF) as a way increase the coverage of the prefetches generated by
+  a baseline prefetcher without negatively impacting accuracy.  PPF
+  enables more aggressive tuning of a given baseline prefetcher,
+  leading to increased coverage by filtering out the growing numbers
+  of inaccurate prefetches such an aggressive tuning implies.  We also
+  explore a range of features to use to train PPF's perceptron layer
+  to identify inaccurate prefetches.  PPF improves performance on a
+  memory-intensive subset of the SPEC CPU 2017 benchmarks by 6.84\%
+  for a single-core configuration, and by 11.9\% for a 4-core
+  configuration, compared to the baseline prefetcher alone.  We also
+  show that performance continues to scale as the number of cores
+  sharing a last level cache increases.
+
+  % In this paper, we introduce Perceptron-based Prefetch Filtering to
+  % help make prefetching decisions accurately. The perceptron layer
+  % acts as a check to filter out the unnecessary prefetches
+  % recommended by the underlying prefetcher.  We have also explored a
+  % range of features that can be used to train the perceptron layer.
+  % Our results show that perceptron-based filtering improves
+  % performance on the memory intensive subset of the SPEC CPU 2017
+  % benchmark suite by 6.84\% on single-core and by 11.9\% on
+  % multi-core traces, as compared to a state-of-the art prefetcher.
+  % We also demonstrate that the performance gained from using our
+  % efficient filter continues to scale as the number of cores sharing
+  % a last level cache increases.
 

### 2018-12-06 back.tex
Mostly related to images, methodology, typos etc

@@
 any look-ahead prefetcher, we develop a practical implementation of the idea
-using SPP as our baseline.  Here we describe the basic architecture of SPP.
+using SPP as our baseline. Here we describe the basic architecture of SPP.\\
 
-\textbf{Signature Table:} The Signature Table keeps track of 256 most recently
+\noindent \textbf{Signature Table:} The Signature Table keeps track of 256 most recently
 accessed pages. It is meant to capture memory access patterns within a page
@@
 found, the stored signature retrieved and used to index into the Pattern
-Table. This process is illustrated by figure \ref{fig:spp_update}.
+Table. This process is illustrated by figure \ref{fig:spp_update}.\\
 
-\textbf{Pattern Table:} The Pattern Table is indexed by the signature
+\noindent \textbf{Pattern Table:} The Pattern Table is indexed by the signature
 generated from the Signature Table. Pattern Table holds predicted delta
@@
 holds up to 4 unique delta predictions. This is implemented by making the
-Pattern Table 4-way associative.
+Pattern Table 4-way associative.\\
 
@@
 
-\textbf{Prefetch Table:} The Prefetch Table is a 1024 entry 1-way associative
+\noindent \textbf{Prefetch Table:} The Prefetch Table is a 1024 entry 1-way associative
 table that keeps a record of last few entries prefetched. This proves useful
@@
 filter, we will be calling this structure as the Prefetch Table to avoid any
-confusion.
+confusion.\\
 
-\textbf{Lookahead Prefetching:} On each trigger, SPP goes down the program
+\noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes down the program
 speculation path using its own prefetch suggestion. Using current prefetch as
@@
 compounding the confidence in each depth. Thus as depth increases, overall
-confidence keeps decreasing.
+confidence keeps decreasing.\\
 
-\textbf{Confidence Tracking}: The Pattern Table keeps a count of hits to each
+\noindent \textbf{Confidence Tracking}: The Pattern Table keeps a count of hits to each
 signature through a counter C\textsubscript{sig}. The number of hits for a
@@
 calculated as the ratio of the number of prefetches which led to a demand hit
-to the number of prefetches recommended in total. The range of $\alpha$ is
-[0,1].
-
-`D' is the lookahead depth. For $d = 1$, when SPP is in
+to the number of prefetches recommended in total. The range of $\alpha$ is [0,1].
+$D$ is the lookahead depth. For $d = 1$, when SPP is in
 non-speculative mode, P\textsubscript{0} can be thought of as 1. 
-The final P\textsubscript{d} is thresholded against PF\_THRESHOLD 
+The final P\textsubscript{d} is thresholded against prefetch threshold 
 (T\textsubscript{p}) to reject the low confidence suggestions and 
-then against a numerically bigger FILL\_THRESHOLD (T\textsubscript{f}) to 
+then against a numerically bigger fill threshold (T\textsubscript{f}) to 
 decide whether the prefetch should be sent to L2 Cache 
@@
   \begin{center}
-  \includegraphics[width=9cm]{SPP_Prefetch_Description}
+  \includegraphics[width=\columnwidth]{SPP_Prefetch_Description}
     \label{fig:spp_strcture}
-  \caption{SPP architecture}
+  \caption{SPP Architecture}
   \end{center}
@@
   \begin{center}
-  \includegraphics[width=9cm]{SPP_Update_Description}
+  \includegraphics[width=\columnwidth]{SPP_Update_Description}
   \label{fig:spp_update}
-  \caption{SPP data-path flow}
+  \caption{SPP Data-path Flow}
   \end{center}

### 2018-12-06 design_arch.tex
Mostly related to images, methodology, typos etc

@@
   \begin{center}
-  \includegraphics[width=8cm]{PPF_Hierarchy.pdf}
+  \includegraphics[width=8cm]{PPF_Hierarchy}
     \label{fig:PPF_Hierarchy}
@@
 placed between the prefetcher and the prefetch insertion queue, as illustrated
-in figure ~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
+in figure~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
 polluting the higher levels of the memory hierarchy.
@@
 
+\begin{figure}[ht]
+  \begin{center}
+  \includegraphics[width=\columnwidth]{Datapath_Separate}
+    \label{fig:PPF_Datapath}
+  \caption{PPF Data Path and Operation}
+  \end{center}
+\end{figure}
+
 \subsection{The Perceptron Filter}
 \label{Arch-Perceptron}
-\pg{rewrite this section as talking to the figures and walking through
-  them}
+%\pg{rewrite this section as talking to the figures and walking through them}
 
@@
 accuracy and area. A detailed explanation of the storage overhead of PPF can
-be found in Figure~\ref{tab:PPF_overhead}.
+be found in Figure~\ref{tab:PPF_overhead}.\\
 
-\begin{figure}
-  \begin{center}
-  \includegraphics[width=9cm]{Datapath.pdf}
-    \label{fig:PPF_Datapath}
-  \caption{PPF Data Path and Operation}
-  \end{center}
-\end{figure}
-
-\noindent{\textbf{Inference}\\
+\noindent{\textbf{Inferencing}\\
 %ETG:is this like an explanation of how prefetching happens? 
@@
 prioritized. A prefetch for which the features result in a confidence value
-between PERC\_THRESHOLD\_LO and PERC\_THRESHOLD\_HI is allocated in the larger
+between PERC\_THRESHOLD\_LO and PERC\_THRESHOLD\_ HI is allocated in the larger
 LLC, as the filter is moderately confident of the future reuse of the cache
@@
 
-\noindent{2. Recording}\\
-As shown in step 2 of figure ~\ref{fig:PPF_Datapath} the prefetches that make
+\noindent \textbf{Recording}\\
+As shown in step 2 of figure~\ref{fig:PPF_Datapath} the prefetches that make
 it through the inference stage are recorded in the ``Prefetch Table''. The
@@
 In general, PPF can be adapted to a new base prefetcher with only a
-few modifications: \\ \\
+few modifications:\\
 %

### 2018-12-06 method.tex
Mostly related to images, methodology, typos etc

@@
 levels of cache hierarchies. Branch prediction is done using the perceptron
-branch predictor~\cite{PerceptronPredictor}.  The page size is 4KB.  ChampSim operates
-all the prefetchers strictly in the physical address space.
+branch predictor~\cite{PerceptronPredictor}. The page size is 4KB. 
+ChampSim operates all the prefetchers strictly in the physical address space.
 
@@
 
-\textit{Single-core performance:} For single-core simulations, we use 200
+\noindent \textit{Single-core performance:} For single-core simulations, we use 200
 million instructions to warm-up the microarchitectural structures and the next
@@
 statistics. We report the IPC speedup over the baseline of no prefetching.
-The final number reported is the geometric mean of the speedup achieved on
-individual traces.
+The final numbers reported are the geometric mean of the weighted mean speedup 
+achieved per application using the SimPoint methodology.\\
 
-\textit{Multi-core performance:} For multi-application workloads, we generate
+\noindent \textit{Multi-core performance:} For multi-application workloads, we generate
 100 random mixes and another 100 mixes from the memory intensive subset of
@@
 instructions after warm-up. For collecting IPC and other data, only the first
-billion instructions are considered as the region of interest.
+billion instructions are considered as the region of interest.\\
 
@@
 simulation run-time within reasonable limits as a single 8-core mix takes up
-to 3 days to simulate one billion instructions.
+to 3 days to simulate one billion instructions.\\
 
-\textit{Validation:} We cross-validated our PPF model using SPEC
+\noindent \textit{Validation:} We cross-validated our PPF model using SPEC
 CPU 2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks. For
@@
 prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern
-Matching (DA-AMPM)~\cite{DA_AMPM} and Signature Path Prefetcher (SPP).  BOP was
+Matching (DA-AMPM) \cite{DA_AMPM} and Signature Path Prefetcher (SPP). BOP was
 the winner of 2nd Data Prefetching Championship. DA-AMPM is the enhanced
@@
 %
-\textbf{Global Pearson's Correlation}\newline For this experiment, we examine
+\textbf{Global Pearson's Correlation}\newline 
+For this experiment, we examine
 the perceptron weights at the end of all trace execution by which time the
@@
 perfectly linear positive / negative correlation respectively.
-Figure~\ref{Fig:Hist} depicts the histogram distribution of trained weights,
-for a feature with high and another with moderate Pearson's coefficient.
 
-Figure~\ref{Fig:Pvals} shows all the features used, arranged in the increasing
-order of their Pearson's factor.  As can be seen 5 out of the 9
-features provide a moderate to high correlation, with the magnitude of
-P-value > 0.6.  The single most important feature, 
-\textit{Confidence XOR Page address} helps provide
+As a part of our perceptron feature selection methodology, we explored a wide 
+variety of features to begin with. Features with a low 
+Pearson's coefficient were rejected as they didn't provide much useful 
+correlation. Figure~\ref{Fig:Hist} depicts the histogram distribution of 
+trained weights for two features. The first feature, 
+\textit{Confidence XOR Page address} has the highest observed P-value, 
+hence was retained. On the other hand, the second feature, \textit{Last Signature}
+did not provide any meaningful correlation and hence was rejected. 
+This is visible from the bulk of its trained weights settling to near zero values.
+
+Figure~\ref{Fig:Pvals} shows all the features which are finally used, 
+arranged in the increasing order of their Pearson's factor. 
+As can be seen 5 out of the 9 features provide a moderate to high 
+correlation, with the magnitude of P-value > 0.6. The single most 
+important feature, \textit{Confidence XOR Page address} helps provide
 a correlation to prefetch outcome with a factor of 0.90. \\ \\
@@
 varies across the traces. Here we give special attention to features
-with low P-values in.  Figure~\ref{Pvals_Traces} shows the
+with low P-values in. Figure~\ref{Fig:Pvals_Trace} shows the
 variation of P-values for three features : \textit{PC XOR Delta},
@@
 For simplicity, the traces are arranged in an
-increasing order of contribution made by the feature and the trace
-names have been omitted.  It can be seen that even features with a low
+increasing order of contribution made by the. 
+It can be seen that even features with a low
 overall correlation provide useful correlation (magnitude > 0.5) for
@@
 %
-\textbf{Trimming Features Using Correlation} \newline Beside
+\textbf{Trimming Features Using Cross-correlation} \newline Beside
 providing interesting insights into prefetching behavior, P-value can
@@
 
-As we examine correlation of each feature with the final outcome, we also
-study correlation between the features.  We used the above methodology to
+As we examined correlation of each feature with the final outcome, we also
+studied correlation between the features. We used the above methodology to
 eliminate features providing little information that has already been captured
@@
 
-We initially came up a mix of 23 features- primary and composite.  By studying
+We initially came up a mix of 23 features. By studying
 cross correlation of each of these features against others in a 23$\times$23
@@
 magnitude and eliminated redundant features, reducing the feature count to 9.
-Thus, in the final implementation of PPF, no two features have high
+Thus, in the final implementation of PPF, no two features have a high
 correlation between them. This way we can be sure that each feature makes a
@@
 correlation, \textit{cache line} and \textit{page address} were given most
-importance and allowed full 12-bits of indexing.  Features like \textit{IP XOR
-delta} and \textit{IP XOR depth} with a low overall P-value were allocated
+importance and allowed full 12-bits of indexing. Features like \textit{PC XOR
+delta} and \textit{PC XOR depth} with a low overall P-value were allocated
 fewer entries in the feature table.
@@
     \hline
-        PC 		& 12 & \multirow{4}{4.8cm}{Meta-data required for perceptron}\\
-        Address 	& 24 & \multirow{4}{1.1cm}{training}\\
+        PC 		& 12 & \\
+        Address 	& 24 & \\
+        Curr Signature 	& 10 & \multirow{2}{4.8cm}{Meta-data required for perceptron}\\
+	PC$_i$ Hash	& 12 & \multirow{2}{1.1cm}{training}\\
         Delta 		& 7  & \\
-        Curr Signature 	& 12 & \\
         Confidence 	& 7  & \\
+	Depth		& 4  & \\
     \hline
-        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
+        \multicolumn{3}{|c|}{\textbf{Total 85 bits}}\\
     \hline

### 2018-12-06 method.tex
Modified to reflect SimPoint methodology

@@
 instructions after warm-up. For collecting IPC and other data, only the first
-billion instructions are considered as the region of interest.\\
+billion instructions are considered as the region of interest.
 
@@
 Throughout the paper, we consider memory intensive subset as the
-traces with LLC MPKI > 1. This includes 48 out of the 95 simpoints
-developed for SPEC CPU 2017. For SPEC CPU 2006, this includes 48 out of 94 simpoints.
+applications with SimPoint weighted LLC MPKI > 1. This includes 11 out of 
+20 SPEC CPU 2017 applications. For SPEC CPU 2006, this includes 16 out of 
+29 applications.
 

### 2018-12-06 results.tex
Mostly related to images, methodology, typos etc
Updating Single Core Results
...to reflect simpoint weighted results

@@
 Figure~\ref{Fig:SPEC2017_1core} shows the IPC speedup obtained by PPF,
-compared to BOP, DA-AMPM and SPP.  In the figure we have depicted selected
-traces where PPF shows a noticeable improvement or shows some degradation,
-along with the geometric mean improvement on the memory intensive subset of
-SPEC CPU 2017, and finally for the complete SPEC CPU 2017 benchmark.  All the
-results have been normalized to the baseline of no prefetching.
+compared to BOP, DA-AMPM and SPP; for the individual SPEC CPU 2017 
+applications, followed by memory intensive subset and finally the full
+suite. All the results have been normalized to the baseline of no prefetching.
 
@@
 
-PPF yields a geometric mean speedup of \textbf{42.8\%} over the baseline. 
-This is equivalent to \textbf{7.8\%} over DA-AMPM, \textbf{9.7\%} over BOP 
-and \textbf{6.92\%} over SPP.  Out of the 95 simpoints developed for SPEC 
+PPF yields a geometric mean speedup of \textbf{27\%} over the baseline. 
+This is equivalent to \textbf{4.63\%} over DA-AMPM, \textbf{4.61\%} over BOP 
+and \textbf{3.78\%} over SPP. Out of the 95 simpoints developed for SPEC 
 CPU 2017, PPF nearly matches or outperforms most of the prefetchers on 91 
 simpoints. PPF fails to match the improvement offered by BOP only for 
-{\tt 607.cactuBSSN\_s} and {\tt 654.roms\_s}.
+{\tt 607.cactuBSSN\_s} and {\tt 654.roms\_s} traces.
 
-At its peak, PPF manages a speedup of a factor of over \textbf{3.39x} on the
-trace {\tt 602.gcc\_s-2226B}.  This also corresponds to speedup gain of
-\textbf{60\%} over the next best prefetcher -- BOP.  In general, benchmarks
+At its peak, PPF manages a speedup of a factor of over \textbf{3.39x} on one of the
+{\tt 602.gcc\_s} traces. This also corresponds to speedup gain of
+\textbf{60\%} over the next best prefetcher for that trace -- BOP. In general, benchmarks
 {\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt 623. xalancbmk\_s} and {\tt
@@
 On the full SPEC CPU 2017 suite, PPF improves the geometric mean IPC of 
-the baseline by \textbf{20.9\%}, which is \textbf{3.44\%} better 
+the baseline by \textbf{15.24\%}, which is \textbf{2.27\%} better 
 than the next best prefetcher -- SPP.
@@
 \begin{figure}[ht]
-\includegraphics[width=2.75in]{4CORE_IPC_PPF}
+\includegraphics[width=2.75in]{4Core_SPEC2017}
 \caption{Speedup for 4-core SPEC CPU 2017}
@@
 \begin{figure}[ht]
-\includegraphics[width=2.75in]{8CORE_IPC_PPF}
+\includegraphics[width=2.75in]{8Core_SPEC2017}
 \caption{Speedup for 8-core SPEC CPU 2017}
@@
 consistently outperforms the best performing prefetcher.
-
-%[EB] How to justify only a small margin over DA-AMPM?
 
@@
 Figure~\ref{Fig:SPEC2006_1core} shows the speed-up achieved by BOP, SPP and
-PPF on traces from SPEC CPU 2006, along with the overall geometric mean over
-the whole suite and over a memory-intensive subset.  PPF provides a speedup of
-\textbf{46.1\%} over the baseline on the memory intensive subset of SPEC CPU
-2006 benchmark, giving an improvement of \textbf{6.66\%} over SPP and
-\textbf{9.88\%} over DA-AMPM and \textbf{9.82} over BOP.  On the whole of 
-the SPEC CPU 2006 suite, the speedup is \textbf{22.4\%}, an improvement 
-of \textbf{3.34\%} over SPP.
+PPF on the memory intensive subset and the full SPEC CPU 2006 suite. 
+PPF provides a speedup of \textbf{36.3\%} over the baseline on the 
+memory intensive subset of SPEC CPU
+2006 benchmark, giving an improvement of \textbf{6.1\%} over SPP and
+\textbf{8.44\%} over DA-AMPM and \textbf{9.93} over BOP. On the whole of 
+the SPEC CPU 2006 suite, the speedup is \textbf{19.6\%}, an improvement 
+of \textbf{3.33\%} over SPP.
 

### 2018-12-06 results.tex
Updated IPC Speedup numbers
...to reflect the final configuration

@@
 
-PPF yields a geometric mean speedup of \textbf{27\%} over the baseline. 
+PPF yields a geometric mean speedup of \textbf{26.95\%} over the baseline. 
 This is equivalent to \textbf{4.63\%} over DA-AMPM, \textbf{4.61\%} over BOP 
@@
 prefetchers, normalized to the baseline. The workloads have been sorted in
-increasing order of the speedup. PPF offers a speedup of \textbf{51.7\%} on
-these traces, and improvement of \textbf{11.9\%} over the baseline SPP,
-\textbf{10.2\%} over the next DA-AMPM, and \textbf{17.4\%} over BOP.
+increasing order of the speedup. PPF offers a speedup of \textbf{51.2\%} on
+these traces, and improvement of \textbf{11.4\%} over the baseline SPP,
+\textbf{9.7\%} over the next DA-AMPM, and \textbf{16.9\%} over BOP.
 
 On a different set of fully random SPEC CPU 2017 4-core mixes (not illustrated
-for space reasons), PPF provides an IPC speedup of \textbf{26.2}\% over the
-baseline, which is an improvement of \textbf{5.7}\% over SPP.
+for space reasons), PPF provides an IPC speedup of \textbf{26.07}\% over the
+baseline, which is an improvement of \textbf{5.6}\% over SPP.
 
@@
 shown in Figure~\ref{Fig:8Core_SPEC2017}. PPF improves baseline performance
-by \textbf{38.6\%}, an improvement of \textbf{10.7\%} over SPP. For a random
+by \textbf{37.6\%}, an improvement of \textbf{9.7\%} over SPP. For a random
 set of SPEC CPU 2017 mixes (not illustrated for space reasons), PPF improves
-performance by \textbf{23.6\%} over the baseline, corresponding to
-\textbf{4.8\%} over SPP. This increased improvement achieved by PPF over the
+performance by \textbf{23.4\%} over the baseline, corresponding to
+\textbf{4.6\%} over SPP. This increased improvement achieved by PPF over the
 base engine SPP in a multi-core environment is expected as PPF has a very
@@
 For 4-core memory intensive mixes, PPF improves the baseline by
-\textbf{59.2\%}, \textbf{8.7\%} ahead of SPP. For 8-core memory intensive
-mixes, the speedup over the baseline is \textbf{47.9\%}, \textbf{11.5\%} ahead
+\textbf{59.1\%}, \textbf{8.6\%} ahead of SPP. For 8-core memory intensive
+mixes, the speedup over the baseline is \textbf{47.8\%}, \textbf{11.3\%} ahead
 of SPP.
@@
 In general, these applications are prefetch agnostic. Even so, PPF manages a
-\textbf{3.7\%} improvement over no prefetching, putting it ahead of the next
+\textbf{3.78\%} improvement over no prefetching, putting it ahead of the next
 best prefetcher, SPP, which provides a 3.08\% speedup.

### 2018-12-07 abs.tex
Mostly related to images, methodology, typos etc
Small changes throughout the paper

@@
 to identify inaccurate prefetches. PPF improves performance on a
-  memory-intensive subset of the SPEC CPU 2017 benchmarks by 6.84\%
-  for a single-core configuration, and by 11.9\% for a 4-core
+memory-intensive subset of the SPEC CPU 2017 benchmarks by 3.78\%
+for a single-core configuration, and by 11.4\% for a 4-core
 configuration, compared to the baseline prefetcher alone. We also

### 2018-12-07 back.tex
Small changes throughout the paper

@@
-\section{Motivation and Prior Work}
+\section{Motivation}
 \label{Background}
 
-In this section we discuss the most closely related work to our proposed
-technique. See Section~\ref{related} for other relevant work.
-
+In this section we discuss the most closely related work to our proposed technique. 
 The idea of prefetching begins with Jouppi's \textit{Instruction Stream
@@
 locality~\cite{Spatial_Pattern,SMS,Temporal_Instruction_Fetch,Off_Chip,STMS,SMS_JILP},
-control-flow speculation~\cite{BFetch,MTBFetch}, and many other aspects to
-detect complex access patterns in modern applications.
+control-flow speculation~\cite{BFetch,MTBFetch}, and other other aspects to
+detect complex memory access patterns.
+See Section~\ref{related} for other relevant work.
 
-\subsection{Lookahead-based Prefetchers}
-
+% [EB:] Removing this section title and moving relevant text to next section
+%       We weren't talking about LA prefetchers here anyway, just SPP
+% \subsection{Lookahead-based Prefetchers}
 %% Simple stride prefetching techniques only detect sequences of addresses that differ by a constant value and fail to capture diverse memory access patterns. 
-Lookahead prefetchers can learn complicated patterns by collecting histories of
-observed data access patterns.
+%Lookahead prefetchers can learn complicated patterns by collecting histories of observed data access patterns.
 %EB: Didn't really understand this: 
 %% "and correlating these with the strides, by considering each new page in a vacuum."
-Kim {\em et. al.} proposed
-SPP~\cite{SPP}, a confidence-based lookahead prefetcher. SPP creates a
-signature associated to a memory request by compressing the history of
-accesses. By correlating the signature with future likely delta patterns, SPP
-learn both simple and complicated memory access patterns quickly. Our proposed
-technique builds upon SPP. Thus, we describe SPP in more detail.
 
-\subsection{Baseline Prefetcher: The Signature Path Prefetcher}
+\subsection{Baseline Prefetcher: SPP}
 \label{Background-SPP}
 
+Kim {\em et. al.} proposed Signature Path Prefetcher (SPP)~\cite{SPP}, 
+a confidence-based lookahead prefetcher. 
+SPP creates a signature associated with a page address by compressing the history of 
+accesses. By correlating the signature with future likely delta patterns, SPP learns 
+both simple and complicated memory access patterns quickly. 
 While the basic idea of perceptron based prefetch filtering is applicable to
@@
 using SPP as our baseline. Here we describe the basic architecture of SPP.\\
-
-\noindent \textbf{Signature Table:} The Signature Table keeps track of 256 most recently
+%
+\noindent \textbf{Signature Table:} As shown in figure \ref{fig:spp_structure}, 
+the Signature Table keeps track of 256 most recently
 accessed pages. It is meant to capture memory access patterns within a page
@@
 Table. This process is illustrated by figure \ref{fig:spp_update}.\\
-
+%
 \noindent \textbf{Pattern Table:} The Pattern Table is indexed by the signature
@@
 patterns and their confidence estimates. Each entry indexed by the signature
-holds up to 4 unique delta predictions. This is implemented by making the
-Pattern Table 4-way associative.\\
+holds up to 4 unique delta predictions.\\
+%
 
@@
 %% 	is stored in the Pattern Table.
-
 \noindent \textbf{Prefetch Table:} The Prefetch Table is a 1024 entry 1-way associative
@@
 confusion.\\
-
+%
 \noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes down the program
@@
 confidence keeps decreasing.\\
-
+%
 \noindent \textbf{Confidence Tracking}: The Pattern Table keeps a count of hits to each
@@
   \includegraphics[width=\columnwidth]{SPP_Prefetch_Description}
-    \label{fig:spp_strcture}
   \caption{SPP Architecture}
+  \label{fig:spp_structure}
   \end{center}
@@
   \includegraphics[width=\columnwidth]{SPP_Update_Description}
+  \caption{SPP Data-path Flow}
   \label{fig:spp_update}
-  \caption{SPP Data-path Flow}
   \end{center}
@@
 orthogonal to the underlying prefetching scheme and can be used
-to prune out the useful prefetcher from the useless ones.
+to prune out the useful prefetches from the useless ones.
 

### 2018-12-07 intro.tex
Mostly related to images, methodology, typos etc
Small changes throughout the paper

@@
 \includegraphics[width=\columnwidth]{Motivation}
-\caption{Variation of IPC, Total Prefetches and Useful Prefetches}
+\caption{Impact of aggressive prefetching on performance. X-asis represents increasingly 
+	 aggressive prefetch lookahead depth. All the metrics have been normalized
+	 to the least aggressive configuration. As can be seen, total number of prefetches increases at a
+	 higher rate than the number of good prefetches, leading to IPC reduction.}
 \label{Fig:Motivation}
@@
 depth coverage. The figure depicts the behavior of the prefetcher variants running 
-on the trace {\tt 603.bwaves\_s-1740B}. The IPC, total number of prefetches 
+on the trace {\tt 603.bwaves\_s}. The IPC, total number of prefetches 
 suggested by the prefetcher (TOTAL\_PF) and the actual useful predictions (GOOD\_PF),
@@
 to cache pollution and bandwidth contention, as is clearly visible with the reduction in 
-the IPC. Similar behavior was observed in 12 other SPEC CPU 2017 simpoints.
+the IPC. 
+%%Similar behavior was observed in 12 other SPEC CPU 2017 simpoints.
 
@@
 mechanism, we are free to re-tune the rest of SPP's design around maximizing
-coverage.  The result is an increase in both accuracy and coverage, and
-a notable increase in performance.
+coverage. This results in a notable increase in system performance.
 
@@
 
-\item Implementing PPF filtering a state-of-the-art prefetchier, giving a
+\item Implementing PPF filtering on a state-of-the-art prefetcher, giving a
 significant performance improvement compared to previous work. PPF learns to
 adapt itself to shared resource constraints, leading to further increased
-performance in multi-core and bandwidth-constrained environments.
+performance in multi-core and small LLC environments.
 
-\item A methodology for determining an appropriate set of features for
+\item A methodology to determine an appropriate set of features for
 prediction, regardless of the underlying prefetcher used. More details are
@@
 
-In a single core configuration, PPF increases performance by 6.84\%
+In a single core configuration, PPF increases performance by 3.78\%
 compared to the underlying prefetcher, SPP. In a multi-core system running a
 mixes of memory intensive SPEC CPU 2017 traces, PPF saw an improvement
-of 11.9\% over SPP for a 4-core system, and 10.67\% for an 8-core system.
+of 11.4\% over SPP for a 4-core system, and 9.65\% for an 8-core system.
 

### 2018-12-07 results.tex
Small changes throughout the paper

@@
 the baseline by \textbf{15.24\%}, which is \textbf{2.27\%} better 
-than the next best prefetcher -- SPP.
-\newline
-
+than the next best prefetcher -- SPP.\\ \\
+%
 \begin{figure}[h]
@@
 \end{figure}
-
+%
 % djimenez: this is a weird way to make a header for a section. use \paragraph or \subsection
-
-\noindent \textbf{COVERAGE}
-\newline
+\noindent \textbf{Coverage:}
 Prefetcher coverage is defined as the ratio of the number of misses avoided
@@
 In this section, we demonstrate the improvement achieved by PPF for a mix of
-multi-programmed workloads.
-\newline
-
-\noindent \textbf{4-CORE ENVIRONMENT}
-\newline
+multi-programmed workloads.\\ \\
+%
+\noindent \textbf{4-Core Environment:}
 Figure~\ref{Fig:4Core_SPEC2017} shows a comparison of speedups obtained on
@@
 \textbf{9.7\%} over the next DA-AMPM, and \textbf{16.9\%} over BOP.
-
 On a different set of fully random SPEC CPU 2017 4-core mixes (not illustrated
 for space reasons), PPF provides an IPC speedup of \textbf{26.07}\% over the
-baseline, which is an improvement of \textbf{5.6}\% over SPP.
-
-%\begin{figure*}[ht]
-%\includegraphics[width=\textwidth]{8Core_SPEC2017}
-%\caption{Normalized Speedup for 8-core SPEC CPU 2017 Workloads}
-%\label{Fig:8Core_SPEC2017}
-%\end{figure*}
-
+baseline, which is an improvement of \textbf{5.6}\% over SPP.\\ \\
+%
 \begin{figure}[ht]
@@
 \end{figure}
-
-\noindent \textbf{8-CORE ENVIRONMENT}
-\newline
+%
+\noindent \textbf{8-Core Environment:}
 The sorted comparison of speedups on the memory intensive 8-core mixes is
@@
 pollution in the shared LLC.
-
 BOP offers a better improvement than SPP for the memory intensive mixes. This
@@
 \begin{figure}[ht]
-\includegraphics[width=\columnwidth]{SPEC2006}
+\includegraphics[width=0.8\columnwidth,height=4cm,keepaspectratio]{SPEC2006}
 \caption{SPEC CPU 2006 Single-Core IPC Speedup}
@@
 \begin{figure}[ht]
-\includegraphics[width=\columnwidth,]{CloudSuite}
-\caption{IPC Speedup for Multi-core CloudSuite Workloads}
+\includegraphics[width=0.9\columnwidth,height=3.5cm,keepaspectratio]{CloudSuite}
+\caption{IPC Speedup for CloudSuite Workloads}
 \label{Fig:CloudSuite}
 \end{figure}
+
+%% \begin{figure}[ht]
+%%   \centering
+%%   \begin{subfigure}{0.5\columnwidth}
+%%     \centering
+%%     \includegraphics[width=0.45\columnwidth]{CloudSuite}
+%%     \caption{CloudSuite Workloads}
+%%   \end{subfigure}
+%%   %
+%%   \begin{subfigure}{0.5\columnwidth}
+%%     \centering
+%%     \includegraphics[width=0.45\columnwidth]{SPEC2006}
+%%     \caption{Single Core SPEC CPU 2006}
+%%   \end{subfigure}
+%%   \caption{Performance with Unseen Benchmarks}
+%%   \label{Fig:CrossVal}
+%% \end{figure}
 
@@
 %        scale-out workloads on modern hardware" in methodology
-
-Here we demonstrate the robustness of PPF by testing it on different
-benchmarks: SPEC CPU 2006 and CloudSuite.
-
+%
+% Here we demonstrate the robustness of PPF by testing it on different
+% benchmarks: SPEC CPU 2006 and CloudSuite.
+%
 % djimenez: is the improvement over BOP and DA-AMPM exactly the same 9.86%

### 2018-12-07 intro.tex
my pass

--- 2018-12-05 intro.tex
@@
 Processor and memory technologies have been developed with different
-goals in mind.  While processor scaling has focused on speed improvements,
-memory scaling has primarily focused on increasing capacity.  The
-difference in each technology's scaling has led to the Memory Wall~\cite{MemWall}
--- the increasing gap between processor and memory performance. Data prefetching
-is one important technique that has been developed to minimize the effects 
-of this trend.
+goals in mind. While processor scaling has focused on speed
+improvements, memory scaling has primarily focused on increasing
+capacity. The difference in each technology's scaling has led to the
+Memory Wall~\cite{MemWall} -- the increasing gap between processor and
+memory performance. Data prefetching is one important technique that
+has been developed to minimize the effects of this trend.
 
@@
 
-An ideal prefetching scheme would perfectly capture a program's memory access pattern, and then
-predict and pre-load the needed data into the processor's caches in a timely manner. 
-Memory access patterns may be simple, such as accessing every item in an array with a for-loop,
-or very complex, such as chasing pointers through dynamically-allocated memory.
-Practical prefetchers face challenges not only in predicting what data will be useful in the future,
-but also in timing when data should be prefetched, deciding which level of the cache hierarchy
-the data should be stored in, and deciding what data should be evicted from the caches to accommodate
-the prefetched data.
-
-All prefetchers are designed around a fundamental trade-off between two important metrics:
-coverage and accuracy.  Prefetcher coverage refers to the number of cache hits that would have
-been cache misses without the prefetcher.  For example, if an application experiences 1,000 cache misses
-without a prefetcher, but only 200 cache misses with a prefetcher, then the prefetcher has 80\%
-coverage for that application.  Prefetcher accuracy refers to the fraction of prefetched cache lines
-that end up being used by the application.  So if a prefetcher prefetches 1,200 cache lines, 
-but only 800 of them are used by the application, then that prefetcher's accuracy is 66.7\%.
+An ideal prefetching scheme would perfectly capture a program's memory
+access pattern, and then predict and pre-load the needed data into the
+processor's caches in a timely manner.  Memory access patterns may be
+simple, such as accessing every item in an array with a for-loop, or
+very complex, such as chasing pointers through dynamically-allocated
+memory.  % Practical prefetchers face challenges not only in predicting
+% what data will be useful in the future, but also in timing when data
+% should be prefetched, deciding which level of the cache hierarchy the
+% data should be stored in, and deciding what data should be evicted
+% from the caches to accommodate the prefetched data.
+All prefetchers are designed around a fundamental trade-off between
+two important metrics: coverage and accuracy. Prefetcher coverage
+refers to the fraction of baseline cache misses that the prefetcher
+pulls into the cache prior to their reference.  For example, if an
+application experiences 1,000 cache misses without a prefetcher, while
+800 of those cache misses become hits with a prefetcher, then the
+prefetcher has 80\% coverage for that application.  Prefetcher
+accuracy refers to the fraction of prefetched cache lines that end up
+being used by the application. So if a prefetcher prefetches 1,200
+cache lines, but only 800 of them are used by the application, then
+that prefetcher's accuracy is 66.7\%.
 
@@
 \includegraphics[width=\columnwidth]{Motivation}
-\caption{Variation of IPC, Total Prefetches and Useful Prefetches}
+\caption{Variation of IPC, Total Prefetches and Useful Prefetches with
+  lookahead depth.}
 \label{Fig:Motivation}
@@
 
-Coverage and accuracy are generally at odds with one another, and as one metric improves, the other
-usually gets worse.  For example, when an application accesses a new region of memory for the first
-time, a na\"ive prefetcher may predict that all data in that region will be used by the application.
-This will clearly result in 100\% coverage for that region, but with possibly a very low accuracy.
-In fact, so much bandwidth may be wasted prefetching unused data that performance can ultimately be
-harmed by this strategy.  On the other extreme, another prefetcher may be overly
-conservative and never prefetch anything, wasting no bandwidth, and achieving 0\% prefetch
-coverage.
+Coverage and accuracy are generally at odds with one another, and as
+one metric improves, the other usually gets worse. For example, when
+an application accesses a new region of memory for the first time, a
+na\"ive prefetcher may predict that all data in that region will be
+used by the application.  This will clearly result in 100\% coverage
+for that region, but with possibly a very low accuracy.  In fact, so
+much cache capacity and bandwidth may be wasted prefetching unused
+data that performance can ultimately be harmed by this strategy. On
+the other extreme, another prefetcher may be overly conservative and
+never prefetch anything, wasting no capacity or bandwidth, and
+achieving 0\% prefetch coverage.
 
-Figure~\ref{Fig:Motivation} illustrates the above scenario in a working manner.
-Here we consider a state-of-the-art lookahead prefetcher -- SPP. We keep on 
-re-tuning its thresholds, each time getting a variant with an increasing lookahead 
-depth coverage. The figure depicts the behavior of the prefetcher variants running 
-on the trace {\tt 603.bwaves\_s-1740B}. The IPC, total number of prefetches 
-suggested by the prefetcher (TOTAL\_PF) and the actual useful predictions (GOOD\_PF),
-all have been normalized to the least aggressive version of the prefetcher. 
-As the look-ahead depth of the prefetch suggestions increase, so do useful prefetches. 
-But that comes at a cost of total prefetches increasing at an even higher rate. This leads 
-to cache pollution and bandwidth contention, as is clearly visible with the reduction in 
-the IPC. Similar behavior was observed in 12 other SPEC CPU 2017 simpoints.
+Figure~\ref{Fig:Motivation} illustrates the above scenario.  Here we
+consider a state-of-the-art. lookahead prefetcher -- SPP~\cite{SPP}.
+Lookahead prefetchers such as SPP provide a mechanism to speculate an
+arbitrary number of references ahead of the initial triggering miss.
+In SPP, a throttling confidence threshold is then used to ensure that
+the lookahead stops when confidence falls too low to ensure that
+prefetches are accurate.  In the figure, we remove this threshold and
+instead allow the prefetcher to lookahead an of arbitrary depth from 7
+to 15.  The figure depicts the behavior of the {\tt 603.bwaves\_s}
+SPEC CPU 2017 benchmark.  The IPC, total number of prefetches
+suggested by the prefetcher (TOTAL\_PF) and the actual useful
+predictions (GOOD\_PF), all have been normalized to the least
+aggressive version of the prefetcher (depth 7).  As the look-ahead
+depth increases, so do useful prefetches, and hence coverage.  This
+coverage, however, comes at a cost of total prefetches increasing at
+an even higher rate.  This leads to cache pollution and bandwidth
+contention, as is clearly visible with the reduction in IPC.  Similar
+behavior was observed in 12 other SPEC CPU 2017 benchmarks.
 
-Therefore, a delicate balance between coverage and accuracy is required for a prefetcher to maximize
-its performance impact.  Prefetchers are generally designed with internal mechanisms to
-monitor their accuracy, with throttling thresholds that can be tuned for either coverage 
-or accuracy.  The more irregular an application's memory access pattern is, the more difficult
-it is to accurately predict every access, so a prefetcher will have to be tuned more toward
-coverage (and away from accuracy) in order to gain any benefit.  This may be especially dangerous
-to do in the context of a multi-core processor, where overly aggressive prefetching in one core can
-waste shared resources, such as last-level cache (LLC) capacity, and off-chip bandwidth~\cite{Friendly}.
+Therefore, a delicate balance between coverage and accuracy is
+required for a prefetcher to maximize its performance impact.
+Prefetchers are generally designed with internal mechanisms to monitor
+their accuracy, and throttling mechanisms that can be tuned for either
+coverage or accuracy.  The more irregular an application's memory
+access pattern is, the more difficult it is to accurately predict
+every access, so a prefetcher will have to be tuned more toward
+coverage (and away from accuracy) in order to gain any benefit. This
+may be especially dangerous to do in the context of a multi-core
+processor, where overly aggressive prefetching in one core can waste
+shared resources, such as last-level cache (LLC) capacity, and
+off-chip bandwidth, impacting the performance of other
+cores~\cite{Friendly}.
 
-%Predicting future accesses requires a trade-off between coverage, the percent of predictions that 
-%prevented a cache miss, and accuracy, the percent of predictions made that were correct. 
-%%Gino: Just added a short explanation of coverage and accuracy tradeoff, not sure if necessary
-%A prefetcher may have high coverage by reducing the number of misses by simply requesting a large 
-%amount of data, however this does not imply accuracy. Even if the prefetcher brought in data accessed in the future,
-%a high percentage of the data requested could never be accessed meaning the prefetcher had low accuracy, 
-%resulting in wasted resources as well as cache pollution. Likewise, a prefetcher can have a 
-%higher accuracy by being completely sure the data requested is used in the future, 
-%but may not affect performance since it did not have the coverage needed to make an impact. %end addition.
-%Most prefetchers maintain an internal confidence mechanism by keeping counters
-%to track events such as cache hits, misses, references to an entry in the prefetcher's
-%structure, etc. By comparing the confidence indicated by the counters against different 
-%threshold values, the aggressiveness of the prefetcher can be adjusted, resulting in lower
-%coverage and higher accuracy.
+% Predicting future accesses requires a trade-off between coverage,
+% the percent of predictions that prevented a cache miss, and
+% accuracy, the percent of predictions made that were correct.
+%% Gino: Just added a short explanation of coverage and accuracy
+%% tradeoff, not sure if necessary
+%A prefetcher may have high coverage by reducing the number of misses
+% by simply requesting a large amount of data, however this does not
+% imply accuracy. Even if the prefetcher brought in data accessed in
+% the future, a high percentage of the data requested could never be
+% accessed meaning the prefetcher had low accuracy, resulting in
+% wasted resources as well as cache pollution. Likewise, a prefetcher
+% can have a higher accuracy by being completely sure the data
+% requested is used in the future, but may not affect performance
+% since it did not have the coverage needed to make an
+% impact. %end addition.
+% Most prefetchers maintain an internal confidence mechanism by
+% keeping counters to track events such as cache hits, misses,
+% references to an entry in the prefetcher's structure, etc. By
+% comparing the confidence indicated by the counters against different
+% threshold values, the aggressiveness of the prefetcher can be
+% adjusted, resulting in lower coverage and higher accuracy.
 
-%For a prefetcher to capture highly irregular access patterns, it needs to be
-%highly aggressive with high coverage. This causes the prefetcher to recommend many
-%low accuracy prefetches that lead to the cache being polluted with useless data.  
-%This effect is more pronounced in a multi-core scenario where the LLC is a
-%shared resource~\cite{Friendly}.  Conversely, a conservative prefetcher would
-%be highly accurate but might not make enough useful prefetches due to its lower coverage.
+%For a prefetcher to capture highly irregular access patterns, it
+% needs to be highly aggressive with high coverage. This causes the
+% prefetcher to recommend many low accuracy prefetches that lead to
+% the cache being polluted with useless data.  This effect is more
+% pronounced in a multi-core scenario where the LLC is a shared
+% resource~\cite{Friendly}. Conversely, a conservative prefetcher
+% would be highly accurate but might not make enough useful prefetches
+% due to its lower coverage.
 
-We propose Perceptron-based Prefetch Filtering (PPF) as an enhancement to
-existing state-of-the-art prefetchers.  PPF works by observing the stream of
-candidate prefetches generated by a prefetcher, and then rejects those that
-are predicted by the online-trained neural model to be inaccurate.
-The state-of-the-art prefetcher that we 
-use to evaluate PPF in this paper is the Signature Path Prefetcher 
-(SPP)~\cite{SPP}, however in principle, PPF could benefit any prefetcher.
-In this design, PPF replaces SPP's existing confidence-based throttling mechanism,
-which itself was a highly tuned feature of that prefetcher.  Because PPF is
-so much more effective at rejecting inaccurate prefetches than SPP's baseline
-mechanism, we are free to re-tune the rest of SPP's design around maximizing
-coverage.  The result is an increase in both accuracy and coverage, and
-a notable increase in performance.
+Here, we propose Perceptron-based Prefetch Filtering (PPF) as an
+enhancement to existing state-of-the-art prefetchers, allowing them to
+speculate deeply to achieve high coverage while filtering out the
+inaccurate prefetches this deep speculation implies.  PPF works by
+observing the stream of candidate prefetches generated by a
+prefetcher, and then rejects those that are predicted by the
+online-trained neural model to be inaccurate.  The state-of-the-art
+prefetcher that we use to evaluate PPF in this paper is the Signature
+Path Prefetcher (SPP)~\cite{SPP}, however as we describe, PPF can be
+designed to benefit any prefetcher.  In this design, PPF replaces
+SPP's existing confidence-based throttling mechanism, which itself was
+a highly tuned feature of that prefetcher.  Because PPF is so much
+more effective at rejecting inaccurate prefetches than SPP's baseline
+mechanism, we are free to re-tune the rest of SPP's design around
+maximizing coverage. The result is an increase in both accuracy and
+coverage, and a notable increase in performance.
 
@@
 
-\item An on-line neural model used for hardware data prefetching.  Previous
-work in this area either relied on program semantics~\cite{Semantics} or were
-application specific~\cite{Datacenter}.
+\item An on-line neural model used for hardware data prefetching.
+  Previous work in this area either relied on program
+  semantics~\cite{Semantics} or were application
+  specific~\cite{Datacenter}.
 
-\item Implementing PPF filtering a state-of-the-art prefetchier, giving a
-significant performance improvement compared to previous work.  PPF learns to
-adapt itself to shared resource constraints, leading to further increased
-performance in multi-core and bandwidth-constrained environments.
+\item Implementing PPF filtering a state-of-the-art prefetcher, giving
+  a significant performance improvement compared to previous work. PPF
+  learns to adapt itself to shared resource constraints, leading to
+  further increased performance in multi-core and
+  bandwidth-constrained environments.
 
 \item A methodology for determining an appropriate set of features for
-prediction, regardless of the underlying prefetcher used.  More details are
-explained in Section \ref{Method-Features}.
+  prediction, regardless of the underlying prefetcher used.  More
+  details are explained in Section \ref{Method-Features}.
 
@@
 In a single core configuration, PPF increases performance by 6.84\%
-compared to the underlying prefetcher, SPP.  In a multi-core system running a
-mixes of memory intensive SPEC CPU 2017 traces, PPF saw an improvement
-of 11.9\% over SPP for a 4-core system, and 10.67\% for an 8-core system.
+compared to the underlying prefetcher, SPP. In a multi-core system
+running a mixes of memory intensive SPEC CPU 2017 traces, PPF saw an
+improvement of 11.9\% over SPP for a 4-core system, and 10.67\% for an
+8-core system.  \pg{update numbers before submit}
 

### 2018-12-07 back.tex
minor

--- 2018-12-06 back.tex
@@
 
-In this section we discuss the most closely related work to our proposed
-technique. See Section~\ref{related} for other relevant work.
+In this section we discuss the most closely related work to our
+proposed technique. See Section~\ref{related} for other relevant work.
 
-The idea of prefetching begins with Jouppi's \textit{Instruction Stream
-Buffers}~\cite{ISB}. Early prefetchers detected stride access patterns in order to
-predict future memory requests~\cite{Smith,Baer,Stride}. Modern prefetching
-mechanisms are more sophisticated as they look into past memory
+The idea of prefetching begins with Jouppi's \textit{Instruction
+  Stream Buffers}~\cite{ISB}.  Early prefetchers detected stride access
+patterns in order to predict future memory
+requests~\cite{Smith,Baer,Stride}. Modern prefetching mechanisms are
+more sophisticated as they look into past memory
 behavior~\cite{Address_Correlated,AMPM},
 locality~\cite{Spatial_Pattern,SMS,Temporal_Instruction_Fetch,Off_Chip,STMS,SMS_JILP},
-control-flow speculation~\cite{BFetch,MTBFetch}, and many other aspects to
-detect complex access patterns in modern applications.
+control-flow speculation~\cite{BFetch,MTBFetch}, and many other
+aspects to detect complex access patterns in modern applications.
 
@@
 
-%% Simple stride prefetching techniques only detect sequences of addresses that differ by a constant value and fail to capture diverse memory access patterns. 
-Lookahead prefetchers can learn complicated patterns by collecting histories of
-observed data access patterns.
+%% Simple stride prefetching techniques only detect sequences of
+%% addresses that differ by a constant value and fail to capture
+%% diverse memory access patterns.
+Lookahead prefetchers can learn complicated patterns by collecting
+histories of observed data access patterns.
 %EB: Didn't really understand this: 
-%% "and correlating these with the strides, by considering each new page in a vacuum."
-Kim {\em et. al.} proposed
-SPP~\cite{SPP}, a confidence-based lookahead prefetcher. SPP creates a
-signature associated to a memory request by compressing the history of
-accesses. By correlating the signature with future likely delta patterns, SPP
-learn both simple and complicated memory access patterns quickly. Our proposed
+%% "and correlating these with the strides, by considering each new
+%% page in a vacuum."
+Kim {\em et. al.} proposed SPP~\cite{SPP}, a confidence-based
+lookahead prefetcher.  SPP creates a per-page signature by compressing
+the history of recent accesses to a given page.  By correlating the
+signature with future likely access patterns, SPP learn both simple
+and complicated memory access patterns quickly.  Our proposed
 technique builds upon SPP. Thus, we describe SPP in more detail.

### 2018-12-07 back.tex

--- 2018-12-07 back.tex
@@
 
-In this section we discuss the most closely related work to our proposed technique. 
-The idea of prefetching begins with Jouppi's \textit{Instruction Stream
-Buffers}~\cite{ISB}. Early prefetchers detected stride access patterns in order to
-predict future memory requests~\cite{Smith,Baer,Stride}. Modern prefetching
-mechanisms are more sophisticated as they look into past memory
+In this section we discuss the most closely related work to our
+proposed technique.  The idea of prefetching begins with Jouppi's
+\textit{Instruction Stream Buffers}~\cite{ISB}. Early prefetchers
+detected stride access patterns in order to predict future memory
+requests~\cite{Smith,Baer,Stride}. Modern prefetching mechanisms are
+more sophisticated as they look into past memory
 behavior~\cite{Address_Correlated,AMPM},
 locality~\cite{Spatial_Pattern,SMS,Temporal_Instruction_Fetch,Off_Chip,STMS,SMS_JILP},
-control-flow speculation~\cite{BFetch,MTBFetch}, and other other aspects to
-detect complex memory access patterns.
-See Section~\ref{related} for other relevant work.
+control-flow speculation~\cite{BFetch,MTBFetch}, and other other
+aspects to detect complex memory access patterns.  See
+Section~\ref{related} for other relevant work.
 
@@
 %% Simple stride prefetching techniques only detect sequences of addresses that differ by a constant value and fail to capture diverse memory access patterns. 
-%Lookahead prefetchers can learn complicated patterns by collecting histories of observed data access patterns.
-%EB: Didn't really understand this: 
+% Lookahead prefetchers can learn complicated patterns by collecting
+% histories of observed data access patterns.  EB: Didn't really
+% understand this:
 %% "and correlating these with the strides, by considering each new page in a vacuum."
@@
 Kim {\em et. al.} proposed Signature Path Prefetcher (SPP)~\cite{SPP},
-a confidence-based lookahead prefetcher. 
-SPP creates a signature associated with a page address by compressing the history of 
-accesses. By correlating the signature with future likely delta patterns, SPP learns 
-both simple and complicated memory access patterns quickly. 
-While the basic idea of perceptron based prefetch filtering is applicable to
-any look-ahead prefetcher, we develop a practical implementation of the idea
-using SPP as our baseline. Here we describe the basic architecture of SPP.\\
-%
-\noindent \textbf{Signature Table:} As shown in figure \ref{fig:spp_structure}, 
-the Signature Table keeps track of 256 most recently
-accessed pages. It is meant to capture memory access patterns within a page
-boundary. SPP indexes into an entry of the Signature Table using the page
-number. For each entry corresponding to a page, the table stores a `last
-block offset' and an `old signature'. Last block offset is the block offset
-of the last memory access of that given page. The block offset is calculated
-with respect to the page boundary. The signature is a 12-bit compressed
-representation of the past few memory accesses for that page. The signature
-is calculated as: $$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\;
-(\,Delta\,)$$ Delta is the numerical difference between the block offset of
-the current and the previous memory access. In case a matching page entry is
-found, the stored signature retrieved and used to index into the Pattern
-Table. This process is illustrated by figure \ref{fig:spp_update}.\\
-%
-\noindent \textbf{Pattern Table:} The Pattern Table is indexed by the signature
-generated from the Signature Table. Pattern Table holds predicted delta
-patterns and their confidence estimates. Each entry indexed by the signature
-holds up to 4 unique delta predictions.\\
-%
+a confidence-based lookahead prefetcher.  SPP creates a signature
+associated with a page address by compressing the history of
+accesses. By correlating the signature with future likely delta
+patterns, SPP learns both simple and complicated memory access
+patterns quickly.  While the basic idea of perceptron based prefetch
+filtering is applicable to any look-ahead prefetcher, we develop a
+practical implementation of the idea
+using SPP as our baseline. Here we describe the basic architecture of SPP.
+
+\noindent \textbf{Signature Table:} As shown in figure
+\ref{fig:spp_structure}, the Signature Table keeps track of 256 most
+recently accessed pages. It is meant to capture memory access patterns
+within a page boundary. SPP indexes into an entry of the Signature
+Table using the page number. For each entry corresponding to a page,
+the table stores a `last block offset' and an `old signature'. Last
+block offset is the block offset of the last memory access of that
+given page. The block offset is calculated with respect to the page
+boundary. The signature is a 12-bit compressed representation of the
+past few memory accesses for that page. The signature is calculated
+as:
+$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\;
+(\,Delta\,)$$ Delta is the numerical difference between the block
+offset of the current and the previous memory access. In case a
+matching page entry is found, the stored signature retrieved and used
+to index into the Pattern
+Table. This process is illustrated by figure \ref{fig:spp_update}.
+
+\noindent \textbf{Pattern Table:} The Pattern Table is indexed by the
+signature generated from the Signature Table. Pattern Table holds
+predicted delta patterns and their confidence estimates. Each entry
+indexed by the signature holds up to 4 unique delta predictions.
 
@@
 %% 	is stored in the Pattern Table.
-\noindent \textbf{Prefetch Table:} The Prefetch Table is a 1024 entry 1-way associative
-table that keeps a record of last few entries prefetched. This proves useful
-for updating the prefetcher states if a tracked prefetch leads to a demand hit
-or a cache eviction. \textit{Note:} The original paper refers to this as the
-Prefetch Filter but since in our context filter refers to the perceptron
-filter, we will be calling this structure as the Prefetch Table to avoid any
-confusion.\\
-%
-\noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes down the program
-speculation path using its own prefetch suggestion. Using current prefetch as
-the baseline, it re-accesses the Pattern Table to generate further prefetches.
-It repeats the cycle of accessing Pattern Table and updating the signature
-based on highest confidence prefetch from the last iteration. The iteration
-counter on which SPP manages to predict prefetch entries in the lookahead
-manner is characterized as its `depth'. While doing so, SPP also keeps
-compounding the confidence in each depth. Thus as depth increases, overall
-confidence keeps decreasing.\\
-%
-\noindent \textbf{Confidence Tracking}: The Pattern Table keeps a count of hits to each
-signature through a counter C\textsubscript{sig}. The number of hits for a
-given delta per signature are tracked using a counter C\textsubscript{delta}.
-The confidence for a given delta is approximated through C\textsubscript{d} =
-C\textsubscript{delta} / C\textsubscript{sig}. When SPP enters into a
-lookahead mode, the path confidence P\textsubscript{d} is given as:
+\noindent \textbf{Prefetch Table:} The Prefetch Table is a 1024 entry
+1-way associative table that keeps a record of last few entries
+prefetched. This proves useful for updating the prefetcher states if a
+tracked prefetch leads to a demand hit or a cache
+eviction. \textit{Note:} The original paper refers to this as the
+Prefetch Filter but since in our context filter refers to the
+perceptron filter, we will be calling this structure as the Prefetch
+Table to avoid any confusion.
+
+\noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes
+down the program speculation path using its own prefetch
+suggestion. Using current prefetch as the baseline, it re-accesses the
+Pattern Table to generate further prefetches.  It repeats the cycle of
+accessing Pattern Table and updating the signature based on highest
+confidence prefetch from the last iteration. The iteration counter on
+which SPP manages to predict prefetch entries in the lookahead manner
+is characterized as its `depth'. While doing so, SPP also keeps
+compounding the confidence in each depth. Thus as depth increases,
+overall confidence keeps decreasing.
+
+\noindent \textbf{Confidence Tracking}: The Pattern Table keeps a
+count of hits to each signature through a counter
+C\textsubscript{sig}. The number of hits for a given delta per
+signature are tracked using a counter C\textsubscript{delta}.  The
+confidence for a given delta is approximated through
+C\textsubscript{d} = C\textsubscript{delta} /
+C\textsubscript{sig}. When SPP enters into a lookahead mode, the path
+confidence P\textsubscript{d} is given as:
 $$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;

### 2018-12-07 design_arch.tex

@@
   \includegraphics[width=8cm]{PPF_Hierarchy}
+  \caption{PPF Architecture in the Memory Hierarchy}
   \label{fig:PPF_Hierarchy}
-  \caption{PPF Architecture in the Memory Hierarchy}
   \end{center}
@@
   \includegraphics[width=\columnwidth]{Datapath_Separate}
+  \caption{PPF Data Path and Operation}
   \label{fig:PPF_Datapath}
-  \caption{PPF Data Path and Operation}
   \end{center}
@@
 accuracy and area. A detailed explanation of the storage overhead of PPF can
-be found in Figure~\ref{tab:PPF_overhead}.\\
+be found in Section~\ref{Method-Overheads}.\\
 
@@
 for the suggested prefetch, and is thresholded against two different
-magnitudes: PERC\_THRESHOLD\_HI and PERC\_ THRESHOLD\_ LO.
+magnitudes: $\tau_{hi}$ and $\tau_{lo}$.
 
-Prefetches whose sum exceeds PERC\_THRESHOLD\_HI are placed into the L2 cache.
+Prefetches whose sum exceeds $\tau_{hi}$ are placed into the L2 cache.
 The higher confidence value hints the prefetch would be useful and should be
 prioritized. A prefetch for which the features result in a confidence value
-between PERC\_THRESHOLD\_LO and PERC\_THRESHOLD\_ HI is allocated in the larger
+between $\tau_{lo}$ and $\tau_{hi}$ is allocated in the larger
 LLC, as the filter is moderately confident of the future reuse of the cache
@@
 Suggested prefetches for which the features lead to a confidence value lower
-than PERC\_THRESHOLD\_LO are not prefetched, as the low confidence value
+than $\tau_{lo}$ are not prefetched, as the low confidence value
 represents that the perceptron has learn that similar set of features are
@@
 helping the filter adapt quickly to changes in memory behavior. These
-thresholds are referred to as POS\_UPDATE\_THRESHOLD ($\theta_p$) and \newline
-NEG\_UPDATE\_THRESHOLD ($\theta_n$), respectively for the positive and
-negative values of training saturation.
+thresholds are referred to as $\theta_p$ and $\theta_n$, 
+respectively for the positive and negative values of training saturation.
 

### 2018-12-07 intro.tex

@@
 \includegraphics[width=\columnwidth]{Motivation}
-\caption{Variation of IPC, Total Prefetches and Useful Prefetches with
-  lookahead depth.}
+\caption{Impact of aggressive prefetching on performance. X-asis
+  represents increasingly aggressive prefetch lookahead depth. All the
+  metrics have been normalized to the least aggressive
+  configuration. As can be seen, total number of prefetches increases
+  at a higher rate than the number of good prefetches, leading to IPC
+  reduction.}
 \label{Fig:Motivation}
@@
 
-In a single core configuration, PPF increases performance by 6.84\%
-compared to the underlying prefetcher, SPP. In a multi-core system
-running a mixes of memory intensive SPEC CPU 2017 traces, PPF saw an
-improvement of 11.9\% over SPP for a 4-core system, and 10.67\% for an
-8-core system.  \pg{update numbers before submit}
+In a single core configuration, PPF increases performance by 3.78\%
+compared to the underlying prefetcher, SPP. In a multi-core system running a
+mixes of memory intensive SPEC CPU 2017 traces, PPF saw an improvement
+of 11.4\% over SPP for a 4-core system, and 9.65\% for an 8-core system.
 

### 2018-12-07 method.tex

@@
 methodology, we identified 95 different program segments of 1 Billion
-instructions each.
-
+instructions each.\\
+%
 \noindent \textit{Single-core performance:} For single-core simulations, we use 200
@@
 achieved per application using the SimPoint methodology.\\
-
+%
 \noindent \textit{Multi-core performance:} For multi-application workloads, we generate
@@
 to 3 days to simulate one billion instructions.\\
-
+%
 \noindent \textit{Validation:} We cross-validated our PPF model using SPEC
@@
 
+\subsection{Overhead for PPF}
+\label{Method-Overheads}
 In this section, we analyze the hardware overhead required to

### 2018-12-07 back.tex
finished a pass

@@
 filtering is applicable to any look-ahead prefetcher, we develop a
-practical implementation of the idea
-using SPP as our baseline. Here we describe the basic architecture of SPP.
+practical implementation of our proposed perceptron prefetch filter
+using SPP as our baseline.  Here we describe the basic architecture of
+SPP.
 
-\noindent \textbf{Signature Table:} As shown in figure
-\ref{fig:spp_structure}, the Signature Table keeps track of 256 most
-recently accessed pages. It is meant to capture memory access patterns
-within a page boundary. SPP indexes into an entry of the Signature
-Table using the page number. For each entry corresponding to a page,
-the table stores a `last block offset' and an `old signature'. Last
-block offset is the block offset of the last memory access of that
-given page. The block offset is calculated with respect to the page
-boundary. The signature is a 12-bit compressed representation of the
-past few memory accesses for that page. The signature is calculated
-as:
+\noindent \textbf{Signature Table:} As shown on the right side of
+figure \ref{fig:spp_update}, the Signature Table keeps track of 256
+most recently accessed pages. It is meant to capture memory access
+patterns within a page boundary. SPP indexes into an entry of the
+Signature Table using the page number. For each entry corresponding to
+a page, the table stores a `last block offset' and an `old
+signature'. Last block offset is the block offset of the last memory
+access of that given page. The block offset is calculated with respect
+to the page boundary. The signature is a 12-bit compressed
+representation of the past few memory accesses for that page. The
+signature is calculated as:
 $$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\;
@@
 matching page entry is found, the stored signature retrieved and used
-to index into the Pattern
-Table. This process is illustrated by figure \ref{fig:spp_update}.
+to index into the Pattern Table.  This process is illustrated in
+figure \ref{fig:spp_update}.
 
-\noindent \textbf{Pattern Table:} The Pattern Table is indexed by the
-signature generated from the Signature Table. Pattern Table holds
-predicted delta patterns and their confidence estimates. Each entry
-indexed by the signature holds up to 4 unique delta predictions.
+\noindent \textbf{Pattern Table:} The Pattern Table, shown on the left
+side of figure~\ref{fig:spp_update} is indexed by the signature
+generated from the Signature Table.  Pattern Table holds predicted
+delta patterns and their confidence estimates. Each entry indexed by
+the signature holds up to 4 unique delta predictions.
 
@@
 %% 	is stored in the Pattern Table.
-\noindent \textbf{Prefetch Table:} The Prefetch Table is a 1024 entry
-1-way associative table that keeps a record of last few entries
-prefetched. This proves useful for updating the prefetcher states if a
-tracked prefetch leads to a demand hit or a cache
-eviction. \textit{Note:} The original paper refers to this as the
-Prefetch Filter but since in our context filter refers to the
-perceptron filter, we will be calling this structure as the Prefetch
-Table to avoid any confusion.
+
+%  -- This looks like a detail that is confusing and not necessary to
+%  understand SPP:
+% \noindent \textbf{Prefetch Table:} The Prefetch Table is a 1024 entry
+% direct-mapped table that keeps a record of last few entries
+% prefetched.  This proves useful for updating the prefetcher state if a
+% tracked prefetch leads to a demand hit or a cache
+% eviction. \textit{Note:} The original paper refers to this as the
+% Prefetch Filter but since in our context filter refers to the
+% perceptron filter, we will be calling this structure as the Prefetch
+% Table to avoid any confusion.
 
 \noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes
-down the program speculation path using its own prefetch
-suggestion. Using current prefetch as the baseline, it re-accesses the
-Pattern Table to generate further prefetches.  It repeats the cycle of
-accessing Pattern Table and updating the signature based on highest
-confidence prefetch from the last iteration. The iteration counter on
-which SPP manages to predict prefetch entries in the lookahead manner
-is characterized as its `depth'. While doing so, SPP also keeps
+down the program speculation path using its own prefetch suggestion.
+Using current prefetch as a starting point, it re-accesses the Pattern
+Table to generate further prefetches.  As illustrated in
+figure~\ref{fig:spp_structure}, it repeats the cycle of accessing
+Pattern Table and updating the signature based on highest confidence
+prefetch from the last iteration.  The iteration counter on which SPP
+manages to predict prefetch entries in the lookahead manner is
+characterized as its `depth'. While doing so, SPP also keeps
 compounding the confidence in each depth. Thus as depth increases,
@@
 
-\noindent \textbf{Confidence Tracking}: The Pattern Table keeps a
-count of hits to each signature through a counter
-C\textsubscript{sig}. The number of hits for a given delta per
-signature are tracked using a counter C\textsubscript{delta}.  The
-confidence for a given delta is approximated through
-C\textsubscript{d} = C\textsubscript{delta} /
+\noindent \textbf{Confidence Tracking}: As shown in
+figure~\ref{fig:spp_structure}, the Pattern Table keeps a count of
+hits to each signature through a counter C\textsubscript{sig}. The
+number of hits for a given delta per signature are tracked using a
+counter C\textsubscript{delta}.  The confidence for a given delta is
+approximated through C\textsubscript{d} = C\textsubscript{delta} /
 C\textsubscript{sig}.  When SPP enters into a lookahead mode, the path
@@
 P\textsubscript{d-1}$$ Here $\alpha$ represents the global accuracy,
-calculated as the ratio of the number of prefetches which led to a demand hit
-to the number of prefetches recommended in total. The range of $\alpha$ is [0,1].
-$D$ is the lookahead depth. For $d = 1$, when SPP is in
-non-speculative mode, P\textsubscript{0} can be thought of as 1. 
-The final P\textsubscript{d} is thresholded against prefetch threshold 
-(T\textsubscript{p}) to reject the low confidence suggestions and 
-then against a numerically bigger fill threshold (T\textsubscript{f}) to 
-decide whether the prefetch should be sent to L2 Cache 
-(high confidence prefetch) or Last Level Cache (low confidence prefetch).
-The two thresholds were empirically set to 25 and 90 respectively, 
-on the scale of 0 to 100.
+calculated as the ratio of the number of prefetches which led to a
+demand hit to the number of prefetches recommended in total. The range
+of $\alpha$ is [0,1].  $D$ is the lookahead depth. For $d = 1$, when
+SPP is in non-speculative mode, P\textsubscript{0} can be thought of
+as 1.  The final P\textsubscript{d} is thresholded against prefetch
+threshold (T\textsubscript{p}) to reject the low confidence
+suggestions and then against a numerically bigger fill threshold
+(T\textsubscript{f}) to decide whether the prefetch should be sent to
+L2 Cache (high confidence prefetch) or Last Level Cache (low
+confidence prefetch).  The two thresholds were empirically set to 25
+and 90 respectively, on the scale of 0 to 100.
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=\columnwidth]{SPP_Update_Description}
+  \caption{SPP Data-path Flow}
+  \label{fig:spp_update}
+  \end{center}
+\end{figure}
 
@@
   \label{fig:spp_structure}
-  \end{center}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-  \includegraphics[width=\columnwidth]{SPP_Update_Description}
-  \caption{SPP Data-path Flow}
-  \label{fig:spp_update}
   \end{center}
@@
 As was noted in Figure~\ref{Fig:Motivation}, an aggressive lookahead
-prefetching if done without any accuracy check, can harm the performance 
-of the system. In this given case, the effective performance gained
-by the original prefetcher is lost by almost 9\% by the aggressive version 
-of the same prefetcher. Most modern prefetchers are getting increasingly 
-aggressive as that enables them to capture the complex memory access patterns 
-seen in pointer-chasing applications. Thus, we need a mechanism that is 
-orthogonal to the underlying prefetching scheme and can be used
-to prune out the useful prefetches from the useless ones.
+prefetching, if done without any accuracy check, can harm the
+performance of the system. As the figure shows, aggressive lookahead
+and its accompanied loss of accuracy degrades performance by almost
+9\%.  This is despite a growing number of useful prefetches generated
+by the prefetcher.  Thus, we need a mechanism that is orthogonal to
+the underlying prefetching scheme and can be used to prune out the
+useful prefetches from the useless ones.
 
-Moreover, the on-line confidence mechanism used by most prefetchers is very
-rudimentary. For example, SPP's internal confidence mechanism is based on
-taking the ratio C\textsubscript{d} = C\textsubscript{delta} /
-C\textsubscript{sig}. This confidence was used to make the decision of
-whether to prefetch or not to prefetch; and which level to prefetch. While
-this approximation was shown to work in the original implementation, we
-believe that a better form of generalized on-line decision making was
-possible. Hence, it was necessary to build a robust and adaptable learning
-mechanism to accept / reject the prefetch suggestions; and to decide the fill
-level (L2 Cache vs Last Level Cache). Thus, we introduce an independent
-on-line perceptron based filtering mechanism.
+Moreover, the on-line confidence mechanism used by most prefetchers is
+very rudimentary. For example, SPP's internal confidence mechanism is
+based on taking the ratio C\textsubscript{d} = C\textsubscript{delta}
+/ C\textsubscript{sig}. This confidence was used to make the decision
+of whether to prefetch or not to prefetch; and which level to
+prefetch.  While this approximation was shown to work in the original
+implementation, we believe that a better form of generalized on-line
+decision making was possible. Hence, it was necessary to build a
+robust and adaptable learning mechanism to accept / reject the
+prefetch suggestions; and to decide the fill level (L2 Cache vs Last
+Level Cache). Thus, we introduce an independent on-line perceptron
+based filtering mechanism.
 
@@
 \label{sec:Background-Perceptron}
-Perceptron learning for microarchitectural prediction was introduced for
-branch prediction~\cite{PerceptronPredictor}. Our predictor uses a version of
-microarchitectural perceptron prediction known as the ``hashed perceptron''
-organization~\cite{HashedPerceptron}. As an abstract idea, a hashed perceptron
-predictor hashes several different features into values that index several
-distinct tables. Small integer weights are read out from the tables and
-summed. If the sum exceeds some threshold, a positive prediction is made, {\em
-e.g.} ``predict branch taken'' or ``allow the prefetch.'' Otherwise, a
-negative prediction is made. Once the ground truth is known, the weights
-corredponding to the prediction are incremented if the outcome was positive,
-or decremented if it was negative. This update only occurs if the prediction
-was incorrect or if the magnitude of the sum failed to exceed a threshold.
-Beyond branch prediction, perceptron learning has been applied to last-level
-cache reuse prediction~\cite{Perc_Reuse,Multiperspective}. In this paper, we
-apply it for the first time to prefetch filtering.
+Perceptron learning for microarchitectural prediction was introduced
+for branch prediction~\cite{PerceptronPredictor}. Our predictor uses a
+version of microarchitectural perceptron prediction known as the
+``hashed perceptron'' organization~\cite{HashedPerceptron}. As an
+abstract idea, a hashed perceptron predictor hashes several different
+features into values that index several distinct tables. Small integer
+weights are read out from the tables and summed. If the sum exceeds
+some threshold, a positive prediction is made, {\em e.g.} ``predict
+branch taken'' or ``allow the prefetch.'' Otherwise, a negative
+prediction is made. Once the ground truth is known, the weights
+corredponding to the prediction are incremented if the outcome was
+positive, or decremented if it was negative. This update only occurs
+if the prediction was incorrect or if the magnitude of the sum failed
+to exceed a threshold.  Beyond branch prediction, perceptron learning
+has been applied to last-level cache reuse
+prediction~\cite{Perc_Reuse,Multiperspective}. In this paper, we apply
+it for the first time to prefetch filtering.
 

### 2018-12-07 back.tex
Condensing for space, Coverage and CrossVal Images

@@
 suggestions and then against a numerically bigger fill threshold
-(T\textsubscript{f}) to decide whether the prefetch should be sent to
+(T\textsubscript{f}) to decide whether to send the prefetch to
 L2 Cache (high confidence prefetch) or Last Level Cache (low

### 2018-12-07 results.tex
Condensing for space, Coverage and CrossVal Images

@@
 \textbf{60\%} over the next best prefetcher for that trace -- BOP. In general, benchmarks
-{\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt 623. xalancbmk\_s} and {\tt
+{\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt {623.xalancbmk}\_s}~~and~~{\tt
 649.fotonik3d\_s}  benefit the most from PPF, with the speedup over SPP ranging
@@
 %
-\begin{figure}[h]
+\begin{figure}[ht]
 \includegraphics[width=\columnwidth]{Coverage}
@@
 
-\begin{figure}[ht]
-\includegraphics[width=0.8\columnwidth,height=4cm,keepaspectratio]{SPEC2006}
-\caption{SPEC CPU 2006 Single-Core IPC Speedup}
-\label{Fig:SPEC2006_1core}
-\end{figure}
+%% \begin{figure}[ht]
+%% \includegraphics[width=0.8\columnwidth,height=4cm,keepaspectratio]{SPEC2006}
+%% \caption{SPEC CPU 2006 Single-Core IPC Speedup}
+%% \label{Fig:SPEC2006_1core}
+%% \end{figure}
+%% 
+%% \begin{figure}[ht]
+%% \includegraphics[width=0.9\columnwidth,height=3.5cm,keepaspectratio]{CloudSuite}
+%% \caption{IPC Speedup for CloudSuite Workloads}
+%% \label{Fig:CloudSuite}
+%% \end{figure}
 
 \begin{figure}[ht]
-\includegraphics[width=0.9\columnwidth,height=3.5cm,keepaspectratio]{CloudSuite}
-\caption{IPC Speedup for CloudSuite Workloads}
-\label{Fig:CloudSuite}
+\begin{adjustwidth}{-0.4cm}{}
+\includegraphics[width=1.1\columnwidth]{CrossVal}
+\caption{IPC Speedup for Unseen Workloads}
+\label{Fig:CrossVal}
+\end{adjustwidth}
 \end{figure}
-
-%% \begin{figure}[ht]
-%%   \centering
-%%   \begin{subfigure}{0.5\columnwidth}
-%%     \centering
-%%     \includegraphics[width=0.45\columnwidth]{CloudSuite}
-%%     \caption{CloudSuite Workloads}
-%%   \end{subfigure}
-%%   %
-%%   \begin{subfigure}{0.5\columnwidth}
-%%     \centering
-%%     \includegraphics[width=0.45\columnwidth]{SPEC2006}
-%%     \caption{Single Core SPEC CPU 2006}
-%%   \end{subfigure}
-%%   \caption{Performance with Unseen Benchmarks}
-%%   \label{Fig:CrossVal}
-%% \end{figure}
 
@@
 
-Figure~\ref{Fig:SPEC2006_1core} shows the speed-up achieved by BOP, SPP and
-PPF on the memory intensive subset and the full SPEC CPU 2006 suite. 
+Figure~\ref{Fig:CrossVal}(a) shows the speed-up achieved by BOP, SPP and
+PPF on the memory intensive subset and the full SPEC CPU 2006 suite for a
+single-processor machine. 
 PPF provides a speedup of \textbf{36.3\%} over the baseline on the 
@@
 
-Figure~\ref{Fig:CloudSuite} shows the performance benefit comparison of all
+Figure~\ref{Fig:CrossVal}(b) shows the performance benefit comparison of all
 the prefetch schemes on 4 different applications in the CloudSuite benchmark.

### 2018-12-07 back.tex
misc changes

@@
 
-\noindent \textbf{Signature Table:} As shown on the right side of
-figure \ref{fig:spp_update}, the Signature Table keeps track of 256
+\noindent \textbf{Signature Table:} As shown on the left side of
+Figure~\ref{fig:spp_update}, the Signature Table keeps track of 256
 most recently accessed pages. It is meant to capture memory access
@@
 to index into the Pattern Table.  This process is illustrated in
-figure \ref{fig:spp_update}.
+Figure~\ref{fig:spp_update}.
 
-\noindent \textbf{Pattern Table:} The Pattern Table, shown on the left
-side of figure~\ref{fig:spp_update} is indexed by the signature
+\noindent \textbf{Pattern Table:} The Pattern Table, shown on the right
+side in Figure~\ref{fig:spp_update} is indexed by the signature
 generated from the Signature Table.  Pattern Table holds predicted
@@
 Table to generate further prefetches.  As illustrated in
-figure~\ref{fig:spp_structure}, it repeats the cycle of accessing
+Figure~\ref{fig:spp_structure}, it repeats the cycle of accessing
 Pattern Table and updating the signature based on highest confidence
@@
 \noindent \textbf{Confidence Tracking}: As shown in
-figure~\ref{fig:spp_structure}, the Pattern Table keeps a count of
+Figure~\ref{fig:spp_structure}, the Pattern Table keeps a count of
 hits to each signature through a counter C\textsubscript{sig}. The

### 2018-12-07 design_arch.tex
misc changes

@@
 placed between the prefetcher and the prefetch insertion queue, as illustrated
-in figure~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
+in Figure~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
 polluting the higher levels of the memory hierarchy.
@@
 \noindent \textbf{Recording}\\
-As shown in step 2 of figure~\ref{fig:PPF_Datapath} the prefetches that make
+As shown in step 2 of Figure~\ref{fig:PPF_Datapath} the prefetches that make
 it through the inference stage are recorded in the ``Prefetch Table''. The
@@
 the filter should have categorized as a useless prefetch. Thus, the
-corresponding features of the prefetch request that cause these block to be
-placed in the cache are used to index the tables of weights, and those weights
-are adjusted accordingly.
+corresponding features of the prefetch request are used to re-index the tables 
+of weights, and those weights are adjusted accordingly.
 

### 2018-12-07 intro.tex
Typo
misc changes

@@
 \includegraphics[width=\columnwidth]{Motivation}
-\caption{Impact of aggressive prefetching on performance. X-asis
+\caption{Impact of aggressive prefetching on performance. X-axis
   represents increasingly aggressive prefetch lookahead depth. All the
   metrics have been normalized to the least aggressive
-  configuration. As can be seen, total number of prefetches increases
+  configuration. As can be seen, total number of prefetches increase
   at a higher rate than the number of good prefetches, leading to IPC
@@
 the lookahead stops when confidence falls too low to ensure that
-prefetches are accurate.  In the figure, we remove this threshold and
-instead allow the prefetcher to lookahead an of arbitrary depth from 7
-to 15.  The figure depicts the behavior of the {\tt 603.bwaves\_s}
-SPEC CPU 2017 benchmark.  The IPC, total number of prefetches
-suggested by the prefetcher (TOTAL\_PF) and the actual useful
-predictions (GOOD\_PF), all have been normalized to the least
-aggressive version of the prefetcher (depth 7).  As the look-ahead
+prefetches are accurate.  In the figure, we iteratively re-tuned 
+this threshold and allowed the prefetcher to lookahead an of 
+arbitrary depth from 7 to 15. The figure depicts the behavior 
+of the {\tt 603.bwaves\_s} SPEC CPU 2017 benchmark. The IPC, the 
+total number of prefetches suggested by the prefetcher (TOTAL\_PF) 
+and the actual useful predictions (GOOD\_PF), all have been normalized 
+to the least aggressive version of the prefetcher. As the look-ahead
 depth increases, so do useful prefetches, and hence coverage. This
@@
 an even higher rate. This leads to cache pollution and bandwidth
-contention, as is clearly visible with the reduction in IPC.  Similar
-behavior was observed in 12 other SPEC CPU 2017 benchmarks.
+contention, as is clearly visible with the reduction in IPC.  
+%Similar behavior was observed in 12 other SPEC CPU 2017 benchmarks.
 

### 2018-12-07 method.tex
Condensing for space, Coverage and CrossVal Images
misc changes

@@
     \centering
+    \begin{adjustbox}{width=0.9\columnwidth}
     \begin{tabular}{|c|c|}
@@
     \end{tabular}
+    \end{adjustbox}
     \caption{Simulation Parameters}
@@
 variations: Low Bandwidth DRAM, where DRAM bandwidth is limited to 3.2 GB/s,
-and small LLC, where the LLC size is reduced to 512 KB All the multi-core
+and small LLC, where the LLC size is reduced to 512 KB. All the multi-core
 simulations are only done in the default configuration.
@@
 %has been used to characterize and measure the prefetch performance.
-We use the SPEC CPU 2017 benchmark suite~\cite{SPEC2017}. We use all the 20
-workloads available in the SPEC CPU 2017 suite. Using the SimPoint~\cite{SimPoint}
+%We use the SPEC CPU 2017 benchmark suite~\cite{SPEC2017}. 
+We use all the 20 workloads available in the SPEC CPU 2017 suite~\cite{SPEC 2017}. Using the SimPoint~\cite{SimPoint}
 methodology, we identified 95 different program segments of 1 Billion
@@
 For simplicity, the traces are arranged in an
-increasing order of contribution made by the. 
+increasing order of contribution made by the feature. 
 It can be seen that even features with a low
@@
 
-We initially came up a mix of 23 features. By studying
+We initially came up a set of 23 features. By studying
 cross correlation of each of these features against others in a 23$\times$23
@@
     \centering
+    \begin{adjustbox}{width=0.9\columnwidth}
     \begin{tabular}{|c|c|c|}
@@
     \end{tabular}
+    \end{adjustbox}
     \caption{Meta-data Stored in Prefetch Table}
@@
 \begin{table}[h]
-\begin{adjustwidth}{-0.1cm}{}
+\begin{adjustwidth}{}{}
     \centering
+    \begin{adjustbox}{width=0.8\columnwidth}
     \begin{tabular}{|c|c|c|c|}
@@
     \end{tabular}
+    \end{adjustbox}
     \caption{SPP-Perc Storage Overhead}

### 2018-12-07 related.tex
Mostly related to images, methodology, typos etc
Added description of DA-AMPM
misc changes

@@
 fixed-stride pattern can be identified and prefetched that is centered on 
-the current access. Pugsley \textit{et al.} propose the Sandbox Prefetcher~\cite{Sandbox},
+the current access. DRAM-Aware AMPM (DA-AMPM)~\cite{DA_AMPM}
+is a variant of AMPM that delays some prefetches so they can be issued 
+together with others in the same DRAM row, increasing bandwidth utilization.
+Pugsley \textit{et al.} propose the Sandbox Prefetcher~\cite{Sandbox},
 which analyzes candidate fixed-offset prefetchers in a sandboxed environment
@@
 Unlike spatial prefetchers, lookahead prefetchers take program order into
-account when they make predictions.  Shevgoor \textit{et al.} propose the Variable 
+account when they make predictions. Shevgoor \textit{et al.} proposed the Variable 
 Length Delta Prefetcher (VLDP)~\cite{VLDP} that correlates histories of

### 2018-12-07 results.tex
misc changes

@@
 prefetchers simulated. On the SPEC CPU 2017 benchmarks, PPF reduces misses by
-\textbf{62.5\%} and \textbf{82.8\%} in the L2 and LLC respectively. For the
-same benchmark, the next best prefetcher, DA-AMPM, covers \textbf{48.6\%} and
-\textbf{72.8\%} of the misses respectively.
+\textbf{75.5\%} and \textbf{86.9\%} in the L2 and LLC respectively. For the
+same benchmark, the next best prefetcher, DA-AMPM, covers \textbf{54.3\%} and
+\textbf{78.5\%} of the misses respectively.
 
@@
 increasing order of the speedup. PPF offers a speedup of \textbf{51.2\%} on
-these traces, and improvement of \textbf{11.4\%} over the baseline SPP,
+these traces, an improvement of \textbf{11.4\%} over the underlying SPP,
 \textbf{9.7\%} over the next DA-AMPM, and \textbf{16.9\%} over BOP.
@@
 
-Figure~\ref{Fig:CrossVal}(a) shows the speed-up achieved by BOP, SPP and
-PPF on the memory intensive subset and the full SPEC CPU 2006 suite for a
+Figure~\ref{Fig:CrossVal}(a) shows the performance benefit comparison of all
+the prefetch schemes on 4 different applications in the CloudSuite benchmark.
+In general, these applications are prefetch agnostic. Even so, PPF manages a
+\textbf{3.78\%} improvement over no prefetching, putting it ahead of the next
+best prefetcher, SPP, which provides a 3.08\% speedup.
+
+Figure~\ref{Fig:CrossVal}(b) shows the speed-up achieved 
+on the memory intensive subset and the full SPEC CPU 2006 suite for a
 single-processor machine. 
@@
 
-Figure~\ref{Fig:CrossVal}(b) shows the performance benefit comparison of all
-the prefetch schemes on 4 different applications in the CloudSuite benchmark.
-In general, these applications are prefetch agnostic. Even so, PPF manages a
-\textbf{3.78\%} improvement over no prefetching, putting it ahead of the next
-best prefetcher, SPP, which provides a 3.08\% speedup.
-
 We developed PPF to yield good performance on the SPEC CPU 2017 benchmarks.

### 2018-12-07 design_impl.tex
Mostly related to images, methodology, typos etc
done with pass

--- 2018-12-04 design_impl.tex
@@
 
-This section describes our implementation of PPF and the range of
-features that are used by the perceptron for correlating the
-prefetches.  For a practical implementation, we have used SPP as the
-base engine prefetcher.
+This section describes a case study implementation of PPF and the
+range of features that are used to determine the usefulness of
+prefetches.  Here, we have selected SPP as the base engine prefetcher.
 
-\subsection{Changes made to original SPP}
+\subsection{Changes Made to SPP}
 \label{Impl-Changes}
@@
 
-\begin{itemize}
-\item \textbf{Original Thresholds discarded}\newline In PPF, the
+\noindent \textbf{Original Thresholds Discarded:} In PPF, the
 perceptron sum is used to decide whether to prefetch or not, and the
 fill-level in case of prefetch.  Thus, the confidence thresholds --
-  T\textsubscript{f} and T\textsubscript{p} are no longer needed.
-
+T\textsubscript{f} and T\textsubscript{p} are no longer needed to
+throttle the prefetcher.  \pg{Note, however, as we discuss later, we
+  use these confidence thresholds a feature in PPF so the mechanism is
+  retained. -- right?}
+% djimenez: keeping comments on every line here so it doesn't screw up my
+%"roll-your-own" itemized list. don't put in any newlines.
+%
+%
 % EB: Can this point be removed?
@@
 %    this a big deal?  Seems like a detail}
+%
 
-\item \textbf{Enhanced Prefetch Table}\newline The prefetch table
+\pg{since you actually moved this table into PPF, most this discussion
+  should actually go in the PPF arch section, I think its generic to
+  all PPFs likely anyways.  Instead here I'd talk about how you
+  removed the prefetch filter that was part of SPP since its redundant
+  with the function of PPF}
+\noindent \textbf{Enhanced Prefetch Table:} The prefetch table
 introduced in the original SPP keeps track of up to 1024 L2 cache
-  prefetch suggestions.  For PPF, the prefetch table is modified to
-  store the features used for perceptron indexing..
-  This metadata is needed to retrieve the
-  values of the features to train the perceptron at a later stage in
-  time. The prefetch table mentioned above keeps track of the
+prefetch suggestions. For PPF, the prefetch table is modified to store
+the features used for perceptron indexing.  This metadata is needed to
+retrieve the values of the features to train the perceptron at a later
+stage in time. The prefetch table mentioned above keeps track of the
 prefetches that actually happened in the L2 Cache.  We introduce the
-  reject table to keep a track of the prefetches that were suggested
-  by the base SPP engine but were rejected by PPF.  In case it is
-  detected that a future demand fetch could have used this rejected
-  prefetch, the perceptron weights are updated to reflect that
-  misprediction.
-\end{itemize}
+reject table to keep a track of the prefetches that were suggested by
+the base SPP engine but were rejected by PPF. In case it is detected
+that a future demand fetch could have used this rejected prefetch, the
+perceptron weights are updated to reflect that misprediction.
 
@@
 decision with the program behavior.  All the features we used can be
-derived from the information available in the L2 Cache access stream.
-Our feature selection involved searching over a large space of 
-relevant perceptron features. Using the statistical methodology 
-outlined in Section~\ref{Method-Features}, we pruned the feature 
-set to a minimal yet relevant set of features
+derived from the information available in the L2 Cache access stream
+or are taken as metadata derived from the baseline prefetcher.  Our
+feature selection involved searching over a large space of relevant
+perceptron features.  Note that part of the process of tuning PPF to a
+specific prefetcher involves examining the available metadata in the
+prefetcher itself, and thus PPF is attuned to the baseline
+prefetcher's design. Using the statistical methodology outlined in
+Section~\ref{Method-Features}, we pruned the feature set to a minimal
+yet relevant set of features.
 % \pg{again you want to get at why
@@
 % space of possible features and these were the best.}
-
+%
 % [EB] Tried addressing. Does it look redundant when read after 'Feature Selection'
@@
 
-\begin{itemize}
-\item \textbf{Base Address}: The base address of the demand 
-access that triggers the prefetch corresponds to a stream of 
-future accesses that SPP and the PPF have seen before. 
-Therefore, PPF will correlate the past behavior of this address 
-to the confidence of a prefetch.
+\noindent \textbf{Base Address}: The base address of the demand access
+that triggers the prefetch corresponds to a stream of future accesses
+that SPP and the PPF have seen before. Therefore, PPF will correlate
+the past behavior of this address to the confidence of a
+prefetch. \pg{what do you mean by ``base address'' the full physical
+  address?  Be specific.  Which bits and why? You should probably also
+  talk about what you do when the number of bits is unwieldy.}
 
-\item \textbf{Cache Line} and \textbf{Page Address}: These two
-  separate features are derived from the base address that triggered
-  the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and
-  base\_addr >> LOG2\_PAGE\_SIZE respectively. The idea behind using
-  three different shifted versions of the same feature is that it
-  allows us to look into a wider range of bits than with a single
-  version.  It also helps give more importance to the overlapping bits
-  and lesser importance to most and least significance bits. This
-  approach can also eliminate destructive interference that can be
-  caused by directly folding the address bits into half.
+\noindent \textbf{Cache Line} and \textbf{Page Address}: These two
+separate features are derived from the base address that triggered the
+prefetch, as follows: $base\_addr >> LOG2\_BLOCK \_SIZE$ and
+$base\_addr >> LOG2\_PAGE\_SIZE$ respectively. The idea behind using
+three different shifted versions of the same feature is that it allows
+the filter to focus its examination in more detail on different
+aspects of the address than with a single version.  It also helps give
+more importance to the overlapping bits and lesser importance to most
+and least significance bits.  This approach can also eliminate
+destructive interference that can be caused by directly folding
+the address bits into half.
 
+%
 % djimenez: this is a nice intuition and explanation. 
 % [EB] Thank you!
+\noindent \textbf{Program Counter XOR Depth}: The PC is for the
+instruction that triggered the prefetch chain.  Depth refers to the
+iteration count of the lookahead stages.  By itself, we find the PC to
+not be a good basis for filtering a lookahead prefetcher, as all the
+prefetches with depth >= 1 are aliased into the same PC, which will
+not be the PC of the eventual actual demand access.  This feature
+resolves a PC into a different value for each lookahead depth of
+prefetch speculation, giving a more accurate correlation in lookahead
+cases.  This is akin to the concept of Virtual Program
+Counters~\cite{VPC} introduced by Kim \textit{et. al.} for indirect
+branch prediction.
 
-\item \textbf{Program Counter XOR Depth}: The PC is for the
-  instruction that triggered the prefetch chain.  Depth refers to the
-  iteration count of the lookahead stages.  In general, the PC is not
-  considered as a good basis for doing lookahead prefetching as all
-  the prefetches with depth >= 1 are aliased into the same PC which
-  is not true in an actual demand access.  This feature resolves
-  a PC into a different value for each lookahead depth of prefetch
-  speculation, giving a more accurate correlation in lookahead cases.
-  This is akin to the concept of Virtual Program Counters~\cite{VPC}
-  introduced by Kim \textit{et. al.} for indirect branch prediction.
+\noindent \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
+  PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC
+before the instruction that triggered the current prefetch.  Hashing
+together the last three PCs informs PPF about the path that led to the
+current demand access and helps capture and branching information of
+the current basic block.  PCs are shifted in the increasing order of
+history before being hashed together. This is done to avoid the
+resultant value of zero when 2 or more PCs are the same. Additionally,
+blurring the information as it gets older allows us to get a wider and
+yet more approximate look into the program's history.
 
-\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
-    PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$
-  PC before the instruction that triggered the current prefetch.
-  Hashing together the last three PCs tell PPF about the path that led
-  to the current demand access and helps capture and branching
-  information of the current basic block.  PCs are shifted in the
-  increasing order of history before being hashed together.  This is
-  done to avoid the resultant value of zero when 2 or more PCs are the
-  same.  Additionally, obfuscating the information as it gets old
-  allows us to get a wider and approximate look into the program's
-  history.
+\noindent \textbf{Program Counter XOR Delta}: This feature tells us if
+a given PC favors particular value(s) of delta.  As noted earlier,
+while the PC alone does not convey useful information, this hash
+resolves the PC into different values based on the tendency of that PC
+to favor a certain delta.  Thus, the dynamic nature of different
+instances of the same memory access instruction is captured here.
 
-\item \textbf{Program Counter XOR Delta}: This feature tells us if a
-  given PC favors particular value(s) of delta.  As noted earlier,
-  while the PC alone does not convey useful information, this
-  hash resolves the PC into different values based on the tendency of
-  that PC to favor a certain delta.  Thus, the dynamic nature of
-  different instances of the same memory access instruction is 
-  captured here.
  % GAC: Seems more predictable than dynamic if its capturing 
@@
  % correct in the past.
-\item \textbf{Confidence}: The integer confidence, on a scale of 0 to
-  100, that was used in the original SPP design. While the original 
-  confidence does not directly make the decision to prefetch,
-  PPF correlates it to the correctness of a proposed prefetch.
+\noindent \textbf{Confidence}: The confidence, on a scale of 0 to 100,
+used to throttle lookahead depth in the original SPP design.  While
+the original confidence does not directly make the decision to
+prefetch, PPF correlates it to the correctness of a proposed prefetch.
 While the original SPP may have dismissed a prefetch due to running
-  further into speculation, PPF can use the original confidence as indicator
-  of not only when prefetches become less confident, but also how likely
-  a low confident speculation is correct in the context of other features.
+further into speculation, PPF can use the original confidence as
+indicator of not only when prefetches become less confident, but also
+how likely a low confident speculation is correct in the context of
+other features.
 
-\item \textbf{Page Address XOR Confidence}: This feature scores the
-  tendency of each page to be prefetch friendly or prefetch averse. It
-  helps resolve a page into different entries depending on its
-  confidence for prefetching, which can vary during phases of a
-  program execution.
+\noindent \textbf{Page Address XOR Confidence}: This feature scores
+the tendency of each page to be prefetch friendly or prefetch averse.
+It helps resolve a page into different entries depending on its
+confidence for prefetching, which can vary during phases of a program
+execution.
 
-\item \textbf{Current Signature XOR Delta}: Recall from the discussion
-  of SPP in Section \ref{Background-SPP} that the new signature is
-  generated using the old signature and the current block delta.
+\noindent \textbf{Current Signature XOR Delta}: Recall from the
+discussion of SPP in Section~\ref{Background-SPP} that the new
+signature is generated using the old signature and the current block
+delta.  The result of this feature is the next signature that is
+predicted to be accessed based on the delta predicted by SPP. While
+``Current Signature XOR Delta'' is not the exact formula for
+generating the future signature, it gives an approximate idea of the
+path that the combination of these two values can lead to.
+
 %Gino: Added the sentence below, is this correct?
 %[EB]: It gives the next signature instead of next delta
-  The result of this feature is the next signature that is predicted to be
-  accessed based on the delta predicted by SPP. While ``Current
-  Signature XOR Delta'' is not the exact formula for generating the
-  future signature, it gives an approximate idea of the path that the
-  combination of these two values can lead to.
-
+%
 %%%%%%% Below are the features rejected when downsizing 14 -> 9
@@
 
-\end{itemize}
-
-Some of the composite features are derived from simple hashing (XOR)
-of two primary features.  There is always a question of usefulness of
-such composite features and the new information added.  We justify the
-choice of each feature by quantifying the contribution made towards
-predicting prefetch behavior, in Section \ref{Method-Features}.
-Finally, as noted above, each feature indexes into its independent
-entry of perceptron weights.
+As can be noted above, some composite features are derived from simple
+hashing (XOR) of two primary features. There is always a question of
+usefulness of such composite features and the new information added.
+We justify the choice of each feature by quantifying the contribution
+made towards predicting prefetch behavior, in Section
+\ref{Method-Features}.  Finally, as noted above, each feature indexes
+into its independent entry of perceptron weights.
 

### 2018-12-07 design_arch.tex
done with pass

@@
 possible. Often, some useful prefetches are generated long after the
-confidence of the prefetcher has fallen below a point, after which the
-performance degrades, due to the increase of inaccurate prefetches. However,
-in order to allow deep speculation in the prefetcher, inaccurate prefetches
-must be filtered out. We propose to leverage perceptron-based learning as a
-mechanism to differentiate between potentially useful deeply speculated
-prefetches and likely not-useful ones. The Perceptron Prefetch Filter (PPF) is
-placed between the prefetcher and the prefetch insertion queue, as illustrated
-in Figure~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
+confidence of the prefetcher has fallen below the point at which
+performance degrades due to the increase of inaccurate prefetches.  In
+order to allow deep speculation in the prefetcher, however, inaccurate
+prefetches must be filtered out.  We propose to leverage
+perceptron-based learning as a mechanism to differentiate between
+potentially useful deeply speculated prefetches and likely not-useful
+ones. The Perceptron Prefetch Filter (PPF) is placed between the
+prefetcher and the prefetch insertion queue, as illustrated in
+Figure~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
 polluting the higher levels of the memory hierarchy.
 
-Perceptron learning is a light-weight mechanism to pull together disparate
-forms of information and synthesize a decision from them. Our work considers
-a number of features corresponding to a prefetch, such as speculation depth,
-page address and offset, and uses this information as the inputs to our
-perceptron-based filter, in order to predict the usefulness of a prefetch. In
-the remainder of this section, we discuss our proposed technique: PPF. It
-enhances a base prefetcher, by filtering out not-useful prefetches. PPF is a
-generalized prefetch filtering mechanism that may be adapted to any prefetcher
-with appropriate modifications.
+Perceptron learning is a light-weight mechanism to pull together
+disparate forms of information and synthesize a decision from
+them. Our work considers a number of features corresponding to a
+prefetch, such as speculation depth, page address and offset, and uses
+this information as the inputs to our perceptron-based filter in order
+to predict the usefulness of a prefetch.  Here, we discuss our the
+design of our proposed perceptron prefetch filter (PPF).  PPF enhances
+a base prefetcher by filtering out predicted unused prefetches. PPF is
+a generalized prefetch filtering mechanism that may be adapted to any
+prefetcher with appropriate feature selection and modifications which
+we describe below.
 
@@
 
-Figure~\ref{fig:PPF_Datapath} shows the required structures of PPF, as well as
-the steps followed to filter out not-useful prefetches. The filter is
-organized as a set of tables, where each entry in the tables holds a
-\textit{weight}. For a configuration of PPF using $N$ number of features, $N$
-different tables of weights are needed. Each feature is used to index a
-distinct table. The number of entries of each table varies according to their
-corresponding feature, hence, different number of bits are needed to index
-different tables. Each weight is a 5-bit saturating counter ranging from -16
-to +15. We found that having 5-bit weights provide a good trade-off between
-accuracy and area. A detailed explanation of the storage overhead of PPF can
-be found in Section~\ref{Method-Overheads}.\\
+Figure~\ref{fig:PPF_Datapath} shows the microarchitecture of PPF, as
+well as the steps required to filter out not-useful prefetches. The
+perceptron filter is organized as a set of tables, where each entry in
+the tables holds a \textit{weight}.  For a configuration of PPF using
+$N$ number of features, $N$ different tables of weights are needed.
+Each feature is used to index a distinct table. The number of entries
+of each table varies according to the corresponding feature, hence,
+different number of bits are needed to index different tables. Each
+weight is a 5-bit saturating counter ranging from -16 to +15. We found
+that having 5-bit weights provides a good trade-off between accuracy
+and area. A detailed explanation of the storage overhead of PPF can
+be found in Section~\ref{Method-Overheads}.
 
-\noindent{\textbf{Inferencing}\\
+\pg{the text below would be easier to follow if you used the figure
+  more. make reference to the figure at the points that you are
+  talking about}
+
+\noindent \textbf{Inferencing}\\
 %ETG:is this like an explanation of how prefetching happens? 
 %EB: I see this as a short context of the datapath leading to the perc filter.
-The base prefetcher is triggered on every demand access to the L2 Cache. At
-this point, it has the opportunity to trigger a prefetch. If it decides to do
-so, it will also need to decide how many cache blocks to prefetch. These
-blocks can be either placed in the L2 or L3 cache according to the confidence
-of the prefetching mechanism. Once the base prefetcher is triggered, the
-suggested prefetch candidates are fed to the perceptron filter to determine
-the usefulness of these prefetches. The filter makes the decision to issue
-the prefetch suggestions from the base prefetcher. To make the decision, each
-feature corresponding to a suggested prefetch is used to index a table and all
-the corresponding weights are summed. The sum denotes the confidence value
-for the suggested prefetch, and is thresholded against two different
-magnitudes: $\tau_{hi}$ and $\tau_{lo}$.
+The base prefetcher is triggered on every demand access to the L2
+Cache.  At this point, it has the opportunity to trigger a prefetch.
+If it does so, it will also need to decide how many cache blocks to
+prefetch.  These blocks can be either placed in the L2 or L3 cache
+according to the confidence of the prefetching mechanism.  Once the
+base prefetcher is triggered, the suggested prefetch candidates are
+fed to the perceptron filter to determine the usefulness of these
+prefetches. The filter ultimately decides whether to issue the
+prefetch suggestions of the base prefetcher. To make the decision,
+each feature corresponding to a suggested prefetch is used to index a
+table and all the corresponding weights are summed. The sum denotes
+the confidence value for the suggested prefetch, and is thresholded
+against two different magnitudes: $\tau_{hi}$ and $\tau_{lo}$.
 
 Prefetches whose sum exceeds $\tau_{hi}$ are placed into the L2 cache.
-The higher confidence value hints the prefetch would be useful and should be
-prioritized. A prefetch for which the features result in a confidence value
-between $\tau_{lo}$ and $\tau_{hi}$ is allocated in the larger
-LLC, as the filter is moderately confident of the future reuse of the cache
-block, but not enough to possible pollute a significantly smaller L2.
-Suggested prefetches for which the features lead to a confidence value lower
-than $\tau_{lo}$ are not prefetched, as the low confidence value
-represents that the perceptron has learn that similar set of features are
-associated with non-useful prefetches. \newline
+The higher confidence value hints the prefetch would be useful and
+should be prioritized.  A prefetch for which the features result in a
+confidence value between $\tau_{lo}$ and $\tau_{hi}$ is allocated in
+the larger LLC, as the filter is moderately confident of the future
+reuse of the cache block, but not enough to possible pollute a
+significantly smaller L2.  Suggested prefetches for which the features
+lead to a confidence value lower than $\tau_{lo}$ are not prefetched,
+as the low confidence value represents that the perceptron learned
+that a similar set of features are associated with non-useful
+prefetches.
 
 \noindent \textbf{Recording}\\
-As shown in step 2 of Figure~\ref{fig:PPF_Datapath} the prefetches that make
-it through the inference stage are recorded in the ``Prefetch Table''. The
-prefetch table is a 1,024-entry, direct mapped structure that contains all
-metadata required to re-index the perceptron entries for training.
+As shown in step 2 of Figure~\ref{fig:PPF_Datapath} the prefetches
+that make it through the inference stage are recorded in the
+``Prefetch Table''. The prefetch table is a 1,024-entry, direct mapped
+structure that contains all metadata required to re-index the
+perceptron entries for training.
 
-In addition to the prefetch table mentioned above, PPF also maintains a
-1,024-entry direct-mapped ``Reject Table.'' If a prefetch suggestion is
-rejected by the perceptron layer, it is logged into the reject table. The
-table is used to train the perceptron to avoid false negatives \textit{i.e.},
-cases where the prediction suggested to reject the prefetch but the prefetch
-was proven to be useful based on the observed demand accesses to the L2. \\
+In addition to the prefetch table mentioned above, PPF also maintains
+a 1,024-entry direct-mapped ``Reject Table.'' If a prefetch suggestion
+is rejected by the perceptron layer, it is logged into the reject
+table. The table is used to train the perceptron to avoid false
+negatives \textit{i.e.}, cases where the prediction suggested to
+reject the prefetch but the prefetch was proven to be useful based on
+the observed demand accesses to the L2.
 
-\noindent{\textbf{Feedback and Data Retrieval}\\
-When there is an eviction or a demand access to the L2, training for both the
-base prefetcher and our filter mechanism is triggered. The address of the
-cache block that triggered training is used to index both the Prefetch and
-Reject tables. 10 bits of the address are used to index tables, and another 6
-bits to perform tag matching. Once the base prefetcher has suggested a
-prefetch request, the corresponding features are used to index the tables of
-weights. \newline
+\noindent \textbf{Feedback and Data Retrieval}\\
+When there is an eviction or a demand access to the L2, training for
+both the base prefetcher and our filter mechanism is triggered. The
+address of the cache block that triggered training is used to index
+both the Prefetch and Reject tables. 10 bits of the address are used
+to index tables, and another 6 bits to perform tag matching. Once the
+base prefetcher has suggested a prefetch request, the corresponding
+features are used to index the tables of weights.
 
@@
 
-\noindent{\textbf{Training}\newline
-The address from the demand request triggering the training is looked up in
-both tables. If the address is in the prefetch table and marked as valid, this
-hints the previous prediction was correct and this is a useful prefetch. We
+\noindent \textbf{Training}\newline The address from the demand
+request triggering the training is looked up in both tables. If the
+address is in the prefetch table and marked as valid, this hints the
+previous prediction was correct and this is a useful prefetch. We
 compute the sum of the corresponding weights. If the sum falls below a
-specific threshold, training occurs and the corresponding weights are adjusted
-accordingly. These thresholds are introduced in order to avoid over-training,
-helping the filter adapt quickly to changes in memory behavior. These
-thresholds are referred to as $\theta_p$ and $\theta_n$, 
-respectively for the positive and negative values of training saturation.
+specific threshold, training occurs and the corresponding weights are
+adjusted accordingly. These thresholds are introduced in order to
+avoid over-training, helping the filter adapt quickly to changes in
+memory behavior. These thresholds are referred to as $\theta_p$ and
+$\theta_n$, respectively for the positive and negative values of
+training saturation.
 
 On a cache block eviction, we look up the corresponding address in the
-prefetch table. If there is a valid entry with this address, the filter made a
-misprediction. The block was allocated in the L2 with a prefetch request that
-the filter should have categorized as a useless prefetch. Thus, the
-corresponding features of the prefetch request are used to re-index the tables 
-of weights, and those weights are adjusted accordingly.
+prefetch table. If there is a valid entry with this address, the
+filter made a misprediction. The block was allocated in the L2 with a
+prefetch request that the filter should have categorized as a useless
+prefetch. Thus, the corresponding features of the prefetch request are
+used to re-index the tables of weights, and those weights are adjusted
+accordingly.
 
@@
 
-Parallel to accessing the prefetch table, on a demand access, the reject table
-is accessed. Before the demand access triggers the next set of prefetches, the
-reject table is checked for a valid entry. A hit means that the corresponding
-cache block was initially suggested by the base prefetcher, but wrongly
-rejected by the perceptron filter. The perceptron filter learns from this and
-makes use of the corresponding features associated to the original prefetch
-request, which are stored in the reject table, to index the weights tables and
+Parallel to accessing the prefetch table, on a demand access, the
+reject table is accessed. Before the demand access triggers the next
+set of prefetches, the reject table is checked for a valid entry. A
+hit means that the corresponding cache block was initially suggested
+by the base prefetcher, but wrongly rejected by the perceptron
+filter. The perceptron filter learns from this and makes use of the
+corresponding features associated to the original prefetch request,
+which are stored in the reject table, to index the weights tables and
 adjust the weights accordingly.
 
-Prior work in this domain lacks the knowledge of which cache blocks should have 
-been prefetched, but were not. The implementation of the reject table,
-allows us to have this information, which can be use to further optimize
-our prefetching mechanism.
+\pg{what prior work? this reference is too oblique to be useful,
+  either be specific or drop this paragraph} Prior work in this domain
+lacks the knowledge of which cache blocks should have been prefetched,
+but were not. The implementation of the reject table, allows us to
+have this information, which can be use to further optimize our
+prefetching mechanism.
 
@@
 \label{Arch-Generalizing}
-The above discussion of PPF shows that it is highly modular and can be adapted
-to be used over any base prefetcher for increased prefetch accuracy. Only 
-two hooks are required between PPF and the baseline prefetcher.
-The first is to make sure that all the prefetch candidates of the prefetcher
-pass through the perceptron filter and if qualified, the metadata for
-perceptron indexing be stored. The second is needed when the feedback of a
-prior prefetch is available in form of a subsequent demand hit or cache
-eviction. In that case, the stored metadata needs to be retrieved to update
-the state of the perceptrons.
+The above discussion of PPF shows that it is highly modular and can be
+adapted to be used over any base prefetcher for increased prefetch
+accuracy.  Only two hooks are required between PPF and the baseline
+prefetcher.  The first is to make sure that all the prefetch
+candidates of the prefetcher pass through the perceptron filter and if
+qualified, the metadata for perceptron indexing be stored. The second
+is needed when the feedback of a prior prefetch is available in form
+of a subsequent demand hit or cache eviction. In that case, the stored
+metadata needs to be retrieved to update the state of the perceptrons.
+\pg{this isn't quite right, I thought we were taking some metadata
+  from the prefetcher itself like confidence.  You should explicitly
+  mention that prefetcher metadata can be used, update the figure to
+  show that metadata passing from SPP to the prefetcher and point to
+  the figure to show where that connection is}
 
 In general, PPF can be adapted to a new base prefetcher with only a
-few modifications:\\
-%
-\textbf{Enhancing the Base Prefetcher:} By tuning down any internal thresholds
-to increase its inherent aggressiveness. \\
-%
-\textbf{Inferencing and Storing:} All prefetch recommendations are tested
-using the perceptron inferencing algorithm. The perceptron's output,
-\textit{true} or \textit{false}, should be saved appropriately, along with all
-metadata required for perceptron indexing. \\
-%
-\textbf{Retrieving and Training:} When feedback for a prefetch is available,
-the previously stored metadata can be used to re-index into the perceptron
-entries and increment or decrement the weights. \\
-%
-\textbf{Feature Selection:} Perceptrons essentially normalize contributions
-from different features to get a single sum representing the final confidence.
-So the perceptron learning can be as good as the set of features chosen.
-Interestingly, this is what makes perceptron learning scalable, as it can
-easily learn to incorporate newer information in the form of new features.
-Some of the features we developed use information derived directly from
-program execution, agnostic to the baseline prefetcher. Beyond that, the
-feature set can be expanded to convey any useful information from the
-prefetcher to the perceptron filter.
+few modifications:
+
+\noindent \textbf{Make Base Prefetcher More Aggressive:} Tune down any
+internal thresholds or throttling mechanisms to increase its
+aggressiveness.
+
+\noindent \textbf{Inferencing and Storing:} All prefetch
+recommendations are tested using the perceptron inferencing algorithm.
+The perceptron's output, \textit{true} or \textit{false}, should be
+saved appropriately, along with all metadata required for perceptron
+indexing.
+
+\noindent \textbf{Retrieving and Training:} When feedback for a
+prefetch becomes available, the previously stored metadata can be used
+to re-index into the perceptron entries and increment or decrement the
+weights.
+
+\noindent \textbf{Feature Selection:} Perceptrons essentially
+integrate contributions from different features to get a single sum
+representing the final confidence.  Thus, perceptron learning can only
+be as good as the set of features chosen.  Interestingly, this is what
+makes perceptron learning scalable, as it can easily learn to
+incorporate newer information in the form of new features.  Some of
+the features we developed use information derived directly from
+program execution, agnostic to the baseline prefetcher.  Beyond that,
+the feature set can be expanded to convey any useful information or
+metadata available in the baseline prefetcher.
 

### 2018-12-07 design_arch.tex
Updating Pval Image, Incorporating suggestions in Design

@@
 
-\pg{the text below would be easier to follow if you used the figure
-  more. make reference to the figure at the points that you are
-  talking about}
+%%\pg{the text below would be easier to follow if you used the figure
+%%  more. make reference to the figure at the points that you are
+%%  talking about}
 
@@
 The base prefetcher is triggered on every demand access to the L2
-Cache.  At this point, it has the opportunity to trigger a prefetch.
+Cache, as seen in Figure~\ref{fig:PPF_Hierarchy}.  At this point, 
+it has the opportunity to trigger a prefetch.
 If it does so, it will also need to decide how many cache blocks to
@@
 prefetches. The filter ultimately decides whether to issue the
-prefetch suggestions of the base prefetcher. To make the decision,
+prefetch suggestions of the base prefetcher. As shown in step 1
+of Figure~\ref{fig:PPF_Datapath}(a), to make the decision,
 each feature corresponding to a suggested prefetch is used to index a
@@
 the confidence value for the suggested prefetch, and is thresholded
-against two different magnitudes: $\tau_{hi}$ and $\tau_{lo}$.
+against two different values: $\tau_{hi}$ and $\tau_{lo}$.
 
@@
 \noindent \textbf{Recording}\\
-As shown in step 2 of Figure~\ref{fig:PPF_Datapath} the prefetches
+As shown in step 2 of Figure~\ref{fig:PPF_Datapath}(a) the prefetches
 that make it through the inference stage are recorded in the
@@
 \noindent \textbf{Feedback and Data Retrieval}\\
-When there is an eviction or a demand access to the L2, training for
+As depicted in step 3, when there is an eviction or a demand access 
+to the L2, training for
 both the base prefetcher and our filter mechanism is triggered. The
@@
 
-\noindent \textbf{Training}\newline The address from the demand
+\noindent \textbf{Training}\newline As can be seen in step 4 of, 
+the address from the demand
 request triggering the training is looked up in both tables. If the
@@
 adjust the weights accordingly.
-
-\pg{what prior work? this reference is too oblique to be useful,
-  either be specific or drop this paragraph} Prior work in this domain
-lacks the knowledge of which cache blocks should have been prefetched,
-but were not. The implementation of the reject table, allows us to
-have this information, which can be use to further optimize our
-prefetching mechanism.
+The implementation of the reject table, allows us to capture the 
+information of prefetches that were rejected, and that can be used 
+to further optimize our prefetching mechanism.
+%
+%\pg{what prior work? this reference is too oblique to be useful,
+%  either be specific or drop this paragraph} 
+% Prior work in this domain
+% lacks the knowledge of which cache blocks should have been prefetched,
+% but were not.
 
@@
 adapted to be used over any base prefetcher for increased prefetch
-accuracy.  Only two hooks are required between PPF and the baseline
-prefetcher.  The first is to make sure that all the prefetch
-candidates of the prefetcher pass through the perceptron filter and if
-qualified, the metadata for perceptron indexing be stored. The second
-is needed when the feedback of a prior prefetch is available in form
-of a subsequent demand hit or cache eviction. In that case, the stored
+accuracy. As a first step, all the prefetch
+candidates of the prefetcher have to pass through the perceptron filter. 
+If qualified, the metadata for perceptron indexing has to be stored. 
+Next, when the feedback of a prior prefetch is available in form
+of a subsequent demand hit or cache eviction, the stored
 metadata needs to be retrieved to update the state of the perceptrons.
-\pg{this isn't quite right, I thought we were taking some metadata
-  from the prefetcher itself like confidence.  You should explicitly
-  mention that prefetcher metadata can be used, update the figure to
-  show that metadata passing from SPP to the prefetcher and point to
-  the figure to show where that connection is}
+% \pg{this isn't quite right, I thought we were taking some metadata
+%   from the prefetcher itself like confidence.  You should explicitly
+%   mention that prefetcher metadata can be used, update the figure to
+%   show that metadata passing from SPP to the prefetcher and point to
+%   the figure to show where that connection is}
 
@@
 
-\noindent \textbf{Make Base Prefetcher More Aggressive:} Tune down any
-internal thresholds or throttling mechanisms to increase its
+\noindent \textbf{Making Base Prefetcher More Aggressive:} By Tuning 
+down any internal thresholds or throttling mechanisms to increase its
 aggressiveness.
@@
 
+\noindent \textbf{Using Metadata from the Prefetcher:} Some of the 
+internal counters specific to the underlying prefetcher can be suitable candidates
+for the perceptron features. To make sure that the perceptron layer sees that, 
+the relevant meta-data must be exported from the prefetcher to PPF. This way,
+PPF can be optimized to work tightly-knit with the base prefetcher.
+
 % %[EB] This statement fits better in the next section 'Features used by

### 2018-12-07 design_impl.tex
Updating Pval Image, Incorporating suggestions in Design

@@
 
+\noindent \textbf{Exporting Features from SPP:} PPF uses the meta-data 
+specific to SPP, to build some of the perceptron features. These include
+the lookahead depth, signature and the confidence counter. These features 
+were made visible to PPF\\
+%
 \noindent \textbf{Original Thresholds Discarded:} In PPF, the
 perceptron sum is used to decide whether to prefetch or not, and the
-fill-level in case of prefetch.  Thus, the confidence thresholds --
-T\textsubscript{f} and T\textsubscript{p} are no longer needed to
-throttle the prefetcher.  \pg{Note, however, as we discuss later, we
-  use these confidence thresholds a feature in PPF so the mechanism is
-  retained. -- right?}
+fill-level in case of prefetch.  Thus, the confidence thresholds used 
+by SPP -- T\textsubscript{f} and T\textsubscript{p} are no longer needed to
+throttle the prefetcher and can be discarded.
+%
+%\pg{Note, however, as we discuss later, we
+%use these confidence thresholds a feature in PPF so the mechanism is
+%retained. -- right?}
 % djimenez: keeping comments on every line here so it doesn't screw up my
@@
 
-\pg{since you actually moved this table into PPF, most this discussion
-  should actually go in the PPF arch section, I think its generic to
-  all PPFs likely anyways.  Instead here I'd talk about how you
-  removed the prefetch filter that was part of SPP since its redundant
-  with the function of PPF}
-\noindent \textbf{Enhanced Prefetch Table:} The prefetch table
-introduced in the original SPP keeps track of up to 1024 L2 cache
-prefetch suggestions. For PPF, the prefetch table is modified to store
-the features used for perceptron indexing.  This metadata is needed to
-retrieve the values of the features to train the perceptron at a later
-stage in time. The prefetch table mentioned above keeps track of the
-prefetches that actually happened in the L2 Cache.  We introduce the
-reject table to keep a track of the prefetches that were suggested by
-the base SPP engine but were rejected by PPF. In case it is detected
-that a future demand fetch could have used this rejected prefetch, the
-perceptron weights are updated to reflect that misprediction.
+%% \pg{since you actually moved this table into PPF, most this discussion
+%%   should actually go in the PPF arch section, I think its generic to
+%%   all PPFs likely anyways.  Instead here I'd talk about how you
+%%   removed the prefetch filter that was part of SPP since its redundant
+%%   with the function of PPF}
+%% \noindent \textbf{Enhanced Prefetch Table:} The prefetch table
+%% introduced in the original SPP keeps track of up to 1024 L2 cache
+%% prefetch suggestions. For PPF, the prefetch table is modified to store
+%% the features used for perceptron indexing.  This metadata is needed to
+%% retrieve the values of the features to train the perceptron at a later
+%% stage in time. The prefetch table mentioned above keeps track of the
+%% prefetches that actually happened in the L2 Cache.  We introduce the
+%% reject table to keep a track of the prefetches that were suggested by
+%% the base SPP engine but were rejected by PPF. In case it is detected
+%% that a future demand fetch could have used this rejected prefetch, the
+%% perceptron weights are updated to reflect that misprediction.
 
@@
 
-\noindent \textbf{Base Address}: The base address of the demand access
-that triggers the prefetch corresponds to a stream of future accesses
-that SPP and the PPF have seen before. Therefore, PPF will correlate
-the past behavior of this address to the confidence of a
-prefetch. \pg{what do you mean by ``base address'' the full physical
-  address?  Be specific.  Which bits and why? You should probably also
-  talk about what you do when the number of bits is unwieldy.}
+\noindent \textbf{Physical Address}: Here we use the lower bits 
+of the physical address of 
+the demand access that triggers the prefetch. This address corresponds 
+to a  stream of accesses that SPP and the PPF have seen before. 
+Therefore, PPF will correlate the past behavior of this address 
+to the prefetch outcome. 
+%
+%\pg{what do you mean by ``base address'' the full physical
+%  address?  Be specific.  Which bits and why? You should probably also
+%  talk about what you do when the number of bits is unwieldy.}
 

### 2018-12-07 method.tex
fix link
Updating Pval Image, Incorporating suggestions in Design

@@
 %We use the SPEC CPU 2017 benchmark suite~\cite{SPEC2017}. 
-We use all the 20 workloads available in the SPEC CPU 2017 suite~\cite{SPEC 2017}. Using the SimPoint~\cite{SimPoint}
+We use all the 20 workloads available in the SPEC CPU 2017 suite~\cite{SPEC2017}. Using the SimPoint~\cite{SimPoint}
 methodology, we identified 95 different program segments of 1 Billion
@@
 This section describes the intuition and analysis that went behind developing
-the features. As noted earlier, we developed a set of nine features that allow
-the perceptron layer to correlate prefetching decision with the program
-context. To study the correlation across each feature, we study statistically
-the perceptron weights and try to interpret their distribution. \\ \\
+the perceptron features. As noted earlier, we developed a set of 
+nine features that allows the perceptron layer to correlate prefetching decision 
+with the program behaviour. To study the correlation across each feature, 
+we statistically examine the perceptron weights and try to 
+interpret their distribution. \\ \\
 %
 \textbf{Global Pearson's Correlation}\newline 
-For this experiment, we examine
+Here we examine
 the perceptron weights at the end of all trace execution by which time the
@@
 overall correlation provide useful correlation (magnitude > 0.5) for
-a significant number of traces. \\ \\
+a significant number of traces. This study motivated us to choose \textit{PC XOR DELTA}
+over \textit{Last Signature} as it provided useful correlation in at least some of the
+traces.\\ \\
 %
@@
 matrix, we identified pairs of features with correlation factor > 0.9 in
-magnitude and eliminated redundant features, reducing the feature count to 9.
+magnitude and eliminated redundant features, using guidance from Global and per-trace
+Pearson's factor of those features. By doing this, we reduced the feature count to 9.
 Thus, in the final implementation of PPF, no two features have a high

### 2018-12-07 results.tex
Analysis of xalancbmk and cactuBSSN

@@
 CPU 2017, PPF nearly matches or outperforms most of the prefetchers on 91 
-simpoints. PPF fails to match the improvement offered by BOP only for 
-{\tt 607.cactuBSSN\_s} and {\tt 654.roms\_s} traces.
+simpoints. 
+An interesting case to see here is {\tt 623.xalancbmk}. Despite SPP 
+underperforming on that application, PPF manages to considerably outperform 
+the next best prefetcher -- BOP. Since this benchmark has varying prefetch 
+deltas, SPP's conservative throttling mechanism catches that and quickly halts 
+prefetching at an average depth of 2.1. On the other hand, PPF's more efficient 
+accuracy check enables it to prefetch upto a lookahead depth of 3.3. Doing this, PPF 
+suggests 1.61 times more total prefetches and 2.53 times more 
+useful prefetches than SPP.
+
+The only benchmark where PPF fails to match the improvement offered by other prefetchers 
+is {\tt 607.cactuBSSN\_s}. \newline Based on our observation of prefetching behavior, we gather that
+BOP's aggressive and localised nature can tackle this workload really well; as opposed
+to SPPs lookahead nature. As a result, SPP, and hence PPF, underperform on this benchmark.
 

### 2018-12-07 back.tex
Fixed minor grammatical error

@@
 down the program speculation path using its own prefetch suggestion.
-Using current prefetch as a starting point, it re-accesses the Pattern
+Using the current prefetch as a starting point, it re-accesses the Pattern
 Table to generate further prefetches.  As illustrated in
@@
 
-As was noted in Figure~\ref{Fig:Motivation}, an aggressive lookahead
+As was noted in Figure~\ref{Fig:Motivation}, aggressive lookahead
 prefetching, if done without any accuracy check, can harm the

### 2018-12-07 related.tex
updated grammar, last paragraph still needs work

@@
 Unlike spatial prefetchers, lookahead prefetchers take program order into
-account when they make predictions. Shevgoor \textit{et al.} proposed the Variable 
-Length Delta Prefetcher (VLDP)~\cite{VLDP} that correlates histories of
+account when they make predictions. Shevgoor \textit{et al.} propose the Variable 
+Length Delta Prefetcher (VLDP)~\cite{VLDP}, which correlates histories of
 deltas between cache line accesses within memory pages with the next delta

### 2018-12-07 conc.tex
Mostly related to images, methodology, typos etc
Small changes throughout the paper
Update conc.tex

@@
 \label{Conclusion}
-In this paper, we introduce perceptron-based learning for prefetching.  The
-perceptron filter acts independently as a check on the underlying prefetch
-engine.  PPF improves performance over the base prefetcher by 11.9\%.  We
-believe that this is a robust and adaptable technique that can be used to
-enhance any existing and upcoming prefetchers.
+In this paper, we introduce Perceptron-Based Prefetch Filtering (PPF). PPF 
+acts as an independent check on the quality of predictions made by the
+underlying prefetch engine.  We evaluate PPF in the context of using SPP
+as the baseline prefetch engine, but in principle other prefetchers could
+be used.  PPF is so effective at filtering bad prefetches, that it enables
+prefetcher design to focus more on increasing coverage, and allows existing
+prefetchers to be re-tuned for higher coverage.  PPF improves performance 
+over the base prefetcher by up to 11.4\%. We believe that PPF is a robust
+and adaptable technique that can be used to enhance any existing prefetchers
+and can be a valuable tool in the design of future prefetchers.

### 2018-12-07 method.tex
finished pass

@@
 \label{Method-Model}
-For testing and comparing PPF, we use the ChampSim~\cite{Champsim} simulator.
-ChampSim is an enhanced version of the framework that was used for the 2nd
-Data Prefetch Championship (DPC-2)~\cite{DPC_2}. We model 1-core, 4-core, and
-8-core out-of-order machines. The details of the configuration parameters are
-summarized in Table~\ref{tab:Sim_params}.
+We use the ChampSim~\cite{Champsim} simulator for the evaluation of
+PPF against prior work techniques.  ChampSim is an enhanced version of
+the framework that was used for the 2nd Data Prefetch Championship
+(DPC-2)~\cite{DPC_2}, also used in the 2nd Cache Replacement
+Competition (CRC2)~\cite{CRC_2}.  We model 1-core, 4-core, and 8-core
+out-of-order machines.  The details of the configuration parameters
+are summarized in Table~\ref{tab:Sim_params}.
 
@@
 
-The block size is fixed at 64 bytes. Prefetching is only triggered at an L2
-cache demand accesses but could be directed to the L2 or last-level cache. No
-L1 data level prefetching is done. The LRU replacement policy is used on all
-levels of cache hierarchies. Branch prediction is done using the perceptron
-branch predictor~\cite{PerceptronPredictor}. The page size is 4KB. 
-ChampSim operates all the prefetchers strictly in the physical address space.
+The block size is fixed at 64 bytes.  Prefetching is only triggered
+upon L2 cache demand accesses but could be directed to the L2 or
+last-level cache.  No L1 data level prefetching is done. The LRU
+replacement policy is used on all levels of cache hierarchies. Branch
+prediction is done using the perceptron branch
+predictor~\cite{PerceptronPredictor}.  The page size is 4KB.  ChampSim
+operates all the prefetchers strictly in the physical address space.
 
@@
 \label{Method-AdditionalMem}
-The default single-core configuration simulates a 2MB LLC and a single channel
-DRAM with 12.8GB/s bandwidth. We extend the simulations to include memory
-constraints introduced in DPC-2. Specifically we look at the following two
-variations: Low Bandwidth DRAM, where DRAM bandwidth is limited to 3.2 GB/s,
-and small LLC, where the LLC size is reduced to 512 KB. All the multi-core
-simulations are only done in the default configuration.
+The default single-core configuration simulates a 2MB LLC and a single
+channel DRAM with 12.8GB/s bandwidth. We extend the simulations to
+include memory constraints introduced in DPC-2. Specifically we look
+at the following two variations: Low Bandwidth DRAM, where DRAM
+bandwidth is limited to 3.2 GB/s, and small LLC, where the LLC size is
+reduced to 512 KB.  All the multi-core simulations are only done in
+the default configuration.
 
@@
 %We use the SPEC CPU 2017 benchmark suite~\cite{SPEC2017}. 
-We use all the 20 workloads available in the SPEC CPU 2017 suite~\cite{SPEC2017}. Using the SimPoint~\cite{SimPoint}
-methodology, we identified 95 different program segments of 1 Billion
-instructions each.\\
-%
-\noindent \textit{Single-core performance:} For single-core simulations, we use 200
-million instructions to warm-up the microarchitectural structures and the next
-one billion instructions to do detailed simulations and collect run-time
-statistics. We report the IPC speedup over the baseline of no prefetching.
-The final numbers reported are the geometric mean of the weighted mean speedup 
-achieved per application using the SimPoint methodology.\\
-%
-\noindent \textit{Multi-core performance:} For multi-application workloads, we generate
-100 random mixes and another 100 mixes from the memory intensive subset of
-SPEC CPU 2017. For 4-core workload, 200 Million instructions are used for warm-up
-and additional 1 Billion instruction simulated for collecting statistics.
-Each CPU keeps executing its workload till the last CPU completes one billion
-instructions after warm-up. For collecting IPC and other data, only the first
-billion instructions are considered as the region of interest.
+We use all the 20 workloads available in the SPEC CPU 2017
+suite~\cite{SPEC2017}. Using the SimPoint~\cite{SimPoint} methodology,
+we identified 95 different program segments of 1 Billion
+instructions each.
+
+\noindent \textbf{Single-core performance:} For single-core
+simulations, we use the first 200 million instructions to warm-up the
+microarchitectural structures and the next one billion instructions to
+do detailed simulations and collect run-time statistics. We report the
+IPC speedup over the baseline of no prefetching.  The final numbers
+reported are the geometric mean of the weighted mean speedup achieved
+per application using the SimPoint methodology.
+
+\noindent \textbf{Multi-core performance:} For multi-application
+workloads, we generate 100 random mixes and another 100 mixes from the
+memory intensive subset of SPEC CPU 2017.  For 4-core workloads, 200
+Million instructions are used for warm-up and additional 1 Billion
+instruction simulated for collecting statistics.  Each CPU keeps
+executing its workload till the last CPU completes one billion
+instructions after warm-up.  For collecting IPC and other data, only
+the first billion instructions are considered as the region of
+interest.
 
@@
 
-We repeat the same process for 8-core workloads, correspondingly with 16MB
-LLC. The only difference is that 20 million warm-up instructions and 100
-million full instructions are executed. This is done so as to keep the
-simulation run-time within reasonable limits as a single 8-core mix takes up
-to 3 days to simulate one billion instructions.\\
-%
-\noindent \textit{Validation:} We cross-validated our PPF model using SPEC
-CPU 2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks. For
-single-core SPEC CPU 2006, we developed 94 simpoints spread across all the 29
-applications. For multi-core, we followed the same methodology as SPEC CPU 2017.
-For CloudSuite, we used the traces made available for the 2nd Cache
-Replacement Competition (CRC-2)~\cite{CRC_2}. The traces include 4 4-core
-applications with 6 distinct phases per application.
+We repeat the same process for 8-core workloads, correspondingly with
+16MB LLC. The only difference is that 20 million warm-up instructions
+and 100 million full instructions are executed. This is done so as to
+keep the simulation run-time within reasonable limits as a single
+8-core mix takes up to 3 days to simulate one billion instructions.
 
-In total, we used 285 traces representing workloads across 53 applications.
-Throughout the paper, we consider memory intensive subset as the
-applications with SimPoint weighted LLC MPKI > 1. This includes 11 out of 
-20 SPEC CPU 2017 applications. For SPEC CPU 2006, this includes 16 out of 
-29 applications.
+\noindent \textbf{Validation:} We cross-validated our PPF model using
+SPEC CPU 2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite}
+benchmarks. For single-core SPEC CPU 2006, we developed 94 simpoints
+spread across all the 29 applications.  For multi-core, we followed
+the same methodology as SPEC CPU 2017.  For CloudSuite, we used the
+traces made available for the 2nd Cache Replacement Competition
+(CRC-2)~\cite{CRC_2}. The traces include four 4-core applications with
+six distinct phases per application.
+
+In total, we used 285 traces representing workloads across 53
+applications.  Throughout the paper, we consider memory intensive
+subset as the applications with SimPoint weighted LLC MPKI > 1. This
+includes 11 out of 20 SPEC CPU 2017 applications. For SPEC CPU 2006,
+this includes 16 out of 29 applications.
 
@@
 \label{Method-Prefetchers}
-We compared PPF against three of the latest state of the art hardware-only
-prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern
-Matching (DA-AMPM) \cite{DA_AMPM} and Signature Path Prefetcher (SPP). BOP was
-the winner of 2nd Data Prefetching Championship. DA-AMPM is the enhanced
-version of AMPM, modified to account for DRAM row buffer locality. SPP has
-been shown to outperform BOP on SPEC CPU 2006 traces. For each of these, we
-compare their speedups taking no prefetching as the baseline.
+We compared PPF against three of the latest, state of the art
+hardware-only prefetchers: Best Offset Prefetcher (BOP), DRAM Aware -
+Access Map Pattern Matching (DA-AMPM) \cite{DA_AMPM} and Signature
+Path Prefetcher (SPP).  BOP was the winner of 2nd Data Prefetching
+Championship. DA-AMPM is the enhanced version of AMPM, modified to
+account for DRAM row buffer locality. SPP has been shown to outperform
+BOP on SPEC CPU 2006 traces~\cite{SPP}. For each of these, we compare
+their speedups taking the no prefetching case as the baseline.
 
@@
 \label{Method-Features}
-This section describes the intuition and analysis that went behind developing
-the perceptron features. As noted earlier, we developed a set of 
-nine features that allows the perceptron layer to correlate prefetching decision 
-with the program behaviour. To study the correlation across each feature, 
-we statistically examine the perceptron weights and try to 
-interpret their distribution. \\ \\
-%
-\textbf{Global Pearson's Correlation}\newline 
-Here we examine
-the perceptron weights at the end of all trace execution by which time the
-weights have settled to steady values. The weights obtained from running all
-the SPEC CPU 2017 traces are concatenated. The feature with bulk of the
-perceptron weights concentrated around 0 or small magnitude numbers show a
-weak correlation with the prefetching outcome. On the other hand, features
-with most of the weights saturated around highest value (+15) show a high
-positive correlation and the features with weights close to the lowest value
-(-16) show a strong negative correlation.
+This section describes the intuition and analysis that went behind
+developing the perceptron features.  As noted earlier, we developed a
+set of nine features that allow the perceptron layer to correlate
+prefetching decision with the program behavior.  To study the
+correlation across each feature, we statistically examine the
+perceptron weights and try to interpret their distribution.
+
+\noindent \textbf{Global Pearson's Correlation} Here we examine the
+perceptron weights at the end of all trace execution by which time the
+weights have settled to steady values. The weights obtained from
+running all the SPEC CPU 2017 traces are concatenated.  Features with
+a bulk of their perceptron weights concentrated around 0 or small
+magnitude numbers show a weak correlation with the prefetching
+outcome.  On the other hand, features with most of the weights
+saturated around highest value (+15) show a high positive correlation
+and the features with weights close to the lowest value (-16) show a
+strong negative correlation.
 
@@
 
-We plot a histogram for each feature depicting weights distribution from -16
-to +15 and generate the Pearson's correlation factor for that feature.
-Pearson's factor is a numerical measure ranging from -1 to 1 of the degree of
-linear correlation between two variables. The magnitude of Pearson's factor
-gives the extent of correlation and the sign indicates whether it is a
-positive correlation or a negative correlation. Values close to 0 suggest a
-low correlation or at times, noisy data. A value of +1/-1 suggests a
-perfectly linear positive / negative correlation respectively.
+We plot a histogram for each feature depicting weights distribution
+from -16 to +15 and generate the Pearson's correlation factor for that
+feature.  Pearson's factor is a numerical measure ranging from -1 to 1
+of the degree of linear correlation between two variables. The
+magnitude of Pearson's factor gives the extent of correlation and the
+sign indicates whether it is a positive correlation or a negative
+correlation. Values close to 0 suggest a low correlation while a value
+of +1/-1 suggests a perfectly linear positive / negative correlation
+respectively.
 
-As a part of our perceptron feature selection methodology, we explored a wide 
-variety of features to begin with. Features with a low 
+As a part of our perceptron feature selection methodology, we explored
+a wide variety of features to begin with. Features with a low
 Pearson's coefficient were rejected as they didn't provide much useful
-correlation. Figure~\ref{Fig:Hist} depicts the histogram distribution of 
-trained weights for two features. The first feature, 
+correlation.  Figure~\ref{Fig:Hist} depicts the histogram distribution
+of trained weights for two features.  The first feature,
 \textit{Confidence XOR Page address} has the highest observed P-value,
-hence was retained. On the other hand, the second feature, \textit{Last Signature}
-did not provide any meaningful correlation and hence was rejected. 
-This is visible from the bulk of its trained weights settling to near zero values.
+hence was retained.  On the other hand, the second feature,
+\textit{Last Signature} did not provide any meaningful correlation and
+hence was rejected.  This is visible from the bulk of its trained
+weights settling to near zero values.
 
 Figure~\ref{Fig:Pvals} shows all the features which are finally used,
-arranged in the increasing order of their Pearson's factor. 
-As can be seen 5 out of the 9 features provide a moderate to high 
-correlation, with the magnitude of P-value > 0.6. The single most 
-important feature, \textit{Confidence XOR Page address} helps provide
-a correlation to prefetch outcome with a factor of 0.90. \\ \\
-%
-\textbf{Per Trace Correlation} \newline Another important way to look
+arranged in the increasing order of their Pearson's factor.  As can be
+seen 5 out of the 9 features provide a moderate to high correlation,
+with the magnitude of P-value > 0.6. The single most important
+feature, \textit{Confidence XOR Page address} helps provide
+a correlation to prefetch outcome with a factor of 0.90.
+
+\noindent \textbf{Per Trace Correlation} Another important way to look
 at the perceptron features is to see how much their contribution
 varies across the traces. Here we give special attention to features
-with low P-values in. Figure~\ref{Fig:Pvals_Trace} shows the
-variation of P-values for three features : \textit{PC XOR Delta},
-\textit{Signature XOR Delta} and 
-\textit{PC XOR Depth}; across all the SPEC CPU 2017 traces. 
-For simplicity, the traces are arranged in an
-increasing order of contribution made by the feature. 
-It can be seen that even features with a low
-overall correlation provide useful correlation (magnitude > 0.5) for
-a significant number of traces. This study motivated us to choose \textit{PC XOR DELTA}
-over \textit{Last Signature} as it provided useful correlation in at least some of the
-traces.\\ \\
-%
-\textbf{Trimming Features Using Cross-correlation} \newline Beside
+with low P-values in. Figure~\ref{Fig:Pvals_Trace} shows the variation
+of P-values for three features : \textit{PC XOR Delta},
+\textit{Signature XOR Delta} and \textit{PC XOR Depth}; across all the
+SPEC CPU 2017 traces.  For simplicity, the traces are arranged in an
+increasing order of contribution made by the feature.  It can be seen
+that even features with a low overall correlation provide useful
+correlation (magnitude > 0.5) for a significant number of traces. This
+study motivated us to choose \textit{PC XOR DELTA} over \textit{Last
+  Signature} as it provided useful correlation in at least some of the
+traces.
+
+\noindent \textbf{Trimming Features Using Cross-correlation} Beside
 providing interesting insights into prefetching behavior, P-value can
@@
 
-As we examined correlation of each feature with the final outcome, we also
-studied correlation between the features. We used the above methodology to
-eliminate features providing little information that has already been captured
-in other features.
+As we examined correlation of each feature with the final outcome, we
+also studied correlation between the features.  We used the above
+methodology to eliminate features providing little information that
+has already been captured in other features.
 
-We initially came up a set of 23 features. By studying
-cross correlation of each of these features against others in a 23$\times$23
-matrix, we identified pairs of features with correlation factor > 0.9 in
-magnitude and eliminated redundant features, using guidance from Global and per-trace
-Pearson's factor of those features. By doing this, we reduced the feature count to 9.
-Thus, in the final implementation of PPF, no two features have a high
-correlation between them. This way we can be sure that each feature makes a
-contribution that cannot be captured using other features.
+We initially came up a set of 23 features.  By studying cross
+correlation of each of these features against others in a 23$\times$23
+matrix, we identified pairs of features with correlation factor > 0.9
+in magnitude and eliminated redundant features, using guidance from
+Global and per-trace Pearson's factor of those features.  By doing
+this, we reduced the feature count to 9.  Thus, in the final
+implementation of PPF, no two features have a high correlation between
+them. This way we can be sure that each feature makes a contribution
+that cannot be captured using other features.
 
-Secondly, studying the relative importance of each feature enabled us to vary
-the number of entries dedicated for each feature. Features with higher
-correlation, \textit{cache line} and \textit{page address} were given most
-importance and allowed full 12-bits of indexing. Features like \textit{PC XOR
-delta} and \textit{PC XOR depth} with a low overall P-value were allocated
-fewer entries in the feature table.
+Secondly, studying the relative importance of each feature enabled us
+to vary the number of entries dedicated for each feature. Features
+with higher correlation, \textit{cache line} and \textit{page address}
+were given most importance and allowed full 12-bits of indexing.
+Features like \textit{PC XOR delta} and \textit{PC XOR depth} with a
+low overall P-value were allocated fewer entries in the feature table.
 
@@
 In this section, we analyze the hardware overhead required to
-implement PPF. The Prefetch Table was enhanced to accommodate
-storing of meta-data for perceptron training. Table
-\ref{tab:PTable_metadata} depicts the meta-data stored for each entry in
-the Prefetch Table. Table \ref{tab:PPF_overhead} shows the total
-storage overhead of PPF implementation. The hardware budget for
-2nd Data Prefetching championship was 32 KB. Keeping that in mind 
-the considerable speedup PPF obtained over the winner, the extra hardware
-budget can be accounted for. The extra hardware also makes the
-overall scheme more scalable than SPP. In the original SPP paper, it
-was demonstrated that adding extra hardware brings little advantage in
-terms of performance gain. The newly added perceptron tables can be
-scaled to increase / decrease features depending on the permitted
-budget.
+implement PPF. The Prefetch Table was enhanced to accommodate storing
+of meta-data for perceptron training. Table \ref{tab:PTable_metadata}
+depicts the meta-data stored for each entry in the Prefetch Table.
+Table \ref{tab:PPF_overhead} shows the total storage overhead of PPF
+implementation. The hardware budget for 2nd Data Prefetching
+championship was 32 KB.  Keeping that in mind the considerable speedup
+PPF obtained over the winner, the extra hardware budget can be
+accounted for. The extra hardware also makes the overall scheme more
+scalable than SPP. In the original SPP paper, it was demonstrated that
+adding extra hardware brings little advantage in terms of performance
+gain. The newly added perceptron tables can be scaled to increase /
+decrease features depending on the permitted budget.
 
@@
 
-\footnotetext[1]{Components of Prefetch Table can be found in Table \ref{tab:PTable_metadata}}
-\footnotetext[2]{The Reject Table does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}
+\footnotetext[1]{Components of Prefetch Table can be found in Table
+  \ref{tab:PTable_metadata}.}
+
+\footnotetext[2]{The Reject Table does not need to maintain the useful
+  bit as that only applies for prefetches that ultimately made
+  through.}
 
@@
 update by +1 or -1. Thus, all the operations required for perceptron
-inferencing or updating the states of the perceptrons can be
-easily done in the time constraints of L2 Cache Accesses.
+inferencing or updating the states of the perceptrons can be easily
+done in the time constraints of L2 Cache Accesses.

### 2018-12-07 results.tex
done

@@
 
-This section discusses the results obtained from running PPF in terms of
-prefetch cache coverage and speedup. For the SPEC CPU 2017 benchmarks, first
-we present the results for single-threaded workloads then for multi-core
-workloads.
+This section discusses the results obtained from running PPF in terms
+of speedup and prefetch cache, for the SPEC CPU 2017 benchmarks.
+First, we present the results for single-threaded workloads then for
+multi-core workloads.
 
@@
 % djimenez: why is this figure described this way? it's not OK to just
-% highlight noticeable improvement. the benchmarks must be selected by some
-% unbiased criteria, e.g. some minimum MPKI under LRU or some minimum speedup
-% under a baseline prefetcher that isn't PPF.
+% highlight noticeable improvement. the benchmarks must be selected by
+% some unbiased criteria, e.g. some minimum MPKI under LRU or some
+% minimum speedup under a baseline prefetcher that isn't PPF.
 
-Figure~\ref{Fig:SPEC2017_1core} shows the IPC speedup obtained by PPF,
-compared to BOP, DA-AMPM and SPP; for the individual SPEC CPU 2017 
-applications, followed by memory intensive subset and finally the full
-suite. All the results have been normalized to the baseline of no prefetching.
+Figure~\ref{Fig:SPEC2017_1core} shows the single core speedup obtained
+by BOP, DA-AMPM, SPP and PPF for each of the individual SPEC CPU 2017
+applications, followed by the geomean of the memory intensive subset
+and finally the geomean across the full suite.  All the results have
+been normalized to the baseline of no prefetching.
 
@@
 
-PPF yields a geometric mean speedup of \textbf{26.95\%} over the baseline. 
-This is equivalent to \textbf{4.63\%} over DA-AMPM, \textbf{4.61\%} over BOP 
-and \textbf{3.78\%} over SPP. Out of the 95 simpoints developed for SPEC 
-CPU 2017, PPF nearly matches or outperforms most of the prefetchers on 91 
-simpoints. 
-An interesting case to see here is {\tt 623.xalancbmk}. Despite SPP 
-underperforming on that application, PPF manages to considerably outperform 
-the next best prefetcher -- BOP. Since this benchmark has varying prefetch 
-deltas, SPP's conservative throttling mechanism catches that and quickly halts 
-prefetching at an average depth of 2.1. On the other hand, PPF's more efficient 
-accuracy check enables it to prefetch upto a lookahead depth of 3.3. Doing this, PPF 
-suggests 1.61 times more total prefetches and 2.53 times more 
-useful prefetches than SPP.
+PPF yields a geometric mean speedup of \textbf{26.95\%} over the
+baseline.  This is equivalent to \textbf{4.63\%} over DA-AMPM,
+\textbf{4.61\%} over BOP and \textbf{3.78\%} over SPP.  Out of the 95
+simpoints developed for SPEC CPU 2017, PPF nearly matches or
+outperforms most of the prefetchers on 91 simpoints.  One interesting
+case here is {\tt 623.xalancbmk}.  Despite SPP under performing on that
+application, PPF manages to considerably outperform all prefetchers.
+Since this benchmark has varying prefetch deltas, SPP's conservative
+throttling mechanism catches that and quickly halts prefetching at an
+average depth of 2.1.  On the other hand, PPF's more efficient
+accuracy check enables it to prefetch upto a lookahead depth of
+3.3. Doing this, PPF suggests 1.61 times more total prefetches and
+2.53 times more useful prefetches than SPP.
 
-The only benchmark where PPF fails to match the improvement offered by other prefetchers 
-is {\tt 607.cactuBSSN\_s}. \newline Based on our observation of prefetching behavior, we gather that
-BOP's aggressive and localised nature can tackle this workload really well; as opposed
-to SPPs lookahead nature. As a result, SPP, and hence PPF, underperform on this benchmark.
+The only benchmark where PPF fails to match the improvement offered by
+other prefetchers is {\tt 607.cactuBSSN\_s}.  Based on our observation
+of prefetching behavior, we gather that BOP's aggressive and localized
+nature fits this workload very well; as opposed to SPPs lookahead
+nature.  As a result, SPP, and hence PPF, underperform on this
+benchmark.
 
-At its peak, PPF manages a speedup of a factor of over \textbf{3.39x} on one of the
+\pg{is the following text old? doesn't look right:} At its peak, PPF
+manages a speedup of a factor of over \textbf{3.39x} on one of the
 {\tt 602.gcc\_s} traces. This also corresponds to speedup gain of
-\textbf{60\%} over the next best prefetcher for that trace -- BOP. In general, benchmarks
-{\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt {623.xalancbmk}\_s}~~and~~{\tt
-649.fotonik3d\_s}  benefit the most from PPF, with the speedup over SPP ranging
-from \textbf{10\% to 25\%}.
+\textbf{60\%} over the next best prefetcher for that trace -- BOP. In
+general, benchmarks {\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt
+  {623.xalancbmk}\_s}~~and~~{\tt 649.fotonik3d\_s} benefit the most
+from PPF, with the speedup over SPP ranging from \textbf{10\% to
+  25\%}.
 
@@
 % [EB] Resolved
-On the full SPEC CPU 2017 suite, PPF improves the geometric mean IPC of 
-the baseline by \textbf{15.24\%}, which is \textbf{2.27\%} better 
-than the next best prefetcher -- SPP.\\ \\
-%
+On the full SPEC CPU 2017 suite, PPF improves the geometric mean IPC
+of the baseline by \textbf{15.24\%}, which is \textbf{2.27\%} better
+than the next best prefetcher -- SPP.
+
 \begin{figure}[ht]
@@
 % djimenez: this is a weird way to make a header for a section. use \paragraph or \subsection
-\noindent \textbf{Coverage:}
-Prefetcher coverage is defined as the ratio of the number of misses avoided
-through prefetching to the number of misses with no prefetching.
 
-Figure~\ref{Fig:Coverage} shows the fraction of misses in the L2 and LLC
-avoided by the various prefetchers. PPF has the highest coverage of all the
-prefetchers simulated. On the SPEC CPU 2017 benchmarks, PPF reduces misses by
-\textbf{75.5\%} and \textbf{86.9\%} in the L2 and LLC respectively. For the
-same benchmark, the next best prefetcher, DA-AMPM, covers \textbf{54.3\%} and
-\textbf{78.5\%} of the misses respectively.
+\noindent \textbf{Coverage:} Prefetcher coverage is defined as the
+ratio of the number of misses avoided through prefetching over the
+number of misses with no prefetching.  Figure~\ref{Fig:Coverage} shows
+the fraction of misses in the L2 and LLC avoided by the various
+prefetchers.  PPF has the highest coverage of all the prefetchers
+simulated.  On the SPEC CPU 2017 benchmarks, PPF reduces misses by
+\textbf{75.5\%} and \textbf{86.9\%} in the L2 and LLC
+respectively. For the same benchmarks, the next best prefetcher,
+DA-AMPM, covers \textbf{54.3\%} and \textbf{78.5\%} of the misses
+respectively.
 
-This superior coverage of PPF can be attributed to aggressive re-tuning of the
-underlying SPP, enabled by the Perceptron Filter making sure the high coverage
-does not lead to increased cache pollution.
+This superior coverage of PPF can be attributed to aggressive
+re-tuning of the underlying SPP, enabled by the Perceptron Filter
+making sure the high coverage does not lead to increased cache
+pollution.
 
@@
 \label{Results-Multi}
-In this section, we demonstrate the improvement achieved by PPF for a mix of
-multi-programmed workloads.\\ \\
-%
-\noindent \textbf{4-Core Environment:}
-Figure~\ref{Fig:4Core_SPEC2017} shows a comparison of speedups obtained on
-4-core mixes of a memory intensive subset SPEC CPU 2017. We plot all 4
-prefetchers, normalized to the baseline. The workloads have been sorted in
-increasing order of the speedup. PPF offers a speedup of \textbf{51.2\%} on
-these traces, an improvement of \textbf{11.4\%} over the underlying SPP,
+In this section, we demonstrate the improvement achieved by PPF for a
+mix of multi-programmed workloads.
+
+\noindent \textbf{4-Core:} Figure~\ref{Fig:4Core_SPEC2017}
+shows a comparison of speedups obtained on 4-core mixes of a memory
+intensive subset SPEC CPU 2017. We plot all 4 prefetchers, normalized
+to the baseline. The workloads have been sorted in increasing order of
+the speedup. PPF offers a speedup of \textbf{51.2\%} on these traces,
+an improvement of \textbf{11.4\%} over the underlying SPP,
 \textbf{9.7\%} over the next DA-AMPM, and \textbf{16.9\%} over BOP.
-On a different set of fully random SPEC CPU 2017 4-core mixes (not illustrated
-for space reasons), PPF provides an IPC speedup of \textbf{26.07}\% over the
-baseline, which is an improvement of \textbf{5.6}\% over SPP.\\ \\
-%
+On a different set of fully random SPEC CPU 2017 4-core mixes (not
+illustrated for space reasons), PPF provides an IPC speedup of
+\textbf{26.07}\% over the
+baseline, which is an improvement of \textbf{5.6}\% over SPP.
+
 \begin{figure}[ht]
@@
 \end{figure}
-%
-\noindent \textbf{8-Core Environment:}
-The sorted comparison of speedups on the memory intensive 8-core mixes is
-shown in Figure~\ref{Fig:8Core_SPEC2017}. PPF improves baseline performance
-by \textbf{37.6\%}, an improvement of \textbf{9.7\%} over SPP. For a random
-set of SPEC CPU 2017 mixes (not illustrated for space reasons), PPF improves
-performance by \textbf{23.4\%} over the baseline, corresponding to
-\textbf{4.6\%} over SPP. This increased improvement achieved by PPF over the
-base engine SPP in a multi-core environment is expected as PPF has a very
-accurate filter, it eliminates useless prefetches before they can cause
-pollution in the shared LLC.
-BOP offers a better improvement than SPP for the memory intensive mixes. This
-superiority can be attributed to BOP's inherent aggressive nature. DA-AMPM is
-also ahead of SPP in both the mixes. Interestingly, in all these cases, PPF
-consistently outperforms the best performing prefetcher.
+
+\noindent \textbf{8-Core:} The sorted comparison of speedups on the
+memory intensive 8-core mixes is shown in
+Figure~\ref{Fig:8Core_SPEC2017}.  PPF improves baseline performance by
+\textbf{37.6\%}, an improvement of \textbf{9.7\%} over SPP. For a
+random set of SPEC CPU 2017 mixes (not illustrated for space reasons),
+PPF improves performance by \textbf{23.4\%} over the baseline,
+corresponding to \textbf{4.6\%} over SPP. This increased improvement
+achieved by PPF over the base engine SPP in a multi-core environment
+is expected as PPF is a very accurate filter. Thus, it eliminates
+useless prefetches before they can cause pollution in the shared LLC.
+BOP offers a better improvement than SPP for the memory intensive
+mixes. This superiority can be attributed to BOP's inherent aggressive
+nature. DA-AMPM is also ahead of SPP in both the mixes. Interestingly,
+in all these cases, PPF consistently outperforms the best performing
+prefetcher.
 
@@
 
-\subsection{Additional Memory Constraints}
-\label{Results-AdditionalMem}
+% \subsection{Additional Memory Constraints}
+% \label{Results-AdditionalMem}
 
-%Figure~\ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a
-%followed by the overall performance on the complete benchmark.  
-%For the sake of brevity, we have only shown selected traces
+% %Figure~\ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a
+% %followed by the overall performance on the complete benchmark.  
+% %For the sake of brevity, we have only shown selected traces
 
-We also model PPF with reduced LLC and with low bandwidth constraints,
-respectively (not illustrated for space reasons). Benchmark {\tt 605.mcf\_s}
-in low bandwidth conditions is prefetch averse. In general, any prefetcher
-yields a negative speedup on that trace. On {\tt 654.roms\_s} and {\tt
-607.cactuBSSN\_s}, PPF is unable to match the performance achieved by the best
-prefetcher. On the other hand, PPF outperforms all the other prefetchers on
-{\tt 623.xalancbmk\_s} and {\tt 638.imagick\_s} benchmarks. Overall, PPF
-gives a better improvement under small LLC condition and matches the best
-prefetcher, BOP, under low DRAM bandwidth conditions.
+% We also model PPF with reduced LLC and with low bandwidth constraints,
+% respectively (not illustrated for space reasons).  % Benchmark {\tt
+% % 605.mcf\_s} in low bandwidth conditions is prefetch averse. In
+% % general, any prefetcher yields a negative speedup on that trace. On
+% % {\tt 654.roms\_s} and {\tt 607.cactuBSSN\_s}, PPF is unable to match
+% % the performance achieved by the best prefetcher. On the other hand,
+% % PPF outperforms all the other prefetchers on {\tt 623.xalancbmk\_s}
+% % and {\tt 638.imagick\_s} benchmarks.
+% Overall, PPF provides a greater improvement under small LLC condition
+% and matches the best prefetcher, BOP, under low DRAM bandwidth
+% conditions.
 
@@
 
-Figure~\ref{Fig:CrossVal}(a) shows the performance benefit comparison of all
-the prefetch schemes on 4 different applications in the CloudSuite benchmark.
-In general, these applications are prefetch agnostic. Even so, PPF manages a
-\textbf{3.78\%} improvement over no prefetching, putting it ahead of the next
-best prefetcher, SPP, which provides a 3.08\% speedup.
+Figure~\ref{Fig:CrossVal}(a) shows the performance benefit comparison
+of all the prefetch schemes on 4 different applications in the
+CloudSuite benchmark.  In general, these applications are prefetch
+agnostic.  Even so, PPF manages a \textbf{3.78\%} improvement over no
+prefetching, putting it ahead of the next best prefetcher, SPP, which
+provides a 3.08\% speedup.
 
-Figure~\ref{Fig:CrossVal}(b) shows the speed-up achieved 
-on the memory intensive subset and the full SPEC CPU 2006 suite for a
-single-processor machine. 
-PPF provides a speedup of \textbf{36.3\%} over the baseline on the 
-memory intensive subset of SPEC CPU
-2006 benchmark, giving an improvement of \textbf{6.1\%} over SPP and
-\textbf{8.44\%} over DA-AMPM and \textbf{9.93} over BOP. On the whole of 
-the SPEC CPU 2006 suite, the speedup is \textbf{19.6\%}, an improvement 
-of \textbf{3.33\%} over SPP.
+Figure~\ref{Fig:CrossVal}(b) shows the speed-up achieved on the memory
+intensive subset and the full SPEC CPU 2006 suite for a
+single-processor machine.  PPF provides a speedup of \textbf{36.3\%}
+over the baseline on the memory intensive subset of SPEC CPU 2006
+benchmark, giving an improvement of \textbf{6.1\%} over SPP and
+\textbf{8.44\%} over DA-AMPM and \textbf{9.93} over BOP. On the whole
+of the SPEC CPU 2006 suite, the speedup is \textbf{19.6\%}, an
+improvement of \textbf{3.33\%} over SPP.
 
 For 4-core memory intensive mixes, PPF improves the baseline by
-\textbf{59.1\%}, \textbf{8.6\%} ahead of SPP. For 8-core memory intensive
-mixes, the speedup over the baseline is \textbf{47.8\%}, \textbf{11.3\%} ahead
-of SPP.
+\textbf{59.1\%}, \textbf{8.6\%} ahead of SPP. For 8-core memory
+intensive mixes, the speedup over the baseline is \textbf{47.8\%},
+\textbf{11.3\%} ahead of SPP.
 
-We developed PPF to yield good performance on the SPEC CPU 2017 benchmarks.
-Nevertheless, the performance is consistently good on other benchmark suites.
-We attribute this fact to the inherent adaptability of the perceptron model.
-In general, perceptron weights are able to adjust in real-time so as to find
-the best possible correlation between the output and the given set of
-features.
+We developed PPF to yield good performance on the SPEC CPU 2017
+benchmarks.  Nevertheless, the performance is consistently good on
+other benchmark suites.  We attribute this fact to the inherent
+adaptability of the perceptron model.  In general, perceptron weights
+are able to adjust in real-time so as to find the best possible
+correlation between the output and the given set of features.

### 2018-12-07 method.tex
adding method.tex

@@
 
-\noindent \textbf{Global Pearson's Correlation} Here we examine the
+\noindent \textbf{Global Pearson's Correlation:} Here we examine the
 perceptron weights at the end of all trace execution by which time the
@@
 
-\noindent \textbf{Per Trace Correlation} Another important way to look
+\noindent \textbf{Per Trace Correlation:} Another important way to look
 at the perceptron features is to see how much their contribution
@@
 
-\noindent \textbf{Trimming Features Using Cross-correlation} Beside
+\noindent \textbf{Trimming Features Using Cross-correlation:} Beside
 providing interesting insights into prefetching behavior, P-value can

### 2018-12-07 back.tex

@@
 patterns quickly.  While the basic idea of perceptron based prefetch
-filtering is applicable to any look-ahead prefetcher, we develop a
+filtering is applicable to any lookahead prefetcher, we develop a
 practical implementation of our proposed perceptron prefetch filter

### 2018-12-07 intro.tex

@@
 \includegraphics[width=\columnwidth]{Motivation}
-\caption{Impact of aggressive prefetching on performance. X-axis
-  represents increasingly aggressive prefetch lookahead depth. All the
-  metrics have been normalized to the least aggressive
-  configuration. As can be seen, total number of prefetches increase
-  at a higher rate than the number of good prefetches, leading to IPC
-  reduction.}
+\caption{The impact of aggressive prefetching on performance for {\tt 603.bwaves\_s}. 
+The number of useful prefetches increases with aggressiveness
+slower than total prefetches, which wastes bandwidth and 
+harms performance.}
 \label{Fig:Motivation}
@@
 much cache capacity and bandwidth may be wasted prefetching unused
-data that performance can ultimately be harmed by this strategy. On
+data that performance can ultimately be harmed by this strategy. At
 the other extreme, another prefetcher may be overly conservative and
@@
 prefetches are accurate.  In the figure, we iteratively re-tuned 
-this threshold and allowed the prefetcher to lookahead an of 
-arbitrary depth from 7 to 15. The figure depicts the behavior 
+this threshold to allow the prefetcher to lookahead a fixed
+depth from 7 to 15. The figure depicts the behavior 
 of the {\tt 603.bwaves\_s} SPEC CPU 2017 benchmark. The IPC, the 
-total number of prefetches suggested by the prefetcher (TOTAL\_PF) 
+total number of prefetches issued by the prefetcher (TOTAL\_PF), 
 and the actual useful predictions (GOOD\_PF), all have been normalized 
-to the least aggressive version of the prefetcher. As the look-ahead
+to lookahead depth 7. As the lookahead
 depth increases, so do useful prefetches, and hence coverage. This
-coverage, however, comes at a cost of total prefetches increasing at
+coverage, however, comes at the cost of total prefetches increasing at
 an even higher rate. This leads to cache pollution and bandwidth
-contention, as is clearly visible with the reduction in IPC.  
+contention, and leads to a reduction in IPC.  
 %Similar behavior was observed in 12 other SPEC CPU 2017 benchmarks.

### 2018-12-07 results.tex
putting it back

@@
 
-% \subsection{Additional Memory Constraints}
-% \label{Results-AdditionalMem}
+\subsection{Additional Memory Constraints}
+\label{Results-AdditionalMem}
 
-% %Figure~\ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a
-% %followed by the overall performance on the complete benchmark.  
-% %For the sake of brevity, we have only shown selected traces
+%Figure~\ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a
+%followed by the overall performance on the complete benchmark.  
+%For the sake of brevity, we have only shown selected traces
 
-% We also model PPF with reduced LLC and with low bandwidth constraints,
-% respectively (not illustrated for space reasons).  % Benchmark {\tt
-% % 605.mcf\_s} in low bandwidth conditions is prefetch averse. In
-% % general, any prefetcher yields a negative speedup on that trace. On
-% % {\tt 654.roms\_s} and {\tt 607.cactuBSSN\_s}, PPF is unable to match
-% % the performance achieved by the best prefetcher. On the other hand,
-% % PPF outperforms all the other prefetchers on {\tt 623.xalancbmk\_s}
-% % and {\tt 638.imagick\_s} benchmarks.
-% Overall, PPF provides a greater improvement under small LLC condition
-% and matches the best prefetcher, BOP, under low DRAM bandwidth
-% conditions.
+We also model PPF with reduced LLC and with low bandwidth constraints,
+respectively (not illustrated for space reasons).  Benchmark {\tt
+  605.mcf\_s} in low bandwidth conditions is prefetch averse. In
+general, any prefetcher yields a negative speedup on that trace. On
+{\tt 654.roms\_s} and {\tt 607.cactuBSSN\_s}, PPF is unable to match
+the performance achieved by the best prefetcher. On the other hand,
+PPF outperforms all the other prefetchers on {\tt 623.xalancbmk\_s}
+and {\tt 638.imagick\_s} benchmarks.  Overall, PPF provides a greater
+improvement under small LLC condition and matches the best prefetcher,
+BOP, under low DRAM bandwidth conditions.
 

### 2018-12-07 related.tex
minor stuff

@@
 
-In addition to branch prediction~\cite{PerceptronPredictor}, perceptron-based
-learning has been applied to the area of cache management.
-Teran \textit{et al.} propose using perceptrons to predict cache line resuse, 
-bypass, and replacement~\cite{Perc_Reuse}. Perceptron Learning trains weights
-selected by hashes of multiple features, including the PC of the memory
-access instruction, some other recent PCs, and two different shifts of the tag of
-the referenced block. These features are used to index into weight tables, and
-the weights are then thresholded to generate a prediction. When a block from one
-of a few sampled sets~\cite{sdbp} is reused or evicted, the corresponding weights
-are decremented or incremented, according to the perceptron
-learning rule. Multiperspective Reuse Prediction~\cite{Multiperspective}
-improves on Perceptron Learning by contributing many new features.
+In addition to branch prediction~\cite{PerceptronPredictor},
+perceptron-based learning has been applied to the area of cache
+management.  Teran \textit{et al.} propose using perceptrons to
+predict cache line resuse, bypass, and replacement~\cite{Perc_Reuse}.
+Perceptron Learning trains weights selected by hashes of multiple
+features, including the PC of the memory access instruction, some
+other recent PCs, and two different shifts of the tag of the
+referenced block. These features are used to index into weight tables,
+and the weights are then thresholded to generate a prediction. When a
+block from one of a few sampled sets~\cite{sdbp} is reused or evicted,
+the corresponding weights are decremented or incremented, according to
+the perceptron learning rule. Multiperspective Reuse
+Prediction~\cite{Multiperspective} improves on Perceptron Learning by
+contributing many new features.
 
@@
 
-%Two widely used prefetchers are Next-$n$ Lines~\cite{nextn} and
-%Stride~\cite{stride} prefetchers. Both capture regular memory access patterns
-%with low overhead. The Next-$n$ Lines prefetcher queues prefetches for the
-%next $n$ blocks after any given miss, expecting that the principle of spatial
-%locality will hold and those cache blocks after a missed block are likely to
-%be used in the future. The stride prefetcher identifies strided reference
-%patterns in programs based on past behavior of missing loads.  When a load
-%misses, cache blocks ahead of that miss are fetched in the pattern following
-%the previous behavior in the hope of avoiding future misses.  Without further
-%knowledge about temporal locality and application characteristics, these
-%prefetchers cannot do more than detecting and prefetching regular memory
-%access patterns with limited spatial locality.
+% Two widely used prefetchers are Next-$n$ Lines~\cite{nextn} and
+% Stride~\cite{stride} prefetchers. Both capture regular memory access
+% patterns with low overhead. The Next-$n$ Lines prefetcher queues
+% prefetches for the next $n$ blocks after any given miss, expecting
+% that the principle of spatial locality will hold and those cache
+% blocks after a missed block are likely to be used in the future. The
+% stride prefetcher identifies strided reference patterns in programs
+% based on past behavior of missing loads.  When a load misses, cache
+% blocks ahead of that miss are fetched in the pattern following the
+% previous behavior in the hope of avoiding future misses.  Without
+% further knowledge about temporal locality and application
+% characteristics, these prefetchers cannot do more than detecting and
+% prefetching regular memory access patterns with limited spatial
+% locality.
 
-Spatial prefetchers include such well-understood examples as the next-line (or
-next-$n$-line) prefetcher, and the stream prefetcher, and are distinguished by prefetching
-data without regard for the order in which the data will be accessed.
-In addition to simpler these, Somogyi \textit{et al.} 
-propose Spatial Memory Streaming (SMS)~\cite{SMS}. SMS works by learning
-the spatial footprint of all data used by a program within a region of memory,
-and when that region is revisited by the program, the same spatial footprint
-is prefetched. Ishii \textit{et al.} propose the Access Map Pattern Matching 
-prefetcher (AMPM)~\cite{AMPM}, which creates a map of all accessed lines within a region
-of memory, and then analyzes this map on every access to see if any 
-fixed-stride pattern can be identified and prefetched that is centered on 
-the current access. DRAM-Aware AMPM (DA-AMPM)~\cite{DA_AMPM}
-is a variant of AMPM that delays some prefetches so they can be issued 
-together with others in the same DRAM row, increasing bandwidth utilization.
-Pugsley \textit{et al.} propose the Sandbox Prefetcher~\cite{Sandbox},
-which analyzes candidate fixed-offset prefetchers in a sandboxed environment
-to determine which is most suitable for the current program phase. Michaud
-proposes the Best-Offset Prefetcher~\cite{BOP}, which determines the optimal
-offset to prefetch while considering memory latency and prefetch timeliness.
+Spatial prefetchers include such well-understood examples as the
+next-line (or next-$n$-line) prefetcher, and the stream prefetcher,
+and are distinguished by prefetching data without regard for the order
+in which the data will be accessed.  In addition to these simpler
+examples, Somogyi \textit{et al.}  propose Spatial Memory Streaming
+(SMS)~\cite{SMS}.  SMS works by learning the spatial footprint of all
+data used by a program within a region of memory around a given
+missing load, and when the load that causes an new miss elsewhere, the
+same spatial footprint is prefetched.  Ishii \textit{et al.} propose
+the Access Map Pattern Matching prefetcher (AMPM)~\cite{AMPM}, which
+creates a map of all accessed lines within a region of memory, and
+then analyzes this map on every access to see if any fixed-stride
+pattern can be identified and prefetched that is centered on the
+current access. DRAM-Aware AMPM (DA-AMPM)~\cite{DA_AMPM} is a variant
+of AMPM that delays some prefetches so they can be issued together
+with others in the same DRAM row, increasing bandwidth utilization.
+Pugsley \textit{et al.}  propose the Sandbox
+Prefetcher~\cite{Sandbox}, which analyzes candidate fixed-offset
+prefetchers in a sandboxed environment to determine which is most
+suitable for the current program phase.  Michaud proposes the
+Best-Offset Prefetcher~\cite{BOP}, which determines the optimal offset
+to prefetch while considering memory latency and prefetch timeliness.
 
@@
 
-Unlike spatial prefetchers, lookahead prefetchers take program order into
-account when they make predictions. Shevgoor \textit{et al.} propose the Variable 
-Length Delta Prefetcher (VLDP)~\cite{VLDP}, which correlates histories of
-deltas between cache line accesses within memory pages with the next delta
-within that page. SPP~\cite{SPP} and KPC's prefetching component~\cite{KPC} are
-more recent examples of lookahead prefetchers. They try to predict not only what
-data will be used in the future, but also the precise order in which the data
-will be used, within a given page. Predictions made by lookahead prefetchers can
-be fed back into their prediction mechanisms to predict even further down 
-a speculative path of memory acesses. These prefetchers can also generalize their
-learned patterns from one page, and use those patterns to make predictions in 
-other pages.
+Unlike spatial prefetchers, lookahead prefetchers take program order
+into account when they make predictions.  Shevgoor \textit{et al.}
+propose the Variable Length Delta Prefetcher (VLDP)~\cite{VLDP}, which
+correlates histories of deltas between cache line accesses within
+memory pages with the next delta within that page. SPP~\cite{SPP} and
+KPC's prefetching component~\cite{KPC} are more recent examples of
+lookahead prefetchers. They try to predict not only what data will be
+used in the future, but also the precise order in which the data will
+be used, within a given page. Predictions made by lookahead
+prefetchers can be fed back into their prediction mechanisms to
+predict even further down a speculative path of memory acesses. These
+prefetchers can also generalize their learned patterns from one page,
+and use those patterns to make predictions in other pages.
 
@@
 
-A low-accuracy aggressive prefetcher can significantly harm performance. To
-minimize interference from prefetching, Wu \textit{et al.} propose
-PACMan~\cite{pacman}, a prefetch-aware cache management policy. PACMan
-dedicates some LLC sets to each of three competing policies that treat demand
-and prefetch requests differently, using the policy in the rest of the cache that shows the fewest
-misses. Seshadri \textit{et al.} propose ICP~\cite{icp}, which
-demotes a prefetch to the lowest reuse priority on a demand hit, based on the
-observation that most prefetches are dead after their first hit. To address
-prefetcher-caused cache pollution, it also uses a variation of EAF~\cite{eaf}
-to track prefetching accuracy, and inserts only accurate prefetches to the
-higher priority position in the LRU stack. Jain \textit{et al.} propose
-Harmony~\cite{Harmony} to accomodate prefetches in their MIN 
-algorithm-inspired Hawkeye cache management system.
+A low-accuracy aggressive prefetcher can significantly harm
+performance.  To minimize interference from prefetching, Wu \textit{et
+  al.} propose PACMan~\cite{pacman}, a prefetch-aware cache management
+policy. PACMan dedicates some LLC sets to each of three competing
+policies that treat demand and prefetch requests differently, using
+the policy in the rest of the cache that shows the fewest
+misses. Seshadri \textit{et al.} propose ICP~\cite{icp}, which demotes
+a prefetch to the lowest reuse priority on a demand hit, based on the
+observation that most prefetches are dead after their first hit. To
+address prefetcher-caused cache pollution, it also uses a variation of
+EAF~\cite{eaf} to track prefetching accuracy, and inserts only
+accurate prefetches to the higher priority position in the LRU
+stack. Jain \textit{et al.} propose Harmony~\cite{Harmony} to
+accomodate prefetches in their MIN algorithm-inspired Hawkeye cache
+management system.
 
@@
 
-Peled \textit{et al.} introduce interesting ideas for on-line Reinforcement
-Learning and dynamically scaling the magnitude of feedback given to the
-baseline prefetcher~\cite{Semantics}. The prefetcher relies on compiler
-support for getting features to build the context. Liao \textit{et al.}
-focuse on prefetching for data center applications~\cite{Datacenter}. They
-use offline machine learning algorithms such as SVMs and logistic regression
-to do a parametric search for an optimal prefetcher configuration.
+Peled \textit{et al.} introduce interesting ideas for on-line
+Reinforcement Learning and dynamically scaling the magnitude of
+feedback given to the baseline prefetcher~\cite{Semantics}. The
+prefetcher relies on compiler support to receive features and build
+the context.  Liao \textit{et al.}  focus on prefetching for data
+center applications~\cite{Datacenter}.  They use offline machine
+learning algorithms such as SVMs and logistic regression to do a
+parametric search for an optimal prefetcher configuration.

### 2018-12-07 conc.tex
wording

--- 2018-12-05 conc.tex
@@
 \label{Conclusion}
-In this paper, we introduce perceptron-based learning for prefetching.  The
-perceptron filter acts independently as a check on the underlying prefetch
-engine.  PPF improves performance over the base prefetcher by 11.9\%.  We
-believe that this is a robust and adaptable technique that can be used to
-enhance any existing and upcoming prefetchers.
+In this paper, we introduce the Perceptron-Based Prefetch Filtering
+(PPF).  PPF acts as an independent check on the quality of predictions
+made by the underlying prefetch engine.  We also created a case study
+implementation of PPF using SPP as the baseline prefetch engine, while
+in principle other prefetchers could be used.  We show that PPF
+effectively filters bad prefetches, such that the given underlying
+prefetcher can be highly aggressively tuned to achieve increasing
+coverage.  PPF improves performance over the base prefetcher by up to
+11.4\%.  PPF is a robust and adaptable technique that can be used to
+enhance any existing prefetcher and can be a valuable tool in the
+design of future memory latency constrained systems.

### 2018-12-07 results.tex
small work in results

@@
 baseline.  This is equivalent to \textbf{4.63\%} over DA-AMPM,
-\textbf{4.61\%} over BOP and \textbf{3.78\%} over SPP.  Out of the 95
-simpoints developed for SPEC CPU 2017, PPF nearly matches or
-outperforms most of the prefetchers on 91 simpoints.  One interesting
+\textbf{4.61\%} over BOP and \textbf{3.78\%} over SPP.  Out of the 20
+SPEC CPU 2017 applications, PPF nearly matches or
+outperforms all the other prefetchers on 19 applications.  
+Benchmarks {\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt{623.xalancbmk}\_s} 
+and {\tt 649. fotonik3d\_s} benefit the most from PPF, with the 
+speedup over SPP ranging from \textbf{10\% to 25\%}.
+
+One interesting
 case here is {\tt 623.xalancbmk}.  Despite SPP under performing on that
@@
 
-\pg{is the following text old? doesn't look right:} At its peak, PPF
-manages a speedup of a factor of over \textbf{3.39x} on one of the
-{\tt 602.gcc\_s} traces. This also corresponds to speedup gain of
-\textbf{60\%} over the next best prefetcher for that trace -- BOP. In
-general, benchmarks {\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt
-  {623.xalancbmk}\_s}~~and~~{\tt 649.fotonik3d\_s} benefit the most
-from PPF, with the speedup over SPP ranging from \textbf{10\% to
-  25\%}.
+% \pg{is the following text old? doesn't look right:} At its peak, PPF
+% manages a speedup of a factor of over \textbf{3.39x} on one of the
+% {\tt 602.gcc\_s} traces. This also corresponds to speedup gain of
+% \textbf{60\%} over the next best prefetcher for that trace -- BOP.
 
@@
 Figure~\ref{Fig:8Core_SPEC2017}.  PPF improves baseline performance by
-\textbf{37.6\%}, an improvement of \textbf{9.7\%} over SPP. For a
+\textbf{37.6\%}, an improvement of \textbf{9.65\%} over SPP. For a
 random set of SPEC CPU 2017 mixes (not illustrated for space reasons),

### 2018-12-08 results.tex
FINAL(?)

@@
 shows a comparison of speedups obtained on 4-core mixes of a memory
-intensive subset SPEC CPU 2017. We plot all 4 prefetchers, normalized
+intensive subset of SPEC CPU 2017. We plot all 4 prefetchers, normalized
 to the baseline. The workloads have been sorted in increasing order of
@@
 benchmark, giving an improvement of \textbf{6.1\%} over SPP and
-\textbf{8.44\%} over DA-AMPM and \textbf{9.93} over BOP. On the whole
+\textbf{8.44\%} over DA-AMPM and \textbf{9.93\%} over BOP. On the whole
 of the SPEC CPU 2006 suite, the speedup is \textbf{19.6\%}, an

### 2018-12-08 abs.tex
minor
Typo. Had corrected in the submitted paper

@@
   Hardware prefetching is an effective technique for hiding cache miss
-latencies in modern processor designs. Prefetcher performance can
-be characterized by two main metrics that are generally at odds with
+  latencies in modern processor designs. Prefetcher performance can be
+  characterized by two main metrics that are generally at odds with
   one another: coverage, the fraction of baseline cache misses which
@@
   In this paper, we introduce Perceptron-based Prefetch Filtering
-(PPF) as a way increase the coverage of the prefetches generated by
+  (PPF) as a way to increase the coverage of the prefetches generated by
   a baseline prefetcher without negatively impacting accuracy. PPF
@@
   for a single-core configuration, and by 11.4\% for a 4-core
-configuration, compared to the baseline prefetcher alone. We also
-show that performance continues to scale as the number of cores
-sharing a last level cache increases.
+  configuration, compared to the baseline prefetcher alone.
 

### 2019-03-31 abs.tex
Rebuttal Changes

@@
   (PPF) as a way to increase the coverage of the prefetches generated by
-  a baseline prefetcher without negatively impacting accuracy. PPF
-  enables more aggressive tuning of a given baseline prefetcher,
+  {\color{red}an underlying} prefetcher without negatively impacting accuracy. PPF
+  enables more aggressive tuning of {\color{red}the underlying prefetcher},
   leading to increased coverage by filtering out the growing numbers
@@
   for a single-core configuration, and by 11.4\% for a 4-core
-  configuration, compared to the baseline prefetcher alone.
+  configuration, compared to the \color{red}underlying prefetcher alone.
 

### 2019-03-31 back.tex
Rebuttal Changes

@@
 
-\subsection{Baseline Prefetcher: SPP}
+\subsection{{\color{red}Underlying Prefetcher: SPP}}
 \label{Background-SPP}
@@
 practical implementation of our proposed perceptron prefetch filter
-using SPP as our baseline.  Here we describe the basic architecture of
-SPP.
+using SPP as our {\color{red}underlying mechanism}. 
+Here we describe the basic architecture of SPP.
 

### 2019-03-31 conc.tex
Rebuttal Changes

@@
 made by the underlying prefetch engine.  We also created a case study
-implementation of PPF using SPP as the baseline prefetch engine, while
+implementation of PPF using SPP as the {\color{red}underlying prefetcher}, while
 in principle other prefetchers could be used.  We show that PPF
@@
 prefetcher can be highly aggressively tuned to achieve increasing
-coverage.  PPF improves performance over the base prefetcher by up to
-11.4\%.  PPF is a robust and adaptable technique that can be used to
-enhance any existing prefetcher and can be a valuable tool in the
-design of future memory latency constrained systems.
+coverage.  PPF improves performance over the {\color{red}underlying} 
+prefetcher by up to 11.4\%.  PPF is a robust and adaptable technique 
+that can be used to enhance any existing prefetcher and can be a 
+valuable tool in the design of future memory latency constrained systems.

### 2019-03-31 design_arch.tex
Rebuttal Changes

@@
 design of our proposed perceptron prefetch filter (PPF).  PPF enhances
-a base prefetcher by filtering out predicted unused prefetches. PPF is
-a generalized prefetch filtering mechanism that may be adapted to any
-prefetcher with appropriate feature selection and modifications which
-we describe below.
+{\color{red}an underlying prefetcher} by filtering out predicted unused prefetches. 
+PPF is a generalized prefetch filtering mechanism that may be adapted 
+to any prefetcher with appropriate feature selection and modifications 
+which we describe below.
 
@@
 %EB: I see this as a short context of the datapath leading to the perc filter.
-The base prefetcher is triggered on every demand access to the L2
+The {\color{red}prefetcher} is triggered on every demand access to the L2
 Cache, as seen in Figure~\ref{fig:PPF_Hierarchy}.  At this point, 
@@
 according to the confidence of the prefetching mechanism.  Once the
-base prefetcher is triggered, the suggested prefetch candidates are
+{\color{red}underlying} prefetcher is triggered, the suggested prefetch candidates are
 fed to the perceptron filter to determine the usefulness of these
 prefetches. The filter ultimately decides whether to issue the
-prefetch suggestions of the base prefetcher. As shown in step 1
+prefetch suggestions of the {\color{red}underlying} prefetcher. As shown in step 1
 of Figure~\ref{fig:PPF_Datapath}(a), to make the decision,
@@
 structure that contains all metadata required to re-index the
-perceptron entries for training.
+perceptron entries for training. {\color{red}10 bits of the address are used to index 
+the tables, and another 6 bits to perform tag matching.}
 
@@
 As depicted in step 3, when there is an eviction or a demand access 
-to the L2, training for
-both the base prefetcher and our filter mechanism is triggered. The
-address of the cache block that triggered training is used to index
-both the Prefetch and Reject tables. 10 bits of the address are used
-to index tables, and another 6 bits to perform tag matching. Once the
-base prefetcher has suggested a prefetch request, the corresponding
-features are used to index the tables of weights.
+to the L2, training for both the {\color{red}underlying} prefetcher and our filter 
+mechanism is triggered. The address of the cache block that triggered 
+the training is used to index both the Prefetch and Reject tables. 
+{\color{red}If it is a match, the corresponding
+features are retrieved to index the tables of perceptron weights.}
 
@@
 hit means that the corresponding cache block was initially suggested
-by the base prefetcher, but wrongly rejected by the perceptron
+by the {\color{red}underlying} prefetcher, but wrongly rejected by the perceptron
 filter. The perceptron filter learns from this and makes use of the
@@
 The above discussion of PPF shows that it is highly modular and can be
-adapted to be used over any base prefetcher for increased prefetch
+adapted to be used over any {\color{red}underlying} prefetcher for increased prefetch
 accuracy. As a first step, all the prefetch
@@
 
-In general, PPF can be adapted to a new base prefetcher with only a
+In general, PPF can be adapted to a {\color{red}new prefetcher} with only a
 few modifications:
 
-\noindent \textbf{Making Base Prefetcher More Aggressive:} By Tuning 
+\noindent \textbf{Making {\color{red}Underlying} Prefetcher More Aggressive:} By Tuning 
 down any internal thresholds or throttling mechanisms to increase its
@@
 the features we developed use information derived directly from
-program execution, agnostic to the baseline prefetcher.  Beyond that,
+program execution, agnostic to the {\color{red}underlying} prefetcher.  Beyond that,
 the feature set can be expanded to convey any useful information or
-metadata available in the baseline prefetcher.
+metadata available in the {\color{red}underlying} prefetcher.
 
@@
 the relevant meta-data must be exported from the prefetcher to PPF. This way,
-PPF can be optimized to work tightly-knit with the base prefetcher.
+PPF can be optimized to work tightly-knit with the {\color{red}underlying} prefetcher.
 

### 2019-03-31 design_impl.tex
Rebuttal Changes

@@
 derived from the information available in the L2 Cache access stream
-or are taken as metadata derived from the baseline prefetcher.  Our
+or are taken as metadata derived from the {\color{red}underlying} prefetcher.  Our
 feature selection involved searching over a large space of relevant
@@
 specific prefetcher involves examining the available metadata in the
-prefetcher itself, and thus PPF is attuned to the baseline
+prefetcher itself, and thus PPF is attuned to the {\color{red}underlying}
 prefetcher's design. Using the statistical methodology outlined in

### 2019-03-31 intro.tex
Rebuttal Changes

@@
 a highly tuned feature of that prefetcher.  Because PPF is so much
-more effective at rejecting inaccurate prefetches than SPP's baseline
-mechanism, we are free to re-tune the rest of SPP's design around
-maximizing coverage. The result is an increase in both accuracy and
-coverage, and a notable increase in performance.
+more effective at rejecting inaccurate prefetches than SPP's 
+{\color{red}internal} mechanism, we are free to re-tune the rest of 
+SPP's design around maximizing coverage. The result is an increase 
+in both accuracy and coverage, and a notable increase in performance.
 

### 2019-03-31 method.tex
FINAL(?)
Rebuttal Changes

@@
 
-We initially came up a set of 23 features.  By studying cross
+We initially came up {\color{red}with} a set of 23 features.  By studying cross
 correlation of each of these features against others in a 23$\times$23
@@
 
+In terms of computations, the perceptron mechanism only introduces an
+extra adder tree. The hash perceptron mechanism makes sure than there
+is no actual vector multiplication happening in the hardware.
+Obtaining the perceptron sum requires addition of {\color{red}nine} 
+5-bit numbers.
+Using an adder tree of {\color{red}four} 5-bit adders, this can be done in
+ceil($log_{2}9$) = 4 steps. Perceptron update only requires weight
+update by +1 or -1. Thus, all the operations required for perceptron
+inferencing or updating the states of the perceptrons can be easily
+done in the time constraints of L2 Cache Accesses.
+
 \footnotetext[1]{Components of Prefetch Table can be found in Table
@@
   through.}
-
-In terms of computations, the perceptron mechanism only introduces an
-extra adder tree. The hash perceptron mechanism makes sure than there
-is no actual vector multiplication happening in the hardware.
-Obtaining the perceptron sum requires addition of 9 5-bit numbers.
-Using an adder tree of 4 5-bit adders, this can be done in
-ceil($log_{2}9$) = 4 steps. Perceptron update only requires weight
-update by +1 or -1. Thus, all the operations required for perceptron
-inferencing or updating the states of the perceptrons can be easily
-done in the time constraints of L2 Cache Accesses.

### 2019-03-31 related.tex
Rebuttal Changes

@@
 accomodate prefetches in their MIN algorithm-inspired Hawkeye cache
-management system.
+management system. {\color{red}Ebrahimi \textit{et al} introduce
+HPAC~\cite{HPAC} which provides a coordinated control between 
+multiple prefetchers present in A CMP by looking at the 
+prefetcher-induced inter-core interferance.}
 
@@
 parametric search for an optimal prefetcher configuration. 
+{\color{red}Hasheni \textit{et. al.}~\cite{LSTM} categorize prefetching as a 
+regression problem and use LSTM based Deep Learning approach.}

### 2019-03-31 results.tex
Rebuttal Changes

@@
 of the baseline by \textbf{15.24\%}, which is \textbf{2.27\%} better
-than the next best prefetcher -- SPP.
+than the next best prefetcher -- SPP. {\color{red}For PPF, the average 
+lookahead depth over the full benchmark is 3.97, while it is 3.28 for just 
+SPP. It is evident that on average for SPP, our scheme allows the 
+prefetcher to speculate 21\% deeper.}
 
@@
 corresponding to \textbf{4.6\%} over SPP. This increased improvement
-achieved by PPF over the base engine SPP in a multi-core environment
-is expected as PPF is a very accurate filter. Thus, it eliminates
-useless prefetches before they can cause pollution in the shared LLC.
-BOP offers a better improvement than SPP for the memory intensive
+achieved by PPF over the {\color{red}underlying prefetcher, SPP,} 
+in a multi-core environment is expected as PPF is a very accurate filter. 
+Thus, it eliminates useless prefetches before they can cause pollution in the 
+shared LLC. BOP offers a better improvement than SPP for the memory intensive
 mixes. This superiority can be attributed to BOP's inherent aggressive

### 2019-04-02 back.tex
made 'd' lower case

@@
 demand hit to the number of prefetches recommended in total. The range
-of $\alpha$ is [0,1].  $D$ is the lookahead depth. For $d = 1$, when
+of $\alpha$ is [0,1].  The lookahead depth is represented by $d. For $d = 1$, when
 SPP is in non-speculative mode, P\textsubscript{0} can be thought of
@@
 implementation, we believe that a better form of generalized on-line
-decision making was possible. Hence, it was necessary to build a
+decision making is possible. Hence, it was necessary to build a
 robust and adaptable learning mechanism to accept / reject the
-prefetch suggestions; and to decide the fill level (L2 Cache vs Last
+prefetch suggestions, and to decide the fill level (L2 Cache vs Last
 Level Cache). Thus, we introduce an independent on-line perceptron

### 2019-04-02 design_arch.tex
Reworded last two paragraphs.

@@
 the feature set can be expanded to convey any useful information or
-metadata available in the {\color{red}underlying} prefetcher.
+metadata available in the {\color{red}underlying} prefetcher itself.
 
@@
 internal counters specific to the underlying prefetcher can be suitable candidates
-for the perceptron features. To make sure that the perceptron layer sees that, 
-the relevant meta-data must be exported from the prefetcher to PPF. This way,
-PPF can be optimized to work tightly-knit with the {\color{red}underlying} prefetcher.
+for perceptron features. Therefore, all relevant meta-data must be exported from 
+the prefetcher to PPF. In this way, PPF can be optimized to work in a tight-knit fashion
+with the {\color{red}underlying} prefetcher.
 

### 2019-04-02 design_impl.tex
Added one more use of "underlying" to describe SPP
Added missing period and other small typo fixes.

@@
 range of features that are used to determine the usefulness of
-prefetches.  Here, we have selected SPP as the base engine prefetcher.
+prefetches.  Here, we have selected SPP as the underlying prefetcher.
 
@@
 
-\noindent \textbf{Exporting Features from SPP:} PPF uses the meta-data 
-specific to SPP, to build some of the perceptron features. These include
-the lookahead depth, signature and the confidence counter. These features 
-were made visible to PPF\\
+\noindent \textbf{Exporting Features from SPP:} PPF uses metadata 
+specific to SPP to build some of the perceptron features. These include
+the lookahead depth, signature, and the confidence counter. These features 
+were made visible to PPF.\\
 %

### 2019-04-17 design_arch.tex
small changes

--- 2019-03-31 design_arch.tex
@@
 structure that contains all metadata required to re-index the
-perceptron entries for training. {\color{red}10 bits of the address are used to index 
-the tables, and another 6 bits to perform tag matching.}
+perceptron entries for training. {\color{red}Ten bits of the address are 
+used to index into the tables, and another six bits are stored to 
+perform tag matching.}
 
@@
 As depicted in step 3, when there is an eviction or a demand access 
-to the L2, training for both the {\color{red}underlying} prefetcher and our filter 
-mechanism is triggered. The address of the cache block that triggered 
-the training is used to index both the Prefetch and Reject tables. 
-{\color{red}If it is a match, the corresponding
-features are retrieved to index the tables of perceptron weights.}
+to the L2, training for both the {\color{red}underlying} prefetcher 
+and our filter mechanism is triggered. The address of the cache block 
+that triggered the training is used to index both the Prefetch and 
+Reject tables. {\color{red}If it is a match, the corresponding
+features are retrieved to index into the tables of perceptron weights.}
 
@@
 
-\noindent \textbf{Making {\color{red}Underlying} Prefetcher More Aggressive:} By Tuning 
+\noindent \textbf{Making the {\color{red}Underlying} Prefetcher More Aggressive:} By Tuning 
 down any internal thresholds or throttling mechanisms to increase its
@@
 for the perceptron features. To make sure that the perceptron layer sees that, 
-the relevant meta-data must be exported from the prefetcher to PPF. This way,
+the relevant metadata must be exported from the prefetcher to PPF. This way,
 PPF can be optimized to work tightly-knit with the {\color{red}underlying} prefetcher.

### 2019-04-17 method.tex
small changes

@@
         Address 	& 24 & \\
-        Curr Signature 	& 10 & \multirow{2}{4.8cm}{Meta-data required for perceptron}\\
+        Curr Signature 	& 10 & \multirow{2}{4.8cm}{Metadata required for perceptron}\\
 	PC$_i$ Hash	& 12 & \multirow{2}{1.1cm}{training}\\
@@
     \end{adjustbox}
-    \caption{Meta-data Stored in Prefetch Table}
+    \caption{Metadata Stored in Prefetch Table}
     \label{tab:PTable_metadata}
@@
 implement PPF. The Prefetch Table was enhanced to accommodate storing
-of meta-data for perceptron training. Table \ref{tab:PTable_metadata}
-depicts the meta-data stored for each entry in the Prefetch Table.
+of metadata for perceptron training. Table \ref{tab:PTable_metadata}
+depicts the metadata stored for each entry in the Prefetch Table.
 Table \ref{tab:PPF_overhead} shows the total storage overhead of PPF
@@
     \hline
-        \multicolumn{4}{|c|}{\textbf{Total: 3,222,940 bits = 39.34 KB}}\\
+        \multicolumn{4}{|c|}{\textbf{Total: 322,240 bits = 39.34 KB}}\\
     \hline

### 2019-04-17 related.tex
small changes

--- 2018-12-07 related.tex
@@
 accomodate prefetches in their MIN algorithm-inspired Hawkeye cache
-management system.
+management system. {\color{red}Ebrahimi \textit{et al.} introduce
+HPAC~\cite{HPAC} which provides a coordinated control between 
+multiple prefetchers present in a CMP by looking at the 
+prefetcher-induced inter-core interferance.}
 
@@
 parametric search for an optimal prefetcher configuration. 
+{\color{red}Hasheni \textit{et al.}~\cite{LSTM} categorize 
+prefetching as a regression problem and use LSTM based Deep 
+Learning approach.}

### 2019-04-17 intro.tex
removed an extra period '.'

@@
 Figure~\ref{Fig:Motivation} illustrates the above scenario.  Here we
-consider a state-of-the-art. lookahead prefetcher -- SPP~\cite{SPP}.
+consider a state-of-the-art lookahead prefetcher -- SPP~\cite{SPP}.
 Lookahead prefetchers such as SPP provide a mechanism to speculate an
-arbitrary number of references ahead of the initial triggering miss.
+arbitrary number of references ahead of the initial triggering access.
 In SPP, a throttling confidence threshold is then used to ensure that

### 2019-04-17 design_impl.tex
small changes
merging

--- 2019-03-31 design_impl.tex
@@
 
-\noindent \textbf{Exporting Features from SPP:} PPF uses the meta-data 
+\noindent \textbf{Exporting Features from SPP:} PPF uses the metadata 
 specific to SPP, to build some of the perceptron features. These include

### 2019-04-17 abs.tex
Paragraph spacing

--- 2018-12-08 abs.tex
@@
   (PPF) as a way to increase the coverage of the prefetches generated by
-  a baseline prefetcher without negatively impacting accuracy. PPF
-  enables more aggressive tuning of a given baseline prefetcher,
+  an underlying prefetcher without negatively impacting accuracy. PPF
+  enables more aggressive tuning of the underlying prefetcher,
   leading to increased coverage by filtering out the growing numbers
@@
   for a single-core configuration, and by 11.4\% for a 4-core
-  configuration, compared to the baseline prefetcher alone.
+  configuration, compared to the underlying prefetcher alone.
 

### 2019-04-17 back.tex
Paragraph spacing

@@
 
-\subsection{{\color{red}Underlying Prefetcher: SPP}}
+\subsection{Underlying Prefetcher: SPP}
 \label{Background-SPP}
@@
 filtering is applicable to any lookahead prefetcher, we develop a
-practical implementation of our proposed perceptron prefetch filter
-using SPP as our {\color{red}underlying mechanism}. 
+practical implementation of our proposed prefetch filter
+using SPP as our underlying mechanism. 
 Here we describe the basic architecture of SPP.
-
+\newline
+\newline
 \noindent \textbf{Signature Table:} As shown on the left side of
-Figure~\ref{fig:spp_update}, the Signature Table keeps track of 256
+Figure~\ref{fig:spp_update}, the Signature Table (ST) keeps track of 256
 most recently accessed pages. It is meant to capture memory access
@@
 Figure~\ref{fig:spp_update}.
-
-\noindent \textbf{Pattern Table:} The Pattern Table, shown on the right
+\newline
+\newline
+\noindent \textbf{Pattern Table:} The Pattern Table (PT), shown on the right
 side in Figure~\ref{fig:spp_update} is indexed by the signature
@@
 the signature holds up to 4 unique delta predictions.
-
+%
 % djimenez: it does look like we're pushing it with all the stuff about SPP.
@@
 % details. (I'm not saying cut anything that's already there, just don't expand.)
-
+%
 %% NOTE: Can we do away with this example? 
@@
 %% 	is stored in the Pattern Table.
-
+%
 %  -- This looks like a detail that is confusing and not necessary to
@@
 % Table to avoid any confusion.
-
+\newline
+\newline
 \noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes
@@
 Figure~\ref{fig:spp_structure}, it repeats the cycle of accessing
-Pattern Table and updating the signature based on highest confidence
+the PT and updating the signature based on highest confidence
 prefetch from the last iteration.  The iteration counter on which SPP
@@
 overall confidence keeps decreasing.
-
+\newline
+\newline
 \noindent \textbf{Confidence Tracking}: As shown in
@@
 C\textsubscript{sig}.  When SPP enters into a lookahead mode, the path
-confidence P\textsubscript{d} is given as:
+confidence P\textsubscript{d} is:
 $$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;
@@
 demand hit to the number of prefetches recommended in total. The range
-of $\alpha$ is [0,1].  The lookahead depth is represented by $d. For $d = 1$, when
+of $\alpha$ is [0,1].  The lookahead depth is represented by $d$. 
+For $d = 1$, when
 SPP is in non-speculative mode, P\textsubscript{0} can be thought of

### 2019-04-17 conc.tex
Paragraph spacing

--- 2018-12-07 conc.tex
@@
 made by the underlying prefetch engine.  We also created a case study
-implementation of PPF using SPP as the baseline prefetch engine, while
+implementation of PPF using SPP as the underlying prefetcher, while
 in principle other prefetchers could be used.  We show that PPF
@@
 prefetcher can be highly aggressively tuned to achieve increasing
-coverage.  PPF improves performance over the base prefetcher by up to
-11.4\%.  PPF is a robust and adaptable technique that can be used to
-enhance any existing prefetcher and can be a valuable tool in the
-design of future memory latency constrained systems.
+coverage.  PPF improves performance over the underlying 
+prefetcher by up to 11.4\%.  PPF is a robust and adaptable technique 
+that can be used to enhance any existing prefetcher and can be a 
+valuable tool in the design of future memory latency constrained systems.

### 2019-04-17 design_arch.tex
merging
Paragraph spacing

@@
 design of our proposed perceptron prefetch filter (PPF).  PPF enhances
-{\color{red}an underlying prefetcher} by filtering out predicted unused prefetches. 
+{an underlying prefetcher} by filtering out predicted unused prefetches. 
 PPF is a generalized prefetch filtering mechanism that may be adapted 
@@
 be found in Section~\ref{Method-Overheads}.
-
+%
 %%\pg{the text below would be easier to follow if you used the figure
@@
 %%  talking about}
-
+%
+\newline
+\newline
 \noindent \textbf{Inferencing}\\
@@
 %EB: I see this as a short context of the datapath leading to the perc filter.
-The {\color{red}prefetcher} is triggered on every demand access to the L2
+The {prefetcher} is triggered on every demand access to the L2
 Cache, as seen in Figure~\ref{fig:PPF_Hierarchy}.  At this point, 
@@
 according to the confidence of the prefetching mechanism.  Once the
-{\color{red}underlying} prefetcher is triggered, the suggested prefetch candidates are
+{underlying} prefetcher is triggered, the suggested prefetch candidates are
 fed to the perceptron filter to determine the usefulness of these
 prefetches. The filter ultimately decides whether to issue the
-prefetch suggestions of the {\color{red}underlying} prefetcher. As shown in step 1
+prefetch suggestions of the {underlying} prefetcher. As shown in step 1
 of Figure~\ref{fig:PPF_Datapath}(a), to make the decision,
@@
 prefetches.
-
+\newline
+\newline
 \noindent \textbf{Recording}\\
@@
 structure that contains all metadata required to re-index the
-perceptron entries for training. {\color{red}Ten bits of the address are 
+perceptron entries for training. {Ten bits of the address are 
 used to index into the tables, and another six bits are stored to 
@@
 the observed demand accesses to the L2.
-
+\newline
+\newline
 \noindent \textbf{Feedback and Data Retrieval}\\
 As depicted in step 3, when there is an eviction or a demand access 
-to the L2, training for both the {\color{red}underlying} prefetcher 
+to the L2, training for both the {underlying} prefetcher 
 and our filter mechanism is triggered. The address of the cache block 
 that triggered the training is used to index both the Prefetch and 
-Reject tables. {\color{red}If it is a match, the corresponding
+Reject tables. {If it is a match, the corresponding
 features are retrieved to index into the tables of perceptron weights.}
-
+%
 % Which bits? are the tables accesses sequentially?
-
+%
+\newline
+\newline
 \noindent \textbf{Training}\newline As can be seen in step 4 of, 
@@
 hit means that the corresponding cache block was initially suggested
-by the {\color{red}underlying} prefetcher, but wrongly rejected by the perceptron
+by the {underlying} prefetcher, but wrongly rejected by the perceptron
 filter. The perceptron filter learns from this and makes use of the
@@
 The above discussion of PPF shows that it is highly modular and can be
-adapted to be used over any {\color{red}underlying} prefetcher for increased prefetch
+adapted to be used over any {underlying} prefetcher for increased prefetch
 accuracy. As a first step, all the prefetch
@@
 
-In general, PPF can be adapted to a {\color{red}new prefetcher} with only a
+\noindent In general, PPF can be adapted to a {new prefetcher} with only a
 few modifications:
-
-\noindent \textbf{Making the {\color{red}Underlying} Prefetcher More Aggressive:} By Tuning 
+\newline
+\newline
+\noindent \textbf{Making the {Underlying} Prefetcher More Aggressive:} By Tuning 
 down any internal thresholds or throttling mechanisms to increase its
 aggressiveness.
-
+\newline
+\newline
 \noindent \textbf{Inferencing and Storing:} All prefetch
@@
 indexing.
-
+\newline
+\newline
 \noindent \textbf{Retrieving and Training:} When feedback for a
@@
 weights.
-
+\newline
+\newline
 \noindent \textbf{Feature Selection:} Perceptrons essentially
@@
 the features we developed use information derived directly from
-program execution, agnostic to the {\color{red}underlying} prefetcher.  Beyond that,
+program execution, agnostic to the {underlying} prefetcher.  Beyond that,
 the feature set can be expanded to convey any useful information or
-metadata available in the {\color{red}underlying} prefetcher.
-
+metadata available in the {underlying} prefetcher itself.
+\newline
+\newline
 \noindent \textbf{Using Metadata from the Prefetcher:} Some of the 
@@
 the relevant metadata must be exported from the prefetcher to PPF. This way,
-PPF can be optimized to work tightly-knit with the {\color{red}underlying} prefetcher.
+PPF can be optimized to work tightly-knit with the {underlying} prefetcher.
 

### 2019-04-17 design_impl.tex
Paragraph spacing

@@
 \label{Impl-Changes}
-To modify the SPP design to suit our scheme, the following changes
+\noindent To modify the SPP design to suit our scheme, the following changes
 were made:
-
+\newline
+\newline
 \noindent \textbf{Exporting Features from SPP:} PPF uses the metadata 
@@
 the lookahead depth, signature and the confidence counter. These features 
-were made visible to PPF\\
+were made visible to PPF.
 %
+\newline
+\newline
 \noindent \textbf{Original Thresholds Discarded:} In PPF, the
@@
 derived from the information available in the L2 Cache access stream
-or are taken as metadata derived from the {\color{red}underlying} prefetcher.  Our
+or are taken as metadata derived from the {underlying} prefetcher.  Our
 feature selection involved searching over a large space of relevant
@@
 specific prefetcher involves examining the available metadata in the
-prefetcher itself, and thus PPF is attuned to the {\color{red}underlying}
+prefetcher itself, and thus PPF is attuned to the {underlying}
 prefetcher's design. Using the statistical methodology outlined in
@@
 % bullet from the previous section?
-
+%
+\newline
+\newline
 \noindent \textbf{Physical Address}: Here we use the lower bits 
@@
 %  talk about what you do when the number of bits is unwieldy.}
-
+%
+\newline
+\newline
 \noindent \textbf{Cache Line} and \textbf{Page Address}: These two
@@
 the address bits into half.
-
+%
 %
@@
 % [EB] Thank you!
+\newline
+\newline
 \noindent \textbf{Program Counter XOR Depth}: The PC is for the
@@
 branch prediction.
-
+%
+\newline
+\newline
 \noindent \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
@@
 yet more approximate look into the program's history.
-
+%
+\newline
+\newline
 \noindent \textbf{Program Counter XOR Delta}: This feature tells us if
@@
 instances of the same memory access instruction is captured here.
-
+%
  % GAC: Seems more predictable than dynamic if its capturing 
@@
  % correct in the past.
+\newline
+\newline
 \noindent \textbf{Confidence}: The confidence, on a scale of 0 to 100,
@@
 other features.
-
+%
+\newline
+\newline
 \noindent \textbf{Page Address XOR Confidence}: This feature scores
@@
 execution.
-
+%
+\newline
+\newline
 \noindent \textbf{Current Signature XOR Delta}: Recall from the
@@
 path that the combination of these two values can lead to.
-
+%
 %Gino: Added the sentence below, is this correct?
@@
 %%%%%%%%%%%%%%%%%%%%%%
-
+%
+\newline
+\newline
 As can be noted above, some composite features are derived from simple

### 2019-04-17 intro.tex
Paragraph spacing

@@
 %of the cache hierarchy into faster higher levels of cache. %phrasing?
+
+\begin{figure}[t]
+\includegraphics[width=0.95\columnwidth]{Motivation}
+\caption{The impact of aggressive prefetching on performance for {\tt 603.bwaves\_s}. 
+The number of useful prefetches increases with aggressiveness
+slower than total prefetches, which wastes bandwidth and 
+harms performance.}
+\label{Fig:Motivation}
+\end{figure}
 
@@
 that prefetcher's accuracy is 66.7\%.
-
-\begin{figure}[t]
-\includegraphics[width=\columnwidth]{Motivation}
-\caption{The impact of aggressive prefetching on performance for {\tt 603.bwaves\_s}. 
-The number of useful prefetches increases with aggressiveness
-slower than total prefetches, which wastes bandwidth and 
-harms performance.}
-\label{Fig:Motivation}
-\end{figure}
 
@@
 more effective at rejecting inaccurate prefetches than SPP's 
-{\color{red}internal} mechanism, we are free to re-tune the rest of 
+{internal} mechanism, we are free to re-tune the rest of 
 SPP's design around maximizing coverage. The result is an increase 

### 2019-04-17 method.tex
Paragraph spacing

@@
 instructions each.
-
+\newline
+\newline
 \noindent \textbf{Single-core performance:} For single-core
@@
 per application using the SimPoint methodology.
-
+\newline
+\newline
 \noindent \textbf{Multi-core performance:} For multi-application
@@
 8-core mix takes up to 3 days to simulate one billion instructions.
-
+\newline
+\newline
 \noindent \textbf{Validation:} We cross-validated our PPF model using
@@
 perceptron weights and try to interpret their distribution.
-
+\newline
+\newline
 \noindent \textbf{Global Pearson's Correlation:} Here we examine the
@@
 a correlation to prefetch outcome with a factor of 0.90. 
-
+\newline
+\newline
 \noindent \textbf{Per Trace Correlation:} Another important way to look
@@
 traces.
-
+\newline
+\newline
 \noindent \textbf{Trimming Features Using Cross-correlation:} Beside
@@
 
-We initially came up {\color{red}with} a set of 23 features.  By studying cross
+We initially came up {with} a set of 23 features.  By studying cross
 correlation of each of these features against others in a 23$\times$23
@@
 is no actual vector multiplication happening in the hardware.
-Obtaining the perceptron sum requires addition of {\color{red}nine} 
+Obtaining the perceptron sum requires addition of {nine} 
 5-bit numbers.
-Using an adder tree of {\color{red}four} 5-bit adders, this can be done in
+Using an adder tree of {four} 5-bit adders, this can be done in
 ceil($log_{2}9$) = 4 steps. Perceptron update only requires weight

### 2019-04-17 related.tex
Paragraph spacing

--- 2018-12-07 related.tex
@@
 accomodate prefetches in their MIN algorithm-inspired Hawkeye cache
-management system.
+management system. Ebrahimi \textit{et al.} introduce
+HPAC~\cite{HPAC} which provides a coordinated control between 
+multiple prefetchers present in a CMP by looking at the 
+prefetcher-induced inter-core interferance.
 
@@
 parametric search for an optimal prefetcher configuration. 
+Hasheni \textit{et al.}~\cite{LSTM} categorize 
+prefetching as a regression problem and use LSTM based Deep 
+Learning approach.

### 2019-04-17 results.tex
Paragraph spacing

--- 2018-12-08 results.tex
@@
 The only benchmark where PPF fails to match the improvement offered by
-other prefetchers is {\tt 607.cactuBSSN\_s}.  Based on our observation
+any other prefetcher is {\tt 607.cactuBSSN\_s}. Based on our observation
 of prefetching behavior, we gather that BOP's aggressive and localized
@@
 of the baseline by \textbf{15.24\%}, which is \textbf{2.27\%} better
-than the next best prefetcher -- SPP.
+than the next best prefetcher -- SPP. {For PPF, the average 
+lookahead depth over the full benchmark is 3.97, while it is 3.28 for just 
+SPP. It is evident that on average for SPP, our scheme allows the 
+prefetcher to speculate 21\% deeper.}
 
@@
 \label{Results-Multi}
-In this section, we demonstrate the improvement achieved by PPF for a
+\noindent In this section, we demonstrate the improvement achieved by PPF for a
 mix of multi-programmed workloads.
-
+\newline
+\newline
 \noindent \textbf{4-Core:} Figure~\ref{Fig:4Core_SPEC2017}
@@
 baseline, which is an improvement of \textbf{5.6}\% over SPP.
-
+%
 \begin{figure}[ht]
@@
 \end{figure}
-
+%
+\newline
+\newline
 \noindent \textbf{8-Core:} The sorted comparison of speedups on the
@@
 corresponding to \textbf{4.6\%} over SPP. This increased improvement
-achieved by PPF over the base engine SPP in a multi-core environment
-is expected as PPF is a very accurate filter. Thus, it eliminates
-useless prefetches before they can cause pollution in the shared LLC.
-BOP offers a better improvement than SPP for the memory intensive
+achieved by PPF over the {underlying prefetcher, SPP,} 
+in a multi-core environment is expected as PPF is a very accurate filter. 
+Thus, it eliminates useless prefetches before they can cause pollution in the 
+shared LLC. BOP offers a better improvement than SPP for the memory intensive
 mixes. This superiority can be attributed to BOP's inherent aggressive

### 2019-04-26 conc.tex
minor
djimenez added acks

@@
 valuable tool in the design of future memory latency constrained systems.
+
+\section{Acknowledgments}
+
+We thank the anonymous reviewers for their valuable comments and feedback to
+improve the content and quality of this paper.  We also thank the National
+Science Foundation, which supports this work through grants FoMR-1823403 and
+CCF-1649242.  We thank Intel Corp. for their generous support. We thank
+Samsung for supporting this work through their Global Outreach Program.

### 2019-05-15 conc.tex
ack hprc
minor

@@
 
-We thank the anonymous reviewers for their valuable comments and feedback to
-improve the content and quality of this paper.  We also thank the National
-Science Foundation, which supports this work through grants FoMR-1823403 and
-CCF-1649242.  We thank Intel Corp. for their generous support. We thank
-Samsung for supporting this work through their Global Outreach Program.
+We thank the anonymous reviewers for their valuable comments and
+feedback to improve the content and quality of this paper.  We also
+thank the National Science Foundation, which supports this work
+through grants FoMR-1823403, I/UCRC-1439722 and CCF-1649242.  We thank
+Intel Corp. for their generous support. We thank Samsung for
+supporting this work through their Global Outreach Program.  Portions
+of this research were conducted with the advanced computing resources
+provided by Texas A\&M High Performance Research Computing.

### 2019-05-21 related.tex
Cleaning up comments
Added commentary on related perceptron prefetcher

@@
 \subsection{Perceptrons in Cache Management}
-
-%Perceptron-based prediction mechanisms have proven to yield superior accuracy, 
-%which represents an opportunity for optimizations in cache management. 
-%Khan \textit{et. al.} proposed SDBP~\cite{sdbp}, in this work, a sampler 
-%structure keeps partial tags of sampled sets separate from the cache. 
-%Three tables of two-bit saturating counters are accessed using a technique 
-%similar to a skewed branch predictor~\cite{Piece_Linear}. For each hit to 
-%a block in the sampled set, the program counter (PC) of the relevant memory 
-%instruction is hashed into the three tables and the corresponding counters 
-%are decremented. For each eviction from a sampled set, the counters corresponding 
-%to the PC of the last instruction to access the victim block are incremented. 
-%For an LLC access, the predictor is consulted by hashing the PC of the memory 
-%access instruction into the three tables and taking the sum of the indexed 
-%counters. When the sum exceeds some threshold, the accessed block is 
-%predicted to be dead. Tags in sampled sets are managed with true LRU and 
-%a reduced associativity, but the LLC may be managed by any policy. 
-%The paper applies its predictor to do block replacement and bypass for LLC.
 
@@
 contributing many new features. 
-
-%features. Compared to previous work that focuses on one or two features at a
-%time, Multiperspective prediction uses many features, each contributing to the
-%overall prediction. The prediction enables placement, promotion and bypass
-%optimizations that further improves performance.
+A similar work also proposed using perceptrons for filtering useless 
+prefetches~\cite{ArxivPerc}. They implemented a basic Rosenblatt perceptron, using 
+general error-correction learning. While they are able to reduce useless 
+prefetches, they do not explore allowing the base-prefetcher to prefetch
+unrestrained down the speculative path. Their design results in
+increased accuracy but lowered cache coverage with low impact on
+ overall performance.
 
 \subsection{Spatial Prefetchers}
-
-% Two widely used prefetchers are Next-$n$ Lines~\cite{nextn} and
-% Stride~\cite{stride} prefetchers. Both capture regular memory access
-% patterns with low overhead. The Next-$n$ Lines prefetcher queues
-% prefetches for the next $n$ blocks after any given miss, expecting
-% that the principle of spatial locality will hold and those cache
-% blocks after a missed block are likely to be used in the future. The
-% stride prefetcher identifies strided reference patterns in programs
-% based on past behavior of missing loads.  When a load misses, cache
-% blocks ahead of that miss are fetched in the pattern following the
-% previous behavior in the hope of avoiding future misses.  Without
-% further knowledge about temporal locality and application
-% characteristics, these prefetchers cannot do more than detecting and
-% prefetching regular memory access patterns with limited spatial
-% locality.
 
@@
 
-%Somogyi et al. proposed a practical, low-overhead prefetcher: Spatial Memory
-%Streaming (SMS) prefetcher~\cite{SMS}. SMS leverages code-based correlation to
-%take advantage of spatial locality in the applications over larger regions of
-%memory called spatial regions. It predicts future access pattern within a
-%spatial region around a miss based on a history of access patterns initiated
-%by that missing instruction in the past. SMS is effective, but it indirectly
-%infers future program control-flow when it speculates on the misses in a
-%spatial region. Thus, the state overhead of this predictor can be higher than
-%the others in this class. The idea was further extended by the class of
-%related prefetchers~\cite{STMS,Temporal_Instruction_Fetch}.
-
-%Other work also considers prefetch timeliness. Michaud proposed the
-%Best-Offset prefetcher~\cite{BOP}. A prefetch offset is set automatically and
-%dynamically tries to adapt to the application behavior. A good prefetch
-%offset is determined by looking at two recent access to a particular block
-%that did not happen too recently. The time between both accesses should be
-%greater than the latency for completing a prefetch request.
-
 \subsection{Lookahead Prefetchers}
@@
 
-%Previous prefetchers have only limited ability to capture memory delta
-%patterns. Moreover, a good prefetch proves to be useful only if the prefetcher
-%manages to bring the block before its next demand access. Both these issues
-%are overcome by Lookahead Prefetchers that use their own predictions to create
-%the speculative path of the program's memory access pattern. By going down the
-%speculation trail to a certain depth, the prefetcher can bring in cache blocks
-%well before they are demand accessed by the processor. VLDP~\cite{VLDP}
-%prefetches a static depth ahead of the demand fetch, without considering the
-%prefetching confidence. More recent Lookahead Prefetchers include SPP and
-%KPC~\cite{KPC}
-
-%Talk about DA-AMPM
-
 \subsection{Managing Prefetched Data}
-
-%not sure this is where this section was going, I just kidna went with it.
 

### 2019-05-22 related.tex
minor

@@
 \label{related}
-
-\subsection{Perceptrons in Cache Management}
-
-In addition to branch prediction~\cite{PerceptronPredictor},
-perceptron-based learning has been applied to the area of cache
-management.  Teran \textit{et al.} propose using perceptrons to
-predict cache line resuse, bypass, and replacement~\cite{Perc_Reuse}.
-Perceptron Learning trains weights selected by hashes of multiple
-features, including the PC of the memory access instruction, some
-other recent PCs, and two different shifts of the tag of the
-referenced block. These features are used to index into weight tables,
-and the weights are then thresholded to generate a prediction. When a
-block from one of a few sampled sets~\cite{sdbp} is reused or evicted,
-the corresponding weights are decremented or incremented, according to
-the perceptron learning rule. Multiperspective Reuse
-Prediction~\cite{Multiperspective} improves on Perceptron Learning by
-contributing many new features. 
-A similar work also proposed using perceptrons for filtering useless 
-prefetches~\cite{ArxivPerc}. They implemented a basic Rosenblatt perceptron, using 
-general error-correction learning. While they are able to reduce useless 
-prefetches, they do not explore allowing the base-prefetcher to prefetch
-unrestrained down the speculative path. Their design results in
-increased accuracy but lowered cache coverage with low impact on
- overall performance.
 
@@
 learning algorithms such as SVMs and logistic regression to do a
-parametric search for an optimal prefetcher configuration. 
-Hasheni \textit{et al.}~\cite{LSTM} categorize 
-prefetching as a regression problem and use LSTM based Deep 
-Learning approach.
+parametric search for an optimal prefetcher configuration.  Hasheni
+\textit{et al.}~\cite{LSTM} categorize prefetching as a regression
+problem and use LSTM based Deep Learning approach.
+
+Wang and Lou propose a similar work where perceptrons filter useless
+prefetches~\cite{ArxivPerc}.  In their design's primary focus was on
+improving the accuracy of an unmodified baseline prefetcher.  Unlike
+the scheme presented here, they implement a basic Rosenblatt
+perceptron, with general error-correction learning.  While they are
+able to increase accuracy, their design results in lower coverage, and
+hence has low impact on overall performance.
+
+\subsection{Perceptrons in Cache Management}
+
+In addition to branch prediction~\cite{PerceptronPredictor},
+perceptron-based learning has been applied to the area of cache
+management.  Teran \textit{et al.} propose using perceptrons to
+predict cache line resuse, bypass, and replacement~\cite{Perc_Reuse}.
+Perceptron Learning trains weights selected by hashes of multiple
+features, including the PC of the memory access instruction, some
+other recent PCs, and two different shifts of the tag of the
+referenced block. These features are used to index into weight tables,
+and the weights are then thresholded to generate a prediction. When a
+block from one of a few sampled sets~\cite{sdbp} is reused or evicted,
+the corresponding weights are decremented or incremented, according to
+the perceptron learning rule. Multiperspective Reuse
+Prediction~\cite{Multiperspective} improves on Perceptron Learning by
+contributing many new features.

### 2019-05-22 back.tex
Minor typo
Cleaning up comments
ran spellcheck

@@
 Section~\ref{related} for other relevant work.
-
-% [EB:] Removing this section title and moving relevant text to next section
-%       We weren't talking about LA prefetchers here anyway, just SPP
-% \subsection{Lookahead-based Prefetchers}
-%% Simple stride prefetching techniques only detect sequences of addresses that differ by a constant value and fail to capture diverse memory access patterns. 
-% Lookahead prefetchers can learn complicated patterns by collecting
-% histories of observed data access patterns.  EB: Didn't really
-% understand this:
-%% "and correlating these with the strides, by considering each new page in a vacuum."
 
@@
 %
-% djimenez: it does look like we're pushing it with all the stuff about SPP.
-% Let's describe it enough to give the reader the idea that he/she has a vague
-% understanding of how it works and can refer to the MICRO paper for more
-% details. (I'm not saying cut anything that's already there, just don't expand.)
-%
-%% NOTE: Can we do away with this example? 
-%% If it looks like we are writing too much about SPP
-%% EDIT: Decided to reorganize and remove this for now.
-%% 	Consider the case that incoming page number 10 with a block offset 3
-%% 	finds a match in the ST. The retrieved pattern signature is 0x30 and
-%% 	the Last Offset is 1. Since now there is the Last Block Offset (1),
-%% 	Incoming Block Offset (3), Old Signature (0x30), and New Signature
-%% 	calculated as per above equation (0x182), SPP can infer
-%% 	non-speculatively that the given pattern of memory accesses (as
-%% 	captured in the Old Signature) leads to the particular Delta. In
-%% 	general, Delta is defined as the difference between the prefetch
-%% 	suggestion block and the initial block which triggered the prefetch.
-%% 	In this case, since SPP is in learning phase, it is defined as the
-%% 	difference between the Incoming Block Offset and the Last Block Offset
-%% 	(+2 in this case). This, in turn generates the new memory pattern
-%% 	(New Signature). This newly learned signature pattern and the delta
-%% 	is stored in the Pattern Table.
-%
-%  -- This looks like a detail that is confusing and not necessary to
-%  understand SPP:
-% \noindent \textbf{Prefetch Table:} The Prefetch Table is a 1024 entry
-% direct-mapped table that keeps a record of last few entries
-% prefetched.  This proves useful for updating the prefetcher state if a
-% tracked prefetch leads to a demand hit or a cache
-% eviction. \textit{Note:} The original paper refers to this as the
-% Prefetch Filter but since in our context filter refers to the
-% perceptron filter, we will be calling this structure as the Prefetch
-% Table to avoid any confusion.
 \newline
@@
 
-% djimenez: this detail is unnecessary and likely to invite questions.
-% reviewers might ask "what about linear separability?" but the proof is in the
-% pudding.
-%This form of indexing makes sure that the hyper-plane learned by the
-%perceptron weights is able to differentiate between linearly
-%inseparable outcomes.
-
-%\textit{Deviations From Actual Perceptrons:} Traditionally, a perceptron
-%prediction involves multiplying the vector of input features:
-%F\textsubscript{1xN}, with the corresponding weight vector:
-%W\textsubscript{Nx1} in a dot product to obtain the sum y\textsubscript{out}.
-%Here what we use is a perceptron-like structure.  The feature is used to hash
-%into the weight of perceptrons and the retrieved weights are added straight
-%away.  Thus, the perceptron algorithm doesn't introduce any multiplication
-%operations in the inference or training process.  Hence what we adapt in this
-%work is a perceptron-like learning algorithm as it involves the same principle
-%involved in perceptron inferencing and training.
-
 \subsection{Case for an On-line Filter}
 \label{Background-Case}
-
-%% Compared to some of the other state of the art prefetchers, SPP is
-%% less aggressive.  In a single-core environment, there is no resource
-%% contention among the different cores, hence aggressive prefetching is bound to prove beneficial. 
-%% As we increase the core
-%% count, we observe that SPP starts outperforming rest of the prefetchers.  This
-%% can be attributed to the fact that each prefetch suggested by SPP is a
-%% carefully calculated and prevents cache pollution.  \textit{<Figure
-%% showing the variation of SPP vs BOP wrt core count>} For 4-core applications,
-%% SPP suggests XX\% fewer prefetches than Best Offset Prefetcher ~\cite{bop} and yet leads to higher IPC.
-%% 
-%% The above analysis shows that with a more careful filtering mechanism, any
-%% prefetcher like SPP can be tuned to become much more aggressive, yielding higher coverage.  The responsibility of maintaining high accuracy now falls on the
-%% independent filter.  To test the hypothesis, we tuned down SPP to the
-%% minimum possible threshold, with the effect of increasing prefetch suggestions
-%% made by SPP by XX\%.  \textit{<Comparison of SPP-Unleased with SPP / BOP>}
-%% Obviously this came at a cost of increased DRAM traffic and cache pollution.
-
-% EB: This remains one of the untouched subsections of the paper
 
@@
 prediction is made. Once the ground truth is known, the weights
-corredponding to the prediction are incremented if the outcome was
+corresponding to the prediction are incremented if the outcome was
 positive, or decremented if it was negative. This update only occurs
@@
 prediction~\cite{Perc_Reuse,Multiperspective}. In this paper, we apply
-it for the first time to prefetch filtering.
-
-%\textit{Perceptron Learning Rule:} In a perceptron-based mechanism, a uniform
-%perceptron update principle is followed.  The weights need to be updated if
-%the prediction was wrong or the magnitude of the dot product does not exceed a
-%certain threshold.  After a certain training period, these weights are
-%proportional to the probability of a correct prediction.  Thresholding the
-%outcome to a
-% magnitude makes sure the perceptrons are trained until a certain level of
-%confidence and yet they are not over-trained to the given set.  The perceptron
-%learning rule will be applied to PPF.
-
-%ETG: this is what I had added about perceptrons, we're trying to 
-%figure out what to leave and what to get rid of ...  
-% djimenez: I ended up cutting everything you wrote as well as the figure. I
-% wrote something similar above, and the perceptron model described here isn't
-% the hashed perceptron that we're using but rather the dot-product model from
-% the original paper.
-
-%\subsection{Perceptron-based Prediction}
-%
-%Perceptrons in microarchitecture were first introduced by Jimenez and Lin~\cite{PerceptronPredictor} 
-%for branch prediction. In this work, a perceptron weights vector is selected by a hash of 
-%the branch address and its dot product with an input vector of the previous branch history 
-%outcomes. If the product is at least 0, the branch is predicted as taken, otherwise it is 
-%predicted not taken. The weights are updated with the perceptron update rule: if the prediction 
-%is incorrect, or the dot product does not exceed some threshold magnitude, then the weights 
-%are incremented or decremented based on their correlation to the corresponding history bits. 
-%Perceptron learning yields to superior accuracy, which results in increased opportunity for 
-%optimizations.
-
-%\begin{figure}[ht]
-%\includegraphics[width=\columnwidth]{Perc_Model}
-%\caption{The Basic Perceptron Model}
-%\label{Fig:Perc_Model}
-%\end{figure}
-
-%\subsubsection{Perceptron Learning}
-%
-%In a perceptron-based mechanism, a uniform perceptron update principle is
-%followed. As shown in the figure~\ref{Fig:Perc_Model} Perceptrons are a simple
-%model of neurons in neural networks \cite{perceptrons1,perceptrons2} modeled
-%by vectors of signed weights learned through online training. The output of a
-%perceptron is the dot product of the weights and a vector of inputs. In this
-%work, we make use of the perceptron learning algorithm. There are two
-%components to the abstract perceptron learning algorithm:
-%
-%\begin{enumerate}
-%\item To predict true or false, a vector of signed weights is chosen according to some criteria. 
-%The dot product of the weights and input vectors, called $y_{\mbox{\small out}}$, is computed. 
-%If $y_{\mbox{\small out}}$ exceeds some threshold, the prediction is true, otherwise it is false.
-%
-%\item To update the weights after the true outcome of the predicted event is
-%known, we first consider the value of $y_{\mbox{\small out}}$. If the
-%prediction was correct and $|y_{\mbox{\small out}}|$ exceeds a threshold
-%$\theta$, then the weights remain unchanged. Otherwise, the inputs are used to
-%update the corresponding weights.  If there is positive correlation between
-%the input and the outcome of the event, the corresponding weight is
-%incremented; otherwise, it is decremented. Over time, the weights are
-%proportional to the probability that the outcome of the event is true in the
-%context of the input and the criterion for choosing that weight. The weights
-%saturate at maximum and minimum values so they will fit in a fixed bit width.
-%\end{enumerate}
+it for the first time to do prefetch filtering.

### 2019-05-22 related.tex
ran spellcheck

@@
 prefetchers can be fed back into their prediction mechanisms to
-predict even further down a speculative path of memory acesses. These
+predict even further down a speculative path of memory accesses. These
 prefetchers can also generalize their learned patterns from one page,
@@
 stack. Jain \textit{et al.} propose Harmony~\cite{Harmony} to
-accomodate prefetches in their MIN algorithm-inspired Hawkeye cache
+accommodate prefetches in their MIN algorithm-inspired Hawkeye cache
 management system. Ebrahimi \textit{et al.} introduce
@@
 multiple prefetchers present in a CMP by looking at the 
-prefetcher-induced inter-core interferance.
+prefetcher-induced inter-core interference.
 
@@
 management.  Teran \textit{et al.} propose using perceptrons to
-predict cache line resuse, bypass, and replacement~\cite{Perc_Reuse}.
+predict cache line reuse, bypass, and replacement~\cite{Perc_Reuse}.
 Perceptron Learning trains weights selected by hashes of multiple
